### 2020-07-30 paper/main.tex
add paper skeleton

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%f
\documentclass[runningheads]{llncs}
%
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amssymb}
\usepackage{multirow}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{tabularx,booktabs}
\usepackage{adjustbox}
\usepackage{xspace}
\usepackage{url}
\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\let\OldStatex\Statex
\renewcommand{\Statex}[1][3]{%
  \setlength\@tempdima{\algorithmicindent}%
  \OldStatex\hskip\dimexpr#1\@tempdima\relax}
\makeatother

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=black,          % color of internal links
    citecolor=black,        % color of links to bibliography
    filecolor=black,      % color of file links
    urlcolor=black           % color of external links
}

\usepackage[square,numbers,sort&compress]{natbib}
% Fixes for using natbib and llncs
\makeatletter
\renewcommand\bibsection%
{
  \section*{\refname
    \@mkboth{\MakeUppercase{\refname}}{\MakeUppercase{\refname}}}
}
\renewcommand\@biblabel[1]{#1.}
\makeatother

% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}

\DeclareMathOperator*{\argmax}{arg\,max}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
\newcommand{\FEmax}{\ensuremath{FE^\text{max}}\xspace}

\usepackage[dvipsnames]{xcolor}
\usepackage[normalem]{ulem}
% \sout is not very good. It cannot handle commands within its argument.
\newcommand{\SuggestEdit}[3][red]{\textcolor{#1}{\sout{#2}}\textcolor{#1}{#3}}
\let\svthefootnote\thefootnote
\newcommand\colorfootnote[2][black]{\def\thefootnote{\color{#1}\svthefootnote}%
  \footnote{\color{#1}#2}\def\thefootnote{\color{black}\svthefootnote}}
\newcommand\RevComment[3][red]{\protect\colorfootnote[#1]{{\textbf{[#2: #3]}}}}

% Command for edits, command for commenting and color
\newcommand\newrevisor[3]{%%
  \colorlet{#1}{#3}
  \expandafter\newcommand\csname#1\endcsname[2]{\SuggestEdit[#1]{##1}{##2}}%%
  \expandafter\newcommand\csname#2\endcsname[1]{\RevComment[#1]{#2}{##1}}%%
}
\newrevisor{manuel}{MANUEL}{Purple}

\begin{document}
%
\title{A Cool Title}
\titlerunning{A Cool Title}

\author{Ekhine Irurozki\inst{1}\and
Manuel L{\'o}pez-Ib{\'a}{\~n}ez\inst{2}}
%\authorrunning{ et al.}

\institute{ \and
  University of Málaga, Spain
   \\\email{manuel.lopez-ibanez@uma.es}
 }
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}

\keywords{Combinatorial optimization \and Bayesian optimization \and Expensive black-box optimization \and Estimation of distribution algorithms}
\end{abstract}
%
%
%
\section{Introduction}
% minimize the uncertainty in the final solutions or

\citet{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}
\citet{PerLopStu2015si}

\citep{LopDubPerStuBir2016irace}

\section{Background}

\section{Methods}

\section{Experimental setup}

\section{Experimental analysis}

\section{Conclusions}

\section*{Acknowledgements}

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}
\bibliographystyle{splncs04nat}
\bibliography{optbib/abbrev,optbib/authors,optbib/journals,optbib/biblio,optbib/crossref}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2020-09-02 paper/main.tex
some writing

@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%f
 \documentclass[runningheads]{llncs}
-%
-\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage[utf8x]{inputenc}
 \usepackage[T1]{fontenc}
+
+\usepackage{amsmath,amssymb}
+%\usepackage[square,numbers,sort&compress]{natbib}
+\usepackage{xspace}
+%% Useful packages
 \usepackage{graphicx}
-\usepackage{amsmath,amssymb,amssymb}
+\usepackage{pdflscape}
 \usepackage{multirow}
-\usepackage{multirow}
-\usepackage{booktabs}
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
-\usepackage{tabularx,booktabs}
-\usepackage{adjustbox}
-\usepackage{xspace}
-\usepackage{url}
-\makeatletter
-\def\BState{\State\hskip-\ALG@thistlm}
-\let\OldStatex\Statex
-\renewcommand{\Statex}[1][3]{%
-  \setlength\@tempdima{\algorithmicindent}%
-  \OldStatex\hskip\dimexpr#1\@tempdima\relax}
-\makeatother
-
-\usepackage{hyperref}
-\hypersetup{
-    colorlinks=true,       % false: boxed links; true: colored links
-    linkcolor=black,          % color of internal links
-    citecolor=black,        % color of links to bibliography
-    filecolor=black,      % color of file links
-    urlcolor=black           % color of external links
-}
-
-\usepackage[square,numbers,sort&compress]{natbib}
+\usepackage{booktabs} % For formal tables
+\usepackage{enumitem}
+\usepackage{afterpage}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage[square,numbers,sort]{natbib}
 % Fixes for using natbib and llncs
 \makeatletter
+% required for natbib to have "References" printed and as section*, not chapter*
 \renewcommand\bibsection%
-{
-  \section*{\refname
-    \@mkboth{\MakeUppercase{\refname}}{\MakeUppercase{\refname}}}
-}
+{\section*{\refname\@mkboth{\MakeUppercase{\refname}}{\MakeUppercase{\refname}}}}
 \renewcommand\@biblabel[1]{#1.}
@@
 \DeclareMathOperator*{\argmax}{arg\,max}
-\renewcommand{\Re}{\ensuremath{\mathbb{R}}}
-\newcommand{\FEmax}{\ensuremath{FE^\text{max}}\xspace}
 
-\usepackage[dvipsnames]{xcolor}
-\usepackage[normalem]{ulem}
-% \sout is not very good. It cannot handle commands within its argument.
-\newcommand{\SuggestEdit}[3][red]{\textcolor{#1}{\sout{#2}}\textcolor{#1}{#3}}
-\let\svthefootnote\thefootnote
-\newcommand\colorfootnote[2][black]{\def\thefootnote{\color{#1}\svthefootnote}%
-  \footnote{\color{#1}#2}\def\thefootnote{\color{black}\svthefootnote}}
-\newcommand\RevComment[3][red]{\protect\colorfootnote[#1]{{\textbf{[#2: #3]}}}}
+\hyphenation{%%% Merriam-Webster
+  di-men-sion-al %%
+  op-tical net-works semi-conduc-tor %%
+  pher-o-mone non-dom-i-nance %%
+  non-dom-i-nat-ed chro-mo-some %%
+  sto-chas-tic make-span an-a-lys-ing}
 
-% Command for edits, command for commenting and color
-\newcommand\newrevisor[3]{%%
-  \colorlet{#1}{#3}
-  \expandafter\newcommand\csname#1\endcsname[2]{\SuggestEdit[#1]{##1}{##2}}%%
-  \expandafter\newcommand\csname#2\endcsname[1]{\RevComment[#1]{#2}{##1}}%%
+\title{Mallows, Black-box combinatorial optimization, limited budget,  ???}
+\author{Ekhiñe Irurozki\inst{1} \and Manuel López-Ibáñez\inst{2}\orcidID{0000-0001-9974-1295}}
+\institute{
+   Basque Center for Applied Mathematics\\
+   \email{eirurozki@bcamath.org}
+   \and
+   University of Málaga, Málaga, Spain\\
+   \email{manuel.lopez-ibanez@uma.es}
 }
-\newrevisor{manuel}{MANUEL}{Purple}
+\date{}%
 
 \begin{document}
-%
-\title{A Cool Title}
-\titlerunning{A Cool Title}
 
-\author{Ekhine Irurozki\inst{1}\and
-Manuel L{\'o}pez-Ib{\'a}{\~n}ez\inst{2}}
-%\authorrunning{ et al.}
+\maketitle
 
-\institute{ \and
-  University of Málaga, Spain
-   \\\email{manuel.lopez-ibanez@uma.es}
- }
-%
-\maketitle              % typeset the header of the contribution
-%
 \begin{abstract}
+%The abstract should briefly summarize the contents of the paper in
+%150--250 words.
+  Black-box combinatorial optimization problems arise in practice when the
+  objective function is evaluated by means of a simulator or a real-world
+  experiment. In such cases, classical techniques such as mixed-integer
+  programming and local search cannot be applied. Moreover, often each solution
+  evaluation is expensive in terms of time or resources, thus only a limited
+  number of evaluations is possible, typically several order of magnitude
+  smaller than in white-box optimization problems. In the continuous case,
+  Bayesian optimization, in particular using Gaussian processes, has proven
+  very effective under these conditions. Much less research is available in the
+  combinatorial case. In this paper, we propose and analyze an
+  estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
+  model and compare it with CEGO, a Bayesian optimization algorithm for
+  combinatorial optimization. Experimental results show that the Bayesian
+  algorithm is able to obtain very good solutions with very few evaluations,
+  however, at a significant computational cost, whereas the proposed EDA
+  outperforms CEGO when the number of solutions evaluated approaches 400, and
+  it is significantly faster. These results suggest that the combination of
+  Bayesian optimization and a Mallows-based EDA may be an interesting direction
+  for future research.
 
@@
 \end{abstract}
-%
-%
-%
+
 \section{Introduction}
@@
 
+Description of UMM and CEGO
+
+Combinatorial Efficient Global Optimization
+(CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
+method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
+optimization problems. In EGO, Gaussian process models are used as a surrogate
+of the landscape of the expensive original problem. An optimization method
+searches for solutions in the surrogate model by optimizing the expected
+improvement criterion, which balances the expected mean and variance of the
+chosen solution. Once a solution is chosen, it is evaluated on the actual
+objective function and the result is used to update the surrogate-model,
+hopefully increasing its predictive power.
+
+CEGO replaces the Euclidean distance measure, used by the surrogate model in
+EGO, with a distance measure appropriate to combinatorial
+landscapes~\citep{ZaeStoBar2014:ppsn}, such as the Kendall distance for
+permutations~\citep{?}. In CEGO, the surrogate model is explored by a GA with
+crossover and mutation operators appropriate for the particular combinatorial
+problem. The original paper notes that coupling the GA with local search does
+not improve the results significantly since the model is anyway an inexact
+estimation of the original objective
+function~\citep[p.~875]{ZaeStoFriFisNauBar2014}.
+
+What else do we need to say?
+
+\newcommand{\minit}{\ensuremath{m_\text{ini}}}
+
+\newcommand{\FEmax}{\ensuremath{FE_{\max}}}
+
 \section{Experimental setup}
+
+We use the implementation of GECO provided by the
+authors\footnote{\url{https://cran.r-project.org/package=CEGO}}. Following the
+original paper~\citep{ZaeStoFriFisNauBar2014}, we use and the GA that optimizes
+the surrogate models uses a population size of 20, crossover rate of 0.5,
+mutation rate of 1, tournament selection of size 2 and probability of 0.9,
+interchange mutation (i.e., exchanging two randomly selected elements of the
+permutation) and cycle crossover for permutations. The budget of each run of
+the GA is \textcolor{red}{10^5} evaluations using the surrogate-model. Although
+it is never stated in the original paper, the implementation of CEGO generates
+a set of initial solutions of size \minit by means of a
+max-min-distance sequential design: new solutions are added to the set
+sequentially by maximizing the minimum distance to solutions already in the
+set. These initial solutions are then evaluated on the actual objective
+function and the result is used to build the initial surrogate model.
+
+Here we use $\minit=20$.
+
+In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
+the actual objective function. In a white-box context, state-of-the-art
+algorithms for the problems considered here typically evaluate
+\textcolor{red}{hundreds? thousands?} of solutions, thus, the budget considered
+here for the black-box context is extremely limited.
+
+We implemented UMM in Python. The parameter settings that we use are...
 
@@
 
-%
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
 \renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}

### 2020-09-02 paper/main.tex
more writing

@@
 \renewcommand\@biblabel[1]{#1.}
+\def\bibfont{\small}
 \makeatother
@@
 permutation) and cycle crossover for permutations. The budget of each run of
-the GA is \textcolor{red}{10^5} evaluations using the surrogate-model. Although
+the GA is \textcolor{red}{$10^5$} evaluations using the surrogate-model. Although
 it is never stated in the original paper, the implementation of CEGO generates
@@
 
+Bayesian optimization methods using a global GP model, such as CEGO, are know
+to have trouble optimizing locally \citep{EriPeaGar2019scalable}. Our
+intuition is that this problem becomes worse in rugged combinatorial
+landscapes, where small steps may produce drastic changes.
+
 \section{Conclusions}
 
-\section*{Acknowledgements}
+\paragraph*{Acknowledgements.}
+
+Thanks to Hao Wang (Leiden University) for pointing us to the arguments of
+\citet{EriPeaGar2019scalable}.
 

### 2020-09-02 cego_lop.py
FG

@@ -113,10 +113,10 @@ def solve_one_umm(instance, ms, rep,  m_ini, best_sol,worst_sol,budgetMM,ratio_s
     return df
 
 
-def run_and_save(n,rep,phi_instance,budgetGA,budgetMM,ratio_samples_learn=0.25,weight_mass_learn=0.9,SLURM_JOB_ID="Local",m_max=400):
+def run_and_save(n,rep,phi_instance,budgetGA,budgetMM,ratio_samples_learn=0.25,weight_mass_learn=0.9,SLURM_JOB_ID="Local",m_max=400,m_ini=10):
     np.random.seed(rep)
-    m_inst = 200
-    m_ini = 10
+    m_inst = 200 # for LOP instance
+    #m_ini = 10
     # rhos=[1e-3,1e-5,1e-7,1e-9]
     # rhos = [1e-9]
     #rhos = [1e-7]
@@ -131,13 +131,15 @@ def run_and_save(n,rep,phi_instance,budgetGA,budgetMM,ratio_samples_learn=0.25,w
     dfuMM = solve_one_umm(instance, m_max, rep, m_ini, instance.best_sol, instance.worst_sol, budgetMM,ratio_samples_learn,weight_mass_learn)
     dfuMM['run_time'] = time.time() - start_time
     df = pd.concat([df,dfuMM],sort=False)
-    
+
     df['best_known'] = instance.get_fitness(instance.best_sol)
     df['worst_known'] = instance.get_fitness(instance.worst_sol)
     df['phi_instance'] = phi_instance
-    df['budget'] = budgetGA
+    df['budgetGA'] = budgetGA
+    df['budgetMM'] = budgetMM
     df['n'] = n
     df['m_max'] = m_max
+    df['m_ini'] = m_ini
     df['ratio_samples_learn'] = ratio_samples_learn
     df['weight_mass_learn'] = weight_mass_learn
     df.to_pickle('pickles/pick'+str(SLURM_JOB_ID)+'.pkl')
@@ -147,6 +149,6 @@ if __name__ == '__main__':
     print(sys.argv)
     params = [float(p) for p in sys.argv[1:]]
     print(params)
-    [n,rep,phi_instance,budgetGA,budgetMM,ratio_samples_learn,weight_mass_learn,SLURM_JOB_ID,m_max] = params
[…]

### 2020-09-03 cego.py
comments paper y code con cego

@@ -34,7 +34,7 @@ def runCEGO(instance, m_ini, m, rep, best_known_sol, worst_known_sol, budgetGA):
     #creation
     cF <- function() sample(n)
     # start optimization
-    print("antes del optimCEGO")
+#    print("antes del optimCEGO")
     res <- optimCEGO(x = NULL,
                      fun = fun,
                      control = list(creationFunction=cF,

### 2020-09-03 cego_lop.py

@@ -124,7 +124,7 @@ def run_and_save(n,rep,phi_instance,budgetGA,budgetMM,ratio_samples_learn=0.25,w
     problem = instance.problem_name
     start_time = time.time()
     df = pd.DataFrame()
-    #df = cego.runCEGO(instance = instance, m_ini = m_ini, m = m_max,rep=rep, best_known_sol=instance.best_sol, worst_known_sol=instance.worst_sol, budgetGA=budgetGA)
+    df = cego.runCEGO(instance = instance, m_ini = m_ini, m = m_max,rep=rep, best_known_sol=instance.best_sol, worst_known_sol=instance.worst_sol, budgetGA=budgetGA)
     df['run_time'] = time.time() - start_time
     #for rho in rhos:
     start_time = time.time()

### 2020-09-03 paper/main.tex
comments paper y code con cego
split work on paper.

@@
 \section{Introduction}
+
+Motivation: Manuel
+
+In many practical optimization problems, the objective function contexts
 % minimize the uncertainty in the final solutions or
+
+Related work: Ekhine
+
+Our contribution: Ekhine
+\begin{itemize}
+\item 
+\end{itemize}
 
@@
 
+MM, GMM, uMM; Ekhine
+
+LOP: Ekhine
+
+Quizás luego: PFSP y QAP
+
 \section{Methods}
 
-Description of UMM and CEGO
+CEGO: Manuel.
 
@@
 
-\newcommand{\minit}{\ensuremath{m_\text{ini}}}
+\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}
 
 \newcommand{\FEmax}{\ensuremath{FE_{\max}}}
+
+Description of UMM: Ekhine
+
+new approach: add sample, learn distri, sample distri
+
+learn distri: uMM with param $\rho$ and binary search rho
+
+sample: peakig distri
+
+Quality guarantees
 
@@
 permutation) and cycle crossover for permutations. The budget of each run of
-the GA is \textcolor{red}{$10^5$} evaluations using the surrogate-model. Although
-it is never stated in the original paper, the implementation of CEGO generates
-a set of initial solutions of size \minit by means of a
-max-min-distance sequential design: new solutions are added to the set
-sequentially by maximizing the minimum distance to solutions already in the
+the GA is \textcolor{red}{$10^5$} evaluations using the
+surrogate-model. Although it is never stated in the original paper, the
+implementation of CEGO generates a set of initial solutions of size \minit by
+means of a max-min-distance sequential design: new solutions are added to the
+set sequentially by maximizing the minimum distance to solutions already in the
 set. These initial solutions are then evaluated on the actual objective
@@
 
+Complete this with details of uMM: MANUEL
+
+Computing system details: MANUEL
+
+Describir como se generan las intancias de LOP: Ekhine
+
+Describir las instancias: Ekhine
+
 \section{Experimental analysis}
 
-Bayesian optimization methods using a global GP model, such as CEGO, are know
+Bayesian optimization methods using a global GP model, such as CEGO, are known
 to have trouble optimizing locally \citep{EriPeaGar2019scalable}. Our

### 2020-09-04 cego_lop.py
instances are now files

@@ -1,3 +1,4 @@
+import pickle
 from scipy.spatial import distance
 import time
 import sys
@@ -112,15 +113,22 @@ def solve_one_umm(instance, ms, rep,  m_ini, best_sol,worst_sol,budgetMM,ratio_s
     # autorho.propor.plot()
     return df
 
+def read_instance(iname):
+    print(iname)
+    if  "LOPsynt_" in iname:
+        pf = open(iname, 'rb')
+        return pickle.load(pf)
 
-def run_and_save(n,rep,phi_instance,budgetGA,budgetMM,ratio_samples_learn=0.25,weight_mass_learn=0.9,SLURM_JOB_ID="Local",m_max=400,m_ini=10):
+
+def run_and_save(n,rep,instance_name,budgetGA,budgetMM,ratio_samples_learn=0.25,weight_mass_learn=0.9,SLURM_JOB_ID="Local",m_max=400,m_ini=10):
     np.random.seed(rep)
     m_inst = 200 # for LOP instance
     #m_ini = 10
     # rhos=[1e-3,1e-5,1e-7,1e-9]
     # rhos = [1e-9]
     #rhos = [1e-7]
-    instance = LOP.generate_synthetic(n, m_inst, phi_instance)
+    #instance = LOP.generate_synthetic(n, m_inst, instance_name)
+    instance = read_instance(instance_name)
     problem = instance.problem_name
     start_time = time.time()
     df = pd.DataFrame()
@@ -134,7 +142,7 @@ def run_and_save(n,rep,phi_instance,budgetGA,budgetMM,ratio_samples_learn=0.25,w
 
     df['best_known'] = instance.get_fitness(instance.best_sol)
     df['worst_known'] = instance.get_fitness(instance.worst_sol)
-    df['phi_instance'] = phi_instance
+    df['instance_name'] = instance_name
     df['budgetGA'] = budgetGA
     df['budgetMM'] = budgetMM
     df['n'] = n
@@ -149,6 +157,6 @@ if __name__ == '__main__':
     print(sys.argv)
     params = [float(p) for p in sys.argv[1:]]
     print(params)
-    [n,rep,phi_instance,budgetGA,budgetMM,ratio_samples_learn,weight_mass_learn,SLURM_JOB_ID,m_max,m_ini] = params
[…]

### 2020-09-04 paper/main.tex
paper, background section

@@
 
-MM, GMM, uMM; Ekhine
+Permutations are defined as bijections of the set $[n]$ integer onto itself.
+The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$.
+We denote permutations with Greek letters with exception of the inverse permutation denoted as $e=1, 2, 3, \ldots,n$.
+We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma_{-1}=e$ always holds.
+
+Distributions over permutations are functions that assign a probability value to each of the permutations in $S_n$, $p(\sigma)\in[0,1]$ \cite{critchlow91}.
+One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
+The MM defines the probability of each permutation $\sigma$ as follows:
+
+\begin{equation}
+p(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
+\end{equation}
+
+with two parameters, $\theta$ and $\sigma_0$, 
+where permutation $\sigma_0$ a reference permutation that has the largest probability value, i.e., the mode of the distribution.
+The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and $\theta$, the dispersion parameter controls this decay.
+The distance $d(\sigma,\sigma_0)$ is the Kendall's-$\tau$ distance.
+The normalization constant $\psi$ can be easily computed for the Kendall's-$\tau$ distance as well as for the Hamming, Cayley and Ulam distance~\cite{Irurozki2016b}.
+
+One of the most common problems related ti probability distributions is that of learning the maximum likelihood parameters  given a sample of data $S$.
+For the MM, this problem tranlates to learning $\theta$ and $\sigma_0$ that best describe a sample of permutations.
+
+The learning process is divide in two stages: first, we estimate the central permutation of the distribution, $\hat\sigma0$ and, second, compute the dispersion parameter, $\hat\theta$.
+
+The exact maximum likelihood estimation is computationally hard~\cite{Dwork:2001:RAM:371920.372165}.
+However, the approximate learning requires polynomilal computational time and is guarantied to obtain high quality parameters~\cite{Caragiannis2013,Coppersmith:2010}.
+
+This process of sample $S$ is as follows: first, compute $\hat\sigma0$ with the Borda count algorithm.
+Borda orders the items $[n]$ by their \textit{Borda score} increasingly, where the Borda score $B(i)$ is the average of each position $i$, $B(i) \propto \sum_{t\in S}  \sigma_t(i)$.
+Second, the computation of $\theta$ is casted as a numerical optimization problem~\cite{Irurozki2016b}.
+
+Recently, it has been proposed an extension of Borda by the name of uBorda.
+uBorda considers a sample of permutations $S$ along with a weight $w(\sigma)$ for each permutation $\sigma$~\cite{}.
+Intuitively it is equivalent to replicating the permutations in the sample proportionally to their weight.
+It has been used to learn a MM in an evolving preference context.
+In this paper, we propose to use uBorda where each permutation in the sample $\sigma\in S$ is weighted by its fitness function $w(\sigma)=f(\sigma)$.
 

### 2020-09-06 cego.py
update to setup umm experiments

@@ -54,7 +54,7 @@ def cego(instance, seed, budget, # m: number of evaluations
     """
     rcode = STAP(rstring, "rcode")
     r_fitness = instance.make_r_fitness()
-    best_x, best_fitness, x , y = rcode.my_cego(r_fitness,
+    best_x, best_fitness, x, y = rcode.my_cego(r_fitness,
                                                 dist = r_kendallTau,
                                                 n = instance.n,
                                                 m_ini = m_ini,

### 2020-09-06 launch_exps.sh

@@ -62,45 +62,46 @@ for n in $LOP_n; do
     done
 done
 
-m_ini=10
-budgetGA=4 # Actually, 10**budgetGA
 budget=400
 nruns=10
 
 LAUNCHER=qsub_job
 #LAUNCHER=launch_local
 
-mkdir -p results
+cego_m_ini=10
+budgetGA=4 # Actually, 10**budgetGA
 
+mkdir -p results
 counter=0
 for instance in $INSTANCES; do
     counter=$((counter+1))
     for run in $(seq 1 $nruns); do
-        $LAUNCHER cego-$run-$counter ./target-runner-cego.py cego $counter $run $instance --m_ini $m_ini --budgetGA $budgetGA --budget 400 --output results/cego-$run-$counter
+        #$LAUNCHER cego-$run-$counter ./target-runner-cego.py cego $counter $run $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $budget --output results/cego-$run-$counter
+        rsl=0.15
+	wml=0.84
+	budgetMM=15
+	umm_m_ini=23
+	$LAUNCHER umm-$run-$counter ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $rsl --wml $wml --budget $budget --output results/umm-$run-$counter
         
-      # ratio_samples_learn=0.15
-      # weight_mass_learn=0.84
-      # budgetMM=15
-      # m_ini=23
-      # $LAUNCHER python3 cego_lop.py $n $repe $phi_instance $budgetGA $budgetMM $ratio_samples_learn $weight_mass_learn $SLURM_JOB_ID $m_max $m_ini
+	# $LAUNCHER python3 cego_lop.py $n $repe $phi_instance $budgetGA $budgetMM $ratio_samples_learn $weight_mass_learn $SLURM_JOB_ID $m_max $m_ini
+	
+	# ratio_samples_learn=0.23
+	# weight_mass_learn=0.88
+	# budgetMM=16
+	# m_ini=38
+	# $LAUNCHER python3 cego_lop.py $n $repe $phi_instance $budgetGA $budgetMM $ratio_samples_learn $weight_mass_learn $SLURM_JOB_ID $m_max $m_ini
 
-      # ratio_samples_learn=0.23
-      # weight_mass_learn=0.88
-      # budgetMM=16
-      # m_ini=38
[…]

### 2020-09-06 target-runner-umm.py

@@ -18,48 +18,47 @@
 ###############################################################################
 import sys
 import os
-import numpy as np
 import pandas as pd
-import mallows_kendall as mk
-from lop import LOP
-from cego_lop import solve_one_umm
-import re
+import runner
+from umm import uMM
 
 
 from argparse import ArgumentParser,RawDescriptionHelpFormatter,_StoreTrueAction,ArgumentDefaultsHelpFormatter,Action
-parser = ArgumentParser(description = "CEGO")
+parser = ArgumentParser(description = "uMM")
 parser.add_argument('configuration_id', type=int, help='configuration_id')
 parser.add_argument('instance_id', type=int, help='instance_id')
 parser.add_argument('algo_seed', type=int, help='random seed')
-# These are part of the instance definition
-parser.add_argument('-seed', type=int, help='inst_seed')
-parser.add_argument('-m', type=int, help='inst_m')
-parser.add_argument('-n', type=int, help='inst_n')
-parser.add_argument('-phi', type=float, help='inst_n')
+parser.add_argument('instance_name', type=str, help='instance name')
+
 # Parameters for the target algorithm
+parser.add_argument('--m_ini', type=int, default=0, help='m_ini')
+parser.add_argument('--budget', type=int, default=400, help='budget')
+parser.add_argument('--budgetMM', type=int, default=0, help='budgetMM')
 parser.add_argument('--rsl', type=float, default=0, help='rsl')
 parser.add_argument('--wml', type=float, default=0, help='wml')
-parser.add_argument('--m_ini', type=int, default=0, help='m_ini')
-parser.add_argument('--budgetMM', type=int, default=0, help='budgetMM')
+parser.add_argument("--output", type=str, default=None, help="output file")
 
 args = parser.parse_args()
-# inst_seed, inst_m, inst_n, phi = re.search("seed=([0-9]+)\s+m=([0-9]+)\s+n=([0-9]+)\s+phi=([^ ]+)", args.instance).group(1,2,3,4)
-# inst_seed = int(inst_seed)
-# inst_m = int(inst_m)
-# inst_n = int(inst_n)
-# phi = float(phi)
 
-np.random.seed(args.seed)
-instance = LOP.generate_synthetic(args.n, args.m, args.phi)
+problem = runner.get_problem(args.instance_name)
+instance = problem.read_instance(args.instance_name)
 
-budget = 400
+budget = args.budget
[…]

### 2020-09-06 umm.py

@@ -90,8 +90,9 @@ def uMM(instance,
                       columns=['rho','phi_estim','phi_sample','Distance'])
     df['Fitness'] = fitnesses
     df['x'] = sample
-    df['budget'] = budget
     df['m_ini'] = m_ini
+    df['seed'] = seed
+    df['budget'] = budget
     df['budgetMM'] = budgetMM
     df['ratio_samples_learn'] = ratio_samples_learn
     df['weight_mass_learn'] = weight_mass_learn

### 2020-09-06 paper/main.tex
more writing

@@
 \DeclareMathOperator*{\argmax}{arg\,max}
+
+%% Revision documents
+\usepackage[dvipsnames]{xcolor}
+\usepackage[normalem]{ulem}
+% FIXME: \sout is not very good. It cannot handle commands within its argument.
+\newcommand{\SuggestEdit}[3][red]{\textcolor{#1}{\sout{#2}}\textcolor{#1}{#3}}
+\let\svthefootnote\thefootnote
+\newcommand\colorfootnote[2][black]{\def\thefootnote{\color{#1}\svthefootnote}%
+  \footnote{\color{#1}#2}\def\thefootnote{\color{black}\svthefootnote}}
+\newcommand\RevComment[3][red]{\protect\colorfootnote[#1]{{\textbf{[#2: #3]}}}}
+% Command for edits, command for commenting and color
+\newcommand\newrevisor[3]{%%
+  \colorlet{#1}{#3}
+  \expandafter\newcommand\csname#1\endcsname[2]{\SuggestEdit[#1]{##1}{##2}}%%
+  \expandafter\newcommand\csname#2\endcsname[1]{\RevComment[#1]{#2}{##1}}%%
+}
+\newrevisor{manuel}{MANUEL}{Purple}
 
@@
 
-CEGO: Manuel.
+\manuel{CEGO: Manuel.}{}
 
@@
 We use the implementation of GECO provided by the
-authors\footnote{\url{https://cran.r-project.org/package=CEGO}}. Following the
-original paper~\citep{ZaeStoFriFisNauBar2014}, we use and the GA that optimizes
+authors\footnote{\url{https://cran.r-project.org/package=CEGO}}. Following the authors of CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use and the GA that optimizes
 the surrogate models uses a population size of 20, crossover rate of 0.5,
@@
 permutation) and cycle crossover for permutations. The budget of each run of
-the GA is \textcolor{red}{$10^5$} evaluations using the
+the GA is \textcolor{red}{$10^4$} evaluations using the
 surrogate-model. Although it is never stated in the original paper, the
@@
 set. These initial solutions are then evaluated on the actual objective
-function and the result is used to build the initial surrogate model.
+function and the result is used to build the initial surrogate model.  Here we
+use $\minit=20$.\MANUEL{\citet{ZaeStoBar2014:ppsn} uses 10 so we may want to change this}
 
-Here we use $\minit=20$.
+The CPU time required for a single run of CEGO using these settings is larger
+than a week.  For simplicity, we use Kendall Tau distance~\citep{?} in all
+experiments\MANUEL{\citet{ZaeStoBar2014:ppsn} claims this is called Swap
+  distance, could you double-check?}. Nevertheless,
+\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
+when using Hamming distance. We plan to extend our analysis of uMM to other
+distance measures and to the dynamic selection of distance
+measures~\citep{ZaeStoBar2014:ppsn}.
 
@@
 
-We implemented UMM in Python. The parameter settings that we use are...
+We implemented UMM in Python 3. The parameter settings that we use are...
 
@@
 
-Computing system details: MANUEL
+\manuel{Computing system details: MANUEL}{%
+  All experiments were run on Intel Xeon E5-2640 CPUs at 2.50\,GHz, 64\,GB RAM
+  running CentOS Linux release 7.4.1708 (Core).  }
 
@@
 Describir las instancias: Ekhine
+
+For the LOP, we generated synthetic instances of size $n=20$, $m=200$, $\phi\in\{0.5,0.7,0.9\}$\MANUEL{what is $m$?}
+
+In the case of the QAP and PFSP, we consider the same instances as
+\citet{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, i.e.,
+\texttt{nug12},\texttt{nug30},\texttt{tho30} and \texttt{kra32} for the QAP,
+with $n$ equal to the number in the instance name, and \texttt{reC05},
+\texttt{reC13}, \texttt{reC19}, \texttt{reC31} for the PFSP, with
+$n \in \{20, 20, 30, 50\}$, respectively.
 
@@
 \section{Conclusions}
+
+Open questions:
+\begin{itemize}
+\item How difficult is to extend uMM to other distance metrics?
+\item Can we plot the posterior probability of the optimal solution?
+\end{itemize}
 

### 2020-09-06 paper/main.tex
more writing
some minor changes to simplify the code

@@
 \title{Mallows, Black-box combinatorial optimization, limited budget,  ???}
-\author{Ekhiñe Irurozki\inst{1} \and Manuel López-Ibáñez\inst{2}\orcidID{0000-0001-9974-1295}}
-\institute{
-   Basque Center for Applied Mathematics\\
-   \email{eirurozki@bcamath.org}
-   \and
-   University of Málaga, Málaga, Spain\\
-   \email{manuel.lopez-ibanez@uma.es}
-}
+% Double-blind
+\author{Double-blind}
+% \author{Ekhiñe Irurozki\inst{1} \and Manuel López-Ibáñez\inst{2}\orcidID{0000-0001-9974-1295}}
+% \institute{
+%    Basque Center for Applied Mathematics\\
+%    \email{eirurozki@bcamath.org}
+%    \and
+%    University of Málaga, Málaga, Spain\\
+%    \email{manuel.lopez-ibanez@uma.es}
+% }
 \date{}%
@@
   for future research.
-
+\textcolor{red}{16 pages max, deadline: 1 November 2020}
 \keywords{Combinatorial optimization \and Bayesian optimization \and Expensive black-box optimization \and Estimation of distribution algorithms}
@@
 We use the implementation of GECO provided by the
-authors\footnote{\url{https://cran.r-project.org/package=CEGO}}. Following the authors of CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use and the GA that optimizes
-the surrogate models uses a population size of 20, crossover rate of 0.5,
-mutation rate of 1, tournament selection of size 2 and probability of 0.9,
-interchange mutation (i.e., exchanging two randomly selected elements of the
-permutation) and cycle crossover for permutations. The budget of each run of
-the GA is \textcolor{red}{$10^4$} evaluations using the
+authors\footnote{\url{https://cran.r-project.org/package=CEGO}}. Following the
+authors of CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use and
+the GA that optimizes the surrogate models uses a population size of 20,
+crossover rate of 0.5, mutation rate of 1, tournament selection of size 2 and
+probability of 0.9, interchange mutation (i.e., exchanging two randomly
+selected elements of the permutation) and cycle crossover for permutations. The
+budget of each run of the GA is \textcolor{red}{$10^4$} evaluations using the
 surrogate-model. Although it is never stated in the original paper, the
-implementation of CEGO generates a set of initial solutions of size \minit by
-means of a max-min-distance sequential design: new solutions are added to the
-set sequentially by maximizing the minimum distance to solutions already in the
-set. These initial solutions are then evaluated on the actual objective
-function and the result is used to build the initial surrogate model.  Here we
-use $\minit=20$.\MANUEL{\citet{ZaeStoBar2014:ppsn} uses 10 so we may want to change this}
+implementation of CEGO generates a set of initial solutions of size
+$\minit=10$ by means of a max-min-distance sequential design: new solutions
+are added to the set sequentially by maximizing the minimum distance to
+solutions already in the set. These initial solutions are then evaluated on the
+actual objective function and the result is used to build the initial surrogate
+model.% \citet{ZaeStoBar2014:ppsn} uses 10
 

### 2020-09-06 cego.py

@@ -64,7 +64,6 @@ def cego(instance, seed, budget, # m: number of evaluations
 
     # We use what instance recorded because CEGO may not get us what was
     # actually evaluated.
-    
     return pd.DataFrame(dict(
         Fitness=instance.evaluations, x=instance.solutions, m_ini = m_ini, seed = seed, budget = budget, budgetGA = budgetGA,
         Distance = [ instance.distance_to_best(perm) for perm in instance.solutions]))

### 2020-09-06 launch_exps.sh

@@ -56,7 +56,7 @@ for n in $LOP_n; do
     for m in $LOP_m; do
         for seed in $LOP_seed; do
             for phi in $LOP_phi; do
-                INSTANCES="$INSTANCES LOP-synthetic,seed=${seed},n=${n},m=${m},phi=0.5"
+                INSTANCES="$INSTANCES LOP-synthetic,seed=${seed},n=${n},m=${m},phi=${phi}"
             done
         done
     done

### 2020-09-06 lop.py

@@ -18,28 +18,32 @@ class LOP(Problem):
   problem_name = "LOP"
   
   @classmethod
-  def generate_synthetic(cls, n, m, phi):
+  def generate_synthetic(cls, n, m, phi, seed = None):
+    if seed is None:
+      seed = np.random.randint(1, 123456789, 1)[0]
+    np.random.seed(seed)
     instance = synthetic_LOP(n, m, phi)
     best_sol = list(range(n))
     worst_sol = best_sol[::-1]
-    return cls(n, instance, best_sol = best_sol, worst_sol = worst_sol)
+    return cls(n, instance, best_sol = best_sol, worst_sol = worst_sol,
+               instance_name = f"LOP-synthetic,seed={seed},n={n},m={m},phi={phi}")
+    )
 
   @classmethod
   def read_instance(cls, filename):
     if "synthetic" in filename:
       seed, n, m, phi = re.search("seed=([0-9]+),n=([0-9]+),m=([0-9]+),phi=([^ ]+)", filename).group(1,2,3,4)
-      np.random.seed(int(seed))
-      return cls.generate_synthetic(int(n), int(m), float(phi))
+      return cls.generate_synthetic(int(n), int(m), float(phi), int(seed))
     else:
       # FIXME: How to read LOPLIB instances?
       pass
     
   # Methods
-  def __init__(self, n, instance, best_sol = None, worst_sol = None):
+  def __init__(self, n, instance, best_sol = None, worst_sol = None, instance_name = None):
     self.n = n
     # FIXME: Can we find a better name than instance?
     self.instance = instance
-    super().__init__(best_sol = best_sol, worst_sol = worst_sol)
+    super().__init__(best_sol = best_sol, worst_sol = worst_sol, instance_name = instance_name)
         
   def fitness_nosave(self, x):
     # In case it is not numpy array.

### 2020-09-06 pfsp.py

@@ -39,9 +39,8 @@ class PFSP(Problem):
         self.P = np.asarray(P)
         # jobs, machines
         self.n, self.m = self.P.shape
-        self.instance_name = instance_name
-        super().__init__(best_sol = best_sol, worst_sol = worst_sol)
-
+        super().__init__(best_sol = best_sol, worst_sol = worst_sol,
+                         instance_name = instance_name)
 
     def completion_times(self, x):
         C = self.P[x, :]

### 2020-09-06 problem.py

@@ -5,9 +5,10 @@ import rpy2.rinterface as ri
 from mallows_kendall import kendallTau
 
 class Problem:
-    def __init__(self, best_sol = None, worst_sol = None):
+    def __init__(self, best_sol = None, worst_sol = None, instance_name = None):
         self.best_sol = best_sol
         self.worst_sol = worst_sol
+        self.instance_name = instance_name
         if best_sol is not None:
             self.best_fitness = self.fitness_nosave(best_sol)
         if worst_sol is not None:

### 2020-09-06 qap.py

@@ -31,8 +31,8 @@ class QAP(Problem):
         assert self.B.shape[0] == self.B.shape[1]
         assert self.A.shape[0] == self.B.shape[0]
         self.n = self.A.shape[0]
-        self.instance_name = instance_name
-        super().__init__(best_sol = best_sol, worst_sol = worst_sol)
+        super().__init__(best_sol = best_sol, worst_sol = worst_sol,
+                         instance_name = instance_name)
         
     def fitness_nosave(self, x):
         return np.sum(self.A * self.B[np.ix_(x, x)])

### 2020-09-06 runner.py

@@ -27,23 +27,30 @@ def get_problem(instance_name):
     raise ValueError("Unknown problem: " + instance_name)
 
 
-def run_once(algo, instance, seed, **algo_params):
+def run_once(algo_name, instance_name, seed, out_filename = None,
+             **algo_params):
+    if algo_name == "uMM":
+        from umm import uMM
+        algo = uMM
+    elif algo_name == "CEGO":
+        from umm import cego
+        algo = cego
+    else:
+        raise ValueError("Unknown algo: " + algo_name)
+    
+    problem = get_problem(instance_name)
+    instance = problem.read_instance(instance_name)
     timer = Timer()
     df = algo(instance, seed, **algo_params)
     if instance.best_fitness is not None and instance.worst_fitness is not None:
         df['Fitness'] = (df.Fitness - instance.best_fitness) / (instance.worst_fitness - instance.best_fitness)
     
     df['run_time'] = timer.elapsed()
+    df['Problem'] = instance.problem_name
+    df['instance'] = instance.instance_name
+    df['Solver'] = algo_name
+    if out_filename is not None:
+        df.to_csv(out_filename + '.csv', index=False)
+        df.to_pickle(out_filename + '.pkl')
     return df
-
-# from cego import cego
-# from lop import LOP
-# from umm import uMM
-
-# instance = LOP.read_instance("synthetic,seed=1,n=20,m=200,phi=0.5")
-
-# #df = run_one(cego, instance, 1, budget = 10, m_ini = 5, budgetGA=100)
-# df = run_one(uMM, instance, 1, budget = 100, m_ini = 5, budgetMM=10,
-#              ratio_samples_learn = 0.22,
-#              weight_mass_learn = .83)
 

### 2020-09-06 target-runner-cego.py

@@ -19,7 +19,6 @@
 import sys
 import os
 import pandas as pd
-from cego import cego
 import runner
 
 from argparse import ArgumentParser,RawDescriptionHelpFormatter,_StoreTrueAction,ArgumentDefaultsHelpFormatter,Action
@@ -37,9 +36,6 @@ parser.add_argument("--output", type=str, default=None, help="output file")
 
 args = parser.parse_args()
 
-problem = runner.get_problem(args.instance_name)
-instance = problem.read_instance(args.instance_name)
-
 budget = args.budget
 assert budget > 2 * args.m_ini
 
@@ -48,13 +44,7 @@ budgetGA = 10**args.budgetGA
 stdout = sys.stdout
 outfilename = f'c{args.configuration_id}-{args.instance_id}-{args.seed}.stdout' 
 with open(outfilename, 'w') as sys.stdout:
-    df = runner.run_once(cego, instance, args.seed, budget = budget, m_ini = args.m_ini, budgetGA = budgetGA)
-    if args.output is not None:
-        df['Problem'] = instance.problem_name
-        df['instance'] = args.instance_name
-        df['Solver'] = "CEGO"
-        df.to_csv(args.output + '.csv', index=False)
-        df.to_pickle(args.output + '.pkl')
+    df = runner.run_once("CEGO", instance_name, args.seed, budget = budget, m_ini = args.m_ini, budgetGA = budgetGA, out_filename = args.output)
         
 sys.stdout = stdout
 print(df["Fitness"].min())

### 2020-09-06 target-runner-umm.py

@@ -20,8 +20,6 @@ import sys
 import os
 import pandas as pd
 import runner
-from umm import uMM
-
 
 from argparse import ArgumentParser,RawDescriptionHelpFormatter,_StoreTrueAction,ArgumentDefaultsHelpFormatter,Action
 parser = ArgumentParser(description = "uMM")
@@ -40,24 +38,15 @@ parser.add_argument("--output", type=str, default=None, help="output file")
 
 args = parser.parse_args()
 
-problem = runner.get_problem(args.instance_name)
-instance = problem.read_instance(args.instance_name)
-
 budget = args.budget
 assert budget > 2 * args.m_ini
 
 stdout = sys.stdout
 outfilename = f'c{args.configuration_id}-{args.instance_id}-{args.seed}.stdout'
 with open(outfilename, 'w') as sys.stdout:
-    df = runner.run_once(uMM, instance, args.seed, budget = budget, m_ini = args.m_ini, 
-                         budgetMM = args.budgetMM, ratio_samples_learn = args.rsl, weight_mass_learn = args.wml)
-    if args.output is not None:
-        df['Problem'] = instance.problem_name
-        df['instance'] = args.instance_name
-        df['Solver'] = "uMM"
-        df.to_csv(args.output + '.csv', index=False)
-        df.to_pickle(args.output + '.pkl')
-        
+    df = runner.run_once("uMM", args.instance_name, args.seed, budget = budget, m_ini = args.m_ini, 
+                         budgetMM = args.budgetMM, ratio_samples_learn = args.rsl, weight_mass_learn = args.wml, out_filename = args.output)
+            
 sys.stdout = stdout
 print(df["Fitness"].min())
 # remove tmp file.

### 2020-09-06 umm.py

@@ -32,7 +32,7 @@ def binary_search_rho(w, ratio_samples_learn, weight_mass_learn,
   return binary_search_rho(w, ratio_samples_learn, weight_mass_learn, mid, last)
 
 def get_expected_distance(iterat,n,m):
-    N = (n-1) * n / 2
+    N = (n - 1) * n / 2
     f_ini, f_end = N / 4, 1
     iter_decrease = m - 10
     salto = (f_ini - f_end ) / iter_decrease
@@ -48,7 +48,6 @@ def uMM(instance,
         weight_mass_learn):
     np.random.seed(seed)
     n = instance.n
-    N = (n - 1) * n / 2
     sample = [np.random.permutation(range(n)) for _ in range(m_ini)]
     fitnesses = [instance.fitness(perm) for perm in sample]
     # ['rho','phi_estim','phi_sample','Distance']
@@ -76,6 +75,7 @@ def uMM(instance,
         #perm = perm[borda]
         perms = [perm[borda] for perm in perms]
         dists = distance.cdist(perms, sample, metric=mk.kendallTau)
+        # FIXME: We probably do not need to sort, just find the min per axis=1.
         dists = np.sort(dists, axis=1)
         indi = np.argmax(dists[:, 0]) #index of the perm with the farthest closest permutation. Maximizes the min dist to the sample
         perm = perms[indi]

### 2020-09-06 launch_exps.sh
more fixes for launching experiments.

@@ -45,6 +45,9 @@ qap/nug12.dat \
 qap/nug30.dat \
 qap/tho30.dat \
 pfsp/rec05.txt \
+pfsp/rec13.txt \
+pfsp/rec19.txt \
+pfsp/rec31.txt \
 "
 
 LOP_n=20
@@ -76,12 +79,13 @@ counter=0
 for instance in $INSTANCES; do
     counter=$((counter+1))
     for run in $(seq 1 $nruns); do
-        #$LAUNCHER cego-$run-$counter ./target-runner-cego.py cego $counter $run $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $budget --output results/cego-$run-$counter
         rsl=0.15
 	wml=0.84
 	budgetMM=15
 	umm_m_ini=23
-	$LAUNCHER umm-$run-$counter ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $rsl --wml $wml --budget $budget --output results/umm-$run-$counter
+	$LAUNCHER umm-$counter-$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $rsl --wml $wml --budget $budget --output results/umm-$counter-$run
+        
+	$LAUNCHER cego-$counter-$run ./target-runner-cego.py cego $counter $run $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $budget --output results/cego-$counter-$run
         
 	# $LAUNCHER python3 cego_lop.py $n $repe $phi_instance $budgetGA $budgetMM $ratio_samples_learn $weight_mass_learn $SLURM_JOB_ID $m_max $m_ini
 	

### 2020-09-06 pfsp.py

@@ -56,4 +56,6 @@ class PFSP(Problem):
         return C[self.n - 1, self.m - 1]
     
     def fitness_nosave(self, x):
+        # In case it is not numpy array.
+        x = np.asarray(x, dtype=int)
         return self.makespan(x)

### 2020-09-06 problem.py

@@ -9,9 +9,13 @@ class Problem:
         self.best_sol = best_sol
         self.worst_sol = worst_sol
         self.instance_name = instance_name
-        if best_sol is not None:
+        if best_sol is None:
+            self.best_fitness = None
+        else:
             self.best_fitness = self.fitness_nosave(best_sol)
-        if worst_sol is not None:
+        if worst_sol is None:
+            self.worst_fitness = None
+        else:
             self.worst_fitness = self.fitness_nosave(worst_sol)
         self.reset()
     

### 2020-09-06 qap.py

@@ -35,5 +35,7 @@ class QAP(Problem):
                          instance_name = instance_name)
         
     def fitness_nosave(self, x):
+        # In case it is not numpy array.
+        x = np.asarray(x, dtype=int)
         return np.sum(self.A * self.B[np.ix_(x, x)])
     

### 2020-09-06 runner.py

@@ -33,7 +33,7 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
         from umm import uMM
         algo = uMM
     elif algo_name == "CEGO":
-        from umm import cego
+        from cego import cego
         algo = cego
     else:
         raise ValueError("Unknown algo: " + algo_name)

### 2020-09-06 target-runner-cego.py

@@ -27,12 +27,12 @@ parser.add_argument('configuration_id', type=str, help='configuration_id')
 parser.add_argument('instance_id', type=str, help='instance_id')
 parser.add_argument('seed', type=int, help='random seed')
 parser.add_argument('instance_name', type=str, help='instance name')
+parser.add_argument("--output", type=str, default=None, help="output file")
 
 # Parameters for the target algorithm
 parser.add_argument('--m_ini', type=int, default=0, help='m_ini')
 parser.add_argument('--budget', type=int, default=400, help='budget')
 parser.add_argument('--budgetGA', type=int, default=0, help='budgetGA')
-parser.add_argument("--output", type=str, default=None, help="output file")
 
 args = parser.parse_args()
 
@@ -44,7 +44,7 @@ budgetGA = 10**args.budgetGA
 stdout = sys.stdout
 outfilename = f'c{args.configuration_id}-{args.instance_id}-{args.seed}.stdout' 
 with open(outfilename, 'w') as sys.stdout:
-    df = runner.run_once("CEGO", instance_name, args.seed, budget = budget, m_ini = args.m_ini, budgetGA = budgetGA, out_filename = args.output)
+    df = runner.run_once("CEGO", args.instance_name, args.seed, budget = budget, m_ini = args.m_ini, budgetGA = budgetGA, out_filename = args.output)
         
 sys.stdout = stdout
 print(df["Fitness"].min())

### 2020-09-06 target-runner-umm.py

@@ -23,10 +23,11 @@ import runner
 
 from argparse import ArgumentParser,RawDescriptionHelpFormatter,_StoreTrueAction,ArgumentDefaultsHelpFormatter,Action
 parser = ArgumentParser(description = "uMM")
-parser.add_argument('configuration_id', type=int, help='configuration_id')
-parser.add_argument('instance_id', type=int, help='instance_id')
-parser.add_argument('algo_seed', type=int, help='random seed')
+parser.add_argument('configuration_id', type=str, help='configuration_id')
+parser.add_argument('instance_id', type=str, help='instance_id')
+parser.add_argument('seed', type=int, help='random seed')
 parser.add_argument('instance_name', type=str, help='instance name')
+parser.add_argument("--output", type=str, default=None, help="output file")
 
 # Parameters for the target algorithm
 parser.add_argument('--m_ini', type=int, default=0, help='m_ini')
@@ -34,7 +35,6 @@ parser.add_argument('--budget', type=int, default=400, help='budget')
 parser.add_argument('--budgetMM', type=int, default=0, help='budgetMM')
 parser.add_argument('--rsl', type=float, default=0, help='rsl')
 parser.add_argument('--wml', type=float, default=0, help='wml')
-parser.add_argument("--output", type=str, default=None, help="output file")
 
 args = parser.parse_args()
 

### 2020-09-06 umm.py

@@ -79,7 +79,8 @@ def uMM(instance,
         dists = np.sort(dists, axis=1)
         indi = np.argmax(dists[:, 0]) #index of the perm with the farthest closest permutation. Maximizes the min dist to the sample
         perm = perms[indi]
-
+        # FIXME: This should already be an array of int type. 
+        perm = np.asarray(perm, dtype='int')
         sample.append(perm)
         fitnesses.append(instance.fitness(perm))
 

### 2020-09-06 runner.py
* runner.py: Compress output files by default.

@@ -50,7 +50,7 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
     df['instance'] = instance.instance_name
     df['Solver'] = algo_name
     if out_filename is not None:
-        df.to_csv(out_filename + '.csv', index=False)
-        df.to_pickle(out_filename + '.pkl')
+        df.to_csv(out_filename + '.csv.gz', index=False)
+        df.to_pickle(out_filename + '.pkl.gz')
     return df
 

### 2020-09-07 umm.py
* umm.py: whitespace.

@@ -64,8 +64,8 @@ def uMM(instance,
         ws = rho ** ws #MINIMIZE
         # print(ws,co[:int(len(co)/4)].sum(),co.sum())
 
-        borda = mk.uborda(np.array(sample),ws)
-        phi_estim = mk.u_phi(sample,borda, ws)
+        borda = mk.uborda(np.array(sample), ws)
+        phi_estim = mk.u_phi(sample, borda, ws)
         expected_dist = get_expected_distance(m, n, budget)
         phi_sample = mk.find_phi(n, expected_dist, expected_dist + 1)
         #phi_estim = 1 - (m+1)/(budget)

### 2020-09-07 paper/main.tex
update cpu info.

@@
 \manuel{Computing system details: MANUEL}{%
-  All experiments were run on Intel Xeon E5-2640 CPUs at 2.50\,GHz, 64\,GB RAM
+  All experiments were run on Intel Xeon ``Ivybridge'' E5-2650v2 CPUs at 2.60\,GHz, 64\,GB RAM
   running CentOS Linux release 7.4.1708 (Core).  }

### 2020-09-08 paper/main.tex
paper ideasq

@@
 
-Description of UMM: Ekhine
+cosas de intro : 
+This new methodology relies on the assumption that the fitness landscatpe is correlated with a distribution over the same space. 
+Previous paper have confirmed this point
+For this case study, we show that this case for the Mallows model under the Kendall distance and
 
-new approach: add sample, learn distri, sample distri
+In this section, we introduce our main contribution: We present a new methodology for the optimization of black box functions with a minimal number of queries.
+This methodology is a probabilistic, population-based algorithm.
+
+Essentially, the algorithm  considers a sample of permutations along with the fitness of each of them. It is an iterative process in which the number of iterations is fixed. 
+Then, at each iteration, UMM (i) estimates a surrogate distribution of the fitness function and (ii) proposes a good fitted individual.
+
+\section{Learning stage}
+In this section
 
@@
 
-sample: peakig distri
+sample: peakig distri. Borda is the
 
 Quality guarantees
+
+prrof, probem-distri correlation
 
@@
 budget of each run of the GA is \textcolor{red}{$10^4$} evaluations using the
-surrogate-model. Although it is never stated in the original paper, the
+surrogate-model. Although it is never stated in the original paper, the	`
 implementation of CEGO generates a set of initial solutions of size

### 2020-09-10 paper/main.tex
a bit of intro

@@
 
-In many practical optimization problems, the objective function contexts
+In many practical optimization problems, the objective function is not
+explicitly available and solutions are evaluated by means of expensive
+prediction models, simulations or physical experiments. When decision variables
+are continuous, the use of Bayesian surrogate-models, e.g., Gaussian processes,
+in optimization has become widespread~\citep{JonSchWel98go,ForKea2009surrogate}.
+Motivated by this success, there have been attempts at adapting such surrogate models to the combinatorial case~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014},
+
 % minimize the uncertainty in the final solutions or

### 2020-09-10 problem.py
* problem.py: use 0.5 to divide by 2.

@@ -36,7 +36,7 @@ class Problem:
     def distance_to_best(self, perm):
         if self.best_sol is None:
             return np.nan
-        return kendallTau(perm, self.best_sol) / (self.n * (self.n - 1) / 2)
+        return kendallTau(perm, self.best_sol) / (self.n * (self.n - 1) * 0.5)
   
     # Returns a closure function that can be called from R.
     # WARNING: this function minimizes for CEGO

### 2020-09-10 runner.py
save function evaluations explicitly.

@@ -44,7 +44,7 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
     df = algo(instance, seed, **algo_params)
     if instance.best_fitness is not None and instance.worst_fitness is not None:
         df['Fitness'] = (df.Fitness - instance.best_fitness) / (instance.worst_fitness - instance.best_fitness)
-    
+    df['Function evaluations'] = np.arange(1, 10)
     df['run_time'] = timer.elapsed()
     df['Problem'] = instance.problem_name
     df['instance'] = instance.instance_name

### 2020-09-10 umm.py

@@ -71,6 +71,7 @@ def uMM(instance,
         #phi_estim = 1 - (m+1)/(budget)
         perms = mk.samplingMM(budgetMM, n, phi=phi_sample, k=None)
         #perm = perm[borda]
+        # Transforms from sampling space to board space.
         perms = [perm[borda] for perm in perms]
         dists = distance.cdist(perms, sample, metric=mk.kendallTau)
         # FIXME: We probably do not need to sort, just find the min per axis=1.

### 2020-09-10 runner.py
save function evaluations explicitly.

@@ -44,7 +44,7 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
     df = algo(instance, seed, **algo_params)
     if instance.best_fitness is not None and instance.worst_fitness is not None:
         df['Fitness'] = (df.Fitness - instance.best_fitness) / (instance.worst_fitness - instance.best_fitness)
-    df['Function evaluations'] = np.arange(1, 10)
+    df['Function evaluations'] = np.arange(1, len(df['Fitness'])+1)
     df['run_time'] = timer.elapsed()
     df['Problem'] = instance.problem_name
     df['instance'] = instance.instance_name

### 2020-09-10 paper/main.tex
save function evaluations explicitly.
import numpy

@@
 are continuous, the use of Bayesian surrogate-models, e.g., Gaussian processes,
-in optimization has become widespread~\citep{JonSchWel98go,ForKea2009surrogate}.
-Motivated by this success, there have been attempts at adapting such surrogate models to the combinatorial case~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014},
+in optimization has become
+widespread~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
+success, there have been attempts at adapting such surrogate models to the
+combinatorial case~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}.
 

### 2020-09-10 runner.py

@@ -12,6 +12,7 @@ class Timer:
         self.elapsed()
 
 import pandas as pd
+import numpy as np
 
 def get_problem(instance_name):
     instance_name = instance_name.lower()

### 2020-09-11 lop.py
lop real data

@@ -1,3 +1,4 @@
+import pandas as pd
 from problem import Problem
 import mallows_kendall as mk
 import numpy as np
@@ -6,49 +7,72 @@ import re
 def synthetic_LOP(n, m, phi):
   instance = np.zeros((n,n))
   s = np.asarray(mk.samplingMM(m, n, phi=phi, k=None))
+  central = np.random.permutation(n)
+  s = s[:,central] #compose
   for i in range(n):
       for j in range(i+1,n):
           instance[i,j] = (s[:,i] < s[:,j]).sum() / m
           instance[j,i] = 1 - instance[i,j]
-  return instance
+  return instance, central
 
 # Linear Ordering Problem
 class LOP(Problem):
   # Class attributes
   problem_name = "LOP"
-  
+
   @classmethod
   def generate_synthetic(cls, n, m, phi, seed = None):
     if seed is None:
       seed = np.random.randint(1, 123456789, 1)[0]
     np.random.seed(seed)
-    instance = synthetic_LOP(n, m, phi)
-    best_sol = list(range(n))
+    instance, best_sol = synthetic_LOP(n, m, phi)
+    # best_sol = list(range(n))
     worst_sol = best_sol[::-1]
     return cls(n, instance, best_sol = best_sol, worst_sol = worst_sol,
                instance_name = f"LOP-synthetic,seed={seed},n={n},m={m},phi={phi}")
-    
+
   @classmethod
   def read_instance(cls, filename):
     if "synthetic" in filename:
       seed, n, m, phi = re.search("seed=([0-9]+),n=([0-9]+),m=([0-9]+),phi=([^ ]+)", filename).group(1,2,3,4)
       return cls.generate_synthetic(int(n), int(m), float(phi), int(seed))
     else:
-      # FIXME: How to read LOPLIB instances?
-      pass
-    
+      # The instances in http://grafo.etsii.urjc.es/optsicom/lolib/#instances and best sols try to maximize the superdiagonal.
+      # We minimize the subdiagonal, which is equivalent, but the best sol needs to be updated
+      print(f"Reading instance from {filename}")
+      with open(filename) as f:
+          n = int(f.readline().strip())
+          instance = np.loadtxt(f, max_rows=n)
+      best_sol, best_fitness = None, None
+      best = pd.read_csv('lop/best_knowns.csv', sep='\t')
+      best = best[best.instance_name==filename.split('/')[-1]]#.best_fval.values[0]
[…]

### 2020-09-11 problem.py

@@ -5,12 +5,12 @@ import rpy2.rinterface as ri
 from mallows_kendall import kendallTau
 
 class Problem:
-    def __init__(self, best_sol = None, worst_sol = None, instance_name = None):
+    def __init__(self, best_sol = None, worst_sol = None, instance_name = None, best_fitness=None):
         self.best_sol = best_sol
         self.worst_sol = worst_sol
         self.instance_name = instance_name
         if best_sol is None:
-            self.best_fitness = None
+            self.best_fitness = best_fitness
         else:
             self.best_fitness = self.fitness_nosave(best_sol)
         if worst_sol is None:
@@ -18,7 +18,7 @@ class Problem:
         else:
             self.worst_fitness = self.fitness_nosave(worst_sol)
         self.reset()
-    
+
     def reset(self):
         self.evaluations = []
         self.solutions = []
@@ -37,7 +37,7 @@ class Problem:
         if self.best_sol is None:
             return np.nan
         return kendallTau(perm, self.best_sol) / (self.n * (self.n - 1) * 0.5)
-  
+
     # Returns a closure function that can be called from R.
     # WARNING: this function minimizes for CEGO
     def make_r_fitness(self):

### 2020-09-11 qap.py

@@ -16,12 +16,12 @@ class QAP(Problem):
             with open(opt_filename) as f:
                 for line in f:
                     name, best_sol = line.strip().split(" ")
-                    if filename.find(name) >= 0: 
+                    if filename.find(name) >= 0:
                         best_sol = np.fromstring(best_sol, dtype=int, sep=",")
                         best_sol -= 1 # 0-indexed
                         print(f"Reading best solution {best_sol} from {opt_filename}")
                         break
-        
+        # print(best_sol, type(best_sol))
         return QAP(A, B, instance_name = filename, best_sol = best_sol)
 
     def __init__(self, A, B, best_sol = None, worst_sol = None, instance_name = "(generated)"):
@@ -33,9 +33,8 @@ class QAP(Problem):
         self.n = self.A.shape[0]
         super().__init__(best_sol = best_sol, worst_sol = worst_sol,
                          instance_name = instance_name)
-        
+
     def fitness_nosave(self, x):
         # In case it is not numpy array.
         x = np.asarray(x, dtype=int)
         return np.sum(self.A * self.B[np.ix_(x, x)])
-    

### 2020-09-11 runner.py

@@ -1,3 +1,4 @@
+from imp import reload
 import time
 class Timer:
     def __init__(self):
@@ -20,7 +21,12 @@ def get_problem(instance_name):
         from qap import QAP
         return QAP
     elif "lop" in instance_name:
+        import lop as lop
+        reload(lop)
+
         from lop import LOP
+
+
         return LOP
     elif "pfsp" in instance_name:
         from pfsp import PFSP
@@ -31,6 +37,8 @@ def get_problem(instance_name):
 def run_once(algo_name, instance_name, seed, out_filename = None,
              **algo_params):
     if algo_name == "uMM":
+        import umm as umm
+        reload(umm)
         from umm import uMM
         algo = uMM
     elif algo_name == "CEGO":
@@ -38,11 +46,15 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
         algo = cego
     else:
         raise ValueError("Unknown algo: " + algo_name)
-    
+
     problem = get_problem(instance_name)
     instance = problem.read_instance(instance_name)
+    # return instance
+
+
     timer = Timer()
     df = algo(instance, seed, **algo_params)
+
     if instance.best_fitness is not None and instance.worst_fitness is not None:
         df['Fitness'] = (df.Fitness - instance.best_fitness) / (instance.worst_fitness - instance.best_fitness)
     df['Function evaluations'] = np.arange(1, len(df['Fitness'])+1)
@@ -54,4 +66,3 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
         df.to_csv(out_filename + '.csv.gz', index=False)
         df.to_pickle(out_filename + '.pkl.gz')
     return df
-

### 2020-09-11 umm.py

@@ -1,3 +1,4 @@
+from imp import reload
 import numpy as np
 import mallows_kendall as mk
 from scipy.spatial import distance
@@ -51,7 +52,7 @@ def uMM(instance,
     # ['rho','phi_estim','phi_sample','Distance']
     res = [ [np.nan, np.nan, np.nan,
              instance.distance_to_best(perm)] for perm in sample]
-    
+
     for m in range(budget - m_ini):
         ws = np.asarray(fitnesses).copy()
         ws = ws - ws.min()
@@ -60,8 +61,11 @@ def uMM(instance,
         co.sort()
         rho = binary_search_rho(co, ratio_samples_learn, weight_mass_learn)
 
-        # print(rho)
+        # print(fitnesses)
+        # print(ws)
         ws = rho ** ws #MINIMIZE
+        # print(ws)
+        # ws = rho ** (1-ws) #MAXIMIZE
         # print(ws,co[:int(len(co)/4)].sum(),co.sum())
 
         borda = mk.uborda(np.array(sample), ws)
@@ -78,16 +82,16 @@ def uMM(instance,
         dists = np.sort(dists, axis=1)
         indi = np.argmax(dists[:, 0]) #index of the perm with the farthest closest permutation. Maximizes the min dist to the sample
         perm = perms[indi]
-        # FIXME: This should already be an array of int type. 
+        # FIXME: This should already be an array of int type.
         perm = np.asarray(perm, dtype='int')
         sample.append(perm)
         fitnesses.append(instance.fitness(perm))
+        # print("Fitness, best, desde UMM.py", fitnesses[-1], instance.best_fitness)
+        # print(fitnesses,ws)
 
         # This is only used for reporting stats.
         res.append([rho, phi_estim, phi_sample, instance.distance_to_best(borda)])
-        #print(perm,fitnesses[-1])
-    df = pd.DataFrame(res,
-                      columns=['rho','phi_estim','phi_sample','Distance'])
+    df = pd.DataFrame(res, columns=['rho','phi_estim','phi_sample','Distance'])
     df['Fitness'] = fitnesses
     df['x'] = sample
     df['m_ini'] = m_ini

### 2020-09-11 paper/main.tex

@@
 widespread~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
-success, there have been attempts at adapting such surrogate models to the
-combinatorial case~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}.
+success, there have been attempts at adapting such Bayesian optimization
+algorithms to the combinatorial case, a notable example being Combinatorial
+Efficient Global Optimization
+(CEGO)~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}. However, the
+ruggedness of combinatorial landscapes, which makes local search particularly
+effective in the white-box context, lessens the effectiveness of global
+surrogate models~\citep{EriPeaGar2019scalable}. Moreover, surrogate models are
+expensive to train and optimize. Unless each function evaluation requires a day
+or more, the time required by the Bayesian optimizer itself may impose a
+significant overhead in computation time. \citet{PerLopStu2015si} recently
+showed that ant colony optimization (ACO)~\citep{DorStu2004:book} is
+competitive with CEGO on a black-box version of the travelling salesman problem
+under a budget of $1\,000$ function evaluations. ACO basically builds a
+probability distribution model from which solutions are sampled with a bias
+towards the best solutions evaluated so far. As such, ACO may be considered a
+type of estimation-of-distribution algorithm (EDA) for combinatorial
+optimization.
+
+In this work, we propose and analyze uMM, an EDA specifically designed for
+black-box combinatorial optimization under a limited budget. The proposed uMM \ldots\MANUEL{what is the inspiration}
+\
 
@@
 \end{itemize}
-
-\citet{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}
-\citet{PerLopStu2015si}
 
@@
 
-\newcommand{\FEmax}{\ensuremath{FE_{\max}}}
+\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}
 

### 2020-10-05 paper/main.tex
paper, section UMM

@@
 
-\section{Background}
+\section{Background}\label{sec:backgroud}
 
@@
 
+\subsection{Unbalanced Borda}\label{sec:uborda}
 Recently, it has been proposed an extension of Borda by the name of uBorda.
@@
 
-\section{Learning stage}
-In this section
+\section{UMM }
+In this section we describe our main contribution: a novel algorithm for the optimization of a black box function over permutation spaces.
+This algorithm is called Unbalanced Mallows model (UMM) since it a probabilistic method based on a well-known probability distribution for permutations: the Mallows model.
+
+The high-level behavior of UMM is as follows.
+It starts with an \textit{small} sample of $m'$ permutations that are randomly generated.
+UMM also keeps a probability distribution of a transformation of the sample.
+Then, for $m$ iterations, it generates a random, promising candidate solution and includes it in the sample.
+Each time a new permutation is included its fitness is computed and the internal probabilistic model is updated.
+
+The probabilistic model is a MM that assigns higher probability values to the permutations in which the fitness value is known or expected to be smaller.\footnote{We assume a minimization problem.} The main contribution of our approach is that the distribution does not represent the current sample considered by the algorithm but instead, represents a variation of it.
+This variation is abstractly equivalent to replicating in the sample the permutations inversely proportionally to their fitness values.
+One key property of UMM is the computational complexity of working with the original sample and the transformed sample is the same, i.e., polynomial.
+However, the question os how many times should we replicate each permutation in the sample?
+UMM includes a parameter $\rho$ that controls this point and will be detailed in Section~\ref{sec:learning}
+
+At each iteration of the algorithm the following steps take place: (1) randomly generate a permutation and include it in the sample and (2) evaluate the new permutation and update the probabilistic model.
+These steps correspond to two statistical operations: (1) sample a probability distribution and (2) learn a probability distribution of the transformation of the sample.
+Both are describe in the subsequent sections.
+
+\subsection{Learning step}\label{sec:learning}
+
+As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two steps (1) estimating the location parameter $\sigma_0$, usually done with Borda algorithm, and (2) computing the scale parameter $\theta$.
+Since the scale parameter is set automatically, we focus on learning the first step.
+
+%In order to minimize the number of permutations in the sample (and consequently the number of function evaluations) our proposed UMM learns the distribution of a transformation of the sample rather than the sample.
+This transformation consists on replicating the best fitted permutations.
+However, this is done abstractly:
+The location parameter can be shown to be the permutation in $S_n$ that minimizes the sum of the distances to the permutations in the sample~\cite{gMallows}, which is usually approximated with the Borda algorithm.
+Since we are interested on biasing the location towards the regions of the space where the fitness function is smaller, we would like to replicate the permutations a number of times that is inversely proportionally to their fitness values.
+We can simulate the replication of the best fitted individuals by using uBorda, described in Section~\ref{sec:uborda} and setting $w(\sigma)=\rho^{f(\sigma)}$.
+This generalization requires setting a value for parameter $\rho$.
+We show how to choose this value in the experimental section.
+
+\subsection{Sampling step}
+The Mallows model is defined by two parameters.
+The location $\sigma_0$ is the permutation with the largest probability value, and in our problem, the permutation that is supposed to be the global optimum.
+The scale parameter $\theta$ controls both the variance and the expected distance of a random permutation.
+Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{Irurozki2016b}\cite{Regenwetter2004} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
+
+At each iteration of UMM $\sigma_0$ is given by the permutation obtained as a solution in the previous learning step.
+The scale $\theta$, in the other hand, is set automatically by the algorithm to control the diversification and intensification trade off: At the first iteration it is set so that the expected distance equals  ${n\choose 4}$ (half of the distance of the uniform distribution), and then it linearly decrease until the last iteration where the expected distance equals 1.
+Similar approaches have been taken in different contexts~\cite{arza2019approaching}.
+
+This setting reflects the fact that, despite beliving that the best global optimum at each iteration is $\sigma_0$, the uncertainty at the fisrt stages of the algorithm is high and
+
+\subsection{Probavle convergence}
+
+\subsection{Comparison to standard probabilistic approaches.}
+UMM is a population based algorithm.
+There exists a wide body of research on population based probabilistic algorithms, where most prominent examples are Estimation of Distribution Algorithms (EDAs).
+Our approach drastically differs from EDAs since EDAs assume that function evaluations are cheap and they can do many of them.
+On the contrary, UMM minimizes the number of function evaluations.
+Moreover, UMM does not rely on an evolving sequence of samples (called generations in evolutionary approaches).
+Instead, UMM considers a single sample of permutations during the whole execution that is increased iteratively, with one permutation each iteration.
 

### 2020-10-09 paper/main.tex
umm real data:

@@
 The probabilistic model is a MM that assigns higher probability values to the permutations in which the fitness value is known or expected to be smaller.\footnote{We assume a minimization problem.} The main contribution of our approach is that the distribution does not represent the current sample considered by the algorithm but instead, represents a variation of it.
-This variation is abstractly equivalent to replicating in the sample the permutations inversely proportionally to their fitness values.
+This variation is abstractly equivalent to replicating in the sample the permutations inversely proportionally to their fitness values, i.e., there will be many copies of the good solutions and few of the bad ones.
 One key property of UMM is the computational complexity of working with the original sample and the transformed sample is the same, i.e., polynomial.
@@
 
-This setting reflects the fact that, despite beliving that the best global optimum at each iteration is $\sigma_0$, the uncertainty at the fisrt stages of the algorithm is high and
+This setting reflects the fact that, despite believing that the best global optimum at each iteration is $\sigma_0$, the uncertainty at the fisrt stages of the algorithm is high and
 
-\subsection{Probavle convergence}
-
-\subsection{Comparison to standard probabilistic approaches.}
+\subsection{Comparison to standard probabilistic optimization approaches.}
 UMM is a population based algorithm.

### 2020-10-14 cego.py
results-loplip

@@ -25,6 +25,141 @@ def cego(instance, seed, budget, # m: number of evaluations
     
     rstring = """
     library(CEGO)
+    # This is identical to the function in the package but takes also a maxTime parameter.
+optimCEGO <- function (x = NULL, fun, control = list()) 
+{
+    con <- list(evalInit = 2, vectorized = FALSE, verbosity = 0, 
+        plotting = FALSE, targetY = -Inf, budget = 100, creationRetries = 100, 
+        distanceFunction = distancePermutationHamming, creationFunction = solutionFunctionGeneratorPermutation(6), 
+        infill = infillExpectedImprovement, model = modelKriging, 
+        modelSettings = list(), optimizer = optimEA, optimizerSettings = list(), 
+        initialDesign = designMaxMinDist, archiveModelInfo = NULL, 
+        initialDesignSettings = list(), maxTime = 3600 * 24 * 3)
+    con[names(control)] <- control
+    control <- con
+    rm(con)
+    maxTime <- control$maxTime + proc.time()["elapsed"] 
+    count <- control$evalInit
+    archiveModelInfo <- control$archiveModelInfo
+    vectorized <- control$vectorized
+    verbosity <- control$verbosity
+    plotting <- control$plotting
+    creationFunction <- control$creationFunction
+    distanceFunction <- control$distanceFunction
+    if (is.null(control$initialDesignSettings$distanceFunction)) 
+        control$initialDesignSettings$distanceFunction <- distanceFunction
+    fun
+    if (!vectorized) 
+        fn <- function(x) unlist(lapply(x, fun))
+    else fn <- fun
+    res <- list(xbest = NA, ybest = NA, x = NA, y = NA, distances = NA, 
+        modelArchive = NA, count = count, convergence = 0, message = "")
+    msg <- "Termination message:"
+    res$x <- control$initialDesign(x, creationFunction, count, 
+        control$initialDesignSettings)
+    distanceHasParam <- FALSE
+    if (is.function(distanceFunction)) {
+        if (length(distanceFunction) == 1) 
+            distanceHasParam <- length(formalArgs(distanceFunction)) > 
+                2
+        else distanceHasParam <- any(sapply(sapply(distanceFunction, 
+            formalArgs, simplify = FALSE), length) > 2)
+        if (!distanceHasParam) 
+            res$distances <- CEGO:::distanceMatrixWrapper(res$x, distanceFunction)
+    }
+    res$y <- fn(res$x)
[…]

### 2020-10-14 paper/main.tex
:wq
figuras real-LOP CEGO

@@
 
-learn distri: uMM with param $\rho$ and binary search rho
-
-sample: peakig distri. Borda is the
-
-Quality guarantees
-
-prrof, probem-distri correlation
-
 \section{Experimental setup}
@@
 
+Te dejo aqui un comentario argumentando por qué he elegido estas instancias, puedes ver todas las graficas en el github
+
+LOP syntéticas, todas en una gráfica.
+
+LOP reales: he puesto 3 de diferentes tamanos.
+Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
+
+kra32 es el unico del QAP en el que hay un comportaiento no random (considera distancia euclidea)
+
+PFSF, rec05, en un problema muy pequeno ninguno hace nada
+
 \section{Conclusions}

### 2020-10-14 paper/main.tex
table paper included

@@
 
+\begin{tabular}{llrrr}
+\toprule
+Problem &                instance &  Fitness CEGO &  Fitness UMM &  Improvement \\
+\midrule
+    LOP &  lop/RandA1/N-t1d100.01 &       78696.6 &      76119.6 &     0.032746 \\
+    LOP &  lop/RandA1/N-t1d100.02 &       79686.5 &      76827.7 &     0.035876 \\
+    LOP &  lop/RandA1/N-t1d150.01 &      187605.5 &     179508.3 &     0.043161 \\
+    LOP &  lop/RandA1/N-t1d150.02 &      184320.9 &     177592.8 &     0.036502 \\
+    LOP &  lop/RandA1/N-t1d200.01 &      335391.1 &     323513.6 &     0.035414 \\
+    LOP &  lop/RandA1/N-t1d200.02 &      333149.5 &     319675.8 &     0.040443 \\
+    LOP &  lop/RandA2/N-t2d150.01 &       24682.7 &      15297.7 &     0.380226 \\
+    LOP &  lop/RandA2/N-t2d150.02 &       24250.8 &      14724.9 &     0.392808 \\
+   PFSP &          pfsp/rec05.txt &        1317.9 &       1330.7 &    -0.009712 \\
+   PFSP &          pfsp/rec13.txt &        2128.2 &       2158.7 &    -0.014331 \\
+   PFSP &          pfsp/rec19.txt &        2398.6 &       2405.1 &    -0.002710 \\
+    QAP &           qap/kra32.dat &      115907.0 &     116320.0 &    -0.003563 \\
+    QAP &           qap/nug12.dat &         624.2 &        660.4 &    -0.057994 \\
+    QAP &           qap/nug30.dat &        7486.8 &       7474.0 &     0.001710 \\
+    QAP &           qap/tho30.dat &      192416.0 &     191482.0 &     0.004854 \\
+\bottomrule
+\end{tabular}
+
 \section{Conclusions}

### 2020-10-15 paper/main.tex
update

@@
 surrogate models~\citep{EriPeaGar2019scalable}. Moreover, surrogate models are
-expensive to train and optimize. Unless each function evaluation requires a day
-or more, the time required by the Bayesian optimizer itself may impose a
-significant overhead in computation time. \citet{PerLopStu2015si} recently
-showed that ant colony optimization (ACO)~\citep{DorStu2004:book} is
-competitive with CEGO on a black-box version of the travelling salesman problem
-under a budget of $1\,000$ function evaluations. ACO basically builds a
-probability distribution model from which solutions are sampled with a bias
-towards the best solutions evaluated so far. As such, ACO may be considered a
-type of estimation-of-distribution algorithm (EDA) for combinatorial
-optimization.
+expensive to train and optimize. The additional time required by the Bayesian
+optimizer may impose a significant overhead in computation time, specially if
+computation time of each function evaluation is measured in ``few'' minutes or
+hours rather than days and when ``expensive'' refers to resources or economical
+cost rather than wall-clock time. \citet{PerLopStu2015si} recently showed that
+ant colony optimization (ACO)~\citep{DorStu2004:book} is competitive with CEGO
+on a black-box version of the travelling salesman problem under a budget of
+$1\,000$ function evaluations. ACO basically builds a probability distribution
+model from which solutions are sampled with a bias towards the best solutions
+evaluated so far. As such, ACO may be considered a type of
+estimation-of-distribution algorithm (EDA) for combinatorial optimization. On
+the other hand, $1\,000$ evaluations is a relatively large budget for CEGO and
+ACO was not designed for such short budgets.
 
@@
 
-Bayesian optimization methods using a global GP model, such as CEGO, are known
-to have trouble optimizing locally \citep{EriPeaGar2019scalable}. Our
-intuition is that this problem becomes worse in rugged combinatorial
-landscapes, where small steps may produce drastic changes.
-
-Te dejo aqui un comentario argumentando por qué he elegido estas instancias, puedes ver todas las graficas en el github
-
 LOP syntéticas, todas en una gráfica.
@@
 
+Table~\ref{tab:results} shows
+
+\begin{table}
+\caption{Results\label{tab:results}}
 \begin{tabular}{llrrr}
@@
 \end{tabular}
+\end{table}
 
 \section{Conclusions}
+
+Bayesian optimization methods using a global GP model, such as CEGO, are known
+to have trouble optimizing locally \citep{EriPeaGar2019scalable}. Our
+intuition is that this problem becomes worse in rugged combinatorial
+landscapes, where small steps may produce drastic changes.
 

### 2020-10-15 paper/main.tex
new files added

--- 2020-10-14 paper/main.tex
@@
 \usepackage{graphicx}
-\usepackage{pdflscape}
+\usepackage{pdflscape,adjustbox}
 \usepackage{multirow}
@@
 \usepackage{enumitem}
-\usepackage{afterpage}
+\usepackage{afterpage,tabularx}
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
@@
 
+Table~\ref{tab:results} shows
+
+\begin{adjustbox}{angle=90}
+\begin{tabular}{lllrrrrrrr}
+\toprule
+{} & Problem &                instance &  Fitness\_CEGO\_mean &  Fitness\_CEGO\_std &  Fitness\_UMM\_mean &  Fitness\_UMM\_std &  Improvement &  ttest statistic &  ttest pvalue \\
+\midrule
+0  &     LOP &  lop/RandA1/N-t1d100.01 &            78696.6 &        811.368145 &           76119.6 &       915.392836 &     0.032746 &         6.662091 &  2.989433e-06 \\
+1  &     LOP &  lop/RandA1/N-t1d100.02 &            79686.5 &        367.876262 &           76827.7 &       890.969391 &     0.035876 &         9.378614 &  2.374606e-08 \\
+2  &     LOP &  lop/RandA1/N-t1d150.01 &           187605.5 &       1906.941953 &          179508.3 &      1647.878437 &     0.043161 &        10.159726 &  6.994357e-09 \\
+3  &     LOP &  lop/RandA1/N-t1d150.02 &           184320.9 &        973.457920 &          177592.8 &      1196.159298 &     0.036502 &        13.795848 &  5.183671e-11 \\
+4  &     LOP &  lop/RandA1/N-t1d200.01 &           335391.1 &       2288.960096 &          323513.6 &      1371.235712 &     0.035414 &        14.076557 &  3.714220e-11 \\
+5  &     LOP &  lop/RandA1/N-t1d200.02 &           333149.5 &       2497.207474 &          319675.8 &      2242.631629 &     0.040443 &        12.694417 &  2.029610e-10 \\
+6  &     LOP &  lop/RandA2/N-t2d150.01 &            24682.7 &        976.639596 &           15297.7 &       563.527787 &     0.380226 &        26.320561 &  8.038269e-16 \\
+7  &     LOP &  lop/RandA2/N-t2d150.02 &            24250.8 &       1354.897110 &           14724.9 &       667.454276 &     0.392808 &        19.944373 &  1.010663e-13 \\
+8  &    PFSP &          pfsp/rec05.txt &             1317.9 &         25.075220 &            1330.7 &        17.901272 &    -0.009712 &        -1.313790 &  2.054160e-01 \\
+9  &    PFSP &          pfsp/rec13.txt &             2128.2 &         24.965754 &            2158.7 &        29.533597 &    -0.014331 &        -2.494040 &  2.258601e-02 \\
+10 &    PFSP &          pfsp/rec19.txt &             2398.6 &         19.839355 &            2405.1 &        12.582616 &    -0.002710 &        -0.874932 &  3.931359e-01 \\
+11 &     QAP &           qap/kra32.dat &           115907.0 &       4663.427566 &          116320.0 &      4673.468852 &    -0.003563 &        -0.197816 &  8.454055e-01 \\
+12 &     QAP &           qap/nug12.dat &              624.2 &         26.288147 &             660.4 &        17.276831 &    -0.057994 &        -3.639054 &  1.876810e-03 \\
+13 &     QAP &           qap/nug30.dat &             7486.8 &         40.411220 &            7474.0 &        83.910005 &     0.001710 &         0.434612 &  6.690098e-01 \\
+14 &     QAP &           qap/tho30.dat &           192416.0 &       2584.174401 &          191482.0 &      4419.722264 &     0.004854 &         0.576896 &  5.711548e-01 \\
+\bottomrule
+\end{tabular}
+\end{adjustbox}
+
 \section{Conclusions}
+
+Bayesian optimization methods using a global GP model, such as CEGO, are known
+to have trouble optimizing locally \citep{EriPeaGar2019scalable}. Our
+intuition is that this problem becomes worse in rugged combinatorial
+landscapes, where small steps may produce drastic changes.
 

### 2020-10-16 paper/main.tex
add plots to latex

@@
 
-LOP syntéticas, todas en una gráfica.
+Fig.~\ref{fig:lop_synth}: LOP syntéticas, todas en una gráfica.
 
-LOP reales: he puesto 3 de diferentes tamanos.
+Fig.~\ref{fig:loplib}: LOP reales: he puesto 3 de diferentes tamanos.
 Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
 
-kra32 es el unico del QAP en el que hay un comportaiento no random (considera distancia euclidea)
+kra32 es el unico del QAP en el que hay un comportaiento no random (considera distancia euclidea) Fig~\ref{fig:kra32}
 
-PFSF, rec05, en un problema muy pequeno ninguno hace nada
+PFSF, rec05, en un problema muy pequeno ninguno hace nada Fig.~\ref{fig:rec05}
 
@@
 
-\begin{adjustbox}{angle=90}
+\begin{table}
+  \caption{Results.\label{tab:results}}
+%\begin{adjustbox}
+\resizebox{\textwidth}{!}{%
 \begin{tabular}{lllrrrrrrr}
@@
 \bottomrule
-\end{tabular}
-\end{adjustbox}
+\end{tabular}}
+% \end{adjustbox}
+ \end{table}
+
+\begin{figure}
+  \centering%
+  \includegraphics[width=\textwidth]{img/synthetic_LOP_combined}
+  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
+\end{figure}
+
+\begin{figure}
+  \centering%
+  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d100_01}\\
+  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d150_01}\\
+  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d200_01}\\
+    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}
+  \end{figure}
+
+\begin{figure}
+  \centering%
+  \includegraphics[width=\textwidth]{img/fitness_real_pfsp_rec05_txt}
+  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
+\end{figure}
+\begin{figure}
+  \centering%
+  \includegraphics[width=\textwidth]{img/fitness_real_qap_kra32_dat}
+    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.\label{fig:kra32}}
+\end{figure}
 

### 2020-10-16 paper/main.tex
add plots to latex

@@
 
-Fig.~\ref{fig:lop_synth}: LOP syntéticas, todas en una gráfica.
+Ekhine: How many runs do we do? 10 20?
+
+To better understand the behaviour of the algorithms, we record the fitness of
+the solution evaluated at each step of each run and we average those values
+over X runs on the same problem instances.  As an example, Fig.~\ref{fig:kra32}
+shows the mean fitness of the solutions evaluated by uMM and CEGO on one QAP
+instance. The shaded area shows one standard deviation around the
+mean. Ideally, each solution evaluated helps to refine the model and leads to a
+better solution being evaluated in the next step.  However, the plot shows that
+the fitness of the solutions does not improve monotonically and new solutions
+are often worse than their predecessors. Nevertheless, in this example, there
+is a clear overall improvement as more evaluations are performed. The minimum
+values achieved in the plot up to a particular number of function evaluations
+gives an estimation of the mean fitness of the best-so-far solution up to that
+point in the run. The algorithms do not return the final solution shown but
+rather the best one (minimum fitness) of each run.
+
+When we compare the behaviour of CEGO and uMM on synthetic LOP instances (Fig.~\ref{fig:lop_synth}).
+
+ we 
+: LOP syntéticas, todas en una gráfica.
 

### 2020-10-16 paper/main.tex
results

@@
 selected elements of the permutation) and cycle crossover for permutations. The
-budget of each run of the GA is \textcolor{red}{$10^4$} evaluations using the
-surrogate-model. Although it is never stated in the original paper, the	`
-implementation of CEGO generates a set of initial solutions of size
-$\minit=10$ by means of a max-min-distance sequential design: new solutions
-are added to the set sequentially by maximizing the minimum distance to
-solutions already in the set. These initial solutions are then evaluated on the
-actual objective function and the result is used to build the initial surrogate
+authors of CEGO used a budget of $10^4$ evaluations using the surrogate-model
+for each run of the GA, however, with such setting, each run of CEGO requires
+more than a week of computation time, which was infeasible for us, hence, we
+set a value of \textcolor{red}{$10^3$}.
+
+Although it is never stated in the original paper, the implementation of CEGO
+generates a set of initial solutions of size $\minit=10$ by means of a
+max-min-distance sequential design: new solutions are added to the set
+sequentially by maximizing the minimum distance to solutions already in the
+set. These initial solutions are then evaluated on the actual objective
+function and the result is used to build the initial surrogate
 model.% \citet{ZaeStoBar2014:ppsn} uses 10
@@
 
-When we compare the behaviour of CEGO and uMM on synthetic LOP instances (Fig.~\ref{fig:lop_synth}).
+%: LOP syntéticas, todas en una gráfica. 
+\begin{figure}
+  \centering%
+  \includegraphics[width=\textwidth]{img/synthetic_LOP_combined}
+  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
+\end{figure}
 
- we 
-: LOP syntéticas, todas en una gráfica.
+When we compare the behaviour of CEGO and uMM on synthetic LOP instances
+(Fig.~\ref{fig:lop_synth}), we observe a clear pattern independently of the
+uniformity of the instances. In particular, CEGO starts with a random sampling
+of X solutions, with are quite poor as expected, followed by building and
+updating the Kriging model based new evaluations, which leads to an impressive
+improvement in less than 50 evaluations. However, progress is more or less
+halted at this point and the model seems to have trouble generating better
+solutions. Lower uniformity values (e.g., $\mu=5$) show worsening evaluations
+despite the increase in solutions used to build the model.
 
-Fig.~\ref{fig:loplib}: LOP reales: he puesto 3 de diferentes tamanos.
+On the other hand, uMM converges much more slowly than CEGO, only reaching the
+same fitness values around evaluation 250. Yet, uMM is able to keep the rate of
+improvement and completely overtake CEGO around evaluation 350. As in the case
+of CEGO, lower uniformity values lead to worse performance, although the
+difference is much smaller for uMM than it was for CEGO.
+
+In summary, the Kriging model clearly helps CEGO to quickly identify good
+solutions in very few evaluations. However, for some reason, CEGO is not able to
+further improve those solutions given more data. This may due to the
+model not being able to produce accurate predictions or the underlying
+optimizer not being able to find improved solutions using the model.\MANUEL{perhaps we should measure the prediction error and see if it grows or decreases. That would answer this question.}
+
+\newcommand{\supplement}{\citep{Supplementary}}
+
+%: LOP reales: he puesto 3 de diferentes tamanos.
 Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
+\begin{figure}
+  \centering%
+  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d100_01}\\
+  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d150_01}\\
+  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d200_01}\\
+    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}
+  \end{figure}
 
-kra32 es el unico del QAP en el que hay un comportaiento no random (considera distancia euclidea) Fig~\ref{fig:kra32}
+The fast initial convergence of CEGO for synthetic LOP instances does not seem
+to translate to the instances available in LOPLIB~\citep{}. We show in
+Fig.~\ref{fig:loplib} three examples with different permutation size, however,
+the results are consistent for other LOPLIB instances (complete results
+available as supplementary material \supplement). Due to extremely long
+runtimes of CEGO and constraints in our computing system, we set a maximum
+wall-clock time of 5 days for each run of CEGO. On these instances, the
+behaviour of CEGO and uMM is much more similar (at least up to the termination
+point of CEGO). It appears that our synthetic LOP instances have a fitness
+landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
+instances do not present such landscape. Comparing CEGO and uMM results, the
+differences get smaller, in favour of uMM, with larger instance size. Our
+conjecture is that building an accurate Bayesian model and searching for good
+solutions on it becomes harder for larger fitness landscapes.
 
-PFSF, rec05, en un problema muy pequeno ninguno hace nada Fig.~\ref{fig:rec05}
+In the case of the PFSP, none of the algorithms are able to achieve a
+  progress better than a random search would, as illustrated by the results
+  (Fig.~\ref{fig:rec05}) obtained on \texttt{rec05}, which is the smallest PFSP
+  instance considered here. The lack of any apparent convergence suggests that
+  the models (both Bayesian and probabilistic) are not learning anything about
+  the fitness landscape. Results for larger PFSP corroborate this conclusions
+  (\supplement). Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
+  that other distance measures are more suited than Kendall's-$\tau$ distance
+  for the PFSP, the reported fitness differences between various distance
+  metrics are small, hence, we conjecture that the same behaviour will be
+  observed with other distance metrics.
 
-Table~\ref{tab:results} shows
+\begin{figure}
+  \centering%
+  \includegraphics[width=\textwidth]{img/fitness_real_pfsp_rec05_txt}
+  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
+\end{figure}
+
+The situation is very similar for the QAP, where both algorithms show the same
+behaviour as in the PFSP for all instances tested, except for
+\texttt{kra32}. As shown in Fig~\ref{fig:kra32}, in this instance both
+algorithms are able to improve over the number of evaluations, although not
+with ease. As in the synthetic LOP instances, we again observe that CEGO does
+improve faster than UMM initially, yet, around evaluation 250, it appears to
+get stuck and around evaluation 300, UMM overtakes it. We believe the reason is
+the same as before, that is, the model is not able to adequately reflect the
+ruggedness of the actual landscape and, hence, the best solutions found by the
+GA searching the model are not improved solutions for the actual problem. Interestingly, \texttt{kra32} is rectangular
+whereas \texttt{tho32} contains rectangular distances and \texttt{nug12} and \texttt{nug30} contain Manhattan distances of rectangular grids.\MANUEL{so they all have the same distance type!!!}
+
+\begin{figure}
+  \centering%
+  \includegraphics[width=\textwidth]{img/fitness_real_qap_kra32_dat}
+    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.\label{fig:kra32}}
+\end{figure}
+
+In terms of best fitness achieved on each run, and taking into account the
+random behaviour observed in some instances and discussed above,
+Table~\ref{tab:results} shows\MANUEL{TODO}
 
@@
 
-\begin{figure}
-  \centering%
-  \includegraphics[width=\textwidth]{img/synthetic_LOP_combined}
-  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
-\end{figure}
-
-\begin{figure}
-  \centering%
-  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d100_01}\\
-  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d150_01}\\
-  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d200_01}\\
-    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}
-  \end{figure}
-
-\begin{figure}
-  \centering%
-  \includegraphics[width=\textwidth]{img/fitness_real_pfsp_rec05_txt}
-  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
-\end{figure}
-\begin{figure}
-  \centering%
-  \includegraphics[width=\textwidth]{img/fitness_real_qap_kra32_dat}
-    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.\label{fig:kra32}}
-\end{figure}
-
 \section{Conclusions}

### 2020-10-18 paper/main.tex
new figures and table

@@
 
-\begin{table}
-  \caption{Results.\label{tab:results}}
-%\begin{adjustbox}
+%\begin{table}
+%  \caption{Results.\label{tab:results}}
+%%\begin{adjustbox}
+%\resizebox{\textwidth}{!}{%
+%\begin{tabular}{lllrrrrrrr}
+%\toprule
+%{} & Problem &                instance &  Fitness\_CEGO\_mean &  Fitness\_CEGO\_std &  Fitness\_UMM\_mean &  Fitness\_UMM\_std &  Improvement &  ttest statistic &  ttest pvalue \\
+%\midrule
+%0  &     LOP &  lop/RandA1/N-t1d100.01 &            78696.6 &        811.368145 &           76119.6 &       915.392836 &     0.032746 &         6.662091 &  2.989433e-06 \\
+%1  &     LOP &  lop/RandA1/N-t1d100.02 &            79686.5 &        367.876262 &           76827.7 &       890.969391 &     0.035876 &         9.378614 &  2.374606e-08 \\
+%2  &     LOP &  lop/RandA1/N-t1d150.01 &           187605.5 &       1906.941953 &          179508.3 &      1647.878437 &     0.043161 &        10.159726 &  6.994357e-09 \\
+%3  &     LOP &  lop/RandA1/N-t1d150.02 &           184320.9 &        973.457920 &          177592.8 &      1196.159298 &     0.036502 &        13.795848 &  5.183671e-11 \\
+%4  &     LOP &  lop/RandA1/N-t1d200.01 &           335391.1 &       2288.960096 &          323513.6 &      1371.235712 &     0.035414 &        14.076557 &  3.714220e-11 \\
+%5  &     LOP &  lop/RandA1/N-t1d200.02 &           333149.5 &       2497.207474 &          319675.8 &      2242.631629 &     0.040443 &        12.694417 &  2.029610e-10 \\
+%6  &     LOP &  lop/RandA2/N-t2d150.01 &            24682.7 &        976.639596 &           15297.7 &       563.527787 &     0.380226 &        26.320561 &  8.038269e-16 \\
+%7  &     LOP &  lop/RandA2/N-t2d150.02 &            24250.8 &       1354.897110 &           14724.9 &       667.454276 &     0.392808 &        19.944373 &  1.010663e-13 \\
+%8  &    PFSP &          pfsp/rec05.txt &             1317.9 &         25.075220 &            1330.7 &        17.901272 &    -0.009712 &        -1.313790 &  2.054160e-01 \\
+%9  &    PFSP &          pfsp/rec13.txt &             2128.2 &         24.965754 &            2158.7 &        29.533597 &    -0.014331 &        -2.494040 &  2.258601e-02 \\
+%10 &    PFSP &          pfsp/rec19.txt &             2398.6 &         19.839355 &            2405.1 &        12.582616 &    -0.002710 &        -0.874932 &  3.931359e-01 \\
+%11 &     QAP &           qap/kra32.dat &           115907.0 &       4663.427566 &          116320.0 &      4673.468852 &    -0.003563 &        -0.197816 &  8.454055e-01 \\
+%12 &     QAP &           qap/nug12.dat &              624.2 &         26.288147 &             660.4 &        17.276831 &    -0.057994 &        -3.639054 &  1.876810e-03 \\
+%13 &     QAP &           qap/nug30.dat &             7486.8 &         40.411220 &            7474.0 &        83.910005 &     0.001710 &         0.434612 &  6.690098e-01 \\
+%14 &     QAP &           qap/tho30.dat &           192416.0 &       2584.174401 &          191482.0 &      4419.722264 &     0.004854 &         0.576896 &  5.711548e-01 \\
+%\bottomrule
+%\end{tabular}}
+%% \end{adjustbox}
+% \end{table}
+
 \resizebox{\textwidth}{!}{%
-\begin{tabular}{lllrrrrrrr}
+\begin{tabular}{llllllrrr}
 \toprule
-{} & Problem &                instance &  Fitness\_CEGO\_mean &  Fitness\_CEGO\_std &  Fitness\_UMM\_mean &  Fitness\_UMM\_std &  Improvement &  ttest statistic &  ttest pvalue \\
+{} & Problem &     Instance &        UMM fitness &       CEGO fitness &                        Confidence interval &  UMM Runtime &  CEGO Runtime &  CEGO F. evaluations \\
 \midrule
-0  &     LOP &  lop/RandA1/N-t1d100.01 &            78696.6 &        811.368145 &           76119.6 &       915.392836 &     0.032746 &         6.662091 &  2.989433e-06 \\
-1  &     LOP &  lop/RandA1/N-t1d100.02 &            79686.5 &        367.876262 &           76827.7 &       890.969391 &     0.035876 &         9.378614 &  2.374606e-08 \\
-2  &     LOP &  lop/RandA1/N-t1d150.01 &           187605.5 &       1906.941953 &          179508.3 &      1647.878437 &     0.043161 &        10.159726 &  6.994357e-09 \\
-3  &     LOP &  lop/RandA1/N-t1d150.02 &           184320.9 &        973.457920 &          177592.8 &      1196.159298 &     0.036502 &        13.795848 &  5.183671e-11 \\
-4  &     LOP &  lop/RandA1/N-t1d200.01 &           335391.1 &       2288.960096 &          323513.6 &      1371.235712 &     0.035414 &        14.076557 &  3.714220e-11 \\
-5  &     LOP &  lop/RandA1/N-t1d200.02 &           333149.5 &       2497.207474 &          319675.8 &      2242.631629 &     0.040443 &        12.694417 &  2.029610e-10 \\
-6  &     LOP &  lop/RandA2/N-t2d150.01 &            24682.7 &        976.639596 &           15297.7 &       563.527787 &     0.380226 &        26.320561 &  8.038269e-16 \\
-7  &     LOP &  lop/RandA2/N-t2d150.02 &            24250.8 &       1354.897110 &           14724.9 &       667.454276 &     0.392808 &        19.944373 &  1.010663e-13 \\
-8  &    PFSP &          pfsp/rec05.txt &             1317.9 &         25.075220 &            1330.7 &        17.901272 &    -0.009712 &        -1.313790 &  2.054160e-01 \\
-9  &    PFSP &          pfsp/rec13.txt &             2128.2 &         24.965754 &            2158.7 &        29.533597 &    -0.014331 &        -2.494040 &  2.258601e-02 \\
-10 &    PFSP &          pfsp/rec19.txt &             2398.6 &         19.839355 &            2405.1 &        12.582616 &    -0.002710 &        -0.874932 &  3.931359e-01 \\
-11 &     QAP &           qap/kra32.dat &           115907.0 &       4663.427566 &          116320.0 &      4673.468852 &    -0.003563 &        -0.197816 &  8.454055e-01 \\
-12 &     QAP &           qap/nug12.dat &              624.2 &         26.288147 &             660.4 &        17.276831 &    -0.057994 &        -3.639054 &  1.876810e-03 \\
-13 &     QAP &           qap/nug30.dat &             7486.8 &         40.411220 &            7474.0 &        83.910005 &     0.001710 &         0.434612 &  6.690098e-01 \\
-14 &     QAP &           qap/tho30.dat &           192416.0 &       2584.174401 &          191482.0 &      4419.722264 &     0.004854 &         0.576896 &  5.711548e-01 \\
+0  &     LOP &  N-t1d100.01 &    76119.6 (915.4) &    78696.6 (811.4) &   (1763.4904574743996, 3390.5095425256004) &      11436.2 &      260954.5 &                111.8 \\
+1  &     LOP &  N-t1d100.02 &    76827.7 (891.0) &    79686.5 (367.9) &    (2194.5410893080507, 3523.058910691955) &      11353.3 &      262421.6 &                110.7 \\
+2  &     LOP &  N-t1d150.01 &  179508.3 (1647.9) &  187605.5 (1906.9) &     (6420.259663264915, 9774.140336735109) &      23395.6 &      261712.4 &                 73.9 \\
+3  &     LOP &  N-t1d150.02 &  177592.8 (1196.2) &   184320.9 (973.5) &     (5700.461476832616, 7755.738523167395) &      23361.2 &      263389.9 &                 74.1 \\
+4  &     LOP &  N-t1d200.01 &  323513.6 (1371.2) &  335391.1 (2289.0) &     (10076.07401989604, 13678.92598010396) &      38476.8 &      263467.5 &                 55.6 \\
+5  &     LOP &  N-t1d200.02 &  319675.8 (2242.6) &  333149.5 (2497.2) &   (11241.971680405779, 15705.428319594244) &      38268.3 &      265217.5 &                 56.5 \\
+6  &     LOP &  N-t2d150.01 &    15297.7 (563.5) &    24682.7 (976.6) &    (8622.206776092156, 10147.793223907844) &      25318.6 &      262324.3 &                 74.6 \\
+7  &     LOP &  N-t2d150.02 &    14724.9 (667.5) &   24250.8 (1354.9) &    (8495.057408450935, 10556.742591549064) &      25425.6 &      262079.2 &                 74.8 \\
+8  &     LOP &  N-t2d200.01 &   33465.2 (2001.0) &   57098.6 (4452.8) &   (20284.508914101127, 26982.291085898876) &      40340.9 &      263908.9 &                 56.3 \\
+9  &     LOP &  N-t2d200.02 &   35901.8 (2077.5) &   58123.0 (2334.9) &     (9515.181030361815, 34927.21896963818) &      40153.1 &      261723.6 &                 55.0 \\
+10 &    PFSP &    rec05.txt &      1330.7 (17.9) &      1317.9 (25.1) &  (-33.424760001214985, 7.8247600012150755) &        506.1 &      139893.0 &                400.0 \\
+11 &    PFSP &    rec13.txt &      2158.7 (29.5) &      2128.2 (25.0) &   (-56.24362499249925, -4.756375007500754) &        506.8 &      140167.2 &                400.0 \\
+12 &    PFSP &    rec19.txt &      2405.1 (12.6) &      2398.6 (19.8) &   (-22.313869704790818, 9.313869704790818) &       1078.3 &      295442.7 &                400.0 \\
+13 &     QAP &    kra32.dat &  116320.0 (4673.5) &  115907.0 (4663.4) &    (-4799.295389697205, 3973.295389697205) &       1194.8 &      344766.2 &                400.0 \\
+14 &     QAP &    nug12.dat &       660.4 (17.3) &       624.2 (26.3) &  (-57.337505259978144, -15.06249474002172) &        180.0 &       53149.3 &                400.0 \\
+15 &     QAP &    nug30.dat &      7474.0 (83.9) &      7486.8 (40.4) &     (-50.84533806724974, 76.4453380672501) &       1073.5 &      303352.6 &                400.0 \\
+16 &     QAP &    tho30.dat &  191482.0 (4419.7) &  192416.0 (2584.2) &   (-2527.0213832396516, 4395.021383239651) &       1066.5 &      306451.3 &                400.0 \\
 \bottomrule
-\end{tabular}}
-% \end{adjustbox}
- \end{table}
+\end{tabular}
+}
 

### 2020-10-18 paper/main.tex
update table

@@
 
-%\begin{table}
-%  \caption{Results.\label{tab:results}}
 %%\begin{adjustbox}
@@
 
+    \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
+\begin{table}
+  \caption{Results.\label{tab:results}}
   \resizebox{\textwidth}{!}{%
-\begin{tabular}{llllllrrr}
+\begin{tabular}{*{6}{r}c*{3}{r}}
   \toprule
-{} & Problem &     Instance &        UMM fitness &       CEGO fitness &                        Confidence interval &  UMM Runtime &  CEGO Runtime &  CEGO F. evaluations \\
-\midrule
-0  &     LOP &  N-t1d100.01 &    76119.6 (915.4) &    78696.6 (811.4) &   (1763.4904574743996, 3390.5095425256004) &      11436.2 &      260954.5 &                111.8 \\
-1  &     LOP &  N-t1d100.02 &    76827.7 (891.0) &    79686.5 (367.9) &    (2194.5410893080507, 3523.058910691955) &      11353.3 &      262421.6 &                110.7 \\
-2  &     LOP &  N-t1d150.01 &  179508.3 (1647.9) &  187605.5 (1906.9) &     (6420.259663264915, 9774.140336735109) &      23395.6 &      261712.4 &                 73.9 \\
-3  &     LOP &  N-t1d150.02 &  177592.8 (1196.2) &   184320.9 (973.5) &     (5700.461476832616, 7755.738523167395) &      23361.2 &      263389.9 &                 74.1 \\
-4  &     LOP &  N-t1d200.01 &  323513.6 (1371.2) &  335391.1 (2289.0) &     (10076.07401989604, 13678.92598010396) &      38476.8 &      263467.5 &                 55.6 \\
-5  &     LOP &  N-t1d200.02 &  319675.8 (2242.6) &  333149.5 (2497.2) &   (11241.971680405779, 15705.428319594244) &      38268.3 &      265217.5 &                 56.5 \\
-6  &     LOP &  N-t2d150.01 &    15297.7 (563.5) &    24682.7 (976.6) &    (8622.206776092156, 10147.793223907844) &      25318.6 &      262324.3 &                 74.6 \\
-7  &     LOP &  N-t2d150.02 &    14724.9 (667.5) &   24250.8 (1354.9) &    (8495.057408450935, 10556.742591549064) &      25425.6 &      262079.2 &                 74.8 \\
-8  &     LOP &  N-t2d200.01 &   33465.2 (2001.0) &   57098.6 (4452.8) &   (20284.508914101127, 26982.291085898876) &      40340.9 &      263908.9 &                 56.3 \\
-9  &     LOP &  N-t2d200.02 &   35901.8 (2077.5) &   58123.0 (2334.9) &     (9515.181030361815, 34927.21896963818) &      40153.1 &      261723.6 &                 55.0 \\
-10 &    PFSP &    rec05.txt &      1330.7 (17.9) &      1317.9 (25.1) &  (-33.424760001214985, 7.8247600012150755) &        506.1 &      139893.0 &                400.0 \\
-11 &    PFSP &    rec13.txt &      2158.7 (29.5) &      2128.2 (25.0) &   (-56.24362499249925, -4.756375007500754) &        506.8 &      140167.2 &                400.0 \\
-12 &    PFSP &    rec19.txt &      2405.1 (12.6) &      2398.6 (19.8) &   (-22.313869704790818, 9.313869704790818) &       1078.3 &      295442.7 &                400.0 \\
-13 &     QAP &    kra32.dat &  116320.0 (4673.5) &  115907.0 (4663.4) &    (-4799.295389697205, 3973.295389697205) &       1194.8 &      344766.2 &                400.0 \\
-14 &     QAP &    nug12.dat &       660.4 (17.3) &       624.2 (26.3) &  (-57.337505259978144, -15.06249474002172) &        180.0 &       53149.3 &                400.0 \\
-15 &     QAP &    nug30.dat &      7474.0 (83.9) &      7486.8 (40.4) &     (-50.84533806724974, 76.4453380672501) &       1073.5 &      303352.6 &                400.0 \\
-16 &     QAP &    tho30.dat &  191482.0 (4419.7) &  192416.0 (2584.2) &   (-2527.0213832396516, 4395.021383239651) &       1066.5 &      306451.3 &                400.0 \\
+            &                  & \mcolc{4}{Mean fitness (sd)} &                 & \mcolc{2}{Mean runtime}             & \bf     Mean FE\\\cmidrule{3-6}\cmidrule(lr){8-9}
+\bf Problem & \bf     Instance & \mcolc{2}{UMM}               & \mcolc{2}{CEGO} & \bf 95\% CI for the mean difference & \bf UMM  & \bf CEGO                                  & \bf CEGO           \\\midrule
+    LOP     & N-t1d100.01      & 76119.6                      & (915.4)         & 78696.6                             & (811.4)  & (1763.4904574743996, 3390.5095425256004)  & 11436.2 & 260954.5 & 111.8 \\
+    LOP     & N-t1d100.02      & 76827.7                      & (891.0)         & 79686.5                             & (367.9)  & (2194.5410893080507, 3523.058910691955)   & 11353.3 & 262421.6 & 110.7 \\
+    LOP     & N-t1d150.01      & 179508.3                     & (1647.9)        & 187605.5                            & (1906.9) & (6420.259663264915, 9774.140336735109)    & 23395.6 & 261712.4 & 73.9  \\
+    LOP     & N-t1d150.02      & 177592.8                     & (1196.2)        & 184320.9                            & (973.5)  & (5700.461476832616, 7755.738523167395)    & 23361.2 & 263389.9 & 74.1  \\
+    LOP     & N-t1d200.01      & 323513.6                     & (1371.2)        & 335391.1                            & (2289.0) & (10076.07401989604, 13678.92598010396)    & 38476.8 & 263467.5 & 55.6  \\
+    LOP     & N-t1d200.02      & 319675.8                     & (2242.6)        & 333149.5                            & (2497.2) & (11241.971680405779, 15705.428319594244)  & 38268.3 & 265217.5 & 56.5  \\
+    LOP     & N-t2d150.01      & 15297.7                      & (563.5)         & 24682.7                             & (976.6)  & (8622.206776092156, 10147.793223907844)   & 25318.6 & 262324.3 & 74.6  \\
+    LOP     & N-t2d150.02      & 14724.9                      & (667.5)         & 24250.8                             & (1354.9) & (8495.057408450935, 10556.742591549064)   & 25425.6 & 262079.2 & 74.8  \\
+    LOP     & N-t2d200.01      & 33465.2                      & (2001.0)        & 57098.6                             & (4452.8) & (20284.508914101127, 26982.291085898876)  & 40340.9 & 263908.9 & 56.3  \\
+    LOP     & N-t2d200.02      & 35901.8                      & (2077.5)        & 58123.0                             & (2334.9) & (9515.181030361815, 34927.21896963818)    & 40153.1 & 261723.6 & 55.0  \\\midrule
+   PFSP     & rec05            & 1330.7                       & (17.9)          & 1317.9                              & (25.1)   & (-33.424760001214985, 7.8247600012150755) & 506.1   & 139893.0 & 400.0 \\
+   PFSP     & rec13            & 2158.7                       & (29.5)          & 2128.2                              & (25.0)   & (-56.24362499249925, -4.756375007500754)  & 506.8   & 140167.2 & 400.0 \\
+   PFSP     & rec19            & 2405.1                       & (12.6)          & 2398.6                              & (19.8)   & (-22.313869704790818, 9.313869704790818)  & 1078.3  & 295442.7 & 400.0 \\\midrule
+    QAP     & kra32            & 116320.0                     & (4673.5)        & 115907.0                            & (4663.4) & (-4799.295389697205, 3973.295389697205)   & 1194.8  & 344766.2 & 400.0 \\
+    QAP     & nug12            & 660.4                        & (17.3)          & 624.2                               & (26.3)   & (-57.337505259978144, -15.06249474002172) & 180.0   & 53149.3  & 400.0 \\
+    QAP     & nug30            & 7474.0                       & (83.9)          & 7486.8                              & (40.4)   & (-50.84533806724974, 76.4453380672501)    & 1073.5  & 303352.6 & 400.0 \\
+    QAP     & tho30            & 191482.0                     & (4419.7)        & 192416.0                            & (2584.2) & (-2527.0213832396516, 4395.021383239651)  & 1066.5  & 306451.3 & 400.0 \\
 \bottomrule
-\end{tabular}
-}
+\end{tabular}}
+\end{table}
 

### 2020-10-19 paper/main.tex
new figures added

--- 2020-10-16 paper/main.tex
@@
 
-Ekhine: How many runs do we do? 10 20?
+Ekhine: How many runs do we do? 10 20? 10, cambio el texto
 
@@
 the solution evaluated at each step of each run and we average those values
-over X runs on the same problem instances.  As an example, Fig.~\ref{fig:kra32}
+over 10 runs on the same problem instances.  As an example, Fig.~\ref{fig:kra32} \textcolor{red}{por qué no mepezamos con la fig1?}
 shows the mean fitness of the solutions evaluated by uMM and CEGO on one QAP
@@
 mean. Ideally, each solution evaluated helps to refine the model and leads to a
-better solution being evaluated in the next step.  However, the plot shows that
+better solution being evaluated in the next step.  However, the plot shows that \textcolor{red}{yo diría solo que lo esperable es que baje monotonically, sin entrar a valorar un ejemplo concreto}
 the fitness of the solutions does not improve monotonically and new solutions
@@
 solutions on it becomes harder for larger fitness landscapes. 
+\textcolor{red}{Parece que el paper va del CEGO en vez del UMM :p.
+Yo diría que tanto UMM como CEGO encuentran soluciones que mejoran según avanzan, pero con diferencias de comportamiento: (1) las los optimos de UMM son '3 veces mejor', y esto yo creo que es importante, aunque en gran parte se debe a que (2) UMM hace 400 ejecuciones en 3, 6 y 11 horas y CEGO entre 50 y 130 en 5 dias (3) cego empeora segun n aumenta (y n=200 tampoco es enorme)}
 
@@
 
-\begin{table}
-  \caption{Results.\label{tab:results}}
-%\begin{adjustbox}
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{lllrrrrrrr}
+%
+%    \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
+%\begin{table}
+%  \caption{Results.\label{tab:results}}
+%  \resizebox{\textwidth}{!}{%
+%\begin{tabular}{*{6}{r}c*{3}{r}}
+%  \toprule
+%            &                  & \mcolc{4}{Mean fitness (sd)} &                 & \mcolc{2}{Mean runtime}             & \bf     Mean FE\\\cmidrule{3-6}\cmidrule(lr){8-9}
+%\bf Problem & \bf     Instance & \mcolc{2}{UMM}               & \mcolc{2}{CEGO} & \bf 95\% CI for the mean difference & \bf UMM  & \bf CEGO                                  & \bf CEGO           \\\midrule
+%    LOP     & N-t1d100.01      & 76119.6                      & (915.4)         & 78696.6                             & (811.4)  & (1763.4904574743996, 3390.5095425256004)  & 11436.2 & 260954.5 & 111.8 \\
+%    LOP     & N-t1d100.02      & 76827.7                      & (891.0)         & 79686.5                             & (367.9)  & (2194.5410893080507, 3523.058910691955)   & 11353.3 & 262421.6 & 110.7 \\
+%    LOP     & N-t1d150.01      & 179508.3                     & (1647.9)        & 187605.5                            & (1906.9) & (6420.259663264915, 9774.140336735109)    & 23395.6 & 261712.4 & 73.9  \\
+%    LOP     & N-t1d150.02      & 177592.8                     & (1196.2)        & 184320.9                            & (973.5)  & (5700.461476832616, 7755.738523167395)    & 23361.2 & 263389.9 & 74.1  \\
+%    LOP     & N-t1d200.01      & 323513.6                     & (1371.2)        & 335391.1                            & (2289.0) & (10076.07401989604, 13678.92598010396)    & 38476.8 & 263467.5 & 55.6  \\
+%    LOP     & N-t1d200.02      & 319675.8                     & (2242.6)        & 333149.5                            & (2497.2) & (11241.971680405779, 15705.428319594244)  & 38268.3 & 265217.5 & 56.5  \\
+%    LOP     & N-t2d150.01      & 15297.7                      & (563.5)         & 24682.7                             & (976.6)  & (8622.206776092156, 10147.793223907844)   & 25318.6 & 262324.3 & 74.6  \\
+%    LOP     & N-t2d150.02      & 14724.9                      & (667.5)         & 24250.8                             & (1354.9) & (8495.057408450935, 10556.742591549064)   & 25425.6 & 262079.2 & 74.8  \\
+%    LOP     & N-t2d200.01      & 33465.2                      & (2001.0)        & 57098.6                             & (4452.8) & (20284.508914101127, 26982.291085898876)  & 40340.9 & 263908.9 & 56.3  \\
+%    LOP     & N-t2d200.02      & 35901.8                      & (2077.5)        & 58123.0                             & (2334.9) & (9515.181030361815, 34927.21896963818)    & 40153.1 & 261723.6 & 55.0  \\\midrule
+%   PFSP     & rec05            & 1330.7                       & (17.9)          & 1317.9                              & (25.1)   & (-33.424760001214985, 7.8247600012150755) & 506.1   & 139893.0 & 400.0 \\
+%   PFSP     & rec13            & 2158.7                       & (29.5)          & 2128.2                              & (25.0)   & (-56.24362499249925, -4.756375007500754)  & 506.8   & 140167.2 & 400.0 \\
+%   PFSP     & rec19            & 2405.1                       & (12.6)          & 2398.6                              & (19.8)   & (-22.313869704790818, 9.313869704790818)  & 1078.3  & 295442.7 & 400.0 \\\midrule
+%    QAP     & kra32            & 116320.0                     & (4673.5)        & 115907.0                            & (4663.4) & (-4799.295389697205, 3973.295389697205)   & 1194.8  & 344766.2 & 400.0 \\
+%    QAP     & nug12            & 660.4                        & (17.3)          & 624.2                               & (26.3)   & (-57.337505259978144, -15.06249474002172) & 180.0   & 53149.3  & 400.0 \\
+%    QAP     & nug30            & 7474.0                       & (83.9)          & 7486.8                              & (40.4)   & (-50.84533806724974, 76.4453380672501)    & 1073.5  & 303352.6 & 400.0 \\
+%    QAP     & tho30            & 191482.0                     & (4419.7)        & 192416.0                            & (2584.2) & (-2527.0213832396516, 4395.021383239651)  & 1066.5  & 306451.3 & 400.0 \\
+%\bottomrule
+%\end{tabular}}
+%\end{table}
+
+  \resizebox{\textwidth}{!}{
+\begin{tabular}{lllllrrr}
 \toprule
-{} & Problem &                instance &  Fitness\_CEGO\_mean &  Fitness\_CEGO\_std &  Fitness\_UMM\_mean &  Fitness\_UMM\_std &  Improvement &  ttest statistic &  ttest pvalue \\
+Problem &       Instance &           UMM fitness &           CEGO fitness &     Confidence interval &  UMM Runtime &  CEGO Runtime &  CEGO F. evaluations \\
 \midrule
-0  &     LOP &  lop/RandA1/N-t1d100.01 &            78696.6 &        811.368145 &           76119.6 &       915.392836 &     0.032746 &         6.662091 &  2.989433e-06 \\
-1  &     LOP &  lop/RandA1/N-t1d100.02 &            79686.5 &        367.876262 &           76827.7 &       890.969391 &     0.035876 &         9.378614 &  2.374606e-08 \\
-2  &     LOP &  lop/RandA1/N-t1d150.01 &           187605.5 &       1906.941953 &          179508.3 &      1647.878437 &     0.043161 &        10.159726 &  6.994357e-09 \\
-3  &     LOP &  lop/RandA1/N-t1d150.02 &           184320.9 &        973.457920 &          177592.8 &      1196.159298 &     0.036502 &        13.795848 &  5.183671e-11 \\
-4  &     LOP &  lop/RandA1/N-t1d200.01 &           335391.1 &       2288.960096 &          323513.6 &      1371.235712 &     0.035414 &        14.076557 &  3.714220e-11 \\
-5  &     LOP &  lop/RandA1/N-t1d200.02 &           333149.5 &       2497.207474 &          319675.8 &      2242.631629 &     0.040443 &        12.694417 &  2.029610e-10 \\
-6  &     LOP &  lop/RandA2/N-t2d150.01 &            24682.7 &        976.639596 &           15297.7 &       563.527787 &     0.380226 &        26.320561 &  8.038269e-16 \\
-7  &     LOP &  lop/RandA2/N-t2d150.02 &            24250.8 &       1354.897110 &           14724.9 &       667.454276 &     0.392808 &        19.944373 &  1.010663e-13 \\
-8  &    PFSP &          pfsp/rec05.txt &             1317.9 &         25.075220 &            1330.7 &        17.901272 &    -0.009712 &        -1.313790 &  2.054160e-01 \\
-9  &    PFSP &          pfsp/rec13.txt &             2128.2 &         24.965754 &            2158.7 &        29.533597 &    -0.014331 &        -2.494040 &  2.258601e-02 \\
-10 &    PFSP &          pfsp/rec19.txt &             2398.6 &         19.839355 &            2405.1 &        12.582616 &    -0.002710 &        -0.874932 &  3.931359e-01 \\
-11 &     QAP &           qap/kra32.dat &           115907.0 &       4663.427566 &          116320.0 &      4673.468852 &    -0.003563 &        -0.197816 &  8.454055e-01 \\
-12 &     QAP &           qap/nug12.dat &              624.2 &         26.288147 &             660.4 &        17.276831 &    -0.057994 &        -3.639054 &  1.876810e-03 \\
-13 &     QAP &           qap/nug30.dat &             7486.8 &         40.411220 &            7474.0 &        83.910005 &     0.001710 &         0.434612 &  6.690098e-01 \\
-14 &     QAP &           qap/tho30.dat &           192416.0 &       2584.174401 &          191482.0 &      4419.722264 &     0.004854 &         0.576896 &  5.711548e-01 \\
+    LOP &    N-t1d100.01 &       76119.6 (915.4) &        78696.6 (811.4) &        (1763.5, 3390.5) &          3.2 &          72.5 &                111.8 \\
+    LOP &    N-t1d100.02 &       76827.7 (891.0) &        79686.5 (367.9) &        (2194.5, 3523.1) &          3.2 &          72.9 &                110.7 \\
+    LOP &    N-t1d150.01 &     179508.3 (1647.9) &      187605.5 (1906.9) &        (6420.3, 9774.1) &          6.5 &          72.7 &                 73.9 \\
+    LOP &    N-t1d150.02 &     177592.8 (1196.2) &       184320.9 (973.5) &        (5700.5, 7755.7) &          6.5 &          73.2 &                 74.1 \\
+    LOP &    N-t1d200.01 &     323513.6 (1371.2) &      335391.1 (2289.0) &      (10076.1, 13678.9) &         10.7 &          73.2 &                 55.6 \\
+    LOP &    N-t1d200.02 &     319675.8 (2242.6) &      333149.5 (2497.2) &      (11242.0, 15705.4) &         10.6 &          73.7 &                 56.5 \\
+    LOP &    N-t2d150.01 &       15297.7 (563.5) &        24682.7 (976.6) &       (8622.2, 10147.8) &          7.0 &          72.9 &                 74.6 \\
+    LOP &    N-t2d150.02 &       14724.9 (667.5) &       24250.8 (1354.9) &       (8495.1, 10556.7) &          7.1 &          72.8 &                 74.8 \\
+    LOP &    N-t2d200.01 &      33465.2 (2001.0) &       57098.6 (4452.8) &      (20284.5, 26982.3) &         11.2 &          73.3 &                 56.3 \\
+    LOP &    N-t2d200.02 &      35901.8 (2077.5) &       54895.6 (4589.7) &      (15539.1, 22448.5) &         11.2 &          73.5 &                 55.6 \\
+    LOP &       N-p50-01 &       21027.6 (711.8) &        21431.2 (299.5) &         (-128.1, 935.3) &          0.9 &          72.3 &                219.9 \\
+    LOP &       N-p50-02 &       21613.6 (596.4) &        22021.1 (275.6) &          (-42.5, 857.5) &          0.9 &          72.3 &                217.3 \\
+    LOP &       N-atp111 &          594.0 (15.4) &           693.1 (10.3) &           (86.7, 111.5) &          4.1 &          72.5 &                 98.7 \\
+    LOP &       N-atp134 &          756.3 (21.3) &           880.5 (18.6) &          (105.4, 143.0) &          5.7 &          72.8 &                 83.0 \\
+    LOP &  N-be75eec\_150 &   1415208.5 (58593.5) &    1666193.1 (72310.8) &    (188960.6, 313008.6) &          6.9 &          72.9 &                 74.0 \\
+    LOP &  N-be75eec\_250 &  4431271.0 (109756.7) &   5214413.0 (166701.4) &    (649040.8, 917243.2) &         15.5 &          73.7 &                 44.9 \\
+    LOP &   N-be75np\_150 &   3204878.3 (80704.1) &    3742444.0 (88500.2) &    (457944.2, 617187.2) &          6.9 &          72.9 &                 74.2 \\
+    LOP &   N-be75np\_250 &  9544754.9 (260523.9) &  10887427.8 (159642.5) &  (1136635.8, 1548710.0) &         15.5 &          74.0 &                 45.3 \\
+   PFSP &      rec05.txt &         1330.7 (17.9) &          1317.9 (25.1) &            (-33.4, 7.8) &          0.1 &          38.9 &                400.0 \\
+   PFSP &      rec13.txt &         2158.7 (29.5) &          2128.2 (25.0) &           (-56.2, -4.8) &          0.1 &          38.9 &                400.0 \\
+   PFSP &      rec19.txt &         2405.1 (12.6) &          2398.6 (19.8) &            (-22.3, 9.3) &          0.3 &          82.1 &                400.0 \\
+    QAP &      kra32.dat &     116320.0 (4673.5) &      115907.0 (4663.4) &       (-4799.3, 3973.3) &          0.3 &          95.8 &                400.0 \\
+    QAP &      nug12.dat &          660.4 (17.3) &           624.2 (26.3) &          (-57.3, -15.1) &          0.0 &          14.8 &                400.0 \\
+    QAP &      nug30.dat &         7474.0 (83.9) &          7486.8 (40.4) &           (-50.8, 76.4) &          0.3 &          84.3 &                400.0 \\
+    QAP &      tho30.dat &     191482.0 (4419.7) &      192416.0 (2584.2) &       (-2527.0, 4395.0) &          0.3 &          85.1 &                400.0 \\
 \bottomrule
 \end{tabular}}
-% \end{adjustbox}
- \end{table}
 
@@
 \bibliographystyle{splncs04nat}
-\bibliography{optbib/abbrev,optbib/authors,optbib/journals,optbib/biblio,optbib/crossref}
+\bibliography{optbib/abbrev,optbib/authors,optbib/journals,optbib/biblio,optbib/crossref,./mendeley}
 

### 2020-10-19 paper/main.tex
table

@@
 %: LOP syntéticas, todas en una gráfica. 
-\begin{figure}
+\begin{figure}[tb]
   \centering%
@@
 
-  \resizebox{\textwidth}{!}{
-\begin{tabular}{lllllrrr}
+   \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
+\begin{table}
+ \caption{Results.\label{tab:results}}
+ \resizebox{\textwidth}{!}{%
+\begin{tabular}{*{6}{r}rl*{3}{r}}
  \toprule
-Problem &       Instance &           UMM fitness &           CEGO fitness &     Confidence interval &  UMM Runtime &  CEGO Runtime &  CEGO F. evaluations \\
-\midrule
-    LOP &    N-t1d100.01 &       76119.6 (915.4) &        78696.6 (811.4) &        (1763.5, 3390.5) &          3.2 &          72.5 &                111.8 \\
-    LOP &    N-t1d100.02 &       76827.7 (891.0) &        79686.5 (367.9) &        (2194.5, 3523.1) &          3.2 &          72.9 &                110.7 \\
-    LOP &    N-t1d150.01 &     179508.3 (1647.9) &      187605.5 (1906.9) &        (6420.3, 9774.1) &          6.5 &          72.7 &                 73.9 \\
-    LOP &    N-t1d150.02 &     177592.8 (1196.2) &       184320.9 (973.5) &        (5700.5, 7755.7) &          6.5 &          73.2 &                 74.1 \\
-    LOP &    N-t1d200.01 &     323513.6 (1371.2) &      335391.1 (2289.0) &      (10076.1, 13678.9) &         10.7 &          73.2 &                 55.6 \\
-    LOP &    N-t1d200.02 &     319675.8 (2242.6) &      333149.5 (2497.2) &      (11242.0, 15705.4) &         10.6 &          73.7 &                 56.5 \\
-    LOP &    N-t2d150.01 &       15297.7 (563.5) &        24682.7 (976.6) &       (8622.2, 10147.8) &          7.0 &          72.9 &                 74.6 \\
-    LOP &    N-t2d150.02 &       14724.9 (667.5) &       24250.8 (1354.9) &       (8495.1, 10556.7) &          7.1 &          72.8 &                 74.8 \\
-    LOP &    N-t2d200.01 &      33465.2 (2001.0) &       57098.6 (4452.8) &      (20284.5, 26982.3) &         11.2 &          73.3 &                 56.3 \\
-    LOP &    N-t2d200.02 &      35901.8 (2077.5) &       54895.6 (4589.7) &      (15539.1, 22448.5) &         11.2 &          73.5 &                 55.6 \\
-    LOP &       N-p50-01 &       21027.6 (711.8) &        21431.2 (299.5) &         (-128.1, 935.3) &          0.9 &          72.3 &                219.9 \\
-    LOP &       N-p50-02 &       21613.6 (596.4) &        22021.1 (275.6) &          (-42.5, 857.5) &          0.9 &          72.3 &                217.3 \\
-    LOP &       N-atp111 &          594.0 (15.4) &           693.1 (10.3) &           (86.7, 111.5) &          4.1 &          72.5 &                 98.7 \\
-    LOP &       N-atp134 &          756.3 (21.3) &           880.5 (18.6) &          (105.4, 143.0) &          5.7 &          72.8 &                 83.0 \\
-    LOP &  N-be75eec\_150 &   1415208.5 (58593.5) &    1666193.1 (72310.8) &    (188960.6, 313008.6) &          6.9 &          72.9 &                 74.0 \\
-    LOP &  N-be75eec\_250 &  4431271.0 (109756.7) &   5214413.0 (166701.4) &    (649040.8, 917243.2) &         15.5 &          73.7 &                 44.9 \\
-    LOP &   N-be75np\_150 &   3204878.3 (80704.1) &    3742444.0 (88500.2) &    (457944.2, 617187.2) &          6.9 &          72.9 &                 74.2 \\
-    LOP &   N-be75np\_250 &  9544754.9 (260523.9) &  10887427.8 (159642.5) &  (1136635.8, 1548710.0) &         15.5 &          74.0 &                 45.3 \\
-   PFSP &      rec05.txt &         1330.7 (17.9) &          1317.9 (25.1) &            (-33.4, 7.8) &          0.1 &          38.9 &                400.0 \\
-   PFSP &      rec13.txt &         2158.7 (29.5) &          2128.2 (25.0) &           (-56.2, -4.8) &          0.1 &          38.9 &                400.0 \\
-   PFSP &      rec19.txt &         2405.1 (12.6) &          2398.6 (19.8) &            (-22.3, 9.3) &          0.3 &          82.1 &                400.0 \\
-    QAP &      kra32.dat &     116320.0 (4673.5) &      115907.0 (4663.4) &       (-4799.3, 3973.3) &          0.3 &          95.8 &                400.0 \\
-    QAP &      nug12.dat &          660.4 (17.3) &           624.2 (26.3) &          (-57.3, -15.1) &          0.0 &          14.8 &                400.0 \\
-    QAP &      nug30.dat &         7474.0 (83.9) &          7486.8 (40.4) &           (-50.8, 76.4) &          0.3 &          84.3 &                400.0 \\
-    QAP &      tho30.dat &     191482.0 (4419.7) &      192416.0 (2584.2) &       (-2527.0, 4395.0) &          0.3 &          85.1 &                400.0 \\
+           &                  & \mcolc{4}{Mean fitness (sd)} &  \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI for the\\mean difference}}}         & \mcolc{2}{Mean runtime}             & \bf \multirow[b]{2.45}{*}{\shortstack{Mean FE\\CEGO}}\\\cmidrule{3-6}\cmidrule(lr){9-10}
+\bf Problem & \bf     Instance & \mcolc{2}{UMM}               & \mcolc{2}{CEGO} & & & \bf UMM  & \bf CEGO                                  & \bf            \\\midrule
+    LOP &    N-t1d100.01 &       76119.6&(915.4) &        78696.6&(811.4) &       $[$1763.5,&3390.5$]$ &          3.2 &          72.5 &                111.8 \\
+    LOP &    N-t1d100.02 &       76827.7&(891.0) &        79686.5&(367.9) &       $[$2194.5,&3523.1$]$ &          3.2 &          72.9 &                110.7 \\
+    LOP &    N-t1d150.01 &     179508.3&(1647.9) &      187605.5&(1906.9) &       $[$6420.3,&9774.1$]$ &          6.5 &          72.7 &                 73.9 \\
+    LOP &    N-t1d150.02 &     177592.8&(1196.2) &       184320.9&(973.5) &       $[$5700.5,&7755.7$]$ &          6.5 &          73.2 &                 74.1 \\
+    LOP &    N-t1d200.01 &     323513.6&(1371.2) &      335391.1&(2289.0) &     $[$10076.1,&13678.9$]$ &         10.7 &          73.2 &                 55.6 \\
+    LOP &    N-t1d200.02 &     319675.8&(2242.6) &      333149.5&(2497.2) &     $[$11242.0,&15705.4$]$ &         10.6 &          73.7 &                 56.5 \\
+    LOP &    N-t2d150.01 &       15297.7&(563.5) &        24682.7&(976.6) &      $[$8622.2,&10147.8$]$ &          7.0 &          72.9 &                 74.6 \\
+    LOP &    N-t2d150.02 &       14724.9&(667.5) &       24250.8&(1354.9) &      $[$8495.1,&10556.7$]$ &          7.1 &          72.8 &                 74.8 \\
+    LOP &    N-t2d200.01 &      33465.2&(2001.0) &       57098.6&(4452.8) &     $[$20284.5,&26982.3$]$ &         11.2 &          73.3 &                 56.3 \\
+    LOP &    N-t2d200.02 &      35901.8&(2077.5) &       54895.6&(4589.7) &     $[$15539.1,&22448.5$]$ &         11.2 &          73.5 &                 55.6 \\
+    LOP &       N-p50-01 &       21027.6&(711.8) &        21431.2&(299.5) &        $[$-128.1,&935.3$]$ &          0.9 &          72.3 &                219.9 \\
+    LOP &       N-p50-02 &       21613.6&(596.4) &        22021.1&(275.6) &         $[$-42.5,&857.5$]$ &          0.9 &          72.3 &                217.3 \\
+    LOP &       N-atp111 &          594.0&(15.4) &           693.1&(10.3) &          $[$86.7,&111.5$]$ &          4.1 &          72.5 &                 98.7 \\
+    LOP &       N-atp134 &          756.3&(21.3) &           880.5&(18.6) &         $[$105.4,&143.0$]$ &          5.7 &          72.8 &                 83.0 \\
+    LOP &  N-be75eec\_150 &   1415208.5&(58593.5) &    1666193.1&(72310.8) &   $[$188960.6,&313008.6$]$ &          6.9 &          72.9 &                 74.0 \\
+    LOP &  N-be75eec\_250 &  4431271.0&(109756.7) &   5214413.0&(166701.4) &   $[$649040.8,&917243.2$]$ &         15.5 &          73.7 &                 44.9 \\
+    LOP &   N-be75np\_150 &   3204878.3&(80704.1) &    3742444.0&(88500.2) &   $[$457944.2,&617187.2$]$ &          6.9 &          72.9 &                 74.2 \\
+    LOP &   N-be75np\_250 &  9544754.9&(260523.9) &  10887427.8&(159642.5) & $[$1136635.8,&1548710.0$]$ &         15.5 &          74.0 &                 45.3 \\
+   PFSP &      rec05.txt &         1330.7&(17.9) &          1317.9&(25.1) &           $[$-33.4,&7.8$]$ &          0.1 &          38.9 &                400.0 \\
+   PFSP &      rec13.txt &         2158.7&(29.5) &          2128.2&(25.0) &          $[$-56.2,&-4.8$]$ &          0.1 &          38.9 &                400.0 \\
+   PFSP &      rec19.txt &         2405.1&(12.6) &          2398.6&(19.8) &           $[$-22.3,&9.3$]$ &          0.3 &          82.1 &                400.0 \\
+    QAP &      kra32.dat &     116320.0&(4673.5) &      115907.0&(4663.4) &      $[$-4799.3,&3973.3$]$ &          0.3 &          95.8 &                400.0 \\
+    QAP &      nug12.dat &          660.4&(17.3) &           624.2&(26.3) &         $[$-57.3,&-15.1$]$ &          0.0 &          14.8 &                400.0 \\
+    QAP &      nug30.dat &         7474.0&(83.9) &          7486.8&(40.4) &          $[$-50.8,&76.4$]$ &          0.3 &          84.3 &                400.0 \\
+    QAP &      tho30.dat &     191482.0&(4419.7) &      192416.0&(2584.2) &      $[$-2527.0,&4395.0$]$ &          0.3 &          85.1 &                400.0 \\
 \bottomrule
 \end{tabular}}
+\end{table}
 

### 2020-10-19 paper/main.tex
3 days
reply to comments

@@
 \newrevisor{manuel}{MANUEL}{Purple}
+\newrevisor{ekhine}{EKHINE}{red}
 
@@
 
-Ekhine: How many runs do we do? 10 20? 10, cambio el texto
-
 To better understand the behaviour of the algorithms, we record the fitness of
 the solution evaluated at each step of each run and we average those values
-over 10 runs on the same problem instances.  As an example, Fig.~\ref{fig:kra32} \textcolor{red}{por qué no mepezamos con la fig1?}
-shows the mean fitness of the solutions evaluated by uMM and CEGO on one QAP
+over 10 runs on the same problem instances.  As an example, each line in Fig.~\ref{fig:lop_synth} 
+shows the mean fitness of the solutions evaluated by either uMM or CEGO on one synthetic LOP
 instance. The shaded area shows one standard deviation around the
-mean. Ideally, each solution evaluated helps to refine the model and leads to a
-better solution being evaluated in the next step.  However, the plot shows that \textcolor{red}{yo diría solo que lo esperable es que baje monotonically, sin entrar a valorar un ejemplo concreto}
-the fitness of the solutions does not improve monotonically and new solutions
+mean. Ideally, each new evaluation will monotonically improve in fitness, since each solution evaluated helps to refine the model and leads to a
+better solution being evaluated in the next step.
+However, it is often the case that  \EKHINE{yo diría solo que lo esperable es que baje monotonically, sin entrar a valorar un ejemplo concreto}\MANUEL{Así está mejor?
+Quiero llamar la atención porque normalmente se muestra la evolución de la mejor solución encontrada pero nosotros estamos mostrando el fitness de cada solución evaluada}
+% the fitness of the solutions does not improve monotonically and
+new solutions
 are often worse than their predecessors. Nevertheless, in this example, there
@@
 solutions. Lower uniformity values (e.g., $\mu=5$) show worsening evaluations
-despite the increase in solutions used to build the model.
+despite the increase in solutions used to build the model. In summary, the
+Kriging model clearly helps CEGO to quickly identify good solutions in very few
+evaluations. However, for some reason, CEGO is not able to further improve
+those solutions given more data. This may due to the model not being able to
+produce accurate predictions or the underlying optimizer not being able to find
+improved solutions using the model.\MANUEL{perhaps we should measure the
+  prediction error and see if it grows or decreases. That would answer this
+  question.}
 
@@
 of CEGO, lower uniformity values lead to worse performance, although the
-difference is much smaller for uMM than it was for CEGO.
-
-In summary, the Kriging model clearly helps CEGO to quickly identify good
-solutions in very few evaluations. However, for some reason, CEGO is not able to
-further improve those solutions given more data. This may due to the
-model not being able to produce accurate predictions or the underlying
-optimizer not being able to find improved solutions using the model.\MANUEL{perhaps we should measure the prediction error and see if it grows or decreases. That would answer this question.}
+difference is much smaller for uMM than it was for CEGO.\MANUEL{If there is anything else we can say here, please go ahead and say it}
 
@@
 Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
-\begin{figure}
+\begin{figure}[tb]
   \centering%
@@
 
-The fast initial convergence of CEGO for synthetic LOP instances does not seem
-to translate to the instances available in LOPLIB~\citep{}. We show in
+Results are quite different on the 
+%The fast initial convergence of CEGO for synthetic LOP instances does not seem
+%to translate to the
+instances available in LOPLIB~\citep{}. We show in
 Fig.~\ref{fig:loplib} three examples with different permutation size, however,
@@
 runtimes of CEGO and constraints in our computing system, we set a maximum
-wall-clock time of 5 days for each run of CEGO. On these instances, the
+wall-clock time of 3 days for each run of CEGO. On these instances, the
 behaviour of CEGO and uMM is much more similar (at least up to the termination
@@
 solutions on it becomes harder for larger fitness landscapes.
-\textcolor{red}{Parece que el paper va del CEGO en vez del UMM :p.
+\EKHINE{Parece que el paper va del CEGO en vez del UMM :p.
 Yo diría que tanto UMM como CEGO encuentran soluciones que mejoran según avanzan, pero con diferencias de comportamiento: (1) las los optimos de UMM son '3 veces mejor', y esto yo creo que es importante, aunque en gran parte se debe a que (2) UMM hace 400 ejecuciones en 3, 6 y 11 horas y CEGO entre 50 y 130 en 5 dias (3) cego empeora segun n aumenta (y n=200 tampoco es enorme)}
+\MANUEL{Eso es porque puedo interpretar los resultados de CEGO mejor que los de UMM. Si crees que hay alguna interpretación o conjetura sobre UMM que podemos mencionar aquí, sería bueno mencionarla.}
+\MANUEL{Prefiero dejar el comentario sobre las mejores soluciones encontradas y el tiempo requerido para cuando hable de la tabla, así se puede ver todo junto}
 
-In the case of the PFSP, none of the algorithms are able to achieve a
-  progress better than a random search would, as illustrated by the results
+In the case of the PFSP, none of the algorithms are able to achieve a progress
+better than a random search would, as illustrated by the results
 (Fig.~\ref{fig:rec05}) obtained on \texttt{rec05}, which is the smallest PFSP
@@
 (\supplement). Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
-  that other distance measures are more suited than Kendall's-$\tau$ distance
-  for the PFSP, the reported fitness differences between various distance
-  metrics are small, hence, we conjecture that the same behaviour will be
-  observed with other distance metrics.
+that other distance measures are more suited than Kendall's-$\tau$ distance for
+the PFSP, the reported fitness differences between various distance metrics are
+small, hence, we conjecture that the same behaviour will be observed with other
+distance metrics.
 
-\begin{figure}
+\begin{figure}[tb]
   \centering%
@@
 GA searching the model are not improved solutions for the actual problem. Interestingly, \texttt{kra32} is rectangular
-whereas \texttt{tho32} contains rectangular distances and \texttt{nug12} and \texttt{nug30} contain Manhattan distances of rectangular grids.\MANUEL{so they all have the same distance type!!!}
+whereas \texttt{tho32} contains rectangular distances and \texttt{nug12} and \texttt{nug30} contain Manhattan distances of rectangular grids.\MANUEL{so they all have the same distance type!!!
+What is the difference then?
+Ask Thomas?}
 

### 2020-10-20 paper/main.tex
Sections 2, 3

@@
 \usepackage{multirow}
+\usepackage[]{algorithm2e}
 \usepackage{booktabs} % For formal tables
@@
 The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$.
-We denote permutations with Greek letters with exception of the inverse permutation denoted as $e=1, 2, 3, \ldots,n$.
+We denote permutations with Greek letters with exception of the identity permutation denoted as $e=1, 2, 3, \ldots,n$.
 We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma_{-1}=e$ always holds.
 
-Distributions over permutations are functions that assign a probability value to each of the permutations in $S_n$, $p(\sigma)\in[0,1]$ \cite{critchlow91}.
+Distributions over permutations~\cite{critchlow91} are functions that assign a probability value to each of the permutations in $S_n$, $p(\sigma)\in[0,1]$.
 One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
@@
 \end{equation}
-
-with two parameters, $\theta$ and $\sigma_0$, 
-where permutation $\sigma_0$ a reference permutation that has the largest probability value, i.e., the mode of the distribution.
+with two parameters, $\theta$ and $\sigma_0$, Permutation $\sigma_0$ a reference permutation that has the largest probability value, i.e., the mode of the distribution.
 The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and $\theta$, the dispersion parameter controls this decay.
@@
 
-One of the most common problems related ti probability distributions is that of learning the maximum likelihood parameters  given a sample of data $S$.
-For the MM, this problem tranlates to learning $\theta$ and $\sigma_0$ that best describe a sample of permutations.
+The learning process is divide in two stages: first, we estimate the central permutation of the distribution, $\hat\sigma0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{Irurozki2016b}.
 
-The learning process is divide in two stages: first, we estimate the central permutation of the distribution, $\hat\sigma0$ and, second, compute the dispersion parameter, $\hat\theta$.
+The exact maximum likelihood estimator (MLE) for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{Dwork:2001:RAM:371920.372165}.
+Unfortunately,  obteinig such ranking is NP-hard~\cite{Dwork:2001:RAM:371920.372165}.
+An alternative to the Kemeny ranking is the Borda ranking~\cite{Ali2011}, which has several advantages.
+First, Borda requires polynomial computational time.
+Second, Borda as an estimator for the MM is guarantied to obtain high quality parameters~\cite{Caragiannis2013}.
+Finally, in a general optimization perspective, Borda error is bounded~\cite{Coppersmith:2010}.
 
-The exact maximum likelihood estimation is computationally hard~\cite{Dwork:2001:RAM:371920.372165}.
-However, the approximate learning requires polynomilal computational time and is guarantied to obtain high quality parameters~\cite{Caragiannis2013,Coppersmith:2010}.
-
-This process of sample $S$ is as follows: first, compute $\hat\sigma0$ with the Borda count algorithm.
-Borda orders the items $[n]$ by their \textit{Borda score} increasingly, where the Borda score $B(i)$ is the average of each position $i$, $B(i) \propto \sum_{t\in S}  \sigma_t(i)$.
-Second, the computation of $\theta$ is casted as a numerical optimization problem~\cite{Irurozki2016b}.
+The Borda ranking of sample $S$ is computed as follows: first, for each item $i$ compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
+Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.
 
 \subsection{Unbalanced Borda}\label{sec:uborda}
-Recently, it has been proposed an extension of Borda by the name of uBorda.
-uBorda considers a sample of permutations $S$ along with a weight $w(\sigma)$ for each permutation $\sigma$~\cite{}.
-Intuitively it is equivalent to replicating the permutations in the sample proportionally to their weight.
-It has been used to learn a MM in an evolving preference context.
-In this paper, we propose to use uBorda where each permutation in the sample $\sigma\in S$ is weighted by its fitness function $w(\sigma)=f(\sigma)$.
+In these lines we introduce the core for our probabilisitc algorithm, the uBorda algorithm.
+Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
+The uBorda algorithm is a recent generalization of Borda~\cite{}.
+In uBorda, the score of item $i$ is defined as follows:
+
+\begin{equation}
+\begin{split}
+B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i),
+\end{split}
+\label{eq:uborda_score}
+\end{equation}
+and, the same as Borda, orders the items increasingly by their score.
+
+Intuitively, the uBorda ranking of sample $S$ is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.
+In this paper, we would like to replicate each permutation $\sigma$ a number of times that is proportional to $f(\sigma)$.
+Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
+
+By setting $\rho=1$ the original Borda is recovered.
+However, for $\rho<1$ uBorda ranking will be closer to the rankings in $S$ with smaller value of $f(\sigma)$.
+Interestingly, Borda is computed in polynomial time.
+Clearly, the choice of the weight depends on the relevant property of a ranking in a particular domain.
+For example, it has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift.
 
@@
 
-\section{Methods}
+\subsection{Previous method for black box optimization in combinatorial spaces: CEGO}
 
@@
 
-\section{UMM }
+\section{Unbalanced Mallows model: probabilistic BB optimization}
 In this section we describe our main contribution: a novel algorithm for the optimization of a black box function over permutation spaces.
@@
 
-The high-level behavior of UMM is as follows.
-It starts with an \textit{small} sample of $m'$ permutations that are randomly generated.
-UMM also keeps a probability distribution of a transformation of the sample.
-Then, for $m$ iterations, it generates a random, promising candidate solution and includes it in the sample.
-Each time a new permutation is included its fitness is computed and the internal probabilistic model is updated.
+\begin{algorithm}[t]
+ \KwData{$n$, $m$,}
+ \KwResult{how to write algorithm with \LaTeX2e }
+ $S$\;
+ $F$\;
+ $\hat\sigma_0$ = uBorda($S, F,\rho$)\;
+ \For{$m$ \text{iterations}}{
+  linear decrease on $\theta$\;
+  $\sigma$ = sample from a $MM(\hat\sigma_0, \theta)$\;
+  Append $\sigma$ and $f(\sigma)$ to $S$ and $F$\;
+   $\hat\sigma_0$ = uBorda($S, F,\rho$)\;
+ }
+ \Return $S, F$
+ \caption{Unbalanced Mallows model for black box optimization}
+ \label{alg:umm}
+\end{algorithm}
 
-The probabilistic model is a MM that assigns higher probability values to the permutations in which the fitness value is known or expected to be smaller.\footnote{We assume a minimization problem.} The main contribution of our approach is that the distribution does not represent the current sample considered by the algorithm but instead, represents a variation of it.
-This variation is abstractly equivalent to replicating in the sample the permutations inversely proportionally to their fitness values, i.e., there will be many copies of the good solutions and few of the bad ones.
+The high-level behavior of UMM is shown in Algorithm~\ref{alg:umm}.
+It starts with an \textit{small} sample of $m'$ permutations, $S=\{\sigma_1, \ldots,  \sigma_{m'}\}$ that are randomly generated.
+UMM also keeps the function evaluation of each solution, $F=\{f(\sigma_1), \ldots, f(\sigma_{m'})\}$.
+An probabilistic model is used internally to guide the search.
+Then, for $m$ iterations\EKHINE{esto es exactamente así o es $m-m'$?}, UMM generates a random, promising candidate solution $\sigma$ and includes it in the sample $S$ while $f(\sigma)$ in included in $F$.
+At each iteration, the internal probabilistic model is updated.
+
+The probabilistic model is a MM that assigns higher probability values to the permutations in which the fitness value is known or expected to be smaller.\footnote{We assume a minimization problem.} The main contribution of our approach is that the distribution does not represent the current sample $S$ considered by the algorithm but instead, represents a variation of it.
+This variation is based on the uBorda algorithm in Section~\ref{sec:uBorda} and is abstractly equivalent to replicating in the sample the permutations inversely proportionally to their fitness values, i.e., there will be many copies of the good solutions and few of the bad ones.
 One key property of UMM is the computational complexity of working with the original sample and the transformed sample is the same, i.e., polynomial.
 However, the question os how many times should we replicate each permutation in the sample?
-UMM includes a parameter $\rho$ that controls this point and will be detailed in Section~\ref{sec:learning}
+UMM includes a parameter $\rho$ that controls this point and will be detailed in Section~\ref{sec:learning}.
 
-At each iteration of the algorithm the following steps take place: (1) randomly generate a permutation and include it in the sample and (2) evaluate the new permutation and update the probabilistic model.
+At each iteration of the algorithm the following steps take place: (1) randomly generate a permutation distributed according to the internal model and include it in the sample and (2) evaluate the new permutation and update the probabilistic model.
 These steps correspond to two statistical operations: (1) sample a probability distribution and (2) learn a probability distribution of the transformation of the sample.
@@
 
-As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two steps (1) estimating the location parameter $\sigma_0$, usually done with Borda algorithm, and (2) computing the scale parameter $\theta$.
+As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two steps (1) estimating the location parameter $\sigma_0$ and (2) computing the scale parameter $\theta$.
 Since the scale parameter is set automatically, we focus on learning the first step.
@@
 However, this is done abstractly:
-The location parameter can be shown to be the permutation in $S_n$ that minimizes the sum of the distances to the permutations in the sample~\cite{gMallows}, which is usually approximated with the Borda algorithm.
-Since we are interested on biasing the location towards the regions of the space where the fitness function is smaller, we would like to replicate the permutations a number of times that is inversely proportionally to their fitness values.
+The location parameter of a MM is usually approximated with the Borda algorithm.
+Since we are interested on biasing the model towards the regions of the space where the fitness function is smaller, we would like to replicate the permutations a number of times that is inversely proportionally to their fitness values.
 We can simulate the replication of the best fitted individuals by using uBorda, described in Section~\ref{sec:uborda} and setting $w(\sigma)=\rho^{f(\sigma)}$.
@@
 \subsection{Sampling step}
-The Mallows model is defined by two parameters.
-The location $\sigma_0$ is the permutation with the largest probability value, and in our problem, the permutation that is supposed to be the global optimum.
-The scale parameter $\theta$ controls both the variance and the expected distance of a random permutation.
+Sampling can be done efficiently~\cite{Irurozki2016b} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$.
 Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{Irurozki2016b}\cite{Regenwetter2004} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
 
-At each iteration of UMM $\sigma_0$ is given by the permutation obtained as a solution in the previous learning step.
+At each iteration of UMM $\sigma_0$ is the permutation obtained as a solution in the previous learning step,~\ref{sec:learning}.
 The scale $\theta$, in the other hand, is set automatically by the algorithm to control the diversification and intensification trade off: At the first iteration it is set so that the expected distance equals  ${n\choose 4}$ (half of the distance of the uniform distribution), and then it linearly decrease until the last iteration where the expected distance equals 1.
-Similar approaches have been taken in different contexts~\cite{arza2019approaching}.
+Similar approaches consisting of changing the variance to control the exploration-explotation tradeoff have been taken in different contexts~\cite{arza2019approaching}.
 
@@
 Instead, UMM considers a single sample of permutations during the whole execution that is increased iteratively, with one permutation each iteration.
+\ekhine{bayesian optim}{ff}
+\EKHINE{subir a arxiv uBorda}
 

### 2020-10-21 paper/main.tex
table

@@
 the solution evaluated at each step of each run and we average those values
-over 10 runs on the same problem instances.  As an example, each line in Fig.~\ref{fig:lop_synth} 
+over 10 runs with different random seeds on the same problem instance.  As an example, each line in Fig.~\ref{fig:lop_synth} 
 shows the mean fitness of the solutions evaluated by either uMM or CEGO on one synthetic LOP
@@
 
+Table~\ref{tab:results} shows overall results over all instances evaluated (excluding the synthetic LOP instances).
+In particular, the mean fitness (and standard deviation) of the best solution found at the end of each run,
+\MANUEL{is the CI for CEGO $-$ UMM or UMM $-$ CEGO? I would prefer}
 In terms of best fitness achieved on each run, and taking into account the
 random behaviour observed in some instances and discussed above,
-Table~\ref{tab:results} shows\MANUEL{TODO}
-
-%
-%    \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
-%\begin{table}
-%  \caption{Results.\label{tab:results}}
-%  \resizebox{\textwidth}{!}{%
-%\begin{tabular}{*{6}{r}c*{3}{r}}
-%  \toprule
-%            &                  & \mcolc{4}{Mean fitness (sd)} &                 & \mcolc{2}{Mean runtime}             & \bf     Mean FE\\\cmidrule{3-6}\cmidrule(lr){8-9}
-%\bf Problem & \bf     Instance & \mcolc{2}{UMM}               & \mcolc{2}{CEGO} & \bf 95\% CI for the mean difference & \bf UMM  & \bf CEGO                                  & \bf CEGO           \\\midrule
-%    LOP     & N-t1d100.01      & 76119.6                      & (915.4)         & 78696.6                             & (811.4)  & (1763.4904574743996, 3390.5095425256004)  & 11436.2 & 260954.5 & 111.8 \\
-%    LOP     & N-t1d100.02      & 76827.7                      & (891.0)         & 79686.5                             & (367.9)  & (2194.5410893080507, 3523.058910691955)   & 11353.3 & 262421.6 & 110.7 \\
-%    LOP     & N-t1d150.01      & 179508.3                     & (1647.9)        & 187605.5                            & (1906.9) & (6420.259663264915, 9774.140336735109)    & 23395.6 & 261712.4 & 73.9  \\
-%    LOP     & N-t1d150.02      & 177592.8                     & (1196.2)        & 184320.9                            & (973.5)  & (5700.461476832616, 7755.738523167395)    & 23361.2 & 263389.9 & 74.1  \\
-%    LOP     & N-t1d200.01      & 323513.6                     & (1371.2)        & 335391.1                            & (2289.0) & (10076.07401989604, 13678.92598010396)    & 38476.8 & 263467.5 & 55.6  \\
-%    LOP     & N-t1d200.02      & 319675.8                     & (2242.6)        & 333149.5                            & (2497.2) & (11241.971680405779, 15705.428319594244)  & 38268.3 & 265217.5 & 56.5  \\
-%    LOP     & N-t2d150.01      & 15297.7                      & (563.5)         & 24682.7                             & (976.6)  & (8622.206776092156, 10147.793223907844)   & 25318.6 & 262324.3 & 74.6  \\
-%    LOP     & N-t2d150.02      & 14724.9                      & (667.5)         & 24250.8                             & (1354.9) & (8495.057408450935, 10556.742591549064)   & 25425.6 & 262079.2 & 74.8  \\
-%    LOP     & N-t2d200.01      & 33465.2                      & (2001.0)        & 57098.6                             & (4452.8) & (20284.508914101127, 26982.291085898876)  & 40340.9 & 263908.9 & 56.3  \\
-%    LOP     & N-t2d200.02      & 35901.8                      & (2077.5)        & 58123.0                             & (2334.9) & (9515.181030361815, 34927.21896963818)    & 40153.1 & 261723.6 & 55.0  \\\midrule
-%   PFSP     & rec05            & 1330.7                       & (17.9)          & 1317.9                              & (25.1)   & (-33.424760001214985, 7.8247600012150755) & 506.1   & 139893.0 & 400.0 \\
-%   PFSP     & rec13            & 2158.7                       & (29.5)          & 2128.2                              & (25.0)   & (-56.24362499249925, -4.756375007500754)  & 506.8   & 140167.2 & 400.0 \\
-%   PFSP     & rec19            & 2405.1                       & (12.6)          & 2398.6                              & (19.8)   & (-22.313869704790818, 9.313869704790818)  & 1078.3  & 295442.7 & 400.0 \\\midrule
-%    QAP     & kra32            & 116320.0                     & (4673.5)        & 115907.0                            & (4663.4) & (-4799.295389697205, 3973.295389697205)   & 1194.8  & 344766.2 & 400.0 \\
-%    QAP     & nug12            & 660.4                        & (17.3)          & 624.2                               & (26.3)   & (-57.337505259978144, -15.06249474002172) & 180.0   & 53149.3  & 400.0 \\
-%    QAP     & nug30            & 7474.0                       & (83.9)          & 7486.8                              & (40.4)   & (-50.84533806724974, 76.4453380672501)    & 1073.5  & 303352.6 & 400.0 \\
-%    QAP     & tho30            & 191482.0                     & (4419.7)        & 192416.0                            & (2584.2) & (-2527.0213832396516, 4395.021383239651)  & 1066.5  & 306451.3 & 400.0 \\
-%\bottomrule
-%\end{tabular}}
-%\end{table}
 
    \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
-\begin{table}
- \caption{Results.\label{tab:results}}
+\begin{table}[tb]
+ \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance.
+The 95\% confidence interval corresponds to the two independent samples t-test on the mean of the difference between the fitness of UMM minus the one of CEGO.
+Mean runtime is measured in seconds.
+If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 3 days, it was terminated earlier.
+This was the case for all LOP instances shown.\label{tab:results}}
  \resizebox{\textwidth}{!}{%
@@
  \toprule
-           &                  & \mcolc{4}{Mean fitness (sd)} &  \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI for the\\mean difference}}}         & \mcolc{2}{Mean runtime}             & \bf \multirow[b]{2.45}{*}{\shortstack{Mean FE\\CEGO}}\\\cmidrule{3-6}\cmidrule(lr){9-10}
-\bf Problem & \bf     Instance & \mcolc{2}{UMM}               & \mcolc{2}{CEGO} & & & \bf UMM  & \bf CEGO                                  & \bf            \\\midrule
-    LOP &    N-t1d100.01 &       76119.6&(915.4) &        78696.6&(811.4) &       $[$1763.5,&3390.5$]$ &          3.2 &          72.5 &                111.8 \\
-    LOP &    N-t1d100.02 &       76827.7&(891.0) &        79686.5&(367.9) &       $[$2194.5,&3523.1$]$ &          3.2 &          72.9 &                110.7 \\
-    LOP &    N-t1d150.01 &     179508.3&(1647.9) &      187605.5&(1906.9) &       $[$6420.3,&9774.1$]$ &          6.5 &          72.7 &                 73.9 \\
-    LOP &    N-t1d150.02 &     177592.8&(1196.2) &       184320.9&(973.5) &       $[$5700.5,&7755.7$]$ &          6.5 &          73.2 &                 74.1 \\
-    LOP &    N-t1d200.01 &     323513.6&(1371.2) &      335391.1&(2289.0) &     $[$10076.1,&13678.9$]$ &         10.7 &          73.2 &                 55.6 \\
-    LOP &    N-t1d200.02 &     319675.8&(2242.6) &      333149.5&(2497.2) &     $[$11242.0,&15705.4$]$ &         10.6 &          73.7 &                 56.5 \\
-    LOP &    N-t2d150.01 &       15297.7&(563.5) &        24682.7&(976.6) &      $[$8622.2,&10147.8$]$ &          7.0 &          72.9 &                 74.6 \\
-    LOP &    N-t2d150.02 &       14724.9&(667.5) &       24250.8&(1354.9) &      $[$8495.1,&10556.7$]$ &          7.1 &          72.8 &                 74.8 \\
-    LOP &    N-t2d200.01 &      33465.2&(2001.0) &       57098.6&(4452.8) &     $[$20284.5,&26982.3$]$ &         11.2 &          73.3 &                 56.3 \\
-    LOP &    N-t2d200.02 &      35901.8&(2077.5) &       54895.6&(4589.7) &     $[$15539.1,&22448.5$]$ &         11.2 &          73.5 &                 55.6 \\
-    LOP &       N-p50-01 &       21027.6&(711.8) &        21431.2&(299.5) &        $[$-128.1,&935.3$]$ &          0.9 &          72.3 &                219.9 \\
-    LOP &       N-p50-02 &       21613.6&(596.4) &        22021.1&(275.6) &         $[$-42.5,&857.5$]$ &          0.9 &          72.3 &                217.3 \\
-    LOP &       N-atp111 &          594.0&(15.4) &           693.1&(10.3) &          $[$86.7,&111.5$]$ &          4.1 &          72.5 &                 98.7 \\
-    LOP &       N-atp134 &          756.3&(21.3) &           880.5&(18.6) &         $[$105.4,&143.0$]$ &          5.7 &          72.8 &                 83.0 \\
-    LOP &  N-be75eec\_150 &   1415208.5&(58593.5) &    1666193.1&(72310.8) &   $[$188960.6,&313008.6$]$ &          6.9 &          72.9 &                 74.0 \\
-    LOP &  N-be75eec\_250 &  4431271.0&(109756.7) &   5214413.0&(166701.4) &   $[$649040.8,&917243.2$]$ &         15.5 &          73.7 &                 44.9 \\
-    LOP &   N-be75np\_150 &   3204878.3&(80704.1) &    3742444.0&(88500.2) &   $[$457944.2,&617187.2$]$ &          6.9 &          72.9 &                 74.2 \\
-    LOP &   N-be75np\_250 &  9544754.9&(260523.9) &  10887427.8&(159642.5) & $[$1136635.8,&1548710.0$]$ &         15.5 &          74.0 &                 45.3 \\
-   PFSP &      rec05.txt &         1330.7&(17.9) &          1317.9&(25.1) &           $[$-33.4,&7.8$]$ &          0.1 &          38.9 &                400.0 \\
-   PFSP &      rec13.txt &         2158.7&(29.5) &          2128.2&(25.0) &          $[$-56.2,&-4.8$]$ &          0.1 &          38.9 &                400.0 \\
-   PFSP &      rec19.txt &         2405.1&(12.6) &          2398.6&(19.8) &           $[$-22.3,&9.3$]$ &          0.3 &          82.1 &                400.0 \\
-    QAP &      kra32.dat &     116320.0&(4673.5) &      115907.0&(4663.4) &      $[$-4799.3,&3973.3$]$ &          0.3 &          95.8 &                400.0 \\
-    QAP &      nug12.dat &          660.4&(17.3) &           624.2&(26.3) &         $[$-57.3,&-15.1$]$ &          0.0 &          14.8 &                400.0 \\
-    QAP &      nug30.dat &         7474.0&(83.9) &          7486.8&(40.4) &          $[$-50.8,&76.4$]$ &          0.3 &          84.3 &                400.0 \\
-    QAP &      tho30.dat &     191482.0&(4419.7) &      192416.0&(2584.2) &      $[$-2527.0,&4395.0$]$ &          0.3 &          85.1 &                400.0 \\
+            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI for the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule{3-6}\cmidrule(lr){9-10}
+\bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule
+    LOP     & N-t1d100.01      & (915.4)         & 76119.6          & 78696.6                 & (811.4)    & $[$1763.5,    & 3390.5$]$     & 111.8& 72.5 & 3.2  \\
+    LOP     & N-t1d100.02      & (891.0)         & 76827.7          & 79686.5                 & (367.9)    & $[$2194.5,    & 3523.1$]$     & 110.7& 72.9 & 3.2  \\
+    LOP     & N-t1d150.01      & (1647.9)        & 179508.3         & 187605.5                & (1906.9)   & $[$6420.3,    & 9774.1$]$     & 73.9 & 72.7 & 6.5  \\
+    LOP     & N-t1d150.02      & (1196.2)        & 177592.8         & 184320.9                & (973.5)    & $[$5700.5,    & 7755.7$]$     & 74.1 & 73.2 & 6.5  \\
+    LOP     & N-t1d200.01      & (1371.2)        & 323513.6         & 335391.1                & (2289.0)   & $[$10076.1,   & 13678.9$]$    & 55.6 & 73.2 & 10.7 \\
+    LOP     & N-t1d200.02      & (2242.6)        & 319675.8         & 333149.5                & (2497.2)   & $[$11242.0,   & 15705.4$]$    & 56.5 & 73.7 & 10.6 \\
+    LOP     & N-t2d150.01      & (563.5)         & 15297.7          & 24682.7                 & (976.6)    & $[$8622.2,    & 10147.8$]$    & 74.6 & 72.9 & 7.0  \\
+    LOP     & N-t2d150.02      & (667.5)         & 14724.9          & 24250.8                 & (1354.9)   & $[$8495.1,    & 10556.7$]$    & 74.8 & 72.8 & 7.1  \\
+    LOP     & N-t2d200.01      & (2001.0)        & 33465.2          & 57098.6                 & (4452.8)   & $[$20284.5,   & 26982.3$]$    & 56.3 & 73.3 & 11.2 \\
+    LOP     & N-t2d200.02      & (2077.5)        & 35901.8          & 54895.6                 & (4589.7)   & $[$15539.1,   & 22448.5$]$    & 55.6 & 73.5 & 11.2 \\
+    LOP     & N-p50-01         & (711.8)         & 21027.6          & 21431.2                 & (299.5)    & $[$-128.1,    & 935.3$]$      & 219.9& 72.3 & 0.9  \\
+    LOP     & N-p50-02         & (596.4)         & 21613.6          & 22021.1                 & (275.6)    & $[$-42.5,     & 857.5$]$      & 217.3& 72.3 & 0.9  \\
+    LOP     & N-atp111         & (15.4)          & 594.0            & 693.1                   & (10.3)     & $[$86.7,      & 111.5$]$      & 98.7 & 72.5 & 4.1  \\
+    LOP     & N-atp134         & (21.3)          & 756.3            & 880.5                   & (18.6)     & $[$105.4,     & 143.0$]$      & 83.0 & 72.8 & 5.7  \\
+    LOP     & N-be75eec\_150   & (58593.5)       & 1415208.5        & 1666193.1               & (72310.8)  & $[$188960.6,  & 313008.6$]$   & 74.0 & 72.9 & 6.9  \\
+    LOP     & N-be75eec\_250   & (109756.7)      & 4431271.0        & 5214413.0               & (166701.4) & $[$649040.8,  & 917243.2$]$   & 44.9 & 73.7 & 15.5 \\
+    LOP     & N-be75np\_150    & (80704.1)       & 3204878.3        & 3742444.0               & (88500.2)  & $[$457944.2,  & 617187.2$]$   & 74.2 & 72.9 & 6.9  \\
+    LOP     & N-be75np\_250    & (260523.9)      & 9544754.9        & 10887427.8              & (159642.5) & $[$1136635.8, & 1548710.0$]$  & 45.3 & 74.0 & 15.5 \\
+   PFSP     & rec05.txt        & (17.9)          & 1330.7           & 1317.9                  & (25.1)     & $[$-33.4,     & 7.8$]$        & 400.0& 38.9 & 0.1  \\
+   PFSP     & rec13.txt        & (29.5)          & 2158.7           & 2128.2                  & (25.0)     & $[$-56.2,     & -4.8$]$       & 400.0& 38.9 & 0.1  \\
+   PFSP     & rec19.txt        & (12.6)          & 2405.1           & 2398.6                  & (19.8)     & $[$-22.3,     & 9.3$]$        & 400.0& 82.1 & 0.3  \\
+    QAP     & kra32.dat        & (4673.5)        & 116320.0         & 115907.0                & (4663.4)   & $[$-4799.3,   & 3973.3$]$     & 400.0& 95.8 & 0.3  \\
+    QAP     & nug12.dat        & (17.3)          & 660.4            & 624.2                   & (26.3)     & $[$-57.3,     & -15.1$]$      & 400.0& 14.8 & 0.0  \\
+    QAP     & nug30.dat        & (83.9)          & 7474.0           & 7486.8                  & (40.4)     & $[$-50.8,     & 76.4$]$       & 400.0& 84.3 & 0.3  \\
+    QAP     & tho30.dat        & (4419.7)        & 191482.0         & 192416.0                & (2584.2)   & $[$-2527.0,   & 4395.0$]$     & 400.0& 85.1 & 0.3  \\
 \bottomrule

### 2020-10-21 paper/main.tex
fix table

@@
  \toprule
-            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI for the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule{3-6}\cmidrule(lr){9-10}
+            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI for the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule{3-6}\cmidrule(lr){10-11}
 \bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule
-    LOP     & N-t1d100.01      & (915.4)         & 76119.6          & 78696.6                 & (811.4)    & $[$1763.5,    & 3390.5$]$     & 111.8& 72.5 & 3.2  \\
-    LOP     & N-t1d100.02      & (891.0)         & 76827.7          & 79686.5                 & (367.9)    & $[$2194.5,    & 3523.1$]$     & 110.7& 72.9 & 3.2  \\
-    LOP     & N-t1d150.01      & (1647.9)        & 179508.3         & 187605.5                & (1906.9)   & $[$6420.3,    & 9774.1$]$     & 73.9 & 72.7 & 6.5  \\
-    LOP     & N-t1d150.02      & (1196.2)        & 177592.8         & 184320.9                & (973.5)    & $[$5700.5,    & 7755.7$]$     & 74.1 & 73.2 & 6.5  \\
-    LOP     & N-t1d200.01      & (1371.2)        & 323513.6         & 335391.1                & (2289.0)   & $[$10076.1,   & 13678.9$]$    & 55.6 & 73.2 & 10.7 \\
-    LOP     & N-t1d200.02      & (2242.6)        & 319675.8         & 333149.5                & (2497.2)   & $[$11242.0,   & 15705.4$]$    & 56.5 & 73.7 & 10.6 \\
-    LOP     & N-t2d150.01      & (563.5)         & 15297.7          & 24682.7                 & (976.6)    & $[$8622.2,    & 10147.8$]$    & 74.6 & 72.9 & 7.0  \\
-    LOP     & N-t2d150.02      & (667.5)         & 14724.9          & 24250.8                 & (1354.9)   & $[$8495.1,    & 10556.7$]$    & 74.8 & 72.8 & 7.1  \\
-    LOP     & N-t2d200.01      & (2001.0)        & 33465.2          & 57098.6                 & (4452.8)   & $[$20284.5,   & 26982.3$]$    & 56.3 & 73.3 & 11.2 \\
-    LOP     & N-t2d200.02      & (2077.5)        & 35901.8          & 54895.6                 & (4589.7)   & $[$15539.1,   & 22448.5$]$    & 55.6 & 73.5 & 11.2 \\
-    LOP     & N-p50-01         & (711.8)         & 21027.6          & 21431.2                 & (299.5)    & $[$-128.1,    & 935.3$]$      & 219.9& 72.3 & 0.9  \\
-    LOP     & N-p50-02         & (596.4)         & 21613.6          & 22021.1                 & (275.6)    & $[$-42.5,     & 857.5$]$      & 217.3& 72.3 & 0.9  \\
-    LOP     & N-atp111         & (15.4)          & 594.0            & 693.1                   & (10.3)     & $[$86.7,      & 111.5$]$      & 98.7 & 72.5 & 4.1  \\
-    LOP     & N-atp134         & (21.3)          & 756.3            & 880.5                   & (18.6)     & $[$105.4,     & 143.0$]$      & 83.0 & 72.8 & 5.7  \\
-    LOP     & N-be75eec\_150   & (58593.5)       & 1415208.5        & 1666193.1               & (72310.8)  & $[$188960.6,  & 313008.6$]$   & 74.0 & 72.9 & 6.9  \\
-    LOP     & N-be75eec\_250   & (109756.7)      & 4431271.0        & 5214413.0               & (166701.4) & $[$649040.8,  & 917243.2$]$   & 44.9 & 73.7 & 15.5 \\
-    LOP     & N-be75np\_150    & (80704.1)       & 3204878.3        & 3742444.0               & (88500.2)  & $[$457944.2,  & 617187.2$]$   & 74.2 & 72.9 & 6.9  \\
-    LOP     & N-be75np\_250    & (260523.9)      & 9544754.9        & 10887427.8              & (159642.5) & $[$1136635.8, & 1548710.0$]$  & 45.3 & 74.0 & 15.5 \\
-   PFSP     & rec05.txt        & (17.9)          & 1330.7           & 1317.9                  & (25.1)     & $[$-33.4,     & 7.8$]$        & 400.0& 38.9 & 0.1  \\
-   PFSP     & rec13.txt        & (29.5)          & 2158.7           & 2128.2                  & (25.0)     & $[$-56.2,     & -4.8$]$       & 400.0& 38.9 & 0.1  \\
-   PFSP     & rec19.txt        & (12.6)          & 2405.1           & 2398.6                  & (19.8)     & $[$-22.3,     & 9.3$]$        & 400.0& 82.1 & 0.3  \\
-    QAP     & kra32.dat        & (4673.5)        & 116320.0         & 115907.0                & (4663.4)   & $[$-4799.3,   & 3973.3$]$     & 400.0& 95.8 & 0.3  \\
-    QAP     & nug12.dat        & (17.3)          & 660.4            & 624.2                   & (26.3)     & $[$-57.3,     & -15.1$]$      & 400.0& 14.8 & 0.0  \\
-    QAP     & nug30.dat        & (83.9)          & 7474.0           & 7486.8                  & (40.4)     & $[$-50.8,     & 76.4$]$       & 400.0& 84.3 & 0.3  \\
-    QAP     & tho30.dat        & (4419.7)        & 191482.0         & 192416.0                & (2584.2)   & $[$-2527.0,   & 4395.0$]$     & 400.0& 85.1 & 0.3  \\
+    LOP     & N-t1d100.01      & 78696.6                 & (811.4)    & 76119.6   & (915.4)                & $[$1763.5,    & 3390.5$]$     & 111.8& 72.5 & 3.2  \\
+    LOP     & N-t1d100.02      & 79686.5                 & (367.9)    & 76827.7   & (891.0)                & $[$2194.5,    & 3523.1$]$     & 110.7& 72.9 & 3.2  \\
+    LOP     & N-t1d150.01      & 187605.5                & (1906.9)   & 179508.3  & (1647.9)               & $[$6420.3,    & 9774.1$]$     & 73.9 & 72.7 & 6.5  \\
+    LOP     & N-t1d150.02      & 184320.9                & (973.5)    & 177592.8  & (1196.2)               & $[$5700.5,    & 7755.7$]$     & 74.1 & 73.2 & 6.5  \\
+    LOP     & N-t1d200.01      & 335391.1                & (2289.0)   & 323513.6  & (1371.2)               & $[$10076.1,   & 13678.9$]$    & 55.6 & 73.2 & 10.7 \\
+    LOP     & N-t1d200.02      & 333149.5                & (2497.2)   & 319675.8  & (2242.6)               & $[$11242.0,   & 15705.4$]$    & 56.5 & 73.7 & 10.6 \\
+    LOP     & N-t2d150.01      & 24682.7                 & (976.6)    & 15297.7   & (563.5)                & $[$8622.2,    & 10147.8$]$    & 74.6 & 72.9 & 7.0  \\
+    LOP     & N-t2d150.02      & 24250.8                 & (1354.9)   & 14724.9   & (667.5)                & $[$8495.1,    & 10556.7$]$    & 74.8 & 72.8 & 7.1  \\
+    LOP     & N-t2d200.01      & 57098.6                 & (4452.8)   & 33465.2   & (2001.0)               & $[$20284.5,   & 26982.3$]$    & 56.3 & 73.3 & 11.2 \\
+    LOP     & N-t2d200.02      & 54895.6                 & (4589.7)   & 35901.8   & (2077.5)               & $[$15539.1,   & 22448.5$]$    & 55.6 & 73.5 & 11.2 \\
+    LOP     & N-p50-01         & 21431.2                 & (299.5)    & 21027.6   & (711.8)                & $[$-128.1,    & 935.3$]$      & 219.9& 72.3 & 0.9  \\
+    LOP     & N-p50-02         & 22021.1                 & (275.6)    & 21613.6   & (596.4)                & $[$-42.5,     & 857.5$]$      & 217.3& 72.3 & 0.9  \\
+    LOP     & N-atp111         & 693.1                   & (10.3)     & 594.0     & (15.4)                 & $[$86.7,      & 111.5$]$      & 98.7 & 72.5 & 4.1  \\
+    LOP     & N-atp134         & 880.5                   & (18.6)     & 756.3     & (21.3)                 & $[$105.4,     & 143.0$]$      & 83.0 & 72.8 & 5.7  \\
+    LOP     & N-be75eec\_150   & 1666193.1               & (72310.8)  & 1415208.5 & (58593.5)              & $[$188960.6,  & 313008.6$]$   & 74.0 & 72.9 & 6.9  \\
+    LOP     & N-be75eec\_250   & 5214413.0               & (166701.4) & 4431271.0 & (109756.7)             & $[$649040.8,  & 917243.2$]$   & 44.9 & 73.7 & 15.5 \\
+    LOP     & N-be75np\_150    & 3742444.0               & (88500.2)  & 3204878.3 & (80704.1)              & $[$457944.2,  & 617187.2$]$   & 74.2 & 72.9 & 6.9  \\
+    LOP     & N-be75np\_250    & 10887427.8              & (159642.5) & 9544754.9 & (260523.9)             & $[$1136635.8, & 1548710.0$]$  & 45.3 & 74.0 & 15.5 \\
+   PFSP     & rec05.txt        & 1317.9                  & (25.1)     & 1330.7    & (17.9)                 & $[$-33.4,     & 7.8$]$        & 400.0& 38.9 & 0.1  \\
+   PFSP     & rec13.txt        & 2128.2                  & (25.0)     & 2158.7    & (29.5)                 & $[$-56.2,     & -4.8$]$       & 400.0& 38.9 & 0.1  \\
+   PFSP     & rec19.txt        & 2398.6                  & (19.8)     & 2405.1    & (12.6)                 & $[$-22.3,     & 9.3$]$        & 400.0& 82.1 & 0.3  \\
+    QAP     & kra32.dat        & 115907.0                & (4663.4)   & 116320.0  & (4673.5)               & $[$-4799.3,   & 3973.3$]$     & 400.0& 95.8 & 0.3  \\
+    QAP     & nug12.dat        & 624.2                   & (26.3)     & 660.4     & (17.3)                 & $[$-57.3,     & -15.1$]$      & 400.0& 14.8 & 0.0  \\
+    QAP     & nug30.dat        & 7486.8                  & (40.4)     & 7474.0    & (83.9)                 & $[$-50.8,     & 76.4$]$       & 400.0& 84.3 & 0.3  \\
+    QAP     & tho30.dat        & 192416.0                & (2584.2)   & 191482.0  & (4419.7)               & $[$-2527.0,   & 4395.0$]$     & 400.0& 85.1 & 0.3  \\
 \bottomrule

### 2020-10-21 paper/main.tex
paper

--- 2020-10-20 paper/main.tex
@@
 In this work, we propose and analyze uMM, an EDA specifically designed for
-black-box combinatorial optimization under a limited budget. The proposed uMM \ldots\MANUEL{what is the inspiration}
-\
+black-box combinatorial optimization under a limited budget. The proposed uMM 
+\EKHINE{yo no diria que es un EDA porque  no hay generaciones de muestras que compiten }
+\ldots\MANUEL{what is the inspiration}
+\EKHINE{we propose a probabilistic algrithm that (like cego?) (1) iteratively adds solutions to the sample and (2) fits a surrogate model for every new sample obtained.
+However,  different fom cego, our surrogate model is an estimation of the predicted fitness function at each point.
+CEGO controls the  intensification-explotation balance with the expected improvement???
+We mimic the intensification-explotation balance by automatically controllong the variance of our surrogate model.}
 

### 2020-10-21 paper/main.tex
update paper

@@
 The 95\% confidence interval corresponds to the two independent samples t-test on the mean of the difference between the fitness of UMM minus the one of CEGO.
-Mean runtime is measured in seconds.
-If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 3 days, it was terminated earlier.
+Mean runtime is measured in hours.
+If a run went over 72 hours, of  CEGO was not able to reach 400 function evaluations (FEs) in 72 hours, it was terminated earlier.
 This was the case for all LOP instances shown.\label{tab:results}}

### 2020-10-21 paper/main.tex
some comments

@@
 
+\newcommand{\Prob}{\ensuremath{p}}
+
 \section{Background}\label{sec:backgroud}
 
-Permutations are defined as bijections of the set $[n]$ integer onto itself.
+Permutations are defined as bijections of the set $[n]$ \manuel{integer}{of integers} onto itself.
 The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$.
-We denote permutations with Greek letters with exception of the identity permutation denoted as $e=1, 2, 3, \ldots,n$.
-We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma_{-1}=e$ always holds.
+We denote permutations with Greek letters with exception of the identity permutation denoted as $e=\manuel{}{(}1, 2, 3, \dotsc,n\manuel{}{)}$.
+We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=e$ always holds.
 
-Distributions over permutations~\cite{critchlow91} are functions that assign a probability value to each of the permutations in $S_n$, $p(\sigma)\in[0,1]$.
+Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$.
 One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
 The MM defines the probability of each permutation $\sigma$ as follows:
+%
+\begin{equation}
+\Prob(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
+\end{equation}
+%
+with two parameters, $\theta$ and $\sigma_0$, \manuel{Permutation $\sigma_0$ a reference}{where $\sigma_0$ is the}  permutation that has the largest probability value, i.e., the mode of the distribution.
+The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.
+The distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance.
+The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
 
-\begin{equation}
-p(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
-\end{equation}
-with two parameters, $\theta$ and $\sigma_0$, Permutation $\sigma_0$ a reference permutation that has the largest probability value, i.e., the mode of the distribution.
-The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and $\theta$, the dispersion parameter controls this decay.
-The distance $d(\sigma,\sigma_0)$ is the Kendall's-$\tau$ distance.
-The normalization constant $\psi$ can be easily computed for the Kendall's-$\tau$ distance as well as for the Hamming, Cayley and Ulam distance~\cite{Irurozki2016b}.
+The learning process\MANUEL{what learning process?} is divide in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, \manuel{compute}{estimate} the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.
 
-The learning process is divide in two stages: first, we estimate the central permutation of the distribution, $\hat\sigma0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{Irurozki2016b}.
-
-The exact maximum likelihood estimator (MLE) for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{Dwork:2001:RAM:371920.372165}.
-Unfortunately,  obteinig such ranking is NP-hard~\cite{Dwork:2001:RAM:371920.372165}.
+The exact maximum likelihood estimator (MLE) for $\sigma_0$\MANUEL{for $\sigma_0$ o for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
+Unfortunately,  obteinig such ranking is NP-hard~\cite{DwoKumNao2001rank}.
 An alternative to the Kemeny ranking is the Borda ranking~\cite{Ali2011}, which has several advantages.
@@
 \subsection{Sampling step}
-Sampling can be done efficiently~\cite{Irurozki2016b} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$.
-Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{Irurozki2016b}\cite{Regenwetter2004} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
+Sampling can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$.
+Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{IruCalLoz2016permallows}\cite{Regenwetter2004} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
 

### 2020-10-21 paper/main.tex
more comments

@@
 The exact maximum likelihood estimator (MLE) for $\sigma_0$\MANUEL{for $\sigma_0$ o for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
-Unfortunately,  obteinig such ranking is NP-hard~\cite{DwoKumNao2001rank}.
-An alternative to the Kemeny ranking is the Borda ranking~\cite{Ali2011}, which has several advantages.
-First, Borda requires polynomial computational time.
-Second, Borda as an estimator for the MM is guarantied to obtain high quality parameters~\cite{Caragiannis2013}.
-Finally, in a general optimization perspective, Borda error is bounded~\cite{Coppersmith:2010}.
+Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}.
+An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda error is bounded~\cite{CopFleRur2010ordering}.
 
-The Borda ranking of sample $S$ is computed as follows: first, for each item $i$ compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
-Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.
+The Borda ranking of \manuel{sample $S$}{a sample of permutations $S \subset S_n$} is computed as follows.
+First, for each item $i$ compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
+Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.\MANUEL{sorry, I don't follow here, what is an item?
+What does $\sigma_t(i)$ means?}
 
 \subsection{Unbalanced Borda}\label{sec:uborda}
-In these lines we introduce the core for our probabilisitc algorithm, the uBorda algorithm.
+We introduce now the core for our probabilistic algorithm, the uBorda algorithm.
 Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
-The uBorda algorithm is a recent generalization of Borda~\cite{}.
-In uBorda, the score of item $i$ is defined as follows:
-
+The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{This paper is missing} In uBorda, the score of item $i$ is defined as follows:
+%
 \begin{equation}
@@
 \end{equation}
-and, the same as Borda, orders the items increasingly by their score.
+%
+and, the same as Borda, orders the items increasingly by their score.\MANUEL{what is $w()$?}
 
-Intuitively, the uBorda ranking of sample $S$ is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.
-In this paper, we would like to replicate each permutation $\sigma$ a number of times that is proportional to $f(\sigma)$.
-Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
-
+Intuitively, the uBorda ranking of sample $S$\MANUEL{What is $S$?} is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.\MANUEL{so $w$ is a weight function from $w\colon S_n \to \mathbb{R}$?} In this paper, we would like to replicate each permutation $\sigma$ a number of times that is proportional to $f(\sigma)$.\MANUEL{We have not defined $f$ but I guess it is fintess, however, we minimize $f$, so how you handle that?
+I guess we need to define the black-box optimization problem earlier than here} Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
+%
 By setting $\rho=1$ the original Borda is recovered.
-However, for $\rho<1$ uBorda ranking will be closer to the rankings in $S$ with smaller value of $f(\sigma)$.
-Interestingly, Borda is computed in polynomial time.
-Clearly, the choice of the weight depends on the relevant property of a ranking in a particular domain.
-For example, it has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift.
+However, for $\rho<1$ uBorda ranking will be closer to the rankings in $S$ with smaller value of $f(\sigma)$.\MANUEL{by ``with smaller value'' do you mean ranking for increasing values of $f(\sigma)$?} Interestingly, Borda is computed in polynomial time.
+Clearly, the choice of the weight\MANUEL{by weight you mean $\rho$?} depends on the relevant property of a ranking in a particular domain.
+For example, it has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift.\MANUEL{where?
+Do we miss a citation here?}
 

### 2020-10-21 paper/main.tex
more comments

@@
 \newcommand{\Prob}{\ensuremath{p}}
+\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}
+\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}
 
 \section{Background}\label{sec:backgroud}
+
+\subsection{Black-box Combinatorial Optimization under a limited budget}
 
@@
 
+\subsection{CEGO}
+
+Combinatorial Efficient Global Optimization
+(CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
+method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
+optimization problems. In EGO, Gaussian process models are used as a surrogate
+of the landscape of the expensive original problem. An optimization method
+searches for solutions in the surrogate model by optimizing the expected
+improvement criterion, which balances the expected mean and variance of the
+chosen solution. Once a solution is chosen, it is evaluated on the actual
+objective function and the result is used to update the surrogate-model,
+hopefully increasing its predictive power.
+
+CEGO replaces the Euclidean distance measure, used by the surrogate model in
+EGO, with a distance measure appropriate to combinatorial
+landscapes~\citep{ZaeStoBar2014:ppsn}, such as the Kendall distance for
+permutations~\citep{?}. In CEGO, the surrogate model is explored by a GA with
+crossover and mutation operators appropriate for the particular combinatorial
+problem. The original paper notes that coupling the GA with local search does
+not improve the results significantly since the model is anyway an inexact
+estimation of the original objective
+function~\citep[p.~875]{ZaeStoFriFisNauBar2014}.
+
+What else do we need to say?
+
+\subsection{Distributions over permutations}
 Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$.
@@
 Quizás luego: PFSP y QAP
-
-\subsection{Previous method for black box optimization in combinatorial spaces: CEGO}
-
-\manuel{CEGO: Manuel.}{}
-
-Combinatorial Efficient Global Optimization
-(CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
-method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
-optimization problems. In EGO, Gaussian process models are used as a surrogate
-of the landscape of the expensive original problem. An optimization method
-searches for solutions in the surrogate model by optimizing the expected
-improvement criterion, which balances the expected mean and variance of the
-chosen solution. Once a solution is chosen, it is evaluated on the actual
-objective function and the result is used to update the surrogate-model,
-hopefully increasing its predictive power.
-
-CEGO replaces the Euclidean distance measure, used by the surrogate model in
-EGO, with a distance measure appropriate to combinatorial
-landscapes~\citep{ZaeStoBar2014:ppsn}, such as the Kendall distance for
-permutations~\citep{?}. In CEGO, the surrogate model is explored by a GA with
-crossover and mutation operators appropriate for the particular combinatorial
-problem. The original paper notes that coupling the GA with local search does
-not improve the results significantly since the model is anyway an inexact
-estimation of the original objective
-function~\citep[p.~875]{ZaeStoFriFisNauBar2014}.
-
-What else do we need to say?
-
-\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}
-
-\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}
 

### 2020-10-22 paper/main.tex
more comments

@@
 \usepackage{multirow}
-\usepackage[]{algorithm2e}
+
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 \usepackage{booktabs} % For formal tables
@@
 
+\newcommand{\true}{\textsc{true}}
+\newcommand{\false}{\textsc{false}}
+\newcommand{\NULL}{\textsc{null}}
+\newcommand{\NA}{---}
+\newcommand{\assign}{\ensuremath{:=}}
+\newcommand{\procedure}[1]{\textsf{#1}}
+\newcommand{\Inf}{\ensuremath{\infty}}
 \DeclareMathOperator*{\argmax}{arg\,max}
@@
 
+Let us assume a black-box ``fitness'' function that must be minimized over a
+space of permutations $f\colon S_n \to \mathbb{R}$. Being black-box means that
+we can evaluate any candidate permutation $\sigma \in S_n$ to obtain
+$f(\sigma)$, however, we do not know anything else about the form of
+$f$. Moreover, the evaluation of $f$ is expensive in computation time or
+resources and, thus, limits the budget of candidate permutations that we can
+evaluate to a relatively low number. Here, we will study budgets lower than
+400. Such expensive black-box combinatorial problems arise in
+manufacturing~\citep{FerAlvDiaIglEna2014ants}, where the evaluation function
+may involve expensive simulations for which no closed-form mathematical
+description is available.
+
 \subsection{CEGO}
@@
 The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.
-The distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance.
-The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
+The distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance.\MANUEL{Can we define briefly Kendall's tau?} The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
 
@@
 
+\section{Unbalanced Mallows model: probabilistic BB optimization}
+
 LOP: Ekhine
@@
 Then, at each iteration, UMM (i) estimates a surrogate distribution of the fitness function and (ii) proposes a good fitted individual.
+\MANUEL{Hay un poco de duplicación entre lo anterior y lo siguiente, no?}
 
-\section{Unbalanced Mallows model: probabilistic BB optimization}
 In this section we describe our main contribution: a novel algorithm for the optimization of a black box function over permutation spaces.
-This algorithm is called Unbalanced Mallows model (UMM) since it a probabilistic method based on a well-known probability distribution for permutations: the Mallows model.
+This algorithm is called Unbalanced Mallows model (UMM) since it a probabilistic method based on a well-known probability distribution for permutations: the Mallows model.\MANUEL{this sounds as if this was the first time you talk about MM, but you already described it above, no?
+}
 
 \begin{algorithm}[t]
- \KwData{$n$, $m$,}
- \KwResult{how to write algorithm with \LaTeX2e }
- $S$\;
- $F$\;
- $\hat\sigma_0$ = uBorda($S, F,\rho$)\;
- \For{$m$ \text{iterations}}{
-  linear decrease on $\theta$\;
-  $\sigma$ = sample from a $MM(\hat\sigma_0, \theta)$\;
-  Append $\sigma$ and $f(\sigma)$ to $S$ and $F$\;
-   $\hat\sigma_0$ = uBorda($S, F,\rho$)\;
- }
- \Return $S, F$
- \caption{Unbalanced Mallows model for black box optimization}
+ \caption{UMM: Unbalanced Mallows Model for Black-Box Optimization}
  \label{alg:umm}
+  \begin{algorithmic}[1]
+    \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
+    \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
+    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
+    \State linear decrease on $\theta$
+    \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
+    \State $S \assign S \cup \{\sigma\}$
+    \State $F \assign F \cup \{f(\sigma)\}$
+    \EndFor 
+    \State \Return $S$, $F$
+  \end{algorithmic}
 \end{algorithm}
 
-The high-level behavior of UMM is shown in Algorithm~\ref{alg:umm}.
-It starts with an \textit{small} sample of $m'$ permutations, $S=\{\sigma_1, \ldots,  \sigma_{m'}\}$ that are randomly generated.
+As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $m'$ permutations,\MANUEL{but the algorithm requires $m$, no $m'$?} $S=\{\sigma_1, \dotsc,  \sigma_{m'}\}$ that are randomly generated.
 UMM also keeps the function evaluation of each solution, $F=\{f(\sigma_1), \ldots, f(\sigma_{m'})\}$.
-An probabilistic model is used internally to guide the search.
-Then, for $m$ iterations\EKHINE{esto es exactamente así o es $m-m'$?}, UMM generates a random, promising candidate solution $\sigma$ and includes it in the sample $S$ while $f(\sigma)$ in included in $F$.
+Then, for $m$ iterations\EKHINE{esto es exactamente así o es $m-m'$?}, UMM generates a random, promising candidate solution $\sigma$ from a probabilistic model and includes it in the sample $S$ while $f(\sigma)$ is included in $F$.
 At each iteration, the internal probabilistic model is updated.

### 2020-10-22 paper/main.tex
manuel

@@
 
-The probabilistic model is a MM that assigns higher probability values to the permutations in which the fitness value is known or expected to be smaller.\footnote{We assume a minimization problem.} The main contribution of our approach is that the distribution does not represent the current sample $S$ considered by the algorithm but instead, represents a variation of it.
-This variation is based on the uBorda algorithm in Section~\ref{sec:uBorda} and is abstractly equivalent to replicating in the sample the permutations inversely proportionally to their fitness values, i.e., there will be many copies of the good solutions and few of the bad ones.
+The probabilistic model is a MM that assigns higher probability values to the permutations in which the fitness value is known or expected to be minimized.
+The main contribution of our approach is that the distribution does not represent the current sample $S$ considered by the algorithm but instead, represents a variation of it.
+This variation is based on the uBorda algorithm described in Section~\ref{sec:uBorda} and is abstractly equivalent to replicating in the sample the permutations \manuel{inversely proportionally}{a number of times inversely proportional} to their fitness values, i.e., there will be many copies of the good solutions and few of the bad ones.
 One key property of UMM is the computational complexity of working with the original sample and the transformed sample is the same, i.e., polynomial.
-However, the question os how many times should we replicate each permutation in the sample?
-UMM includes a parameter $\rho$ that controls this point and will be detailed in Section~\ref{sec:learning}.
+However, the question of how many times should we replicate each permutation in the sample?
+UMM includes a parameter $\rho$ that controls this point and will be detailed in Section~\ref{sec:learning}.\MANUEL{is this $\rho$ the same we explained earlier?}
 
-At each iteration of the algorithm the following steps take place: (1) randomly generate a permutation distributed according to the internal model and include it in the sample and (2) evaluate the new permutation and update the probabilistic model.
-These steps correspond to two statistical operations: (1) sample a probability distribution and (2) learn a probability distribution of the transformation of the sample.
-Both are describe in the subsequent sections.
+At each iteration of the algorithm the following steps take place: (1) randomly
+generate a permutation distributed according to the internal model and include
+it in the sample and (2) evaluate the new permutation and update the
+probabilistic model. These steps correspond to two statistical operations: (1)
+sample a probability distribution and (2) learn a probability distribution of
+the transformation of the sample. Both are describe in the subsequent sections.
 
-\subsection{Learning step}\label{sec:learning}
-
+\subsubsection{Learning step.}\label{sec:learning}
 As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two steps (1) estimating the location parameter $\sigma_0$ and (2) computing the scale parameter $\theta$.
-Since the scale parameter is set automatically, we focus on learning the first step.
-
+Since the scale parameter is set automatically,\MANUEL{How is it set automatically?} we focus on estimating $\sigma_0$.
 %In order to minimize the number of permutations in the sample (and consequently the number of function evaluations) our proposed UMM learns the distribution of a transformation of the sample rather than the sample.
@@
 This generalization requires setting a value for parameter $\rho$.
-We show how to choose this value in the experimental section.
+We show how to choose this value in the experimental section.\MANUEL{I thought we select it automatically now?}
 
-\subsection{Sampling step}
+\subsubsection{Sampling step.}
 Sampling can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$.
-Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{IruCalLoz2016permallows}\cite{Regenwetter2004} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
-
-At each iteration of UMM $\sigma_0$ is the permutation obtained as a solution in the previous learning step,~\ref{sec:learning}.
+Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{IruCalLoz2016permallows}\cite{Regenwetter2004}\MANUEL{You call this Regenwetter, but the author is Doignon?} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
+%
+At each iteration of UMM, $\sigma_0$ is the permutation obtained as a solution in the previous learning step.
 The scale $\theta$, in the other hand, is set automatically by the algorithm to control the diversification and intensification trade off: At the first iteration it is set so that the expected distance equals  ${n\choose 4}$ (half of the distance of the uniform distribution), and then it linearly decrease until the last iteration where the expected distance equals 1.
-Similar approaches consisting of changing the variance to control the exploration-explotation tradeoff have been taken in different contexts~\cite{arza2019approaching}.
-
-This setting reflects the fact that, despite believing that the best global optimum at each iteration is $\sigma_0$, the uncertainty at the fisrt stages of the algorithm is high and
+Similar approaches consisting of changing the variance to control the exploration-explotation tradeoff have been taken in different contexts~\cite{arza2019approaching}.\MANUEL{there is something wrong with this reference}
+%
+This setting reflects the fact that, despite believing that the best global optimum at each iteration is $\sigma_0$, the uncertainty at the first stages of the algorithm is high and\MANUEL{and?}
 
@@
 Moreover, UMM does not rely on an evolving sequence of samples (called generations in evolutionary approaches).
-Instead, UMM considers a single sample of permutations during the whole execution that is increased iteratively, with one permutation each iteration.
+Instead, UMM considers a single sample of permutations during the whole execution that is increased iteratively, with one permutation each iteration.\MANUEL{I would suggest to move this subsection to the intro, is that OK?}
 \ekhine{bayesian optim}{ff}

### 2020-10-22 paper/main.tex
paper

--- 2020-10-21 paper/main.tex
@@
 
-\newcommand{\Prob}{\ensuremath{p}}
-\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}
-\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}
-
 \section{Background}\label{sec:backgroud}
 
-\subsection{Black-box Combinatorial Optimization under a limited budget}
+Permutations are defined as bijections of the set $[n]$ integer onto itself.
+The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$.
+We denote permutations with Greek letters with exception of the identity permutation denoted as $e=1, 2, 3, \ldots,n$.
+We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma_{-1}=e$ always holds.
 
-Permutations are defined as bijections of the set $[n]$ \manuel{integer}{of integers} onto itself.
-The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$.
-We denote permutations with Greek letters with exception of the identity permutation denoted as $e=\manuel{}{(}1, 2, 3, \dotsc,n\manuel{}{)}$.
-We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=e$ always holds.
+Distributions over permutations~\cite{critchlow91} are functions that assign a probability value to each of the permutations in $S_n$, $p(\sigma)\in[0,1]$.
+One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
+The MM defines the probability of each permutation $\sigma$ as follows:
 
-\subsection{CEGO}
+\begin{equation}
+p(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
+\end{equation}
+with two parameters, $\theta$ and $\sigma_0$, Permutation $\sigma_0$ a reference permutation that has the largest probability value, i.e., the mode of the distribution.
+The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and $\theta$, the dispersion parameter controls this decay.
+The distance $d(\sigma,\sigma_0)$ is the Kendall's-$\tau$ distance.
+The normalization constant $\psi$ can be easily computed for the Kendall's-$\tau$ distance as well as for the Hamming, Cayley and Ulam distance~\cite{Irurozki2016b}.
+
+The learning process is divide in two stages: first, we estimate the central permutation of the distribution, $\hat\sigma_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{Irurozki2016b}.
+
+The exact maximum likelihood estimator (MLE) for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{Dwork:2001:RAM:371920.372165}.
+Unfortunately,  obteinig such ranking is NP-hard~\cite{Dwork:2001:RAM:371920.372165}.
+An alternative to the Kemeny ranking is the Borda ranking~\cite{Ali2011}, which has several advantages.
+First, Borda requires polynomial computational time.
+Second, Borda as an estimator for the MM is guarantied to obtain high quality parameters~\cite{Caragiannis2013}.
+Finally, in a general optimization perspective, Borda error is bounded~\cite{Coppersmith:2010}.
+
+The Borda ranking of sample $S$ is computed as follows: first, for each item $i$ compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
+Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.
+
+\subsection{Unbalanced Borda}\label{sec:uborda}
+In these lines we introduce the core for our probabilisitc algorithm, the uBorda algorithm.
+Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
+The uBorda algorithm is a recent generalization of Borda~\cite{}.
+In uBorda, the score of item $i$ is defined as follows:
+
+\begin{equation}
+\begin{split}
+B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i),
+\end{split}
+\label{eq:uborda_score}
+\end{equation}
+and, the same as Borda, orders the items increasingly by their score.
+
+Intuitively, the uBorda ranking of sample $S$ is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.
+In this paper, we would like to replicate each permutation $\sigma$ a number of times that is proportional to $f(\sigma)$.
+Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
+
+By setting $\rho=1$ the original Borda is recovered.
+However, for $\rho<1$ uBorda ranking will be closer to the rankings in $S$ with smaller value of $f(\sigma)$.
+Interestingly, Borda is computed in polynomial time.
+Clearly, the choice of the weight depends on the relevant property of a ranking in a particular domain.
+For example, it has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift.
+
+LOP: Ekhine
+
+Quizás luego: PFSP y QAP
+
+\subsection{Previous method for black box optimization in combinatorial spaces: CEGO}
+
+\manuel{CEGO: Manuel.}{}
 
@@
 
-\subsection{Distributions over permutations}
-Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$.
-One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
-The MM defines the probability of each permutation $\sigma$ as follows:
-%
-\begin{equation}
-\Prob(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
-\end{equation}
-%
-with two parameters, $\theta$ and $\sigma_0$, \manuel{Permutation $\sigma_0$ a reference}{where $\sigma_0$ is the}  permutation that has the largest probability value, i.e., the mode of the distribution.
-The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.
-The distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance.
-The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
+\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}
 
-The learning process\MANUEL{what learning process?} is divide in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, \manuel{compute}{estimate} the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.
-
-The exact maximum likelihood estimator (MLE) for $\sigma_0$\MANUEL{for $\sigma_0$ o for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
-Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}.
-An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda error is bounded~\cite{CopFleRur2010ordering}.
-
-The Borda ranking of \manuel{sample $S$}{a sample of permutations $S \subset S_n$} is computed as follows.
-First, for each item $i$ compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
-Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.\MANUEL{sorry, I don't follow here, what is an item?
-What does $\sigma_t(i)$ means?}
-
-\subsection{Unbalanced Borda}\label{sec:uborda}
-We introduce now the core for our probabilistic algorithm, the uBorda algorithm.
-Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
-The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{This paper is missing} In uBorda, the score of item $i$ is defined as follows:
-%
-\begin{equation}
-\begin{split}
-B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i),
-\end{split}
-\label{eq:uborda_score}
-\end{equation}
-%
-and, the same as Borda, orders the items increasingly by their score.\MANUEL{what is $w()$?}
-
-Intuitively, the uBorda ranking of sample $S$\MANUEL{What is $S$?} is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.\MANUEL{so $w$ is a weight function from $w\colon S_n \to \mathbb{R}$?} In this paper, we would like to replicate each permutation $\sigma$ a number of times that is proportional to $f(\sigma)$.\MANUEL{We have not defined $f$ but I guess it is fintess, however, we minimize $f$, so how you handle that?
-I guess we need to define the black-box optimization problem earlier than here} Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
-%
-By setting $\rho=1$ the original Borda is recovered.
-However, for $\rho<1$ uBorda ranking will be closer to the rankings in $S$ with smaller value of $f(\sigma)$.\MANUEL{by ``with smaller value'' do you mean ranking for increasing values of $f(\sigma)$?} Interestingly, Borda is computed in polynomial time.
-Clearly, the choice of the weight\MANUEL{by weight you mean $\rho$?} depends on the relevant property of a ranking in a particular domain.
-For example, it has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift.\MANUEL{where?
-Do we miss a citation here?}
-
-LOP: Ekhine
-
-Quizás luego: PFSP y QAP
+\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}
 
@@
 \subsection{Sampling step}
-Sampling can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$.
-Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{IruCalLoz2016permallows}\cite{Regenwetter2004} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
+Sampling can be done efficiently~\cite{Irurozki2016b} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$.
+Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{Irurozki2016b}\cite{Regenwetter2004} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
 
@@
 The 95\% confidence interval corresponds to the two independent samples t-test on the mean of the difference between the fitness of UMM minus the one of CEGO.
-Mean runtime is measured in hours.
-If a run went over 72 hours, of  CEGO was not able to reach 400 function evaluations (FEs) in 72 hours, it was terminated earlier.
+Mean runtime is measured in seconds.
+If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 3 days, it was terminated earlier.
 This was the case for all LOP instances shown.\label{tab:results}}

### 2020-10-22 paper/main.tex
remove conflict markers

--- 2020-10-22 paper/main.tex
@@
 
+\section{Background}\label{sec:backgroud}
+
+Permutations are defined as bijections of the set $[n]$ integer onto itself.
+The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$.
+We denote permutations with Greek letters with exception of the identity permutation denoted as $e=1, 2, 3, \ldots,n$.
+We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma_{-1}=e$ always holds.
+
+Distributions over permutations~\cite{critchlow91} are functions that assign a probability value to each of the permutations in $S_n$, $p(\sigma)\in[0,1]$.
+One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
+The MM defines the probability of each permutation $\sigma$ as follows:
+
+\begin{equation}
+p(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
+\end{equation}
+with two parameters, $\theta$ and $\sigma_0$, Permutation $\sigma_0$ a reference permutation that has the largest probability value, i.e., the mode of the distribution.
+The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and $\theta$, the dispersion parameter controls this decay.
+The distance $d(\sigma,\sigma_0)$ is the Kendall's-$\tau$ distance.
+The normalization constant $\psi$ can be easily computed for the Kendall's-$\tau$ distance as well as for the Hamming, Cayley and Ulam distance~\cite{Irurozki2016b}.
+
+The learning process is divide in two stages: first, we estimate the central permutation of the distribution, $\hat\sigma_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{Irurozki2016b}.
+
+The exact maximum likelihood estimator (MLE) for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{Dwork:2001:RAM:371920.372165}.
+Unfortunately,  obteinig such ranking is NP-hard~\cite{Dwork:2001:RAM:371920.372165}.
+An alternative to the Kemeny ranking is the Borda ranking~\cite{Ali2011}, which has several advantages.
+First, Borda requires polynomial computational time.
+Second, Borda as an estimator for the MM is guarantied to obtain high quality parameters~\cite{Caragiannis2013}.
+Finally, in a general optimization perspective, Borda error is bounded~\cite{Coppersmith:2010}.
+
+The Borda ranking of sample $S$ is computed as follows: first, for each item $i$ compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
+Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.
+
+\subsection{Unbalanced Borda}\label{sec:uborda}
+In these lines we introduce the core for our probabilisitc algorithm, the uBorda algorithm.
+Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
+The uBorda algorithm is a recent generalization of Borda~\cite{}.
+In uBorda, the score of item $i$ is defined as follows:
+
+\begin{equation}
+\begin{split}
+B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i),
+\end{split}
+\label{eq:uborda_score}
+\end{equation}
+and, the same as Borda, orders the items increasingly by their score.
+
+Intuitively, the uBorda ranking of sample $S$ is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.
+In this paper, we would like to replicate each permutation $\sigma$ a number of times that is proportional to $f(\sigma)$.
+Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
+
+By setting $\rho=1$ the original Borda is recovered.
+However, for $\rho<1$ uBorda ranking will be closer to the rankings in $S$ with smaller value of $f(\sigma)$.
+Interestingly, Borda is computed in polynomial time.
+Clearly, the choice of the weight depends on the relevant property of a ranking in a particular domain.
+For example, it has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift.
+
 \newcommand{\Prob}{\ensuremath{p}}
@@
 At each iteration, the internal probabilistic model is updated.
+
+The probabilistic model is a MM that assigns higher probability values to the permutations in which the fitness value is known or expected to be minimized.
+The main contribution of our approach is that the distribution does not represent the current sample $S$ considered by the algorithm but instead, represents a variation of it.
+This variation is based on the uBorda algorithm described in Section~\ref{sec:uBorda} and is abstractly equivalent to replicating in the sample the permutations \manuel{inversely proportionally}{a number of times inversely proportional} to their fitness values, i.e., there will be many copies of the good solutions and few of the bad ones.
+One key property of UMM is the computational complexity of working with the original sample and the transformed sample is the same, i.e., polynomial.
+However, the question of how many times should we replicate each permutation in the sample?
+UMM includes a parameter $\rho$ that controls this point and will be detailed in Section~\ref{sec:learning}.\MANUEL{is this $\rho$ the same we explained earlier?}
 

### 2020-10-23 paper/main.tex
paper corrections

--- 2020-10-22 paper/main.tex
@@
 black-box combinatorial optimization under a limited budget. The proposed uMM 
-\EKHINE{yo no diria que es un EDA porque  no hay generaciones de muestras que compiten }
 \ldots\MANUEL{what is the inspiration}
@@
 
-\citep{LopDubPerStuBir2016irace}
-
+%
+%\citep{LopDubPerStuBir2016irace}
+%
+%\section{Background}\label{sec:backgroud}
+%
+%Permutations are defined as bijections of the set $[n]$ integer onto itself.
+The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$.
+We denote permutations with Greek letters with exception of the identity permutation denoted as $e=1, 2, 3, \ldots,n$.
+We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma_{-1}=e$ always holds.
+%
+%Distributions over permutations~\cite{critchlow91} are functions that assign a probability value to each of the permutations in $S_n$, $p(\sigma)\in[0,1]$.
+One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
+The MM defines the probability of each permutation $\sigma$ as follows:
+%
+%\begin{equation}
+%p(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
+%\end{equation}
+%with two parameters, $\theta$ and $\sigma_0$, Permutation $\sigma_0$ a reference permutation that has the largest probability value, i.e., the mode of the distribution.
+The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and $\theta$, the dispersion parameter controls this decay.
+The distance $d(\sigma,\sigma_0)$ is the Kendall's-$\tau$ distance.
+The normalization constant $\psi$ can be easily computed for the Kendall's-$\tau$ distance as well as for the Hamming, Cayley and Ulam distance~\cite{Irurozki2016b}.
+%
+%The learning process is divide in two stages: first, we estimate the central permutation of the distribution, $\hat\sigma_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{Irurozki2016b}.
+%
+%The exact maximum likelihood estimator (MLE) for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{Dwork:2001:RAM:371920.372165}.
+Unfortunately,  obteinig such ranking is NP-hard~\cite{Dwork:2001:RAM:371920.372165}.
+An alternative to the Kemeny ranking is the Borda ranking~\cite{Ali2011}, which has several advantages.
+First, Borda requires polynomial computational time.
+Second, Borda as an estimator for the MM is guarantied to obtain high quality parameters~\cite{Caragiannis2013}.
+Finally, in a general optimization perspective, Borda error is bounded~\cite{Coppersmith:2010}.
+%
+%The Borda ranking of sample $S$ is computed as follows: first, for each item $i$ compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
+Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.
+%
+%\subsection{Unbalanced Borda}\label{sec:uborda}
+%In these lines we introduce the core for our probabilisitc algorithm, the uBorda algorithm.
+Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
+The uBorda algorithm is a recent generalization of Borda~\cite{}.
+In uBorda, the score of item $i$ is defined as follows:
+%
+%\begin{equation}
+%\begin{split}
+%B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i),
+%\end{split}
+%\label{eq:uborda_score}
+%\end{equation}
+%and, the same as Borda, orders the items increasingly by their score. 
+%
+%Intuitively, the uBorda ranking of sample $S$ is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.
+In this paper, we would like to replicate each permutation $\sigma$ a number of times that is proportional to $f(\sigma)$.
+Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
+%
+%By setting $\rho=1$ the original Borda is recovered.
+However, for $\rho<1$ uBorda ranking will be closer to the rankings in $S$ with smaller value of $f(\sigma)$.
+Interestingly, Borda is computed in polynomial time.
+Clearly, the choice of the weight depends on the relevant property of a ranking in a particular domain.
+For example, it has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift.
+%
 \newcommand{\Prob}{\ensuremath{p}}
@@
 
-Permutations are defined as bijections of the set $[n]$ \manuel{integer}{of integers} onto itself.
+Permutations are defined as bijections of the set $[n]$ of integers onto itself.
 The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$.
-We denote permutations with Greek letters with exception of the identity permutation denoted as $e=\manuel{}{(}1, 2, 3, \dotsc,n\manuel{}{)}$.
+We represent permutations as an ordered vector and, following the standard convention, we say that item $i$ is ranked at position $j$ whenever $\sigma(i)=j$.
+We denote permutations with Greek letters with exception of the identity permutation denoted as $e=(1, 2, 3, \dotsc,n)$.
 We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=e$ always holds.
@@
 One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
-The MM defines the probability of each permutation $\sigma$ as follows:
+The MM defines the probability of each permutation $\sigma\in S_n$ as follows:
 %
@@
 %
-with two parameters, $\theta$ and $\sigma_0$, \manuel{Permutation $\sigma_0$ a reference}{where $\sigma_0$ is the}  permutation that has the largest probability value, i.e., the mode of the distribution.
+with two parameters, $\theta$ and $\sigma_0$.
+Permutation $\sigma_0\S_n$ is a reference permutation and has the largest probability value, i.e., the mode of the distribution.
 The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.
-The distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance.\MANUEL{Can we define briefly Kendall's tau?} The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
+The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
+we assume that the distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance, which counts the number of items ranked in different order in both permutations, \MANUEL{X Can we define briefly Kendall's tau?}
 
-The learning process\MANUEL{what learning process?} is divide in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, \manuel{compute}{estimate} the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.
+\begin{equation}
+\begin{split}
+d(\sigma _{1},\sigma _{2})=|\{(i,j):i<j,&(\sigma _{1}(i)<\sigma _{1}(j)\wedge \sigma _{2}(i)>\sigma _{2}(j))\vee \\
+&(\sigma _{1}(i)>\sigma _{1}(j)\wedge \sigma _{2}(i)<\sigma _{2}(j))\}|
+ \end{split}
+ \label{eq:kendall_decomp}
+\end{equation}
 
-The exact maximum likelihood estimator (MLE) for $\sigma_0$\MANUEL{for $\sigma_0$ o for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
+An alternative definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made in $\sigma_1^{-1}$ to convert it in $\sigma_2^{-1}$ and therefore it is sometimes called \ekhine{swap distance?
+como la llamas tú?}{}
+
+One of the fundamental problems in statistics is to learn the parameters that best fit a given sample of data.
+We will use the standard maximum likelihood estimation (MLE) approach to fit the parameters of a MM for a given collection of permutations.
+The MLE process\MANUEL{X what learning process?} is divided in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.
+
+The exact MLE for $\hat\sigma_0$\MANUEL{X for $\sigma_0$ o for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
 Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}.
@@
 
-The Borda ranking of \manuel{sample $S$}{a sample of permutations $S \subset S_n$} is computed as follows.
-First, for each item $i$ compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
-Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.\MANUEL{sorry, I don't follow here, what is an item?
-What does $\sigma_t(i)$ means?}
+The Borda ranking of a sample of permutations $S \subset S_n$ is computed as follows. 
+\ekhine{First, for each item $i$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
+Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.}
+{First, for each position $i$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
+Then, assign label 1 to the position with the smallest Borda score, label 2 to the position with the second smallest score and so on, i.e., order the positions by Borda score increasingly.}
+\MANUEL{sorry, I don't follow here, what is an item? What does $\sigma_t(i)$ means? EKHINE: he cambiado las definiciones. $\sigma(i)=j$ es  $j$ es el ranking del item $i$}
 
@@
 Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
-The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{This paper is missing} In uBorda, the score of item $i$ is defined as follows:
+The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R_+}$.
+Then, the score of item $i$ is defined as follows:
 %
@@
 %
-and, the same as Borda, orders the items increasingly by their score.\MANUEL{what is $w()$?}
+and, the same as Borda, orders the \ekhine{positions}{items} increasingly by their score.\MANUEL{X what is $w()$?}
 
-Intuitively, the uBorda ranking of sample $S$\MANUEL{What is $S$?} is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.\MANUEL{so $w$ is a weight function from $w\colon S_n \to \mathbb{R}$?} In this paper, we would like to replicate each permutation $\sigma$ a number of times that is proportional to $f(\sigma)$.\MANUEL{We have not defined $f$ but I guess it is fintess, however, we minimize $f$, so how you handle that?
-I guess we need to define the black-box optimization problem earlier than here} Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
-%
-By setting $\rho=1$ the original Borda is recovered.
-However, for $\rho<1$ uBorda ranking will be closer to the rankings in $S$ with smaller value of $f(\sigma)$.\MANUEL{by ``with smaller value'' do you mean ranking for increasing values of $f(\sigma)$?} Interestingly, Borda is computed in polynomial time.
-Clearly, the choice of the weight\MANUEL{by weight you mean $\rho$?} depends on the relevant property of a ranking in a particular domain.
-For example, it has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift.\MANUEL{where?
-Do we miss a citation here?}
+Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$.
+However, in other case uBorda ranking will be closer to the rankings in $S$ with \ekhine{lower}{smaller} value of $w(\sigma)$.
+Rougly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
+%Intuitively, the uBorda ranking a given sample of permutations $S$\MANUEL{X What is $S$?} is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.\MANUEL{X so $w$ is a weight function from $w\colon S_n \to \mathbb{R}$?} %
+Clearly, the choice of  $w(\sigma)$ \MANUEL{X by weight you mean $\rho$?} depends on the relevant property of a ranking in a particular domain.
+For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{}.\MANUEL{X where?
+Do we miss a citation here?\ekhine{}{to be processed, arxiv}}
+
+In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$.\MANUEL{We have not defined $f$ but I guess it is fintess, however, we minimize $f$, so how you handle that?
+X I guess we need to define the black-box optimization problem earlier than here} Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
+Interestingly, uBorda is computed in polynomial time.
+\MANUEL{by ``with smaller value'' do you mean ranking for increasing values of $f(\sigma)$?}
 
@@
 
+\ekhine{
 cosas de intro : 
@@
 For this case study, we show that this case for the Mallows model under the Kendall distance and 
-
 In this section, we introduce our main contribution: We present a new methodology for the optimization of black box functions with a minimal number of queries.
 This methodology is a probabilistic, population-based algorithm.
-
 Essentially, the algorithm  considers a sample of permutations along with the fitness of each of them. It is an iterative process in which the number of iterations is fixed. 
-Then, at each iteration, UMM (i) estimates a surrogate distribution of the fitness function and (ii) proposes a good fitted individual.
-\MANUEL{Hay un poco de duplicación entre lo anterior y lo siguiente, no?}
+Then, at each iteration, UMM (i) estimates a surrogate distribution of the fitness function and (ii) proposes a good fitted individual.}{aqui ha habido mas magia de git, esta seccion empieza en el siguiere paraafo} \MANUEL{Hay un poco de duplicacion entre lo anterior y lo siguiente, no?}
 
@@
 However, the question of how many times should we replicate each permutation in the sample?
+This is controlled with the $\rho$ parameter in uBorda and its tunning and will be detailed in Section~\ref{sec:learning} and the experimental part.\MANUEL{X is this $\rho$ the same we explained earlier?}
+
+%The probabilistic model is a MM that assigns higher probability values to the permutations in which the fitness value is known or expected to be minimized.
+The main contribution of our approach is that the distribution does not represent the current sample $S$ considered by the algorithm but instead, represents a variation of it.
+This variation is based on the uBorda algorithm described in Section~\ref{sec:uBorda} and is abstractly equivalent to replicating in the sample the permutations \manuel{inversely proportionally}{a number of times inversely proportional} to their fitness values, i.e., there will be many copies of the good solutions and few of the bad ones.
+One key property of UMM is the computational complexity of working with the original sample and the transformed sample is the same, i.e., polynomial.
+However, the question of how many times should we replicate each permutation in the sample?
 UMM includes a parameter $\rho$ that controls this point and will be detailed in Section~\ref{sec:learning}.\MANUEL{is this $\rho$ the same we explained earlier?}
@@
 As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two steps (1) estimating the location parameter $\sigma_0$ and (2) computing the scale parameter $\theta$.
-Since the scale parameter is set automatically,\MANUEL{How is it set automatically?} we focus on estimating $\sigma_0$.
-%In order to minimize the number of permutations in the sample (and consequently the number of function evaluations) our proposed UMM learns the distribution of a transformation of the sample rather than the sample.
-This transformation consists on replicating the best fitted permutations.
-However, this is done abstractly:
+Since the scale parameter $\theta$ is set in the sampling step,\MANUEL{X How is it set automatically?} we focus on estimating $\sigma_0$.
+
 The location parameter of a MM is usually approximated with the Borda algorithm.
 Since we are interested on biasing the model towards the regions of the space where the fitness function is smaller, we would like to replicate the permutations a number of times that is inversely proportionally to their fitness values.
-We can simulate the replication of the best fitted individuals by using uBorda, described in Section~\ref{sec:uborda} and setting $w(\sigma)=\rho^{f(\sigma)}$.
+We can simulate the replication of the best fitted individuals by using uBorda and setting $w(\sigma)=\rho^{f(\sigma)}$, as described in Section~\ref{sec:uborda}.
 This generalization requires setting a value for parameter $\rho$.
-We show how to choose this value in the experimental section.\MANUEL{I thought we select it automatically now?}
+We show how to choose this value in the experimental section.\MANUEL{I thought we select it automatically now?\ekhine{that is $\theta$}}
 

### 2020-10-23 paper/main.tex
paper update

@@
 Sampling can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$.
-Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{IruCalLoz2016permallows}\cite{Regenwetter2004}\MANUEL{You call this Regenwetter, but the author is Doignon?} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
+Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{IruCalLoz2016permallows}\cite{Regenwetter2004}\MANUEL{You call this Regenwetter, but the author is Doignon?
+\ekhine{}{he is another coauthor}} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
 %
-At each iteration of UMM, $\sigma_0$ is the permutation obtained as a solution in the previous learning step.
-The scale $\theta$, in the other hand, is set automatically by the algorithm to control the diversification and intensification trade off: At the first iteration it is set so that the expected distance equals  ${n\choose 4}$ (half of the distance of the uniform distribution), and then it linearly decrease until the last iteration where the expected distance equals 1.
-Similar approaches consisting of changing the variance to control the exploration-explotation tradeoff have been taken in different contexts~\cite{arza2019approaching}.\MANUEL{there is something wrong with this reference}
-%
-This setting reflects the fact that, despite believing that the best global optimum at each iteration is $\sigma_0$, the uncertainty at the first stages of the algorithm is high and\MANUEL{and?}
+At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained as a solution in the previous learning step.
+The scale $\theta$ controls the variance and expected distance $\mathbb{E}[D]$ of the MM.
+It is set automatically by the algorithm to control the diversification and intensification trade off: At the first iteration it is set so that the expected distance $\mathbb{E}[D]$ equals  ${n\choose 4}$ (half of the distance of the uniform distribution), and then the expected distance linearly decrease until the last iteration where the expected distance equals 1.
+The expected distance, shown in the next line, does not have a closed-form as function of $\theta$ but its monotonicity guarantees low-cost, high-quality approximations.
 
-\subsection{Comparison to standard probabilistic optimization approaches.}
+\begin{equation}
+\begin{split}
+\mathbb{E}[D] = \frac{k \cdot \exp(-\theta)}{1-\exp(-\theta)} - \sum_{j=n-k+1}^{n}  \frac{j \exp(-j \theta)}{1-\exp(-j\theta)}
+\label{eq:expectation}
+\end{split}
+\end{equation}
+
+A purely explotation approach\ekhine{esto se dice asi?
+uno greedy igual esta mejor dicho} would systematically generate $\sigma_0$ in the sampling stage.
+However, this setting  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
+Similar approaches have been taken in different contexts~\cite{arza2019approaching}.\MANUEL{there is something wrong with this reference\ekhine{}{qué le pasa?}}
+
+\subsection{Comparison to standard probabilistic optimization approaches.}\ekhine{esto lo dejamos para el final? intro, discussion...}{}
 UMM is a population based algorithm.
@@
 \section{Experimental analysis}
+
+\ekhine{}{In prev sections we say that we show here how to tune $rho$. Le ves sitio? con los heatmaps? al principio de la sección? an outline here}
+\ekhine{}{recall that $w(\sigma) = \rho^{f(sigma)}$. We select $\rho$in such a way that the 100$r\%$ of the samples have the 100$w\%$ of the weight}
 

### 2020-10-23 paper/main.tex
paper update
more comments

--- 2020-10-23 paper/main.tex
@@
   Bayesian optimization and a Mallows-based EDA may be an interesting direction
-  for future research.
+  for future research.\sloppy  
 \textcolor{red}{16 pages max, deadline: 1 November 2020}
@@
 
-In this work, we propose and analyze uMM, an EDA specifically designed for
-black-box combinatorial optimization under a limited budget. The proposed uMM 
+In this work, we propose and analyze UMM, an EDA specifically designed for
+black-box combinatorial optimization under a limited budget. The proposed UMM 
+is a population based algorithm.
+There exists a wide body of research on population based probabilistic algorithms, where most prominent examples are Estimation of Distribution Algorithms (EDAs).
+Our approach drastically differs from EDAs since EDAs assume that function evaluations are cheap and they can do many of them.
+On the contrary, UMM minimizes the number of function evaluations.
+Moreover, UMM does not rely on an evolving sequence of samples (called generations in evolutionary approaches).
+Instead, UMM considers a single sample of permutations during the whole execution that is increased iteratively, with one permutation each iteration.
+
 \ldots\MANUEL{what is the inspiration}
@@
 evaluate to a relatively low number. Here, we will study budgets lower than
-400. Such expensive black-box combinatorial problems arise in
-manufacturing~\citep{FerAlvDiaIglEna2014ants}, where the evaluation function
+400. Such expensive black-box combinatorial problems arise, for example, in
+manufacturing~\citep{FerAlvDiaIglEna2014ants}, where the fitness function
 may involve expensive simulations for which no closed-form mathematical
 description is available.
+
+Distances in permutation space can be measured in various ways.  Kendall's $\tau$ distance counts the number of items ranked in different order in both permutations:
+%
+\begin{equation}
+\begin{split}
+d(\sigma _{1},\sigma_{2})=\left\lvert\Bigl\{(i,j):i<j \land \phantom{}\Bigl(\right.&\bigl(\sigma_{1}(i) < \sigma_{1}(j)\land \sigma_{2}(i)>\sigma_{2}(j)\bigr)\lor \phantom{} \\
+ &\bigl(\sigma_{1}(i)>\sigma_{1}(j)\land \sigma_{2}(i)<\sigma_{2}(j)\bigr)\Bigr)\Bigr\}\Bigr\rvert\\
+ \end{split}
+ \label{eq:kendall_decomp}
+\end{equation}
+%
+An alternative\MANUEL{Alternative or equivalent?
+That is, it is the same numbers or it will produce a different result?} definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made in $\sigma_1^{-1}$ to obtain $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
 
@@
 chosen solution. Once a solution is chosen, it is evaluated on the actual
-objective function and the result is used to update the surrogate-model,
+fitness function and the result is used to update the surrogate-model,
 hopefully increasing its predictive power.
@@
 EGO, with a distance measure appropriate to combinatorial
-landscapes~\citep{ZaeStoBar2014:ppsn}, such as the Kendall distance for
-permutations~\citep{?}. In CEGO, the surrogate model is explored by a GA with
+landscapes~\citep{ZaeStoBar2014:ppsn}, such as Kendall's $\tau$ distance for
+permutations (Eq.~\ref{eq:kendall_decomp}). In CEGO, the surrogate model is explored by a GA with
 crossover and mutation operators appropriate for the particular combinatorial
@@
 not improve the results significantly since the model is anyway an inexact
-estimation of the original objective
+estimation of the original fitness
 function~\citep[p.~875]{ZaeStoFriFisNauBar2014}.
-
-What else do we need to say?
 
@@
 One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
-The MM defines the probability of each permutation $\sigma\in S_n$ as follows:
+The MM defines the probability of each permutation $\sigma\in S_n$ as follows:\MANUEL{Could you check this?
+I wanted to introduce the notation $MM(\sigma_0, \theta)$ that we use later}
 %
-\begin{equation}
-\Prob(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
+\begin{equation}\label{eq:MM}
+\Prob(\sigma \mid \sigma_0, \theta )= \frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi} \sim MM(\sigma_0, \theta)
 \end{equation}
@@
 with two parameters, $\theta$ and $\sigma_0$.
-Permutation $\sigma_0\S_n$ is a reference permutation and has the largest probability value, i.e., the mode of the distribution.
+Permutation $\sigma_0\in S_n$ is a reference permutation and has the largest probability value, i.e., the mode of the distribution.
 The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.
-The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
-we assume that the distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance, which counts the number of items ranked in different order in both permutations, \MANUEL{X Can we define briefly Kendall's tau?}
+The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance (Eq.~\ref{eq:kendall_decomp}) as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
 
-\begin{equation}
-\begin{split}
-d(\sigma _{1},\sigma _{2})=|\{(i,j):i<j,&(\sigma _{1}(i)<\sigma _{1}(j)\wedge \sigma _{2}(i)>\sigma _{2}(j))\vee \\
-&(\sigma _{1}(i)>\sigma _{1}(j)\wedge \sigma _{2}(i)<\sigma _{2}(j))\}|
- \end{split}
- \label{eq:kendall_decomp}
-\end{equation}
+\manuel{One of the fundamental problems in statistics is to learn the parameters that best fit a given sample of data.}{}\MANUEL{I think we don't need to emphasize this, it is known} %
+We will use the standard maximum likelihood estimation (MLE) approach to fit the parameters of a MM for a given collection of permutations. 
+The MLE process is divided in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.
 
-An alternative definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made in $\sigma_1^{-1}$ to convert it in $\sigma_2^{-1}$ and therefore it is sometimes called \ekhine{swap distance?
-como la llamas tú?}{}
-
-One of the fundamental problems in statistics is to learn the parameters that best fit a given sample of data.
-We will use the standard maximum likelihood estimation (MLE) approach to fit the parameters of a MM for a given collection of permutations.
-The MLE process\MANUEL{X what learning process?} is divided in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.
-
-The exact MLE for $\hat\sigma_0$\MANUEL{X for $\sigma_0$ o for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
+The exact MLE for $\hat\sigma_0$\MANUEL{MLE for $\sigma_0$ or for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
 Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}.
-An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda error is bounded~\cite{CopFleRur2010ordering}.
+An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda error is bounded~\cite{CopFleRur2010ordering}.\MANUEL{I'm not sure what the latter means}
 
 The Borda ranking of a sample of permutations $S \subset S_n$ is computed as follows. 
-\ekhine{First, for each item $i$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
-Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.}
-{First, for each position $i$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
-Then, assign label 1 to the position with the smallest Borda score, label 2 to the position with the second smallest score and so on, i.e., order the positions by Borda score increasingly.}
-\MANUEL{sorry, I don't follow here, what is an item? What does $\sigma_t(i)$ means? EKHINE: he cambiado las definiciones. $\sigma(i)=j$ es  $j$ es el ranking del item $i$}
+First, for each position $i \in [n]$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
+Then, assign label 1 to the position with the smallest Borda score, label 2 to the position with the second smallest score and so on, i.e., order the positions by Borda score increasingly.
 
@@
 Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
-The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R_+}$.
-Then, the score of item $i$ is defined as follows:
+The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R^{+}}$.
+Then, the score of \manuel{item}{position} $i$ is defined as follows:
 %
-\begin{equation}
-\begin{split}
-B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i),
-\end{split}
-\label{eq:uborda_score}
+\begin{equation}\label{eq:uborda_score}
+B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i) \enspace,
 \end{equation}
 %
-and, the same as Borda, orders the \ekhine{positions}{items} increasingly by their score.\MANUEL{X what is $w()$?}
+and, the same as Borda, orders the \ekhine{positions}{items}\MANUEL{why items? we talk about positions above. We could talk about decision variables if that is clearer} increasingly by their score.
 
 Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$.
-However, in other case uBorda ranking will be closer to the rankings in $S$ with \ekhine{lower}{smaller} value of $w(\sigma)$.
-Rougly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
+However, in other case uBorda ranking will be closer to the rankings in $S$ with smaller value of $w(\sigma)$.
+Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
 %Intuitively, the uBorda ranking a given sample of permutations $S$\MANUEL{X What is $S$?} is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.\MANUEL{X so $w$ is a weight function from $w\colon S_n \to \mathbb{R}$?} %
-Clearly, the choice of  $w(\sigma)$ \MANUEL{X by weight you mean $\rho$?} depends on the relevant property of a ranking in a particular domain.
+Clearly, the choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
 For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{}.\MANUEL{X where?
@@
 
-In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$.\MANUEL{We have not defined $f$ but I guess it is fintess, however, we minimize $f$, so how you handle that?
-X I guess we need to define the black-box optimization problem earlier than here} Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
+In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$.
+Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
 Interestingly, uBorda is computed in polynomial time.
-\MANUEL{by ``with smaller value'' do you mean ranking for increasing values of $f(\sigma)$?}
 
-\section{Unbalanced Mallows model: probabilistic BB optimization}
+\renewcommand{\subsubsection}{\smallskip{}\noindent\textbf}
 
-LOP: Ekhine
+\section{Unbalanced Mallows model (UMM)}
 
-Quizás luego: PFSP y QAP
+% \ekhine{
+% cosas de intro : 
+% This new methodology relies on the assumption that the fitness landscatpe is correlated with a distribution over the same space. 
+% Previous paper have confirmed this point
+% For this case study, we show that this case for the Mallows model under the Kendall distance and 
+% In this section, we introduce our main contribution: We present a new methodology for the optimization of black box functions with a minimal number of queries.
+This methodology is a probabilistic, population-based algorithm.
+% Essentially, the algorithm  considers a sample of permutations along with the fitness of each of them. It is an iterative process in which the number of iterations is fixed. 
+% Then, at each iteration, UMM (i) estimates a surrogate distribution of the fitness function and (ii) proposes a good fitted individual.}{aqui ha habido mas magia de git, esta seccion empieza en el siguiere paraafo} \MANUEL{Hay un poco de duplicacion entre lo anterior y lo siguiente, no?}
 
-\ekhine{
-cosas de intro : 
-This new methodology relies on the assumption that the fitness landscatpe is correlated with a distribution over the same space. 
-Previous paper have confirmed this point
-For this case study, we show that this case for the Mallows model under the Kendall distance and 
-In this section, we introduce our main contribution: We present a new methodology for the optimization of black box functions with a minimal number of queries.
-This methodology is a probabilistic, population-based algorithm.
-Essentially, the algorithm  considers a sample of permutations along with the fitness of each of them. It is an iterative process in which the number of iterations is fixed. 
-Then, at each iteration, UMM (i) estimates a surrogate distribution of the fitness function and (ii) proposes a good fitted individual.}{aqui ha habido mas magia de git, esta seccion empieza en el siguiere paraafo} \MANUEL{Hay un poco de duplicacion entre lo anterior y lo siguiente, no?}
-
-In this section we describe our main contribution: a novel algorithm for the optimization of a black box function over permutation spaces.
-This algorithm is called Unbalanced Mallows model (UMM) since it a probabilistic method based on a well-known probability distribution for permutations: the Mallows model.\MANUEL{this sounds as if this was the first time you talk about MM, but you already described it above, no?
-}
+In this section we describe our main contribution: Unbalanced Mallows model
+(UMM), a novel algorithm for the optimization of a black-box function over
+permutation spaces. This algorithm is a probabilistic method based on the
+Mallows model and uBorda learning.
 
@@
   \begin{algorithmic}[1]
-    \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
-    \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
-    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
+    \Require $\minit$: number of initial permutations, $m$: total budget, $\rho$: learning rate
+    %\\ ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
+    \State $S \assign $ generate $\minit$ permutations uniformly at random
+    \State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
+    \For{$i \assign \minit$ \textbf{to} $m$ \text{evaluations}}
+    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}
     \State linear decrease on $\theta$
-    \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
-    \State $S \assign S \cup \{\sigma\}$
-    \State $F \assign F \cup \{f(\sigma)\}$
+    \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
+    \State $S \assign S \cup \{\sigma\}$, $\quad F \assign F \cup \{f(\sigma)\}$
     \EndFor 
     \State \Return $S$, $F$
+  %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
+  %   \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
+  %   \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
+  %   \State linear decrease on $\theta$
+  %   \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
+  %   \State $S \assign S \cup \{\sigma\}$
+  %   \State $F \assign F \cup \{f(\sigma)\}$
+  %   \EndFor 
+  %   \State \Return $S$, $F$
  \end{algorithmic}
@@
 
-As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $m'$ permutations,\MANUEL{but the algorithm requires $m$, no $m'$?} $S=\{\sigma_1, \dotsc,  \sigma_{m'}\}$ that are randomly generated.
-UMM also keeps the function evaluation of each solution, $F=\{f(\sigma_1), \ldots, f(\sigma_{m'})\}$.
-Then, for $m$ iterations\EKHINE{esto es exactamente así o es $m-m'$?}, UMM generates a random, promising candidate solution $\sigma$ from a probabilistic model and includes it in the sample $S$ while $f(\sigma)$ is included in $F$.
-At each iteration, the internal probabilistic model is updated.
+As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are randomly generated (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated.
+Then, using the current sample $S$, the uBorda algorithm is applied to find a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}).
+Together with the parameter $\theta$, they define a Mallows model (Eq.~\ref{eq:MM}) from which we sample a new candidate permutation (line~\ref{line:sample}).
+This candidate permutation (solution) is evaluated and added to the sample $S$.
+The process is repeated up to a given maximum budget of evaluations $m$.\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it.
+Can we give a bit more detail?}
 
-The probabilistic model is a MM that assigns higher probability values to the permutations in which the fitness value is known or expected to be minimized.
-The main contribution of our approach is that the distribution does not represent the current sample $S$ considered by the algorithm but instead, represents a variation of it.
-This variation is based on the uBorda algorithm described in Section~\ref{sec:uBorda} and is abstractly equivalent to replicating in the sample the permutations \manuel{inversely proportionally}{a number of times inversely proportional} to their fitness values, i.e., there will be many copies of the good solutions and few of the bad ones.
-One key property of UMM is the computational complexity of working with the original sample and the transformed sample is the same, i.e., polynomial.
-However, the question of how many times should we replicate each permutation in the sample?
-This is controlled with the $\rho$ parameter in uBorda and its tunning and will be detailed in Section~\ref{sec:learning} and the experimental part.\MANUEL{X is this $\rho$ the same we explained earlier?}
+The key idea behind UMM is that the probabilistic Mallows model (MM) used in
+UMM does not simply represent the current sample $S$. Instead, the model
+assigns higher probability values to the permutations with known (or expected)
+minimal fitness values. This is achieved by means of the uBorda ranking
+(Section~\ref{sec:uBorda}), which can be seen as replicating each permutation
+$\sigma$ in the sample a number of times inversely proportional to its fitness
+$f(\sigma)$, i.e., it behaves as if there were many copies of the best
+solutions and few of the bad ones. A crucial property of UMM is that the
+computational complexity of working with the original sample and with the
+transformed sample is the same, i.e., polynomial. The most important steps are
+the sampling of a new permutation from $MM(\sigma_0,\rho)$ and the update
+(\emph{learning}) of the parameters $\sigma_0$ and $\rho$.\MANUEL{Is
+  this $\rho$ the same we explained earlier for the weights?}
+
+% 'However, the question of how
+% many times should we replicate each permutation in the sample? This is
+% controlled with the $\rho$ parameter in uBorda and its tunning and will be
+% detailed in Section~\ref{sec:learning} and the experimental part.
 
@@
 
-At each iteration of the algorithm the following steps take place: (1) randomly
+\manuel{At each iteration of the algorithm the following steps take place: (1) randomly
 generate a permutation distributed according to the internal model and include
@@
 sample a probability distribution and (2) learn a probability distribution of
-the transformation of the sample. Both are describe in the subsequent sections.
+the transformation of the sample.
+Both are describe in the subsequent sections.}{}\MANUEL{I don't think we need this, it is said above (if some detail is missing we should integrate it the paragraph above)}
 
-\subsubsection{Learning step.}\label{sec:learning}
-As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two steps (1) estimating the location parameter $\sigma_0$ and (2) computing the scale parameter $\theta$.
-Since the scale parameter $\theta$ is set in the sampling step,\MANUEL{X How is it set automatically?} we focus on estimating $\sigma_0$.
+\newcommand{\Edist}{\ensuremath{\text{E}[D]}}
 
-The location parameter of a MM is usually approximated with the Borda algorithm.
+\subsubsection{Sampling step.}
+Sampling from $MM(\sigma_0,\rho)$  can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{Collas,IruCalLoz2016permallows,Regenwetter2004}\MANUEL{Which paper is Collas?
+It is not found.}\MANUEL{You call this Regenwetter, but the author is Doignon?
+\ekhine{}{he is another coauthor} But the entry shows only 1 author.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{respectively?}
+%
+At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained as a solution in the previous learning step.
+The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM.\MANUEL{I'm not sure what is $D$, distance from where to where?} It is set automatically by the algorithm to control the diversification and intensification trade off.
+At the first iteration, it is set so that the expected distance $\Edist$ equals  $\binom{n}{4}$ (half of the distance of the uniform distribution), and then the expected distance linearly decrease until the last iteration where the expected distance equals 1.
+The expected distance, shown in the next line, does not have a closed-form as function of $\theta$ but its monotonicity guarantees low-cost, high-quality approximations:\MANUEL{I don't understand this sentence, sorry.
+You say it does not have a closed-form, but you show a closed-form.
+Do you mean that the formula below is only an approximation of $\Edist$?
+What is low-cost here?
+what is high-quality?}\MANUEL{What is $k$ in the equation?}
+%
+\begin{equation}\label{eq:expectation}
+\Edist = \frac{k \cdot \exp(-\theta)}{1-\exp(-\theta)} - \sum_{j=n-k+1}^{n}  \frac{j \exp(-j \theta)}{1-\exp(-j\theta)}
+\end{equation}
+
+A purely explotation approach\ekhine{esto se dice asi?
+uno greedy igual esta mejor dicho} would systematically generate $\sigma_0$ in the sampling stage.\MANUEL{I'm not sure what ``systematically generate $\sigma_0$'' means.} However, this setting\MANUEL{which setting?
+$\theta$?}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
+Similar approaches have been taken in different contexts~\cite{arza2019approaching}.\MANUEL{there is something wrong with this reference\ekhine{}{qué le pasa?} Don't you see a complete mess of the authors list?
+It looks like it has 20 authors.}
+
+\subsubsection{Learning step.}
+As explained above, the parameter $\theta$ is updated deterministically at the
+sampling step. The parameter $\sigma_0$, however, is learned from the current
+sample $S$ by using the uBorda algorithm with a weight
+$w(\sigma)=\rho^{f(\sigma)}$ (Section~\ref{sec:uborda}). The learning rate
+is controlled by parameter $\rho \in (0,1)$, which currently is user-defined.
+Higher values or $\rho$ bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas smaller values make the sampling more uniform, thus slowing down the learning.
+
+% As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two steps (1) estimating the location parameter $\sigma_0$ and (2) computing the scale parameter $\theta$.
+Since the scale parameter $\theta$ is set in the sampling step,\MANUEL{X How is it set?} we focus on estimating $\sigma_0$.
+
+% The location parameter of a MM is usually approximated with the Borda algorithm.
 Since we are interested on biasing the model towards the regions of the space where the fitness function is smaller, we would like to replicate the permutations a number of times that is inversely proportionally to their fitness values.
-We can simulate the replication of the best fitted individuals by using uBorda and setting $w(\sigma)=\rho^{f(\sigma)}$, as described in Section~\ref{sec:uborda}.
+% We can simulate the replication of the best fitted individuals by using uBorda and setting $w(\sigma)=\rho^{f(\sigma)}$, as described in .
 This generalization requires setting a value for parameter $\rho$.
@@
 
-\subsubsection{Sampling step.}
-Sampling can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$.
-Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{IruCalLoz2016permallows}\cite{Regenwetter2004}\MANUEL{You call this Regenwetter, but the author is Doignon?} where the computational complexities are $O(n\log n)$ and $O(n^2)$.
-%
-At each iteration of UMM, $\sigma_0$ is the permutation obtained as a solution in the previous learning step.
-The scale $\theta$, in the other hand, is set automatically by the algorithm to control the diversification and intensification trade off: At the first iteration it is set so that the expected distance equals  ${n\choose 4}$ (half of the distance of the uniform distribution), and then it linearly decrease until the last iteration where the expected distance equals 1.
-Similar approaches consisting of changing the variance to control the exploration-explotation tradeoff have been taken in different contexts~\cite{arza2019approaching}.\MANUEL{there is something wrong with this reference}
-%
-This setting reflects the fact that, despite believing that the best global optimum at each iteration is $\sigma_0$, the uncertainty at the first stages of the algorithm is high and\MANUEL{and?}
-
-\subsection{Comparison to standard probabilistic optimization approaches.}
-UMM is a population based algorithm.
-There exists a wide body of research on population based probabilistic algorithms, where most prominent examples are Estimation of Distribution Algorithms (EDAs).
-Our approach drastically differs from EDAs since EDAs assume that function evaluations are cheap and they can do many of them.
-On the contrary, UMM minimizes the number of function evaluations.
-Moreover, UMM does not rely on an evolving sequence of samples (called generations in evolutionary approaches).
-Instead, UMM considers a single sample of permutations during the whole execution that is increased iteratively, with one permutation each iteration.\MANUEL{I would suggest to move this subsection to the intro, is that OK?}
 \ekhine{bayesian optim}{ff}
 \EKHINE{subir a arxiv uBorda}
+
+LOP: Ekhine
+
+Quizás luego: PFSP y QAP
 
@@
 \citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
-when using Hamming distance. We plan to extend our analysis of uMM to other
+when using Hamming distance. We plan to extend our analysis of UMM to other
 distance measures and to the dynamic selection of distance
@@
 
-Complete this with details of uMM: MANUEL
+Complete this with details of UMM: MANUEL
 
@@
 
+\ekhine{}{In prev sections we say that we show here how to tune $rho$. Le ves sitio? con los heatmaps? al principio de la sección? an outline here}
+\ekhine{}{recall that $w(\sigma) = \rho^{f(sigma)}$. We select $\rho$ in such a way that the 100$r\%$ of the samples have the 100$w\%$ of the weight}
+
 To better understand the behaviour of the algorithms, we record the fitness of
@@
 over 10 runs with different random seeds on the same problem instance.  As an example, each line in Fig.~\ref{fig:lop_synth} 
-shows the mean fitness of the solutions evaluated by either uMM or CEGO on one synthetic LOP
+shows the mean fitness of the solutions evaluated by either UMM or CEGO on one synthetic LOP
 instance. The shaded area shows one standard deviation around the
@@
 
-When we compare the behaviour of CEGO and uMM on synthetic LOP instances
+When we compare the behaviour of CEGO and UMM on synthetic LOP instances
 (Fig.~\ref{fig:lop_synth}), we observe a clear pattern independently of the
@@
 
-On the other hand, uMM converges much more slowly than CEGO, only reaching the
-same fitness values around evaluation 250. Yet, uMM is able to keep the rate of
+On the other hand, UMM converges much more slowly than CEGO, only reaching the
+same fitness values around evaluation 250. Yet, UMM is able to keep the rate of
 improvement and completely overtake CEGO around evaluation 350. As in the case
 of CEGO, lower uniformity values lead to worse performance, although the
-difference is much smaller for uMM than it was for CEGO.\MANUEL{If there is anything else we can say here, please go ahead and say it}
+difference is much smaller for UMM than it was for CEGO.\MANUEL{If there is anything else we can say here, please go ahead and say it}
 
@@
 wall-clock time of 3 days for each run of CEGO. On these instances, the
-behaviour of CEGO and uMM is much more similar (at least up to the termination
+behaviour of CEGO and UMM is much more similar (at least up to the termination
 point of CEGO). It appears that our synthetic LOP instances have a fitness
 landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
-instances do not present such landscape. Comparing CEGO and uMM results, the
-differences get smaller, in favour of uMM, with larger instance size. Our
+instances do not present such landscape. Comparing CEGO and UMM results, the
+differences get smaller, in favour of UMM, with larger instance size. Our
 conjecture is that building an accurate Bayesian model and searching for good
@@
 \begin{itemize}
-\item How difficult is to extend uMM to other distance metrics?
+\item How difficult is to extend UMM to other distance metrics?
 \item Can we plot the posterior probability of the optimal solution?

### 2020-10-23 paper/main.tex
more edits

@@
 is a population based algorithm.
-There exists a wide body of research on population based probabilistic algorithms, where most prominent examples are Estimation of Distribution Algorithms (EDAs).
+There exists a wide body of research on population based probabilistic algorithms, where most prominent examples are Estimation of Distribution Algorithms (EDAs)~\citep{???}.
 Our approach drastically differs from EDAs since EDAs assume that function evaluations are cheap and they can do many of them.
@@
 
-Let us assume a black-box ``fitness'' function that must be minimized over a
-space of permutations $f\colon S_n \to \mathbb{R}$. Being black-box means that
-we can evaluate any candidate permutation $\sigma \in S_n$ to obtain
-$f(\sigma)$, however, we do not know anything else about the form of
-$f$. Moreover, the evaluation of $f$ is expensive in computation time or
-resources and, thus, limits the budget of candidate permutations that we can
-evaluate to a relatively low number. Here, we will study budgets lower than
-400. Such expensive black-box combinatorial problems arise, for example, in
-manufacturing~\citep{FerAlvDiaIglEna2014ants}, where the fitness function
-may involve expensive simulations for which no closed-form mathematical
-description is available.
-
 Distances in permutation space can be measured in various ways.  Kendall's $\tau$ distance counts the number of items ranked in different order in both permutations:
@@
 That is, it is the same numbers or it will produce a different result?} definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made in $\sigma_1^{-1}$ to obtain $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
+
+Let us assume a black-box ``fitness'' function that must be minimized over a
+space of permutations $f\colon S_n \to \mathbb{R}$. Being black-box means that
+we can evaluate any candidate permutation $\sigma \in S_n$ to obtain
+$f(\sigma)$, however, we do not know anything else about the form of
+$f$. Moreover, the evaluation of $f$ is expensive in computation time or
+resources and, thus, we can only evaluate a limited budget of candidate
+permutations. Here, we will study budgets lower than 400. Such expensive
+black-box combinatorial problems arises in diverse contexts, for example,
+protein folding~\citep{RomKraArn2012protein} and industrial
+production~\citep{FerAlvDiaIglEna2014ants}, where the fitness function may
+involve expensive simulations for which no closed-form mathematical description
+is available.
 
@@
 A purely explotation approach\ekhine{esto se dice asi?
-uno greedy igual esta mejor dicho} would systematically generate $\sigma_0$ in the sampling stage.\MANUEL{I'm not sure what ``systematically generate $\sigma_0$'' means.} However, this setting\MANUEL{which setting?
+uno greedy igual esta mejor dicho} would systematically generate $\sigma_0$ in the sampling stage.\MANUEL{I'm not sure what ``systematically generate $\sigma_0$'' means in this context.} However, this setting\MANUEL{which setting?
 $\theta$?}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
@@
 \EKHINE{subir a arxiv uBorda}
-
-LOP: Ekhine
-
-Quizás luego: PFSP y QAP
 
@@
 
-Complete this with details of UMM: MANUEL
-
 \manuel{Computing system details: MANUEL}{%
@@
   running CentOS Linux release 7.4.1708 (Core).  }
+
+\subsubsection{Benchmark problems.} Experiments with real-world expensive black-box problems would be computationally infeasible.
+Instead, we consider classical combinatorial optimization problems as black-box optimization problems.
+
+The \textbf{LOP} \ldots
+
+The \emph{Quadratic Assignment Problem (QAP)} is defined by:
+\begin{equation}\label{eq:qap}
+  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^n \sum_{j=1}^n a_{i,j} b_{\sigma(i), \sigma(j)}
+\end{equation}
+%
+where $[a_{i,j}]$ and $[b_{i,j}]$ are two matrices of size $n \times n$.
+
+The \emph{Permutation Flowshop Scheduling Problem (PFSP)} is defined by a matrix $[p_{ij}]$ of size $n \times M$ that gives the processing time of a job $i \in [n]$ on a machine $j \in [M]$.
+All jobs must be processed by all machines in the same order.
+Given a permutation of the jobs $\sigma$, $C_{i,j}$ denotes the completion time of job at position $\sigma(i)$ on machine $j$, and $C_{\max} = C_{n,M}$  is the completion time of the last job on the last machine, i.e., the makespan.
+The objective of the PFSP is to find:
+\begin{equation}\label{eq:qap}
+  \begin{split}
+    \min_{\sigma \in S_n}\quad& f(\sigma) = C_{n,M}\\
+    \text{s.t.}\quad & C_{1,j} = 0 \quad j\in [M], \qquad C_{i,1} = 0\quad i \in [n]\\
+    &C_{i,j} = p_{\sigma(i),j} + \max\{C_{i-1, j}, C_{i, j-1}\}  \ \  i \in \{2,\dotsc,n\},\ j \in\{2,\dotsc, M\}\\
+\end{split}
+\end{equation}
 
@@
 \texttt{nug12},\texttt{nug30},\texttt{tho30} and \texttt{kra32} for the QAP,
-with $n$ equal to the number in the instance name, and \texttt{reC05},
+with $n$ (the permutation length) equal to the number in the instance name, and \texttt{reC05},
 \texttt{reC13}, \texttt{reC19}, \texttt{reC31} for the PFSP, with
@@
 If a run went over 72 hours, of  CEGO was not able to reach 400 function evaluations (FEs) in 72 hours, it was terminated earlier.
-This was the case for all LOP instances shown.\label{tab:results}}
+This was the case for all LOP instances shown.\manuel{}{Don't we have results for rec31?}\label{tab:results}}
  \resizebox{\textwidth}{!}{%
-\begin{tabular}{*{6}{r}rl*{3}{r}}
+\begin{tabular}{r@{\hskip -2ex}*{5}{r}rl*{3}{r}}
  \toprule
-            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI for the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule{3-6}\cmidrule(lr){10-11}
+            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI for the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){10-11}
 \bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule
     LOP     & N-t1d100.01      & 78696.6                 & (811.4)    & 76119.6   & (915.4)                & $[$1763.5,    & 3390.5$]$     & 111.8& 72.5 & 3.2  \\
-    LOP     & N-t1d100.02      & 79686.5                 & (367.9)    & 76827.7   & (891.0)                & $[$2194.5,    & 3523.1$]$     & 110.7& 72.9 & 3.2  \\
-    LOP     & N-t1d150.01      & 187605.5                & (1906.9)   & 179508.3  & (1647.9)               & $[$6420.3,    & 9774.1$]$     & 73.9 & 72.7 & 6.5  \\
-    LOP     & N-t1d150.02      & 184320.9                & (973.5)    & 177592.8  & (1196.2)               & $[$5700.5,    & 7755.7$]$     & 74.1 & 73.2 & 6.5  \\
-    LOP     & N-t1d200.01      & 335391.1                & (2289.0)   & 323513.6  & (1371.2)               & $[$10076.1,   & 13678.9$]$    & 55.6 & 73.2 & 10.7 \\
-    LOP     & N-t1d200.02      & 333149.5                & (2497.2)   & 319675.8  & (2242.6)               & $[$11242.0,   & 15705.4$]$    & 56.5 & 73.7 & 10.6 \\
-    LOP     & N-t2d150.01      & 24682.7                 & (976.6)    & 15297.7   & (563.5)                & $[$8622.2,    & 10147.8$]$    & 74.6 & 72.9 & 7.0  \\
-    LOP     & N-t2d150.02      & 24250.8                 & (1354.9)   & 14724.9   & (667.5)                & $[$8495.1,    & 10556.7$]$    & 74.8 & 72.8 & 7.1  \\
-    LOP     & N-t2d200.01      & 57098.6                 & (4452.8)   & 33465.2   & (2001.0)               & $[$20284.5,   & 26982.3$]$    & 56.3 & 73.3 & 11.2 \\
-    LOP     & N-t2d200.02      & 54895.6                 & (4589.7)   & 35901.8   & (2077.5)               & $[$15539.1,   & 22448.5$]$    & 55.6 & 73.5 & 11.2 \\
-    LOP     & N-p50-01         & 21431.2                 & (299.5)    & 21027.6   & (711.8)                & $[$-128.1,    & 935.3$]$      & 219.9& 72.3 & 0.9  \\
-    LOP     & N-p50-02         & 22021.1                 & (275.6)    & 21613.6   & (596.4)                & $[$-42.5,     & 857.5$]$      & 217.3& 72.3 & 0.9  \\
-    LOP     & N-atp111         & 693.1                   & (10.3)     & 594.0     & (15.4)                 & $[$86.7,      & 111.5$]$      & 98.7 & 72.5 & 4.1  \\
-    LOP     & N-atp134         & 880.5                   & (18.6)     & 756.3     & (21.3)                 & $[$105.4,     & 143.0$]$      & 83.0 & 72.8 & 5.7  \\
-    LOP     & N-be75eec\_150   & 1666193.1               & (72310.8)  & 1415208.5 & (58593.5)              & $[$188960.6,  & 313008.6$]$   & 74.0 & 72.9 & 6.9  \\
-    LOP     & N-be75eec\_250   & 5214413.0               & (166701.4) & 4431271.0 & (109756.7)             & $[$649040.8,  & 917243.2$]$   & 44.9 & 73.7 & 15.5 \\
-    LOP     & N-be75np\_150    & 3742444.0               & (88500.2)  & 3204878.3 & (80704.1)              & $[$457944.2,  & 617187.2$]$   & 74.2 & 72.9 & 6.9  \\
-    LOP     & N-be75np\_250    & 10887427.8              & (159642.5) & 9544754.9 & (260523.9)             & $[$1136635.8, & 1548710.0$]$  & 45.3 & 74.0 & 15.5 \\
-   PFSP     & rec05.txt        & 1317.9                  & (25.1)     & 1330.7    & (17.9)                 & $[$-33.4,     & 7.8$]$        & 400.0& 38.9 & 0.1  \\
-   PFSP     & rec13.txt        & 2128.2                  & (25.0)     & 2158.7    & (29.5)                 & $[$-56.2,     & -4.8$]$       & 400.0& 38.9 & 0.1  \\
-   PFSP     & rec19.txt        & 2398.6                  & (19.8)     & 2405.1    & (12.6)                 & $[$-22.3,     & 9.3$]$        & 400.0& 82.1 & 0.3  \\
-    QAP     & kra32.dat        & 115907.0                & (4663.4)   & 116320.0  & (4673.5)               & $[$-4799.3,   & 3973.3$]$     & 400.0& 95.8 & 0.3  \\
-    QAP     & nug12.dat        & 624.2                   & (26.3)     & 660.4     & (17.3)                 & $[$-57.3,     & -15.1$]$      & 400.0& 14.8 & 0.0  \\
-    QAP     & nug30.dat        & 7486.8                  & (40.4)     & 7474.0    & (83.9)                 & $[$-50.8,     & 76.4$]$       & 400.0& 84.3 & 0.3  \\
-    QAP     & tho30.dat        & 192416.0                & (2584.2)   & 191482.0  & (4419.7)               & $[$-2527.0,   & 4395.0$]$     & 400.0& 85.1 & 0.3  \\
+     & N-t1d100.02      & 79686.5                 & (367.9)    & 76827.7   & (891.0)                & $[$2194.5,    & 3523.1$]$     & 110.7& 72.9 & 3.2  \\
+     & N-t1d150.01      & 187605.5                & (1906.9)   & 179508.3  & (1647.9)               & $[$6420.3,    & 9774.1$]$     & 73.9 & 72.7 & 6.5  \\
+     & N-t1d150.02      & 184320.9                & (973.5)    & 177592.8  & (1196.2)               & $[$5700.5,    & 7755.7$]$     & 74.1 & 73.2 & 6.5  \\
+     & N-t1d200.01      & 335391.1                & (2289.0)   & 323513.6  & (1371.2)               & $[$10076.1,   & 13678.9$]$    & 55.6 & 73.2 & 10.7 \\
+     & N-t1d200.02      & 333149.5                & (2497.2)   & 319675.8  & (2242.6)               & $[$11242.0,   & 15705.4$]$    & 56.5 & 73.7 & 10.6 \\
+     & N-t2d150.01      & 24682.7                 & (976.6)    & 15297.7   & (563.5)                & $[$8622.2,    & 10147.8$]$    & 74.6 & 72.9 & 7.0  \\
+     & N-t2d150.02      & 24250.8                 & (1354.9)   & 14724.9   & (667.5)                & $[$8495.1,    & 10556.7$]$    & 74.8 & 72.8 & 7.1  \\
+     & N-t2d200.01      & 57098.6                 & (4452.8)   & 33465.2   & (2001.0)               & $[$20284.5,   & 26982.3$]$    & 56.3 & 73.3 & 11.2 \\
+     & N-t2d200.02      & 54895.6                 & (4589.7)   & 35901.8   & (2077.5)               & $[$15539.1,   & 22448.5$]$    & 55.6 & 73.5 & 11.2 \\
+     & N-p50-01         & 21431.2                 & (299.5)    & 21027.6   & (711.8)                & $[$-128.1,    & 935.3$]$      & 219.9& 72.3 & 0.9  \\
+     & N-p50-02         & 22021.1                 & (275.6)    & 21613.6   & (596.4)                & $[$-42.5,     & 857.5$]$      & 217.3& 72.3 & 0.9  \\
+     & N-atp111         & 693.1                   & (10.3)     & 594.0     & (15.4)                 & $[$86.7,      & 111.5$]$      & 98.7 & 72.5 & 4.1  \\
+     & N-atp134         & 880.5                   & (18.6)     & 756.3     & (21.3)                 & $[$105.4,     & 143.0$]$      & 83.0 & 72.8 & 5.7  \\
+     & N-be75eec\_150   & 1666193.1               & (72310.8)  & 1415208.5 & (58593.5)              & $[$188960.6,  & 313008.6$]$   & 74.0 & 72.9 & 6.9  \\
+     & N-be75eec\_250   & 5214413.0               & (166701.4) & 4431271.0 & (109756.7)             & $[$649040.8,  & 917243.2$]$   & 44.9 & 73.7 & 15.5 \\
+     & N-be75np\_150    & 3742444.0               & (88500.2)  & 3204878.3 & (80704.1)              & $[$457944.2,  & 617187.2$]$   & 74.2 & 72.9 & 6.9  \\
+     & N-be75np\_250    & 10887427.8              & (159642.5) & 9544754.9 & (260523.9)             & $[$1136635.8, & 1548710.0$]$  & 45.3 & 74.0 & 15.5 \\\midrule
+   PFSP     & rec05        & 1317.9                  & (25.1)     & 1330.7    & (17.9)                 & $[$-33.4,     & 7.8$]$        & 400.0& 38.9 & 0.1  \\
+       & rec13        & 2128.2                  & (25.0)     & 2158.7    & (29.5)                 & $[$-56.2,     & -4.8$]$       & 400.0& 38.9 & 0.1  \\
+       & rec19        & 2398.6                  & (19.8)     & 2405.1    & (12.6)                 & $[$-22.3,     & 9.3$]$        & 400.0& 82.1 & 0.3  \\\midrule
+    QAP     & kra32        & 115907.0                & (4663.4)   & 116320.0  & (4673.5)               & $[$-4799.3,   & 3973.3$]$     & 400.0& 95.8 & 0.3  \\
+        & nug12        & 624.2                   & (26.3)     & 660.4     & (17.3)                 & $[$-57.3,     & -15.1$]$      & 400.0& 14.8 & 0.0  \\
+        & nug30        & 7486.8                  & (40.4)     & 7474.0    & (83.9)                 & $[$-50.8,     & 76.4$]$       & 400.0& 84.3 & 0.3  \\
+        & tho30        & 192416.0                & (2584.2)   & 191482.0  & (4419.7)               & $[$-2527.0,   & 4395.0$]$     & 400.0& 85.1 & 0.3  \\
 \bottomrule

### 2020-10-23 paper/main.tex
use figures from ../img/

@@
   \centering%
-  \includegraphics[width=\textwidth]{img/synthetic_LOP_combined}
+  \includegraphics[width=\textwidth]{../img/synthetic_LOP_combined}
   \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
@@
   \centering%
-  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d100_01}\\
-  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d150_01}\\
-  \includegraphics[width=\textwidth]{img/fitness_real_lop_RandA1_N-t1d200_01}\\
+  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d100_01}\\
+  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d150_01}\\
+  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d200_01}\\
     \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}
@@
   \centering%
-  \includegraphics[width=\textwidth]{img/fitness_real_pfsp_rec05_txt}
+  \includegraphics[width=\textwidth]{../img/fitness_real_pfsp_rec05_txt}
   \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
@@
   \centering%
-  \includegraphics[width=\textwidth]{img/fitness_real_qap_kra32_dat}
+  \includegraphics[width=\textwidth]{../img/fitness_real_qap_kra32_dat}
     \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.\label{fig:kra32}}

### 2020-10-24 paper/main.tex
improve intro

@@
 \usepackage[T1]{fontenc}
-
+\usepackage{relsize}
 \usepackage{amsmath,amssymb}
@@
 
-Motivation: Manuel
-
 In many practical optimization problems, the objective function is not
@@
 prediction models, simulations or physical experiments. When decision variables
-are continuous, the use of Bayesian surrogate-models, e.g., Gaussian processes,
-in optimization has become
-widespread~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
+are continuous, the use of Bayesian surrogate-models (e.g., Gaussian processes)
+is widespread in
+optimization~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
 success, there have been attempts at adapting such Bayesian optimization
@@
 on a black-box version of the travelling salesman problem under a budget of
-$1\,000$ function evaluations. ACO basically builds a probability distribution
-model from which solutions are sampled with a bias towards the best solutions
-evaluated so far. As such, ACO may be considered a type of
-estimation-of-distribution algorithm (EDA) for combinatorial optimization. On
-the other hand, $1\,000$ evaluations is a relatively large budget for CEGO and
-ACO was not designed for such short budgets.
+$1\,000$ function evaluations.  On the other hand, $1\,000$ evaluations is
+still a relatively large budget for CEGO, whereas ACO was not designed for such
+short budgets.
 
-In this work, we propose and analyze UMM, an EDA specifically designed for
-black-box combinatorial optimization under a limited budget. The proposed UMM 
-is a population based algorithm.
-There exists a wide body of research on population based probabilistic algorithms, where most prominent examples are Estimation of Distribution Algorithms (EDAs)~\citep{???}.
-Our approach drastically differs from EDAs since EDAs assume that function evaluations are cheap and they can do many of them.
-On the contrary, UMM minimizes the number of function evaluations.
-Moreover, UMM does not rely on an evolving sequence of samples (called generations in evolutionary approaches).
-Instead, UMM considers a single sample of permutations during the whole execution that is increased iteratively, with one permutation each iteration.
+In this work, we propose and analyze Unbalanced Mallows Model (UMM), a
+population-based probabilistic algorithm specifically designed for black-box
+combinatorial optimization under a limited budget. There exists a wide body of
+research on probabilistic algorithms~\citep{???},\MANUEL{what can we cite from
+  that body of research?} with the most prominent examples being estimation of
+distribution algorithms (EDAs)~\citep{???}.\MANUEL{what would be the best work
+  to cite about EDAs for combinatorial optimization?} ACO may also be
+considered a type of EDA, since it builds a probability distribution model from
+which solutions are sampled with a bias towards the best solutions evaluated so
+far. The proposed UMM drastically differs from EDAs since EDAs assume that
+function evaluations are cheap and they can do many of them. On the contrary,
+UMM aims at learning as much as possible with a limited number of function
+evaluations. Moreover, unlike EDAs, UMM does not rely on a fixed-size
+population of solutions that evolves over ``generations''. Instead, taking
+inspiration from Bayesian optimizers, UMM considers a sample of permutations
+during the whole execution that is increased iteratively by one new permutation
+each iteration. Finally, the balance between intensification and exploration is
+dynamically adapted by controlling the variance of the probabilistic Mallows model.
+\manuel{}{
+  \begin{itemize}
+  \item Who else has used Mallows model for black-box optimization?
+  \item Who else has used unbalanced borda for optimization?
+  \end{itemize}
+}
 
-\ldots\MANUEL{what is the inspiration}
-\EKHINE{we propose a probabilistic algrithm that (like cego?) (1) iteratively adds solutions to the sample and (2) fits a surrogate model for every new sample obtained.
+% \ldots\MANUEL{what is the inspiration}
+% \EKHINE{we propose a probabilistic algrithm that (like cego?) (1) iteratively adds solutions to the sample and (2) fits a surrogate model for every new sample obtained.
 However,  different fom cego, our surrogate model is an estimation of the predicted fitness function at each point.
@@
 
-% minimize the uncertainty in the final solutions or
+% % minimize the uncertainty in the final solutions or
 
-Related work: Ekhine
+% Related work: Ekhine
 
-Our contribution: Ekhine
-\begin{itemize}
-\item 
-\end{itemize}
+% Our contribution: Ekhine
+% \begin{itemize}
+% \item 
+% \end{itemize}
 
@@
 \citet{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, i.e.,
-\texttt{nug12},\texttt{nug30},\texttt{tho30} and \texttt{kra32} for the QAP,
+\texttt{nug12}, \texttt{nug30}, \texttt{tho30} and \texttt{kra32} for the QAP,
 with $n$ (the permutation length) equal to the number in the instance name, and \texttt{reC05},
@@
 
+\begin{smaller}
 \paragraph*{Acknowledgements.}
-
+M.\@ L\'opez-Ib\'a\~nez is a ``Beatriz Galindo'' Senior Distinguished Researcher (BEAGAL 18/00053) funded by the Ministry of Science and Innovation of the Spanish Government.
+%
 Thanks to Hao Wang (Leiden University) for pointing us to the arguments of
 \citet{EriPeaGar2019scalable}.
+\end{smaller}
 

### 2020-10-24 paper/main.tex

@@
 landscapes~\citep{ZaeStoBar2014:ppsn}, such as Kendall's $\tau$ distance for
-permutations (Eq.~\ref{eq:kendall_decomp}). In CEGO, the surrogate model is explored by a GA with
-crossover and mutation operators appropriate for the particular combinatorial
-problem. The original paper notes that coupling the GA with local search does
-not improve the results significantly since the model is anyway an inexact
-estimation of the original fitness
+permutations (Eq.~\ref{eq:kendall_decomp}). In CEGO, the surrogate model is
+explored by a GA with crossover and mutation operators appropriate for the
+particular combinatorial problem. The original paper notes that coupling the GA
+with local search does not improve the results significantly since the model is
+anyway an inexact estimation of the original fitness
 function~\citep[p.~875]{ZaeStoFriFisNauBar2014}.
@@
 For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{}.\MANUEL{X where?
-Do we miss a citation here?\ekhine{}{to be processed, arxiv}}
+Do we miss a citation here?\ekhine{}{to be processed, arxiv} Maybe you can create a bibentry with the details that you know already.}
 
@@
 
+\subsubsection{CEGO.}
 We use the implementation of GECO provided by the
-authors\footnote{\url{https://cran.r-project.org/package=CEGO}}. Following the
-authors of CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use and
-the GA that optimizes the surrogate models uses a population size of 20,
-crossover rate of 0.5, mutation rate of 1, tournament selection of size 2 and
-probability of 0.9, interchange mutation (i.e., exchanging two randomly
-selected elements of the permutation) and cycle crossover for permutations. The
-authors of CEGO used a budget of $10^4$ evaluations using the surrogate-model
-for each run of the GA, however, with such setting, each run of CEGO requires
-more than a week of computation time, which was infeasible for us, hence, we
-set a value of \textcolor{red}{$10^3$}.
+authors.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
+Following the authors of
+CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use a GA to optimize
+the surrogate models, with population size of 20, crossover rate of 0.5,
+mutation rate of 1, tournament selection of size 2 and probability of 0.9,
+interchange mutation (i.e., exchanging two randomly selected elements) and
+cycle crossover for permutations. To generate each new solution, CEGO runs the
+GA using the model as a surrogate of the true fitness function.  The authors of
+CEGO used a budget of $10^4$ evaluations of the surrogate-model for each run of
+the GA. However, with such a setting, one run of CEGO with a budget of $m=400$
+evaluations of the true fitness function requires more than a week of
+computation time, which was infeasible for our study.
+Hence, we set the budget of the GA to $10^3$.
+Although this means that the results of CEGO reported here could, in principle, improved with a higher GA budget  we believe
 
@@
 The CPU time required for a single run of CEGO using these settings is larger
-than a week.  For simplicity, we use Kendall Tau distance~\citep{?} in all
+than a week.  For simplicity, we use Kendall's $\tau$ distance~(Eq.~\ref{eq:kendall_decomp}) in all
 experiments\MANUEL{\citet{ZaeStoBar2014:ppsn} claims this is called Swap
@@
 In particular, the mean fitness (and standard deviation) of the best solution found at the end of each run,
-\MANUEL{is the CI for CEGO $-$ UMM or UMM $-$ CEGO? I would prefer}
 In terms of best fitness achieved on each run, and taking into account the
@@
  \resizebox{\textwidth}{!}{%
-\begin{tabular}{r@{\hskip -2ex}*{5}{r}rl*{3}{r}}
+\begin{tabular}{r@{\hskip -2ex}*{5}{r}rl@{\hskip -2ex}*{3}{r}}
  \toprule
-            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI for the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){10-11}
+            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack[r]{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){10-11}
 \bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule

### 2020-10-24 paper/main.tex
improve 5

@@
 \DeclareMathOperator*{\argmax}{arg\,max}
+
+\newcommand{\Prob}{\ensuremath{p}}
+\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}
+%\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}
+\newcommand{\FEmax}{\ensuremath{m}}
 
@@
 %
-\newcommand{\Prob}{\ensuremath{p}}
-\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}
-\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}
 
@@
   \begin{algorithmic}[1]
-    \Require $\minit$: number of initial permutations, $m$: total budget, $\rho$: learning rate
-    %\\ ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
+    \Require $\minit$: number of initial permutations, $\FEmax$: total budget, $\rho$: learning rate
+    %\\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
     \State $S \assign $ generate $\minit$ permutations uniformly at random
     \State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
-    \For{$i \assign \minit$ \textbf{to} $m$ \text{evaluations}}
+    \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
     \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}
@@
 This candidate permutation (solution) is evaluated and added to the sample $S$.
-The process is repeated up to a given maximum budget of evaluations $m$.\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it.
+The process is repeated up to a given maximum budget of evaluations $\FEmax$.\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it.
 Can we give a bit more detail?}
@@
 authors.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
-Following the authors of
-CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use a GA to optimize
-the surrogate models, with population size of 20, crossover rate of 0.5,
-mutation rate of 1, tournament selection of size 2 and probability of 0.9,
-interchange mutation (i.e., exchanging two randomly selected elements) and
-cycle crossover for permutations. To generate each new solution, CEGO runs the
-GA using the model as a surrogate of the true fitness function.  The authors of
-CEGO used a budget of $10^4$ evaluations of the surrogate-model for each run of
-the GA. However, with such a setting, one run of CEGO with a budget of $m=400$
-evaluations of the true fitness function requires more than a week of
-computation time, which was infeasible for our study.
-Hence, we set the budget of the GA to $10^3$.
-Although this means that the results of CEGO reported here could, in principle, improved with a higher GA budget  we believe
-
 Although it is never stated in the original paper, the implementation of CEGO
@@
 sequentially by maximizing the minimum distance to solutions already in the
-set. These initial solutions are then evaluated on the actual objective
-function and the result is used to build the initial surrogate
-model.% \citet{ZaeStoBar2014:ppsn} uses 10
+set. These initial solutions are then evaluated on the true fitness function
+and the result is used to build the initial surrogate
+model.% \citet{ZaeStoBar2014:ppsn} uses \minit=10
+Following the authors of
+CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use a GA to optimize
+the surrogate models with population size of 20, crossover rate of 0.5,
+mutation rate of $1/n$, tournament selection of size 2 with probability of 0.9,
+interchange mutation (i.e., exchanging two randomly selected elements) and
+cycle crossover for permutations. To generate each new solution, CEGO runs the
+GA using the model as a surrogate of the true fitness function.  The authors of
+CEGO used a budget of $10^4$ surrogate-model evaluations for each run of the
+GA. However, with such a setting, a single run of CEGO with a budget of $\FEmax=400$
+true fitness evaluations requires \emph{more than a week} of computation time,
+which was infeasible for our study. Hence, we set the budget of the GA to
+$10^3$. Although a higher GA budget could improve the results of CEGO reported
+here, our results are similar (and sometimes better) than the ones reported in
+the original papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to
+running CEGO with $\FEmax=400$ instead of $\FEmax=200$. In fact, we found that the
+increase in runtime caused by doubling $\FEmax$ is much smaller than by increasing
+the GA budget by a factor of $10$. Nevertheless, we caution that choosing a
+different trade-off may lead to different conclusions and more work is needed
+to understand CEGO parameters.
 
-The CPU time required for a single run of CEGO using these settings is larger
-than a week.  For simplicity, we use Kendall's $\tau$ distance~(Eq.~\ref{eq:kendall_decomp}) in all
+\subsubsection{Other}.  For simplicity, we use Kendall's $\tau$
+distance~(Eq.~\ref{eq:kendall_decomp}) in all
 experiments\MANUEL{\citet{ZaeStoBar2014:ppsn} claims this is called Swap
@@
 distance measures and to the dynamic selection of distance
-measures~\citep{ZaeStoBar2014:ppsn}.
-
-In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
+measures~\citep{ZaeStoBar2014:ppsn}. xIn all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
 the actual objective function. In a white-box context, state-of-the-art
@@
 
-For the LOP, we generated synthetic instances of size $n=20$, $m=200$, $\phi\in\{0.5,0.7,0.9\}$\MANUEL{what is $m$?}
+For the LOP, we generated synthetic instances of size $n=20$, $m=200$, $\phi\in\{0.5,0.7,0.9\}$\MANUEL{what is $m$ here? We already use $\FEmax$ for the budget, but we could change it. }
 

### 2020-10-25 paper/main.tex
remove stray x
paper

@@
 
-Distances in permutation space can be measured in various ways.  Kendall's $\tau$ distance counts the number of items ranked in different order in both permutations:
+Distances in permutation space can be measured in various ways.  Kendall's-$\tau$ distance counts the number of items ranked in different order in both permutations:
 %
@@
 d(\sigma _{1},\sigma_{2})=\left\lvert\Bigl\{(i,j):i<j \land \phantom{}\Bigl(\right.&\bigl(\sigma_{1}(i) < \sigma_{1}(j)\land \sigma_{2}(i)>\sigma_{2}(j)\bigr)\lor \phantom{} \\
- &\bigl(\sigma_{1}(i)>\sigma_{1}(j)\land \sigma_{2}(i)<\sigma_{2}(j)\bigr)\Bigr)\Bigr\}\Bigr\rvert\\
+ &\bigl(\sigma_{1}(i)>\sigma_{1}(j)\land \sigma_{2}(i)<\sigma_{2}(j)\bigr)\Bigr)\Bigr\}\Bigr\rvert\\.
  \end{split}
@@
 \end{equation}
-%
-An alternative\MANUEL{Alternative or equivalent?
+
+An equivalent\MANUEL{X Alternative or equivalent?
 That is, it is the same numbers or it will produce a different result?} definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made in $\sigma_1^{-1}$ to obtain $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
@@
 One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
-The MM defines the probability of each permutation $\sigma\in S_n$ as follows:\MANUEL{Could you check this?
+The MM defines the probability of each permutation $\sigma\in S_n$ as follows:\MANUEL{X Could you check this?
 I wanted to introduce the notation $MM(\sigma_0, \theta)$ that we use later}
@@
 The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.
-The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance (Eq.~\ref{eq:kendall_decomp}) as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
+The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance (Eq.~\eqref{eq:kendall_decomp}) as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
 
-\manuel{One of the fundamental problems in statistics is to learn the parameters that best fit a given sample of data.}{}\MANUEL{I think we don't need to emphasize this, it is known} %
 We will use the standard maximum likelihood estimation (MLE) approach to fit the parameters of a MM for a given collection of permutations. 
@@
 
-The exact MLE for $\hat\sigma_0$\MANUEL{MLE for $\sigma_0$ or for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
+The exact MLE for $\sigma_0$\MANUEL{X MLE for $\sigma_0$ or for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
 Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}.
-An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda error is bounded~\cite{CopFleRur2010ordering}.\MANUEL{I'm not sure what the latter means}
+An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda is an approximation to Kemeny~\cite{CopFleRur2010ordering}.\MANUEL{X I'm not sure what the latter means}
 
@@
 %
-and, the same as Borda, orders the \ekhine{positions}{items}\MANUEL{why items? we talk about positions above. We could talk about decision variables if that is clearer} increasingly by their score.
+and, the same as Borda, orders the \ekhine{items}{positions}\MANUEL{X why items? we talk about positions above. We could talk about decision variables if that is clearer} increasingly by their score.
 
@@
 Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
-%Intuitively, the uBorda ranking a given sample of permutations $S$\MANUEL{X What is $S$?} is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.\MANUEL{X so $w$ is a weight function from $w\colon S_n \to \mathbb{R}$?} %
 Clearly, the choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
@@
     \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}
-    \State linear decrease on $\theta$
+    \State decrease on $\theta$\label{line:theta}
     \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
@@
 Then, using the current sample $S$, the uBorda algorithm is applied to find a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}).
-Together with the parameter $\theta$, they define a Mallows model (Eq.~\ref{eq:MM}) from which we sample a new candidate permutation (line~\ref{line:sample}).
+Together with the parameter $\theta$, they define the Mallows model $MM(\hat\sigma_0, \theta)$.
+A decreasing $\theta$ parameter (line~\ref{line:theta}) implies a decreasing variance of the Mallows model (Eq.~\ref{eq:MM}), more details in Section~\ref{sec:sampling}.
+Then, we sample a new candidate permutation from  $MM(\hat\sigma_0, \theta)$ (line~\ref{line:sample}).
 This candidate permutation (solution) is evaluated and added to the sample $S$.
@@
 transformed sample is the same, i.e., polynomial. The most important steps are
-the sampling of a new permutation from $MM(\sigma_0,\rho)$ and the update
-(\emph{learning}) of the parameters $\sigma_0$ and $\rho$.\MANUEL{Is
+the sampling of a new permutation from $MM(\sigma_0,\theta)$ (line~\ref{line:sample} in Algorithm~\ref{alg:umm}) and the update
+(\emph{learning}) of the parameters $\sigma_0$ $\theta$ and $\rho$ (line~\ref{line:uborda} in Algorithm~\ref{alg:umm}).\MANUEL{X Is
   this $\rho$ the same we explained earlier for the weights?}
@@
 
-\newcommand{\Edist}{\ensuremath{\text{E}[D]}}
+\newcommand{\Edist}{\ensuremath{\mathbb{E}[D]}}
 
-\subsubsection{Sampling step.}
-Sampling from $MM(\sigma_0,\rho)$  can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{Collas,IruCalLoz2016permallows,Regenwetter2004}\MANUEL{Which paper is Collas?
-It is not found.}\MANUEL{You call this Regenwetter, but the author is Doignon?
-\ekhine{}{he is another coauthor} But the entry shows only 1 author.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{respectively?}
-%
-At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained as a solution in the previous learning step.
-The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM.\MANUEL{I'm not sure what is $D$, distance from where to where?} It is set automatically by the algorithm to control the diversification and intensification trade off.
-At the first iteration, it is set so that the expected distance $\Edist$ equals  $\binom{n}{4}$ (half of the distance of the uniform distribution), and then the expected distance linearly decrease until the last iteration where the expected distance equals 1.
-The expected distance, shown in the next line, does not have a closed-form as function of $\theta$ but its monotonicity guarantees low-cost, high-quality approximations:\MANUEL{I don't understand this sentence, sorry.
-You say it does not have a closed-form, but you show a closed-form.
-Do you mean that the formula below is only an approximation of $\Edist$?
-What is low-cost here?
-what is high-quality?}\MANUEL{What is $k$ in the equation?}
-%
-\begin{equation}\label{eq:expectation}
-\Edist = \frac{k \cdot \exp(-\theta)}{1-\exp(-\theta)} - \sum_{j=n-k+1}^{n}  \frac{j \exp(-j \theta)}{1-\exp(-j\theta)}
-\end{equation}
-
-A purely explotation approach\ekhine{esto se dice asi?
-uno greedy igual esta mejor dicho} would systematically generate $\sigma_0$ in the sampling stage.\MANUEL{I'm not sure what ``systematically generate $\sigma_0$'' means in this context.} However, this setting\MANUEL{which setting?
-$\theta$?}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
-Similar approaches have been taken in different contexts~\cite{arza2019approaching}.\MANUEL{there is something wrong with this reference\ekhine{}{qué le pasa?} Don't you see a complete mess of the authors list?
-It looks like it has 20 authors.}
-
-\subsubsection{Learning step.}
-As explained above, the parameter $\theta$ is updated deterministically at the
-sampling step. The parameter $\sigma_0$, however, is learned from the current
+\subsubsection{Learning step.}\ekhine{ }{conceptuaplemet sampling va después}
+\ekhine{As explained above, the parameter $\theta$ is updated deterministically at the
+sampling step.}{At each iteration we start by fitting a MM to the sample $S$, which implies estimating $\hat\sigma_0$ and $\theta$. $\theta$ is updated deterministically at the sampling step.} 
+The parameter $\sigma_0$, however, is learned from the current
 sample $S$ by using the uBorda algorithm with a weight
 $w(\sigma)=\rho^{f(\sigma)}$ (Section~\ref{sec:uborda}). The learning rate
-is controlled by parameter $\rho \in (0,1)$, which currently is user-defined.
+is controlled by parameter $\rho \in (0,1)$.
+\ekhine{, which currently is user-defined.
 Higher values or $\rho$ bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas smaller values make the sampling more uniform, thus slowing down the learning.
+}{}
+In this paper, the value of $\rho$ is set at each iteration and depends on the set of function evaluations $F$.
+Parameter $\rho$ is set such that the largest $r_1$\% of the mass of the weights is concentrated in the best $r_2$\% of the solutions in $S$, i.e.,
+\begin{equation}
+\begin{split}
+r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma\in S'}\rho^{f(\sigma)} \text{where} |S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\geq f(\sigma)
+\end{split}
+\end{equation}
 
@@
 
-\ekhine{bayesian optim}{ff}
-\EKHINE{subir a arxiv uBorda}
+\subsubsection{Sampling step.}\label{sec:sampling}
+Sampling from $MM(\sigma_0,\theta)$  can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{Collas,IruCalLoz2016permallows,Regenwetter2004}\MANUEL{Which paper is Collas?
+It is not found.}\MANUEL{You call this Regenwetter, but the author is Doignon?
+\ekhine{}{he is another coauthor} But the entry shows only 1 author.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{respectively?}
+%
+At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained as a solution in the previous learning step.
+The scale $\theta$ controls the variance and expected distance of the MM $\Edist$, i.e., the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$.
+\MANUEL{X I'm not sure what is $D$, distance from where to where?} It is set automatically by the algorithm to control the diversification and intensification trade off.
+At the first iteration, it is set so that the expected distance $\Edist$ equals  $\binom{n}{4}$ (half of the distance of the uniform distribution), and then the expected distance linearly decrease until the last iteration where the expected distance equals 1.
+Despite it is not possible to isolate $\theta$ in the $\Edist$ expression, Equation~\eqref{eq:expectation}, its monotonicity allows the use of bisection methods for efficient solution approximations.
+\MANUEL{X I don't understand this sentence, sorry.
+You say it does not have a closed-form, but you show a closed-form.
+Do you mean that the formula below is only an approximation of $\Edist$?
+What is low-cost here?
+what is high-quality?}\MANUEL{X What is $k$ in the equation?}
+%
+\begin{equation}\label{eq:expectation}
+\Edist = \frac{n \cdot \exp(-\theta)}{1-\exp(-\theta)} - \sum_{j=1}^{n}  \frac{j \exp(-j \theta)}{1-\exp(-j\theta)}
+\end{equation}
+
+A purely explotation approach would always generate $\sigma_0$ in the sampling stage.\MANUEL{X I'm not sure what ``systematically generate $\sigma_0$'' means in this context.} However, the setting of a decreasing variance \MANUEL{X which setting?
+$\theta$?}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
+Similar approaches have been taken in different contexts~\cite{arza2019approaching}.\MANUEL{there is something wrong with this reference\ekhine{}{en mi mendeley esta bien :s que puede ser?} }
 
@@
 distance measures and to the dynamic selection of distance
-measures~\citep{ZaeStoBar2014:ppsn}. xIn all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
+measures~\citep{ZaeStoBar2014:ppsn}. In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
 the actual objective function. In a white-box context, state-of-the-art
@@
 \ekhine{}{In prev sections we say that we show here how to tune $rho$. Le ves sitio? con los heatmaps? al principio de la sección? an outline here}
-\ekhine{}{recall that $w(\sigma) = \rho^{f(sigma)}$. We select $\rho$ in such a way that the 100$r\%$ of the samples have the 100$w\%$ of the weight}
+\ekhine{}{recall that $w(\sigma) = \rho^{f(sigma)}$. We select $\rho$ in such a way that the 100$r_1\%$ of the samples have the 100$w\%$ of the weight}
+
+The first part of the experimental analysis concerns the tunning of parameter $\rho$.
+As stated in Section~\ref{sec:sampling}, $\rho$ is automatically set at each iteration in such a way that the 100$r_1\%$ of the samples have the 100$r_2\%$ of the weight.
+In this regard, we tried different configurations of parameters $r_1$ and $r_2$ in all the instances.
+Figure~\ref{fig:heatmaps} shows the best solution (averaged for 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the PFSP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP.
+
+\begin{figure}[H]
+  \centering%
+  \includegraphics[width=0.24\textwidth]{../img/heatmap_N-t1d100_01}
+  \includegraphics[width=0.24\textwidth]{../img/heatmap_N-t1d150_01}
+  \includegraphics[width=0.24\textwidth]{../img/heatmap_rec05}
+  \includegraphics[width=0.24\textwidth]{../img/heatmap_kra32_dat}
+  \label{fig:heatmaps}
+  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
+\end{figure}
+
+Overall the instances we see that the configuration $(r_1, r_2)=(0.1, 0.9)$ has good performance.
+Therefore, for the lack consistency along the experimental part, we use this configuration.
+Moreover, we recommend this configuration as a first approach with UMM.
+
+Moreover, we can see that the behavior of the  two left figures differs from the two in the right: while the two in the left appear to be increase homogeneously from the top right corner to the bottom left one, the two in the right appear to be closer to randomness.
+We point out that this performance is related to the adequacy of UMM to the problem: homogeneousness is associated to the problem (not to the particular instances) and the more homogeneous the plot, the better performance of UMM.
+We elaborate this argument in the following sections with detail.
 

### 2020-10-26 paper/main.tex
some updates

@@
 \usepackage{amsmath,amssymb}
-%\usepackage[square,numbers,sort&compress]{natbib}
 \usepackage{xspace}
@@
 and the result is used to build the initial surrogate
-model.% \citet{ZaeStoBar2014:ppsn} uses \minit=10
+model. % \citet{ZaeStoBar2014:ppsn} uses \minit=10
 Following the authors of
@@
 
+\subsubsection{UMM}. UMM uses the same budget of evaluation ($\FEmax$) as CEGO
+and initially samples $\minit=10$ solutions uniformly. The other two parameters
+of UMM are $r_1$ and $r_2$, which determine the balance between exploration and
+exploitation. In the experimental section we will study their impact. We
+implemented UMM using Python 3.
+
 \subsubsection{Other}.  For simplicity, we use Kendall's $\tau$
@@
 \citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
-when using Hamming distance. We plan to extend our analysis of UMM to other
-distance measures and to the dynamic selection of distance
-measures~\citep{ZaeStoBar2014:ppsn}. In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
+when using Hamming distance.  We plan to extend UMM to other distance measures
+and to the dynamic selection of distance measures~\citep{ZaeStoBar2014:ppsn}.
+In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
 the actual objective function. In a white-box context, state-of-the-art
@@
 here for the black-box context is extremely limited.
-
-We implemented UMM in Python 3. The parameter settings that we use are...
-
-\manuel{Computing system details: MANUEL}{%
-  All experiments were run on Intel Xeon ``Ivybridge'' E5-2650v2 CPUs at 2.60\,GHz, 64\,GB RAM
-  running CentOS Linux release 7.4.1708 (Core).  }
 
@@
 $n \in \{20, 20, 30, 50\}$, respectively.
+
+\subsubsection{Computing environment.}
+All experiments were run on Intel Xeon ``Ivybridge'' E5-2650v2 CPUs at
+2.60\,GHz, 64\,GB RAM running CentOS Linux release 7.4.1708 (Core).
+\MANUEL{This may need updating since you ran some experiments? didn't you?}
 

### 2020-10-26 paper/main.tex
add title
intro

@@
 
-\title{Mallows, Black-box combinatorial optimization, limited budget,  ???}
+\title{Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems}
 % Double-blind
@@
 In this work, we propose and analyze Unbalanced Mallows Model (UMM), a
-population-based probabilistic algorithm specifically designed for black-box
-combinatorial optimization under a limited budget. There exists a wide body of
-research on probabilistic algorithms~\citep{???},\MANUEL{what can we cite from
-  that body of research?} with the most prominent examples being estimation of
-distribution algorithms (EDAs)~\citep{???}.\MANUEL{what would be the best work
-  to cite about EDAs for combinatorial optimization?} ACO may also be
-considered a type of EDA, since it builds a probability distribution model from
-which solutions are sampled with a bias towards the best solutions evaluated so
-far. The proposed UMM drastically differs from EDAs since EDAs assume that
-function evaluations are cheap and they can do many of them. On the contrary,
-UMM aims at learning as much as possible with a limited number of function
-evaluations. Moreover, unlike EDAs, UMM does not rely on a fixed-size
-population of solutions that evolves over ``generations''. Instead, taking
-inspiration from Bayesian optimizers, UMM considers a sample of permutations
-during the whole execution that is increased iteratively by one new permutation
-each iteration. Finally, the balance between intensification and exploration is
-dynamically adapted by controlling the variance of the probabilistic Mallows model.
-\manuel{}{
+population-based probabilistic algorithm specifically designed for optimizing
+black-box problems over a permutation landscape under a limited budget. There
+exists a wide body of research on probabilistic
+algorithms~\citep{???},\MANUEL{what can we cite from that body of research?}
+with the most prominent examples being estimation of distribution algorithms
+(EDAs)~\citep{???}.\MANUEL{what would be the best work to cite about EDAs for
+  combinatorial optimization?} ACO may also be considered a type of EDA, since
+it builds a probability distribution model from which solutions are sampled
+with a bias towards the best solutions evaluated so far. The proposed UMM
+drastically differs from EDAs since EDAs assume that function evaluations are
+cheap and they can do many of them. On the contrary, UMM aims at learning as
+much as possible with a limited number of function evaluations. Moreover,
+unlike EDAs, UMM does not rely on a fixed-size population of solutions that
+evolves over ``generations''. Instead, taking inspiration from Bayesian
+optimizers, UMM considers a sample of permutations during the whole execution
+that is increased by one new permutation at each iteration. Finally,
+the balance between intensification and exploration is dynamically adapted by
+controlling the variance of the probabilistic Mallows model.  \manuel{}{
   \begin{itemize}

### 2020-10-26 paper/main.tex
more comments and fixes

@@
 \usepackage[noend]{algpseudocode}
+%\algrenewcommand{\algorithmiccomment}[1]{\hfill\texttt{//} #1}
+
 \usepackage{booktabs} % For formal tables
@@
 
-Distances in permutation space can be measured in various ways.  Kendall's-$\tau$ distance counts the number of items ranked in different order in both permutations:
+Distance in permutation space can be measured in various ways.  Kendall's-$\tau$ counts the number of items ranked in different order in both permutations:
 %
@@
 d(\sigma _{1},\sigma_{2})=\left\lvert\Bigl\{(i,j):i<j \land \phantom{}\Bigl(\right.&\bigl(\sigma_{1}(i) < \sigma_{1}(j)\land \sigma_{2}(i)>\sigma_{2}(j)\bigr)\lor \phantom{} \\
- &\bigl(\sigma_{1}(i)>\sigma_{1}(j)\land \sigma_{2}(i)<\sigma_{2}(j)\bigr)\Bigr)\Bigr\}\Bigr\rvert\\.
+ &\bigl(\sigma_{1}(i)>\sigma_{1}(j)\land \sigma_{2}(i)<\sigma_{2}(j)\bigr)\Bigr)\Bigr\}\Bigr\rvert\\
  \end{split}
@@
 
-An equivalent\MANUEL{X Alternative or equivalent?
-That is, it is the same numbers or it will produce a different result?} definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made in $\sigma_1^{-1}$ to obtain $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
+An equivalent definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made in $\sigma_1^{-1}$ to obtain $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
 
@@
 One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
-The MM defines the probability of each permutation $\sigma\in S_n$ as follows:\MANUEL{X Could you check this?
-I wanted to introduce the notation $MM(\sigma_0, \theta)$ that we use later}
+The MM defines the probability of each permutation $\sigma\in S_n$ as follows:
 %
@@
 
-The exact MLE for $\sigma_0$\MANUEL{X MLE for $\sigma_0$ or for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
+The exact MLE for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
 Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}.
-An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda is an approximation to Kemeny~\cite{CopFleRur2010ordering}.\MANUEL{X I'm not sure what the latter means}
+An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda is an approximation to Kemeny~\cite{CopFleRur2010ordering}.\MANUEL{I'm not sure ``in a general optimization perspective'' means}
 
@@
     \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}
-    \State decrease on $\theta$\label{line:theta}
+    \State Decrease  $\theta$\label{line:theta}\Comment{Eq.~\ref{eq:expectation:}}
     \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
@@
 
-As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are randomly generated (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated.
+As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are uniformly randomly generated (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated.
 Then, using the current sample $S$, the uBorda algorithm is applied to find a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}).
 Together with the parameter $\theta$, they define the Mallows model $MM(\hat\sigma_0, \theta)$.
-A decreasing $\theta$ parameter (line~\ref{line:theta}) implies a decreasing variance of the Mallows model (Eq.~\ref{eq:MM}), more details in Section~\ref{sec:sampling}.
+A decreasing $\theta$ parameter (line~\ref{line:theta}) implies a decreasing variance of the Mallows model (Eq.~\ref{eq:MM}), as explained in  Section~\ref{sec:sampling} below.
 Then, we sample a new candidate permutation from  $MM(\hat\sigma_0, \theta)$ (line~\ref{line:sample}).
 This candidate permutation (solution) is evaluated and added to the sample $S$.
-The process is repeated up to a given maximum budget of evaluations $\FEmax$.\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it.
-Can we give a bit more detail?}
+The process is repeated up to a given maximum budget of evaluations $\FEmax$.\MANUEL{If the learning step happens later, when does it happen here?}
+\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it. Can we give a bit more detail?}
 
@@
 $f(\sigma)$, i.e., it behaves as if there were many copies of the best
-solutions and few of the bad ones. A crucial property of UMM is that the
+solutions and few of the bad ones. A \manuel{crucial}{} property of UMM is that the
 computational complexity of working with the original sample and with the
@@
 the sampling of a new permutation from $MM(\sigma_0,\theta)$ (line~\ref{line:sample} in Algorithm~\ref{alg:umm}) and the update
-(\emph{learning}) of the parameters $\sigma_0$ $\theta$ and $\rho$ (line~\ref{line:uborda} in Algorithm~\ref{alg:umm}).\MANUEL{X Is
-  this $\rho$ the same we explained earlier for the weights?}
+(\emph{learning}) of the parameters $\sigma_0$, $\theta$ and $\rho$ (line~\ref{line:uborda} in Algorithm~\ref{alg:umm}).
 
@@
 
-\subsubsection{Learning step.}\ekhine{ }{conceptuaplemet sampling va después}
-\ekhine{As explained above, the parameter $\theta$ is updated deterministically at the
-sampling step.}{At each iteration we start by fitting a MM to the sample $S$, which implies estimating $\hat\sigma_0$ and $\theta$. $\theta$ is updated deterministically at the sampling step.} 
-The parameter $\sigma_0$, however, is learned from the current
-sample $S$ by using the uBorda algorithm with a weight
-$w(\sigma)=\rho^{f(\sigma)}$ (Section~\ref{sec:uborda}). The learning rate
-is controlled by parameter $\rho \in (0,1)$.
-\ekhine{, which currently is user-defined.
-Higher values or $\rho$ bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas smaller values make the sampling more uniform, thus slowing down the learning.
-}{}
-In this paper, the value of $\rho$ is set at each iteration and depends on the set of function evaluations $F$.
-Parameter $\rho$ is set such that the largest $r_1$\% of the mass of the weights is concentrated in the best $r_2$\% of the solutions in $S$, i.e.,
+\subsubsection{Learning step.}
+At each iteration we start by fitting a MM to the sample $S$, which implies
+estimating $\hat\sigma_0$ and $\theta$. Parameter $\theta$ is updated
+deterministically at the sampling step.  The parameter $\sigma_0$, however, is
+learned from the current sample $S$ by using the uBorda algorithm with a weight
+$w(\sigma)=\rho^{f(\sigma)}$ (Section~\ref{sec:uborda}). The learning rate is
+controlled by parameter $\rho \in (0,1)$. %
+In UMM, the value of $\rho$ depends on the set of true fitness evaluations $F$
+and it is set at each iteration such that the largest $r_1$\% of the mass of
+the weights is concentrated in the best $r_2$\% of the solutions in $S$, i.e.,
+\MANUEL{Let's talk about this Equation tomorrow}
+%
 \begin{equation}
 \begin{split}
-r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma\in S'}\rho^{f(\sigma)} \text{where} |S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\geq f(\sigma)
+r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma\in S'}\rho^{f(\sigma)} \quad\text{where}\quad |S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\geq f(\sigma)
 \end{split}
 \end{equation}
+%
+\MANUEL{should the $\sigma$ be the same on both sides of $=$?}
+where $S'$ contains the $r_2|S|$ elements of $S$ with the best corresponding fitness values in $F$.
+The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification: higher values bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas smaller values make the sampling more uniform, thus slowing down the learning.
 
-% As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two steps (1) estimating the location parameter $\sigma_0$ and (2) computing the scale parameter $\theta$.
+% As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two= steps (1) estimating the location parameter $\sigma_0$ and (2) computing the scale parameter $\theta$.
 Since the scale parameter $\theta$ is set in the sampling step,\MANUEL{X How is it set?} we focus on estimating $\sigma_0$.
@@
 \subsubsection{Sampling step.}\label{sec:sampling}
-Sampling from $MM(\sigma_0,\theta)$  can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{Collas,IruCalLoz2016permallows,Regenwetter2004}\MANUEL{Which paper is Collas?
-It is not found.}\MANUEL{You call this Regenwetter, but the author is Doignon?
-\ekhine{}{he is another coauthor} But the entry shows only 1 author.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{respectively?}
+Sampling from $MM(\sigma_0,\theta)$  can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{Collas,IruCalLoz2016permallows,DoiPekReg2004rank}\MANUEL{Which paper is Collas?
+It is not found.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{respectively?}
 %
-At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained as a solution in the previous learning step.
-The scale $\theta$ controls the variance and expected distance of the MM $\Edist$, i.e., the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$.
-\MANUEL{X I'm not sure what is $D$, distance from where to where?} It is set automatically by the algorithm to control the diversification and intensification trade off.
-At the first iteration, it is set so that the expected distance $\Edist$ equals  $\binom{n}{4}$ (half of the distance of the uniform distribution), and then the expected distance linearly decrease until the last iteration where the expected distance equals 1.
-Despite it is not possible to isolate $\theta$ in the $\Edist$ expression, Equation~\eqref{eq:expectation}, its monotonicity allows the use of bisection methods for efficient solution approximations.
-\MANUEL{X I don't understand this sentence, sorry.
-You say it does not have a closed-form, but you show a closed-form.
-Do you mean that the formula below is only an approximation of $\Edist$?
-What is low-cost here?
-what is high-quality?}\MANUEL{X What is $k$ in the equation?}
+At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained in the previous learning step.
+The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM, i.e., $D$ is the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$:\MANUEL{where does this formula come from?}
 %
 \begin{equation}\label{eq:expectation}
-\Edist = \frac{n \cdot \exp(-\theta)}{1-\exp(-\theta)} - \sum_{j=1}^{n}  \frac{j \exp(-j \theta)}{1-\exp(-j\theta)}
+\Edist = \frac{n \cdot \exp(-\theta)}{1-\exp(-\theta)} - \sum_{j=1}^{n}  \frac{j\cdot\exp(-j  \theta)}{1-\exp(-j\theta)}
 \end{equation}
 
-A purely explotation approach would always generate $\sigma_0$ in the sampling stage.\MANUEL{X I'm not sure what ``systematically generate $\sigma_0$'' means in this context.} However, the setting of a decreasing variance \MANUEL{X which setting?
-$\theta$?}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
-Similar approaches have been taken in different contexts~\cite{arza2019approaching}.\MANUEL{there is something wrong with this reference\ekhine{}{en mi mendeley esta bien :s que puede ser?} }
+Parameter $\theta$ is set automatically by the algorithm to control the diversification and intensification trade off.
+At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$ (half of the distance of the uniform distribution), and then the expected distance linearly decreases until the last iteration where $\theta$ is set such that $\Edist = 1$.
+Despite it is not possible to isolate $\theta$ in Eq.~\eqref{eq:expectation}, its monotonicity allows the use of bisection methods for efficiently finding an approximation.
+
+A purely explotation approach would always generate $\sigma_0$ in the sampling stage.
+However, the setting of a decreasing variance \MANUEL{But we are controlling the expectation not the variance???}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
+Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.
 

### 2020-10-26 paper/main.tex
write conclusions

@@
 cost rather than wall-clock time. \citet{PerLopStu2015si} recently showed that
-ant colony optimization (ACO)~\citep{DorStu2004:book} is competitive with CEGO
+ant colony optimization (ACO) %~\citep{DorStu2004:book}
+is competitive with CEGO
 on a black-box version of the travelling salesman problem under a budget of
@@
 permutations. Here, we will study budgets lower than 400. Such expensive
-black-box combinatorial problems arises in diverse contexts, for example,
+black-box combinatorial problems arise in diverse contexts, for example,
 protein folding~\citep{RomKraArn2012protein} and industrial
@@
 involve expensive simulations for which no closed-form mathematical description
-is available.
+is available. The black-box formulation and limited budget preclude most of the
+optimization techniques that are successful in combinatorial optimization, such
+as constructive and perturbation heuristics and local search, which makes this
+a particularly challenging scenario, perhaps even more than its continuous
+counter-part.
 
@@
 In UMM, the value of $\rho$ depends on the set of true fitness evaluations $F$
-and it is set at each iteration such that the largest $r_1$\% of the mass of
-the weights is concentrated in the best $r_2$\% of the solutions in $S$, i.e.,
+and it is set at each iteration such that the largest $100r_1$\% of the mass of
+the weights is concentrated in the best $100r_2$\% of the solutions in $S$, i.e.,
 \MANUEL{Let's talk about this Equation tomorrow}
@@
 algorithms for the problems considered here typically evaluate
-\textcolor{red}{hundreds? thousands?} of solutions, thus, the budget considered
+thousands of solutions, thus, the budget considered
 here for the black-box context is extremely limited.
@@
 
+\subsection{Analysis of $r_1$ and $r_2$ parameters}
+
 \ekhine{}{In prev sections we say that we show here how to tune $rho$. Le ves sitio? con los heatmaps? al principio de la sección? an outline here}
@@
 
-The first part of the experimental analysis concerns the tunning of parameter $\rho$.
-As stated in Section~\ref{sec:sampling}, $\rho$ is automatically set at each iteration in such a way that the 100$r_1\%$ of the samples have the 100$r_2\%$ of the weight.
-In this regard, we tried different configurations of parameters $r_1$ and $r_2$ in all the instances.
-Figure~\ref{fig:heatmaps} shows the best solution (averaged for 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the PFSP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP.
+The first part of the experimental analysis concerns the \manuel{tunning}{analyses} of parameter $\rho$.
+As stated in Section~\ref{sec:sampling}, $\rho$ is automatically set at each iteration in such a way that the $100r_1\%$ of the samples have the $100r_2\%$ of the weight.
+In this regard, we tried different configurations of parameters $r_1$ and $r_2$ in all the instances.\MANUEL{why did we try this values.} Figure~\ref{fig:heatmaps} shows the best solution\MANUEL{best solution found at the end of a run?} (averaged for 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP.
+\MANUEL{The caption says synthetic instances}
 
-\begin{figure}[H]
+\begin{figure}[htb]
   \centering%
@@
   \includegraphics[width=0.24\textwidth]{../img/heatmap_kra32_dat}
-  \label{fig:heatmaps}
-  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
+    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:heatmaps}}
 \end{figure}
 
-Overall the instances we see that the configuration $(r_1, r_2)=(0.1, 0.9)$ has good performance.
-Therefore, for the lack consistency along the experimental part, we use this configuration.
-Moreover, we recommend this configuration as a first approach with UMM.
+The plot shows that UMM is indeed very sensitive to the values of $r_1$ and
+$r_2$, yet, low values of $r_1$ and high values of $r_2$ almost always produce
+the best results, independently of the problem. Nevertheless, there are a few
+cases where an interaction seems to occur and particular combinations perform
+remarkably well.\MANUEL{why?}  In particular, we observe that the pattern on
+the two left-most figures differs from the two right-most ones. While on the
+left the fitness gets worse homogeneously from the top right corner to the
+bottom left one, the pattern on the right is not as homogeneous, yet still
+fitness gets worse with decreasing $r_2$.  We observe that this behavior is
+related to the adequacy of UMM to the problem: the homogeneous behavior of
+parameters $r_1$ and $r_2$ depends on the problem (not to the particular
+instances) and the more homogeneous the behavior, the better performance of
+UMM. We elaborate this argument in the following sections with detail.
 
-Moreover, we can see that the behavior of the  two left figures differs from the two in the right: while the two in the left appear to be increase homogeneously from the top right corner to the bottom left one, the two in the right appear to be closer to randomness.
-We point out that this performance is related to the adequacy of UMM to the problem: homogeneousness is associated to the problem (not to the particular instances) and the more homogeneous the plot, the better performance of UMM.
-We elaborate this argument in the following sections with detail.
+Over all the instances \MANUEL{(which instances we tested? all synthetic?)} we
+see that the configuration $r_1 =0.1$, $r_2=0.9$ has good
+performance\MANUEL{how do you define this? The best mean?}. From this
+preliminary experiment, it is not obvious how the best settings are related to
+problem features, such as permutation length ($n$). On the other hand, given
+the sharp transitions in the plots, it is clear that further fine-tuning may
+improve performance and uncover further patterns.  Although automatic parameter
+tuning may likely provide even more fine-tuned settings, our goal here is to
+understand how UMM performs in comparison to an existing Bayesian method (CEGO)
+and automatically tuning the parameters of CEGO is not computationally
+feasible.  For consistency, in the following sections, we run UMM with this
+parameter configuration. Moreover, we recommend this configuration as the
+current default of UMM.
+
+\subsection{Analysis of UMM and CEGO}
 
@@
 Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
-\begin{figure}[tb]
+\begin{figure}[tbp]
   \centering%
@@
 
-\begin{figure}
+\begin{figure}[tp]
   \centering%
@@
 
-Bayesian optimization methods using a global GP model, such as CEGO, are known
-to have trouble optimizing locally \citep{EriPeaGar2019scalable}. Our
-intuition is that this problem becomes worse in rugged combinatorial
-landscapes, where small steps may produce drastic changes.
+In this paper, we have introduced UMM, a population-based probabilistic
+algorithm based on an unbalanced Mallows model. The algorithm is designed for
+black-box combinatorial problems on permutation landscapes and when the budget
+of fitness evaluations is severely limited (up to $1000$ or less).  UMM is
+specially well-suited for budget-limited scenarios that are still
+time-sensitive, e.g., when the overhead incurred by the optimizer should be no
+more than a few minutes added to each fitness evaluation. As shown by our
+results, a Bayesian optimizer such as CEGO may sometimes require more than hour
+per fitness evaluation. The time required to train the surrogate-model of CEGO
+also sharply increases with permutation length, hence, UMM would be a
+computationally feasible alternative for relatively large problem sizes.
 
-Open questions:
-\begin{itemize}
-\item How difficult is to extend UMM to other distance metrics?
-\item Can we plot the posterior probability of the optimal solution?
-\end{itemize}
+Although we were inspired by previous work on ACO for expensive black-box
+combinatorial problems~\citep{PerLopStu2015si}, UMM is, to the best of our
+knowledge, the first EDA specifically designed for such problems that is able
+to obtain competitive results when compared with Bayesian optimizers. The
+results presented here show that, despite the intrinsic difficulty of this scenario, there are still significant advances yet to be made. Thus, we hope
+that this paper will motivate further research. % on such problems.
 
+A somewhat surprising result was that the behavior of UMM (and the same for
+CEGO) on the PFSP is close to random. This result may be due to the distance
+metric used, but further analysis is needed to corroborate this
+conjecture. Here, we have focused on Kendall's $\tau$ distance, however, it is
+possible to extend UMM to other distance metrics, which will allow us to
+dynamically select among various distance metrics for an unknown black-box
+permutation landscape~\citep{ZaeStoBar2014:ppsn}. Also, as shown by our
+experiments, the behavior of the parameters ($r_1$ and $r_2$) of UMM is not
+always obvious, thus a more detailed analysis would be needed to provide either
+generally good static values or an online adaptation approach. Finally, we have
+used here three very different combinatorial problems (LOP, PFSP and QAP) as
+black-box benchmarks. However, an even more diverse range of problems would be
+needed to understand the behavior of UMM on real-world black-box combinatorial
+landscapes.
+
+% Bayesian optimization methods using a global GP model, such as CEGO, are known
+% to have trouble optimizing locally \citep{EriPeaGar2019scalable}. Our
+% intuition is that this problem becomes worse in rugged combinatorial
+% landscapes, where small steps may produce drastic changes.
+
+% Open questions:
+% \begin{itemize}
+% \item How difficult is to extend UMM to other distance metrics?
+% \item Can we plot the posterior probability of the optimal solution?
+% \end{itemize}
+
+%% MANUEL: Blind, so no acknowledgments yet.
 \begin{smaller}
 \paragraph*{Acknowledgements.}
-M.\@ L\'opez-Ib\'a\~nez is a ``Beatriz Galindo'' Senior Distinguished Researcher (BEAGAL 18/00053) funded by the Ministry of Science and Innovation of the Spanish Government.
+% M.\@ L\'opez-Ib\'a\~nez is a ``Beatriz Galindo'' Senior Distinguished Researcher (BEAGAL 18/00053) funded by the Ministry of Science and Innovation of the Spanish Government.
 %

### 2020-10-27 paper/main.tex
paper

@@
 
+Let us assume a black-box ``fitness'' function that must be minimized over a
+space of permutations $f\colon S_n \to \mathbb{R}$. Being black-box means that
+we can evaluate any candidate permutation $\sigma \in S_n$ to obtain
+$f(\sigma)$, however, we do not know anything else about the form of
+$f$. Moreover, the evaluation of $f$ is expensive in computation time or
+resources and, thus, we can only evaluate a limited budget of candidate
+permutations. Here, we will study budgets lower than 400. Such expensive
+black-box combinatorial problems arise in diverse contexts, for example,
+protein folding~\citep{RomKraArn2012protein} and industrial
+production~\citep{FerAlvDiaIglEna2014ants}, where the fitness function may
+involve expensive simulations for which no closed-form mathematical description
+is available. The black-box formulation and limited budget preclude most of the
+optimization techniques that are successful in combinatorial optimization, such
+as constructive and perturbation heuristics and local search, which makes this
+a particularly challenging scenario, perhaps even more than its continuous
+counter-part.
+
+\subsection{Permutations}
+
 Permutations are defined as bijections of the set $[n]$ of integers onto itself.
@@
 
-Let us assume a black-box ``fitness'' function that must be minimized over a
-space of permutations $f\colon S_n \to \mathbb{R}$. Being black-box means that
-we can evaluate any candidate permutation $\sigma \in S_n$ to obtain
-$f(\sigma)$, however, we do not know anything else about the form of
-$f$. Moreover, the evaluation of $f$ is expensive in computation time or
-resources and, thus, we can only evaluate a limited budget of candidate
-permutations. Here, we will study budgets lower than 400. Such expensive
-black-box combinatorial problems arise in diverse contexts, for example,
-protein folding~\citep{RomKraArn2012protein} and industrial
-production~\citep{FerAlvDiaIglEna2014ants}, where the fitness function may
-involve expensive simulations for which no closed-form mathematical description
-is available. The black-box formulation and limited budget preclude most of the
-optimization techniques that are successful in combinatorial optimization, such
-as constructive and perturbation heuristics and local search, which makes this
-a particularly challenging scenario, perhaps even more than its continuous
-counter-part.
+\subsection{Distributions over permutations}
+Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$.
+One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
+The MM defines the probability of each permutation $\sigma\in S_n$ as follows:
+%
+\begin{equation}\label{eq:MM}
+\Prob(\sigma \mid \sigma_0, \theta )= \frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi} \sim MM(\sigma_0, \theta)
+\end{equation}
+%
+with two parameters, $\theta$ and $\sigma_0$.
+Permutation $\sigma_0\in S_n$ is a reference permutation and has the largest probability value, i.e., the mode of the distribution.
+The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.
+The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance (Eq.~\eqref{eq:kendall_decomp}) as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
+
+We will use the standard maximum likelihood estimation (MLE) approach to fit the parameters of a MM for a given collection of permutations. 
+The MLE process is divided in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.
+
+The exact MLE for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
+Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}.
+An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters for a sample distributed according to the MM~\cite{CarProSha2013votes}; and (iii) and also when no distribution is assumed, Borda is an approximation to Kemeny~\cite{CopFleRur2010ordering}.
+
+The Borda ranking of a sample of permutations $S \subset S_n$ is computed as follows. 
+First, for each position $i \in [n]$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
+Then, assign label 1 to the position with the smallest Borda score, label 2 to the position with the second smallest score and so on, i.e., order the positions by Borda score increasingly.
+
+\subsection{Unbalanced Borda}\label{sec:uborda}
+We introduce now the core for our probabilistic algorithm, the uBorda algorithm.
+Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
+The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R^{+}}$.
+Then, the score of \manuel{item}{position} $i$ is defined as follows:
+%
+\begin{equation}\label{eq:uborda_score}
+B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i) \enspace,
+\end{equation}
+%
+and, the same as Borda, orders the \ekhine{items}{positions}\MANUEL{X why items?
+we talk about positions above.
+We could talk about decision variables if that is clearer\ekhine{}{id rather use the simplest possible notation.
+In statistical literature you ll see items/rankings.
+}} increasingly by their score.
+
+Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$.
+However, in other case uBorda ranking will be closer to the rankings in $S$ with smaller value of $w(\sigma)$.
+Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
+Clearly, the choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
+For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{}.\MANUEL{X where?
+Do we miss a citation here?\ekhine{}{to be processed, arxiv} Maybe you can create a bibentry with the details that you know already.}
+
+In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$.
+Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
+Interestingly, uBorda is computed in polynomial time.
 
@@
 function~\citep[p.~875]{ZaeStoFriFisNauBar2014}.
-
-\subsection{Distributions over permutations}
-Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$.
-One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
-The MM defines the probability of each permutation $\sigma\in S_n$ as follows:
-%
-\begin{equation}\label{eq:MM}
-\Prob(\sigma \mid \sigma_0, \theta )= \frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi} \sim MM(\sigma_0, \theta)
-\end{equation}
-%
-with two parameters, $\theta$ and $\sigma_0$.
-Permutation $\sigma_0\in S_n$ is a reference permutation and has the largest probability value, i.e., the mode of the distribution.
-The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.
-The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance (Eq.~\eqref{eq:kendall_decomp}) as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
-
-We will use the standard maximum likelihood estimation (MLE) approach to fit the parameters of a MM for a given collection of permutations. 
-The MLE process is divided in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.
-
-The exact MLE for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}.
-Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}.
-An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda is an approximation to Kemeny~\cite{CopFleRur2010ordering}.\MANUEL{I'm not sure ``in a general optimization perspective'' means}
-
-The Borda ranking of a sample of permutations $S \subset S_n$ is computed as follows. 
-First, for each position $i \in [n]$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
-Then, assign label 1 to the position with the smallest Borda score, label 2 to the position with the second smallest score and so on, i.e., order the positions by Borda score increasingly.
-
-\subsection{Unbalanced Borda}\label{sec:uborda}
-We introduce now the core for our probabilistic algorithm, the uBorda algorithm.
-Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
-The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R^{+}}$.
-Then, the score of \manuel{item}{position} $i$ is defined as follows:
-%
-\begin{equation}\label{eq:uborda_score}
-B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i) \enspace,
-\end{equation}
-%
-and, the same as Borda, orders the \ekhine{items}{positions}\MANUEL{X why items? we talk about positions above. We could talk about decision variables if that is clearer} increasingly by their score.
-
-Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$.
-However, in other case uBorda ranking will be closer to the rankings in $S$ with smaller value of $w(\sigma)$.
-Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
-Clearly, the choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
-For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{}.\MANUEL{X where?
-Do we miss a citation here?\ekhine{}{to be processed, arxiv} Maybe you can create a bibentry with the details that you know already.}
-
-In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$.
-Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
-Interestingly, uBorda is computed in polynomial time.
 
@@
 The process is repeated up to a given maximum budget of evaluations $\FEmax$.\MANUEL{If the learning step happens later, when does it happen here?}
-\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it. Can we give a bit more detail?}
+\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it. Can we give a bit more detail?\ekhine{no without explaining the expecteattions, too messy}}
 
@@
 
-\manuel{At each iteration of the algorithm the following steps take place: (1) randomly
-generate a permutation distributed according to the internal model and include
-it in the sample and (2) evaluate the new permutation and update the
-probabilistic model. These steps correspond to two statistical operations: (1)
-sample a probability distribution and (2) learn a probability distribution of
-the transformation of the sample.
-Both are describe in the subsequent sections.}{}\MANUEL{I don't think we need this, it is said above (if some detail is missing we should integrate it the paragraph above)}
-
 \newcommand{\Edist}{\ensuremath{\mathbb{E}[D]}}
@@
 %
-\MANUEL{should the $\sigma$ be the same on both sides of $=$?}
+\MANUEL{X should the $\sigma$ be the same on both sides of $=$?}
 where $S'$ contains the $r_2|S|$ elements of $S$ with the best corresponding fitness values in $F$.
@@
 Sampling from $MM(\sigma_0,\theta)$  can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{Collas,IruCalLoz2016permallows,DoiPekReg2004rank}\MANUEL{Which paper is Collas?
-It is not found.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{respectively?}
+It is not found.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{X respectively?
+more than 2}
 %
 At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained in the previous learning step.
-The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM, i.e., $D$ is the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$:\MANUEL{where does this formula come from?}
+The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM, i.e., $D$ is the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$~\cite{gMallows}:\MANUEL{X where does this formula come from?\ekhine{}{}}
 %
@@
 A purely explotation approach would always generate $\sigma_0$ in the sampling stage.
-However, the setting of a decreasing variance \MANUEL{But we are controlling the expectation not the variance???}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
+However, the setting of a decreasing variance \MANUEL{But we are controlling the expectation not the variance???\ekhine{}{both are a function of $\theta$}}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
 Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.
@@
 experiments\MANUEL{\citet{ZaeStoBar2014:ppsn} claims this is called Swap
-  distance, could you double-check?}. Nevertheless,
+  distance, could you double-check?\ekhine{}{es la kendall among the inverses en realidad}}. Nevertheless,
 \citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
@@
 \MANUEL{This may need updating since you ran some experiments? didn't you?}
+\ekhine{}{esto es lo q se: Hipatia (Cluster)
+18 nodes (1 with Nvidia Tesla K40 GPU)
+624 cores
+4TB of RAM Memory
+Infiniband network}
 
@@
 
-\ekhine{}{In prev sections we say that we show here how to tune $rho$. Le ves sitio? con los heatmaps? al principio de la sección? an outline here}
-\ekhine{}{recall that $w(\sigma) = \rho^{f(sigma)}$. We select $\rho$ in such a way that the 100$r_1\%$ of the samples have the 100$w\%$ of the weight}
-
-The first part of the experimental analysis concerns the \manuel{tunning}{analyses} of parameter $\rho$.
+\subsection{Setting the learning rate}
+The first part of the experimental analysis concerns the {analyses} of the learning rate, which corresponds to parameter $\rho$ of uBorda.
 As stated in Section~\ref{sec:sampling}, $\rho$ is automatically set at each iteration in such a way that the $100r_1\%$ of the samples have the $100r_2\%$ of the weight.
-In this regard, we tried different configurations of parameters $r_1$ and $r_2$ in all the instances.\MANUEL{why did we try this values.} Figure~\ref{fig:heatmaps} shows the best solution\MANUEL{best solution found at the end of a run?} (averaged for 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP.
+In this regard, we tried different configurations of parameters $r_1$ and $r_2$ in all the instances.\MANUEL{??
+why did we try this values.} Figure~\ref{fig:heatmaps} shows the best solution\MANUEL{best solution found at the end of a run?} (averaged for 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP.
+This experiment for all the real datasets can be bound in the supplemental material.
 \MANUEL{The caption says synthetic instances}
@@
   \includegraphics[width=0.24\textwidth]{../img/heatmap_kra32_dat}
-    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:heatmaps}}
+    \caption{Mean fitness  for different configuration for the learning rate \label{fig:heatmaps}}
 \end{figure}
@@
 
-Over all the instances \MANUEL{(which instances we tested? all synthetic?)} we
+Over all the instances \MANUEL{X (which instances we tested? all synthetic?)} we
 see that the configuration $r_1 =0.1$, $r_2=0.9$ has good
@@
 mean. Ideally, each new evaluation will monotonically improve in fitness, since each solution evaluated helps to refine the model and leads to a
-better solution being evaluated in the next step.
-However, it is often the case that  \EKHINE{yo diría solo que lo esperable es que baje monotonically, sin entrar a valorar un ejemplo concreto}\MANUEL{Así está mejor?
-Quiero llamar la atención porque normalmente se muestra la evolución de la mejor solución encontrada pero nosotros estamos mostrando el fitness de cada solución evaluada}
-% the fitness of the solutions does not improve monotonically and
+better solution being evaluated in the next step.  However, it is often the case that 
 new solutions
@@
 
+\EKHINE{cambio un poco al argumento}
 When we compare the behaviour of CEGO and UMM on synthetic LOP instances
 (Fig.~\ref{fig:lop_synth}), we observe a clear pattern independently of the
-uniformity of the instances. In particular, CEGO starts with a random sampling
-of X solutions, with are quite poor as expected, followed by building and
+uniformity of the instances.
+In particular, we see a linear convergence to the optimal value for UMM.
+This is due to the linear decrease in the parameter $\theta$ (line~\ref{line:theta} in Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the internal probability model of the algorithm over the $m$ functions evaluations.
+Overall, UMM found the best solution in all the repetitions.
+
+%UMM converges much more slowly than CEGO, only reaching the
+%same fitness values around evaluation 250.
+
+%Yet, UMM is able to keep the rate of
+%improvement and completely overtake CEGO around evaluation 350. As in the case
+%of CEGO, lower uniformity values lead to worse performance, although the
+%difference is much smaller for UMM than it was for CEGO.\MANUEL{If there is anything else we can say here, please go ahead and say it}
+
+On the other hand, \EKHINE{cambio el orden}
+CEGO starts with a random sampling
+of X\ekhine{}{10?} solutions, with are quite poor as expected, followed by building and
 updating the Kriging model based new evaluations, which leads to an impressive
-improvement in less than 50 evaluations. However, progress is more or less
+improvement in less than 50 evaluations. However, progress is %more or less 
 halted at this point and the model seems to have trouble generating better
-solutions. Lower uniformity values (e.g., $\mu=5$) show worsening evaluations
+solutions. Lower uniformity values (e.g., $\phi=.5$) show worsening evaluations
 despite the increase in solutions used to build the model. In summary, the
@@
   question.}
-
-On the other hand, UMM converges much more slowly than CEGO, only reaching the
-same fitness values around evaluation 250. Yet, UMM is able to keep the rate of
-improvement and completely overtake CEGO around evaluation 350. As in the case
-of CEGO, lower uniformity values lead to worse performance, although the
-difference is much smaller for UMM than it was for CEGO.\MANUEL{If there is anything else we can say here, please go ahead and say it}
 
@@
 
-Results are quite different on the 
+%Results are quite different on the 
 %The fast initial convergence of CEGO for synthetic LOP instances does not seem
 %to translate to the
-instances available in LOPLIB~\citep{}. We show in
-Fig.~\ref{fig:loplib} three examples with different permutation size, however,
-the results are consistent for other LOPLIB instances (complete results
-available as supplementary material \supplement). Due to extremely long
-runtimes of CEGO and constraints in our computing system, we set a maximum
-wall-clock time of 3 days for each run of CEGO. On these instances, the
-behaviour of CEGO and UMM is much more similar (at least up to the termination
-point of CEGO). It appears that our synthetic LOP instances have a fitness
-landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
-instances do not present such landscape. Comparing CEGO and UMM results, the
-differences get smaller, in favour of UMM, with larger instance size. Our
-conjecture is that building an accurate Bayesian model and searching for good
-solutions on it becomes harder for larger fitness landscapes.
-\EKHINE{Parece que el paper va del CEGO en vez del UMM :p.
+%instances available in LOPLIB~\citep{}. We show in
+%Fig.~\ref{fig:loplib} three examples with different permutation size, however,
+%the results are consistent for other LOPLIB instances (complete results
+%available as supplementary material \supplement). Due to extremely long
+%runtimes of CEGO and constraints in our computing system, we set a maximum
+%wall-clock time of 3 days for each run of CEGO. On these instances, the
+%behaviour of CEGO and UMM is much more similar (at least up to the termination
+%point of CEGO). It appears that our synthetic LOP instances have a fitness
+%landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
+%instances do not present such landscape. Comparing CEGO and UMM results, the
+%differences get smaller, in favour of UMM, with larger instance size. Our
+%conjecture is that building an accurate Bayesian model and searching for good
+%solutions on it becomes harder for larger fitness landscapes.
+%\EKHINE{Parece que el paper va del CEGO en vez del UMM :p.
 Yo diría que tanto UMM como CEGO encuentran soluciones que mejoran según avanzan, pero con diferencias de comportamiento: (1) las los optimos de UMM son '3 veces mejor', y esto yo creo que es importante, aunque en gran parte se debe a que (2) UMM hace 400 ejecuciones en 3, 6 y 11 horas y CEGO entre 50 y 130 en 5 dias (3) cego empeora segun n aumenta (y n=200 tampoco es enorme)}
-\MANUEL{Eso es porque puedo interpretar los resultados de CEGO mejor que los de UMM. Si crees que hay alguna interpretación o conjetura sobre UMM que podemos mencionar aquí, sería bueno mencionarla.}
-\MANUEL{Prefiero dejar el comentario sobre las mejores soluciones encontradas y el tiempo requerido para cuando hable de la tabla, así se puede ver todo junto}
+%\MANUEL{Eso es porque puedo interpretar los resultados de CEGO mejor que los de UMM. Si crees que hay alguna interpretación o conjetura sobre UMM que podemos mencionar aquí, sería bueno mencionarla.}
+%\MANUEL{Prefiero dejar el comentario sobre las mejores soluciones encontradas y el tiempo requerido para cuando hable de la tabla, así se puede ver todo junto}
+
+In the following paragraphs we analyze the result over real rata.
+Fig.~\ref{fig:loplib} shows the results on the instances available in LOPLIB~\citep{}.
+The distribution of this instances differs from the symthetic LOP instances, and thus so does the landscape.
+However, UMM shows a very similar performance: there is a (close to) linear decrease on the function evaluations of new solutions along iterations.
+This suggests that UMM very good fits the fitness landscape.
+Regarding CEGO, we can take similar conclussions for the first 50-100 iterations.
+However, the processes had to be halted after hitting the wall time, so no conlusion can be taken for further iterations.
+Overall, the improvement from the first (random) to the last solutions is 2-3 times larger in UMM than CEGO.
 
@@
 In terms of best fitness achieved on each run, and taking into account the
-random behaviour observed in some instances and discussed above,
+random behaviour observed in some instances and discussed above,\ekhine{}{falta un cacho aqui}
 
@@
 results, a Bayesian optimizer such as CEGO may sometimes require more than hour
+\EKHINE{me parece mas de CEGO q UMM y parece q UMM solo tiene de ventaja ir rápido.
+Primero destacaría lo bueno de UMM.
+Creo q es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos  y luego que addemás es más rápido}
 per fitness evaluation. The time required to train the surrogate-model of CEGO
@@
 knowledge, the first EDA specifically designed for such problems that is able
-to obtain competitive results when compared with Bayesian optimizers. The
+to obtain competitive\EKHINE{hhmmm lo hacemos mejor o igual} results when compared with Bayesian optimizers. The
 results presented here show that, despite the intrinsic difficulty of this scenario, there are still significant advances yet to be made. Thus, we hope
@@
 possible to extend UMM to other distance metrics, which will allow us to
-dynamically select among various distance metrics for an unknown black-box
+dynamically select among various distance metrics for an unknown black-box \ekhine{}{estoy totalmente en contra de esto :) }
 permutation landscape~\citep{ZaeStoBar2014:ppsn}. Also, as shown by our

### 2020-10-27 paper/main.tex
paper

@@
 black-box problems over a permutation landscape under a limited budget. There
-exists a wide body of research on probabilistic
+exists a wide body of research on probabilistic models for permutations including  Plackett-Luce and Mallows
 algorithms~\citep{???},\MANUEL{what can we cite from that body of research?}
 with the most prominent examples being estimation of distribution algorithms
-(EDAs)~\citep{???}.\MANUEL{what would be the best work to cite about EDAs for
+(EDAs)~\cite{Ceberio2013}.\MANUEL{X what would be the best work to cite about EDAs for
   combinatorial optimization?} ACO may also be considered a type of EDA, since

### 2020-10-27 paper/main.tex
fix title

@@
 \title{Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems}
+\titlerunning{Unbalanced Mallows Models for Expensive Black-Box Permutation Problems}
 % Double-blind
@@
 knowledge, the first EDA specifically designed for such problems that is able
-to obtain competitive\EKHINE{hhmmm lo hacemos mejor o igual} results when compared with Bayesian optimizers. The
+to match (and sometimes even improve) the results of a Bayesian optimizer\EKHINE{hhmmm lo hacemos mejor o igual}. The
 results presented here show that, despite the intrinsic difficulty of this scenario, there are still significant advances yet to be made. Thus, we hope

### 2020-10-27 paper/main.tex
fix intro

@@
 exists a wide body of research on probabilistic models for permutations including  Plackett-Luce and Mallows
-algorithms~\citep{???},\MANUEL{what can we cite from that body of research?}
+algorithms~\citep{???},\MANUEL{what can we cite from that body of research?: Ekhine will fix it.}
 with the most prominent examples being estimation of distribution algorithms
-(EDAs)~\cite{Ceberio2013}.\MANUEL{X what would be the best work to cite about EDAs for
-  combinatorial optimization?} ACO may also be considered a type of EDA, since
+(EDAs)~\cite{Ceberio2013}.  ACO may also be considered a type of EDA, since
 it builds a probability distribution model from which solutions are sampled

### 2020-10-27 paper/main.tex
fix intro

@@
 exists a wide body of research on probabilistic models for permutations including  Plackett-Luce and Mallows
-algorithms~\citep{???},\MANUEL{what can we cite from that body of research?: Ekhine will fix it.}
+algorithms~\citep{???},\MANUEL{what can we cite from that body of research?: Ekhine will fix this.}
 with the most prominent examples being estimation of distribution algorithms
@@
   \begin{itemize}
-  \item Who else has used Mallows model for black-box optimization?
-  \item Who else has used unbalanced borda for optimization?
+  \item Who else has used Mallows model for black-box optimization? Ekhine: Va a mirarlo.
+  \item Who else has used unbalanced borda for optimization? Ekhine va a añadir 2 frases.
   \end{itemize}

### 2020-10-27 paper/main.tex
fix intro
fix background

@@
 }
+This paper structured as follows \MANUEL{Mío}
 
@@
 
-\subsection{Black-box Combinatorial Optimization under a limited budget}
-
-Let us assume a black-box ``fitness'' function that must be minimized over a
-space of permutations $f\colon S_n \to \mathbb{R}$. Being black-box means that
-we can evaluate any candidate permutation $\sigma \in S_n$ to obtain
-$f(\sigma)$, however, we do not know anything else about the form of
-$f$. Moreover, the evaluation of $f$ is expensive in computation time or
-resources and, thus, we can only evaluate a limited budget of candidate
-permutations. Here, we will study budgets lower than 400. Such expensive
-black-box combinatorial problems arise in diverse contexts, for example,
-protein folding~\citep{RomKraArn2012protein} and industrial
-production~\citep{FerAlvDiaIglEna2014ants}, where the fitness function may
-involve expensive simulations for which no closed-form mathematical description
-is available. The black-box formulation and limited budget preclude most of the
-optimization techniques that are successful in combinatorial optimization, such
-as constructive and perturbation heuristics and local search, which makes this
-a particularly challenging scenario, perhaps even more than its continuous
-counter-part.
-
 \subsection{Permutations}
@@
 The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R^{+}}$.
-Then, the score of \manuel{item}{position} $i$ is defined as follows:
+Then, the score of item $i \in [n]$ is defined as follows:
 %
@@
 %
-and, the same as Borda, orders the \ekhine{items}{positions}\MANUEL{X why items?
-we talk about positions above.
-We could talk about decision variables if that is clearer\ekhine{}{id rather use the simplest possible notation.
-In statistical literature you ll see items/rankings.
-}} increasingly by their score.
+and, the same as Borda, orders the items $i \in [n]$ increasingly by their score.
 
@@
 Do we miss a citation here?\ekhine{}{to be processed, arxiv} Maybe you can create a bibentry with the details that you know already.}
-
-In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$.
+%
+% In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$.
 Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
 Interestingly, uBorda is computed in polynomial time.
+
+\subsection{Black-box Combinatorial Optimization under a limited budget}
+
+Let us assume a black-box ``fitness'' function that must be minimized over a
+space of permutations $f\colon S_n \to \mathbb{R}$. Being black-box means that
+we can evaluate any candidate permutation $\sigma \in S_n$ to obtain
+$f(\sigma)$, however, we do not know anything else about the form of
+$f$. Moreover, the evaluation of $f$ is expensive in computation time or
+resources and, thus, we can only evaluate a limited budget of candidate
+permutations. Here, we will study budgets lower than 400. Such expensive
+black-box combinatorial problems arise in diverse contexts, for example,
+protein folding~\citep{RomKraArn2012protein} and industrial
+production~\citep{FerAlvDiaIglEna2014ants}, where the fitness function may
+involve expensive simulations for which no closed-form mathematical description
+is available. The black-box formulation and limited budget preclude most of the
+optimization techniques that are successful in combinatorial optimization, such
+as constructive and perturbation heuristics and local search, which makes this
+a particularly challenging scenario, perhaps even more than its continuous
+counter-part.
 

### 2020-10-27 paper/main.tex
fix background

@@
 The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$.
-We represent permutations as an ordered vector and, following the standard convention, we say that item $i$ is ranked at position $j$ whenever $\sigma(i)=j$.
+We represent permutations as an ordered vector and, we say that position $i$ contains value $j$ whenever $\sigma(i)=j$.
 We denote permutations with Greek letters with exception of the identity permutation denoted as $e=(1, 2, 3, \dotsc,n)$.
@@
 The Borda ranking of a sample of permutations $S \subset S_n$ is computed as follows. 
-First, for each position $i \in [n]$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
-Then, assign label 1 to the position with the smallest Borda score, label 2 to the position with the second smallest score and so on, i.e., order the positions by Borda score increasingly.
+First, for each item $i \in [n]$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
+Then, assign label 1 to the item with the smallest Borda score, label 2 to the item with the second smallest score and so on, i.e., order the positions by Borda score increasingly.
 

### 2020-10-27 paper/main.tex
fix background

@@
 Permutations are defined as bijections of the set $[n]$ of integers onto itself.
-The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$.
+The set of all permutations of $n$ values is denoted as $S_n$ and has cardinality $n!$.
 We represent permutations as an ordered vector and, we say that position $i$ contains value $j$ whenever $\sigma(i)=j$.
@@
 Clearly, the choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
-For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{}.\MANUEL{X where?
+For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{???}.\MANUEL{X where?
 Do we miss a citation here?\ekhine{}{to be processed, arxiv} Maybe you can create a bibentry with the details that you know already.}
@@
     \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
-    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}
-    \State Decrease  $\theta$\label{line:theta}\Comment{Eq.~\ref{eq:expectation:}}
+    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}\Comment{Learning Step}
+    \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}\Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
     \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
@@
 As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are uniformly randomly generated (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated.
-Then, using the current sample $S$, the uBorda algorithm is applied to find a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}).
+Then, using the current sample $S$, the uBorda algorithm is applied to learn a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}).
 Together with the parameter $\theta$, they define the Mallows model $MM(\hat\sigma_0, \theta)$. 
@@
 This candidate permutation (solution) is evaluated and added to the sample $S$.
-The process is repeated up to a given maximum budget of evaluations $\FEmax$.\MANUEL{If the learning step happens later, when does it happen here?}
-\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it. Can we give a bit more detail?\ekhine{no without explaining the expecteattions, too messy}}
+The process is repeated up to a given maximum budget of evaluations $\FEmax$.
 

### 2020-10-27 paper/main.tex
fix background

@@
 minimal fitness values. This is achieved by means of the uBorda ranking
-(Section~\ref{sec:uBorda}), which can be seen as replicating each permutation
+(Section~\ref{sec:uborda}}), which can be seen as replicating each permutation
 $\sigma$ in the sample a number of times inversely proportional to its fitness
 $f(\sigma)$, i.e., it behaves as if there were many copies of the best
-solutions and few of the bad ones. A \manuel{crucial}{} property of UMM is that the
-computational complexity of working with the original sample and with the
-transformed sample is the same, i.e., polynomial. The most important steps are
+solutions and few of the bad ones. A property of UMM is that the
+computational complexity of working with the original sample (Borda) and with the
+transformed sample (uBorda) is the same, i.e., polynomial. The most important steps are
 the sampling of a new permutation from $MM(\sigma_0,\theta)$ (line~\ref{line:sample} in Algorithm~\ref{alg:umm}) and the update

### 2020-10-27 paper/main.tex
fix sec 3

@@
 minimal fitness values. This is achieved by means of the uBorda ranking
-(Section~\ref{sec:uborda}}), which can be seen as replicating each permutation
+(Section~\ref{sec:uborda}), which can be seen as replicating each permutation
 $\sigma$ in the sample a number of times inversely proportional to its fitness

### 2020-10-27 paper/main.tex
fix sec 3

@@
 the weights is concentrated in the best $100r_2$\% of the solutions in $S$, i.e.,
-\MANUEL{Let's talk about this Equation tomorrow}
 %
 \begin{equation}
-\begin{split}
-r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma\in S'}\rho^{f(\sigma)} \quad\text{where}\quad |S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\geq f(\sigma)
-\end{split}
+r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma'\in S'}\rho^{f(\sigma')} \quad\text{where}\quad |S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\leq f(\sigma)
 \end{equation}
 %
-\MANUEL{X should the $\sigma$ be the same on both sides of $=$?}
 where $S'$ contains the $r_2|S|$ elements of $S$ with the best corresponding fitness values in $F$.

### 2020-10-27 paper/main.tex
fix sec 3

@@
 \subsubsection{Sampling step.}\label{sec:sampling}
-Sampling from $MM(\sigma_0,\theta)$  can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{Collas,IruCalLoz2016permallows,DoiPekReg2004rank}\MANUEL{Which paper is Collas?
-It is not found.} where the computational complexities are $O(n\log n)$ and $O(n^2)$.\MANUEL{X respectively?
+Sampling from $MM(\sigma_0,\theta)$  can be done efficiently given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{IruCalLoz2016permallows}, where the computational complexity is $O(n\log n)$.
+% and $O(n^2)$.\MANUEL{X respectively?
 more than 2}

### 2020-10-27 paper/main.tex
fix sec 3

@@
 At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained in the previous learning step.
-The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM, i.e., $D$ is the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$~\cite{gMallows}:\MANUEL{X where does this formula come from?\ekhine{}{}}
+The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM, where $D$ is the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$.
+Both variance and expected distance increase monotonically with the value of $\rho$.
+In particular, $\Edist$ is given by~\cite{gMallows}:
 %

### 2020-10-27 paper/main.tex
fix sec 3

@@
 The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM, where $D$ is the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$.
-Both variance and expected distance increase monotonically with the value of $\rho$.
+Both variance and expected distance increase monotonically with the value of $\theta$.
 In particular, $\Edist$ is given by~\cite{gMallows}:
@@
 A purely explotation approach would always generate $\sigma_0$ in the sampling stage.
-However, the setting of a decreasing variance \MANUEL{But we are controlling the expectation not the variance???\ekhine{}{both are a function of $\theta$}}  controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
+However, the setting of a decreasing variance   controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
 Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.

### 2020-10-27 paper/main.tex
fix sec 3

@@
 At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained in the previous learning step.
-The scale $\theta$ controls the variance and expected distance $\Edist$ of the MM, where $D$ is the distance of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$.
-Both variance and expected distance increase monotonically with the value of $\theta$.
-In particular, $\Edist$ is given by~\cite{gMallows}:
+The scale $\theta$ controls the variance and expected value and variance of the distance $D$ of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$.
+Both  the expected value of $D$ and its variance increase monotonically with the value of $\theta$.
+In particular, the expected distance $\Edist$ is given by~\cite{gMallows}:
 %

### 2020-10-27 paper/main.tex
fix sec 3

@@
 At each iteration of UMM, $\sigma_0=\hat\sigma_0$ is the permutation obtained in the previous learning step.
-The scale $\theta$ controls the variance and expected value and variance of the distance $D$ of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$.
+The scale $\theta$ controls the expected value and variance of the distance $D$ of a random permutation $\sigma\sim MM(\sigma_0, \theta)$ to the location parameter $\sigma_0$.
 Both  the expected value of $D$ and its variance increase monotonically with the value of $\theta$.

### 2020-10-27 paper/main.tex
fix sec 4

@@
 
-\subsubsection{Other}.  For simplicity, we use Kendall's $\tau$
+\subsubsection{Other settings}.  For simplicity, we use Kendall's $\tau$
 distance~(Eq.~\ref{eq:kendall_decomp}) in all
-experiments\MANUEL{\citet{ZaeStoBar2014:ppsn} claims this is called Swap
-  distance, could you double-check?\ekhine{}{es la kendall among the inverses en realidad}}. Nevertheless,
+experiments. Nevertheless,
 \citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
-when using Hamming distance.  We plan to extend UMM to other distance measures
-and to the dynamic selection of distance measures~\citep{ZaeStoBar2014:ppsn}.
+when using Hamming distance.  We plan to extend UMM to other distance measures.
 In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of

### 2020-10-27 paper/main.tex
fix sec 5

@@
 \citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
-when using Hamming distance.  We plan to extend UMM to other distance measures.
+when using Hamming distance.  We plan to extend UMM to other distance measures in the future.
 In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
@@
 
-The \textbf{LOP} \ldots
+The \textbf{LOP} \ldots\EKHINE{Mío}
 
@@
 
+For the LOP, we generated synthetic instances of size $n=20$, $m=200$, $\phi\in\{0.5,0.7,0.9\}$\MANUEL{what is $m$ here? We already use $\FEmax$ for the budget, but we could change it. }
 Describir como se generan las intancias de LOP: Ekhine
 
-Describir las instancias: Ekhine
-
-For the LOP, we generated synthetic instances of size $n=20$, $m=200$, $\phi\in\{0.5,0.7,0.9\}$\MANUEL{what is $m$ here? We already use $\FEmax$ for the budget, but we could change it. }
+Describir las LOPLIB instancias: Ekhine
 
@@
 
-\subsection{Setting the learning rate}
 The first part of the experimental analysis concerns the {analyses} of the learning rate, which corresponds to parameter $\rho$ of uBorda.
@@
   \centering%
-  \includegraphics[width=0.24\textwidth]{../img/heatmap_N-t1d100_01}
-  \includegraphics[width=0.24\textwidth]{../img/heatmap_N-t1d150_01}
-  \includegraphics[width=0.24\textwidth]{../img/heatmap_rec05}
-  \includegraphics[width=0.24\textwidth]{../img/heatmap_kra32_dat}
+  \begin{minipage}{0.49\linewidth}\centering
+    \texttt{N-t1d100.01}\\[-0.2ex]
+    \includegraphics[width=\textwidth]{../img/heatmap_N-t1d100_01}
+  \end{minipage}
+  \begin{minipage}{0.49\linewidth}\centering
+    \texttt{N-t1d50.01}\\[-0.2ex]
+ \includegraphics[width=\textwidth]{../img/heatmap_N-t1d150_01}
+\end{minipage}\\
+\includegraphics[width=0.49\textwidth]{../img/heatmap_rec05}
+  \includegraphics[width=0.49\textwidth]{../img/heatmap_kra32_dat}
     \caption{Mean fitness  for different configuration for the learning rate \label{fig:heatmaps}}

### 2020-10-27 paper/main.tex
fix sec 5

@@
   \includegraphics[width=0.49\textwidth]{../img/heatmap_kra32_dat}
-    \caption{Mean fitness  for different configuration for the learning rate \label{fig:heatmaps}}
+    \caption{Mean fitness  for different configuration for the learning rate (darker is better). \label{fig:heatmaps}}
 \end{figure}
@@
 cases where an interaction seems to occur and particular combinations perform
-remarkably well.\MANUEL{why?}  In particular, we observe that the pattern on
+remarkably well.\MANUEL{Missing data, whitespaces}  In particular, we observe that the pattern on
 the two left-most figures differs from the two right-most ones. While on the
@@
 
-Over all the instances \MANUEL{X (which instances we tested? all synthetic?)} we
+Over all the instances  we
 see that the configuration $r_1 =0.1$, $r_2=0.9$ has good
@@
 
-\EKHINE{cambio un poco al argumento}
-When we compare the behaviour of CEGO and UMM on synthetic LOP instances
-(Fig.~\ref{fig:lop_synth}), we observe a clear pattern independently of the
-uniformity of the instances.
-In particular, we see a linear convergence to the optimal value for UMM.
-This is due to the linear decrease in the parameter $\theta$ (line~\ref{line:theta} in Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the internal probability model of the algorithm over the $m$ functions evaluations.
-Overall, UMM found the best solution in all the repetitions.
+When we compare the behaviour of CEGO and
+UMM on synthetic LOP instances (Fig.~\ref{fig:lop_synth}), we observe a clear
+pattern independently of the uniformity of the instances. In particular, we see
+a linear convergence to the optimal value for UMM.  This is due to the linear
+decrease in the parameter $\theta$ (line~\ref{line:theta} in
+Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the
+internal probability model of the algorithm over the $m$ functions
+evaluations. Overall, UMM found  the best solutions in all the repetitions.\MANUEL{in all runs UMM found a better solution? }
 

### 2020-10-27 paper/main.tex
heatmaps
crop heatmaps
fix sec 5.1

@@
 
-For the LOP, we generated synthetic instances of size $n=20$, $m=200$, $\phi\in\{0.5,0.7,0.9\}$\MANUEL{what is $m$ here? We already use $\FEmax$ for the budget, but we could change it. }
+For the LOP, we generated synthetic instances of size $n=20$, $m=200$, $\phi\in\{0.5,0.7,0.9\}$\MANUEL{what is $m$ here? We already use $\FEmax$ for the budget, but we could change it.}
 Describir como se generan las intancias de LOP: Ekhine
@@
 As stated in Section~\ref{sec:sampling}, $\rho$ is automatically set at each iteration in such a way that the $100r_1\%$ of the samples have the $100r_2\%$ of the weight.
-In this regard, we tried different configurations of parameters $r_1$ and $r_2$ in all the instances.\MANUEL{??
-why did we try this values.} Figure~\ref{fig:heatmaps} shows the best solution\MANUEL{best solution found at the end of a run?} (averaged for 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP.
+In this regard, we tried a few different values of parameters $r_1$ and $r_2$ in all the instances.
+Only values of $r_1 < r_2$ make sense and $r_1 = r_2 = 0.5$ would produce equal weights in the learning step.
+Thus, we tried values $r_1 = \{0.1, 0.2, \dotsc, 0.5\}$ and $r_2 = \{0.6,\dotsc, 0.9, 0.99\}$.
+Figure~\ref{fig:heatmaps} shows the best solution found by each independent run (averaged over 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP.
 This experiment for all the real datasets can be bound in the supplemental material.
-\MANUEL{The caption says synthetic instances}
 
@@
   \begin{minipage}{0.49\linewidth}\centering
-    \texttt{N-t1d100.01}\\[-0.2ex]
+    \texttt{N-t1d100.01} (LOP)\\[-0.2ex]
     \includegraphics[width=\textwidth]{../img/heatmap_N-t1d100_01}
@@
   \begin{minipage}{0.49\linewidth}\centering
-    \texttt{N-t1d50.01}\\[-0.2ex]
+    \texttt{N-t1d50.01} (LOP)\\[-0.2ex]
  \includegraphics[width=\textwidth]{../img/heatmap_N-t1d150_01}
-\end{minipage}\\
-\includegraphics[width=0.49\textwidth]{../img/heatmap_rec05}
-  \includegraphics[width=0.49\textwidth]{../img/heatmap_kra32_dat}
+\end{minipage}\\\vskip 1ex
+\begin{minipage}{0.49\linewidth}
+  \centering
+  \texttt{reC05} (PFSP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/heatmap_rec05}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+  \centering
+  \texttt{kra32} (QAP)\\[-0.5ex]
+\includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
+\end{minipage}
 \caption{Mean fitness  for different configuration for the learning rate (darker is better). \label{fig:heatmaps}}

### 2020-10-28 paper/main.tex
paper

@@
 black-box problems over a permutation landscape under a limited budget. There
-exists a wide body of research on probabilistic models for permutations including  Plackett-Luce and Mallows
-algorithms~\citep{???},\MANUEL{what can we cite from that body of research?: Ekhine will fix this.}
+exists a wide body of research on probabilistic models for permutations, such as Mallows~\cite{gMallows} that have already been 
+been used in optimization problems successfully,
 with the most prominent examples being estimation of distribution algorithms
-(EDAs)~\cite{Ceberio2013}.  ACO may also be considered a type of EDA, since
+(EDAs)~\cite{Ceberio2013}, \cite{Ceberio2015}. ACO may also be considered a type of EDA, since
 it builds a probability distribution model from which solutions are sampled
-with a bias towards the best solutions evaluated so far. The proposed UMM
+with a bias towards the best solutions evaluated so far. This works fall into at least one of these assumptions: function evaluations are cheap or the function evaluation is a white-box.
+
+The proposed UMM
 drastically differs from EDAs since EDAs assume that function evaluations are
@@
 the balance between intensification and exploration is dynamically adapted by
-controlling the variance of the probabilistic Mallows model.  \manuel{}{
+controlling the variance of the probabilistic Mallows model.
+
+The proposed UMM is based on uBorda~\cite{irurozki2020}, a generalization of the well-known rank aggregation algorithm Borda~\cite{Borda1781}.
+Rank aggregation algorithms summarize in a single permutation the information of a sample of permutations $S$, and they can be seen as functions to compute  the \textit{mean} of sample $S$.
+The uBorda algorithm has been proposed to compute a \textit{weighted mean} of sample $S$.
+UBorda has been show to return the correct estimator with high probability on streaming contexts; in this paper, we show its quality on black-box optimization problems.
+\ekhine{}{revisa esto por favor, Manuel}
+
+\manuel{}{
   \begin{itemize}
@@
 Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
-The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{X This paper is missing \ekhine{}{to be pocessed by arxiv}} In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R^{+}}$.
+The uBorda algorithm is a recent generalization of Borda~\cite{irurozki2020}.
+In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R^{+}}$.
 Then, the score of item $i \in [n]$ is defined as follows:
@@
 Clearly, the choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
-For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{???}.\MANUEL{X where?
-Do we miss a citation here?\ekhine{}{to be processed, arxiv} Maybe you can create a bibentry with the details that you know already.}
+For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{irurozki2020}.
 %
@@
 
-The \textbf{LOP} \ldots\EKHINE{Mío}
+The \textbf{LOP} is defined as 
+\begin{equation}\label{eq:lop}
+  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^{n-1} \sum_{j=i+1}^n  a_{\sigma(i), \sigma(j)}
+\end{equation}
+where $[a_{i,j}]$ is a matrix of size $n \times n$.
 
@@
 
-For the LOP, we generated synthetic instances of size $n=20$, $m=200$, $\phi\in\{0.5,0.7,0.9\}$\MANUEL{what is $m$ here? We already use $\FEmax$ for the budget, but we could change it.}
-Describir como se generan las intancias de LOP: Ekhine
+\paragraph{Datasets}
 
-Describir las LOPLIB instancias: Ekhine
+For the LOP, we generated synthetic instances as follows: sample 200 permutations from a MM of size $n=20$ and $\theta=\{-\log(.5), -\log(.7), -\log(.9)\}$, i.e., from peaked distributions to almost uniform distributions.
+Then, generate matrix $a$ as $a_{i,j} = \sum_{\sigma\in S}\sum_{i<j}\mathbb{I}(\sigma(i)<\sigma(j))$.
+
+Moreover, we consider real LOP datasets from the LOPLIB repository. \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
+In particular, we consider the following types: IO, Random instances of type A and Random instances of type B.
 
@@
 see that the configuration $r_1 =0.1$, $r_2=0.9$ has good
-performance\MANUEL{how do you define this? The best mean?}. From this
+performance: \ekhine{}{the rank of this parameter configuration among all the possible configurations is less than 4 on average for all the instances}. From this
 preliminary experiment, it is not obvious how the best settings are related to
@@
 
-To better understand the behaviour of the algorithms, we record the fitness of
+To better understand the behavior of the algorithms, we record the fitness of
 the solution evaluated at each step of each run and we average those values
@@
 
-When we compare the behaviour of CEGO and
+When we compare the behavior of CEGO and
 UMM on synthetic LOP instances (Fig.~\ref{fig:lop_synth}), we observe a clear
@@
 internal probability model of the algorithm over the $m$ functions
-evaluations. Overall, UMM found  the best solutions in all the repetitions.\MANUEL{in all runs UMM found a better solution? }
+evaluations. Overall, \ekhine{UMM found  the best solutions in all the repetitions. }{the mean best solutions of UMM are better than the mean of CEGO regardless the uniformity of the instances. }
 
@@
 
-On the other hand, \EKHINE{cambio el orden}
+On the other hand, % \EKHINE{cambio el orden}
 CEGO starts with a random sampling
-of X\ekhine{}{10?} solutions, with are quite poor as expected, followed by building and
+of 10 solutions, with are quite poor as expected, followed by building and
 updating the Kriging model based new evaluations, which leads to an impressive
@@
 In the following paragraphs we analyze the result over real rata.
-Fig.~\ref{fig:loplib} shows the results on the instances available in LOPLIB~\citep{}.
-The distribution of this instances differs from the symthetic LOP instances, and thus so does the landscape.
+Fig.~\ref{fig:loplib} shows the results on the instances available in LOPLIB.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} The distribution of this instances differs from the symthetic LOP instances, and thus so does the landscape.
 However, UMM shows a very similar performance: there is a (close to) linear decrease on the function evaluations of new solutions along iterations.
@@
 black-box combinatorial problems on permutation landscapes and when the budget
-of fitness evaluations is severely limited (up to $1000$ or less).  UMM is
-specially well-suited for budget-limited scenarios that are still
+of fitness evaluations is severely limited (up to $1000$ or less).  UMM 
+\ekhine{}{has shown to perform statistically better or equal than CEGO, the state of the art algorithm, in the same instances as in the original paper of CEGO and some new instances. UMM}
+is specially well-suited for budget-limited scenarios that are still
 time-sensitive, e.g., when the overhead incurred by the optimizer should be no

### 2020-10-29 paper/main.tex
paper

@@
 still a relatively large budget for CEGO, whereas ACO was not designed for such
-short budgets.
+short budgets. Finally, we note that learning Budget-constrained Combinatorial is a hot topic for a variety of structures besides permutation~\cite{manchanda2020}.
 
@@
 
-The \textbf{LOP} is defined as 
+The \emph{Linear Ordering Problem (LOP)} is defined as 
 \begin{equation}\label{eq:lop}
@@
 
-For the LOP, we generated synthetic instances as follows: sample 200 permutations from a MM of size $n=20$ and $\theta=\{-\log(.5), -\log(.7), -\log(.9)\}$, i.e., from peaked distributions to almost uniform distributions.
+For the LOP, we generated synthetic instances as follows: sample 200 permutations from a MM of size $n=20$ and $\theta=-\log(uniformity)$, where  $uniformity=\{.5,.7,.9)\}$, i.e., from peaked distributions to almost uniform distributions.
 Then, generate matrix $a$ as $a_{i,j} = \sum_{\sigma\in S}\sum_{i<j}\mathbb{I}(\sigma(i)<\sigma(j))$.
@@
 \subsubsection{Computing environment.}
-All experiments were run on Intel Xeon ``Ivybridge'' E5-2650v2 CPUs at
-2.60\,GHz, 64\,GB RAM running CentOS Linux release 7.4.1708 (Core).
-\MANUEL{This may need updating since you ran some experiments? didn't you?}
-\ekhine{}{esto es lo q se: Hipatia (Cluster)
-18 nodes (1 with Nvidia Tesla K40 GPU)
-624 cores
-4TB of RAM Memory
-Infiniband network}
+All experiments were run on 
+Intel Xeon ``Ivybridge'' E5-2650v2 CPUs at 2.60\,GHz, 64\,GB RAM running CentOS Linux release 7.4.1708 (Core),
+\ekhine{}{
+Intel Xeon ``GenuineIntel'' E5-2680v3 CPUs at 2.50\,GHz, 32???\,GB RAM running Linux,
+Intel Xeon ``GenuineIntel'' E5-2683v4 CPUs at 2.10\,GHz, 32???\,GB RAM running Linux.
+}
+%processor : 23
+%vendor_id : GenuineIntel
+%cpu family : 6
+%model : 63
+%model name : Intel(R) Xeon(R) CPU E5-2680 v3 @ 2.50GHz
+%stepping : 2
+%microcode : 0x43
+%cpu MHz : 1199.951
+%cache size : 30720 KB
+%physical id : 1
+%siblings : 12
+%core id : 13
+%cpu cores : 12
+%apicid : 58
+%initial apicid : 58
+%fpu : yes
+%fpu_exception : yes
+%cpuid level : 15
+%wp : yes
+%flags : fpu vme de pse tsc msr pae mce cx8 apic sep mtrr pge mca cmov pat pse36 clflush dts acpi mmx fxsr sse sse2 ss ht tm pbe syscall nx pdpe1gb rdtscp lm constant_tsc arch_perfmon pebs bts rep_good nopl xtopology nonstop_tsc aperfmperf eagerfpu pni pclmulqdq dtes64 monitor ds_cpl vmx smx est tm2 ssse3 sdbg fma cx16 xtpr pdcm pcid dca sse4_1 sse4_2 x2apic movbe popcnt tsc_deadline_timer aes xsave avx f16c rdrand lahf_lm abm epb intel_ppin ssbd ibrs ibpb stibp tpr_shadow vnmi flexpriority ept vpid fsgsbase tsc_adjust bmi1 avx2 smep bmi2 erms invpcid cqm xsaveopt cqm_llc cqm_occup_llc dtherm ida arat pln pts md_clear spec_ctrl intel_stibp flush_l1d
+%bogomips : 5005.34
+%clflush size : 64
+%cache_alignment : 64
+%address sizes : 46 bits physical, 48 bits virtual
+%
+%The node n015 has 32 processors. This is the info of the 32st processor:
+%processor : 31
+%vendor_id : GenuineIntel
+%cpu family : 6
+%model : 79
+%model name : Intel(R) Xeon(R) CPU E5-2683 v4 @ 2.10GHz
+%stepping : 1
+%microcode : 0xb000022
+%cpu MHz : 2442.608
+%cache size : 40960 KB
+%physical id : 1
+%siblings : 16
+%core id : 15
+%cpu cores : 16
+%apicid : 62
+%initial apicid : 62
+%fpu : yes
+%fpu_exception : yes
+%cpuid level : 20
+%wp : yes
+%flags : fpu vme de pse tsc msr pae mce cx8 apic sep mtrr pge mca cmov pat pse36 clflush dts acpi mmx fxsr sse sse2 ss ht tm pbe syscall nx pdpe1gb rdtscp lm constant_tsc arch_perfmon pebs bts rep_good nopl xtopology nonstop_tsc aperfmperf eagerfpu pni pclmulqdq dtes64 monitor ds_cpl vmx smx est tm2 ssse3 sdbg fma cx16 xtpr pdcm pcid dca sse4_1 sse4_2 x2apic movbe popcnt tsc_deadline_timer aes xsave avx f16c rdrand lahf_lm abm 3dnowprefetch epb cat_l3 cdp_l3 intel_ppin intel_pt tpr_shadow vnmi flexpriority ept vpid fsgsbase tsc_adjust bmi1 hle avx2 smep bmi2 erms invpcid rtm cqm rdt_a rdseed adx smap xsaveopt cqm_llc cqm_occup_llc cqm_mbm_total cqm_mbm_local dtherm ida arat pln pts
+%bogomips : 4207.67
+%clflush size : 64
+%cache_alignment : 64
+%address sizes : 46 bits physical, 48 bits virtual
 

### 2020-10-30 paper/main.tex
comments

@@
 
-Moreover, we consider real LOP datasets from the LOPLIB repository. \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
+Moreover, we consider real LOP datasets from the LOPLIB repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
 In particular, we consider the following types: IO, Random instances of type A and Random instances of type B.
@@
 cases where an interaction seems to occur and particular combinations perform
-remarkably well.\MANUEL{Missing data, whitespaces}  In particular, we observe that the pattern on
+remarkably well.\MANUEL{Missing data, whitespaces where more than 6 hours of wall-clock time}  In particular, we observe that the pattern on
 the two left-most figures differs from the two right-most ones. While on the
@@
 instances) and the more homogeneous the behavior, the better performance of
-UMM. We elaborate this argument in the following sections with detail.
+UMM. We elaborate this argument in the following sections with detail.\MANUEL{top and bottom}
 
@@
 produce accurate predictions or the underlying optimizer not being able to find
-improved solutions using the model.\MANUEL{perhaps we should measure the
-  prediction error and see if it grows or decreases. That would answer this
-  question.}
+improved solutions using the model.
+%\MANUEL{perhaps we should measure the
+%  prediction error and see if it grows or decreases. That would answer this
+%  question.}
 
@@
 
-In the following paragraphs we analyze the result over real rata.
-Fig.~\ref{fig:loplib} shows the results on the instances available in LOPLIB.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} The distribution of this instances differs from the symthetic LOP instances, and thus so does the landscape.
+  In the following paragraphs we analyze the result over real rata. Fig.~\ref{fig:loplib} shows the results on the instances available in LOPLIB.
+  %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
+  % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
+  The distribution of this instances differs from the symthetic LOP instances, and thus so does the landscape.
 However, UMM shows a very similar performance: there is a (close to) linear decrease on the function evaluations of new solutions along iterations.
@@
 the models (both Bayesian and probabilistic) are not learning anything about
-the fitness landscape. Results for larger PFSP corroborate this conclusions
+the fitness landscape. Results for larger PFSP corroborate these conclusions
 (\supplement). Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
@@
 In terms of best fitness achieved on each run, and taking into account the
-random behaviour observed in some instances and discussed above,\ekhine{}{falta un cacho aqui}
+random behaviour observed in some instances and discussed above,\ekhine{}{falta un cacho aqui}\MANUEL{Mío}
 

### 2020-10-30 paper/main.tex
add structure

@@
 
-The proposed UMM
-drastically differs from EDAs since EDAs assume that function evaluations are
-cheap and they can do many of them. On the contrary, UMM aims at learning as
-much as possible with a limited number of function evaluations. Moreover,
-unlike EDAs, UMM does not rely on a fixed-size population of solutions that
-evolves over ``generations''. Instead, taking inspiration from Bayesian
-optimizers, UMM considers a sample of permutations during the whole execution
-that is increased by one new permutation at each iteration. Finally,
-the balance between intensification and exploration is dynamically adapted by
-controlling the variance of the probabilistic Mallows model.
+The proposed UMM drastically differs from EDAs since EDAs assume that function
+evaluations are cheap and they can do many of them. On the contrary, UMM aims
+at learning as much as possible with a limited number of function
+evaluations. Moreover, unlike EDAs, UMM does not rely on a fixed-size
+population of solutions that evolves over ``generations''. Instead, taking
+inspiration from Bayesian optimizers, UMM considers a sample of permutations
+during the whole execution that is increased by one new permutation at each
+iteration. Finally, the balance between intensification and exploration is
+dynamically adapted by controlling the variance of the probabilistic Mallows
+model.
 
-The proposed UMM is based on uBorda~\cite{irurozki2020}, a generalization of the well-known rank aggregation algorithm Borda~\cite{Borda1781}.
-Rank aggregation algorithms summarize in a single permutation the information of a sample of permutations $S$, and they can be seen as functions to compute  the \textit{mean} of sample $S$.
-The uBorda algorithm has been proposed to compute a \textit{weighted mean} of sample $S$.
-UBorda has been show to return the correct estimator with high probability on streaming contexts; in this paper, we show its quality on black-box optimization problems.
-\ekhine{}{revisa esto por favor, Manuel}
+The proposed UMM is based on uBorda~\cite{irurozki2020}, a generalization of
+the well-known rank aggregation algorithm Borda~\cite{Borda1781}. Rank
+aggregation algorithms summarize in a single permutation the information of a
+sample of permutations $S$, and they can be seen as functions to compute the
+\textit{mean} of sample $S$. The uBorda algorithm has been proposed to compute
+a \textit{weighted mean} of sample $S$. UBorda has been show to return the
+correct estimator with high probability on streaming contexts; in this paper,
+we show its quality on black-box optimization problems. \ekhine{}{revisa esto
+  por favor, Manuel}
 
@@
 }
-This paper structured as follows \MANUEL{Mío}
+This paper structured as follows.
+Section~\ref{sec:backgroud} summarizes fundamental concepts about permutation spaces, the uBorda algorithm, black-box combinatorial optimization under a limited budget and CEGO, a Bayesian optimizer for such problems.
+Section~\ref{sec:umm} presents the UMM algorithm, which is the main contribution of this paper.
+UMMA is analyzed using the experimental setup presented in Section~\ref{sec:setup} and results are presented in Section~\ref{sec:analysis}.
+We end the paper with overall conclusions and a call for further research on this type of optimization scenario (Section~\ref{sec:conclusions}).
 
@@
 
-\section{Unbalanced Mallows model (UMM)}
+\section{Unbalanced Mallows model (UMM)}\label{sec:umm}
 
@@
 
-\section{Experimental setup}
+\section{Experimental setup}\label{sec:setup}
 
@@
 
-\section{Experimental analysis}
+\section{Experimental analysis}\label{sec:analysis}
 
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusions}
 

### 2020-10-30 paper/main.tex
minor

@@
 during the whole execution that is increased by one new permutation at each
-iteration. Finally, the balance between intensification and exploration is
-dynamically adapted by controlling the variance of the probabilistic Mallows
-model.
+iteration.
+%
+Finally, the balance between intensification and exploration is dynamically
+adapted by controlling the variance of the probabilistic Mallows model.
 
@@
 }
+
 This paper structured as follows.
@@
 Section~\ref{sec:umm} presents the UMM algorithm, which is the main contribution of this paper.
-UMMA is analyzed using the experimental setup presented in Section~\ref{sec:setup} and results are presented in Section~\ref{sec:analysis}.
+UMM is analyzed using the experimental setup presented in Section~\ref{sec:setup} and results are presented in Section~\ref{sec:analysis}.
 We end the paper with overall conclusions and a call for further research on this type of optimization scenario (Section~\ref{sec:conclusions}).
@@
 knowledge, the first EDA specifically designed for such problems that is able
-to match (and sometimes even improve) the results of a Bayesian optimizer\EKHINE{hhmmm lo hacemos mejor o igual}. The
-results presented here show that, despite the intrinsic difficulty of this scenario, there are still significant advances yet to be made. Thus, we hope
-that this paper will motivate further research. % on such problems.
+to match (and sometimes even improve) the results of a Bayesian
+optimizer\EKHINE{hhmmm lo hacemos mejor o igual}. The results presented here
+show that, despite the intrinsic difficulty of this scenario, there are still
+significant advances yet to be made. Thus, we hope that this paper will
+motivate further research. % on such problems.
 
@@
 possible to extend UMM to other distance metrics, which will allow us to
-dynamically select among various distance metrics for an unknown black-box \ekhine{}{estoy totalmente en contra de esto :) }
+dynamically select among various distance metrics for an unknown black-box
+% \ekhine{}{estoy totalmente en contra de esto :) }
 permutation landscape~\citep{ZaeStoBar2014:ppsn}. Also, as shown by our

### 2020-10-31 paper/main.tex
more edits

@@
 optimization~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
-success, there have been attempts at adapting such Bayesian optimization
-algorithms to the combinatorial case, a notable example being Combinatorial
+success, there have been attempts at adapting such surrogate-based optimization
+algorithms to the combinatorial case~\cite{MorKat2011:evo}, a notable example being Combinatorial
 Efficient Global Optimization
@@
 still a relatively large budget for CEGO, whereas ACO was not designed for such
-short budgets. Finally, we note that learning Budget-constrained Combinatorial is a hot topic for a variety of structures besides permutation~\cite{manchanda2020}.
+short budgets.
+%Finally, we note that learning Budget-constrained Combinatorial is a hot topic for a variety of structures besides permutation~\cite{manchanda2020}.
 
@@
 black-box problems over a permutation landscape under a limited budget. There
-exists a wide body of research on probabilistic models for permutations, such as Mallows~\cite{gMallows} that have already been 
-been used in optimization problems successfully,
-with the most prominent examples being estimation of distribution algorithms
-(EDAs)~\cite{Ceberio2013}, \cite{Ceberio2015}. ACO may also be considered a type of EDA, since
-it builds a probability distribution model from which solutions are sampled
-with a bias towards the best solutions evaluated so far. This works fall into at least one of these assumptions: function evaluations are cheap or the function evaluation is a white-box.
+exists a wide body of research on probabilistic models for permutations, such
+as Mallows~\cite{FliVer1986}, that have already been been used in optimization
+problems successfully, with the most prominent examples being estimation of
+distribution algorithms
+(EDAs)~\cite{CebIruMen2014eda,CebMenLoz2015mallows}. ACO may also be considered
+a type of EDA, since it builds a probability distribution model from which
+solutions are sampled with a bias towards the best solutions evaluated so
+far. These classical EDAs  assume that function
+evaluations are cheap and/or the fitness function is white-box.\MANUEL{Do EDAs really assume that the fitness function is white-box? ACO doesn't.}
+%The proposed UMM drastically differs from the above algorithms.
 
-The proposed UMM drastically differs from EDAs since EDAs assume that function
-evaluations are cheap and they can do many of them. On the contrary, UMM aims
-at learning as much as possible with a limited number of function
-evaluations. Moreover, unlike EDAs, UMM does not rely on a fixed-size
+In contrast, UMM aims at learning as much as possible with a limited number of
+function evaluations. Moreover, unlike EDAs, UMM does not rely on a fixed-size
 population of solutions that evolves over ``generations''. Instead, taking
@@
 during the whole execution that is increased by one new permutation at each
-iteration.
+iteration. The sample is used to learn a Mallows model from which a single new
+permutation is sampled. The model is learned by using
+uBorda~\cite{IruLobPer2020arxiv}, a generalization of the well-known rank
+aggregation algorithm Borda~\cite{Borda1781}. Rank aggregation algorithms
+summarize in a single permutation the information of a sample of permutations,
+and they can be seen as functions to compute the \emph{sample mean}. The uBorda
+algorithm computes a \textit{weighted sample mean} and has been shown to return
+the correct estimator with high probability on streaming
+contexts~\cite{IruLobPer2020arxiv}.
 %
@@
 
-The proposed UMM is based on uBorda~\cite{irurozki2020}, a generalization of
-the well-known rank aggregation algorithm Borda~\cite{Borda1781}. Rank
-aggregation algorithms summarize in a single permutation the information of a
-sample of permutations $S$, and they can be seen as functions to compute the
-\textit{mean} of sample $S$. The uBorda algorithm has been proposed to compute
-a \textit{weighted mean} of sample $S$. UBorda has been show to return the
-correct estimator with high probability on streaming contexts; in this paper,
-we show its quality on black-box optimization problems. \ekhine{}{revisa esto
-  por favor, Manuel}
+% The proposed UMM is based on uBorda~\cite{IruLobPer2020arxiv}, a generalization
+% of the well-known rank aggregation algorithm Borda~\cite{Borda1781}. Rank
+% aggregation algorithms summarize in a single permutation the information of a
+% sample of permutations, and they can be seen as functions to compute the
+% \emph{sample mean}. The uBorda algorithm has been proposed to compute a
+% \textit{weighted sample mean} and shown to return the correct estimator with
+% high probability on streaming contexts.  in this paper, we show its quality on
+% black-box optimization problems.
+%\ekhine{}{revisa esto por
+%  favor, Manuel}
 
-\manuel{}{
-  \begin{itemize}
-  \item Who else has used Mallows model for black-box optimization? Ekhine: Va a mirarlo.
-  \item Who else has used unbalanced borda for optimization? Ekhine va a añadir 2 frases.
-  \end{itemize}
-}
+% \manuel{}{
+%   \begin{itemize}
+%   \item Who else has used Mallows model for black-box optimization? Ekhine: Va a mirarlo.
+%   \item Who else has used unbalanced borda for optimization? Ekhine va a añadir 2 frases.
+%   \end{itemize}
+% }
 
 This paper structured as follows.
-Section~\ref{sec:backgroud} summarizes fundamental concepts about permutation spaces, the uBorda algorithm, black-box combinatorial optimization under a limited budget and CEGO, a Bayesian optimizer for such problems.
+Section~\ref{sec:backgroud} summarizes fundamental concepts about permutation spaces, uBorda, black-box combinatorial optimization under a limited budget and CEGO, a Bayesian optimizer for such problems.
 Section~\ref{sec:umm} presents the UMM algorithm, which is the main contribution of this paper.
@@
 We represent permutations as an ordered vector and, we say that position $i$ contains value $j$ whenever $\sigma(i)=j$.
-We denote permutations with Greek letters with exception of the identity permutation denoted as $e=(1, 2, 3, \dotsc,n)$.
-We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=e$ always holds.
+We denote permutations with Greek letters.\MANUEL{Do we use $e$ anywhere?
+If not, we do not need this definition ``with exception of the identity permutation denoted as $e=(1, 2, 3, \dotsc,n)$.''} We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.
 
@@
 Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
-The uBorda algorithm is a recent generalization of Borda~\cite{irurozki2020}.
+The uBorda algorithm is a recent generalization of Borda~\cite{IruLobPer2020arxiv}.
 In uBorda, we consider that each permutation in the sample $\sigma\in S$ has an associated weight $w(\sigma)$ where $w\colon S_n \to \mathbb{R^{+}}$.
@@
 Clearly, the choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
-For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{irurozki2020}.
+For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{IruLobPer2020arxiv}.
 %
@@
 $f$. Moreover, the evaluation of $f$ is expensive in computation time or
-resources and, thus, we can only evaluate a limited budget of candidate
+resources and, thus, we can only evaluate a limited budget $\FEmax$ of candidate
 permutations. Here, we will study budgets lower than 400. Such expensive
@@
 optimization techniques that are successful in combinatorial optimization, such
-as constructive and perturbation heuristics and local search, which makes this
-a particularly challenging scenario, perhaps even more than its continuous
+as constructive  heuristics and local search, leading to
+a particularly challenging scenario, perhaps even more challenging than its continuous
 counter-part.
@@
 
-\renewcommand{\subsubsection}{\smallskip{}\noindent\textbf}
+\renewcommand{\subsubsection}[1]{\smallskip{}\noindent\textbf{#1}.}
 
@@
 (Section~\ref{sec:uborda}), which can be seen as replicating each permutation
-$\sigma$ in the sample a number of times inversely proportional to its fitness
+$\sigma \in S$ a number of times inversely proportional to its fitness
 $f(\sigma)$, i.e., it behaves as if there were many copies of the best
@@
 computational complexity of working with the original sample (Borda) and with the
-transformed sample (uBorda) is the same, i.e., polynomial. The most important steps are
-the sampling of a new permutation from $MM(\sigma_0,\theta)$ (line~\ref{line:sample} in Algorithm~\ref{alg:umm}) and the update
-(\emph{learning}) of the parameters $\sigma_0$, $\theta$ and $\rho$ (line~\ref{line:uborda} in Algorithm~\ref{alg:umm}).
+transformed sample (uBorda) is the same, i.e., polynomial. The most important steps are the update
+(\emph{learning}) of the parameters $\sigma_0$, $\theta$ and $\rho$ (line~\ref{line:uborda} in Algorithm~\ref{alg:umm}) and
+the sampling of a new permutation from $MM(\sigma_0,\theta)$ (line~\ref{line:sample}). % in Algorithm~\ref{alg:umm}).
 
@@
 
-\subsubsection{Learning step.}
-At each iteration we start by fitting a MM to the sample $S$, which implies
+\subsubsection{Learning step}
+%
+At each iteration, we start by fitting a MM to the sample $S$, which implies
 estimating $\hat\sigma_0$ and $\theta$. Parameter $\theta$ is updated
@@
 \begin{equation}
-r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma'\in S'}\rho^{f(\sigma')} \quad\text{where}\quad |S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\leq f(\sigma)
+  \begin{split}
+  r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma'\in S'}\rho^{f(\sigma')} \quad\text{where}\; &|S'| = r_2|S| \\[-1.2em]
+  &\land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\leq f(\sigma)
+  \end{split}
 \end{equation}
 %
-where $S'$ contains the $r_2|S|$ elements of $S$ with the best corresponding fitness values in $F$.
+i.e., $S'$ contains the $r_2|S|$ elements of $S$ with the best corresponding fitness values in $F$.
 The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification: higher values bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas smaller values make the sampling more uniform, thus slowing down the learning.
@@
 
-\subsubsection{Sampling step.}\label{sec:sampling}
+\subsubsection{Sampling step}\label{sec:sampling}
+%
 Sampling from $MM(\sigma_0,\theta)$  can be done efficiently given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{IruCalLoz2016permallows}, where the computational complexity is $O(n\log n)$.
@@
 Both  the expected value of $D$ and its variance increase monotonically with the value of $\theta$.
-In particular, the expected distance $\Edist$ is given by~\cite{gMallows}:
+In particular, the expected distance $\Edist$ is given by~\cite{FliVer1986}:
 %
@@
 
-\subsubsection{CEGO.}
+\subsubsection{CEGO}
+%
 We use the implementation of GECO provided by the
@@
 
-\subsubsection{UMM}. UMM uses the same budget of evaluation ($\FEmax$) as CEGO
+\subsubsection{UMM}
+%
+UMM uses the same budget of evaluations ($\FEmax$) as CEGO
 and initially samples $\minit=10$ solutions uniformly. The other two parameters
 of UMM are $r_1$ and $r_2$, which determine the balance between exploration and
-exploitation. In the experimental section we will study their impact. We
-implemented UMM using Python 3.
+exploitation. In the experimental section we will study their impact. implemented UMM using Python 3.
 
-\subsubsection{Other settings}.  For simplicity, we use Kendall's $\tau$
-distance~(Eq.~\ref{eq:kendall_decomp}) in all
-experiments. Nevertheless,
+\subsubsection{Other settings}
+%
+For simplicity, we use Kendall's $\tau$
+distance~(Eq.~\ref{eq:kendall_decomp}) in all experiments. Nevertheless,
 \citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
-when using Hamming distance.  We plan to extend UMM to other distance measures in the future.
-In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
-the actual objective function. In a white-box context, state-of-the-art
-algorithms for the problems considered here typically evaluate
-thousands of solutions, thus, the budget considered
-here for the black-box context is extremely limited.
+when using Hamming distance and the same may apply to UMM.
+We plan to extend UMM to other distance measures in the future.  In all
+experiments, we consider a maximum budget of $\FEmax=400$ evaluations of the
+actual objective function. In a white-box context, state-of-the-art algorithms
+for the problems considered here typically evaluate thousands of solutions,
+thus, the budget considered here for the black-box context is extremely
+limited.
 
-\subsubsection{Benchmark problems.} Experiments with real-world expensive black-box problems would be computationally infeasible.
-Instead, we consider classical combinatorial optimization problems as black-box optimization problems.
+\subsubsection{Benchmark problems}
+%
+Experiments with real-world expensive black-box problems would be computationally infeasible. Instead, we consider classical combinatorial optimization problems as black-box optimization problems.
 
@@
 The objective of the PFSP is to find:
-\begin{equation}\label{eq:qap}
+\begin{equation}\label{eq:pfsp}
   \begin{split}
@@
 
-\paragraph{Datasets}
+\subsubsection{Datasets}
+%
+For the LOP, we generated synthetic instances as follows.
+First, we create a sample $S$ of 200 permutations of size $n=20$ from a MM with $\sigma_0$ equal to the identity permutation and $\theta=-\log(\textit{Uniformity})$, where  $\textit{Uniformity}\in\{0.5,$ $0.7,$ $0.9\}$, i.e., from peaked distributions to almost uniform distributions.
+Then, we generate matrix $a$ as $a_{i,j} = \sum_{\sigma\in S}\sum_{i<j}\mathbf{I}\{\sigma(i)<\sigma(j)\}$, where $\mathbf{I}\colon S \times S \to \{0,1\}$ is the indicator function.
+In principle, these should be best-case instances for UMM since their fitness landscape closely matches the Mallows model.\MANUEL{Ekhine: Could you revise this text?
+I wanted to explain why we use synthetic datasets.}
 
-For the LOP, we generated synthetic instances as follows: sample 200 permutations from a MM of size $n=20$ and $\theta=-\log(uniformity)$, where  $uniformity=\{.5,.7,.9)\}$, i.e., from peaked distributions to almost uniform distributions.
-Then, generate matrix $a$ as $a_{i,j} = \sum_{\sigma\in S}\sum_{i<j}\mathbb{I}(\sigma(i)<\sigma(j))$.
-
-Moreover, we consider real LOP datasets from the LOPLIB repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
-In particular, we consider the following types: IO, Random instances of type A and Random instances of type B.
+Moreover, we consider real LOP datasets from the LOPLIB
+repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} In
+particular, we consider the following types: IO, Random instances of type A and
+Random instances of type B of various sizes $n \in \{50, 100, 150, 200,
+250\}$. These are harder problems than the synthetic instances, not only
+because of the larger search space, but also because of the different instance
+types.
 
@@
 \subsubsection{Computing environment.}
-All experiments were run on 
-Intel Xeon ``Ivybridge'' E5-2650v2 CPUs at 2.60\,GHz, 64\,GB RAM running CentOS Linux release 7.4.1708 (Core),
-\ekhine{}{
-Intel Xeon ``GenuineIntel'' E5-2680v3 CPUs at 2.50\,GHz, 32???\,GB RAM running Linux,
-Intel Xeon ``GenuineIntel'' E5-2683v4 CPUs at 2.10\,GHz, 32???\,GB RAM running Linux.
-}
+Due to the computational effort required, we used several computing systems to
+run experiments. All UMM experiments were either on run on Intel Xeon
+\emph{Broadwell} E5-2683v4 CPUs at 2.10\,GHz or Intel Xeon \emph{Haswell} E5-2680v3
+CPUs at 2.50\,GHz, both of them with 32\,GB RAM running Linux.\MANUEL{If you enter in a node and do \texttt{cat /etc/lsb-release} what does it say?} %
+Experiments with
+CEGO used a more powerful system: Intel Xeon \emph{Ivybridge} E5-2650v2 CPUs at
+2.60\,GHz, 64\,GB RAM running CentOS Linux release 7.4.1708 (Core).
+
 %processor : 23
@@
 Figure~\ref{fig:heatmaps} shows the best solution found by each independent run (averaged over 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP.
-This experiment for all the real datasets can be bound in the supplemental material.
+We terminated runs that took longer than 6 hours and those combinations are shown as whitespace.
+Results for all the datasets can be found in the supplemental material.
 
@@
 \end{minipage}
-\caption{Mean fitness  for different configuration for the learning rate (darker is better). \label{fig:heatmaps}}
+\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better). Whitespace denotes missing data (runs that took longer than 6 hours).\label{fig:heatmaps}}
 \end{figure}

### 2020-10-31 launch_exps.sh
results on pfsp/rec31

@@ -60,7 +60,7 @@ gen_lop_synthetic() {
     echo $INSTANCES
 }
 
-# INSTANCES="\
+#INSTANCES="\
 # qap/kra32.dat \
 # qap/nug12.dat \
 # qap/nug30.dat \
@@ -69,9 +69,11 @@ gen_lop_synthetic() {
 # pfsp/rec13.txt \
 # pfsp/rec19.txt \
 # pfsp/rec31.txt \
-# "
+#"
+
+INSTANCES=" pfsp/rec31.txt "
 #INSTANCES=$(gen_lop_synthetic $INSTANCES)
-INSTANCES="$(tr '\n' ' ' < loplib-instances.txt)"
+#INSTANCES="$(tr '\n' ' ' < loplib-instances.txt)"
 
 
 

### 2020-10-31 paper/main.tex
more comments.

@@
 \newcommand{\FEmax}{\ensuremath{m}}
+
+\newcommand{\supplement}{~(\href{http://doi.org/10.5281/zenodo.4167148}{doi:10.5281/zenodo.4167148})}
 
@@
   experiment. In such cases, classical techniques such as mixed-integer
-  programming and local search cannot be applied. Moreover, often each solution
-  evaluation is expensive in terms of time or resources, thus only a limited
-  number of evaluations is possible, typically several order of magnitude
-  smaller than in white-box optimization problems. In the continuous case,
-  Bayesian optimization, in particular using Gaussian processes, has proven
-  very effective under these conditions. Much less research is available in the
-  combinatorial case. In this paper, we propose and analyze an
+  programming and local search cannot be applied. Moreover, each fitness
+  evaluation is often expensive in terms of time or resources, thus only a
+  limited number of evaluations is possible, typically several orders of
+  magnitude smaller than in white-box optimization problems. In the continuous
+  case, Bayesian optimization, in particular using Gaussian processes, has
+  proven very effective under these conditions. Much less research is available
+  in the combinatorial case. In this paper, we propose and analyze UMM, an
   estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
-  model and compare it with CEGO, a Bayesian optimization algorithm for
-  combinatorial optimization. Experimental results show that the Bayesian
-  algorithm is able to obtain very good solutions with very few evaluations,
-  however, at a significant computational cost, whereas the proposed EDA
-  outperforms CEGO when the number of solutions evaluated approaches 400, and
-  it is significantly faster. These results suggest that the combination of
-  Bayesian optimization and a Mallows-based EDA may be an interesting direction
-  for future research.\sloppy  
+  model and unbalanced rank aggregation (uBorda). Experimental results on
+  black-box versions of LOP, QAP and PFSP show that UMM is able to match, and
+  sometimes surpass, the solutions obtained by CEGO, a Bayesian optimization
+  algorithm for combinatorial optimization, specially when the budget of
+  solutions evaluated approaches 400. The computational cost of UMM is at least
+  an order of magnitude smaller than CEGO, which makes it specially useful when
+  both computation time and number of evaluations are of concern.\sloppy
   \textcolor{red}{16 pages max, deadline: 1 November 2020}
-\keywords{Combinatorial optimization \and Bayesian optimization \and Expensive black-box optimization \and Estimation of distribution algorithms}
+  \keywords{Combinatorial optimization \and Bayesian optimization \and
+    Expensive black-box optimization \and Estimation of distribution
+    algorithms}
 \end{abstract}
@@
 anyway an inexact estimation of the original fitness
-function~\citep[p.~875]{ZaeStoFriFisNauBar2014}.
+function~\citep[p.~875]{ZaeStoFriFisNauBar2014}. A thorough comparison of CEGO
+with other optimizers for expensive black-box combinatorial optimization ranked
+it as the best-performing~\citep{ZaeStoFriFisNauBar2014}.
 
@@
 CEGO used a budget of $10^4$ surrogate-model evaluations for each run of the
-GA. However, with such a setting, a single run of CEGO with a budget of $\FEmax=400$
-true fitness evaluations requires \emph{more than a week} of computation time,
-which was infeasible for our study. Hence, we set the budget of the GA to
-$10^3$. Although a higher GA budget could improve the results of CEGO reported
-here, our results are similar (and sometimes better) than the ones reported in
-the original papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to
-running CEGO with $\FEmax=400$ instead of $\FEmax=200$. In fact, we found that the
-increase in runtime caused by doubling $\FEmax$ is much smaller than by increasing
-the GA budget by a factor of $10$. Nevertheless, we caution that choosing a
+GA. However, with such a setting, a single run of CEGO with a budget of
+$\FEmax=400$ true fitness evaluations requires \emph{more than a week} of
+computation time, which was infeasible for our study. Due to extremely long
+runtimes of CEGO and constraints in our computing system, we set a maximum
+wall-clock time of 3 days for each run of CEGO. Hence, we set the budget of the
+GA to $10^3$. Although a higher GA budget could improve the results of CEGO
+reported here, our results are similar (and sometimes better) than the ones
+reported in the original
+papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to running CEGO
+with $\FEmax=400$ instead of $\FEmax=200$. In fact, we found that the increase
+in runtime caused by doubling $\FEmax$ is much smaller than by increasing the
+GA budget by a factor of $10$. Nevertheless, we caution that choosing a
 different trade-off may lead to different conclusions and more work is needed
@@
 Figure~\ref{fig:heatmaps} shows the best solution found by each independent run (averaged over 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP.
-We terminated runs that took longer than 6 hours and those combinations are shown as whitespace.
-Results for all the datasets can be found in the supplemental material.
+We terminated runs that took longer than 6 hours and those combinations are shown as white.
+Results for all the datasets can be found in the supplemental material\supplement.
 
@@
  \includegraphics[width=\textwidth]{../img/heatmap_N-t1d150_01}
-\end{minipage}\\\vskip 1ex
+\end{minipage}\\\vskip 1em
 \begin{minipage}{0.49\linewidth}
@@
 $r_2$, yet, low values of $r_1$ and high values of $r_2$ almost always produce
-the best results, independently of the problem. Nevertheless, there are a few
-cases where an interaction seems to occur and particular combinations perform
-remarkably well.\MANUEL{Missing data, whitespaces where more than 6 hours of wall-clock time}  In particular, we observe that the pattern on
-the two left-most figures differs from the two right-most ones. While on the
-left the fitness gets worse homogeneously from the top right corner to the
-bottom left one, the pattern on the right is not as homogeneous, yet still
-fitness gets worse with decreasing $r_2$.  We observe that this behavior is
-related to the adequacy of UMM to the problem: the homogeneous behavior of
-parameters $r_1$ and $r_2$ depends on the problem (not to the particular
+the best results, independently of the problem.  In particular, we observe that
+the pattern on the two top figures differs from the two bottom ones. While on
+the top figures the fitness gets worse homogeneously from the top-right corner
+to the bottom-left one, the pattern on the bottom figures is not as homogeneous, yet
+still fitness gets worse with decreasing $r_2$.  We observe that this behavior
+is related to the adequacy of UMM to the problem: the homogeneous behavior of
+parameters $r_1$ and $r_2$ depends on the problem (not on the particular
 instances) and the more homogeneous the behavior, the better performance of
-UMM. We elaborate this argument in the following sections with detail.\MANUEL{top and bottom}
-
-Over all the instances  we
-see that the configuration $r_1 =0.1$, $r_2=0.9$ has good
-performance: \ekhine{}{the rank of this parameter configuration among all the possible configurations is less than 4 on average for all the instances}. From this
-preliminary experiment, it is not obvious how the best settings are related to
-problem features, such as permutation length ($n$). On the other hand, given
-the sharp transitions in the plots, it is clear that further fine-tuning may
-improve performance and uncover further patterns.  Although automatic parameter
-tuning may likely provide even more fine-tuned settings, our goal here is to
-understand how UMM performs in comparison to an existing Bayesian method (CEGO)
-and automatically tuning the parameters of CEGO is not computationally
-feasible.  For consistency, in the following sections, we run UMM with this
-parameter configuration. Moreover, we recommend this configuration as the
-current default of UMM.
-
-\subsection{Analysis of UMM and CEGO}
+UMM. We elaborate this argument in detail in the next section.
+
+From this preliminary experiment, it is not obvious how the best settings are
+related to problem features, such as permutation length ($n$). On the other
+hand, given the sharp transitions in the plots, it is clear that further
+fine-tuning may improve performance and uncover further patterns.
+
+Over all the instances, we observe that the configuration $r_1 =0.1$, $r_2=0.9$
+has good performance, i.e., its mean rank (over 10 runs) relative to all the
+other configurations is less than 4 for all the instances.  Although automatic
+parameter tuning may likely provide even more fine-tuned settings, our goal
+here is to understand how UMM performs in comparison to an existing Bayesian
+method (CEGO) and automatically tuning the parameters of CEGO is not
+computationally feasible.  For consistency, in the following sections, we run
+UMM with this parameter configuration.  Moreover, we recommend this
+configuration as the current default of UMM.
+
+\subsection{Analysis of UMM and CEGO on synthetic LOP instances}
 
@@
 Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the
-internal probability model of the algorithm over the $m$ functions
-evaluations. Overall, \ekhine{UMM found  the best solutions in all the repetitions. }{the mean best solutions of UMM are better than the mean of CEGO regardless the uniformity of the instances. }
+internal probability model of the algorithm over the $\FEmax$ functions
+evaluations. Overall, the mean fitness returned by UMM after $\FEmax$ is better one returned by  CEGO, regardless the uniformity of the instances.
 
@@
 halted at this point and the model seems to have trouble generating better
-solutions. Lower uniformity values (e.g., $\phi=.5$) show worsening evaluations
+solutions. Lower uniformity values (e.g., \manuel{$\phi=.5$}{0.5})\MANUEL{We never use $\phi$, not even in the plots, so less have one less symbol to worry about.} show worsening evaluations
 despite the increase in solutions used to build the model. In summary, the
@@
 %  question.}
-
-\newcommand{\supplement}{\citep{Supplementary}}
 
@@
 
-  In the following paragraphs we analyze the result over real rata. Fig.~\ref{fig:loplib} shows the results on the instances available in LOPLIB.
+\subsection{Analysis of UMM and CEGO on LOP, QAP and PFSP}
+
+In the following paragraphs we analyze the result over real-world instances. %
+Figure~\ref{fig:loplib} shows the results on the instances taken from LOPLIB.
   %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
   % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
-  The distribution of this instances differs from the symthetic LOP instances, and thus so does the landscape.
-However, UMM shows a very similar performance: there is a (close to) linear decrease on the function evaluations of new solutions along iterations.
-This suggests that UMM very good fits the fitness landscape.
-Regarding CEGO, we can take similar conclussions for the first 50-100 iterations.
-However, the processes had to be halted after hitting the wall time, so no conlusion can be taken for further iterations.
-Overall, the improvement from the first (random) to the last solutions is 2-3 times larger in UMM than CEGO.
+The distribution\MANUEL{what do you mean the distribution? Is this a feature of
+  the matrix $a$?} of these instances differs from the synthetic LOP instances,
+and, thus, so does the landscape. However, UMM shows a very similar behavior:
+there is a (close to) linear decrease \manuel{on the function evaluations of
+  new solutions along iterations}{of the fitness of new solutions over each new
+  evaluation}. This suggests that UMM \manuel{very good fits}{is able to model}
+the fitness landscape. Regarding CEGO, we \manuel{can take similar
+  conclusions}{observe a similar behavior} for the first 50-100
+\manuel{iterations}{evaluations}. At this point, CEGO reaches the time limit of
+3 days, so no observations can be made for longer budgets.  Overall, the
+improvement from the first (random) to the last solutions is 2-3 times larger
+in UMM than CEGO. %
+\MANUEL{But this is a bit misleading because UMM runs for 4 times more
+  evaluations, perhaps we should also measure up to the evaluation that CEGO
+  reaches so we report both values?}
 
@@
 better than a random search would, as illustrated by the results
-(Fig.~\ref{fig:rec05}) obtained on \texttt{rec05}, which is the smallest PFSP
+ obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP
 instance considered here. The lack of any apparent convergence suggests that
 the models (both Bayesian and probabilistic) are not learning anything about
-the fitness landscape. Results for larger PFSP corroborate these conclusions
-(\supplement). Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
+the fitness landscape. Results for larger PFSP corroborate these conclusions\supplement.
+\manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
 that other distance measures are more suited than Kendall's-$\tau$ distance for
 the PFSP, the reported fitness differences between various distance metrics are
-small, hence, we conjecture that the same behaviour will be observed with other
-distance metrics.
+small.
+Hence, we conjecture that the same behavior will be observed with other distance metrics.}{}\MANUEL{On second thought, I'm not sure we should say this, because we really have no idea and we actually hope it is not true.}
 
@@
 behaviour as in the PFSP for all instances tested, except for
-\texttt{kra32}. As shown in Fig~\ref{fig:kra32}, in this instance both
+\texttt{kra32}. As shown in Fig~\ref{fig:kra32}, for \texttt{kra32} both
 algorithms are able to improve over the number of evaluations, although not
 with ease. As in the synthetic LOP instances, we again observe that CEGO does
-improve faster than UMM initially, yet, around evaluation 250, it appears to
-get stuck and around evaluation 300, UMM overtakes it. We believe the reason is
-the same as before, that is, the model is not able to adequately reflect the
+improve faster than UMM initially, yet CEGO appears to
+get stuck around evaluation 250,  and UMM overtakes it around evaluation 300. We believe the reason is
+the same as before, that is, the CEGO model is not able to adequately reflect the
 ruggedness of the actual landscape and, hence, the best solutions found by the
@@
 
-Table~\ref{tab:results} shows overall results over all instances evaluated (excluding the synthetic LOP instances).
-In particular, the mean fitness (and standard deviation) of the best solution found at the end of each run,
-In terms of best fitness achieved on each run, and taking into account the
-random behaviour observed in some instances and discussed above,\ekhine{}{falta un cacho aqui}\MANUEL{Mío}
+Table~\ref{tab:results} shows results over all instances evaluated (excluding
+the synthetic LOP instances). In particular, we show the mean fitness (and
+standard deviation) of the best solution found at the end of each run.
+We also report the mean runtime in hours and, since CEGO sometimes reached the limit of 72 hours, also the number of fitness evaluations reached by CEGO.
+
+In the LOP, differences are mostly in favour of UMM, due to the time limit of
+72 hours, which resulted in CEGO typically achieving less than 200
+evaluations. In contrast, UMM was always able to reach 400 evaluations in less
+than 16 hours. If UMM was allowed to reach the 72 hours limit, the differences
+would be even stronger.
+
+In the case of the PFSP, the differences are minor and/or not statistically
+significant, with the qualification that both algorithms appear to be
+performing a random search. Similar observations can be made for the QAP, where
+only on the smallest instance \texttt{nug12}, there is a significant different
+in favour of CEGO.
+
+The overall conclusion from Table~\ref{tab:results} is that there is not a
+significant difference in performance between UMM and CEGO. For the larger LOP
+instances, CEGO reaches the time limit of 72 hours before consuming the budget
+of 400 evaluations and thus, obtains quite poor solutions compared to UMM,
+which is at least an order of magnitude faster.
 
@@
 Mean runtime is measured in hours.
-If a run went over 72 hours, of  CEGO was not able to reach 400 function evaluations (FEs) in 72 hours, it was terminated earlier.
+If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 72 hours, it was terminated earlier.
 This was the case for all LOP instances shown.\manuel{}{Don't we have results for rec31?}\label{tab:results}}
@@
        & rec19        & 2398.6                  & (19.8)     & 2405.1    & (12.6)                 & $[$-22.3,     & 9.3$]$        & 400.0& 82.1 & 0.3  \\\midrule
-    QAP     & kra32        & 115907.0                & (4663.4)   & 116320.0  & (4673.5)               & $[$-4799.3,   & 3973.3$]$     & 400.0& 95.8 & 0.3  \\
-        & nug12        & 624.2                   & (26.3)     & 660.4     & (17.3)                 & $[$-57.3,     & -15.1$]$      & 400.0& 14.8 & 0.0  \\
+    QAP  & nug12        & 624.2                   & (26.3)     & 660.4     & (17.3)                 & $[$-57.3,     & -15.1$]$      & 400.0& 14.8 & 0.0  \\
          & nug30        & 7486.8                  & (40.4)     & 7474.0    & (83.9)                 & $[$-50.8,     & 76.4$]$       & 400.0& 84.3 & 0.3  \\
         & tho30        & 192416.0                & (2584.2)   & 191482.0  & (4419.7)               & $[$-2527.0,   & 4395.0$]$     & 400.0& 85.1 & 0.3  \\
+         & kra32        & 115907.0                & (4663.4)   & 116320.0  & (4673.5)               & $[$-4799.3,   & 3973.3$]$     & 400.0& 95.8 & 0.3  \\
 \bottomrule
@@
 Creo q es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos  y luego que addemás es más rápido}
+\MANUEL{Pero UMM es peor en rec13 y nug12, y en LOP es mejor porque CEGO no llega a 400, es decir, porque UMM es más rápido.
+No podemos saber qué pasaría si CEGO llegara a 400.
+Podríamos medir lo que pasa en el momento en que matamos a CEGO y dar también ese valor entonces la diferencia en velocidad no importaría.
+}
 per fitness evaluation. The time required to train the surrogate-model of CEGO
@@
 to match (and sometimes even improve) the results of a Bayesian
-optimizer\EKHINE{hhmmm lo hacemos mejor o igual}. The results presented here
+optimizer.
+% MANUEL: OK?
+% \EKHINE{hhmmm lo hacemos mejor o igual}
+The results presented here
 show that, despite the intrinsic difficulty of this scenario, there are still

### 2020-10-31 cego.py
Rename uMM to UMM in .py files

@@ -271,7 +271,7 @@ def runCEGO(instance, m_ini, m, rep, best_known_sol, worst_known_sol, budgetGA):
     df['Solver'] = 'CEGO'
     df['Sample size'] = range(m)
     df['rep'] = rep
-    #    df['budgetGA'] = budgetGA this must be set for all, including uMM, so that we can filter appropriately
+    #    df['budgetGA'] = budgetGA this must be set for all, including UMM, so that we can filter appropriately
     # FIXME: We should calculate the distance outside here so we do not need the best_known_sol here.
     if best_known_sol is not None:
         df['Distance'] = [ kendallTau(perm, best_known_sol) / (instance.n * (instance.n - 1) / 2) for perm in instance.solutions]

### 2020-10-31 runner.py

@@ -33,11 +33,11 @@ def get_problem(instance_name):
 
 def run_once(algo_name, instance_name, seed, out_filename = None,
              **algo_params):
-    if algo_name == "uMM":
+    if algo_name == "UMM":
         import umm as umm
         reload(umm)
-        from umm import uMM
-        algo = uMM
+        from umm import UMM
+        algo = UMM
     elif algo_name == "CEGO":
         from cego import cego
         algo = cego

### 2020-10-31 target-runner-umm.py

@@ -22,7 +22,7 @@ import pandas as pd
 import runner
 
 from argparse import ArgumentParser,RawDescriptionHelpFormatter,_StoreTrueAction,ArgumentDefaultsHelpFormatter,Action
-parser = ArgumentParser(description = "uMM")
+parser = ArgumentParser(description = "UMM")
 parser.add_argument('configuration_id', type=str, help='configuration_id')
 parser.add_argument('instance_id', type=str, help='instance_id')
 parser.add_argument('seed', type=int, help='random seed')
@@ -44,7 +44,7 @@ assert budget > 2 * args.m_ini
 stdout = sys.stdout
 outfilename = f'c{args.configuration_id}-{args.instance_id}-{args.seed}.stdout'
 with open(outfilename, 'w') as sys.stdout:
-    df = runner.run_once("uMM", args.instance_name, args.seed, budget = budget, m_ini = args.m_ini, 
+    df = runner.run_once("UMM", args.instance_name, args.seed, budget = budget, m_ini = args.m_ini, 
                          budgetMM = args.budgetMM, ratio_samples_learn = args.rsl, weight_mass_learn = args.wml, out_filename = args.output)
             
 sys.stdout = stdout

### 2020-10-31 umm.py

@@ -39,7 +39,7 @@ def get_expected_distance(iterat,n,m):
     return max(a, f_end)
 
 
-def uMM(instance,
+def UMM(instance,
         seed, budget,
         m_ini,
         budgetMM,

### 2020-10-31 setup.sh
add some files for reproduction.

@@ -1,0 +1,11 @@
+#!/bin/sh
+
+# Install the R package `CEGO` v2.4.0. Original is available at:
+#   https://CRAN.R-project.org/package=CEGO
+#
+# We used R 3.4.4 for our experiments but later versions should also work.
+R CMD INSTALL CEGO_2.4.0.tar.gz
+
+# Install rpy2, we used rpy2-3.3.5
+pip install --force-reinstall --update --user rpy2
+

### 2020-10-31 1-launch_exps.sh
* rename launch_exps.sh to 1-launch_exps.sh

@@ -60,6 +60,7 @@ gen_lop_synthetic() {
     echo $INSTANCES
 }
 
+## For QAP, PFSP instances
 #INSTANCES="\
 # qap/kra32.dat \
 # qap/nug12.dat \
@@ -71,11 +72,12 @@ gen_lop_synthetic() {
 # pfsp/rec31.txt \
 #"
 
-INSTANCES=" pfsp/rec31.txt "
+###### For synthetic LOP instances
 #INSTANCES=$(gen_lop_synthetic $INSTANCES)
-#INSTANCES="$(tr '\n' ' ' < loplib-instances.txt)"
 
 
+###### For LOPLIB instances
+INSTANCES="$(tr '\n' ' ' < loplib-instances.txt)"
 
 budget=400
 nruns=10
@@ -86,41 +88,22 @@ LAUNCHER=qsub_job
 cego_m_ini=10
 budgetGA=3 # Actually, 10**budgetGA
 
-RESULTS=results-loplib
-mkdir -p $RESULTS
 counter=0
 for instance in $INSTANCES; do
     counter=$((counter+1))
+    RESULTS="results/$instance"
+    mkdir -p $RESULTS
     for run in $(seq 1 $nruns); do
-	$LAUNCHER cego-$counter-$run ./target-runner-cego.py cego $counter $run $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $budget --output $RESULTS/cego-$counter-$run
+	### Uncomment for running CEGO
+	$LAUNCHER cego-$counter-r$run ./target-runner-cego.py cego $counter $run $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $budget --output $RESULTS/cego-r$run
 
-        rsl=0.15
-	wml=0.84
-	budgetMM=15
-	umm_m_ini=23
+	### Uncomment for running UMM
+	#rsl=0.15
+	#wml=0.84
+	#budgetMM=15
+	#umm_m_ini=23
 	#$LAUNCHER umm-$counter-$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $rsl --wml $wml --budget $budget --output results/umm-$counter-$run
         
-
-        
[…]

### 2020-10-31 paper/main.tex
-a
more edits

@@
 
-In the case of the PFSP, none of the algorithms are able to achieve a progress
-better than a random search would, as illustrated by the results
+In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
  obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP

### 2020-11-07 cego.py
* Run for 6 days.

@@ -34,7 +34,7 @@ optimCEGO <- function (x = NULL, fun, control = list())
         infill = infillExpectedImprovement, model = modelKriging, 
         modelSettings = list(), optimizer = optimEA, optimizerSettings = list(), 
         initialDesign = designMaxMinDist, archiveModelInfo = NULL, 
-        initialDesignSettings = list(), maxTime = 3600 * 24 * 3)
+        initialDesignSettings = list(), maxTime = 3600 * 24 * 6)
     con[names(control)] <- control
     control <- con
     rm(con)

### 2020-11-07 paper/main.tex
remove unused

@@
 far. These classical EDAs  assume that function
-evaluations are cheap and/or the fitness function is white-box.\MANUEL{Do EDAs really assume that the fitness function is white-box? ACO doesn't.}
+evaluations are cheap and/or the fitness function is white-box.\MANUEL{Do EDAs really assume that the fitness function is white-box? ACO doesn't: They are hybridized with local search }
 %The proposed UMM drastically differs from the above algorithms.
@@
 We represent permutations as an ordered vector and, we say that position $i$ contains value $j$ whenever $\sigma(i)=j$.
-We denote permutations with Greek letters.\MANUEL{Do we use $e$ anywhere?
-If not, we do not need this definition ``with exception of the identity permutation denoted as $e=(1, 2, 3, \dotsc,n)$.''} We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.
+We denote permutations with Greek letters.
+% \MANUEL{Do we use $e$ anywhere? If not, we do not need this definition ``with exception of the identity permutation denoted as $e=(1, 2, 3, \dotsc,n)$.''}
+We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.
 
@@
 \emph{Broadwell} E5-2683v4 CPUs at 2.10\,GHz or Intel Xeon \emph{Haswell} E5-2680v3
-CPUs at 2.50\,GHz, both of them with 32\,GB RAM running Linux.\MANUEL{If you enter in a node and do \texttt{cat /etc/lsb-release} what does it say?} %
+CPUs at 2.50\,GHz, both of them with 32\,GB RAM. 
+ %
 Experiments with
 CEGO used a more powerful system: Intel Xeon \emph{Ivybridge} E5-2650v2 CPUs at
-2.60\,GHz, 64\,GB RAM running CentOS Linux release 7.4.1708 (Core).
-
+2.60\,GHz, 64\,GB RAM. % running CentOS Linux release 7 (Core).
+The operating system was CentOS Linux release 7 (Core).
 %processor : 23
@@
 halted at this point and the model seems to have trouble generating better
-solutions. Lower uniformity values (e.g., \manuel{$\phi=.5$}{0.5})\MANUEL{We never use $\phi$, not even in the plots, so less have one less symbol to worry about.} show worsening evaluations
+solutions. Lower uniformity values, e.g., 0.5,
+ show worsening evaluations
 despite the increase in solutions used to build the model. In summary, the
@@
   % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
-The distribution\MANUEL{what do you mean the distribution? Is this a feature of
-  the matrix $a$?} of these instances differs from the synthetic LOP instances,
-and, thus, so does the landscape. However, UMM shows a very similar behavior:
-there is a (close to) linear decrease \manuel{on the function evaluations of
-  new solutions along iterations}{of the fitness of new solutions over each new
-  evaluation}. This suggests that UMM \manuel{very good fits}{is able to model}
-the fitness landscape. Regarding CEGO, we \manuel{can take similar
-  conclusions}{observe a similar behavior} for the first 50-100
-\manuel{iterations}{evaluations}. At this point, CEGO reaches the time limit of
-3 days, so no observations can be made for longer budgets.  Overall, the
-improvement from the first (random) to the last solutions is 2-3 times larger
-in UMM than CEGO. %
+The type of instances from the synthetic LOP instances, and, thus, so does the
+landscape. However, UMM shows a very similar behavior: there is a (close to)
+linear decrease of the fitness of new solutions over each new evaluation. This
+suggests that UMM is able to model the fitness landscape. Regarding CEGO, we
+observe a similar behavior for the first 50-100 evaluations. At this point,
+CEGO reaches the time limit of 3 days, so no observations can be made for
+longer budgets.  Overall, the improvement from the first (random) to the last
+solutions is 2-3 times larger in UMM than CEGO. %
 \MANUEL{But this is a bit misleading because UMM runs for 4 times more
@@
 the fitness landscape. Results for larger PFSP corroborate these conclusions\supplement.
-\manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
-that other distance measures are more suited than Kendall's-$\tau$ distance for
-the PFSP, the reported fitness differences between various distance metrics are
-small.
-Hence, we conjecture that the same behavior will be observed with other distance metrics.}{}\MANUEL{On second thought, I'm not sure we should say this, because we really have no idea and we actually hope it is not true.}
+% \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
+% that other distance measures are more suited than Kendall's-$\tau$ distance for
+% the PFSP, the reported fitness differences between various distance metrics are
+% small.
+% Hence, we conjecture that the same behavior will be observed with other distance metrics.}{}\MANUEL{On second thought, I'm not sure we should say this, because we really have no idea and we actually hope it is not true.}
 
@@
 results, a Bayesian optimizer such as CEGO may sometimes require more than hour
+\MANUEL{UMM is able to model the LOP landscape at least as well as CEGO as shown by the plots where UMM shows a linear decrease in fitness.}
 \EKHINE{me parece mas de CEGO q UMM y parece q UMM solo tiene de ventaja ir rápido.

### 2020-11-07 codeLOP.cpp

@@ -1,21 +1,0 @@
-int kendall_distance(int n_, int*s1, int*s2){
-    int dist = 0 ;
-    int a, b;
-    for(int i = 0 ; i < n_ ; i ++)
-      for (int j=i+1; j < n_ ; j++){
-         a =  s1[i] - s1[j];
-         b =  s2[i] - s2[j];
-         if (a*b < 0 ) dist +=1;
-       }
-    return dist;
-}
-
-
-int main(int argc, char const *argv[]) {
-  /* code */
-  int n_ = 8;
-  int     *sigma_0 = new int[n_];
-    int     *sigma_0_estim = new int[n_];
-  kendall_distance(n_, sigma_0, sigma_0_estim);
-  return 0;
-}

### 2020-11-11 1-launch_exps.sh
more cego results

@@ -87,6 +87,11 @@ LAUNCHER=qsub_job
 cego_m_ini=10
 budgetGA=3 # Actually, 10**budgetGA
 
+r_1=0.15
+r_2=0.84
+budgetMM=15
+umm_m_ini=10
+
 counter=0
 for instance in $INSTANCES; do
     counter=$((counter+1))
@@ -94,14 +99,10 @@ for instance in $INSTANCES; do
     mkdir -p $RESULTS
     for run in $(seq 1 $nruns); do
 	### Uncomment for running CEGO
-	$LAUNCHER cego-$counter-r$run ./target-runner-cego.py cego $counter $run $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $budget --output $RESULTS/cego-r$run
+	#$LAUNCHER cego-$counter-r$run ./target-runner-cego.py cego $counter $run $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $budget --output $RESULTS/cego-r$run
 
 	### Uncomment for running UMM
-	#rsl=0.15
-	#wml=0.84
-	#budgetMM=15
-	#umm_m_ini=23
-	#$LAUNCHER umm-$counter-$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $rsl --wml $wml --budget $budget --output results/umm-$counter-$run
+	$LAUNCHER umm-$counter-$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --output $RESULTS/cego-r$run
         
     done
 done

### 2020-11-11 1-launch_exps.sh
launch tuning params

@@ -76,7 +76,7 @@ INSTANCES="\
 #INSTANCES=$(gen_lop_synthetic $INSTANCES)
 
 ###### For LOPLIB instances
-#INSTANCES="$(tr '\n' ' ' < loplib-instances.txt)"
+INSTANCES="$(tr '\n' ' ' < loplib-instances.txt)"
 
 budget=400
 nruns=10
@@ -87,9 +87,9 @@ LAUNCHER=qsub_job
 cego_m_ini=10
 budgetGA=3 # Actually, 10**budgetGA
 
-r_1=0.15
-r_2=0.84
-budgetMM=15
+r_1=0.1
+r_2=0.9
+budgetMM=10
 umm_m_ini=10
 
 counter=0
@@ -102,8 +102,27 @@ for instance in $INSTANCES; do
 	#$LAUNCHER cego-$counter-r$run ./target-runner-cego.py cego $counter $run $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $budget --output $RESULTS/cego-r$run
 
 	### Uncomment for running UMM
-	$LAUNCHER umm-$counter-$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --output $RESULTS/cego-r$run
+	$LAUNCHER umm-$counter-$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --output $RESULTS/umm-r$run
         
     done
 done
 
+RERUN_params=0
+if [ $RERUN_params -eq 0 ]; then
+    exit 0
+fi
+
+counter=0
+for instance in $INSTANCES; do
+    for r_1 in $(seq 0.1 0.1 0.5); do
+	for r_2 in "$(seq 0.6 0.1 0.9) 0.99"; do
+	    counter=$((counter+1))
+	    RESULTS="results-r1r2/$instance"
+	    mkdir -p $RESULTS
+	    for run in $(seq 1 $nruns); do
+		### Uncomment for running UMM
+		$LAUNCHER umm-$counter-$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --output $RESULTS/umm-r$run
+	    done
+	done
+    done
+done

### 2020-11-11 paper/main.tex
missing things in algorithm
update paper
update plots in paper

@@
 However, the setting of a decreasing variance   controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
-Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.
+Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.\EKHINE{Explain budgetMM.
+we sample several permutations and choose one only!}
 
@@
   \centering%
+  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandB_N-p50-01}\\
   \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d100_01}\\
   \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d150_01}\\
-  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d200_01}\\
+%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d200_01}\\
     \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}

### 2020-11-11 paper/main.tex
some edits

@@
 computation time, which was infeasible for our study. Due to extremely long
-runtimes of CEGO and constraints in our computing system, we set a maximum
-wall-clock time of 3 days for each run of CEGO. Hence, we set the budget of the
-GA to $10^3$. Although a higher GA budget could improve the results of CEGO
-reported here, our results are similar (and sometimes better) than the ones
-reported in the original
-papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to running CEGO
-with $\FEmax=400$ instead of $\FEmax=200$. In fact, we found that the increase
-in runtime caused by doubling $\FEmax$ is much smaller than by increasing the
+runtimes of CEGO and constraints in our computing system, we set the budget of
+the GA to $10^3$ and we stopped each run of CEGO after 144 hours of CPU-time.
+Although a higher GA budget could improve the results of CEGO reported here,
+our results are similar (and sometimes better) than the ones reported in the
+original papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to
+running CEGO with $\FEmax=400$ instead of $\FEmax=200$. In fact, we found that
+doubling $\FEmax$ produces a smaller increase in runtime than multiplying the
 GA budget by a factor of $10$. Nevertheless, we caution that choosing a
@@
 standard deviation) of the best solution found at the end of each run.
-We also report the mean runtime in hours and, since CEGO sometimes reached the limit of 72 hours, also the number of fitness evaluations reached by CEGO.
+We also report the mean runtime in hours and, since CEGO sometimes reached the limit of 144 hours, also the number of fitness evaluations reached by CEGO.
 
 In the LOP, differences are mostly in favour of UMM, due to the time limit of
-72 hours, which resulted in CEGO typically achieving less than 200
+144 hours, which resulted in CEGO typically achieving less than 200
 evaluations. In contrast, UMM was always able to reach 400 evaluations in less
-than 16 hours. If UMM was allowed to reach the 72 hours limit, the differences
+than 16 hours. If UMM was allowed to reach the 144 hours limit, the differences
 would be even stronger.
@@
 significant difference in performance between UMM and CEGO. For the larger LOP
-instances, CEGO reaches the time limit of 72 hours before consuming the budget
+instances, CEGO reaches the time limit of 144 hours before consuming the budget
 of 400 evaluations and thus, obtains quite poor solutions compared to UMM,
@@
 Mean runtime is measured in hours.
-If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 72 hours, it was terminated earlier.
+If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 144 hours, it was terminated earlier.
 This was the case for all LOP instances shown.\manuel{}{Don't we have results for rec31?}\label{tab:results}}

### 2020-11-11 lop.py
update NB

@@ -11,8 +11,8 @@ def generate_list_of_instances(opt_filename):
             instance_name = find_in_lop_folder(name, "path")
             if instance_name:
                 print(instance_name)
-    
-    
+
+
 def find_in_lop_folder(instance_name, ret_value='instance'):
     for fol in os.listdir("lop"):
         subdir = "lop/" + fol
@@ -35,10 +35,10 @@ def read_best_known(opt_filename, instance_name):
                 # The instances in http://grafo.etsii.urjc.es/optsicom/lolib/#instances and best sols try to maximize the superdiagonal.
                 # We minimize the subdiagonal, which is equivalent, but the best sol needs to be updated
                 return int(value)
-                
+
     print(f"Instance {instance_name} not found in {opt_filename}")
     return None
-            
+
 def synthetic_LOP(n, m, phi):
     instance = np.zeros((n,n))
     s = np.asarray(mk.samplingMM(m, n, phi=phi, k=None))
@@ -67,7 +67,7 @@ class LOP(Problem):
                    instance_name = f"LOP-synthetic,seed={seed},n={n},m={m},phi={phi}")
 
     @classmethod
-    def read_instance(cls, filename, opt_filename = "lop/best_knowns.csv"):
+    def read_instance(cls, filename, opt_filename = "./best_knowns.csv"):
         if "synthetic" in filename:
             seed, n, m, phi = re.search("seed=([0-9]+),n=([0-9]+),m=([0-9]+),phi=([^ ]+)", filename).group(1,2,3,4)
             print(f"Generating synthetic LOP instance with seed={seed} n={n} m={m} phi={phi}")
@@ -83,7 +83,7 @@ class LOP(Problem):
                 if best_fitness != None:
                     best_fitness = instance.sum() - best_fitness
                     print(f"Reading best-known fitness {best_fitness} from {opt_filename}")
-                    
+
             return LOP(n, instance, best_fitness = best_fitness, instance_name = filename)
 
     # Methods

### 2020-11-14 1-launch_exps.sh
more new results

@@ -108,7 +108,7 @@ for instance in $INSTANCES; do
 	$LAUNCHER cego-$counter-r$run ./target-runner-cego.py cego $counter $run $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $budget --output $RESULTS/cego-r$run
 
 	### Uncomment for running UMM
-	$LAUNCHER umm-$counter-$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --output $RESULTS/umm-r$run
+	$LAUNCHER umm-$counter-r$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --output $RESULTS/umm-r$run
         
     done
 done
@@ -127,7 +127,7 @@ for instance in $INSTANCES; do
 	    mkdir -p $RESULTS
 	    for run in $(seq 1 $nruns); do
 		### Uncomment for running UMM
-		$LAUNCHER umm-$counter-$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --output $RESULTS/umm-r$run
+		$LAUNCHER umm-$counter-r$run ./target-runner-umm.py umm $counter $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --output $RESULTS/umm-r$run
 	    done
 	done
     done

### 2020-11-14 lop.py
Move best_knowns.csv -> lop/best_knowns.csv 2-An intro to UMM.ipynb: more cleanups

@@ -67,7 +67,7 @@ class LOP(Problem):
                    instance_name = f"LOP-synthetic,seed={seed},n={n},m={m},phi={phi}")
 
     @classmethod
-    def read_instance(cls, filename, opt_filename = "./best_knowns.csv"):
+    def read_instance(cls, filename, opt_filename = "./lop/best_knowns.csv"):
         if "synthetic" in filename:
             seed, n, m, phi = re.search("seed=([0-9]+),n=([0-9]+),m=([0-9]+),phi=([^ ]+)", filename).group(1,2,3,4)
             print(f"Generating synthetic LOP instance with seed={seed} n={n} m={m} phi={phi}")

### 2020-11-14 paper/main.tex
acks
more writing

@@
 
-\newcommand{\supplement}{~(\href{http://doi.org/10.5281/zenodo.4167148}{doi:10.5281/zenodo.4167148})}
+\newcommand{\supplement}{\href{http://doi.org/10.5281/zenodo.4167148}{doi:10.5281/zenodo.4167148}}
 
@@
 \subsection{Distributions over permutations}
-Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$.
+Distributions over permutations~\cite{CriFliVer1996rankings} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$.
 One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
@@
 Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}.
-An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters for a sample distributed according to the MM~\cite{CarProSha2013votes}; and (iii) and also when no distribution is assumed, Borda is an approximation to Kemeny~\cite{CopFleRur2010ordering}.
+An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters for a sample distributed according to the MM~\cite{CarProSha2013votes}; and (iii) when no distribution is assumed, Borda is an approximation to Kemeny~\cite{CopFleRur2010ordering}.
 
@@
 Then, we generate matrix $a$ as $a_{i,j} = \sum_{\sigma\in S}\sum_{i<j}\mathbf{I}\{\sigma(i)<\sigma(j)\}$, where $\mathbf{I}\colon S \times S \to \{0,1\}$ is the indicator function.
-In principle, these should be best-case instances for UMM since their fitness landscape closely matches the Mallows model.\MANUEL{Ekhine: Could you revise this text?
+In principle, these should be best-case instances for UMM since their fitness landscape closely matches the Mallows model.%\MANUEL{Ekhine: Could you revise this text?
 I wanted to explain why we use synthetic datasets.}
@@
 We terminated runs that took longer than 6 hours and those combinations are shown as white.
-Results for all the datasets can be found in the supplemental material\supplement.
+Results for all the datasets can be found in the supplemental material~(\supplement).
 
@@
 with ease. As in the synthetic LOP instances, we again observe that CEGO does
-improve faster than UMM initially, yet CEGO appears to
-get stuck around evaluation 250,  and UMM overtakes it around evaluation 300. We believe the reason is
-the same as before, that is, the CEGO model is not able to adequately reflect the
-ruggedness of the actual landscape and, hence, the best solutions found by the
-GA searching the model are not improved solutions for the actual problem. Interestingly, \texttt{kra32} is rectangular
-whereas \texttt{tho32} contains rectangular distances and \texttt{nug12} and \texttt{nug30} contain Manhattan distances of rectangular grids.\MANUEL{so they all have the same distance type!!!
-What is the difference then?
-Ask Thomas?}
+improve faster than UMM initially, yet CEGO appears to get stuck around
+evaluation 250, and UMM overtakes it around evaluation 300. We believe the
+reason is the same as before, that is, the CEGO model is not able to adequately
+reflect the ruggedness of the actual landscape and, hence, the best solutions
+found by the GA searching the model are not improved solutions for the actual
+problem. Interestingly, there are well-known differences between
+\texttt{kra32}, on one hand, and \texttt{nug12}, \texttt{nug30} and
+\texttt{tho32}, on the other~\citep{AnsBriGou2002qap}. However, we cannot at
+this moment point out the root cause of this different behavior.
 
@@
 black-box combinatorial problems on permutation landscapes and when the budget
-of fitness evaluations is severely limited (up to $1000$ or less).  UMM 
-\ekhine{}{has shown to perform statistically better or equal than CEGO, the state of the art algorithm, in the same instances as in the original paper of CEGO and some new instances. UMM}
-is specially well-suited for budget-limited scenarios that are still
-time-sensitive, e.g., when the overhead incurred by the optimizer should be no
-more than a few minutes added to each fitness evaluation. As shown by our
-results, a Bayesian optimizer such as CEGO may sometimes require more than hour
-\MANUEL{UMM is able to model the LOP landscape at least as well as CEGO as shown by the plots where UMM shows a linear decrease in fitness.}
-\EKHINE{me parece mas de CEGO q UMM y parece q UMM solo tiene de ventaja ir rápido.
-Primero destacaría lo bueno de UMM.
-Creo q es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos  y luego que addemás es más rápido}
-\MANUEL{Pero UMM es peor en rec13 y nug12, y en LOP es mejor porque CEGO no llega a 400, es decir, porque UMM es más rápido.
-No podemos saber qué pasaría si CEGO llegara a 400.
-Podríamos medir lo que pasa en el momento en que matamos a CEGO y dar también ese valor entonces la diferencia en velocidad no importaría.
-}
-per fitness evaluation. The time required to train the surrogate-model of CEGO
-also sharply increases with permutation length, hence, UMM would be a
-computationally feasible alternative for relatively large problem sizes.
+of fitness evaluations is severely limited (here, up to $400$ evaluations).
+UMM \ekhine{}{has shown to perform statistically better or equal than CEGO, the
+  state of the art algorithm, in the same instances as in the original paper of
+  CEGO and some new instances. UMM} is specially well-suited for budget-limited
+scenarios that are still time-sensitive, e.g., when the overhead incurred by
+the optimizer should be no more than a few minutes added to each fitness
+evaluation. As shown by our results, a Bayesian optimizer such as CEGO may
+sometimes require more than hour \MANUEL{UMM is able to model the LOP landscape
+  at least as well as CEGO as shown by the plots where UMM shows a linear
+  decrease in fitness.}  \EKHINE{me parece mas de CEGO q UMM y parece q UMM
+  solo tiene de ventaja ir rápido. Primero destacaría lo bueno de UMM. Creo q
+  es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos y
+  luego que addemás es más rápido} \MANUEL{Pero UMM es peor en rec13 y nug12, y
+  en LOP es mejor porque CEGO no llega a 400, es decir, porque UMM es más
+  rápido. No podemos saber qué pasaría si CEGO llegara a 400. Podríamos medir
+  lo que pasa en el momento en que matamos a CEGO y dar también ese valor
+  entonces la diferencia en velocidad no importaría. } per fitness
+evaluation. The time required to train the surrogate-model of CEGO also sharply
+increases with permutation length, hence, UMM would be a computationally
+feasible alternative for relatively large problem sizes.
 
@@
 % \EKHINE{hhmmm lo hacemos mejor o igual}
-The results presented here
-show that, despite the intrinsic difficulty of this scenario, there are still
-significant advances yet to be made. Thus, we hope that this paper will
-motivate further research. % on such problems.
-
+The results presented here show that, despite the intrinsic difficulty of the
+expensive black-box combinatorial scenario, there are still significant
+advances yet to be made. Thus, we hope that this paper will motivate further
+research. % on such problems.
+\MANUEL{I'm thinking about finishing here and commenting out the following paragraph}
 A somewhat surprising result was that the behavior of UMM (and the same for
@@
 \begin{smaller}
+  \paragraph*{Reproducibility.} Source code, datasets and scripts necessary to reproduce the results are available at \supplement.
+  \vspace*{-1em}
 \paragraph*{Acknowledgements.}
@@
 %
+% MANUEL: We should also thank the COST ACTION: Ask Carlos?
 Thanks to Hao Wang (Leiden University) for pointing us to the arguments of
-\citet{EriPeaGar2019scalable}.
+\citet{EriPeaGar2019scalable} and to Thomas Stützle (Université libre de Bruxelles) for suggesting works
+discussing differences between QAPLIB instances.
 \end{smaller}
@@
 \bibliographystyle{splncs04nat}
-\bibliography{optbib/abbrev,optbib/authors,optbib/journals,optbib/biblio,optbib/crossref,./mendeley}
+\bibliography{optbib/abbrev,optbib/authors,optbib/journals,optbib/biblio,optbib/crossref}%,./mendeley}
 

### 2020-11-15 runner.py
* runner.py: Make sure it compresses.

@@ -58,6 +58,6 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
     df['instance'] = instance.instance_name
     df['Solver'] = algo_name
     if out_filename is not None:
-        df.to_csv(out_filename + '.csv.gz', index=False)
-        df.to_pickle(out_filename + '.pkl.gz')
+        df.to_csv(out_filename + '.csv.gz', index=False, compression = 'gz')
+        df.to_pickle(out_filename + '.pkl.gz', compression = 'gz')
     return df

### 2020-11-15 paper/main.tex
Figures

@@
 The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.
-The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance (Eq.~\eqref{eq:kendall_decomp}) as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
+The normalization constant $\psi$ can be easily computed for the Kendall's-$\tau$ distance (Eq.~\eqref{eq:kendall_decomp}) as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
 
@@
 EGO, with a distance measure appropriate to combinatorial
-landscapes~\citep{ZaeStoBar2014:ppsn}, such as Kendall's $\tau$ distance for
+landscapes~\citep{ZaeStoBar2014:ppsn}, such as Kendall's-$\tau$ distance for
 permutations (Eq.~\ref{eq:kendall_decomp}). In CEGO, the surrogate model is
@@
 %
-For simplicity, we use Kendall's $\tau$
+For simplicity, we use Kendall's-$\tau$
 distance~(Eq.~\ref{eq:kendall_decomp}) in all experiments. Nevertheless,
@@
 UMM \ekhine{}{has shown to perform statistically better or equal than CEGO, the
-  state of the art algorithm, in the same instances as in the original paper of
+  state of the art algorithm for black-box expensive optimization, in the same instances as in the original paper of
   CEGO and some new instances. UMM} is specially well-suited for budget-limited
@@
   entonces la diferencia en velocidad no importaría. } per fitness
-evaluation. The time required to train the surrogate-model of CEGO also sharply
-increases with permutation length, hence, UMM would be a computationally
+evaluation. The time required to train the surrogate-model of \ekhine{esta establecida la complexity de CEGO???}{UMM increases linearly with the permutation length $n$, while in }
+CEGO  sharply
+increases with $n$, hence, UMM would be a computationally
 feasible alternative for relatively large problem sizes.
@@
 metric used, but further analysis is needed to corroborate this
-conjecture. Here, we have focused on Kendall's $\tau$ distance, however, it is
+conjecture. Here, we have focused on Kendall's-$\tau$ distance, however, it is
 possible to extend UMM to other distance metrics, which will allow us to

### 2020-11-15 lop.py

@@ -99,6 +99,6 @@ class LOP(Problem):
         # In case it is not numpy array.
         x = np.asarray(x, dtype=int)
         xinverse = np.argsort(x)#### OJOOOOO se evalua la inversa
-        # Sum of the upper triangle. We have to minimize this.
+        # Sum of the lower triangle. We have to minimize this.
         f = np.tril(self.instance[np.ix_(xinverse, xinverse)]).sum()
         return f

### 2020-11-15 mallows_kendall.py
save normalization as Fitness_norm

@@ -15,21 +15,23 @@ def find_phi_n(n, bins): #NO
     ed, phi_ed = [], []
     #expected dist en la uniforma para este n
     # FIXME: Use max_dist(n) / 2
+    # FIXME: Is this an integer? should we divide by 2?
     ed_uniform = (n*(n-1)/2)/2 # dist max div 2. AKA
     # ed_uniform = np.mean([expected_dist_MM(n,-0.01), expected_dist_MM(n,0.01)])
     # if k is not None : ed_uniform = np.mean([expected_V(n,theta=None, phi=0.99999,k=k).sum(), expected_V(n,theta=None, phi=1.00000001,k=k).sum()])
     # print(ed_uniform, "find")
-    for dmin in np.linspace(0,ed_uniform-1,bins):
+    for dmin in np.linspace(0, ed_uniform - 1, bins):
         ed.append(dmin)
         phi_ed.append(find_phi(n, dmin, dmin+1))
     return ed, phi_ed
 
 def find_phi(n, dmin, dmax): #NO
-    imin, imax = np.float64(0),np.float64(1)
+    # 0.0, 1.0
+    imin, imax = np.float64(0), np.float64(1)
     iterat = 0
     while iterat < 500:
-        med = imin + (imax-imin)/2
-        d = expected_dist_MM(n,phi2theta(med))#mk.phi2theta(med)
+        med = imin + (imax - imin) / 2
+        d = expected_dist_MM(n, phi2theta(med))#mk.phi2theta(med)
         #print(imin, imax, med, d,imin==imax)
         if d < dmax and d > dmin: return med
         elif d < dmin : imin = med
@@ -71,16 +73,23 @@ def borda_partial(rankings):
     # Double argsort should cancel itself?
     borda = np.argsort(np.argsort(np.nanmean(rankings, axis=0))).astype(float)
     mask = np.isnan(rankings).all(axis=0)
-    borda[mask]=np.nan
+    borda[mask] = np.nan
     return borda
 
-
 def check_theta_phi(theta, phi):
-    if not ((phi is None) ^ (theta is None)): print("KAKA, pon valores")
-    if phi is None and type(theta)!=list: phi = theta2phi(theta)
-    if theta is None and type(phi)!=list: theta = phi2theta(phi)
-    if phi is None and type(theta)==list: phi = [theta2phi(t) for t in theta]
[…]

### 2020-11-15 runner.py

@@ -49,9 +49,8 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
 
     timer = Timer()
     df = algo(instance, seed, **algo_params)
-
     if instance.best_fitness is not None and instance.worst_fitness is not None:
-        df['Fitness'] = (df.Fitness - instance.best_fitness) / (instance.worst_fitness - instance.best_fitness)
+        df['Fitness_norm'] = (df.Fitness - instance.best_fitness) / (instance.worst_fitness - instance.best_fitness)
     df['Function evaluations'] = np.arange(1, len(df['Fitness'])+1)
     df['run_time'] = timer.elapsed()
     df['Problem'] = instance.problem_name

### 2020-11-15 lop.py
fix lop.py conflicts

@@ -102,12 +102,7 @@ class LOP(Problem):
     def fitness_nosave(self, x):
         # In case it is not numpy array.
         x = np.asarray(x, dtype=int)
-<<<<<<< HEAD
-        xinverse = np.argsort(x)#### OJOOOOO se evalua la inversa
-        # Sum of the lower triangle. We have to minimize this.
-=======
         xinverse = np.argsort(x) # Important: We evaluate the inverse permutation
-        # Sum of the upper triangle. We have to minimize this.
->>>>>>> 956903d23a481193e4430038eadcd1922190f4e9
+        # Sum of the lower triangle. We have to minimize this.
         f = np.tril(self.instance[np.ix_(xinverse, xinverse)]).sum()
         return f

### 2020-11-16 paper/main.tex
Paper update

@@
 \subsubsection{Sampling step}\label{sec:sampling}
-%
+In this step, the algorithm samples from the distribution the new data point to be evaluated. 
 Sampling from $MM(\sigma_0,\theta)$  can be done efficiently given the parameters  $\sigma_0$ and $\theta$, as studied by \citet{IruCalLoz2016permallows}, where the computational complexity is $O(n\log n)$.
@@
 Parameter $\theta$ is set automatically by the algorithm to control the diversification and intensification trade off.
-At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$ (half of the distance of the uniform distribution), and then the expected distance linearly decreases until the last iteration where $\theta$ is set such that $\Edist = 1$.
+At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$ (half of the expected distance for the uniform distribution), and then the expected distance linearly decreases until the last iteration where $\theta$ is set such that $\Edist = 1$.
 Despite it is not possible to isolate $\theta$ in Eq.~\eqref{eq:expectation}, its monotonicity allows the use of bisection methods for efficiently finding an approximation.
@@
 However, the setting of a decreasing variance   controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
-Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.\EKHINE{Explain budgetMM.
-we sample several permutations and choose one only!}
+Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.
+The algorithm also implements the following strategy to avoid sampling the same points more than once, which will be more likely at the explotation phase.
+The strategy is as follows.
+Instead of sampling just one permutation from the MM, the algorithm generates $budgetMM$ permutations, $\Sigma$.
+Then it selects  for evaluation the farest to the sample, i.e.,  $\sigma\in \Sigma$ such that $\sigma = \arg\max\arg\min d(\sigma, \pi)$ for every $\pi\in S$.
 
@@
 the optimizer should be no more than a few minutes added to each fitness
-evaluation. As shown by our results, a Bayesian optimizer such as CEGO may
-sometimes require more than hour \MANUEL{UMM is able to model the LOP landscape
+evaluation.
+
+\MANUEL{UMM is able to model the LOP landscape
   at least as well as CEGO as shown by the plots where UMM shows a linear
-  decrease in fitness.}  \EKHINE{me parece mas de CEGO q UMM y parece q UMM
+  decrease in fitness.}  
+  \EKHINE{me parece mas de CEGO q UMM y parece q UMM
   solo tiene de ventaja ir rápido. Primero destacaría lo bueno de UMM. Creo q
   es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos y
-  luego que addemás es más rápido} \MANUEL{Pero UMM es peor en rec13 y nug12, y
+  luego que addemás es más rápido} 
+  \MANUEL{Pero UMM es peor en rec13 y nug12, y
   en LOP es mejor porque CEGO no llega a 400, es decir, porque UMM es más
@@
   lo que pasa en el momento en que matamos a CEGO y dar también ese valor
-  entonces la diferencia en velocidad no importaría. } per fitness
-evaluation. The time required to train the surrogate-model of \ekhine{esta establecida la complexity de CEGO???}{UMM increases linearly with the permutation length $n$, while in }
-CEGO  sharply
-increases with $n$, hence, UMM would be a computationally
+  entonces la diferencia en velocidad no importaría. } 
+The time required to train the surrogate-model of \ekhine{esta establecida la complexity de CEGO???}{UMM increases linearly with the permutation length $n$, while in }
+CEGO  it sharply
+increases with $n$, requiring 
+sometimes more than hour per fitness evaluation as shown in the experimental section. 
+Hence, UMM is a computationally
 feasible alternative for relatively large problem sizes.

### 2020-11-17 paper/main.tex
we only use one computing system now

@@
 \subsubsection{Computing environment.}
-Due to the computational effort required, we used several computing systems to
-run experiments. All UMM experiments were either on run on Intel Xeon
-\emph{Broadwell} E5-2683v4 CPUs at 2.10\,GHz or Intel Xeon \emph{Haswell} E5-2680v3
-CPUs at 2.50\,GHz, both of them with 32\,GB RAM. 
- %
-Experiments with
-CEGO used a more powerful system: Intel Xeon \emph{Ivybridge} E5-2650v2 CPUs at
-2.60\,GHz, 64\,GB RAM. % running CentOS Linux release 7 (Core).
-The operating system was CentOS Linux release 7 (Core).
-%processor : 23
-%vendor_id : GenuineIntel
-%cpu family : 6
-%model : 63
-%model name : Intel(R) Xeon(R) CPU E5-2680 v3 @ 2.50GHz
-%stepping : 2
-%microcode : 0x43
-%cpu MHz : 1199.951
-%cache size : 30720 KB
-%physical id : 1
-%siblings : 12
-%core id : 13
-%cpu cores : 12
-%apicid : 58
-%initial apicid : 58
-%fpu : yes
-%fpu_exception : yes
-%cpuid level : 15
-%wp : yes
-%flags : fpu vme de pse tsc msr pae mce cx8 apic sep mtrr pge mca cmov pat pse36 clflush dts acpi mmx fxsr sse sse2 ss ht tm pbe syscall nx pdpe1gb rdtscp lm constant_tsc arch_perfmon pebs bts rep_good nopl xtopology nonstop_tsc aperfmperf eagerfpu pni pclmulqdq dtes64 monitor ds_cpl vmx smx est tm2 ssse3 sdbg fma cx16 xtpr pdcm pcid dca sse4_1 sse4_2 x2apic movbe popcnt tsc_deadline_timer aes xsave avx f16c rdrand lahf_lm abm epb intel_ppin ssbd ibrs ibpb stibp tpr_shadow vnmi flexpriority ept vpid fsgsbase tsc_adjust bmi1 avx2 smep bmi2 erms invpcid cqm xsaveopt cqm_llc cqm_occup_llc dtherm ida arat pln pts md_clear spec_ctrl intel_stibp flush_l1d
-%bogomips : 5005.34
-%clflush size : 64
-%cache_alignment : 64
-%address sizes : 46 bits physical, 48 bits virtual
-%
-%The node n015 has 32 processors. This is the info of the 32st processor:
-%processor : 31
-%vendor_id : GenuineIntel
-%cpu family : 6
-%model : 79
-%model name : Intel(R) Xeon(R) CPU E5-2683 v4 @ 2.10GHz
-%stepping : 1
-%microcode : 0xb000022
-%cpu MHz : 2442.608
-%cache size : 40960 KB
-%physical id : 1
-%siblings : 16
-%core id : 15
-%cpu cores : 16
-%apicid : 62
-%initial apicid : 62
-%fpu : yes
-%fpu_exception : yes
-%cpuid level : 20
-%wp : yes
-%flags : fpu vme de pse tsc msr pae mce cx8 apic sep mtrr pge mca cmov pat pse36 clflush dts acpi mmx fxsr sse sse2 ss ht tm pbe syscall nx pdpe1gb rdtscp lm constant_tsc arch_perfmon pebs bts rep_good nopl xtopology nonstop_tsc aperfmperf eagerfpu pni pclmulqdq dtes64 monitor ds_cpl vmx smx est tm2 ssse3 sdbg fma cx16 xtpr pdcm pcid dca sse4_1 sse4_2 x2apic movbe popcnt tsc_deadline_timer aes xsave avx f16c rdrand lahf_lm abm 3dnowprefetch epb cat_l3 cdp_l3 intel_ppin intel_pt tpr_shadow vnmi flexpriority ept vpid fsgsbase tsc_adjust bmi1 hle avx2 smep bmi2 erms invpcid rtm cqm rdt_a rdseed adx smap xsaveopt cqm_llc cqm_occup_llc cqm_mbm_total cqm_mbm_local dtherm ida arat pln pts
-%bogomips : 4207.67
-%clflush size : 64
-%cache_alignment : 64
-%address sizes : 46 bits physical, 48 bits virtual
+% Due to the computational effort required, we used several computing systems to
+% run experiments.
+% All UMM experiments were either on run on Intel Xeon
+% \emph{Broadwell} E5-2683v4 CPUs at 2.10\,GHz or Intel Xeon \emph{Haswell} E5-2680v3
+% CPUs at 2.50\,GHz, both of them with 32\,GB RAM. 
+ %
+Experiments were run on Intel Xeon \emph{Ivybridge} E5-2650v2 CPUs at
+2.60\,GHz, 64\,GB RAM running CentOS Linux release 7 (Core).
+%The operating system was CentOS Linux release 7 (Core).
 

### 2020-11-17 1-launch_exps.sh
new results

@@ -42,7 +42,7 @@ launch_local() {
 
 # Generate LOP synthetic
 gen_lop_synthetic() {
-    INSTANCES=$@
+    INSTANCES=""
     LOP_n=20
     LOP_m=200
     LOP_seed=123456
@@ -60,6 +60,11 @@ gen_lop_synthetic() {
     echo $INSTANCES
 }
 
+nruns=10
+
+LAUNCHER=qsub_job
+#LAUNCHER=launch_local
+
 ## For QAP, PFSP instances
 INSTANCES="\
   qap/kra30a.dat \
@@ -74,17 +79,14 @@ INSTANCES="\
  pfsp/rec31.txt \
 "
 
-###### For synthetic LOP instances
-INSTANCES=$(gen_lop_synthetic "")
-
 ###### For LOPLIB instances
 #INSTANCES="$(tr '\n' ' ' < loplib-instances.txt)"
 
-budget=400
-nruns=10
+###### For synthetic LOP instances, we generate different instances and one run per instance.
+INSTANCES=$(gen_lop_synthetic)
 
-LAUNCHER=qsub_job
-#LAUNCHER=launch_local
+
+budget=400
 
 cego_m_ini=10
 budgetGA=3 # Actually, 10**budgetGA
@@ -100,11 +102,11 @@ for instance in $INSTANCES; do
     RESULTS="results/$instance"
     mkdir -p $RESULTS
     for run in $(seq 1 $nruns); do
-	### Uncomment for running CEGO
+       	### Uncomment for running CEGO
 	$LAUNCHER cego-$counter-r$run ./target-runner-cego.py cego $counter-r$run-$$ $run $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $budget --output $RESULTS/cego-r$run
 
 	### Uncomment for running UMM
-	#$LAUNCHER umm-$counter-r$run ./target-runner-umm.py umm $counter-r$run-$$ $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --output $RESULTS/umm-r$run
+	$LAUNCHER umm-$counter-r$run ./target-runner-umm.py umm $counter-r$run-$$ $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --output $RESULTS/umm-r$run
         
     done
 done

### 2020-11-17 runner.py

@@ -57,6 +57,6 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
     df['instance'] = instance.instance_name
     df['Solver'] = algo_name
     if out_filename is not None:
-        df.to_csv(out_filename + '.csv.gz', index=False)
-        df.to_pickle(out_filename + '.pkl.gz')
+        df.to_csv(out_filename + '.csv.gz', index=False, compression = "gzip")
+        df.to_pickle(out_filename + '.pkl.gz', compression = "gzip")
     return df

### 2020-11-17 paper/main.tex
update results on paper

@@
 In the case of the QAP and PFSP, we consider the same instances as
-\citet{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, i.e.,
-\texttt{nug12}, \texttt{nug30}, \texttt{tho30} and \texttt{kra32} for the QAP,
-with $n$ (the permutation length) equal to the number in the instance name, and \texttt{reC05},
-\texttt{reC13}, \texttt{reC19}, \texttt{reC31} for the PFSP, with
-$n \in \{20, 20, 30, 50\}$, respectively.
+\citet{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, i.e., \texttt{nug12},
+\texttt{nug30}, \texttt{tho30} and \texttt{kra32} for the QAP, with $n$ (the
+permutation length) equal to the number in the instance name, and
+\texttt{reC05}, \texttt{reC13} and \texttt{reC19} for the PFSP, with
+$n \in \{20, 20, 30\}$, respectively. For reasons that will become obvious
+later, we also consider QAP instances \texttt{kra30a} and \texttt{kra30b}.
 
@@
 the solution evaluated at each step of each run and we average those values
-over 10 runs with different random seeds on the same problem instance.  As an example, each line in Fig.~\ref{fig:lop_synth} 
-shows the mean fitness of the solutions evaluated by either UMM or CEGO on one synthetic LOP
-instance. The shaded area shows one standard deviation around the
-mean. Ideally, each new evaluation will monotonically improve in fitness, since each solution evaluated helps to refine the model and leads to a
-better solution being evaluated in the next step.  However, it is often the case that 
-new solutions
-are often worse than their predecessors. Nevertheless, in this example, there
-is a clear overall improvement as more evaluations are performed. The minimum
-values achieved in the plot up to a particular number of function evaluations
-gives an estimation of the mean fitness of the best-so-far solution up to that
-point in the run. The algorithms do not return the final solution shown but
-rather the best one (minimum fitness) of each run.
+over 10 runs with different random seeds on the same problem instance.  As an
+example, each line in Fig.~\ref{fig:lop_synth} shows the mean fitness of the
+solutions evaluated by either UMM or CEGO on one synthetic LOP instance. The
+shaded area shows one standard deviation around the mean. Ideally, each new
+evaluation will monotonically improve in fitness, since each solution evaluated
+helps to refine the model and leads to a better solution being evaluated in the
+next step.  However, it is often the case that new solutions are often worse
+than their predecessors. Nevertheless, in this example, there is a clear
+overall improvement as more evaluations are performed. The minimum values
+achieved in the plot up to a particular number of function evaluations gives an
+estimation of the mean fitness of the best-so-far solution up to that point in
+the run. The algorithms do not return the final solution shown but rather the
+best one (minimum fitness) of each run.
 
@@
 behaviour as in the PFSP for all instances tested, except for
-\texttt{kra32}. As shown in Fig~\ref{fig:kra32}, for \texttt{kra32} both
+\texttt{kra} instances. As shown in Fig~\ref{fig:kra32}, for \texttt{kra32} both
 algorithms are able to improve over the number of evaluations, although not
@@
 problem. Interestingly, there are well-known differences between
-\texttt{kra32}, on one hand, and \texttt{nug12}, \texttt{nug30} and
+\texttt{kra30a}, \texttt{kra30b}, \texttt{kra32}, on one hand, and \texttt{nug12}, \texttt{nug30} and
 \texttt{tho32}, on the other~\citep{AnsBriGou2002qap}. However, we cannot at
@@
 evaluations. In contrast, UMM was always able to reach 400 evaluations in less
-than 16 hours. If UMM was allowed to reach the 144 hours limit, the differences
+than 12.5 hours. If UMM was allowed to reach the 144 hours limit, the differences
 would be even stronger.
@@
 performing a random search. Similar observations can be made for the QAP, where
-only on the smallest instance \texttt{nug12}, there is a significant different
+only on the smallest instance \texttt{nug12}, there is a significant difference
 in favour of CEGO.
@@
 If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 144 hours, it was terminated earlier.
-This was the case for all LOP instances shown.\manuel{}{Don't we have results for rec31?}\label{tab:results}}
+This was the case for all LOP instances shown.\label{tab:results}}
  \resizebox{\textwidth}{!}{%
@@
             \bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule
-    LOP     & N-t1d100.01      & 78696.6                 & (811.4)    & 76119.6   & (915.4)                & $[$1763.5,    & 3390.5$]$     & 111.8& 72.5 & 3.2  \\
-     & N-t1d100.02      & 79686.5                 & (367.9)    & 76827.7   & (891.0)                & $[$2194.5,    & 3523.1$]$     & 110.7& 72.9 & 3.2  \\
-     & N-t1d150.01      & 187605.5                & (1906.9)   & 179508.3  & (1647.9)               & $[$6420.3,    & 9774.1$]$     & 73.9 & 72.7 & 6.5  \\
-     & N-t1d150.02      & 184320.9                & (973.5)    & 177592.8  & (1196.2)               & $[$5700.5,    & 7755.7$]$     & 74.1 & 73.2 & 6.5  \\
-     & N-t1d200.01      & 335391.1                & (2289.0)   & 323513.6  & (1371.2)               & $[$10076.1,   & 13678.9$]$    & 55.6 & 73.2 & 10.7 \\
-     & N-t1d200.02      & 333149.5                & (2497.2)   & 319675.8  & (2242.6)               & $[$11242.0,   & 15705.4$]$    & 56.5 & 73.7 & 10.6 \\
-     & N-t2d150.01      & 24682.7                 & (976.6)    & 15297.7   & (563.5)                & $[$8622.2,    & 10147.8$]$    & 74.6 & 72.9 & 7.0  \\
-     & N-t2d150.02      & 24250.8                 & (1354.9)   & 14724.9   & (667.5)                & $[$8495.1,    & 10556.7$]$    & 74.8 & 72.8 & 7.1  \\
-     & N-t2d200.01      & 57098.6                 & (4452.8)   & 33465.2   & (2001.0)               & $[$20284.5,   & 26982.3$]$    & 56.3 & 73.3 & 11.2 \\
-     & N-t2d200.02      & 54895.6                 & (4589.7)   & 35901.8   & (2077.5)               & $[$15539.1,   & 22448.5$]$    & 55.6 & 73.5 & 11.2 \\
-     & N-p50-01         & 21431.2                 & (299.5)    & 21027.6   & (711.8)                & $[$-128.1,    & 935.3$]$      & 219.9& 72.3 & 0.9  \\
-     & N-p50-02         & 22021.1                 & (275.6)    & 21613.6   & (596.4)                & $[$-42.5,     & 857.5$]$      & 217.3& 72.3 & 0.9  \\
-     & N-atp111         & 693.1                   & (10.3)     & 594.0     & (15.4)                 & $[$86.7,      & 111.5$]$      & 98.7 & 72.5 & 4.1  \\
-     & N-atp134         & 880.5                   & (18.6)     & 756.3     & (21.3)                 & $[$105.4,     & 143.0$]$      & 83.0 & 72.8 & 5.7  \\
-     & N-be75eec\_150   & 1666193.1               & (72310.8)  & 1415208.5 & (58593.5)              & $[$188960.6,  & 313008.6$]$   & 74.0 & 72.9 & 6.9  \\
-     & N-be75eec\_250   & 5214413.0               & (166701.4) & 4431271.0 & (109756.7)             & $[$649040.8,  & 917243.2$]$   & 44.9 & 73.7 & 15.5 \\
-     & N-be75np\_150    & 3742444.0               & (88500.2)  & 3204878.3 & (80704.1)              & $[$457944.2,  & 617187.2$]$   & 74.2 & 72.9 & 6.9  \\
-     & N-be75np\_250    & 10887427.8              & (159642.5) & 9544754.9 & (260523.9)             & $[$1136635.8, & 1548710.0$]$  & 45.3 & 74.0 & 15.5 \\\midrule
-   PFSP     & rec05        & 1317.9                  & (25.1)     & 1330.7    & (17.9)                 & $[$-33.4,     & 7.8$]$        & 400.0& 38.9 & 0.1  \\
-       & rec13        & 2128.2                  & (25.0)     & 2158.7    & (29.5)                 & $[$-56.2,     & -4.8$]$       & 400.0& 38.9 & 0.1  \\
-       & rec19        & 2398.6                  & (19.8)     & 2405.1    & (12.6)                 & $[$-22.3,     & 9.3$]$        & 400.0& 82.1 & 0.3  \\\midrule
-    QAP  & nug12        & 624.2                   & (26.3)     & 660.4     & (17.3)                 & $[$-57.3,     & -15.1$]$      & 400.0& 14.8 & 0.0  \\
-         & nug30        & 7486.8                  & (40.4)     & 7474.0    & (83.9)                 & $[$-50.8,     & 76.4$]$       & 400.0& 84.3 & 0.3  \\
-        & tho30        & 192416.0                & (2584.2)   & 191482.0  & (4419.7)               & $[$-2527.0,   & 4395.0$]$     & 400.0& 85.1 & 0.3  \\
-         & kra32        & 115907.0                & (4663.4)   & 116320.0  & (4673.5)               & $[$-4799.3,   & 3973.3$]$     & 400.0& 95.8 & 0.3  \\
+LOP
+& N-p50-01       & 21034.3             & (183.9)            & 21161.8            & (657.4)           & [-606.0,             & 351.0]              & 305.8        & 144.6 & 0.5  \\
+& N-p50-02       & 21767.3             & (254.9)            & 21746.6            & (645.4)           & [-458.6,             & 500.0]              & 315.1        & 144.5 & 0.5  \\
+& N-t1d100.01    & 78012.0             & (432.1)            & 76499.3            & (1099.7)          & [696.4,              & 2329.0]             & 158.3        & 145.1 & 1.9  \\
+& N-t1d100.02    & 78834.0             & (477.8)            & 76860.7            & (864.0)           & [1303.8,             & 2642.8]             & 156.0        & 144.7 & 1.8  \\
+& N-atp111       & 668.5               & (9.4)              & 596.6              & (17.0)            & [58.8,               & 85.0]               & 137.9        & 145.1 & 2.3  \\
+& N-atp134       & 856.4               & (12.2)             & 760.1              & (23.4)            & [78.4,               & 114.2]              & 116.5        & 145.0 & 3.3  \\
+& N-t1d150.01    & 186969.0            & (2094.6)           & 179615.6           & (1731.5)          & [5543.3,             & 9163.5]             & 103.9        & 145.7 & 4.4  \\
+& N-t1d150.02    & 183414.7            & (779.2)            & 177524.2           & (1165.6)          & [4949.2,             & 6831.8]             & 103.6        & 145.3 & 4.6  \\
+& N-t2d150.01    & 22207.3             & (1667.3)           & 15136.9            & (745.9)           & [5817.1,             & 8323.7]             & 104.5        & 145.5 & 4.5  \\
+& N-t2d150.02    & 22036.0             & (744.7)            & 14718.4            & (677.5)           & [6648.3,             & 7986.9]             & 103.1        & 145.0 & 4.5  \\
+& N-t1d200.01    & 333804.1            & (1191.9)           & 323501.4           & (1385.7)          & [9086.4,             & 11519.0]            & 77.7         & 145.4 & 7.6  \\
+& N-t1d200.02    & 331730.9            & (2819.7)           & 319786.2           & (2265.6)          & [9533.6,             & 14355.8]            & 78.2         & 145.9 & 7.1  \\
+& N-t2d200.01    & 54043.6             & (1477.6)           & 33389.0            & (1984.0)          & [19001.4,            & 22307.8]            & 78.2         & 146.7 & 7.4  \\
+& N-t2d200.02    & 52787.0             & (4194.2)           & 35589.0            & (2319.6)          & [13948.0,            & 20448.0]            & 78.9         & 145.7 & 7.5  \\
+& N-be75np\_150  & 3669753.9           & (74292.3)          & 3216907.1          & (72239.3)         & [383998.5,           & 521695.1]           & 105.2        & 145.5 & 4.4  \\
+& N-be75np\_250  & 10818828.5          & (127722.3)         & 9582261.7          & (258176.8)        & [1040022.9,          & 1433110.7]          & 62.5         & 146.4 & 12.5 \\
+& N-be75eec\_150 & 1592220.8           & (29945.9)          & 1418578.9          & (56557.7)         & [130141.0,           & 217142.8]           & 103.2        & 145.9 & 4.4  \\
+& N-be75eec\_250 & 5098734.4           & (172874.0)         & 4430408.5          & (110233.0)        & [530352.4,           & 806299.4]           & 63.0         & 146.7 & 12.3 \\\midrule
+PFSP
+& rec05          & 1317.9              & (25.1)             & 1332.5             & (15.6)            & [-34.5,              & 5.3]                & 400.0        & 38.6  & 0.1  \\
+& rec13          & 2128.2              & (25.0)             & 2163.6             & (28.1)            & [-60.4,              & -10.4]              & 400.0        & 38.6  & 0.1  \\
+& rec19          & 2398.6              & (19.8)             & 2405.2             & (12.8)            & [-22.5,              & 9.3]                & 400.0        & 85.2  & 0.2  \\\midrule
+QAP
+& nug12          & 624.2               & (26.3)             & 661.2              & (17.8)            & [-58.3,              & -15.7]              & 400.0        & 14.2  & 0.0  \\
+& kra30b         & 120192.0            & (1778.1)           & 119923.0           & (3459.6)          & [-2379.5,            & 2917.5]             & 400.0        & 85.1  & 0.2  \\
+& tho30          & 192416.0            & (2584.2)           & 191586.4           & (4342.0)          & [-2582.9,            & 4242.1]             & 400.0        & 85.6  & 0.2  \\
+& kra30a         & 119872.0            & (1746.7)           & 119502.0           & (1827.5)          & [-1309.8,            & 2049.8]             & 400.0        & 86.1  & 0.2  \\
+& nug30          & 7486.8              & (40.4)             & 7472.4             & (83.6)            & [-49.0,              & 77.8]               & 400.0        & 86.5  & 0.2  \\
+& kra32          & 115907.0            & (4663.4)           & 116898.0           & (4343.1)          & [-5226.3,            & 3244.3]             & 400.0        & 94.0  & 0.2  \\
 \bottomrule

### 2020-11-17 lop.py
update fit

@@ -104,5 +104,6 @@ class LOP(Problem):
         x = np.asarray(x, dtype=int)
         xinverse = np.argsort(x) # Important: We evaluate the inverse permutation
         # Sum of the lower triangle. We have to minimize this.
-        f = np.tril(self.instance[np.ix_(xinverse, xinverse)]).sum()
+        # f = np.tril(self.instance[np.ix_(xinverse, xinverse)]).sum()
+        f = np.tril(self.instance[np.ix_(x, x)]).sum()
         return f

### 2020-11-17 lop.py
update fit

@@ -62,7 +62,7 @@ class LOP(Problem):
             seed = np.random.randint(1, 123456789, 1)[0]
         np.random.seed(seed)
         instance, best_sol = synthetic_LOP(n, m, phi)
-        worst_sol = best_sol[::-1]
+        worst_sol = np.argsort(np.argsort(bes)[::-1])
         return cls(n, instance, best_sol = best_sol, worst_sol = worst_sol,
                    instance_name = f"LOP-synthetic,seed={seed},n={n},m={m},phi={phi}")
 

### 2020-11-17 lop.py
update fit

@@ -62,7 +62,7 @@ class LOP(Problem):
             seed = np.random.randint(1, 123456789, 1)[0]
         np.random.seed(seed)
         instance, best_sol = synthetic_LOP(n, m, phi)
-        worst_sol = np.argsort(np.argsort(bes)[::-1])
+        worst_sol = np.argsort(np.argsort(best_sol)[::-1])
         return cls(n, instance, best_sol = best_sol, worst_sol = worst_sol,
                    instance_name = f"LOP-synthetic,seed={seed},n={n},m={m},phi={phi}")
 

### 2020-11-17 lop.py
update fit

@@ -63,6 +63,9 @@ class LOP(Problem):
         np.random.seed(seed)
         instance, best_sol = synthetic_LOP(n, m, phi)
         worst_sol = np.argsort(np.argsort(best_sol)[::-1])
+        #OJO, si la fitness NO evalua la inversa, las bests y worst son estas de abajo
+        worst_sol = np.argsort(worst_sol)
+        best_sol = np.argsort(best_sol)
         return cls(n, instance, best_sol = best_sol, worst_sol = worst_sol,
                    instance_name = f"LOP-synthetic,seed={seed},n={n},m={m},phi={phi}")
 

### 2020-11-17 lop.py
update lolib list

@@ -66,7 +66,7 @@ class LOP(Problem):
         np.random.seed(seed)
         instance, best_sol = synthetic_LOP(n, m, phi)
         # The worst is the reverse of the best because the evaluation does not
-        # use the inverse
+        # use the inverse.
         worst_sol = best_sol[::-1]
         return cls(n, instance, best_sol = best_sol, worst_sol = worst_sol,
                    instance_name = f"LOP-synthetic,seed={seed},n={n},m={m},phi={phi}")

### 2020-11-17 umm.py
Avoid NaN

@@ -8,26 +8,40 @@ import pandas as pd
 def binary_search_rho(w, ratio_samples_learn, weight_mass_learn,
                       # 0 <= w_i <= 1, w is sorted increasingly,
                       rho_ini=1, rho_end=0, tol=0.001):
-  w = np.asarray(w)
-  assert np.all(w >= 0.0)
-  assert np.all(w <= 1.0)
-  #if pos is None we take the largest 4th.
-  #find the rho s.t. the largest 25%(ratio_samples) of the weights  (rho**ws) take the 0.9(weight_mass) of the total ws.  rho^w[:pos] = 0.9*rho^w
-  # codes as a recursive binary search in (0,1)
-  pos = int(len(w) * ratio_samples_learn)
-  # print(pos,len(w),ratio_samples_learn)
-  rho_med = rho_ini + (rho_end - rho_ini) / 2
-  acum = np.cumsum(rho_med ** w)
-  a = acum[pos]
-  b = acum[-1]
-  if abs(a/b - weight_mass_learn) < tol:
-      return rho_med
-  if a/b > weight_mass_learn:
-      mid = rho_ini
-      last = rho_med
-  else:
-      mid = rho_med
-      last = rho_end
+  try:
+      w = np.asarray(w)
+      ane = np.ediff1d(w)[np.ediff1d(w)>0].min()/2 #avoid numerical errors caused by having several zeros in w
+      w = w+ane
+      w[0] = 0.0
+      w[w>1] = 1.0
+      assert np.all(w >= 0.0)
+      assert np.all(w <= 1.0)
+
+      #if pos is None we take the largest 4th.
+      #find the rho s.t. the largest 25%(ratio_samples) of the weights  (rho**ws) take the 0.9(weight_mass) of the total ws.  rho^w[:pos] = 0.9*rho^w
+      # codes as a recursive binary search in (0,1)
+      pos = int(len(w) * ratio_samples_learn)
+      # print(pos,len(w),ratio_samples_learn)
+      rho_med = rho_ini + (rho_end - rho_ini) / 2
+      acum = np.cumsum(rho_med ** w)
+      a = acum[pos]
+      b = acum[-1]
+      if abs(a/b - weight_mass_learn) < tol:
+          return rho_med
+      if a/b > weight_mass_learn:
+          mid = rho_ini
+          last = rho_med
+      else:
+          mid = rho_med
+          last = rho_end
+  except:
[…]

### 2020-11-18 1-launch_exps.sh
new results

@@ -42,7 +42,7 @@ launch_local() {
 
 # Generate LOP synthetic
 gen_lop_synthetic() {
-    INSTANCES=""
+    INSTANCES=$@
     LOP_n=20
     LOP_m=200
     LOP_seed=123456
@@ -79,11 +79,11 @@ INSTANCES="\
  pfsp/rec31.txt \
 "
 
-###### For LOPLIB instances
-#INSTANCES="$(tr '\n' ' ' < loplib-instances.txt)"
+###### Synthetic LOP instances
+INSTANCES=$(gen_lop_synthetic "")
 
-###### For synthetic LOP instances, we generate different instances and one run per instance.
-INSTANCES=$(gen_lop_synthetic)
+###### For LOLIB instances
+INSTANCES="$INSTANCES $(tr '\n' ' ' < loplib-instances.txt)"
 
 
 budget=400

### 2020-11-18 umm.py
update NB

@@ -10,10 +10,6 @@ def binary_search_rho(w, ratio_samples_learn, weight_mass_learn,
                       rho_ini=1, rho_end=0, tol=0.001):
   try:
       w = np.asarray(w)
-      ane = np.ediff1d(w)[np.ediff1d(w)>0].min()/2 #avoid numerical errors caused by having several zeros in w
-      w = w+ane
-      w[0] = 0.0
-      w[w>1] = 1.0
       assert np.all(w >= 0.0)
       assert np.all(w <= 1.0)
 
@@ -35,6 +31,7 @@ def binary_search_rho(w, ratio_samples_learn, weight_mass_learn,
           mid = rho_med
           last = rho_end
   except:
+       print(w)
        pos = int(len(w) * ratio_samples_learn)
        # print(pos,len(w),ratio_samples_learn)
        rho_med = rho_ini + (rho_end - rho_ini) / 2

### 2020-11-18 lop.py
revert bc1aa864a3a4fb376f3caed56d4347178295c36b

@@ -66,7 +66,7 @@ class LOP(Problem):
         np.random.seed(seed)
         instance, best_sol = synthetic_LOP(n, m, phi)
         # The worst is the reverse of the best because the evaluation does not
-        # use the inverse.
+        # use the inverse
         worst_sol = best_sol[::-1]
         return cls(n, instance, best_sol = best_sol, worst_sol = worst_sol,
                    instance_name = f"LOP-synthetic,seed={seed},n={n},m={m},phi={phi}")

### 2020-11-18 lop.py
revert back to inverse LOP

@@ -49,10 +49,7 @@ def synthetic_LOP(n, m, phi):
         for j in range(i+1, n):
             instance[i, j] = (s[:, i] < s[:, j]).sum() / m
             instance[j, i] = 1 - instance[i, j]
-    # The best is the inverse central because the evaluation does not use the
-    # inverse
-    best_sol = np.argsort(central)
-    return instance, best_sol
+    return instance, central
 
 # Linear Ordering Problem
 class LOP(Problem):
@@ -65,8 +62,7 @@ class LOP(Problem):
             seed = np.random.randint(1, 123456789, 1)[0]
         np.random.seed(seed)
         instance, best_sol = synthetic_LOP(n, m, phi)
-        # The worst is the reverse of the best because the evaluation does not
-        # use the inverse
+        # FIXME: This worst is not the worst.
         worst_sol = best_sol[::-1]
         return cls(n, instance, best_sol = best_sol, worst_sol = worst_sol,
                    instance_name = f"LOP-synthetic,seed={seed},n={n},m={m},phi={phi}")
@@ -107,6 +103,7 @@ class LOP(Problem):
     def fitness_nosave(self, x):
         # In case it is not numpy array.
         x = np.asarray(x, dtype=int)
+        xinverse = np.argsort(x) # Important: We evaluate the inverse permutation
         # Sum of the lower triangle. We have to minimize this.
-        f = np.tril(self.instance[np.ix_(x, x)]).sum()
+        f = np.tril(self.instance[np.ix_(xinverse, xinverse)]).sum()
         return f

### 2020-11-18 runner.py

@@ -50,7 +50,7 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
     timer = Timer()
     df = algo(instance, seed, **algo_params)
     if instance.best_fitness is not None and instance.worst_fitness is not None:
-        df['Fitness_norm'] = (df.Fitness - instance.best_fitness) / (instance.worst_fitness - instance.best_fitness)
+        df['Fitness'] = (df.Fitness - instance.best_fitness) / (instance.worst_fitness - instance.best_fitness)
     df['Function evaluations'] = np.arange(1, len(df['Fitness'])+1)
     df['run_time'] = timer.elapsed()
     df['Problem'] = instance.problem_name

### 2020-11-18 paper/main.tex
update lop definition
fix suppl ref
recalculate everything after the reverts

@@
 
-The \emph{Linear Ordering Problem (LOP)} is defined as 
+The \emph{Linear Ordering Problem (LOP)}, when minimizing, is defined as
+%\footnote{For consistency with the other problems, we minimise the lower triangle instead of maximising the upper triangle.} 
 \begin{equation}\label{eq:lop}
-  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^{n-1} \sum_{j=i+1}^n  a_{\sigma(i), \sigma(j)}
+  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^{n} \sum_{j=1}^{i-1}  a_{\sigma^{-1}(i), \sigma^{-1}(j)}
 \end{equation}
-where $[a_{i,j}]$ is a matrix of size $n \times n$.
+where $[a_{i,j}]$ is a matrix of size $n \times n$ and $\sigma^{-1}$ is the inverse of $\sigma$.
 
@@
 the models (both Bayesian and probabilistic) are not learning anything about
-the fitness landscape. Results for larger PFSP corroborate these conclusions\supplement.
+the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).
 % \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded

### 2020-11-18 paper/main.tex
update paper

@@
 distribution algorithms
-(EDAs)~\cite{CebIruMen2014eda,CebMenLoz2015mallows}. ACO may also be considered
+(EDAs)~\cite{CebIruMen2014eda}. ACO may also be considered
 a type of EDA, since it builds a probability distribution model from which
@@
 far. These classical EDAs  assume that function
-evaluations are cheap and/or the fitness function is white-box.\MANUEL{Do EDAs really assume that the fitness function is white-box? ACO doesn't: They are hybridized with local search }
+evaluations are cheap and/or the fitness function is white-box.%\MANUEL{Do EDAs really assume that the fitness function is white-box? ACO doesn't: They are hybridized with local search }
 %The proposed UMM drastically differs from the above algorithms.
@@
 uBorda~\cite{IruLobPer2020arxiv}, a generalization of the well-known rank
-aggregation algorithm Borda~\cite{Borda1781}. Rank aggregation algorithms
+aggregation algorithm Borda~\cite{AliMei2011kemeny}. Rank aggregation algorithms
 summarize in a single permutation the information of a sample of permutations,
@@
 \subsection{Distributions over permutations}
-Distributions over permutations~\cite{CriFliVer1996rankings} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$.
+Distributions over permutations are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$.
 One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations.
@@
     \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}\Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
-    \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
+    \State $\Sigma \assign$ sampled $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}\Comment{Generate $\beta$ permutations}
+    \State $\sigma\assign  \arg\max_{\pi'\in \Sigma}\arg\min_{ \pi \in S} d(\pi', \pi)   $ \Comment{Select the farthest from $S$}
+   % \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
     \State $S \assign S \cup \{\sigma\}$, $\quad F \assign F \cup \{f(\sigma)\}$
@@
 The strategy is as follows.
-Instead of sampling just one permutation from the MM, the algorithm generates $budgetMM$ permutations, $\Sigma$.
-Then it selects  for evaluation the farest to the sample, i.e.,  $\sigma\in \Sigma$ such that $\sigma = \arg\max\arg\min d(\sigma, \pi)$ for every $\pi\in S$.
+Instead of sampling just one permutation from the MM, the algorithm generates $\beta$ permutations, $\Sigma$.
+Then it selects  for evaluation the farthest to the sample, i.e.,  $\sigma\in \Sigma$ such that $\sigma = \arg\max\arg\min d(\sigma, \pi)$ for every $\pi\in S$.
 
 \section{Experimental setup}\label{sec:setup}
+
+\subsubsection{UMM}
+%
+UMM uses the same budget of evaluations ($\FEmax$) as CEGO
+and initially samples $\minit=10$ solutions uniformly. The other two parameters
+of UMM are $r_1$ and $r_2$, which determine the balance between exploration and
+exploitation. In the experimental section we will study their impact. implemented UMM using Python 3.
+At each sampling step, we set $\beta=10$, meaning that 10 permutations are generated and the farthest from $S$ is selected for evaluation.
 
@@
 
-\subsubsection{UMM}
-%
-UMM uses the same budget of evaluations ($\FEmax$) as CEGO
-and initially samples $\minit=10$ solutions uniformly. The other two parameters
-of UMM are $r_1$ and $r_2$, which determine the balance between exploration and
-exploitation. In the experimental section we will study their impact. implemented UMM using Python 3.
-
 \subsubsection{Other settings}
@@
 First, we create a sample $S$ of 200 permutations of size $n=20$ from a MM with $\sigma_0$ equal to the identity permutation and $\theta=-\log(\textit{Uniformity})$, where  $\textit{Uniformity}\in\{0.5,$ $0.7,$ $0.9\}$, i.e., from peaked distributions to almost uniform distributions.
-Then, we generate matrix $a$ as $a_{i,j} = \sum_{\sigma\in S}\sum_{i<j}\mathbf{I}\{\sigma(i)<\sigma(j)\}$, where $\mathbf{I}\colon S \times S \to \{0,1\}$ is the indicator function.
+Then, we generate matrix $a$ as $a_{i,j} = \sum_{\sigma\in S}\sum_{i<j}\mathbb{I}\{\sigma(i)<\sigma(j)\}$, where ${\mathbb{I}\colon S \times S \to \{0,1\}}$ is the indicator function.
 In principle, these should be best-case instances for UMM since their fitness landscape closely matches the Mallows model.%\MANUEL{Ekhine: Could you revise this text?
@@
 solutions is 2-3 times larger in UMM than CEGO. %
-\MANUEL{But this is a bit misleading because UMM runs for 4 times more
-  evaluations, perhaps we should also measure up to the evaluation that CEGO
-  reaches so we report both values?}
+%\MANUEL{But this is a bit misleading because UMM runs for 4 times more
+%  evaluations, perhaps we should also measure up to the evaluation that CEGO
+%  reaches so we report both values?}
 
@@
 
-\MANUEL{UMM is able to model the LOP landscape
-  at least as well as CEGO as shown by the plots where UMM shows a linear
-  decrease in fitness.}  
-  \EKHINE{me parece mas de CEGO q UMM y parece q UMM
-  solo tiene de ventaja ir rápido. Primero destacaría lo bueno de UMM. Creo q
-  es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos y
-  luego que addemás es más rápido} 
-  \MANUEL{Pero UMM es peor en rec13 y nug12, y
-  en LOP es mejor porque CEGO no llega a 400, es decir, porque UMM es más
-  rápido. No podemos saber qué pasaría si CEGO llegara a 400. Podríamos medir
-  lo que pasa en el momento en que matamos a CEGO y dar también ese valor
-  entonces la diferencia en velocidad no importaría. } 
+%\MANUEL{UMM is able to model the LOP landscape
+%  at least as well as CEGO as shown by the plots where UMM shows a linear
+%  decrease in fitness.}  
+%  \EKHINE{me parece mas de CEGO q UMM y parece q UMM
+%  solo tiene de ventaja ir rápido. Primero destacaría lo bueno de UMM. Creo q
+%  es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos y
+%  luego que addemás es más rápido} 
+%  \MANUEL{Pero UMM es peor en rec13 y nug12, y
+%  en LOP es mejor porque CEGO no llega a 400, es decir, porque UMM es más
+%  rápido. No podemos saber qué pasaría si CEGO llegara a 400. Podríamos medir
+%  lo que pasa en el momento en que matamos a CEGO y dar también ese valor
+%  entonces la diferencia en velocidad no importaría. } 
 The time required to train the surrogate-model of \ekhine{esta establecida la complexity de CEGO???}{UMM increases linearly with the permutation length $n$, while in }
@@
 optimizer.
-% MANUEL: OK?
-% \EKHINE{hhmmm lo hacemos mejor o igual}
 The results presented here show that, despite the intrinsic difficulty of the
@@
 research. % on such problems.
-\MANUEL{I'm thinking about finishing here and commenting out the following paragraph}
+%\MANUEL{I'm thinking about finishing here and commenting out the following paragraph}
 A somewhat surprising result was that the behavior of UMM (and the same for
@@
 %
-% MANUEL: We should also thank the COST ACTION: Ask Carlos?
+\manuel{}{: We should also thank the COST ACTION: Ask Carlos?} \ekhine{}{Sure!, esto no habra q ponerlo aun por ser blind, no?}
 Thanks to Hao Wang (Leiden University) for pointing us to the arguments of
@@
 discussing differences between QAPLIB instances.
+Thanks to Ahmed Boujaada for the experimental support. 
 \end{smaller}

### 2020-11-19 paper/main.tex
update nb, more edits

@@
   both computation time and number of evaluations are of concern.\sloppy
-  \textcolor{red}{16 pages max, deadline: 1 November 2020}
+  %\textcolor{red}{16 pages max, deadline: 1 November 2020}
   \keywords{Combinatorial optimization \and Bayesian optimization \and
@@
     \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}\Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
-    \State $\Sigma \assign$ sampled $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}\Comment{Generate $\beta$ permutations}
-    \State $\sigma\assign  \arg\max_{\pi'\in \Sigma}\arg\min_{ \pi \in S} d(\pi', \pi)   $ \Comment{Select the farthest from $S$}
+    \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}\Comment{Generate $\beta$ permutations}
+    \State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \Comment{Select the farthest from $S$}\label{line:distance}
    % \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
@@
 (\emph{learning}) of the parameters $\sigma_0$, $\theta$ and $\rho$ (line~\ref{line:uborda} in Algorithm~\ref{alg:umm}) and
-the sampling of a new permutation from $MM(\sigma_0,\theta)$ (line~\ref{line:sample}). % in Algorithm~\ref{alg:umm}).
+the sampling from $MM(\sigma_0,\theta)$ (lines~\ref{line:sample}--\ref{line:distance}). % in Algorithm~\ref{alg:umm}).
 
@@
 Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.
-The algorithm also implements the following strategy to avoid sampling the same points more than once, which will be more likely at the explotation phase.
-The strategy is as follows.
-Instead of sampling just one permutation from the MM, the algorithm generates $\beta$ permutations, $\Sigma$.
-Then it selects  for evaluation the farthest to the sample, i.e.,  $\sigma\in \Sigma$ such that $\sigma = \arg\max\arg\min d(\sigma, \pi)$ for every $\pi\in S$.
+
+To avoid sampling the same points more than once and increase the exploration of the search space, the algorithm samples $\beta$ permutations, $\Sigma$.
+Then it selects  for evaluation the farthest from permutations already evaluated ($S$), as shown in line~\ref{line:distance} of Algorithm~\ref{alg:umm}.
+%i.e.,  $\sigma\in \Sigma$ such that $\sigma = \argmax_{\sigma'\in \Sigma}\min d(\sigma', \pi)$ for every $\pi\in S$.
 
@@
 %
-We use the implementation of GECO provided by the
-authors.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
+We use the original implementation of GECO.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
 Although it is never stated in the original paper, the implementation of CEGO
@@
 original papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to
-running CEGO with $\FEmax=400$ instead of $\FEmax=200$. In fact, we found that
+running CEGO with $\FEmax=400$ instead of $\FEmax=200$. We found that
 doubling $\FEmax$ produces a smaller increase in runtime than multiplying the
@@
 On the other hand, % \EKHINE{cambio el orden}
-CEGO starts with a random sampling
-of 10 solutions, with are quite poor as expected, followed by building and
-updating the Kriging model based new evaluations, which leads to an impressive
-improvement in less than 50 evaluations. However, progress is %more or less 
-halted at this point and the model seems to have trouble generating better
-solutions. Lower uniformity values, e.g., 0.5,
- show worsening evaluations
-despite the increase in solutions used to build the model. In summary, the
-Kriging model clearly helps CEGO to quickly identify good solutions in very few
-evaluations. However, for some reason, CEGO is not able to further improve
-those solutions given more data. This may due to the model not being able to
-produce accurate predictions or the underlying optimizer not being able to find
-improved solutions using the model.
+CEGO starts with a random sampling of 10 solutions, with are quite poor as
+expected, followed by building and updating the Kriging model based new
+evaluations, which leads to an impressive improvement in less than 50
+evaluations. However, progress is halted at this point and the model seems to
+have trouble generating better solutions. Around evaluation 200, the search
+becomes fully random. Lower uniformity values, e.g., 0.5, show worsening
+evaluations despite the increase in solutions used to build the model. In
+summary, the Kriging model clearly helps CEGO to quickly identify good
+solutions in very few evaluations. However, for some reason, CEGO is not able
+to further improve those solutions given more data. Given what we know about
+Kriging~\citep{EriPeaGar2019scalable}, it appears that the surrogate model cannot model the local step
+necessary to reach the optimal solution and switches to an exploration strategy
+that ends up being completely random.
+
+% This may due to the model not being able to produce accurate
+% predictions or the underlying optimizer not being able to find improved
+% solutions using the model.
 %\MANUEL{perhaps we should measure the

### 2020-11-19 paper/main.tex
one ref left

@@
 Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$.
-However, in other case uBorda ranking will be closer to the rankings in $S$ with smaller value of $w(\sigma)$.
+Otherwise, uBorda ranking will be closer to the rankings in $S$ with smaller value of $w(\sigma)$.
 Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
-Clearly, the choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
-For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{IruLobPer2020arxiv}.
+The choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
+%% MANUEL: To save space.
+%For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{IruLobPer2020arxiv}.
 %
@@
 Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
-Interestingly, uBorda is computed in polynomial time.
+Finally, uBorda is computed in polynomial time.
 
@@
 
-\subsection{CEGO}
-
-Combinatorial Efficient Global Optimization
-(CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
+\subsection{Combinatorial Efficient Global Optimization (CEGO)}
+
+CEGO~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
 method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
@@
 
-Moreover, we consider real LOP datasets from the LOPLIB
+Moreover, we consider real LOP datasets from the LOLIB
 repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} In
@@
 
-\begin{figure}[htb]
+\begin{figure}[tb]
   \centering%
@@
 helps to refine the model and leads to a better solution being evaluated in the
-next step.  However, it is often the case that new solutions are often worse
-than their predecessors. Nevertheless, in this example, there is a clear
-overall improvement as more evaluations are performed. The minimum values
+next step.  However, it is possible that new solutions are worse
+than their predecessors. %Nevertheless, in this example, there is a clear
+% overall improvement as more evaluations are performed.
+The minimum values
 achieved in the plot up to a particular number of function evaluations gives an
@@
 Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the
-internal probability model of the algorithm over the $\FEmax$ functions
-evaluations. Overall, the mean fitness returned by UMM after $\FEmax$ is better one returned by  CEGO, regardless the uniformity of the instances.
+internal probability model of the UMM over the $\FEmax$ functions
+evaluations. Overall, the mean fitness returned by UMM at $\FEmax=400$ is better than the one returned by  CEGO, regardless the uniformity of the instances.
 
@@
 On the other hand, % \EKHINE{cambio el orden}
-CEGO starts with a random sampling of 10 solutions, with are quite poor as
-expected, followed by building and updating the Kriging model based new
+CEGO starts with 10 random initial solutions, % with are quite poor as
+% expected,
+followed by building and updating the Kriging model based on new
 evaluations, which leads to an impressive improvement in less than 50
-evaluations. However, progress is halted at this point and the model seems to
-have trouble generating better solutions. Around evaluation 200, the search
+evaluations. However, progress is halted at this point and CEGO seems to
+have trouble finding better solutions. Around evaluation 200, the search
 becomes fully random. Lower uniformity values, e.g., 0.5, show worsening
@@
 summary, the Kriging model clearly helps CEGO to quickly identify good
-solutions in very few evaluations. However, for some reason, CEGO is not able
+solutions in very few evaluations. However, CEGO is not able
 to further improve those solutions given more data. Given what we know about
@@
 Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
-\begin{figure}[tbp]
+\begin{figure}[!tbp]
   \centering%
@@
 %  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d200_01}\\
-    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOPLIB.\label{fig:loplib}}
+    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOLIB.\label{fig:lolib}}
   \end{figure}
@@
 %to translate to the
-%instances available in LOPLIB~\citep{}. We show in
-%Fig.~\ref{fig:loplib} three examples with different permutation size, however,
-%the results are consistent for other LOPLIB instances (complete results
+%instances available in LOLIB~\citep{}. We show in
+%Fig.~\ref{fig:lolib} three examples with different permutation size, however,
+%the results are consistent for other LOLIB instances (complete results
 %available as supplementary material \supplement). Due to extremely long
@@
 %point of CEGO). It appears that our synthetic LOP instances have a fitness
-%landscape that is very amenable to CEGO's Kriging model, whereas the LOPLIB
+%landscape that is very amenable to CEGO's Kriging model, whereas the LOLIB
 %instances do not present such landscape. Comparing CEGO and UMM results, the
@@
 
-In the following paragraphs we analyze the result over real-world instances. %
-Figure~\ref{fig:loplib} shows the results on the instances taken from LOPLIB.
+In the following paragraphs we analyze the results over real-world instances. %
+Figure~\ref{fig:lolib} shows the results on the instances taken from LOLIB.
   %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
   % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
-The type of instances from the synthetic LOP instances, and, thus, so does the
-landscape. However, UMM shows a very similar behavior: there is a (close to)
-linear decrease of the fitness of new solutions over each new evaluation. This
-suggests that UMM is able to model the fitness landscape. Regarding CEGO, we
-observe a similar behavior for the first 50-100 evaluations. At this point,
-CEGO reaches the time limit of 3 days, so no observations can be made for
-longer budgets.  Overall, the improvement from the first (random) to the last
-solutions is 2-3 times larger in UMM than CEGO. %
+The type of instances differs greatly from the synthetic LOP instances, and,
+thus, so does the landscape. However, UMM shows a very similar behavior: there
+is a (close to) linear decrease of the fitness of new solutions over each new
+evaluation. This suggests that UMM is able to model the fitness
+landscape. Regarding CEGO, we observe a similar behavior to UMM for the first
+50-100 evaluations, however, in several instances CEGO reaches the time limit
+of 144 hours, so no observations can be made for longer budgets. In smaller
+instances (e.g., Figure~\ref{fig:lolib}, top), CEGO again appears to stagnate.
+%Overall, the improvement from the first  to the last
+%solutions is 2-3 times larger in UMM than CEGO. %
 %\MANUEL{But this is a bit misleading because UMM runs for 4 times more
@@
 %  reaches so we report both values?}
+%
 
@@
 of fitness evaluations is severely limited (here, up to $400$ evaluations).
-UMM \ekhine{}{has shown to perform statistically better or equal than CEGO, the
-  state of the art algorithm for black-box expensive optimization, in the same instances as in the original paper of
-  CEGO and some new instances. UMM} is specially well-suited for budget-limited
-scenarios that are still time-sensitive, e.g., when the overhead incurred by
-the optimizer should be no more than a few minutes added to each fitness
-evaluation.
+Our experiments presented here show that UMM is able to match the results of
+CEGO, a state-of-the-art Bayesian optimizer for expensive black-box
+optimization if the computation time of the optimizers is not a concern.
+
+In addition, UMM is specially well-suited for budget-limited scenarios that are
+still time-sensitive, e.g., when the overhead incurred by the optimizer should
+be no more than a few minutes added to each fitness evaluation. By comparison,
+CEGO may require up to one hour per fitness evaluation, with the time
+increasing non-linearly with permutation size and larger budgets.  Hence, UMM
+is a computationally feasible alternative for relatively large problem sizes.
 
@@
 %  entonces la diferencia en velocidad no importaría. } 
-The time required to train the surrogate-model of \ekhine{esta establecida la complexity de CEGO???}{UMM increases linearly with the permutation length $n$, while in }
-CEGO  it sharply
-increases with $n$, requiring 
-sometimes more than hour per fitness evaluation as shown in the experimental section. 
-Hence, UMM is a computationally
-feasible alternative for relatively large problem sizes.
+% The time required to train the surrogate-model of \ekhine{esta establecida la complexity de CEGO???}{UMM increases linearly with the permutation length $n$, while in }
+% CEGO  it sharply
+% increases with $n$, requiring 
+% sometimes more than hour per fitness evaluation as shown in the experimental section. 
+% Hence, UMM is a computationally
+% feasible alternative for relatively large problem sizes.
 
@@
 to match (and sometimes even improve) the results of a Bayesian
-optimizer.
-The results presented here show that, despite the intrinsic difficulty of the
-expensive black-box combinatorial scenario, there are still significant
-advances yet to be made. Thus, we hope that this paper will motivate further
-research. % on such problems.
+optimizer. Moreover, it is overwhelmingly faster than the alternatives for
+large problem sizes.  The results presented here show that, despite the
+intrinsic difficulty of the expensive black-box combinatorial scenario, there
+are still significant advances yet to be made. Thus, we hope that this paper
+will motivate further research. % on such problems.
 %\MANUEL{I'm thinking about finishing here and commenting out the following paragraph}
-A somewhat surprising result was that the behavior of UMM (and the same for
-CEGO) on the PFSP is close to random. This result may be due to the distance
+
+An interesting result is that the behavior of UMM (and the same for CEGO) on
+the PFSP (and some QAP instances) is close to random. The same behavior is
+observed for CEGO, which seems to have been never reported until now.
+This result may be due to the distance
 metric used, but further analysis is needed to corroborate this
-conjecture. Here, we have focused on Kendall's-$\tau$ distance, however, it is
+conjecture. Here, we have focused on Kendall's-$\tau$ distance, however, we are certain it is
 possible to extend UMM to other distance metrics, which will allow us to
@@
 %
-\manuel{}{: We should also thank the COST ACTION: Ask Carlos?} \ekhine{}{Sure!, esto no habra q ponerlo aun por ser blind, no?}
-Thanks to Hao Wang (Leiden University) for pointing us to the arguments of
-\citet{EriPeaGar2019scalable} and to Thomas Stützle (Université libre de Bruxelles) for suggesting works
-discussing differences between QAPLIB instances.
-Thanks to Ahmed Boujaada for the experimental support. 
+%\manuel{}{: We should also thank the COST ACTION: Ask Carlos?} \ekhine{}{Sure!, esto no habra q ponerlo aun por ser blind, no?}
+Thanks to: Hao Wang (Leiden University) for pointing us to the arguments of
+\citep{EriPeaGar2019scalable}; Thomas Stützle (Université libre de Bruxelles) for suggesting works
+about differences between QAPLIB instances; and Ahmed Boujaada for the experimental support. 
 \end{smaller}
@@
 \renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}
-\bibliographystyle{splncs04nat}
-\bibliography{optbib/abbrev,optbib/authors,optbib/journals,optbib/biblio,optbib/crossref}%,./mendeley}
+\bibliographystyle{splncs04natnodoi}
+\bibliography{optbib/abbrev,optbib/authors,optbib/abbrevshort,optbib/journals,optbib/biblio,optbib/crossref}%,./mendeley}
 

### 2020-11-19 paper/main.tex
16 pages

@@
   \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.\label{fig:kra32}}
+  \vspace*{-1em}
 \end{figure}
@@
 would be even stronger.
-
-In the case of the PFSP, the differences are minor and/or not statistically
+%
+In the PFSP, the differences are minor and/or not statistically
 significant, with the qualification that both algorithms appear to be
@@
   \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance.
-The 95\% confidence interval corresponds to the two independent samples t-test on the mean of the difference between the fitness of UMM minus the one of CEGO.
-Mean runtime is measured in hours.
-If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 144 hours, it was terminated earlier.
-This was the case for all LOP instances shown.\label{tab:results}}
+The 95\% confidence interval corresponds to the two independent samples t-test on the mean difference between the fitness of UMM minus the one of CEGO.
+Mean runtime is shown in hours.
+CEGO runs were stopped after 144 hours.
+This was the case for all LOP instances.\label{tab:results}}
   \resizebox{\textwidth}{!}{%
@@
 \end{tabular}}
+\vspace*{-1em}
  \end{table}

### 2020-11-19 paper/main.tex
update paper

@@
   sometimes surpass, the solutions obtained by CEGO, a Bayesian optimization
-  algorithm for combinatorial optimization, specially when the budget of
-  solutions evaluated approaches 400. The computational cost of UMM is at least
-  an order of magnitude smaller than CEGO, which makes it specially useful when
-  both computation time and number of evaluations are of concern.\sloppy
+  algorithm for combinatorial optimization. %, specially when the budget of solutions evaluated approaches 400. 
+  Moreover, the computational complexity of UMM increases linearly with both the number of function evaluations and the permutation size.
+%  an order of magnitude smaller than CEGO, which makes it specially useful when
+  %both computation time and number of evaluations are of concern.
+  \sloppy
   %\textcolor{red}{16 pages max, deadline: 1 November 2020}
@@
 
-As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are uniformly randomly generated (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated.
+As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are generated uniformly at random  (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated.
 Then, using the current sample $S$, the uBorda algorithm is applied to learn a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}).
@@
   r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma'\in S'}\rho^{f(\sigma')} \quad\text{where}\; &|S'| = r_2|S| \\[-1.2em]
-  &\land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\leq f(\sigma)
+  &\land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\leq f(\sigma).
   \end{split}
@@
 i.e., $S'$ contains the $r_2|S|$ elements of $S$ with the best corresponding fitness values in $F$.
-The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification: higher values bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas smaller values make the sampling more uniform, thus slowing down the learning.
-
-% As shown in Section~\ref{sec:backgroud}, MM can be approximately learnt in two= steps (1) estimating the location parameter $\sigma_0$ and (2) computing the scale parameter $\theta$.
-Since the scale parameter $\theta$ is set in the sampling step,\MANUEL{X How is it set?} we focus on estimating $\sigma_0$.
-
-% The location parameter of a MM is usually approximated with the Borda algorithm.
-Since we are interested on biasing the model towards the regions of the space where the fitness function is smaller, we would like to replicate the permutations a number of times that is inversely proportionally to their fitness values.
-% We can simulate the replication of the best fitted individuals by using uBorda and setting $w(\sigma)=\rho^{f(\sigma)}$, as described in .
-This generalization requires setting a value for parameter $\rho$.
-We show how to choose this value in the experimental section.\MANUEL{I thought we select it automatically now?\ekhine{that is $\theta$}}
+%The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification.
+%: extreme different values bias the sampling towards the best solutions found so far, which may lead to quick stagnation, whereas values (0.5,0.5) make the sampling more uniform, thus slowing down the learning.
 

### 2020-11-19 paper/main.tex
update paper

@@
 \begin{table}[tb]
-  \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance.
+  \caption{Mean fitness (and standard deviation) of the best solution found.
+% at the end of each run, over 10 independent runs for each problem instance.
 The 95\% confidence interval corresponds to the two independent samples t-test on the mean difference between the fitness of UMM minus the one of CEGO.
   Mean runtime is shown in hours. 
-CEGO runs were stopped after 144 hours.
-This was the case for all LOP instances.\label{tab:results}}
+  %CEGO runs were stopped after 144 hours. This was the case for all LOP instances.
+  \label{tab:results}}
   \resizebox{\textwidth}{!}{%
@@
 An interesting result is that the behavior of UMM (and the same for CEGO) on
-the PFSP (and some QAP instances) is close to random. The same behavior is
-observed for CEGO, which seems to have been never reported until now.
+the PFSP (and some QAP instances) is close to random. %The same behavior is
+%observed for CEGO, which seems to have been never reported until now.
 This result may be due to the distance
@@
   \vspace*{-1em}
-\paragraph*{Acknowledgements.}
+\paragraph*{Acknowledgements.} Blind
 % M.\@ L\'opez-Ib\'a\~nez is a ``Beatriz Galindo'' Senior Distinguished Researcher (BEAGAL 18/00053) funded by the Ministry of Science and Innovation of the Spanish Government.
@@
 %\manuel{}{: We should also thank the COST ACTION: Ask Carlos?} \ekhine{}{Sure!, esto no habra q ponerlo aun por ser blind, no?}
-Thanks to: Hao Wang (Leiden University) for pointing us to the arguments of
-\citep{EriPeaGar2019scalable}; Thomas Stützle (Université libre de Bruxelles) for suggesting works
-about differences between QAPLIB instances; and Ahmed Boujaada for the experimental support. 
+%Thanks to: Hao Wang (Leiden University) for pointing us to the arguments of
+%\citep{EriPeaGar2019scalable}; Thomas Stützle (Université libre de Bruxelles) for suggesting works
+%about differences between QAPLIB instances; and Ahmed Boujaada for the experimental support.
+
 \end{smaller}

### 2020-11-25 lop.py
* umm.py: Use lambda to simplify.  * lop.py: Simplify. * mallows_kendall.py: Comments.

@@ -49,7 +49,8 @@ def synthetic_LOP(n, m, phi):
         for j in range(i+1, n):
             instance[i, j] = (s[:, i] < s[:, j]).sum() / m
             instance[j, i] = 1 - instance[i, j]
-    return instance, central
+    best_sol = np.argsort(central)
+    return instance, best_sol
 
 # Linear Ordering Problem
 class LOP(Problem):
@@ -63,7 +64,6 @@ class LOP(Problem):
         np.random.seed(seed)
         instance, best_sol = synthetic_LOP(n, m, phi)
         # print(instance[np.ix_(best_sol, best_sol)])
-        best_sol = np.argsort(best_sol)
         # print(instance[np.ix_(best_sol, best_sol)])
         # print(np.tril(instance[np.ix_(best_sol, best_sol)]))
         # print(np.tril(instance[np.ix_(best_sol, best_sol)]).sum())

### 2020-11-25 mallows_kendall.py

@@ -4,7 +4,7 @@ import scipy as sp
 from scipy import optimize
 #fit MM
 
-# FIXME: Isn't this specific to Kendall?
+# MANUEL: Isn't this specific to Kendall? If so, perhaps we should have a class for distance metrics.
 def max_dist(n):
     return int(n * (n - 1) / 2)
 

### 2020-11-25 umm.py

@@ -14,7 +14,7 @@ def binary_search_rho(w, ratio_samples_learn, weight_mass_learn,
       assert np.all(w <= 1.0)
 
       #if pos is None we take the largest 4th.
-      #find the rho s.t. the largest 25%(ratio_samples) of the weights  (rho**ws) take the 0.9(weight_mass) of the total ws.  rho^w[:pos] = 0.9*rho^w
+      # Find the rho s.t. the largest 25%(ratio_samples) of the weights  (rho**ws) take the 0.9(weight_mass) of the total ws.  rho^w[:pos] = 0.9*rho^w
       # codes as a recursive binary search in (0,1)
       pos = int(len(w) * ratio_samples_learn)
       # print(pos,len(w),ratio_samples_learn)
@@ -22,15 +22,13 @@ def binary_search_rho(w, ratio_samples_learn, weight_mass_learn,
       acum = np.cumsum(rho_med ** w)
       a = acum[pos]
       b = acum[-1]
-      if abs(a/b - weight_mass_learn) < tol:
+      if abs(a / b - weight_mass_learn) < tol:
           return rho_med
-      if a/b > weight_mass_learn:
-          mid = rho_ini
-          last = rho_med
+      if a / b > weight_mass_learn:
+          mid, last = rho_ini, rho_med
       else:
-          mid = rho_med
-          last = rho_end
-  except:
+          mid, last = rho_med, rho_end
+  except: # MANUEL: How can the above fail?
        print(w)
        pos = int(len(w) * ratio_samples_learn)
        # print(pos,len(w),ratio_samples_learn)
@@ -41,25 +39,31 @@ def binary_search_rho(w, ratio_samples_learn, weight_mass_learn,
        print(a,b,a/b,weight_mass_learn,pos,w, rho_med,ratio_samples_learn, weight_mass_learn,rho_ini,rho_end)
   return binary_search_rho(w, ratio_samples_learn, weight_mass_learn, mid, last)
 
-def get_expected_distance(iterat,n,m):
-    N = (n - 1) * n / 2
-    f_ini, f_end = N / 4, 1
-    iter_decrease = m - 10
-    salto = (f_ini - f_end ) / iter_decrease
-    a = f_ini - salto * iterat
-    return max(a, f_end)
+def get_expected_distance(iterat, n, budget):
+  # MANUEL: Should this be Kendall max dist?
+  N = (n - 1) * n / 2
[…]

### 2020-11-25 1-launch_exps.sh
new results with er0 and er1

@@ -79,16 +79,15 @@ INSTANCES="\
  pfsp/rec31.txt \
 "
 
+###### For LOLIB instances
+#INSTANCES="$INSTANCES $(tr '\n' ' ' < loplib-instances.txt)"
+
 ###### Synthetic LOP instances
 INSTANCES=$(gen_lop_synthetic "")
 
-###### For LOLIB instances
-INSTANCES="$INSTANCES $(tr '\n' ' ' < loplib-instances.txt)"
-
-
 budget=400
 
-eval_ranks=0
+#eval_ranks=0
 eval_ranks=1
 
 cego_m_ini=10

### 2020-11-25 runner.py

@@ -57,6 +57,6 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
     df['instance'] = instance.instance_name
     df['Solver'] = algo_name
     if out_filename is not None:
-        df.to_csv(out_filename + '.csv.gz', index=False, compression = "gzip")
-        df.to_pickle(out_filename + '.pkl.gz', compression = "gzip")
+        df.to_csv(out_filename + '.csv.xz', index=False, compression = "xz")
+        df.to_pickle(out_filename + '.pkl.xz', compression = "xz")
     return df

### 2021-01-20 paper/main.tex
* reduce spacing even more.
fix smaller -> larger in uborda

@@
 \newrevisor{ekhine}{EKHINE}{red}
+%\usepackage{showframe}
 
@@
 Borda and uBorda are equivalent when $w(\sigma)$ is constant for every $\sigma$.
-Otherwise, uBorda ranking will be closer to the rankings in $S$ with smaller value of $w(\sigma)$.
-Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with low values of $w(\sigma$).
+Otherwise, uBorda ranking will be closer to the rankings in $S$ with larger value of $w(\sigma)$.
+Roughly speaking, it is equivalent to replicating in the sample the $\sigma$ with high values of $w(\sigma$).
 The choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
@@
 Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
-\begin{figure}[!tbp]
+\begin{figure}[!tb]
   \centering%

### 2021-01-20 paper/main.tex

@@
 minimal fitness values. This is achieved by means of the uBorda ranking
-(Section~\ref{sec:uborda}), which can be seen as replicating each permutation
+(Section~\ref{sec:uborda}) with $w(\sigma) = \rho^f(\sigma)$, with $\rho \in (0,1)$, which can be seen as replicating each permutation
 $\sigma \in S$ a number of times inversely proportional to its fitness

### 2021-01-20 paper/main.tex

--- 2021-01-20 paper/main.tex
@@
 minimal fitness values. This is achieved by means of the uBorda ranking
-(Section~\ref{sec:uborda}), which can be seen as replicating each permutation
-$\sigma \in S$ a number of times inversely proportional to its fitness
-$f(\sigma)$, i.e., it behaves as if there were many copies of the best
+(Section~\ref{sec:uborda}) with $w(\sigma) = \rho^{f(\sigma)}$ and $\rho \in (0,1)$, which %  can be seen as replicating each permutation
+% $\sigma \in S$ a number of times inversely proportional to its fitness
+% $f(\sigma)$, i.e., it 
+behaves as if there were many copies of the best
 solutions and few of the bad ones. A property of UMM is that the

### 2021-01-20 runCEGO.R
Fix runCEGO.R.

@@ -2,8 +2,8 @@ argv <- commandArgs(trailingOnly=TRUE)
 seed <- as.numeric(argv[1])
 eval_ranks <- as.numeric(argv[2])
 elapsed <- proc.time()
-#seed <- 12345
-#eval_ranks <- 1
+#seed <- 1
+#eval_ranks <- 0
 library(CEGO)
 cat("seed = ", seed, "  eval_ranks = ", eval_ranks, "\n")
 print(sessionInfo())
@@ -32,8 +32,16 @@ my_optimCEGO <- function (x = NULL, fun, control = list())
         control$initialDesignSettings$distanceFunction <- distanceFunction
     fun
     if (!vectorized) { 
-        fn <- if (control$eval_ranks) function(x) unlist(lapply(x, fun)) 
-              else function(x) unlist(lapply(x, function(y) fun(order(y)))) 
+      fn <- if (control$eval_ranks) { function(x) unlist(lapply(x, fun))
+            } else { 
+              function(x) {
+                ## print(x)
+                ## print(lapply(x,order))
+                ## print(lapply(x, fun))
+                ## print(lapply(lapply(x, order), fun))
+                return(unlist(lapply(lapply(x, order), fun)))
+              }
+            }
     } else {
      # fn <- fun
      stop("We do not handle vectorized functions")
@@ -143,7 +151,7 @@ my_optimCEGO <- function (x = NULL, fun, control = list())
     res$message <- msg
     res$distances <- NULL
     res
-    }
+}
 
 n <- 20
 m <- 15
@@ -184,7 +192,7 @@ cF <- function() sample(n)
 #    print("antes del optimCEGO")
 budgetGA <- 10^3
 budget <- 400
-res <- optimCEGO(x = NULL,
+res <- my_optimCEGO(x = NULL,
                  fun = fun,
                  control = list(creationFunction=cF,
                                 distanceFunction = distancePermutationSwap,

### 2021-01-20 paper/main.tex
Reviewer comments.

@@
 and it is set at each iteration such that the largest $100r_1$\% of the mass of
-the weights is concentrated in the best $100r_2$\% of the solutions in $S$, i.e.,
+the weights is concentrated in the best $100r_2$\% of the solutions in $S$, i.e.,\MANUEL{Reviewer 3:  Justify the form of the formula used to learn the "learning rate" $\rho$ (equation 4).
+Why this particular approach was chosen?}
 %
@@
 Parameter $\theta$ is set automatically by the algorithm to control the diversification and intensification trade off.
-At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$ (half of the expected distance for the uniform distribution), and then the expected distance linearly decreases until the last iteration where $\theta$ is set such that $\Edist = 1$.
+At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$ (half of the expected distance for the uniform distribution),\MANUEL{Reviewer 3: "At the first iteration, $\theta$ is set so that $E[D] = {n \choose 4}$".
+Why?} and then the expected distance linearly decreases until the last iteration where $\theta$ is set such that $\Edist = 1$.
 Despite it is not possible to isolate $\theta$ in Eq.~\eqref{eq:expectation}, its monotonicity allows the use of bisection methods for efficiently finding an approximation.
@@
 
-To avoid sampling the same points more than once and increase the exploration of the search space, the algorithm samples $\beta$ permutations, $\Sigma$.
+To increase the exploration of the search space, the algorithm samples $\beta$ permutations, $\Sigma$.
 Then it selects  for evaluation the farthest from permutations already evaluated ($S$), as shown in line~\ref{line:distance} of Algorithm~\ref{alg:umm}.

### 2021-01-20 cego.py
minor updates to scripts

@@ -26,8 +26,8 @@ def cego(instance, seed, budget, m_ini, budgetGA, eval_ranks):
     library(CEGO)
     print(sessionInfo())
     # This is identical to the function in the package but takes also a maxTime parameter.
-optimCEGO <- function (x = NULL, fun, control = list()) 
-{
+    my_optimCEGO <- function (x = NULL, fun, control = list()) 
+    {
     con <- list(evalInit = 2, vectorized = FALSE, verbosity = 0, 
         plotting = FALSE, targetY = -Inf, budget = 100, creationRetries = 100, 
         distanceFunction = distancePermutationHamming, creationFunction = solutionFunctionGeneratorPermutation(6), 
@@ -51,7 +51,7 @@ optimCEGO <- function (x = NULL, fun, control = list())
     fun
     if (!vectorized) { 
         fn <- if (control$eval_ranks) function(x) unlist(lapply(x, fun)) 
-              else function(x) unlist(lapply(x, function(y) fun(order(y) - 1))) 
+              else function(x) unlist(lapply(x, function(y) fun(order(y) - 1)))
     } else {
      # fn <- fun
      stop("We do not handle vectorized functions")
@@ -174,7 +174,7 @@ optimCEGO <- function (x = NULL, fun, control = list())
     cF <- function() sample(n)
     # start optimization
 #    print("antes del optimCEGO")
-    res <- optimCEGO(x = NULL,
+    res <- my_optimCEGO(x = NULL,
                      fun = fun,
                      control = list(creationFunction=cF,
                                      distanceFunction = dist,

### 2021-01-20 collect-data.R

@@ -7,3 +7,4 @@ for(i in 1:30) {
 saveRDS(cegores, file="cegores.rds")
 write.csv(cegores, file="cegores.csv", row.names=FALSE)
 aggregate(cegores$time, list(er=cegores$eval_ranks), mean)
+print(cegores)

### 2021-01-24 2-collect-tuning.sh
new tuning results

@@ -28,23 +28,24 @@ gen_lop_synthetic() {
 
 ## For QAP, PFSP instances
 INSTANCES="\
-  qap/kra30a.dat \
-  qap/kra30b.dat \
- qap/nug30.dat \
- qap/tho30.dat \
- pfsp/rec13.txt \
+qap/kra30a.dat \
+qap/kra30b.dat \
+qap/nug30.dat \
+qap/tho30.dat \
+pfsp/rec13.txt \
 pfsp/rec19.txt \
+lop/RandB/N-p40-02 \
+lop/IO/N-t59d11xx \
+lop/SGB/N-sgb75.02 \
+lop/xLOLIB/N-be75eec_150 \
 "
-
-###### For LOLIB instances
-#INSTANCES="$(grep -v '#' lolib-instances.txt | tr '\n' ' ')"
 
 ###### Synthetic LOP instances
 INSTANCES=$(gen_lop_synthetic $INSTANCES)
 
-OUTFILE="$OUTDIR/tuning.csv"
+OUTFILE="$BINDIR/results-er0/tuning.csv"
 echo "instance,seed,umm_m_ini,budgetMM,r_1,r_2,budget,eval_ranks,fitness" > $OUTFILE 
-counter=0
+
 for instance in $INSTANCES; do
     RESULTS="$OUTDIR/tuning-er?/$instance"
     echo "Reading $RESULTS"

### 2021-01-24 collect-data.R
cegores.csv: Update.

@@ -7,4 +7,5 @@ for(i in 1:30) {
 saveRDS(cegores, file="cegores.rds")
 write.csv(cegores, file="cegores.csv", row.names=FALSE)
 aggregate(cegores$time, list(er=cegores$eval_ranks), mean)
-print(cegores)
+head(cegores[cegores$eval_ranks == 0,])
+head(cegores[cegores$eval_ranks == 1,])

### 2021-01-24 paper/main.tex
Figures

@@
 %Finally, we note that learning Budget-constrained Combinatorial is a hot topic for a variety of structures besides permutation~\cite{manchanda2020}. 
+\manuel{- the authors assigned to black-box optimization problems characteristics that they do not own. For instance they say that "local search cannot be used".
+ they mixed expensive to evaluate functions with black-box functions and of course they are different. In fact much of the metaheuristics do not mind
+ if the problem is withe-box or black-box as they do not use static information from the function in the optimization process. In this case I understand
+ that they have a black-box problem where the objective function is costly to evaluate. This should be clear and I would recommend to delete other comments
+ related with white-box and black-box.
+}{}
 
@@
 The choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
+\ekhine{reesribe y ojo an grater / less}{}
 %% MANUEL: To save space.
@@
 The process is repeated up to a given maximum budget of evaluations $\FEmax$.
+\ekhine{(3) In section 3, second half of the second paragraph (from "Then, we sample a new candidate [...]"): the algorithm 1 is not properly described.
+Please provide a description in line with algorithm 1 and, above all, describe the rationale behind the choice of line 7 of algorithm 1.
+I like it, but it deserves to be better described.
+It seems to me a sort of "novelty-driven" optimization.(4) A small detail, but requires a clarification.
+In page 6, learning step paragraph: the weight formula uses fitness in the exponent, while the base rho is in (0,1).
+Clearly the authors are working with minimization problems and want to increase the weight when fitness improves/decreases.
+This is ok, but this only happens with their formula when fitnesses are greater than 1.
+I know this is quite usual in the problems analyzed by the authors, but it may not be so usual in general.
+Please add a clarification.}{}
 
@@
 %i.e.,  $\sigma\in \Sigma$ such that $\sigma = \argmax_{\sigma'\in \Sigma}\min d(\sigma', \pi)$ for every $\pi\in S$.
+
+\ekhine{The algorithm uses a parameter $\beta$ that determines the number of points sampled, while the authors analyze the sensitivity of the results to the parameters r 1 and r 2, they do not do that with $\beta$.
+In my opinion the value of $\beta$ is quite relevant.}{}
 

### 2021-01-24 paper/main.tex
paper

@@
 
+\ekhine{6.
+When the main contribution of the paper is discussed it is stated that this contribution is the Unbalanced Mallows *model* (which does not seem to be true as the models used in the paper seem to be known already).
+Right after that it is stated that the contribution is *a novel algorithm* which, I think, reflects the contents of the paper much better.
+}{}
 In this work, we propose and analyze Unbalanced Mallows Model (UMM), a
@@
 
-This paper structured as follows.
+This paper is structured as follows.
 Section~\ref{sec:backgroud} summarizes fundamental concepts about permutation spaces, uBorda, black-box combinatorial optimization under a limited budget and CEGO, a Bayesian optimizer for such problems.
@@
 We end the paper with overall conclusions and a call for further research on this type of optimization scenario (Section~\ref{sec:conclusions}).
+
+\manuel{9. Please use some newer references instead of [11] and, especially, [14]}{}
 
@@
 
-\subsection{Permutations}
+\subsection{Permutations: rankings vs orderings}
 
@@
 The set of all permutations of $n$ values is denoted as $S_n$ and has cardinality $n!$.
-We represent permutations as an ordered vector and, we say that position $i$ contains value $j$ whenever $\sigma(i)=j$.
-We denote permutations with Greek letters.
-% \MANUEL{Do we use $e$ anywhere? If not, we do not need this definition ``with exception of the identity permutation denoted as $e=(1, 2, 3, \dotsc,n)$.''}
+We will use the one-line notation and denote permutations with Greek letters, $\sigma  = (\sigma(1), \ldots , \sigma(n))$.
 We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.
 
-Distance in permutation space can be measured in various ways.  Kendall's-$\tau$ counts the number of items ranked in different order in both permutations:
-%
+Permutation $\sigma$ represents a ranking of item:  $\sigma(i)=j$ denotes that item $i$ has rank $j$.
+The ordering associated with $\sigma$ is given by its inverse $\sigma^{-1}$, therefore,  $\sigma(i)=j \Leftrightarrow  \sigma^{-1}(j)=i$.
+This distinction is in order since statistical methodologies are devoted for rankings and optimizaiton literature for orderings and mixing both concepts is a very common mistake.
+We will illustrate in the experimental section the consequences of this confusion.
+
+Kendall's-$\tau$ distance is the metric for rankings that counts the pairs of items ranked in different order in both permutations:
+
 \begin{equation}
@@
 
-An equivalent definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made in $\sigma_1^{-1}$ to obtain $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
+An equivalent definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made to convert $\sigma_1^{-1}$ into $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
+\ekhine{ojo, aqui lo dice bien o  mal?}{}
 
@@
 The choice of  $w(\sigma)$ depends on the relevant property of a ranking in a particular domain.
-\ekhine{reesribe y ojo an grater / less}{}
-%% MANUEL: To save space.
-%For example, uBorda has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift~\cite{IruLobPer2020arxiv}.
-%
-% In this paper, we would like to replicate each permutation $\sigma$ a number of times that is inversely proportional to $f(\sigma)$.
-Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
 Finally, uBorda is computed in polynomial time.
@@
 
-\subsection{Combinatorial Efficient Global Optimization (CEGO)}
-
-CEGO~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
-method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
-optimization problems. In EGO, Gaussian process models are used as a surrogate
-of the landscape of the expensive original problem. An optimization method
-searches for solutions in the surrogate model by optimizing the expected
-improvement criterion, which balances the expected mean and variance of the
-chosen solution. Once a solution is chosen, it is evaluated on the actual
-fitness function and the result is used to update the surrogate-model,
-hopefully increasing its predictive power.
-
-CEGO replaces the Euclidean distance measure, used by the surrogate model in
-EGO, with a distance measure appropriate to combinatorial
-landscapes~\citep{ZaeStoBar2014:ppsn}, such as Kendall's-$\tau$ distance for
-permutations (Eq.~\ref{eq:kendall_decomp}). In CEGO, the surrogate model is
-explored by a GA with crossover and mutation operators appropriate for the
-particular combinatorial problem. The original paper notes that coupling the GA
-with local search does not improve the results significantly since the model is
-anyway an inexact estimation of the original fitness
-function~\citep[p.~875]{ZaeStoFriFisNauBar2014}. A thorough comparison of CEGO
-with other optimizers for expensive black-box combinatorial optimization ranked
-it as the best-performing~\citep{ZaeStoFriFisNauBar2014}.
-
-\renewcommand{\subsubsection}[1]{\smallskip{}\noindent\textbf{#1}.}
-
 \section{Unbalanced Mallows model (UMM)}\label{sec:umm}
@@
 Mallows model and uBorda learning.
+
+\ekhine{10.
+In Algorithm 1 'S' and 'F' are *sets* (thus, unordered).
+Therefore, when adding sigma and f(sigma) the correspondence between the two elements seems to be lost.
+This mapping would be better represented as a set of *pairs* < sigma, f(sigma) >.
+You always add these elements to both sets at the same time in step 8, anyway.
+}{}
 
@@
 
+\ekhine{4. Justify the form of the formula used to learn the "learning rate" $\rho$ (equation 4). Why this particular approach was chosen?}{}
+
 \subsubsection{Sampling step}\label{sec:sampling}
@@
 
+\ekhine{5. "At the first iteration, Theta is set so that $E[D] = {n \choose 4}$". Why?}{}
 A purely explotation approach would always generate $\sigma_0$ in the sampling stage.
@@
 
+\subsection{Convergence guarantees}
+\ekhine{}{completar}
+
 \section{Experimental setup}\label{sec:setup}
@@
 of UMM are $r_1$ and $r_2$, which determine the balance between exploration and
-exploitation. In the experimental section we will study their impact. implemented UMM using Python 3.
+exploitation. In the experimental section we will study their impact. %implemented UMM using Python 3.
 At each sampling step, we set $\beta=10$, meaning that 10 permutations are generated and the farthest from $S$ is selected for evaluation.
 
-\subsubsection{CEGO}
-%
+\subsubsection{CEGO} \ekhine{tenemos una pagina entera dedicada a CEGO :s}{}
+Combinatorial Efficient Global Optimization (CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
+method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
+optimization problems. In EGO, Gaussian process models are used as a surrogate
+of the landscape of the expensive original problem. An optimization method
+searches for solutions in the surrogate model by optimizing the expected
+improvement criterion, which balances the expected mean and variance of the
+chosen solution. Once a solution is chosen, it is evaluated on the actual
+fitness function and the result is used to update the surrogate-model,
+hopefully increasing its predictive power.
+
+CEGO replaces the Euclidean distance measure, used by the surrogate model in
+EGO, with a distance measure appropriate to combinatorial
+landscapes~\citep{ZaeStoBar2014:ppsn}, such as Kendall's-$\tau$ distance for
+permutations (Eq.~\ref{eq:kendall_decomp}). In CEGO, the surrogate model is
+explored by a GA with crossover and mutation operators appropriate for the
+particular combinatorial problem. The original paper notes that coupling the GA
+with local search does not improve the results significantly since the model is
+anyway an inexact estimation of the original fitness
+function~\citep[p.~875]{ZaeStoFriFisNauBar2014}. A thorough comparison of CEGO
+with other optimizers for expensive black-box combinatorial optimization ranked
+it as the best-performing~\citep{ZaeStoFriFisNauBar2014}.
+
+\renewcommand{\subsubsection}[1]{\smallskip{}\noindent\textbf{#1}.}
+
 We use the original implementation of GECO.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
@@
 %
-Experiments with real-world expensive black-box problems would be computationally infeasible. Instead, we consider classical combinatorial optimization problems as black-box optimization problems.
+Experiments with real-world expensive black-box problems would be computationally infeasible.
+Instead, we consider classical combinatorial optimization problems as black-box optimization problems.
+\ekhine{cambio la definition de LOP}{In this section, and following the consensus on the optimization literature, the permutations denote orderings.
+}
 
@@
 \begin{equation}\label{eq:lop}
-  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^{n} \sum_{j=1}^{i-1}  a_{\sigma^{-1}(i), \sigma^{-1}(j)}
+  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^{n} \sum_{j=1}^{i-1}  a_{\sigma(i), \sigma(j)}
 \end{equation}
@@
 %
-For the LOP, we generated synthetic instances as follows.
+%For the LOP, we generated synthetic instances as follows.
 First, we create a sample $S$ of 200 permutations of size $n=20$ from a MM with $\sigma_0$ equal to the identity permutation and $\theta=-\log(\textit{Uniformity})$, where  $\textit{Uniformity}\in\{0.5,$ $0.7,$ $0.9\}$, i.e., from peaked distributions to almost uniform distributions.
@@
 I wanted to explain why we use synthetic datasets.}
-
-Moreover, we consider real LOP datasets from the LOLIB
+We consider real LOP datasets from the LOLIB
 repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} In
@@
 
-\subsection{Analysis of UMM and CEGO on synthetic LOP instances}
-
-To better understand the behavior of the algorithms, we record the fitness of
-the solution evaluated at each step of each run and we average those values
-over 10 runs with different random seeds on the same problem instance.  As an
-example, each line in Fig.~\ref{fig:lop_synth} shows the mean fitness of the
-solutions evaluated by either UMM or CEGO on one synthetic LOP instance. The
-shaded area shows one standard deviation around the mean. Ideally, each new
-evaluation will monotonically improve in fitness, since each solution evaluated
-helps to refine the model and leads to a better solution being evaluated in the
-next step.  However, it is possible that new solutions are worse
-than their predecessors. %Nevertheless, in this example, there is a clear
-% overall improvement as more evaluations are performed.
-The minimum values
-achieved in the plot up to a particular number of function evaluations gives an
-estimation of the mean fitness of the best-so-far solution up to that point in
-the run. The algorithms do not return the final solution shown but rather the
-best one (minimum fitness) of each run.
-
-%: LOP syntéticas, todas en una gráfica. 
-\begin{figure}[tb]
-  \centering%
-  \includegraphics[width=\textwidth]{../img/synthetic_LOP_combined}
-  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
-\end{figure}
-
-When we compare the behavior of CEGO and
-UMM on synthetic LOP instances (Fig.~\ref{fig:lop_synth}), we observe a clear
-pattern independently of the uniformity of the instances. In particular, we see
-a linear convergence to the optimal value for UMM.  This is due to the linear
-decrease in the parameter $\theta$ (line~\ref{line:theta} in
-Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the
-internal probability model of the UMM over the $\FEmax$ functions
-evaluations. Overall, the mean fitness returned by UMM at $\FEmax=400$ is better than the one returned by  CEGO, regardless the uniformity of the instances.
-
-%UMM converges much more slowly than CEGO, only reaching the
-%same fitness values around evaluation 250.
-
-%Yet, UMM is able to keep the rate of
-%improvement and completely overtake CEGO around evaluation 350. As in the case
-%of CEGO, lower uniformity values lead to worse performance, although the
-%difference is much smaller for UMM than it was for CEGO.\MANUEL{If there is anything else we can say here, please go ahead and say it}
-
-On the other hand, % \EKHINE{cambio el orden}
-CEGO starts with 10 random initial solutions, % with are quite poor as
-% expected,
-followed by building and updating the Kriging model based on new
-evaluations, which leads to an impressive improvement in less than 50
-evaluations. However, progress is halted at this point and CEGO seems to
-have trouble finding better solutions. Around evaluation 200, the search
-becomes fully random. Lower uniformity values, e.g., 0.5, show worsening
-evaluations despite the increase in solutions used to build the model. In
-summary, the Kriging model clearly helps CEGO to quickly identify good
-solutions in very few evaluations. However, CEGO is not able
-to further improve those solutions given more data. Given what we know about
-Kriging~\citep{EriPeaGar2019scalable}, it appears that the surrogate model cannot model the local step
-necessary to reach the optimal solution and switches to an exploration strategy
-that ends up being completely random.
-
-% This may due to the model not being able to produce accurate
-% predictions or the underlying optimizer not being able to find improved
-% solutions using the model.
-%\MANUEL{perhaps we should measure the
-%  prediction error and see if it grows or decreases. That would answer this
-%  question.}
-
-%: LOP reales: he puesto 3 de diferentes tamanos.
+%\subsection{Analysis of UMM and CEGO on synthetic LOP instances}
+%
+%To better understand the behavior of the algorithms, we record the fitness of
+%the solution evaluated at each step of each run and we average those values
+%over 10 runs with different random seeds on the same problem instance.  As an
+%example, each line in Fig.~\ref{fig:lop_synth} shows the mean fitness of the
+%solutions evaluated by either UMM or CEGO on one synthetic LOP instance. The
+%shaded area shows one standard deviation around the mean. Ideally, each new
+%evaluation will monotonically improve in fitness, since each solution evaluated
+%helps to refine the model and leads to a better solution being evaluated in the
+%next step.  However, it is possible that new solutions are worse
+%than their predecessors. %Nevertheless, in this example, there is a clear
+%% overall improvement as more evaluations are performed.
+%The minimum values
+%achieved in the plot up to a particular number of function evaluations gives an
+%estimation of the mean fitness of the best-so-far solution up to that point in
+%the run. The algorithms do not return the final solution shown but rather the
+%best one (minimum fitness) of each run.
+%
+%
+%%: LOP syntéticas, todas en una gráfica. 
+%\begin{figure}[tb]
+%  \centering%
+%  \includegraphics[width=\textwidth]{../img/synthetic_LOP_combined}
+%  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
+%\end{figure}
+%
+%When we compare the behavior of CEGO and
+%UMM on synthetic LOP instances (Fig.~\ref{fig:lop_synth}), we observe a clear
+%pattern independently of the uniformity of the instances. In particular, we see
+%a linear convergence to the optimal value for UMM.  This is due to the linear
+%decrease in the parameter $\theta$ (line~\ref{line:theta} in
+%Algorithm~\ref{alg:umm}), which imposes a linear variance decrease on the
+%internal probability model of the UMM over the $\FEmax$ functions
+%evaluations. Overall, the mean fitness returned by UMM at $\FEmax=400$ is better than the one returned by  CEGO, regardless the uniformity of the instances.
+%
+%
+%
+%%UMM converges much more slowly than CEGO, only reaching the
+%%same fitness values around evaluation 250. 
+%
+%%Yet, UMM is able to keep the rate of
+%%improvement and completely overtake CEGO around evaluation 350. As in the case
+%%of CEGO, lower uniformity values lead to worse performance, although the
+%%difference is much smaller for UMM than it was for CEGO.\MANUEL{If there is anything else we can say here, please go ahead and say it}
+%
+%On the other hand, % \EKHINE{cambio el orden}
+%CEGO starts with 10 random initial solutions, % with are quite poor as
+%% expected,
+%followed by building and updating the Kriging model based on new
+%evaluations, which leads to an impressive improvement in less than 50
+%evaluations. However, progress is halted at this point and CEGO seems to
+%have trouble finding better solutions. Around evaluation 200, the search
+%becomes fully random. Lower uniformity values, e.g., 0.5, show worsening
+%evaluations despite the increase in the number of solutions used to build the model. In
+%summary, the Kriging model clearly helps CEGO to quickly identify good
+%solutions in very few evaluations. However, CEGO is not able
+%to further improve those solutions given more data. Given what we know about
+%Kriging~\citep{EriPeaGar2019scalable}, it appears that the surrogate model cannot model the local step
+%necessary to reach the optimal solution and switches to an exploration strategy
+%that ends up being completely random.
+%
+%% This may due to the model not being able to produce accurate
+%% predictions or the underlying optimizer not being able to find improved
+%% solutions using the model.
+%%\MANUEL{perhaps we should measure the
+%%  prediction error and see if it grows or decreases. That would answer this
+%%  question.}
+%
+%
+%
+%%: LOP reales: he puesto 3 de diferentes tamanos.
 Me ha parecido interesante que, a pesar de que CEGO parece que va mejor, esta diferencia se va reduciendo segun se aumenta el tamano del problema (que se puede ver en el nombre del problema y en el títutlo de la gráfica)
-\begin{figure}[!tb]
-  \centering%
-  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandB_N-p50-01}\\
-  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d100_01}\\
-  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d150_01}\\
-%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d200_01}\\
-    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOLIB.\label{fig:lolib}}
-  \end{figure}
+%\begin{figure}[!tb]
+%  \centering%
+%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandB_N-p50-01}\\
+%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d100_01}\\
+%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d150_01}\\
+%%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d200_01}\\
+%    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOLIB.\label{fig:lolib}}
+%  \end{figure}
+%
 
@@
 \subsection{Analysis of UMM and CEGO on LOP, QAP and PFSP}
-
-In the following paragraphs we analyze the results over real-world instances. %
-Figure~\ref{fig:lolib} shows the results on the instances taken from LOLIB.
+To  understand the behavior of the algorithms, we record the fitness of
+the solution evaluated at each step of each run and we average those values
+over 10 runs with different random seeds on the same problem instance.  As an
+example, each line in Fig.~\ref{fig:lop_synth} shows the mean fitness of the
+\ekhine{solutions evaluated by either UMM or CEGO on one synthetic LOP instance}{ojo}. The
+shaded area shows one standard deviation around the mean. Ideally, each new
+evaluation will monotonically improve in fitness, since each solution evaluated
+helps to refine the model and leads to a better solution being evaluated in the
+next step.  However, it is possible that new solutions are worse
+than their predecessors. %Nevertheless, in this example, there is a clear
+% overall improvement as more evaluations are performed.
+The minimum values
+achieved in the plot up to a particular number of function evaluations gives an
+estimation of the mean fitness of the best-so-far solution up to that point in
+the run. The algorithms do not return the final solution shown but rather the
+best one (minimum fitness) of each run.
+
+Figure~\ref{fig:lolib} shows the results on the real-world instances taken from LOLIB. 
   %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
   % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
-The type of instances differs greatly from the synthetic LOP instances, and,
-thus, so does the landscape. However, UMM shows a very similar behavior: there
+%The type of instances differs greatly from the synthetic LOP instances, and,
+%thus, so does the landscape. However, UMM shows a very similar behavior: 
+We can see that for UMM there 
 is a (close to) linear decrease of the fitness of new solutions over each new
 evaluation. This suggests that UMM is able to model the fitness
-landscape. Regarding CEGO, we observe a similar behavior to UMM for the first
+landscape. Regarding CEGO, \ekhine{we observe a similar behavior to UMM for the first
 50-100 evaluations, however, in several instances CEGO reaches the time limit
 of 144 hours, so no observations can be made for longer budgets. In smaller
-instances (e.g., Figure~\ref{fig:lolib}, top), CEGO again appears to stagnate.
-%Overall, the improvement from the first  to the last
-%solutions is 2-3 times larger in UMM than CEGO. %
-%\MANUEL{But this is a bit misleading because UMM runs for 4 times more
-%  evaluations, perhaps we should also measure up to the evaluation that CEGO
-%  reaches so we report both values?}
-%
-
-In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
+instances (e.g., Figure~\ref{fig:lolib}, top), CEGO again appears to stagnate.}{ the objective function decrease slowly?}
+
+\ekhine{In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
  obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP
@@
 the models (both Bayesian and probabilistic) are not learning anything about
-the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).
+the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{}
 % \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
@@
 
+\subsection{Rankings and orderings}
+%\ekhine{}{rellenar}
+In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
+In this lines, we show the consequences of this mistakes.
+We remind the readear that statistical models (Borda or the Kendall distance) are defined for rankings while standard literature in optimization considers orderings.
+As a summary of Algorithm~\ref{alg:umm}, we model rankings in our UMM model, generate a ranking $\sigma$, then invert this ranking to obtain an ordering in line~\ref{line:inversion} and evaluate the ordering $\sigma^{-1}$ in the objective function.
+A natural question is what happens if we did not invert the ranking.
+
+In CEGO, the objects modeled and evaluated are the same: There is no inversion between the objects sampled from the model and the objects evaluated.
+This means that botha re rankings or both are orders.
+For modelling the Kendall distance is used, that as stated is meant for rankings.
+A natural question is what happens if we did invert the ranking.
+
+Figure~\ref{fig:rankorder} answers the above two questions.
+In a setting as the one in previous sections, we find 4 different lines, for the evaluation of UMM and CEGO with and without inversion.
+We see that in case of No inversion (dashed lines) both UMM and CEGO do not seem to converge (despite we can see some decrease along iterations).
+When there is an inversion between modelling and evaluation both algorithms behave better.
+In this case, since UMM is set for a a linear decrease in the the scale parameter for the $m$ iterations, we see the linear decrease in the function evaluations.
+In CEGO, we see an abrupt decrease at the beggining which seems stuck at iteration 200.
+
+\begin{figure}[tp]
+  \centering%
+  \includegraphics[width=\textwidth]{../img/fitness_real_qap_kra32_dat}
+  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.}
+  \label{fig:rankorder}
+\end{figure}
+
+We compare this results with the numerical results provided in the original paper. These results are consistent with the setting of no inversion for CEGO. 
+\ekhine{probar con m=100 para ver ci podemos competir en instancias cortas}{}
+
 \section{Conclusions}\label{sec:conclusions}

### 2021-02-02 3-launch_exps.sh
More new results of CEGO m400-er0 and UMM beta=1

@@ -12,7 +12,8 @@ qsub_job() {
     # We would like to use $BASHPID here, but OS X version of bash does not
     # support it.
     ALGO=$1
-    shift 1
+    OUTPUT=$2
+    shift 2
     JOBNAME=${ALGO}-$counter-$$
     qsub -v PATH <<EOF
 #!/bin/bash --login
@@ -32,18 +33,19 @@ qsub_job() {
 #$ -cwd
 module load apps/anaconda3
 run=\$SGE_TASK_ID
-echo "running: ./target-runner-${ALGO}.py $ALGO $counter-$$-r\$run \$run $@ --output $RESULTS/$ALGO-r\$run"
-./target-runner-${ALGO}.py $ALGO $counter-$$-r\$run \$run $@ --output $RESULTS/$ALGO-r\$run
+echo "running: ./target-runner-${ALGO}.py $ALGO $counter-$$-r\$run \$run $@ --output ${OUTPUT}-r\$run"
+./target-runner-${ALGO}.py $ALGO $counter-$$-r\$run \$run $@ --output "${OUTPUT}-r\$run"
 EOF
 }
 
 # FIXME: Not working right now
 launch_local() {
+    ALGO=$1
+    OUTPUT=$2
+    shift 2
     for run in $(seq 1 $nruns); do
-	echo $1
-	shift 1
-	echo "running: ./target-runner-${ALGO}.py $ALGO $counter-$$-r$run $run $@ --output $RESULTS/$ALGO-r$run"
-	./target-runner-${ALGO}.py $ALGO $counter-$$-r$run $run $@ --output $RESULTS/$ALGO-r$run
+	echo "running: ./target-runner-${ALGO}.py $ALGO $counter-$$-r$run $run $@ --output ${OUTPUT}-r$run"
+	./target-runner-${ALGO}.py $ALGO $counter-$$-r$run $run $@ --output "${OUTPUT}-r$run"
     done
 }
 
@@ -95,8 +97,8 @@ pfsp/rec31.txt \
 
 
 ###### For LOLIB instances
-#INSTANCES="$INSTANCES $(grep -v '#' lolib-instances.txt | tr '\n' ' ')"
-INSTANCES="$(grep -v '#' lolib-instances.txt | tr '\n' ' ')"
+INSTANCES="$INSTANCES $(grep -v '#' lolib-instances.txt | tr '\n' ' ')"
[…]

### 2021-02-02 paper/main.tex
Fix matplotlib fonts in python notebooks

@@
 %150--250 words.
-  Black-box combinatorial optimization problems arise in practice when the
-  objective function is evaluated by means of a simulator or a real-world
-  experiment. In such cases, classical techniques such as mixed-integer
-  programming and local search cannot be applied. Moreover, each fitness
-  evaluation is often expensive in terms of time or resources, thus only a
-  limited number of evaluations is possible, typically several orders of
-  magnitude smaller than in white-box optimization problems. In the continuous
-  case, Bayesian optimization, in particular using Gaussian processes, has
-  proven very effective under these conditions. Much less research is available
-  in the combinatorial case. In this paper, we propose and analyze UMM, an
+  Expensive black-box combinatorial optimization problems arise in practice
+  when the objective function is evaluated by means of a simulator or a
+  real-world experiment. Since each fitness evaluation is expensive in terms of
+  time or resources, only a limited number of evaluations is possible,
+  typically several orders of magnitude smaller than in non-expensive problems. In
+  this scenario, classical optimization methods such as mixed-integer
+  programming and local search are not useful.  In the continuous case,
+  Bayesian optimization, in particular using Gaussian processes, has proven
+  very effective under these conditions. Much less research is available in the
+  combinatorial case. In this paper, we propose and analyze UMM, an
   estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
@@
   sometimes surpass, the solutions obtained by CEGO, a Bayesian optimization
-  algorithm for combinatorial optimization. %, specially when the budget of solutions evaluated approaches 400. 
-  Moreover, the computational complexity of UMM increases linearly with both the number of function evaluations and the permutation size.
+  algorithm for combinatorial
+  optimization. %, specially when the budget of solutions evaluated approaches 400.
+  Moreover, the computational complexity of UMM increases linearly with both
+  the number of function evaluations and the permutation size.
 %  an order of magnitude smaller than CEGO, which makes it specially useful when
@@
 ruggedness of combinatorial landscapes, which makes local search particularly
-effective in the white-box context, lessens the effectiveness of global
+effective for combinatorial problems, lessens the effectiveness of global
 surrogate models~\citep{EriPeaGar2019scalable}. Moreover, surrogate models are
@@
 %Finally, we note that learning Budget-constrained Combinatorial is a hot topic for a variety of structures besides permutation~\cite{manchanda2020}. 
-\manuel{- the authors assigned to black-box optimization problems characteristics that they do not own. For instance they say that "local search cannot be used".
+\MANUEL{- the authors assigned to black-box optimization problems characteristics that they do not own. For instance they say that "local search cannot be used".
  they mixed expensive to evaluate functions with black-box functions and of course they are different. In fact much of the metaheuristics do not mind
@@
  that they have a black-box problem where the objective function is costly to evaluate. This should be clear and I would recommend to delete other comments
- related with white-box and black-box.
-}{}
-
-\ekhine{6.
+ related with white-box and black-box. MANUEL: FIXED
+}
+
+\EKHINE{6.
 When the main contribution of the paper is discussed it is stated that this contribution is the Unbalanced Mallows *model* (which does not seem to be true as the models used in the paper seem to be known already).
 Right after that it is stated that the contribution is *a novel algorithm* which, I think, reflects the contents of the paper much better.
-}{}
-In this work, we propose and analyze Unbalanced Mallows Model (UMM), a
+}
+In this work, we propose and analyze Unbalanced Mallows Model (UMM)\MANUEL{Perhaps we need to rename the algorithm to UMM Algorithm or UMM Optimizer to avoid the criticism of the reviewer that we do not propose the model?}, a
 population-based probabilistic algorithm specifically designed for optimizing
@@
 exists a wide body of research on probabilistic models for permutations, such
-as Mallows~\cite{FliVer1986}, that have already been been used in optimization
+as Mallows~\cite{FliVer1986}, that have already been  used in optimization
 problems successfully, with the most prominent examples being estimation of

### 2021-02-02 paper/main.tex
Update to GECCO format.

@@
-\documentclass[runningheads]{llncs}
+\documentclass[sigconf,dvipsnames]{acmart}
 \usepackage[english]{babel}
@@
 \usepackage{afterpage,tabularx}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage[square,numbers,sort]{natbib}
-% Fixes for using natbib and llncs
-\makeatletter
-% required for natbib to have "References" printed and as section*, not chapter*
-\renewcommand\bibsection%
-{\section*{\refname\@mkboth{\MakeUppercase{\refname}}{\MakeUppercase{\refname}}}}
-\renewcommand\@biblabel[1]{#1.}
-\def\bibfont{\small}
-\makeatother
-
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-\renewcommand\UrlFont{\color{blue}\rmfamily}
+%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
 
@@
 %% Revision documents
-\usepackage[dvipsnames]{xcolor}
+%\usepackage[dvipsnames]{xcolor}
 \usepackage[normalem]{ulem}
@@
 
+% Copyright
+%\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
+
+% DOI
+\acmDOI{10.1145/nnnnnnn.nnnnnnn}
+
+% ISBN
+\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
+
+% Conference
+\acmConference[GECCO '21]{the Genetic and Evolutionary Computation Conference 2021}{July 10--14, 2021}{Lille, France}
+\acmYear{2021}
+\copyrightyear{2021}
+
+%\acmArticle{4}
+\acmPrice{15.00}
+
+\begin{document}
 \title{Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems}
-\titlerunning{Unbalanced Mallows Models for Expensive Black-Box Permutation Problems}
 % Double-blind
 \author{Double-blind}
+% \authornote{Dr.~Trovato insisted his name be first.}
+% \orcid{1234-5678-9012}
+% \affiliation{%
+%   \institution{Institute for Clarity in Documentation}
+%   \streetaddress{P.O. Box 1212}
+%   \city{Dublin} 
+%   \state{Ohio} 
+%   \postcode{43017-6221}
+% }
+% \email{trovato@corporation.com}
+
 % \author{Ekhiñe Irurozki\inst{1} \and Manuel López-Ibáñez\inst{2}\orcidID{0000-0001-9974-1295}}
@@
 %  }
-\date{}%
-
-\begin{document}
-
-\maketitle
-
+%    \renewcommand{\shortauthors}{B. Trovato et al.}
 \begin{abstract}
@@
   %\textcolor{red}{16 pages max, deadline: 1 November 2020}
-  \keywords{Combinatorial optimization \and Bayesian optimization \and
-    Expensive black-box optimization \and Estimation of distribution
-    algorithms}
 \end{abstract}
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below. 
+%
+\begin{CCSXML}
+<ccs2012>
+ <concept>
+  <concept_id>10010520.10010553.10010562</concept_id>
+  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+  <concept_significance>500</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010575.10010755</concept_id>
+  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+  <concept_significance>300</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010553.10010554</concept_id>
+  <concept_desc>Computer systems organization~Robotics</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10003033.10003083.10003095</concept_id>
+  <concept_desc>Networks~Network reliability</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+</ccs2012>  
+\end{CCSXML}
+
+\ccsdesc[500]{Computer systems organization~Embedded systems}
+\ccsdesc[300]{Computer systems organization~Redundancy}
+\ccsdesc{Computer systems organization~Robotics}
+\ccsdesc[100]{Networks~Network reliability}
+
+\keywords{Combinatorial optimization, Bayesian optimization, Expensive
+  black-box optimization, Estimation of distribution algorithms}
+
+\maketitle
 
@@
 \end{minipage}
-\begin{minipage}{0.49\linewidth}
-  \centering
-  \texttt{kra32} (QAP)\\[-0.5ex]
-\includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
-\end{minipage}
+% \begin{minipage}{0.49\linewidth}
+%   \centering
+%   \texttt{kra32} (QAP)\\[-0.5ex]
+% \includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
+% \end{minipage}
 \caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better). Whitespace denotes missing data (runs that took longer than 6 hours).\label{fig:heatmaps}}
@@
 
-\renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}
-\bibliographystyle{splncs04natnodoi}
+%\renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}
+\bibliographystyle{ACM-Reference-Format-abbrv}
 \bibliography{optbib/abbrev,optbib/authors,optbib/abbrevshort,optbib/journals,optbib/biblio,optbib/crossref}%,./mendeley}

### 2021-02-02 paper/main.tex
more updates

@@
   model and unbalanced rank aggregation (uBorda). Experimental results on
-  black-box versions of LOP, QAP and PFSP show that UMM is able to match, and
+  black-box versions of LOP and PFSP show that UMM is able to match, and
   sometimes surpass, the solutions obtained by CEGO, a Bayesian optimization
@@
 The set of all permutations of $n$ values is denoted as $S_n$ and has cardinality $n!$.
-We will use the one-line notation and denote permutations with Greek letters, $\sigma  = (\sigma(1), \ldots , \sigma(n))$.
-We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.
-
-Permutation $\sigma$ represents a ranking of item:  $\sigma(i)=j$ denotes that item $i$ has rank $j$.
-The ordering associated with $\sigma$ is given by its inverse $\sigma^{-1}$, therefore,  $\sigma(i)=j \Leftrightarrow  \sigma^{-1}(j)=i$.
-This distinction is in order since statistical methodologies are devoted for rankings and optimizaiton literature for orderings and mixing both concepts is a very common mistake.
-We will illustrate in the experimental section the consequences of this confusion.
+We will use the one-line notation\MANUEL{what is one-line?} and denote permutations with Greek letters, $\sigma  = (\sigma(1), \ldots , \sigma(n))$.
+We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi = (\sigma(\pi(1)), \ldots, \sigma(\pi(n)))$\MANUEL{Is this correct?} and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.
+
+\manuel{P}{Throughout this paper and for consistency, p}ermutation $\sigma$ represents a \emph{ranking}:  $\sigma(i)=j$ denotes that item $i$ has rank $j$.
+The \emph{ordering} associated with $\sigma$ is given by its inverse $\sigma^{-1}$, therefore,  $\sigma(i)=j \Leftrightarrow  \sigma^{-1}(j)=i$.
+This distinction is in order since statistical methodologies \manuel{are devoted for}{typically assume} rankings \manuel{and}{whereas the} optimization literature \manuel{for}{often assumes} orderings.
+Mixing both concepts is a very common confusion with grave consequences, as we will illustrate in the experimental section.
 
 Kendall's-$\tau$ distance is the metric for rankings that counts the pairs of items ranked in different order in both permutations:
-
+%
 \begin{equation}

### 2021-02-02 paper/main.tex
more updates on paper

@@
 \DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 
-Kendall's-$\tau$ distance is the metric for rankings that counts the pairs of items ranked in different order in both permutations:
+Kendall's-$\tau$ distance is the metric for rankings that counts the pairs of items ranked in different order in two permutations:
 %
@@
 An equivalent definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made to convert $\sigma_1^{-1}$ into $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
-\ekhine{ojo, aqui lo dice bien o  mal?}{}
+\EKHINE{ojo, aqui lo dice bien o  mal?}\MANUEL{Habría que mirarlo bien!}
 
@@
 The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay.
-The normalization constant $\psi$ can be easily computed for the Kendall's-$\tau$ distance (Eq.~\eqref{eq:kendall_decomp}) as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
+The normalization constant $\psi$ can be easily computed for the Kendall's-$\tau$ distance (Eq.~\ref{eq:kendall_decomp}) as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.
 
@@
 In this section we describe our main contribution: Unbalanced Mallows model
-(UMM), a novel algorithm for the optimization of a black-box function over
+(UMM), a novel algorithm for the optimization of an expensive black-box function over
 permutation spaces. This algorithm is a probabilistic method based on the
@@
 
-\ekhine{10.
+\EKHINE{10.
 In Algorithm 1 'S' and 'F' are *sets* (thus, unordered).
@@
 You always add these elements to both sets at the same time in step 8, anyway.
-}{}
+}
+
+\newcommand{\tuple}[1]{\ensuremath{\langle #1\rangle}}
 
 \begin{algorithm}[t]
- \caption{UMM: Unbalanced Mallows Model for Black-Box Optimization}
+ \caption{UMM: Unbalanced Mallows Model Algorithm}
  \label{alg:umm}
   \begin{algorithmic}[1]
-    \Require $\minit$: number of initial permutations, $\FEmax$: total budget, $\rho$: learning rate
-    %\\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
-    \State $S \assign $ generate $\minit$ permutations uniformly at random
-    \State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
+    \Require $\minit$: number of initial permutations, $\FEmax$: total budget,\newline\hspace*{5ex} \manuel{$\rho$: learning rate,}{$r_1, r_2$: parameters controlling the learning rate} 
+    % \\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
+    \State $S \assign \{\}$
+    \For{$i \assign 0$ \textbf{to} $\minit$}
+    \State $\sigma \assign \text{generate uniformly at random}$
+    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
+    \EndFor
+    %\State $S \assign $ generate  $\minit$ permutations uniformly at random
+    %\State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
     \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
     \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}\Comment{Learning Step}
-    \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}\Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
-    \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}\Comment{Generate $\beta$ permutations}
-    \State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \Comment{Select the farthest from $S$}\label{line:distance}
+    \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}
+    \Statex \Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
+    \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
+    % \Comment{Generate $\beta$ permutations}
+    \State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \label{line:distance}
+    \Statex \Comment{Select the farthest from $S$}
    % \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
-    \State $S \assign S \cup \{\sigma\}$, $\quad F \assign F \cup \{f(\sigma)\}$
+    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
     \EndFor 
-    \State \Return $S$, $F$
+    \State \Return $\tuple{\sigma^\ast, f(\sigma^\ast)}\assign \argmin_{\tuple{\sigma, f(\sigma)}\in S} f(\sigma)$
   %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values

### 2021-02-02 mallows_kendall.py

@@ -6,13 +6,16 @@ from scipy import optimize
 def kendall_max_dist(n):
     return int(n * (n - 1) / 2)
 
+# FIXME: Optimize this function
 def kendallTau(A, B=None):
     # if any partial is B
     if B is None : B = list(range(len(A)))
     n = len(A)
+    # FIXME: Since n is typically fixed, this could be done once outside this function.
     pairs = it.combinations(range(n), 2)
     distance = 0
     # print("IIIIMNNMNNN",list(pairs),len(A))
+    # FIXME: Avoid this loop: https://stackoverflow.com/questions/16003217/n-d-version-of-itertools-combinations-in-numpy
     for x, y in pairs:
         #if not A[x]!=A[x] and not A[y]!=A[y]:#OJO no se check B
         a = A[x] - A[y]

### 2021-02-02 paper/main.tex
more updates to the paper

@@
 permutation spaces. This algorithm is a probabilistic method based on the
-Mallows model and uBorda learning.
+Mallows model and uBorda learning.\MANUEL{Perhaps we should call it UMM Algorithm? otherwise people would complain. Or Unbalanced Mallows Algorithm (UMA)?}
 
@@
 You always add these elements to both sets at the same time in step 8, anyway.
+MANUEL: FIXED
 }
@@
 The process is repeated up to a given maximum budget of evaluations $\FEmax$.
-\ekhine{(3) In section 3, second half of the second paragraph (from "Then, we sample a new candidate [...]"): the algorithm 1 is not properly described.
+\EKHINE{(3) In section 3, second half of the second paragraph (from "Then, we sample a new candidate [...]"): the algorithm 1 is not properly described.
 Please provide a description in line with algorithm 1 and, above all, describe the rationale behind the choice of line 7 of algorithm 1.
@@
 I know this is quite usual in the problems analyzed by the authors, but it may not be so usual in general.
-Please add a clarification.}{}
+Please add a clarification.}
 
@@
 
-\subsubsection{Learning step}
+\subsection{Learning step}
 %
@@
 $w(\sigma)=\rho^{f(\sigma)}$ (Section~\ref{sec:uborda}). The learning rate is
-controlled by parameter $\rho \in (0,1)$. %
+controlled by parameter $0 < \rho < 1$. \MANUEL{Don't we have to say something about $f(\sigma) > 1$?}
 In UMM, the value of $\rho$ depends on the set of true fitness evaluations $F$
 and it is set at each iteration such that the largest $100r_1$\% of the mass of
-the weights is concentrated in the best $100r_2$\% of the solutions in $S$, i.e.,\MANUEL{Reviewer 3:  Justify the form of the formula used to learn the "learning rate" $\rho$ (equation 4).
-Why this particular approach was chosen?}
+the weights is concentrated in the best $100r_2$\% of the solutions in $S$:
+%
+\MANUEL{Reviewer 3:  Justify the form of the formula used to learn the "learning rate" $\rho$ (equation 4). Why this particular approach was chosen?}
 %
@@
 
-\ekhine{4. Justify the form of the formula used to learn the "learning rate" $\rho$ (equation 4). Why this particular approach was chosen?}{}
-
-\subsubsection{Sampling step}\label{sec:sampling}
+\subsection{Sampling step}\label{sec:sampling}
 In this step, the algorithm samples from the distribution the new data point to be evaluated. 
@@
 
-\ekhine{5. "At the first iteration, Theta is set so that $E[D] = {n \choose 4}$". Why?}{}
+\EKHINE{5. "At the first iteration, Theta is set so that $E[D] = {n \choose 4}$". Why?}
 A purely explotation approach would always generate $\sigma_0$ in the sampling stage.
@@
 
-\ekhine{The algorithm uses a parameter $\beta$ that determines the number of points sampled, while the authors analyze the sensitivity of the results to the parameters r 1 and r 2, they do not do that with $\beta$.
-In my opinion the value of $\beta$ is quite relevant.}{}
+\EKHINE{The algorithm uses a parameter $\beta$ that determines the number of points sampled, while the authors analyze the sensitivity of the results to the parameters r 1 and r 2, they do not do that with $\beta$.
+In my opinion the value of $\beta$ is quite relevant.}
 
@@
 \ekhine{}{completar}
+
+\renewcommand{\subsubsection}[1]{\smallskip{}\noindent\textbf{#1}.}
 
@@
 
-\subsubsection{CEGO} \ekhine{tenemos una pagina entera dedicada a CEGO :s}{}
+\subsubsection{CEGO} \EKHINE{tenemos una pagina entera dedicada a CEGO :s}\MANUEL{Sugiere qué quitar usando el commando \textbackslash{}ekhine}
 Combinatorial Efficient Global Optimization (CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
@@
 it as the best-performing~\citep{ZaeStoFriFisNauBar2014}.
-
-\renewcommand{\subsubsection}[1]{\smallskip{}\noindent\textbf{#1}.}
 
@@
 
+\subsubsection{Reproducibility.} Source code, datasets and scripts necessary to reproduce the results are available at \supplement.
+%  \vspace*{-1em}
+%\begin{acks}
 %% MANUEL: Blind, so no acknowledgments yet.
-\begin{smaller}
-  \paragraph*{Reproducibility.} Source code, datasets and scripts necessary to reproduce the results are available at \supplement.
-  \vspace*{-1em}
-\paragraph*{Acknowledgements.} Blind
+%\paragraph*{Acknowledgements.} Blind
 % M.\@ L\'opez-Ib\'a\~nez is a ``Beatriz Galindo'' Senior Distinguished Researcher (BEAGAL 18/00053) funded by the Ministry of Science and Innovation of the Spanish Government.
@@
 %about differences between QAPLIB instances; and Ahmed Boujaada for the experimental support. 
-
-\end{smaller}
+  % supported by the \grantsponsor{GS501100001809}{National Natural
+  %   Science Foundation of
+  %   China}{http://dx.doi.org/10.13039/501100001809} under Grant
+  % No.:~\grantnum{GS501100001809}{61273304}
+  % and~\grantnum[http://www.nnsf.cn/youngscientists]{GS501100001809}{Young
+  %   Scientists' Support Program}.
+%\end{acks}
 

### 2021-02-02 paper/main.tex
more updates to the paper

@@
 %
-UMM uses the same budget of evaluations ($\FEmax$) as CEGO
-and initially samples $\minit=10$ solutions uniformly. The other two parameters
-of UMM are $r_1$ and $r_2$, which determine the balance between exploration and
-exploitation. In the experimental section we will study their impact. %implemented UMM using Python 3.
-At each sampling step, we set $\beta=10$, meaning that 10 permutations are generated and the farthest from $S$ is selected for evaluation.
+UMM uses the same budget of evaluations ($\FEmax$) as CEGO and initially
+samples $\minit=10$ solutions uniformly.  At each sampling step, we set
+$\beta=10$, meaning that 10 permutations are generated and the farthest from
+$S$ is selected for evaluation.  The other two parameters of UMM are $r_1$ and
+$r_2$, which determine the balance between exploration and exploitation. In the
+experimental section we will study their impact.
 
@@
 For simplicity, we use Kendall's-$\tau$
-distance~(Eq.~\ref{eq:kendall_decomp}) in all experiments. Nevertheless,
-\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
-when using Hamming distance and the same may apply to UMM.
+distance~(Eq.~\ref{eq:kendall_decomp}) in all experiments.
+%Nevertheless,
+%\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
+%when using Hamming distance and the same may apply to UMM.
 We plan to extend UMM to other distance measures in the future.  In all
@@
 actual objective function. In a white-box context, state-of-the-art algorithms
-for the problems considered here typically evaluate thousands of solutions,
+for the LOP and PFSP typically evaluate thousands of solutions,
 thus, the budget considered here for the black-box context is extremely
@@
 \ekhine{cambio la definition de LOP}{In this section, and following the consensus on the optimization literature, the permutations denote orderings.
-}
+}\MANUEL{That seems strange, shouldn't be use the same everywhere?
+I would rather use rankings or orderings everywhere and I have a slight preference for orders since this is an optimization paper.}
 
@@
 \end{equation}
-where $[a_{i,j}]$ is a matrix of size $n \times n$ and $\sigma^{-1}$ is the inverse of $\sigma$.
-
-The \emph{Quadratic Assignment Problem (QAP)} is defined by:
-\begin{equation}\label{eq:qap}
-  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^n \sum_{j=1}^n a_{i,j} b_{\sigma(i), \sigma(j)}
-\end{equation}
-%
-where $[a_{i,j}]$ and $[b_{i,j}]$ are two matrices of size $n \times n$.
+%
+where $[a_{i,j}]$ is a matrix of size $n \times n$ \manuel{and $\sigma^{-1}$ is the inverse of $\sigma$}{}.
+
+% The \emph{Quadratic Assignment Problem (QAP)} is defined as
+% \begin{equation}\label{eq:qap}
+%   \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^n \sum_{j=1}^n a_{i,j} b_{\sigma(i), \sigma(j)}
+% \end{equation}
+% %
+% where $[a_{i,j}]$ and $[b_{i,j}]$ are two matrices of size $n \times n$.
 

### 2021-02-03 paper/main.tex
update heatmaps

@@
 repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} In
-particular, we consider the following types: IO, Random instances of type A and
+particular, we have considered two instances of each class, focusing on relatively instances (the computation of CEGO grows rapidly with instance size):
+\texttt{N-p40-01}, \texttt{N-p40-02}, \texttt{N-p50-01}, \texttt{N-p50-02} (RandB), 
+lop/RandB/
+lop/IO/N-t59d11xx
+lop/IO/N-t59b11xx
+lop/SGB/N-sgb75.02
+lop/SGB/N-sgb75.01
+RandB/
+RandB/N-p50-02
+
+: IO, Random instances of type A and
 Random instances of type B of various sizes $n \in \{50, 100, 150, 200,
@@
 because of the larger search space, but also because of the different instance
-types.
-
-In the case of the QAP and PFSP, we consider the same instances as
-\citet{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, i.e., \texttt{nug12},
-\texttt{nug30}, \texttt{tho30} and \texttt{kra32} for the QAP, with $n$ (the
-permutation length) equal to the number in the instance name, and
-\texttt{reC05}, \texttt{reC13} and \texttt{reC19} for the PFSP, with
-$n \in \{20, 20, 30\}$, respectively. For reasons that will become obvious
-later, we also consider QAP instances \texttt{kra30a} and \texttt{kra30b}.
+types.\MANUEL{Needs updating}
+
+In the case of the PFSP, we consider the same instances as
+\citet{ZaeStoBar2014:ppsn}, i.e., \texttt{reC05}, \texttt{reC13} and
+\texttt{reC19} for the PFSP, with $n \in \{20, 20, 30\}$, respectively.
+
+% In the case of the QAP and PFSP, we consider the same instances as
+% \citet{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, i.e., \texttt{nug12},
+% \texttt{nug30}, \texttt{tho30} and \texttt{kra32} for the QAP, with $n$ (the
+% permutation length) equal to the number in the instance name, and
+% \texttt{reC05}, \texttt{reC13} and \texttt{reC19} for the PFSP, with
+% $n \in \{20, 20, 30\}$, respectively. For reasons that will become obvious
+% later, we also consider QAP instances \texttt{kra30a} and \texttt{kra30b}.
 
@@
   \begin{minipage}{0.49\linewidth}\centering
-    \texttt{N-t1d100.01} (LOP)\\[-0.2ex]
-    \includegraphics[width=\textwidth]{../img/heatmap_N-t1d100_01}
+    \texttt{N-t59d11xx} (LOP)\\[-0.2ex]
+    \includegraphics[width=\textwidth]{../img/heatmap_lop_IO_N-t59d11xx}
   \end{minipage}
   \begin{minipage}{0.49\linewidth}\centering
-    \texttt{N-t1d50.01} (LOP)\\[-0.2ex]
- \includegraphics[width=\textwidth]{../img/heatmap_N-t1d150_01}
+    \texttt{SGB_N-sgb75_02} (LOP)\\[-0.2ex]
+ \includegraphics[width=\textwidth]{../img/heatmap_lop_SGB_N-sgb75_02}
 \end{minipage}\\\vskip 1em
@@
   \centering
-  \texttt{reC05} (PFSP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/heatmap_rec05}
+  \texttt{reC13} (PFSP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec13_txt}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+  \centering
+  \texttt{reC19} (PFSP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec19_txt}
 \end{minipage}
@@
 % \end{minipage}
-\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better). Whitespace denotes missing data (runs that took longer than 6 hours).\label{fig:heatmaps}}
+\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better).}\label{fig:heatmaps}}
 \end{figure}

### 2021-02-03 paper/main.tex
update heatmaps
updates

@@
 We terminated runs that took longer than 6 hours and those combinations are shown as white.
-Results for all the datasets can be found in the supplemental material~(\supplement).
+Results for all the datasets can be found in the supplemental material~(\supplement).\MANUEL{TODO: Update (mine)}
 
@@
 \centering%
-  \begin{minipage}{0.49\linewidth}\centering
+\begin{minipage}{0.49\linewidth}
+  \centering
     \texttt{N-t59d11xx} (LOP)\\[-0.2ex]
@@
   \end{minipage}
-  \begin{minipage}{0.49\linewidth}\centering
-    \texttt{SGB_N-sgb75_02} (LOP)\\[-0.2ex]
+  \begin{minipage}{0.49\linewidth}
+    \centering
+    \texttt{N-sgb75.02} (LOP)\\[-0.2ex]
  \includegraphics[width=\textwidth]{../img/heatmap_lop_SGB_N-sgb75_02}
@@
 % \end{minipage}
-\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better).}\label{fig:heatmaps}}
+\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better).\label{fig:heatmaps}}
 \end{figure}
@@
 \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
-\begin{table}[tb]
+\begin{table*}[tb]
   \caption{Mean fitness (and standard deviation) of the best solution found.
@@
 \vspace*{-1em}
- \end{table}
+ \end{table*}
 
 \subsection{Rankings and orderings}
-%\ekhine{}{rellenar}
 In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
 In this lines, we show the consequences of this mistakes.
-We remind the readear that statistical models (Borda or the Kendall distance) are defined for rankings while standard literature in optimization considers orderings.
+We remind the reader that statistical models (Borda or the Kendall distance) are defined for rankings while standard literature in optimization considers orderings.
 As a summary of Algorithm~\ref{alg:umm}, we model rankings in our UMM model, generate a ranking $\sigma$, then invert this ranking to obtain an ordering in line~\ref{line:inversion} and evaluate the ordering $\sigma^{-1}$ in the objective function.
 A natural question is what happens if we did not invert the ranking.
+%\ekhine{}{rellenar}
 

### 2021-02-03 paper/main.tex
paper

@@
 repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} In
-particular, we have considered two instances of each class, focusing on relatively instances (the computation of CEGO grows rapidly with instance size):
-\texttt{N-p40-01}, \texttt{N-p40-02}, \texttt{N-p50-01}, \texttt{N-p50-02} (RandB), 
-lop/RandB/
-lop/IO/N-t59d11xx
-lop/IO/N-t59b11xx
-lop/SGB/N-sgb75.02
-lop/SGB/N-sgb75.01
-RandB/
-RandB/N-p50-02
-
-: IO, Random instances of type A and
-Random instances of type B of various sizes $n \in \{50, 100, 150, 200,
-250\}$. These are harder problems than the synthetic instances, not only
-because of the larger search space, but also because of the different instance
-types.\MANUEL{Needs updating}
-
+particular, we have taken two instances of various types focusing on relatively
+small instances, since he computation time of CEGO grows rapidly with instance
+size: \texttt{N-p40-01}, \texttt{N-p40-02}, \texttt{N-p50-01} and
+\texttt{N-p50-02} of type \emph{RandB}; \texttt{N-t59d11xx} and
+\texttt{N-t59b11xx} of type \emph{IO}; and \texttt{N-sgb75.01} and
+\texttt{N-sgb75.02} of type \emph{SGB}.
+% Random instances of type B of various sizes $n \in \{50, 100, 150, 200,
+% 250\}$. These are harder problems than the synthetic instances, not only
+% because of the larger search space, but also because of the different instance
+% types.
 In the case of the PFSP, we consider the same instances as
-\citet{ZaeStoBar2014:ppsn}, i.e., \texttt{reC05}, \texttt{reC13} and
-\texttt{reC19} for the PFSP, with $n \in \{20, 20, 30\}$, respectively.
+\citet{ZaeStoBar2014:ppsn}, i.e., \texttt{reC05}, \texttt{reC13},
+\texttt{reC19} and \texttt{rec31}, with $n \in \{20, 20, 30, 50\}$ and $M \in \{5, 20, 10, 15\}$, respectively.
 
@@
 
-The first part of the experimental analysis concerns the {analyses} of the learning rate, which corresponds to parameter $\rho$ of uBorda.
+The first part of the experimental analysis concerns the analysis of the learning rate, which corresponds to parameter $\rho$ of uBorda.
 As stated in Section~\ref{sec:sampling}, $\rho$ is automatically set at each iteration in such a way that the $100r_1\%$ of the samples have the $100r_2\%$ of the weight.
@@
 Thus, we tried values $r_1 = \{0.1, 0.2, \dotsc, 0.5\}$ and $r_2 = \{0.6,\dotsc, 0.9, 0.99\}$.
-Figure~\ref{fig:heatmaps} shows the best solution found by each independent run (averaged over 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP.
-We terminated runs that took longer than 6 hours and those combinations are shown as white.
-Results for all the datasets can be found in the supplemental material~(\supplement).\MANUEL{TODO: Update (mine)}
+Figure~\ref{fig:heatmaps} shows the best solution found by each independent run (averaged over 10 runs) for instances: \texttt{N-t59d11xx} and \texttt{N-sgb75.02} for the LOP, and \texttt{reC13} and \texttt{reC19} for the PFSP.
+Results for all the datasets can be found in the supplemental material~(\supplement).
 
@@
   \centering%
+  \smaller
 \begin{minipage}{0.49\linewidth}
@@
 
-\subsection{Analysis of UMM and CEGO on LOP, QAP and PFSP}
+\subsection{Analysis of UMM and CEGO on LOP and PFSP}
 To  understand the behavior of the algorithms, we record the fitness of
@@
 
-The situation is very similar for the QAP, where both algorithms show the same
-behaviour as in the PFSP for all instances tested, except for
-\texttt{kra} instances. As shown in Fig~\ref{fig:kra32}, for \texttt{kra32} both
-algorithms are able to improve over the number of evaluations, although not
-with ease. As in the synthetic LOP instances, we again observe that CEGO does
-improve faster than UMM initially, yet CEGO appears to get stuck around
-evaluation 250, and UMM overtakes it around evaluation 300. We believe the
-reason is the same as before, that is, the CEGO model is not able to adequately
-reflect the ruggedness of the actual landscape and, hence, the best solutions
-found by the GA searching the model are not improved solutions for the actual
-problem. Interestingly, there are well-known differences between
-\texttt{kra30a}, \texttt{kra30b}, \texttt{kra32}, on one hand, and \texttt{nug12}, \texttt{nug30} and
-\texttt{tho32}, on the other~\citep{AnsBriGou2002qap}. However, we cannot at
-this moment point out the root cause of this different behavior.
-
-\begin{figure}[tp]
-  \centering%
-  \includegraphics[width=\textwidth]{../img/fitness_real_qap_kra32_dat}
-  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.\label{fig:kra32}}
-  \vspace*{-1em}
-\end{figure}
+% The situation is very similar for the QAP, where both algorithms show the same
+% behaviour as in the PFSP for all instances tested, except for
+% \texttt{kra} instances. As shown in Fig~\ref{fig:kra32}, for \texttt{kra32} both
+% algorithms are able to improve over the number of evaluations, although not
+% with ease. As in the synthetic LOP instances, we again observe that CEGO does
+% improve faster than UMM initially, yet CEGO appears to get stuck around
+% evaluation 250, and UMM overtakes it around evaluation 300. We believe the
+% reason is the same as before, that is, the CEGO model is not able to adequately
+% reflect the ruggedness of the actual landscape and, hence, the best solutions
+% found by the GA searching the model are not improved solutions for the actual
+% problem. Interestingly, there are well-known differences between
+% \texttt{kra30a}, \texttt{kra30b}, \texttt{kra32}, on one hand, and \texttt{nug12}, \texttt{nug30} and
+% \texttt{tho32}, on the other~\citep{AnsBriGou2002qap}. However, we cannot at
+% this moment point out the root cause of this different behavior.
+
+% \begin{figure}[tp]
+%   \centering%
+%   \includegraphics[width=\textwidth]{../img/fitness_real_qap_kra32_dat}
+%   \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.\label{fig:kra32}}
+%   \vspace*{-1em}
+% \end{figure}
 
@@
 significant, with the qualification that both algorithms appear to be
-performing a random search. Similar observations can be made for the QAP, where
-only on the smallest instance \texttt{nug12}, there is a significant difference
-in favour of CEGO.
+performing a random search.
 
@@
 
-\begin{figure}[tp]
-  \centering%
-  \includegraphics[width=\textwidth]{../img/fitness_real_qap_kra32_dat}
-  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.}
-  \label{fig:rankorder}
-\end{figure}
-
 We compare this results with the numerical results provided in the original paper. These results are consistent with the setting of no inversion for CEGO. 
@@
 
-\subsubsection{Reproducibility.} Source code, datasets and scripts necessary to reproduce the results are available at \supplement.
+\subsubsection{Reproducibility} Source code, datasets and scripts necessary to reproduce the results are available at \supplement.
 %  \vspace*{-1em}

### 2021-02-03 1-tuning.sh
More new results of CEGO m400-er0

@@ -11,11 +11,10 @@ qsub_job() {
     PARALLEL_ENV=smp.pe
     # We would like to use $BASHPID here, but OS X version of bash does not
     # support it.
-    JOBNAME=umm-$counter-r$run-$$
-    OUTFILE="${RESULTS}/$JOBNAME"
-    
+    JOBNAME=umm-$counter-$$
     qsub -v PATH <<EOF
 #!/bin/bash --login
+#$ -t 1-$nruns
 #$ -N $JOBNAME
 # -pe $PARALLEL_ENV $NB_PARALLEL_PROCESS 
 #$ -l ivybridge
@@ -30,18 +29,22 @@ qsub_job() {
 #$ -j y
 #$ -cwd
 module load apps/anaconda3
-echo "running: $BINDIR/target-runner-umm.py umm $counter-r$run-$$ $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --eval_ranks $eval_ranks"
-echo -n "$instance,$run,$umm_m_ini,$budgetMM,$r_1,$r_2,$budget,$eval_ranks,$value" > $OUTFILE
-$BINDIR/target-runner-umm.py umm $counter-r$run-$$ $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --eval_ranks $eval_ranks >> $OUTFILE
+run=\$SGE_TASK_ID
+OUTFILE="${RESULTS}/$JOBNAME-r\$run"
+echo "running: $BINDIR/target-runner-umm.py umm $counter-$$-r\$run \$run $@ > \$OUTFILE"
+echo -n "$instance,\$run,$umm_m_ini,$budgetMM,$r_1,$r_2,$budget,$eval_ranks," > \$OUTFILE
+$BINDIR/target-runner-umm.py umm $counter-$$-r\$run \$run $@ >> \$OUTFILE
 EOF
 
 }
 launch_local() {
-    JOBNAME=umm-$counter-r$run-$$
-    OUTFILE=$RESULTS/$JOBNAME
-    echo "running: $BINDIR/target-runner-umm.py umm $counter-r$run-$$ $run $instance --m_ini $umm_m_ini --budgetMM $budgetMM --rsl $r_1 --wml $r_2 --budget $budget --eval_ranks $eval_ranks"
[…]

### 2021-02-03 3-launch_exps.sh

@@ -33,8 +33,8 @@ qsub_job() {
 #$ -cwd
 module load apps/anaconda3
 run=\$SGE_TASK_ID
-echo "running: ./target-runner-${ALGO}.py $ALGO $counter-$$-r\$run \$run $@ --output ${OUTPUT}-r\$run"
-./target-runner-${ALGO}.py $ALGO $counter-$$-r\$run \$run $@ --output "${OUTPUT}-r\$run"
+echo "running: ${BINDIR}/target-runner-${ALGO}.py $ALGO $counter-$$-r\$run \$run $@ --output ${OUTPUT}-r\$run"
+${BINDIR}/target-runner-${ALGO}.py $ALGO $counter-$$-r\$run \$run $@ --output "${OUTPUT}-r\$run"
 EOF
 }
 
@@ -44,8 +44,8 @@ launch_local() {
     OUTPUT=$2
     shift 2
     for run in $(seq 1 $nruns); do
-	echo "running: ./target-runner-${ALGO}.py $ALGO $counter-$$-r$run $run $@ --output ${OUTPUT}-r$run"
-	./target-runner-${ALGO}.py $ALGO $counter-$$-r$run $run $@ --output "${OUTPUT}-r$run"
+	echo "running: ${BINDIR}/target-runner-${ALGO}.py $ALGO $counter-$$-r$run $run $@ --output ${OUTPUT}-r$run"
+	${BINDIR}/target-runner-${ALGO}.py $ALGO $counter-$$-r$run $run $@ --output "${OUTPUT}-r$run"
     done
 }
 
@@ -98,12 +98,12 @@ pfsp/rec31.txt \
 
 ###### For LOLIB instances
 INSTANCES="$INSTANCES $(grep -v '#' lolib-instances.txt | tr '\n' ' ')"
-#INSTANCES="$(grep -v '#' lolib-instances.txt | tr '\n' ' ')"
+INSTANCES="$(grep -v '#' lolib-instances.txt | tr '\n' ' ')"
 
 ###### Synthetic LOP instances
 #INSTANCES="$INSTANCES $(gen_lop_synthetic)"
 
-budget=400
+budget=100
 
 #eval_ranks=1
 eval_ranks="0 1"
@@ -115,18 +115,20 @@ budgetGA=4
 
 r_1=0.1
 r_2=0.9
-#budgetMM=10
-budgetMM=1
+beta=10
+#beta=1
 umm_m_ini=10
 
 counter=0
-for er in $eval_ranks; do
[…]

### 2021-02-03 paper/main.tex
paper table

@@
 \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
-\begin{table*}[tb]
-  \caption{Mean fitness (and standard deviation) of the best solution found.
+%\begin{table*}[tb]
+%  \caption{Mean fitness (and standard deviation) of the best solution found.
 % at the end of each run, over 10 independent runs for each problem instance.
 The 95\% confidence interval corresponds to the two independent samples t-test on the mean difference between the fitness of UMM minus the one of CEGO.
-  Mean runtime is shown in hours. 
-  %CEGO runs were stopped after 144 hours. This was the case for all LOP instances.
-  \label{tab:results}}
-  \resizebox{\textwidth}{!}{%
-\begin{tabular}{r@{\hskip -2ex}*{5}{r}rl@{\hskip -2ex}*{3}{r}}
+%  Mean runtime is shown in hours. 
+%  %CEGO runs were stopped after 144 hours. This was the case for all LOP instances.
+%  \label{tab:results}}
+%  \resizebox{\textwidth}{!}{%
+%\begin{tabular}{r@{\hskip -2ex}*{5}{r}rl@{\hskip -2ex}*{3}{r}}
+% \toprule
+%            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack[r]{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){10-11}
+%            \bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule
+
+\begin{table*}[tb] % resumns
+\begin{tabular}{l@{\hskip -2ex}*{5}{r}rl@{\hskip -2ex}*{3}{r}}%llrlrllrrr}
 \toprule
+%Problem &    Instance &  CEGO\_Fitness\_mean & CEGO\_Fitness\_std &  UMM\_Fitness\_mean & UMM\_Fitness\_std &    Confidence interval &  CEGO F. evaluations &  CEGO Runtime &  UMM Runtime \\
             &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack[r]{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){10-11}
             \bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule
-LOP
-& N-p50-01       & 21034.3             & (183.9)            & 21161.8            & (657.4)           & [-606.0,             & 351.0]              & 305.8        & 144.6 & 0.5  \\
-& N-p50-02       & 21767.3             & (254.9)            & 21746.6            & (645.4)           & [-458.6,             & 500.0]              & 315.1        & 144.5 & 0.5  \\
-& N-t1d100.01    & 78012.0             & (432.1)            & 76499.3            & (1099.7)          & [696.4,              & 2329.0]             & 158.3        & 145.1 & 1.9  \\
-& N-t1d100.02    & 78834.0             & (477.8)            & 76860.7            & (864.0)           & [1303.8,             & 2642.8]             & 156.0        & 144.7 & 1.8  \\
-& N-atp111       & 668.5               & (9.4)              & 596.6              & (17.0)            & [58.8,               & 85.0]               & 137.9        & 145.1 & 2.3  \\
-& N-atp134       & 856.4               & (12.2)             & 760.1              & (23.4)            & [78.4,               & 114.2]              & 116.5        & 145.0 & 3.3  \\
-& N-t1d150.01    & 186969.0            & (2094.6)           & 179615.6           & (1731.5)          & [5543.3,             & 9163.5]             & 103.9        & 145.7 & 4.4  \\
-& N-t1d150.02    & 183414.7            & (779.2)            & 177524.2           & (1165.6)          & [4949.2,             & 6831.8]             & 103.6        & 145.3 & 4.6  \\
-& N-t2d150.01    & 22207.3             & (1667.3)           & 15136.9            & (745.9)           & [5817.1,             & 8323.7]             & 104.5        & 145.5 & 4.5  \\
-& N-t2d150.02    & 22036.0             & (744.7)            & 14718.4            & (677.5)           & [6648.3,             & 7986.9]             & 103.1        & 145.0 & 4.5  \\
-& N-t1d200.01    & 333804.1            & (1191.9)           & 323501.4           & (1385.7)          & [9086.4,             & 11519.0]            & 77.7         & 145.4 & 7.6  \\
-& N-t1d200.02    & 331730.9            & (2819.7)           & 319786.2           & (2265.6)          & [9533.6,             & 14355.8]            & 78.2         & 145.9 & 7.1  \\
-& N-t2d200.01    & 54043.6             & (1477.6)           & 33389.0            & (1984.0)          & [19001.4,            & 22307.8]            & 78.2         & 146.7 & 7.4  \\
-& N-t2d200.02    & 52787.0             & (4194.2)           & 35589.0            & (2319.6)          & [13948.0,            & 20448.0]            & 78.9         & 145.7 & 7.5  \\
-& N-be75np\_150  & 3669753.9           & (74292.3)          & 3216907.1          & (72239.3)         & [383998.5,           & 521695.1]           & 105.2        & 145.5 & 4.4  \\
-& N-be75np\_250  & 10818828.5          & (127722.3)         & 9582261.7          & (258176.8)        & [1040022.9,          & 1433110.7]          & 62.5         & 146.4 & 12.5 \\
-& N-be75eec\_150 & 1592220.8           & (29945.9)          & 1418578.9          & (56557.7)         & [130141.0,           & 217142.8]           & 103.2        & 145.9 & 4.4  \\
-& N-be75eec\_250 & 5098734.4           & (172874.0)         & 4430408.5          & (110233.0)        & [530352.4,           & 806299.4]           & 63.0         & 146.7 & 12.3 \\\midrule
-PFSP
-& rec05          & 1317.9              & (25.1)             & 1332.5             & (15.6)            & [-34.5,              & 5.3]                & 400.0        & 38.6  & 0.1  \\
-& rec13          & 2128.2              & (25.0)             & 2163.6             & (28.1)            & [-60.4,              & -10.4]              & 400.0        & 38.6  & 0.1  \\
-& rec19          & 2398.6              & (19.8)             & 2405.2             & (12.8)            & [-22.5,              & 9.3]                & 400.0        & 85.2  & 0.2  \\\midrule
-QAP
-& nug12          & 624.2               & (26.3)             & 661.2              & (17.8)            & [-58.3,              & -15.7]              & 400.0        & 14.2  & 0.0  \\
-& kra30b         & 120192.0            & (1778.1)           & 119923.0           & (3459.6)          & [-2379.5,            & 2917.5]             & 400.0        & 85.1  & 0.2  \\
-& tho30          & 192416.0            & (2584.2)           & 191586.4           & (4342.0)          & [-2582.9,            & 4242.1]             & 400.0        & 85.6  & 0.2  \\
-& kra30a         & 119872.0            & (1746.7)           & 119502.0           & (1827.5)          & [-1309.8,            & 2049.8]             & 400.0        & 86.1  & 0.2  \\
-& nug30          & 7486.8              & (40.4)             & 7472.4             & (83.6)            & [-49.0,              & 77.8]               & 400.0        & 86.5  & 0.2  \\
-& kra32          & 115907.0            & (4663.4)           & 116898.0           & (4343.1)          & [-5226.3,            & 3244.3]             & 400.0        & 94.0  & 0.2  \\
+\midrule
+    LOP &    N-p40-01 &            10496.6 &          (116.8) &           12369.1 &         (389.9) &     [-2157.1, -1587.9] &                400.0 &          19.9 &          0.3 \\
+    LOP &    N-p40-02 &            11714.7 &           (84.7) &           13160.5 &         (454.8) &     [-1773.5, -1118.1] &                400.0 &          20.1 &          0.3 \\
+    LOP &  N-t59b11xx &           102408.0 &         (3800.2) &          114075.5 &        (5389.2) &    [-16084.2, -7250.8] &                400.0 &          21.1 &          0.4 \\
+    LOP &  N-t59d11xx &            12692.2 &         (1097.9) &           16662.9 &        (2155.3) &     [-5618.5, -2322.9] &                400.0 &          21.2 &          0.4 \\
+    LOP &    N-p50-02 &            19090.5 &          (285.4) &           21746.6 &         (645.4) &     [-3140.6, -2171.6] &                400.0 &          22.9 &          0.5 \\
+    LOP &    N-p50-01 &            18197.7 &          (225.6) &           21161.8 &         (657.4) &     [-3447.4, -2480.8] &                400.0 &          23.2 &          0.5 \\
+    LOP &  N-sgb75.01 &          1893533.2 &        (28594.6) &         1992059.9 &       (40235.7) &  [-131577.1, -65476.3] &                400.0 &          33.2 &          1.0 \\
+    LOP &  N-sgb75.02 &          1895801.8 &        (82847.9) &         1967132.3 &       (49228.1) &   [-136421.2, -6239.8] &                400.0 &          33.3 &          1.1 \\
+   PFSP &       rec05 &             1276.2 &           (15.1) &            1276.3 &           (8.4) &          [-11.8, 11.6] &                400.0 &          18.3 &          0.1 \\
+   PFSP &       rec13 &             1992.9 &           (27.6) &            2057.8 &          (22.8) &         [-88.8, -41.0] &                400.0 &          18.5 &          0.1 \\
+   PFSP &       rec19 &             2264.5 &           (45.0) &            2304.8 &          (37.7) &          [-79.4, -1.2] &                400.0 &          18.2 &          0.2 \\
+   PFSP &       rec31 &             3415.5 &           (62.1) &            3419.3 &          (37.0) &          [-52.6, 45.0] &                400.0 &          23.1 &          0.5 \\
 \bottomrule
-\end{tabular}}
-\vspace*{-1em}
+\end{tabular} 
+\end{table*}
+
+\begin{table*}[tb]%time
+\begin{tabular}{llrrrrrr}
+\toprule
+Problem &    Instance &  CEGO\_inv &  CEGO\_orig &  UMM\_100 &  UMM\_200 &  UMM\_400 &  UMM\_no\_inv \\
+\midrule
+    LOP &  N-t59b11xx &   1265.98 &    1240.36 &     1.37 &     5.78 &    22.24 &       21.89 \\
+    LOP &  N-t59d11xx &   1269.93 &    1252.89 &     1.37 &     5.62 &    21.84 &       22.30 \\
+    LOP &    N-p40-01 &   1195.90 &    1170.81 &     1.14 &     4.61 &    18.13 &       19.06 \\
+    LOP &    N-p40-02 &   1207.24 &    1183.09 &     1.15 &     4.76 &    18.11 &       19.06 \\
+    LOP &    N-p50-01 &   1391.55 &        NaN &      NaN &      NaN &    27.36 &         NaN \\
+    LOP &    N-p50-02 &   1372.71 &        NaN &      NaN &      NaN &    28.36 &         NaN \\
+    LOP &  N-sgb75.01 &   1994.22 &    2004.70 &     3.97 &    15.77 &    62.75 &       65.40 \\
+    LOP &  N-sgb75.02 &   1995.09 &    1975.41 &     4.03 &    15.90 &    64.58 &       62.69 \\
+   PFSP &       rec05 &   1099.63 &     917.81 &     0.30 &     1.14 &     4.54 &        4.53 \\
+   PFSP &       rec13 &   1109.51 &    1043.08 &     0.31 &     1.07 &     4.55 &        4.50 \\
+   PFSP &       rec19 &   1090.06 &    1087.38 &     0.63 &     2.59 &    10.51 &       10.35 \\
+   PFSP &       rec31 &   1387.09 &    1367.26 &     1.79 &     6.90 &    27.84 &       26.91 \\
+\bottomrule
+\end{tabular}
+\end{table*}
+
+\begin{table*}[tb]%fitness
+\begin{tabular}{llrrrrrr}
+\toprule
+Problem &    Instance &   CEGO\_inv &   CEGO\_orig &   UMM\_400 &   UMM\_200 &   UMM\_100 &   UMM\_no\_inv \\
+\midrule
+    LOP &  N-t59b11xx &          102408.0 &           126204.0 &         114075.5 &         115941.5 &         128195.5 &            138819.5 \\
+    LOP &  N-t59d11xx &           12692.2 &            30199.8 &          16662.9 &          22117.4 &          28050.4 &             42971.2 \\
+    LOP &    N-p40-01 &           10496.6 &            15027.7 &          12369.1 &          13043.7 &          14459.3 &             16471.5 \\
+    LOP &    N-p40-02 &           11714.7 &            15150.2 &          13160.5 &          13923.2 &          15027.7 &             16330.7 \\
+    LOP &    N-p50-01 &           18197.7 &                NaN &          21161.8 &              NaN &              NaN &                 NaN \\
+    LOP &    N-p50-02 &           19090.5 &                NaN &          21746.6 &              NaN &              NaN &                 NaN \\
+    LOP &  N-sgb75.01 &         1893533.2 &          2095899.6 &        1992059.9 &        2011271.7 &        2074111.6 &           2111368.5 \\
+    LOP &  N-sgb75.02 &         1895801.8 &          2069286.9 &        1967132.3 &        2011004.2 &        2045556.5 &           2056681.8 \\
+   PFSP &       rec05 &            1276.2 &             1318.7 &           1276.3 &           1283.1 &           1300.4 &              1332.5 \\
+   PFSP &       rec13 &            1992.9 &             2096.0 &           2057.8 &           2113.6 &           2134.9 &              2163.6 \\
+   PFSP &       rec19 &            2264.5 &             2328.4 &           2304.8 &           2346.5 &           2392.4 &              2405.2 \\
+   PFSP &       rec31 &            3415.5 &             3512.3 &           3419.3 &           3438.9 &           3511.9 &              3554.1 \\
+\bottomrule
+\end{tabular}
 \end{table*}

### 2021-02-03 paper/main.tex
paper table

@@
 \usepackage{multirow}
+
+% \usepackage{multirow}
+\usepackage{xcolor}
 
@@
 
-\newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
-%\begin{table*}[tb]
-%  \caption{Mean fitness (and standard deviation) of the best solution found.
-% at the end of each run, over 10 independent runs for each problem instance.
-The 95\% confidence interval corresponds to the two independent samples t-test on the mean difference between the fitness of UMM minus the one of CEGO.
-%  Mean runtime is shown in hours. 
-%  %CEGO runs were stopped after 144 hours. This was the case for all LOP instances.
-%  \label{tab:results}}
-%  \resizebox{\textwidth}{!}{%
-%\begin{tabular}{r@{\hskip -2ex}*{5}{r}rl@{\hskip -2ex}*{3}{r}}
-% \toprule
-%            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack[r]{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){10-11}
-%            \bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule
-
-\begin{table*}[tb] % resumns
-\begin{tabular}{l@{\hskip -2ex}*{5}{r}rl@{\hskip -2ex}*{3}{r}}%llrlrllrrr}
-\toprule
-%Problem &    Instance &  CEGO\_Fitness\_mean & CEGO\_Fitness\_std &  UMM\_Fitness\_mean & UMM\_Fitness\_std &    Confidence interval &  CEGO F. evaluations &  CEGO Runtime &  UMM Runtime \\
-            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack[r]{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){10-11}
-            \bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule
-\midrule
-    LOP &    N-p40-01 &            10496.6 &          (116.8) &           12369.1 &         (389.9) &     [-2157.1, -1587.9] &                400.0 &          19.9 &          0.3 \\
-    LOP &    N-p40-02 &            11714.7 &           (84.7) &           13160.5 &         (454.8) &     [-1773.5, -1118.1] &                400.0 &          20.1 &          0.3 \\
-    LOP &  N-t59b11xx &           102408.0 &         (3800.2) &          114075.5 &        (5389.2) &    [-16084.2, -7250.8] &                400.0 &          21.1 &          0.4 \\
-    LOP &  N-t59d11xx &            12692.2 &         (1097.9) &           16662.9 &        (2155.3) &     [-5618.5, -2322.9] &                400.0 &          21.2 &          0.4 \\
-    LOP &    N-p50-02 &            19090.5 &          (285.4) &           21746.6 &         (645.4) &     [-3140.6, -2171.6] &                400.0 &          22.9 &          0.5 \\
-    LOP &    N-p50-01 &            18197.7 &          (225.6) &           21161.8 &         (657.4) &     [-3447.4, -2480.8] &                400.0 &          23.2 &          0.5 \\
-    LOP &  N-sgb75.01 &          1893533.2 &        (28594.6) &         1992059.9 &       (40235.7) &  [-131577.1, -65476.3] &                400.0 &          33.2 &          1.0 \\
-    LOP &  N-sgb75.02 &          1895801.8 &        (82847.9) &         1967132.3 &       (49228.1) &   [-136421.2, -6239.8] &                400.0 &          33.3 &          1.1 \\
-   PFSP &       rec05 &             1276.2 &           (15.1) &            1276.3 &           (8.4) &          [-11.8, 11.6] &                400.0 &          18.3 &          0.1 \\
-   PFSP &       rec13 &             1992.9 &           (27.6) &            2057.8 &          (22.8) &         [-88.8, -41.0] &                400.0 &          18.5 &          0.1 \\
-   PFSP &       rec19 &             2264.5 &           (45.0) &            2304.8 &          (37.7) &          [-79.4, -1.2] &                400.0 &          18.2 &          0.2 \\
-   PFSP &       rec31 &             3415.5 &           (62.1) &            3419.3 &          (37.0) &          [-52.6, 45.0] &                400.0 &          23.1 &          0.5 \\
-\bottomrule
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{table*}[]
+\begin{tabular}{rrrrrrrrr}
+\multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{4}{c}{\textbf{Mean fitness (sd)}}                       & \multicolumn{1}{c}{\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}95\% CI of the \\ mean difference\end{tabular}}}} & \multicolumn{2}{c}{\textbf{Runtime}} \\ \cline{3-6} \cline{8-9}
+\textbf{Problem}              & \textbf{Instance}             & \multicolumn{2}{r}{\textbf{CEGO}} & \multicolumn{2}{r}{\textbf{UMM}} & \multicolumn{1}{c}{}                                                                                                    & \textbf{CEGO}     & \textbf{UMM}     \\ \hline
+LOP                           & N-p40-01                      & 10496.6          & (116.8)        & 12369.1         & (389.9)        & {[}-2157.1, -1587.9{]}                                                                                                  & 19.9              & 0.3              \\
+LOP                           & N-p40-02                      & 11714.7          & (84.7)         & 13160.5         & (454.8)        & {[}-1773.5, -1118.1{]}                                                                                                  & 20.1              & 0.3              \\
+LOP                           & N-t59b11xx                    & 102408.0         & (3800.2)       & 114075.5        & (5389.2)       & {[}-16084.2, -7250.8{]}                                                                                                 & 21.1              & 0.4              \\
+LOP                           & N-t59d11xx                    & 12692.2          & (1097.9)       & 16662.9         & (2155.3)       & {[}-5618.5, -2322.9{]}                                                                                                  & 21.2              & 0.4              \\
+LOP                           & N-p50-02                      & 19090.5          & (285.4)        & 21746.6         & (645.4)        & {[}-3140.6, -2171.6{]}                                                                                                  & 22.9              & 0.5              \\
+LOP                           & N-p50-01                      & 18197.7          & (225.6)        & 21161.8         & (657.4)        & {[}-3447.4, -2480.8{]}                                                                                                  & 23.2              & 0.5              \\
+LOP                           & N-sgb75.01                    & 1893533.2        & (28594.6)      & 1992059.9       & (40235.7)      & {[}-131577.1, -65476.3{]}                                                                                               & 33.2              & 1.0              \\
+LOP                           & N-sgb75.02                    & 1895801.8        & (82847.9)      & 1967132.3       & (49228.1)      & {[}-136421.2, -6239.8{]}                                                                                                & 33.3              & 1.1              \\
+PFSP                          & rec05                         & 1276.2           & (15.1)         & 1276.3          & (8.4)          & {[}-11.8, 11.6{]}                                                                                                       & 18.3              & 0.1              \\
+PFSP                          & rec13                         & 1992.9           & (27.6)         & 2057.8          & (22.8)         & {[}-88.8, -41.0{]}                                                                                                      & 18.5              & 0.1              \\
+PFSP                          & rec19                         & 2264.5           & (45.0)         & 2304.8          & (37.7)         & {[}-79.4, -1.2{]}                                                                                                       & 18.2              & 0.2              \\
+PFSP                          & rec31                         & 3415.5           & (62.1)         & 3419.3          & (37.0)         & {[}-52.6, 45.0{]}                                                                                                       & 23.1              & 0.5
 \end{tabular}

### 2021-02-03 paper/main.tex
paper table

@@
 \begin{table*}[tb]%time
-\begin{tabular}{llrrrrrr}
+\begin{tabular}{llrrrrr}
 \toprule
-Problem &    Instance &  CEGO\_inv &  CEGO\_orig &  UMM\_100 &  UMM\_200 &  UMM\_400 &  UMM\_no\_inv \\
+Problem &    Instance &  CEGO\_inv &  CEGO\_orig &  UMM\_400 &  UMM\_200 &  UMM\_100 \\
 \midrule
-    LOP &  N-t59b11xx &   1265.98 &    1240.36 &     1.37 &     5.78 &    22.24 &       21.89 \\
-    LOP &  N-t59d11xx &   1269.93 &    1252.89 &     1.37 &     5.62 &    21.84 &       22.30 \\
-    LOP &    N-p40-01 &   1195.90 &    1170.81 &     1.14 &     4.61 &    18.13 &       19.06 \\
-    LOP &    N-p40-02 &   1207.24 &    1183.09 &     1.15 &     4.76 &    18.11 &       19.06 \\
-    LOP &    N-p50-01 &   1391.55 &        NaN &      NaN &      NaN &    27.36 &         NaN \\
-    LOP &    N-p50-02 &   1372.71 &        NaN &      NaN &      NaN &    28.36 &         NaN \\
-    LOP &  N-sgb75.01 &   1994.22 &    2004.70 &     3.97 &    15.77 &    62.75 &       65.40 \\
-    LOP &  N-sgb75.02 &   1995.09 &    1975.41 &     4.03 &    15.90 &    64.58 &       62.69 \\
-   PFSP &       rec05 &   1099.63 &     917.81 &     0.30 &     1.14 &     4.54 &        4.53 \\
-   PFSP &       rec13 &   1109.51 &    1043.08 &     0.31 &     1.07 &     4.55 &        4.50 \\
-   PFSP &       rec19 &   1090.06 &    1087.38 &     0.63 &     2.59 &    10.51 &       10.35 \\
-   PFSP &       rec31 &   1387.09 &    1367.26 &     1.79 &     6.90 &    27.84 &       26.91 \\
+    LOP &  N-t59b11xx &   1265.98 &    1240.36 &    22.24 &     5.78 &     1.37 \\
+    LOP &  N-t59d11xx &   1269.93 &    1252.89 &    21.84 &     5.62 &     1.37 \\
+    LOP &    N-p40-01 &   1195.90 &    1170.81 &    18.13 &     4.61 &     1.14 \\
+    LOP &    N-p40-02 &   1207.24 &    1183.09 &    18.11 &     4.76 &     1.15 \\
+    LOP &    N-p50-01 &   1391.55 &        NaN &    27.36 &      NaN &      NaN \\
+    LOP &    N-p50-02 &   1372.71 &        NaN &    28.36 &      NaN &      NaN \\
+    LOP &  N-sgb75.01 &   1994.22 &    2004.70 &    62.75 &    15.77 &     3.97 \\
+    LOP &  N-sgb75.02 &   1995.09 &    1975.41 &    64.58 &    15.90 &     4.03 \\
+   PFSP &       rec05 &   1099.63 &     917.81 &     4.54 &     1.14 &     0.30 \\
+   PFSP &       rec13 &   1109.51 &    1043.08 &     4.55 &     1.07 &     0.31 \\
+   PFSP &       rec19 &   1090.06 &    1087.38 &    10.51 &     2.59 &     0.63 \\
+   PFSP &       rec31 &   1387.09 &    1367.26 &    27.84 &     6.90 &     1.79 \\
 \bottomrule
@@
 \begin{table*}[tb]%fitness
-\begin{tabular}{llrrrrrr}
+\begin{tabular}{llrrrrr}
 \toprule
-Problem &    Instance &   CEGO\_inv &   CEGO\_orig &   UMM\_400 &   UMM\_200 &   UMM\_100 &   UMM\_no\_inv \\
+Problem &    Instance &   CEGO\_inv &   CEGO\_orig &   UMM\_400 &   UMM\_200 &   UMM\_100 \\
 \midrule
-    LOP &  N-t59b11xx &          102408.0 &           126204.0 &         114075.5 &         115941.5 &         128195.5 &            138819.5 \\
-    LOP &  N-t59d11xx &           12692.2 &            30199.8 &          16662.9 &          22117.4 &          28050.4 &             42971.2 \\
-    LOP &    N-p40-01 &           10496.6 &            15027.7 &          12369.1 &          13043.7 &          14459.3 &             16471.5 \\
-    LOP &    N-p40-02 &           11714.7 &            15150.2 &          13160.5 &          13923.2 &          15027.7 &             16330.7 \\
-    LOP &    N-p50-01 &           18197.7 &                NaN &          21161.8 &              NaN &              NaN &                 NaN \\
-    LOP &    N-p50-02 &           19090.5 &                NaN &          21746.6 &              NaN &              NaN &                 NaN \\
-    LOP &  N-sgb75.01 &         1893533.2 &          2095899.6 &        1992059.9 &        2011271.7 &        2074111.6 &           2111368.5 \\
-    LOP &  N-sgb75.02 &         1895801.8 &          2069286.9 &        1967132.3 &        2011004.2 &        2045556.5 &           2056681.8 \\
-   PFSP &       rec05 &            1276.2 &             1318.7 &           1276.3 &           1283.1 &           1300.4 &              1332.5 \\
-   PFSP &       rec13 &            1992.9 &             2096.0 &           2057.8 &           2113.6 &           2134.9 &              2163.6 \\
-   PFSP &       rec19 &            2264.5 &             2328.4 &           2304.8 &           2346.5 &           2392.4 &              2405.2 \\
-   PFSP &       rec31 &            3415.5 &             3512.3 &           3419.3 &           3438.9 &           3511.9 &              3554.1 \\
+    LOP &  N-t59b11xx &          102408.0 &           126204.0 &         114075.5 &         115941.5 &         128195.5 \\
+    LOP &  N-t59d11xx &           12692.2 &            30199.8 &          16662.9 &          22117.4 &          28050.4 \\
+    LOP &    N-p40-01 &           10496.6 &            15027.7 &          12369.1 &          13043.7 &          14459.3 \\
+    LOP &    N-p40-02 &           11714.7 &            15150.2 &          13160.5 &          13923.2 &          15027.7 \\
+    LOP &    N-p50-01 &           18197.7 &                NaN &          21161.8 &              NaN &              NaN \\
+    LOP &    N-p50-02 &           19090.5 &                NaN &          21746.6 &              NaN &              NaN \\
+    LOP &  N-sgb75.01 &         1893533.2 &          2095899.6 &        1992059.9 &        2011271.7 &        2074111.6 \\
+    LOP &  N-sgb75.02 &         1895801.8 &          2069286.9 &        1967132.3 &        2011004.2 &        2045556.5 \\
+   PFSP &       rec05 &            1276.2 &             1318.7 &           1276.3 &           1283.1 &           1300.4 \\
+   PFSP &       rec13 &            1992.9 &             2096.0 &           2057.8 &           2113.6 &           2134.9 \\
+   PFSP &       rec19 &            2264.5 &             2328.4 &           2304.8 &           2346.5 &           2392.4 \\
+   PFSP &       rec31 &            3415.5 &             3512.3 &           3419.3 &           3438.9 &           3511.9 \\
 \bottomrule

### 2021-02-04 3-launch_exps.sh
New results of umm with beta=1

@@ -103,11 +103,11 @@ INSTANCES="$(grep -v '#' lolib-instances.txt | tr '\n' ' ')"
 ###### Synthetic LOP instances
 #INSTANCES="$INSTANCES $(gen_lop_synthetic)"
 
-budget=100
+budget=400
 
-#eval_ranks=1
 eval_ranks="0 1"
 eval_ranks=0
+eval_ranks=1
 
 cego_m_ini=10
 # Actually, 10**budgetGA
@@ -126,8 +126,8 @@ for m in $budget; do
 	    counter=$((counter+1))
 	    RESULTS="$OUTDIR/results/m${m}-er${er}/$instance"
 	    mkdir -p "$RESULTS"
-	    $LAUNCHER umm "${RESULTS}/umm-b${beta}" $instance --m_ini $umm_m_ini --budgetMM $beta --rsl $r_1 --wml $r_2 --budget $m --eval_ranks $er
-	    #$LAUNCHER cego "${RESULTS}/cego" $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $m --eval_ranks $er
+	    #$LAUNCHER umm "${RESULTS}/umm-b${beta}" $instance --m_ini $umm_m_ini --budgetMM $beta --rsl $r_1 --wml $r_2 --budget $m --eval_ranks $er
+	    $LAUNCHER cego "${RESULTS}/cego" $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $m --eval_ranks $er
 	done
     done
 done

### 2021-02-04 paper/main.tex
minor fix
update zenodo link
update css concepts

@@
 
-\newcommand{\supplement}{\href{http://doi.org/10.5281/zenodo.4167148}{doi:10.5281/zenodo.4167148}}
+\newcommand{\supplement}{\href{http://doi.org/10.5281/zenodo.4500974}{doi:~10.5281/zenodo.4500974}}
 
@@
 <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+<concept_id>10002950.10003624.10003625.10003630</concept_id>
+<concept_desc>Mathematics of computing~Combinatorial optimization</concept_desc>
 <concept_significance>500</concept_significance>
@@
 <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
+<concept_id>10003752.10010061.10011795</concept_id>
+<concept_desc>Theory of computation~Random search heuristics</concept_desc>
+<concept_significance>500</concept_significance>
 </concept>
@@
 
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
+\ccsdesc[500]{Mathematics of computing~Combinatorial optimization}
+\ccsdesc[500]{Theory of computation~Random search heuristics}
 
@@
     \text{s.t.}\quad & C_{1,j} = 0 \quad j\in [M], \qquad C_{i,1} = 0\quad i \in [n]\\
-    &C_{i,j} = p_{\sigma(i),j} + \max\{C_{i-1, j}, C_{i, j-1}\}  \ \  i \in \{2,\dotsc,n\},\ j \in\{2,\dotsc, M\}\\
+    &C_{i,j} = p_{\sigma(i),j} + \max\{C_{i-1, j}, C_{i, j-1}\}  \ \  i \in \{2,\dotsc,n\},\\
+    &\hspace{14.5em} j \in\{2,\dotsc, M\}
   \end{split}

### 2021-02-04 paper/main.tex
paper s1-2

@@
 \newcommand{\supplement}{\href{http://doi.org/10.5281/zenodo.4500974}{doi:~10.5281/zenodo.4500974}}
-
+\newcommand{\ken}{Kendall's-$\tau$}
 %% Revision documents
@@
 
-\EKHINE{6.
-When the main contribution of the paper is discussed it is stated that this contribution is the Unbalanced Mallows *model* (which does not seem to be true as the models used in the paper seem to be known already).
-Right after that it is stated that the contribution is *a novel algorithm* which, I think, reflects the contents of the paper much better.
-}
-In this work, we propose and analyze Unbalanced Mallows Model (UMM)\MANUEL{Perhaps we need to rename the algorithm to UMM Algorithm or UMM Optimizer to avoid the criticism of the reviewer that we do not propose the model?}, a
+In this work, we propose and analyze the Unbalanced Mallows Model (UMM) algorithm, a
 population-based probabilistic algorithm specifically designed for optimizing
-black-box problems over a permutation landscape under a limited budget. There
+black-box problems over a permutation landscape under a limited budget. There\EKHINE{En realidad está disenado para esto  pero podríamos probarlo para otros casos. Trabajo futuro}
 exists a wide body of research on probabilistic models for permutations, such
@@
 An equivalent definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made to convert $\sigma_1^{-1}$ into $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
-\EKHINE{ojo, aqui lo dice bien o  mal?}\MANUEL{Habría que mirarlo bien!}
+\EKHINE{Mirandolo bien, lo que pasa es que dicen que es lo mismo la swap distance que la kendall!}
 
@@
 Parameter $\theta$ is set automatically by the algorithm to control the diversification and intensification trade off.
-At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$ (half of the expected distance for the uniform distribution),\MANUEL{Reviewer 3: "At the first iteration, $\theta$ is set so that $E[D] = {n \choose 4}$".
-Why?} and then the expected distance linearly decreases until the last iteration where $\theta$ is set such that $\Edist = 1$.
+At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$, i.e., half of the expected distance for the uniform distribution.
+Then the expected distance linearly decreases until the last iteration where $\theta$ is set such that $\Edist = 1$.
 Despite it is not possible to isolate $\theta$ in Eq.~\eqref{eq:expectation}, its monotonicity allows the use of bisection methods for efficiently finding an approximation.
 
-\EKHINE{5. "At the first iteration, Theta is set so that $E[D] = {n \choose 4}$". Why?}
-A purely explotation approach would always generate $\sigma_0$ in the sampling stage.
+Setting $\Edist = \binom{n}{2}$ will generate random permutations at early stages, slowing the algorithm convergence. 
+A purely explotation approach will set $\Edist = 0$  generating $\sigma_0$ in the sampling stage.
 However, the setting of a decreasing variance   controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
+It follows that the behaviour is different for different choices of $m$.
 Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.
@@
 the models (both Bayesian and probabilistic) are not learning anything about
-the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{}
+the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{kaka}
 % \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
@@
 
-\begin{figure}[tb]
+\begin{figure*}[tb]
   \centering%
-  \includegraphics[width=\textwidth]{../img/fitness_real_pfsp_rec05_txt}
+    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
+    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
+
+    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p40-01}
+    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p50-01}
+
+    \includegraphics[width=0.49\textwidth]{../img/fitness_pfsp_rec19_txt}
+    \includegraphics[width=0.49\textwidth]{../img/fitness_pfsp_rec31_txt}
+
   \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
-\end{figure}
+\end{figure*}
 

### 2021-02-04 paper/main.tex
beta gone

@@
     \Statex \Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
-    \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
+   % \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
+     \State $\sigma \assign$ sample a permutation from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
     % \Comment{Generate $\beta$ permutations}
-    \State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \label{line:distance}
-    \Statex \Comment{Select the farthest from $S$}
+    %\State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \label{line:distance}
+    %\Statex \Comment{Select the farthest from $S$}
    % \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
@@
 
-To increase the exploration of the search space, the algorithm samples $\beta$ permutations, $\Sigma$.
+% To increase the exploration of the search space, the algorithm samples $\beta$ permutations, $\Sigma$.
 Then it selects  for evaluation the farthest from permutations already evaluated ($S$), as shown in line~\ref{line:distance} of Algorithm~\ref{alg:umm}.
-%i.e.,  $\sigma\in \Sigma$ such that $\sigma = \argmax_{\sigma'\in \Sigma}\min d(\sigma', \pi)$ for every $\pi\in S$.
-
-\EKHINE{The algorithm uses a parameter $\beta$ that determines the number of points sampled, while the authors analyze the sensitivity of the results to the parameters r 1 and r 2, they do not do that with $\beta$.
+% %i.e.,  $\sigma\in \Sigma$ such that $\sigma = \argmax_{\sigma'\in \Sigma}\min d(\sigma', \pi)$ for every $\pi\in S$.
+
+% \EKHINE{The algorithm uses a parameter $\beta$ that determines the number of points sampled, while the authors analyze the sensitivity of the results to the parameters r 1 and r 2, they do not do that with $\beta$.
 In my opinion the value of $\beta$ is quite relevant.}
@@
 UMM uses the same budget of evaluations ($\FEmax$) as CEGO and initially
-samples $\minit=10$ solutions uniformly.  At each sampling step, we set
-$\beta=10$, meaning that 10 permutations are generated and the farthest from
-$S$ is selected for evaluation.  The other two parameters of UMM are $r_1$ and
-$r_2$, which determine the balance between exploration and exploitation. In the
-experimental section we will study their impact.
+samples $\minit=10$ solutions uniformly.
+%At each sampling step, we set
+%$\beta=10$, meaning that 10 permutations are generated and the farthest from
+%$S$ is selected for evaluation.
+The other two parameters of UMM are $r_1$ and $r_2$, which determine the
+balance between exploration and exploitation. In the experimental section we
+will study their impact.
 
@@
 
-\ekhine{In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
- obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP
-instance considered here. The lack of any apparent convergence suggests that
-the models (both Bayesian and probabilistic) are not learning anything about
-the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{kaka}
+% \ekhine{In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
+%  obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP
+% instance considered here. The lack of any apparent convergence suggests that
+% the models (both Bayesian and probabilistic) are not learning anything about
+% the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{}
 % \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded

### 2021-02-04 paper/main.tex
rank vs order

@@
 Instead, we consider classical combinatorial optimization problems as black-box optimization problems.
-\ekhine{cambio la definition de LOP}{In this section, and following the consensus on the optimization literature, the permutations denote orderings.
-}\MANUEL{That seems strange, shouldn't be use the same everywhere?
-I would rather use rankings or orderings everywhere and I have a slight preference for orders since this is an optimization paper.}
-
+\manuel{In this section, and following the consensus on the optimization literature, the permutations denote orderings.}{For consistency with the rest of the paper, a permutation $\sigma$ denotes a ranking also in this section, thus we use the inverse permutation $\sigma^{-1}$ in the definition of the problems.
+Later in the experimental section we discuss the effects of not performing this inversion.}
+%
 The \emph{Linear Ordering Problem (LOP)}, when minimizing, is defined as
@@
 \begin{equation}\label{eq:lop}
-  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^{n} \sum_{j=1}^{i-1}  a_{\sigma(i), \sigma(j)}
+  \min_{\sigma \in S_n} f(\sigma) = \sum_{i=1}^{n} \sum_{j=1}^{i-1}  a_{\sigma^{-1}(i), \sigma^{-1}(j)}
 \end{equation}
@@
 All jobs must be processed by all machines in the same order.
-Given a permutation of the jobs $\sigma$, $C_{i,j}$ denotes the completion time of job at position $\sigma(i)$ on machine $j$, and $C_{\max} = C_{n,M}$  is the completion time of the last job on the last machine, i.e., the makespan.
+Given a permutation of the jobs $\sigma$, $C_{i,j}$ denotes the completion time of job at position $\sigma^{-1}(i)$ on machine $j$, and $C_{\max} = C_{n,M}$  is the completion time of the last job on the last machine, i.e., the makespan.
 The objective of the PFSP is to find:
@@
     \text{s.t.}\quad & C_{1,j} = 0 \quad j\in [M], \qquad C_{i,1} = 0\quad i \in [n]\\
-    &C_{i,j} = p_{\sigma(i),j} + \max\{C_{i-1, j}, C_{i, j-1}\}  \ \  i \in \{2,\dotsc,n\},\\
+    &C_{i,j} = p_{\sigma^{-1}(i),j} + \max\{C_{i-1, j}, C_{i, j-1}\}  \ \  i \in \{2,\dotsc,n\},\\
     &\hspace{14.5em} j \in\{2,\dotsc, M\}

### 2021-02-04 paper/main.tex
paper s5

--- 2021-02-04 paper/main.tex
@@
 
+\subsection{Rankings and orderings}
+In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
+In this lines, we show the consequences of misinterpreting permutations.
+We remind the reader that statistical models (Borda or the \ken distance) are defined for rankings while standard literature in optimization considers orderings.
+This is the reason while in the optimization field we will usually find the Swap distance instead of the \ken distance.
+However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.
+
+As a summary of Algorithm~\ref{alg:umm}, we model rankings in our UMM model, generate a ranking $\sigma$, then invert this ranking to obtain an ordering in line~\ref{line:inversion} and evaluate the ordering $\sigma^{-1}$ in the objective function.
+This might seem an implementation detail but for the sake of consistency with the literature, we will expect objective functions to evaluate ordering.
+A natural question is what happens if we did not invert the ranking and evaluated a ranking with an optimization for orderings.
+For the anxious reader, we advance that the result is similar to the original version of CEGO.
+
+In CEGO, the objects modeled and evaluated are the same.
+They claim that they use the swap distance but they are using the \ken  distance, meaning that the object model is a ranking.
+\ekhine{}{esto es demasido rude?}
+On the function evaluation the authors interpert their permutation as an ordering, as usual. 
+However, there is no inversion between the objects sampled from the model and the objects evaluated. 
+This means that both are rankings (or both are orderings).
+A natural question is what happens if we did invert the ranking.
+We have carried out all the experimentation with this modification of CEGO and we have denote this version CEGO\_inv.
+The original version of CEGO is denoted CEGO\_orig.
+
+%Figure~\ref{fig:rankorder} answers the above two questions.
+In a setting as the one in previous sections, we find 4 different lines, for the evaluation of UMM and CEGO with and without inversion.
+We see that in case of No inversion (dashed lines) both UMM and CEGO do not seem to converge (despite we can see some decrease along iterations).
+When there is an inversion between modelling and evaluation both algorithms behave better.
+In this case, since UMM is set for a a linear decrease in the the scale parameter for the $m$ iterations, we see the linear decrease in the function evaluations.
+In CEGO, we see an abrupt decrease at the beggining which seems stuck at iteration 200.
+
 \subsection{Analysis of UMM and CEGO on LOP and PFSP}
@@
 over 10 runs with different random seeds on the same problem instance.  As an
-example, each line in Fig.~\ref{fig:lop_synth} shows the mean fitness of the
-\ekhine{solutions evaluated by either UMM or CEGO on one synthetic LOP instance}{ojo}. The
+example, each line in Fig.~\ref{fig:results} shows the mean fitness of the
+solutions evaluated by either UMM or CEGO on one instance. The
 shaded area shows one standard deviation around the mean. Ideally, each new
@@
 best one (minimum fitness) of each run.
+This result summarization is not the classical.
+However, due to the limited number of function evaluations, we can have a better picture of the behaviour of the algorithm at each step.
+As a consequence, results displayed here are not comparable with the original paper of CEGO: for this comparison we provide in Table~\ref{table:fitness} the summary of the best solutions found by each algorithm.
 
@@
     \includegraphics[width=0.49\textwidth]{../img/fitness_pfsp_rec31_txt}
-
+\label{fig:results}
   \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
@@
 
-\subsection{Rankings and orderings}
-In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
-In this lines, we show the consequences of this mistakes.
-We remind the reader that statistical models (Borda or the Kendall distance) are defined for rankings while standard literature in optimization considers orderings.
-As a summary of Algorithm~\ref{alg:umm}, we model rankings in our UMM model, generate a ranking $\sigma$, then invert this ranking to obtain an ordering in line~\ref{line:inversion} and evaluate the ordering $\sigma^{-1}$ in the objective function.
-A natural question is what happens if we did not invert the ranking.
-%\ekhine{}{rellenar}
-
-In CEGO, the objects modeled and evaluated are the same: There is no inversion between the objects sampled from the model and the objects evaluated.
-This means that botha re rankings or both are orders.
-For modelling the Kendall distance is used, that as stated is meant for rankings.
-A natural question is what happens if we did invert the ranking.
-
-Figure~\ref{fig:rankorder} answers the above two questions.
-In a setting as the one in previous sections, we find 4 different lines, for the evaluation of UMM and CEGO with and without inversion.
-We see that in case of No inversion (dashed lines) both UMM and CEGO do not seem to converge (despite we can see some decrease along iterations).
-When there is an inversion between modelling and evaluation both algorithms behave better.
-In this case, since UMM is set for a a linear decrease in the the scale parameter for the $m$ iterations, we see the linear decrease in the function evaluations.
-In CEGO, we see an abrupt decrease at the beggining which seems stuck at iteration 200.
-
-We compare this results with the numerical results provided in the original paper. These results are consistent with the setting of no inversion for CEGO. 
-\ekhine{probar con m=100 para ver ci podemos competir en instancias cortas}{}
-
 \section{Conclusions}\label{sec:conclusions}

### 2021-02-04 paper/main.tex
tuples

@@
   \begin{algorithmic}[1]
-    \Require $\minit$: number of initial permutations, $\FEmax$: total budget,\newline\hspace*{5ex} \manuel{$\rho$: learning rate,}{$r_1, r_2$: parameters controlling the learning rate} 
+    \Require $\minit$: number of initial permutations, $\FEmax$: total budget,\newline\hspace*{5ex} $r_1, r_2$: parameters controlling the learning rate 
     % \\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
@@
 \begin{equation}
-  \begin{split}
-  r_1 \sum_{\sigma\in S}\rho^{f(\sigma)} =  \sum_{\sigma'\in S'}\rho^{f(\sigma')} \quad\text{where}\; &|S'| = r_2|S| \\[-1.2em]
-  &\land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\leq f(\sigma).
-  \end{split}
+%  \begin{split}
+    r_1 \cdot \sum_{\tuple{\sigma,f(\sigma)}\in S}\rho^{f(\sigma)} =  \sum_{\tuple{\sigma', f(\sigma')}\in S'}\rho^{f(\sigma')}  %\\
+  %    & \text{where}\;|S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\leq f(\sigma).
+%  \end{split}
 \end{equation}
 %
-i.e., $S'$ contains the $r_2|S|$ elements of $S$ with the best corresponding fitness values in $F$.
+where $|S'| = r_2|S| \land \forall \tuple{\sigma',f(\sigma')}\in S', \tuple{\sigma,f(\sigma)}\in S\setminus S' : f(\sigma')\leq f(\sigma)$, that is, $S'$ contains the $r_2|S|$ elements of $S$ with the best  fitness values.
 %The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification.

### 2021-02-04 paper/main.tex
minor

@@
 
-\subsection{Permutations: rankings vs orderings}
+\subsection{Permutations: rankings vs orderings}\label{sec:permutations}
 
@@
     \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
-    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}\Comment{Learning Step}
+    \State $\hat\sigma_0 \assign \text{uBorda}(S,\rho)$\label{line:uborda}\Comment{Learning Step}
     \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}
@@
 \subsection{Rankings and orderings}
-In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
+In Section~\ref{sec:permutations} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
 In this lines, we show the consequences of misinterpreting permutations.
@@
 %\renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}
+
 \bibliographystyle{ACM-Reference-Format-abbrv}

### 2021-02-04 paper/main.tex
paper s5

--- 2021-02-04 paper/main.tex
@@
 
-\subsection{Rankings and orderings}
+\subsection{Rankings and orderings: experiments}\label{sec:exper_ro}
+\ekhine{}{este primer parago es bastante importante para dejarlo aqui. }
 In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
@@
 However, due to the limited number of function evaluations, we can have a better picture of the behaviour of the algorithm at each step.
-As a consequence, results displayed here are not comparable with the original paper of CEGO: for this comparison we provide in Table~\ref{table:fitness} the summary of the best solutions found by each algorithm.
-
-Figure~\ref{fig:lolib} shows the results on the real-world instances taken from LOLIB. 
+As a consequence, results displayed in the figures are not comparable with the original paper of CEGO: for this comparison we provide in Table~\ref{table:fitness} the summary of the best solutions found by each algorithm, which are similar to the original study.
+
+\paragraph{Figure descriptiopn} Figure~\ref{fig:results} shows the results on the real-world instances from LOLIB and PFSP.
+There is one figure per instance.
+In each figure there is one horizontal line displaying the best known solutions for the problem.
+Besides, there are 5 different lines representing each of the different algorithm/versions considered, in particular:
+\begin{itemize}
+\item UMM\_100, UMM\_200 and UMM\_400 are three different versions for UMM, each considering a different total number of function evaluations, 100, 200 and 400.
+Obviously, the lines corresponding to UMM\_100, UMM\_200 wont reach the right corner of the figure.
+\item CEGO\_orig and CEGO\_inv are, respectively, the original version of CEGO and the modification of Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$.
+\end{itemize}
+
+We start our analysis with the three version of UMM.
+We can see a general trend for all the instances: After the 10 random initial permutations, the fitness of the new solutions decreases with a linear trend along iterations.
+The (close to) linear decrease is given by the linear decrease in the expected distance of the model which is sampled at each step, described in Section~\ref{sec:sampling}.
+Overall, the improvement does not seem to converge not even for the longer runs of 400 evaluations.
+The average runtime for all the instances is around 0.3 minutes, 1.15 minutes and 4.5 minutes respectively for the  versions with 100, 200 and 400 fitness evaluations.
+
   %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí

### 2021-02-04 paper/main.tex
sec 5.2

@@
 \usepackage{afterpage,tabularx}
+\hypersetup{draft}
 %\usepackage[colorlinks=true, allcolors=blue]{hyperref}
@@
 
+\newcommand{\CEGOorig}{CEGO$_\text{ori}$\xspace}
+\newcommand{\CEGOinv}{CEGO$_\text{inv}$\xspace}
+
 \subsection{Rankings and orderings: experiments}\label{sec:exper_ro}
+
+In Section~\ref{sec:permutations}, we highlighted the important difference between rankings and orderings, usually neglected in the optimization literature.
+Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually considers orderings.
+An example of such confusion is the claim that Swap distance is the same metric as  the \ken distance, which is false except for particular permutations.
+In reality, the \emph{swap} distance is the \ken distance of the inverted permutations.
+
+As a consequence of this important distinction, when evaluating classical
+combinatorial problems, we must convert rankings to orderings by inverting the
+permutation, as done in Eq.~\ref{eq:lop} and Eq.~\ref{eq:pfsp}, for the LOP and
+PFSP, respectively. This might seem an implementation detail but a failure to use the correct interpretation has grave consequences.
+
+During our experiments, we realized that the \emph{swap} distance used by CEGO
+assumes rankings, i.e., it actually measures the \ken distance, yet CEGO did
+not invert the permutation before evaluating the PFSP. We verified that this is
+indeed how the implementation works and that those results match the published
+ones~\citep{ZaeStoBar2014:ppsn}. Of course, the obvious question arises of what
+would be the effect of inverting the permutation before calling the objective
+function, so that the rankings are transformed to ordering, as done by our UMM
+algorithm. Therefore, in the following, we show results of both variants, that is, the original CEGO (\CEGOorig) and our modified version that inverts the permutation before evaluation (\CEGOinv).
+
+\manuel{}{The above replaces the text below?}
+
 \ekhine{}{este primer parago es bastante importante para dejarlo aqui. }
 In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
+In this lines, we show the consequences of misinterpreting permutations.
+We remind the reader that statistical models (Borda or the \ken distance) are defined for rankings while standard literature in optimization considers orderings.
+This is the reason while in the optimization field we will usually find the Swap distance instead of the \ken distance.
+However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.
+
+In Section~\ref{sec:permutations} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
 In this lines, we show the consequences of misinterpreting permutations.
@@
 
-  %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
-  % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
-%The type of instances differs greatly from the synthetic LOP instances, and,
-%thus, so does the landscape. However, UMM shows a very similar behavior: 
-We can see that for UMM there 
-is a (close to) linear decrease of the fitness of new solutions over each new
-evaluation. This suggests that UMM is able to model the fitness
-landscape. Regarding CEGO, \ekhine{we observe a similar behavior to UMM for the first
-50-100 evaluations, however, in several instances CEGO reaches the time limit
-of 144 hours, so no observations can be made for longer budgets. In smaller
-instances (e.g., Figure~\ref{fig:lolib}, top), CEGO again appears to stagnate.}{ the objective function decrease slowly?}
-
-% \ekhine{In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
-%  obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP
-% instance considered here. The lack of any apparent convergence suggests that
-% the models (both Bayesian and probabilistic) are not learning anything about
-% the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{}
-% \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
-% that other distance measures are more suited than Kendall's-$\tau$ distance for
-% the PFSP, the reported fitness differences between various distance metrics are
-% small.
-% Hence, we conjecture that the same behavior will be observed with other distance metrics.}{}\MANUEL{On second thought, I'm not sure we should say this, because we really have no idea and we actually hope it is not true.}
+The red
 
@@
 \end{figure*}
+
+\paragraph{Discussion on other problems: the case of QAP}
 

### 2021-02-04 paper/main.tex
paper s5
plots

--- 2021-02-04 paper/main.tex
@@
 Instead, we consider classical combinatorial optimization problems as black-box optimization problems.
-\manuel{In this section, and following the consensus on the optimization literature, the permutations denote orderings.}{For consistency with the rest of the paper, a permutation $\sigma$ denotes a ranking also in this section, thus we use the inverse permutation $\sigma^{-1}$ in the definition of the problems.
-Later in the experimental section we discuss the effects of not performing this inversion.}
+For consistency with the rest of the paper, a permutation $\sigma$ denotes a ranking also in this section, thus we use the inverse permutation $\sigma^{-1}$ in the definition of the problems.
+Later in the experimental section we discuss the effects of not performing this inversion.
 %
@@
 %
-where $[a_{i,j}]$ is a matrix of size $n \times n$ \manuel{and $\sigma^{-1}$ is the inverse of $\sigma$}{}.
+where $[a_{i,j}]$ is a matrix of size $n \times n$.
 
@@
 indeed how the implementation works and that those results match the published
-ones~\citep{ZaeStoBar2014:ppsn}. Of course, the obvious question arises of what
-would be the effect of inverting the permutation before calling the objective
+ones~\citep{ZaeStoBar2014:ppsn}. Of course, the obvious question arises about the effect of inverting the permutation before calling the objective
 function, so that the rankings are transformed to ordering, as done by our UMM
@@
 \begin{figure*}[tb]
+  \small
   \centering%
@@
     \includegraphics[width=0.49\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
-
+\\
     \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p40-01}
-    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p50-01}
-
-    \includegraphics[width=0.49\textwidth]{../img/fitness_pfsp_rec19_txt}
-    \includegraphics[width=0.49\textwidth]{../img/fitness_pfsp_rec31_txt}
+    \begin{minipage}{0.49\linewidth}
+          \texttt{RandB N-p50-01} (LOP)\\[-0.5ex]
+    \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
+  \end{minipage}
+    \\
+    \begin{minipage}{0.49\linewidth}
+    \texttt{reC19} (PFSP)\\[-0.5ex]
+\includegraphics[width=\textwidth]{../img/fitness_pfsp_rec19_txt}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+    \texttt{reC31} (PFSP)\\[-0.5ex]
+    \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec31_txt}
+  \end{minipage}
 \label{fig:results}
-  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
+  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:rec05}}
 \end{figure*}

### 2021-02-04 paper/main.tex
paper s5

--- 2021-02-04 paper/main.tex
@@
 Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually considers orderings.
+For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.
+The concepts of rankings and orderings are usually confused of ignored.
 An example of such confusion is the claim that Swap distance is the same metric as  the \ken distance, which is false except for particular permutations.
@@
 
-The red
+The red and the gold lines correspond to CEGO, the original and our modified versions
 

### 2021-02-04 paper/main.tex
plots
minor

@@
 Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually considers orderings.
-For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.
+For example, the Swap distance $d_s(\sigma,\pi)$ is the counterpart of the  \ken distance  in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.
 The concepts of rankings and orderings are usually confused of ignored.
@@
 As a consequence of this important distinction, when evaluating classical
-combinatorial problems, we must convert rankings to orderings by inverting the
+combinatorial problems, UMM transforms rankings to orderings by inverting the
 permutation, as done in Eq.~\ref{eq:lop} and Eq.~\ref{eq:pfsp}, for the LOP and
@@
 \centering%
-    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
-    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
+\begin{minipage}{0.49\linewidth}
+  \texttt{N-t59b11xx} (LOP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+  \texttt{N-sgb75-01} (LOP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
+\end{minipage}
 \\
-    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p40-01}
 \begin{minipage}{0.49\linewidth}
-          \texttt{RandB N-p50-01} (LOP)\\[-0.5ex]
+  \texttt{N-p40-01} (LOP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-01}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+  \texttt{N-p50-01} (LOP)\\[-0.5ex]
   \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}

### 2021-02-04 paper/main.tex
5.3

@@
 
-\subsection{Analysis of UMM and CEGO on LOP and PFSP}
-To  understand the behavior of the algorithms, we record the fitness of
-the solution evaluated at each step of each run and we average those values
-over 10 runs with different random seeds on the same problem instance.  As an
-example, each line in Fig.~\ref{fig:results} shows the mean fitness of the
-solutions evaluated by either UMM or CEGO on one instance. The
-shaded area shows one standard deviation around the mean. Ideally, each new
+\subsection{Experimental Analysis of UMM and CEGO} % on LOP and PFSP instances}
+
+To better analyze the behavior of the algorithms, we record the fitness of the
+solution evaluated at each step of each run. For each problem instance and each
+algorithm, we plot the mean fitness at each evaluation over 10 runs with
+different random seeds, as shown in Fig.~\ref{fig:results}.
+%
+The shaded area shows one standard deviation around the mean. Ideally, each new
 evaluation will monotonically improve in fitness, since each solution evaluated
 helps to refine the model and leads to a better solution being evaluated in the
-next step.  However, it is possible that new solutions are worse
-than their predecessors. %Nevertheless, in this example, there is a clear
+next step.  However, it is possible that new solutions are worse than their
+predecessors. %Nevertheless, in this example, there is a clear
 % overall improvement as more evaluations are performed.
-The minimum values
-achieved in the plot up to a particular number of function evaluations gives an
-estimation of the mean fitness of the best-so-far solution up to that point in
-the run. The algorithms do not return the final solution shown but rather the
-best one (minimum fitness) of each run.
+
+We chose to plot the fitness of each evaluated solution rather than the fitness
+of the best-so-far to better understand the behavior of the algorithms at each
+step. Nevertheless, the minimum values achieved in the plot up to a particular
+number of function evaluations gives an estimation of the mean fitness of the
+best-so-far solution up to that point in the run. The algorithms do not return
+the final solution shown but rather the best one (minimum fitness) of each run.
+
 This result summarization is not the classical.
@@
 \label{fig:results}
-\caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:rec05}}
+\caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:results}}
 \end{figure*}

### 2021-02-04 paper/main.tex
paper s5

--- 2021-02-04 paper/main.tex
@@
 \subsection{Rankings and orderings: experiments}\label{sec:exper_ro}
+\ekhine{}{este primer parago es bastante importante para dejarlo aqui. }
 
 In Section~\ref{sec:permutations}, we highlighted the important difference between rankings and orderings, usually neglected in the optimization literature.
-Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually considers orderings.
+Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually builds upon orderings.
 For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.
-The concepts of rankings and orderings are usually confused of ignored.
+%The concepts of rankings and orderings are usually confused of ignored.
 An example of such confusion is the claim that Swap distance is the same metric as  the \ken distance, which is false except for particular permutations.
@@
 
-As a consequence of this important distinction, when evaluating classical
+As a consequence of this important distinction, when dealing with data in the form of permutations, in particular in 
 combinatorial problems, we must convert rankings to orderings by inverting the
@@
 
-During our experiments, we realized that the \emph{swap} distance used by CEGO
-assumes rankings, i.e., it actually measures the \ken distance, yet CEGO did
+During our experiments, we realized that CEGO claimed to use the Swap distance. Instead, CEGO implements the \ken distance, yet CEGO did
 not invert the permutation before evaluating the PFSP. We verified that this is
@@
 
-\manuel{}{The above replaces the text below?}
-
-\ekhine{}{este primer parago es bastante importante para dejarlo aqui. }
-In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
+%\manuel{}{The above replaces the text below?}
+%
+%In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
 In this lines, we show the consequences of misinterpreting permutations.
@@
 However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.
-
-In Section~\ref{sec:permutations} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
+%
+%
+%In Section~\ref{sec:permutations} we highlighted the great difference of rankings and orderings, usually ommitted in the literature.
 In this lines, we show the consequences of misinterpreting permutations.
@@
 However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.
-
-As a summary of Algorithm~\ref{alg:umm}, we model rankings in our UMM model, generate a ranking $\sigma$, then invert this ranking to obtain an ordering in line~\ref{line:inversion} and evaluate the ordering $\sigma^{-1}$ in the objective function.
+% 
+%As a summary of Algorithm~\ref{alg:umm}, we model rankings in our UMM model, generate a ranking $\sigma$, then invert this ranking to obtain an ordering in line~\ref{line:inversion} and evaluate the ordering $\sigma^{-1}$ in the objective function.
 This might seem an implementation detail but for the sake of consistency with the literature, we will expect objective functions to evaluate ordering.
@@
 For the anxious reader, we advance that the result is similar to the original version of CEGO.
-
-In CEGO, the objects modeled and evaluated are the same.
+%
+%In CEGO, the objects modeled and evaluated are the same.
 They claim that they use the swap distance but they are using the \ken  distance, meaning that the object model is a ranking.
 \ekhine{}{esto es demasido rude?}
-On the function evaluation the authors interpert their permutation as an ordering, as usual. 
-However, there is no inversion between the objects sampled from the model and the objects evaluated. 
-This means that both are rankings (or both are orderings).
+%On the function evaluation the authors interpert their permutation as an ordering, as usual. 
+%However, there is no inversion between the objects sampled from the model and the objects evaluated. 
+%This means that both are rankings (or both are orderings).
 A natural question is what happens if we did invert the ranking.
@@
 The original version of CEGO is denoted CEGO\_orig.
+%
 
@@
 
-We start our analysis with the three version of UMM.
+\paragraph{Analysis} We start our analysis with the three version of UMM.
 We can see a general trend for all the instances: After the 10 random initial permutations, the fitness of the new solutions decreases with a linear trend along iterations.
@@
 
-The red and the gold lines correspond to CEGO, the original and our modified versions
+The red and the gold lines correspond to CEGO, the original and our modified versions.
+For the reason for the abrupt change between both version for every instance, we refer the reader to Section~\ref{sec:exper_ro}.
+For the original version, there is a global decrease in the fitness function.
+However, a closer analysis of each iteration does not show a real convergence to any point.
+However, with the modified version in which rankings are modeled and orderings are evaluated we can appreciate the a decrease in terms of fitness per evaluation.
+Overall, the modified CEGO has a very abrupt decrease at the first 50 iterations.
+At one point, the algorithm seems to have converged and very little improvement is don from the 150-th iteration onwards.
+The average running time is approximately 22 hours for the 400 function evaluations, which is 300 times slower than UMM for the same number of function evaluations.
+
+The performance of the UMM is clearly better than the performance of the original CEGO for the same problem.
+For CEGO\_inv on the other hand improves the results of the UMM with the current experimental setting.
+However, UMM seems to be improving the quality of the solutions until the last iteration while CEGO\_inv seems to be stucked at a local optimum.
+Further analysis lines will consider a larger budget of fitness evaluation.
+
+Moreover, UMM seems to scale good for larger instances in quality performance. Note that the instances are order from top to bottom in the size of the permutations. 
+The computational complexity of UMM is \ekhine{}{ $O(n \log n)$ }at each iteration while both versions of CEGO need \manuel{tu sabes}{?}.
+This difference is evident in the time performance of the algorithms.
+The details can be seen in Table~\ref{table:time}.
 
@@
 
-Table~\ref{tab:results} shows results over all instances evaluated (excluding
-the synthetic LOP instances). In particular, we show the mean fitness (and
-standard deviation) of the best solution found at the end of each run.
+%Table~\ref{tab:results} shows results over all instances evaluated (excluding
+%the synthetic LOP instances). In particular, we show the mean fitness (and
+%standard deviation) of the best solution found at the end of each run.
 We also report the mean runtime in hours and, since CEGO sometimes reached the limit of 144 hours, also the number of fitness evaluations reached by CEGO.
 
-In the LOP, differences are mostly in favour of UMM, due to the time limit of
-144 hours, which resulted in CEGO typically achieving less than 200
-evaluations. In contrast, UMM was always able to reach 400 evaluations in less
-than 12.5 hours. If UMM was allowed to reach the 144 hours limit, the differences
-would be even stronger.
-%
-In the PFSP, the differences are minor and/or not statistically
-significant, with the qualification that both algorithms appear to be
-performing a random search.
+%In the LOP, differences are mostly in favour of UMM, due to the time limit of
+%144 hours, which resulted in CEGO typically achieving less than 200
+%evaluations. In contrast, UMM was always able to reach 400 evaluations in less
+%than 12.5 hours. If UMM was allowed to reach the 144 hours limit, the differences
+%would be even stronger.
+%%
+%In the PFSP, the differences are minor and/or not statistically
+%significant, with the qualification that both algorithms appear to be
+%performing a random search.
 
@@
 black-box combinatorial problems on permutation landscapes and when the budget
-of fitness evaluations is severely limited (here, up to $400$ evaluations).
+of fitness evaluations is severely limited (here, up to $100$, $200$ and $400$ evaluations).
 Our experiments presented here show that UMM is able to match the results of
-CEGO, a state-of-the-art Bayesian optimizer for expensive black-box
+the state-of-the-art Bayesian optimizer for expensive black-box
 optimization if the computation time of the optimizers is not a concern.
@@
 
-%\MANUEL{UMM is able to model the LOP landscape
-%  at least as well as CEGO as shown by the plots where UMM shows a linear
-%  decrease in fitness.}  
-%  \EKHINE{me parece mas de CEGO q UMM y parece q UMM
-%  solo tiene de ventaja ir rápido. Primero destacaría lo bueno de UMM. Creo q
-%  es importante destaacar q UMM va MEJOR o IGUAL en TODOS los experimentos y
-%  luego que addemás es más rápido} 
-%  \MANUEL{Pero UMM es peor en rec13 y nug12, y
-%  en LOP es mejor porque CEGO no llega a 400, es decir, porque UMM es más
-%  rápido. No podemos saber qué pasaría si CEGO llegara a 400. Podríamos medir
-%  lo que pasa en el momento en que matamos a CEGO y dar también ese valor
-%  entonces la diferencia en velocidad no importaría. } 
-% The time required to train the surrogate-model of \ekhine{esta establecida la complexity de CEGO???}{UMM increases linearly with the permutation length $n$, while in }
-% CEGO  it sharply
-% increases with $n$, requiring 
-% sometimes more than hour per fitness evaluation as shown in the experimental section. 
-% Hence, UMM is a computationally
-% feasible alternative for relatively large problem sizes.
+In this paper, we discuss the importance of the representation of permutations in applied mathematics problems.
+In the context considered in this paper, permutations can bee seen as rankings or orderings of items.
+In other problems they could be matchings or cycles.
+While certain functions are
+this discussion is usually ommited in the literature.
+We discuss the different concepts and show, experimentally, the consequences of misunderstanding both.
 
@@
 knowledge, the first EDA specifically designed for such problems that is able
-to match (and sometimes even improve) the results of a Bayesian
+to match (and sometimes even improve) the state-of-the-art
 optimizer. Moreover, it is overwhelmingly faster than the alternatives for
@@
 
-An interesting result is that the behavior of UMM (and the same for CEGO) on
-the PFSP (and some QAP instances) is close to random. %The same behavior is
-%observed for CEGO, which seems to have been never reported until now.
-This result may be due to the distance
-metric used, but further analysis is needed to corroborate this
-conjecture. Here, we have focused on Kendall's-$\tau$ distance, however, we are certain it is
+In this paper, we focus on the UMM under the \ken distance.
+Therefore, we consider appropiate problems for it, such as the Linear Ordering Problem and the Permutation Flowshop Scheduling Problem, where the pairwise ordering among the items is relevant (as opposed to the actual exat position of each item in the permutation)
+
+However, we are certain it is
 possible to extend UMM to other distance metrics, which will allow us to

### 2021-02-04 paper/main.tex
paper s5

@@
 \label{fig:results}
-  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:rec05}}
+\caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:results}}
 \end{figure*}

### 2021-02-04 paper/main.tex
more

@@
 (\emph{learning}) of the parameters $\sigma_0$, $\theta$ and $\rho$ (line~\ref{line:uborda} in Algorithm~\ref{alg:umm}) and
-the sampling from $MM(\sigma_0,\theta)$ (lines~\ref{line:sample}--\ref{line:distance}). % in Algorithm~\ref{alg:umm}).
+the sampling from $MM(\sigma_0,\theta)$ (line~\ref{line:sample}) %(lines~\ref{line:sample}--\ref{line:distance}). % in Algorithm~\ref{alg:umm}).
 
@@
 At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$, i.e., half of the expected distance for the uniform distribution.
-Then the expected distance linearly decreases until the last iteration where $\theta$ is set such that $\Edist = 1$.
+Then the expected distance linearly decreases until the last iteration $m$ where $\theta$ is set such that $\Edist = 1$.
 Despite it is not possible to isolate $\theta$ in Eq.~\eqref{eq:expectation}, its monotonicity allows the use of bisection methods for efficiently finding an approximation.
@@
 A purely explotation approach will set $\Edist = 0$  generating $\sigma_0$ in the sampling stage.
-However, the setting of a decreasing variance   controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.
-It follows that the behaviour is different for different choices of $m$.
+However, the setting of a decreasing variance   controls the exploration-explotation tradeoff as a function of the number of function evaluations ($m$) that the algorithm performs, thus \manuel{it follows that}{} the behaviour is different for different choices of $m$.
 Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.
@@
 
-\newcommand{\CEGOorig}{CEGO$_\text{ori}$\xspace}
-\newcommand{\CEGOinv}{CEGO$_\text{inv}$\xspace}
+\newcommand{\CEGOorig}{CEGO$_\textsc{ori}$\xspace}
+\newcommand{\CEGOinv}{CEGO$_\textsc{inv}$\xspace}
 
 \subsection{Rankings and orderings: experiments}\label{sec:exper_ro}
-\ekhine{}{este primer parago es bastante importante para dejarlo aqui. }
 
@@
 Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually builds upon orderings.
-For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.
-%The concepts of rankings and orderings are usually confused of ignored.
+For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.\MANUEL{necesitamos las ecuaciones?
+Swap de una permutación es simplemente la distancia de kendall de la permutación inversa.} %The concepts of rankings and orderings are usually confused of ignored.
 An example of such confusion is the claim that Swap distance is the same metric as  the \ken distance, which is false except for particular permutations.
@@
 
-As a consequence of this important distinction, when dealing with data in the form of permutations, in particular in 
+As a consequence of this important distinction, when dealing with \manuel{data in the form of}{} permutations, in particular in 
 combinatorial problems, we must convert rankings to orderings by inverting the
@@
 
+%During our experiments, we realized that the \emph{swap} distance used by CEGO
+%assumes rankings, i.e., it actually measures the \ken distance, yet CEGO did
+% not invert the permutation before evaluating the PFSP.
 During our experiments, we realized that CEGO claimed to use the Swap distance. Instead, CEGO implements the \ken distance, yet CEGO did
-not invert the permutation before evaluating the PFSP. We verified that this is
+not invert the permutation before evaluating the PFSP.\MANUEL{no me convence esta forma de expresarlo porque la crítica obvia sería que por qué invertimos la función objectivo y no la función de distancia?} We verified that this is
 indeed how the implementation works and that those results match the published
@@
 
-To better analyze the behavior of the algorithms, we record the fitness of the
-solution evaluated at each step of each run. For each problem instance and each
-algorithm, we plot the mean fitness at each evaluation over 10 runs with
-different random seeds, as shown in Fig.~\ref{fig:results}.
+In the following, we compare 5 algorithm configurations. Three configurations
+of UMM (UMM$_{100}$, UMM$_{200}$, and UMM$_{400}$) that differ in their budget
+of evaluations ($m$). As mentioned in Sec.~\ref{sec:sampling}, the behavior of
+UMM depends on the value of $m$. The other two configurations correspond 
+to \CEGOorig and \CEGOinv, as defined above.
+
+Since there is a small number of evaluations and the algorithms evaluate just
+one solution at each step, we chose to record and plot the fitness of the
+solution evaluated at each step of each run, instead of the usual plots of the
+best-so-far solution over number of evaluations. We believe that plotting each
+evaluated solution provides better insights on the search dynamics of the
+algorithms analyzed here. Please note that the algorithms do not return the
+final solution shown in the plots, but rather the best one (minimum fitness) of
+each run. Nevertheless, the minimum values achieved in the plot up to a
+particular number of function evaluations gives an estimation of the mean
+fitness of the best-so-far solution up to that point in the run. We provide
+later an analysis of the best solution found by each run of the algorithm.
+
+For each problem instance and each algorithm, we plot the mean fitness of the
+solution evaluated at each step over 10 runs with different random
+seeds, as shown in Fig.~\ref{fig:results}.
 %
@@
 
-We chose to plot the fitness of each evaluated solution rather than the fitness
-of the best-so-far to better understand the behavior of the algorithms at each
-step. Nevertheless, the minimum values achieved in the plot up to a particular
-number of function evaluations gives an estimation of the mean fitness of the
-best-so-far solution up to that point in the run. The algorithms do not return
-the final solution shown but rather the best one (minimum fitness) of each run.
-
+Figure~\ref{fig:results} shows the results on the real-world instances from LOLIB and PFSP.
+There is one figure per instance.
+In each figure there is one horizontal line displaying the best known solutions for the problem.
+Besides, there are 5 different lines representing each of the different algorithm/versions considered, in particular:
+
+\manuel{%
 This result summarization is not the classical.
@@
 As a consequence, results displayed in the figures are not comparable with the original paper of CEGO: for this comparison we provide in Table~\ref{table:fitness} the summary of the best solutions found by each algorithm, which are similar to the original study.
+}{}
 

### 2021-02-04 paper/main.tex
paper s5

--- 2021-02-04 paper/main.tex
@@
 Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually builds upon orderings.
-For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.
-%The concepts of rankings and orderings are usually confused of ignored.
-An example of such confusion is the claim that Swap distance is the same metric as  the \ken distance, which is false except for particular permutations.
-In reality, the \emph{swap} distance is the \ken distance of the inverted permutations.
+For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, ${d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})}$.
 
@@
 
-\subsection{Experimental Analysis of UMM and CEGO} % on LOP and PFSP instances}
+\subsection{Experimental Analysis} % on LOP and PFSP instances}
 
@@
   \texttt{reC19} (PFSP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec05_txt}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+  \texttt{reC31} (PFSP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec13_txt}
+\end{minipage}
+\\
+\begin{minipage}{0.49\linewidth}
+  \texttt{reC19} (PFSP)\\[-0.5ex]
   \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec19_txt}

### 2021-02-04 paper/main.tex
results

--- 2021-02-04 paper/main.tex
@@
 
-\renewcommand{\subsubsection}[1]{\smallskip{}\noindent\textbf{#1}.}
+\renewcommand{\myparagraph}[1]{\smallskip{}\noindent\textbf{#1}.}
 
@@
 
-\subsubsection{UMM}
+\myparagraph{UMM}
 %
@@
 
-\subsubsection{CEGO} \EKHINE{tenemos una pagina entera dedicada a CEGO :s}\MANUEL{Sugiere qué quitar usando el commando \textbackslash{}ekhine}
+\myparagraph{CEGO} \EKHINE{tenemos una pagina entera dedicada a CEGO :s}\MANUEL{Sugiere qué quitar usando el commando \textbackslash{}ekhine}
 Combinatorial Efficient Global Optimization (CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
@@
 
-\subsubsection{Other settings}
+\myparagraph{Other settings}
 %
@@
 
-\subsubsection{Benchmark problems}
+\myparagraph{Benchmark problems}
 %
@@
 
-\subsubsection{Datasets}
+\myparagraph{Datasets}
 %
@@
 
-\subsubsection{Computing environment.}
+\myparagraph{Computing environment.}
 % Due to the computational effort required, we used several computing systems to
@@
 
-For each problem instance and each algorithm, we plot the mean fitness of the
-solution evaluated at each step over 10 runs with different random
-seeds, as shown in Fig.~\ref{fig:results}.
+Figure~\ref{fig:results} shows such plots for a few LOP instances from LOLIB
+and the PFSP.  For each problem instance and each algorithm, we plot the mean
+fitness of the solution evaluated at each step over 10 runs with different
+random seeds. The horizontal line indicates the best-known fitness for that instance.
 % 
@@
 
-Figure~\ref{fig:results} shows the results on the real-world instances from LOLIB and PFSP.
-There is one figure per instance.
-In each figure there is one horizontal line displaying the best known solutions for the problem.
-Besides, there are 5 different lines representing each of the different algorithm/versions considered, in particular:
+\myparagraph{Convergence analysis of UMM} We start our analysis with the three
+configurations of UMM. We can see a common trend in all instances: After
+the 10 random initial permutations, the fitness of new solutions decreases
+with a linear trend along iterations. The (close to) linear decrease is given
+by the linear decrease in the expected distance\MANUEL{distance to the sampling model or variance of the sampling model?} of the sampling model, as described in Section~\ref{sec:sampling}.
+Overall, the rate of improvement
+does not seem to converge, not even for the longer runs of 400 evaluations.
+Therefore, we conjecture that longer runs are likely to produce further improved solutions.
+Moreover, a non-linear decrease in the expected distance should result in a faster convergence for larger evaluation budgets.
+\manuel{The
+average runtime for all the instances is around 0.3 minutes, 1.15 minutes and
+4.5 minutes respectively for the versions with 100, 200 and 400 fitness
+evaluations.}{}\MANUEL{We should discuss the time when discussing the tables.}
+
+\paragraph{Convergence analysis of CEGO}
+The red and the gold lines in the plots shown in Fig.~\ref{fig:results}
+correspond to the two variants of CEGO the original \CEGOorig and our modified
+version \CEGOinv. The different between these two variants is overwhelmingly in
+favour of \CEGOinv. With the exception of the \texttt{reC19} PFSP instance, the
+search behavior of \CEGOorig resembles a random search, after a brief
+improvement phase at the start of the run. By contrast, \CEGOinv shows a very
+fast improvement in fitness within the first 50 evaluations that tends to
+converge around the 150$^\text{th}$ evaluation, showing little improvement
+after that. Since the only difference between these two variants is the
+transformation from rankings to orderings before evaluating the objective
+function (Sec.~\ref{sec:exper_ro}), we must conclude that this difference has
+tremendous effect in performance. Therefore, we will focus our analysis on
+\CEGOinv.
+
+\myparagraph{Comparison of UMM and CEGO}
+
+When comparing the UMM and CEGO configurations shown in Fig.~\ref{fig:results},
+we can observe that all UMM configurations perform much better than the
+original CEGO (\CEGOorig), while they converge slower than our modified
+\CEGOinv. However, the rate of improvement of the UMM variants does not seem to
+slow down near the maximum budget given and, in several instances such as
+\texttt{rec31} and \texttt{N-sgb75-01}, matches and surpasses the solutions
+found by \CEGOinv. Interestingly, this is the case, even for the shorter budget
+configurations of UMM, i.e., UMM$_{100}$ and UMM$_{200}$. We also observe a
+pattern that UMM performs better, or \CEGOinv performs worse, on larger
+instances, as can be seen by comparing \texttt{rec19} with \texttt{rec31}.
+
+\myparagraph{Comparison of the best solutions found} We now turn our analysis
+to the best (minimum) fitness found at the end of the run by each algorithm. We
+show in Table~\ref{tab:results}
+
+For the reason for the abrupt change between both version for
+every instance, we refer the reader to Section~\ref{sec:exper_ro}. For the
+original version, there is a global decrease in the fitness function. However,
+a closer analysis of each iteration does not show a real convergence to any
+point. However, with the modified version in which rankings are modeled and
+orderings are evaluated we can appreciate the a decrease in terms of fitness
+per evaluation. Overall, the modified CEGO has a very abrupt decrease at the
+first 50 iterations. At one point, the algorithm seems to have converged and
+very little improvement is don from the 150-th iteration onwards.  The average
+running time is approximately 22 hours for the 400 function evaluations, which
+is 300 times slower than UMM for the same number of function evaluations.
 
@@
 \begin{minipage}{0.49\linewidth}
+  \texttt{N-p40-01} (LOP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-01}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+  \texttt{N-p50-01} (LOP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
+\end{minipage}
+\\
+\begin{minipage}{0.49\linewidth}
   \texttt{N-t59b11xx} (LOP)\\[-0.5ex]
@@
   \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
-\end{minipage}
-\\
-\begin{minipage}{0.49\linewidth}
-  \texttt{N-p40-01} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-01}
-\end{minipage}
-\begin{minipage}{0.49\linewidth}
-  \texttt{N-p50-01} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
 \end{minipage}
@@
 
-\subsubsection{Reproducibility} Source code, datasets and scripts necessary to reproduce the results are available at \supplement.
+\myparagraph{Reproducibility} Source code, datasets and scripts necessary to reproduce the results are available at \supplement.
 %  \vspace*{-1em}

### 2021-02-04 paper/main.tex
paper s45

--- 2021-02-04 paper/main.tex
@@
 An equivalent definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made to convert $\sigma_1^{-1}$ into $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
-\EKHINE{Mirandolo bien, lo que pasa es que dicen que es lo mismo la swap distance que la kendall!}
+%\EKHINE{Mirandolo bien, lo que pasa es que dicen que es lo mismo la swap distance que la kendall!}
 
@@
 Then, using the current sample $S$, the uBorda algorithm is applied to learn a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}).
-Together with the parameter $\theta$, they define the Mallows model $MM(\hat\sigma_0, \theta)$. 
-A decreasing $\theta$ parameter (line~\ref{line:theta}) implies a decreasing variance of the Mallows model (Eq.~\ref{eq:MM}), as explained in  Section~\ref{sec:sampling} below. 
+Together with the parameter $\theta$, the algorithm defines the Mallows model $MM(\hat\sigma_0, \theta)$. 
+A decreasing $\theta$ parameter along iterations (line~\ref{line:theta}) implies a decreasing variance of the Mallows model (Eq.~\ref{eq:MM}), as explained in  Section~\ref{sec:sampling} below. 
 Then, we sample a new candidate permutation from  $MM(\hat\sigma_0, \theta)$ (line~\ref{line:sample}).
@@
 minimal fitness values. This is achieved by means of the uBorda ranking
-(Section~\ref{sec:uborda}) with $w(\sigma) = \rho^{f(\sigma)}$ and $\rho \in (0,1)$, which %  can be seen as replicating each permutation
-% $\sigma \in S$ a number of times inversely proportional to its fitness
-% $f(\sigma)$, i.e., it 
-behaves as if there were many copies of the best
+(Section~\ref{sec:uborda}) with $w(\sigma) = \rho^{f(\sigma)}$ and $\rho \in (0,1)$. %, which %  can be seen as replicating each permutation
+Intuitively, uBorda behaves as if there were many copies of the best
 solutions and few of the bad ones. A property of UMM is that the
@@
 $w(\sigma)=\rho^{f(\sigma)}$ (Section~\ref{sec:uborda}). The learning rate is
-controlled by parameter $0 < \rho < 1$. \MANUEL{Don't we have to say something about $f(\sigma) > 1$?}
+controlled by parameter $0 < \rho < 1$. 
+\MANUEL{Don't we have to say something about $f(\sigma) > 1$?}
+\EKHINE{es decreciente igual}
 In UMM, the value of $\rho$ depends on the set of true fitness evaluations $F$
@@
 
-\subsection{Convergence guarantees}
-\ekhine{}{completar}
+%\paragraph{Quality guarantees}
 
@@
 %
-UMM uses the same budget of evaluations ($\FEmax$) as CEGO and initially
-samples $\minit=10$ solutions uniformly.
-%At each sampling step, we set
-%$\beta=10$, meaning that 10 permutations are generated and the farthest from
-%$S$ is selected for evaluation.
-The other two parameters of UMM are $r_1$ and $r_2$, which determine the
+UMM uses a budget of evaluations ($\FEmax$) of  $\minit=10$ solutions sampled uniformly.
+The other two parameters of UMM are $r_1$ and $r_2$ that determine the learning rate $\rho$ in uBorda, and consequently, determine the
 balance between exploration and exploitation. In the experimental section we
 will study their impact.
+
+In all
+experiments, we consider a maximum budget of $\FEmax=400$ evaluations of the
+actual objective function. In a white-box context, state-of-the-art algorithms
+for the LOP and PFSP typically evaluate thousands of solutions,
+thus, the budget considered here for the black-box context is extremely
+limited.
 
@@
 CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use a GA to optimize
+\ekhine{}{hay que actualizar essto??}
 the surrogate models with population size of 20, crossover rate of 0.5,
@@
 CEGO used a budget of $10^4$ surrogate-model evaluations for each run of the
-GA. However, with such a setting, a single run of CEGO with a budget of
-$\FEmax=400$ true fitness evaluations requires \emph{more than a week} of
-computation time, which was infeasible for our study. Due to extremely long
-runtimes of CEGO and constraints in our computing system, we set the budget of
-the GA to $10^3$ and we stopped each run of CEGO after 144 hours of CPU-time.
-Although a higher GA budget could improve the results of CEGO reported here,
-our results are similar (and sometimes better) than the ones reported in the
-original papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to
-running CEGO with $\FEmax=400$ instead of $\FEmax=200$. We found that
-doubling $\FEmax$ produces a smaller increase in runtime than multiplying the
-GA budget by a factor of $10$. Nevertheless, we caution that choosing a
-different trade-off may lead to different conclusions and more work is needed
-to understand CEGO parameters.
+GA and we use the same confuguration parameters.% However, with such a setting, a single run of CEGO with a budget of
+%$\FEmax=400$ true fitness evaluations requires \emph{more than a week} of
+%computation time, which was infeasible for our study. 
+%Due to extremely long
+%runtimes of CEGO and constraints in our computing system, we set the budget of
+%the GA to $10^3$ and we stopped each run of CEGO after 144 hours of CPU-time.
+%Although a higher GA budget could improve the results of CEGO reported here,
+%our results are similar (and sometimes better) than the ones reported in the
+%original papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to
+%running CEGO with $\FEmax=400$ instead of $\FEmax=200$. We found that
+%doubling $\FEmax$ produces a smaller increase in runtime than multiplying the
+%GA budget by a factor of $10$. Nevertheless, we caution that choosing a
+%different trade-off may lead to different conclusions and more work is needed
+%to understand CEGO parameters.
 
@@
 %when using Hamming distance and the same may apply to UMM.
-We plan to extend UMM to other distance measures in the future.  In all
-experiments, we consider a maximum budget of $\FEmax=400$ evaluations of the
-actual objective function. In a white-box context, state-of-the-art algorithms
-for the LOP and PFSP typically evaluate thousands of solutions,
-thus, the budget considered here for the black-box context is extremely
-limited.
+We plan to extend UMM to other distance measures in the future.
 
@@
 Instead, we consider classical combinatorial optimization problems as black-box optimization problems.
-For consistency with the rest of the paper, a permutation $\sigma$ denotes a ranking also in this section, thus we use the inverse permutation $\sigma^{-1}$ in the definition of the problems.
+For consistency with the rest of the paper, a permutation $\sigma$ denotes a ranking also in this section.
+Recall that the ordering associated to ranking $\sigma$ is given by its inverse, $\sigma^{-1}$.
+Thus, we will use ordering $\sigma^{-1}$ to formulate the problems.
 Later in the experimental section we discuss the effects of not performing this inversion.
-%
-The \emph{Linear Ordering Problem (LOP)}, when minimizing, is defined as
+
+The \emph{Linear Ordering Problem (LOP)}, when minimizing as in our case, is defined as
 %\footnote{For consistency with the other problems, we minimise the lower triangle instead of maximising the upper triangle.} 
@@
 All jobs must be processed by all machines in the same order.
-Given a permutation of the jobs $\sigma$, $C_{i,j}$ denotes the completion time of job at position $\sigma^{-1}(i)$ on machine $j$, and $C_{\max} = C_{n,M}$  is the completion time of the last job on the last machine, i.e., the makespan.
+Given a permutation (ranking) $\sigma$, $C_{i,j}$ denotes the completion time of job $\sigma^{-1}(i)$ at position $i$ on machine $j$, and $C_{\max} = C_{n,M}$  is the completion time of the last job on the last machine, i.e., the makespan.
 The objective of the PFSP is to find:
@@
 
-\subsubsection{Computing environment.}
+\subsubsection{Computing environment}
 % Due to the computational effort required, we used several computing systems to

### 2021-02-04 paper/main.tex
table

--- 2021-02-04 paper/main.tex
@@
 
-\renewcommand{\myparagraph}[1]{\smallskip{}\noindent\textbf{#1}.}
+\newcommand{\myparagraph}[1]{\smallskip{}\noindent\textbf{#1}.}
 
@@
 
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-\begin{table*}[]
-\begin{tabular}{rrrrrrrrr}
-\multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{4}{c}{\textbf{Mean fitness (sd)}}                       & \multicolumn{1}{c}{\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}c@{}}95\% CI of the \\ mean difference\end{tabular}}}} & \multicolumn{2}{c}{\textbf{Runtime}} \\ \cline{3-6} \cline{8-9}
-\textbf{Problem}              & \textbf{Instance}             & \multicolumn{2}{r}{\textbf{CEGO}} & \multicolumn{2}{r}{\textbf{UMM}} & \multicolumn{1}{c}{}                                                                                                    & \textbf{CEGO}     & \textbf{UMM}     \\ \hline
+\newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
+\begin{table*}[tb]
+  \caption{Mean fitness (and standard deviation) of the best solution found.
+% at the end of each run, over 10 independent runs for each problem instance.
+The 95\% confidence interval corresponds to the two independent samples t-test on the mean difference between the fitness of UMM minus the one of CEGO.
+  Mean runtime is shown in hours. 
+  %CEGO runs were stopped after 144 hours. This was the case for all LOP instances.
+  \label{tab:results}}
+  \resizebox{\textwidth}{!}{%
+\begin{tabular}{r@{\hskip -2ex}*{5}{r}rl@{\hskip -2ex}*{3}{r}}
+ \toprule
+            &              & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the                                                                                                                                                                                                                                                                                           \\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack[r]{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){10-11}
+\bf Problem & \bf Instance & \mcolc{2}{CEGO}              & \mcolc{2}{UMM}                                                                                                                                                                                                               &                                                          &           & \bf                       & \bf CEGO & \bf UMM \\\midrule
 LOP         & N-p40-01     & 10496.6                      & (116.8)                                                                                                                                                                                                                      & 12369.1                                                  & (389.9)   & {[}-2157.1, -1587.9{]}    & 19.9     & 0.3     \\
@@
 PFSP        & rec19        & 2264.5                       & (45.0)                                                                                                                                                                                                                       & 2304.8                                                   & (37.7)    & {[}-79.4, -1.2{]}         & 18.2     & 0.2     \\
-PFSP                          & rec31                         & 3415.5           & (62.1)         & 3419.3          & (37.0)         & {[}-52.6, 45.0{]}                                                                                                       & 23.1              & 0.5
-\end{tabular}
+PFSP        & rec31        & 3415.5                       & (62.1)                                                                                                                                                                                                                       & 3419.3                                                   & (37.0)    & {[}-52.6, 45.0{]}         & 23.1     & 0.5     \\
+\bottomrule
+\end{tabular}}
 \end{table*}
@@
 
+% \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
+% \begin{table}[tb]
+%   \caption{Mean fitness (and standard deviation) of the best solution found.
+% at the end of each run, over 10 independent runs for each problem instance.
+The 95\% confidence interval corresponds to the two independent samples t-test on the mean difference between the fitness of UMM minus the one of CEGO.
+%   Mean runtime is shown in hours. 
+%   %CEGO runs were stopped after 144 hours. This was the case for all LOP instances.
+%   \label{tab:results}}
+%   \resizebox{\textwidth}{!}{%
+% \begin{tabular}{r@{\hskip -2ex}*{5}{r}rl@{\hskip -2ex}*{3}{r}}
+%  \toprule
+%             &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack[r]{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){10-11}
+%             \bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule
+% LOP
+% & N-p50-01       & 21034.3             & (183.9)            & 21161.8            & (657.4)           & [-606.0,             & 351.0]              & 305.8        & 144.6 & 0.5  \\
+% & N-p50-02       & 21767.3             & (254.9)            & 21746.6            & (645.4)           & [-458.6,             & 500.0]              & 315.1        & 144.5 & 0.5  \\
+% & N-t1d100.01    & 78012.0             & (432.1)            & 76499.3            & (1099.7)          & [696.4,              & 2329.0]             & 158.3        & 145.1 & 1.9  \\
+% & N-t1d100.02    & 78834.0             & (477.8)            & 76860.7            & (864.0)           & [1303.8,             & 2642.8]             & 156.0        & 144.7 & 1.8  \\
+% & N-atp111       & 668.5               & (9.4)              & 596.6              & (17.0)            & [58.8,               & 85.0]               & 137.9        & 145.1 & 2.3  \\
+% & N-atp134       & 856.4               & (12.2)             & 760.1              & (23.4)            & [78.4,               & 114.2]              & 116.5        & 145.0 & 3.3  \\
+% & N-t1d150.01    & 186969.0            & (2094.6)           & 179615.6           & (1731.5)          & [5543.3,             & 9163.5]             & 103.9        & 145.7 & 4.4  \\
+% & N-t1d150.02    & 183414.7            & (779.2)            & 177524.2           & (1165.6)          & [4949.2,             & 6831.8]             & 103.6        & 145.3 & 4.6  \\
+% & N-t2d150.01    & 22207.3             & (1667.3)           & 15136.9            & (745.9)           & [5817.1,             & 8323.7]             & 104.5        & 145.5 & 4.5  \\
+% & N-t2d150.02    & 22036.0             & (744.7)            & 14718.4            & (677.5)           & [6648.3,             & 7986.9]             & 103.1        & 145.0 & 4.5  \\
+% & N-t1d200.01    & 333804.1            & (1191.9)           & 323501.4           & (1385.7)          & [9086.4,             & 11519.0]            & 77.7         & 145.4 & 7.6  \\
+% & N-t1d200.02    & 331730.9            & (2819.7)           & 319786.2           & (2265.6)          & [9533.6,             & 14355.8]            & 78.2         & 145.9 & 7.1  \\
+% & N-t2d200.01    & 54043.6             & (1477.6)           & 33389.0            & (1984.0)          & [19001.4,            & 22307.8]            & 78.2         & 146.7 & 7.4  \\
+% & N-t2d200.02    & 52787.0             & (4194.2)           & 35589.0            & (2319.6)          & [13948.0,            & 20448.0]            & 78.9         & 145.7 & 7.5  \\
+% & N-be75np\_150  & 3669753.9           & (74292.3)          & 3216907.1          & (72239.3)         & [383998.5,           & 521695.1]           & 105.2        & 145.5 & 4.4  \\
+% & N-be75np\_250  & 10818828.5          & (127722.3)         & 9582261.7          & (258176.8)        & [1040022.9,          & 1433110.7]          & 62.5         & 146.4 & 12.5 \\
+% & N-be75eec\_150 & 1592220.8           & (29945.9)          & 1418578.9          & (56557.7)         & [130141.0,           & 217142.8]           & 103.2        & 145.9 & 4.4  \\
+% & N-be75eec\_250 & 5098734.4           & (172874.0)         & 4430408.5          & (110233.0)        & [530352.4,           & 806299.4]           & 63.0         & 146.7 & 12.3 \\\midrule
+% PFSP
+% & rec05          & 1317.9              & (25.1)             & 1332.5             & (15.6)            & [-34.5,              & 5.3]                & 400.0        & 38.6  & 0.1  \\
+% & rec13          & 2128.2              & (25.0)             & 2163.6             & (28.1)            & [-60.4,              & -10.4]              & 400.0        & 38.6  & 0.1  \\
+% & rec19          & 2398.6              & (19.8)             & 2405.2             & (12.8)            & [-22.5,              & 9.3]                & 400.0        & 85.2  & 0.2  \\\midrule
+% QAP
+% & nug12          & 624.2               & (26.3)             & 661.2              & (17.8)            & [-58.3,              & -15.7]              & 400.0        & 14.2  & 0.0  \\
+% & kra30b         & 120192.0            & (1778.1)           & 119923.0           & (3459.6)          & [-2379.5,            & 2917.5]             & 400.0        & 85.1  & 0.2  \\
+% & tho30          & 192416.0            & (2584.2)           & 191586.4           & (4342.0)          & [-2582.9,            & 4242.1]             & 400.0        & 85.6  & 0.2  \\
+% & kra30a         & 119872.0            & (1746.7)           & 119502.0           & (1827.5)          & [-1309.8,            & 2049.8]             & 400.0        & 86.1  & 0.2  \\
+% & nug30          & 7486.8              & (40.4)             & 7472.4             & (83.6)            & [-49.0,              & 77.8]               & 400.0        & 86.5  & 0.2  \\
+% & kra32          & 115907.0            & (4663.4)           & 116898.0           & (4343.1)          & [-5226.3,            & 3244.3]             & 400.0        & 94.0  & 0.2  \\
+% \bottomrule
+% \end{tabular}}
+% \vspace*{-1em}
+%  \end{table}
+
 \section{Conclusions}\label{sec:conclusions}

### 2021-02-04 paper/main.tex

--- 2021-02-04 paper/main.tex
@@
 
-In all
-experiments, we consider a maximum budget of $\FEmax=400$ evaluations of the
+We consider 3 different values for the maximum budget of $\FEmax=100, 200,400$ evaluations of the
 actual objective function. In a white-box context, state-of-the-art algorithms
@@
 
-\subsubsection{CEGO} \EKHINE{tenemos una pagina entera dedicada a CEGO :s}\MANUEL{Sugiere qué quitar usando el commando \textbackslash{}ekhine}
 Combinatorial Efficient Global Optimization (CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO

### 2021-02-04 paper/main.tex
table

@@
   \resizebox{\textwidth}{!}{%
-\begin{tabular}{r@{\hskip -2ex}*{5}{r}rl@{\hskip -2ex}*{3}{r}}
+\begin{tabular}{r@{\hskip 0ex}*{5}{r}rl@{\hskip -2ex}*{2}{r}}
  \toprule
-            &              & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the                                                                                                                                                                                                                                                                                           \\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack[r]{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){10-11}
-\bf Problem & \bf Instance & \mcolc{2}{CEGO}              & \mcolc{2}{UMM}                                                                                                                                                                                                               &                                                          &           & \bf                       & \bf CEGO & \bf UMM \\\midrule
-LOP         & N-p40-01     & 10496.6                      & (116.8)                                                                                                                                                                                                                      & 12369.1                                                  & (389.9)   & {[}-2157.1, -1587.9{]}    & 19.9     & 0.3     \\
-LOP         & N-p40-02     & 11714.7                      & (84.7)                                                                                                                                                                                                                       & 13160.5                                                  & (454.8)   & {[}-1773.5, -1118.1{]}    & 20.1     & 0.3     \\
-LOP         & N-t59b11xx   & 102408.0                     & (3800.2)                                                                                                                                                                                                                     & 114075.5                                                 & (5389.2)  & {[}-16084.2, -7250.8{]}   & 21.1     & 0.4     \\
-LOP         & N-t59d11xx   & 12692.2                      & (1097.9)                                                                                                                                                                                                                     & 16662.9                                                  & (2155.3)  & {[}-5618.5, -2322.9{]}    & 21.2     & 0.4     \\
-LOP         & N-p50-02     & 19090.5                      & (285.4)                                                                                                                                                                                                                      & 21746.6                                                  & (645.4)   & {[}-3140.6, -2171.6{]}    & 22.9     & 0.5     \\
-LOP         & N-p50-01     & 18197.7                      & (225.6)                                                                                                                                                                                                                      & 21161.8                                                  & (657.4)   & {[}-3447.4, -2480.8{]}    & 23.2     & 0.5     \\
-LOP         & N-sgb75.01   & 1893533.2                    & (28594.6)                                                                                                                                                                                                                    & 1992059.9                                                & (40235.7) & {[}-131577.1, -65476.3{]} & 33.2     & 1.0     \\
-LOP         & N-sgb75.02   & 1895801.8                    & (82847.9)                                                                                                                                                                                                                    & 1967132.3                                                & (49228.1) & {[}-136421.2, -6239.8{]}  & 33.3     & 1.1     \\
-PFSP        & rec05        & 1276.2                       & (15.1)                                                                                                                                                                                                                       & 1276.3                                                   & (8.4)     & {[}-11.8, 11.6{]}         & 18.3     & 0.1     \\
-PFSP        & rec13        & 1992.9                       & (27.6)                                                                                                                                                                                                                       & 2057.8                                                   & (22.8)    & {[}-88.8, -41.0{]}        & 18.5     & 0.1     \\
-PFSP        & rec19        & 2264.5                       & (45.0)                                                                                                                                                                                                                       & 2304.8                                                   & (37.7)    & {[}-79.4, -1.2{]}         & 18.2     & 0.2     \\
-PFSP        & rec31        & 3415.5                       & (62.1)                                                                                                                                                                                                                       & 3419.3                                                   & (37.0)    & {[}-52.6, 45.0{]}         & 23.1     & 0.5     \\
+            &              & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the                                                                                       \\mean difference}}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){9-10}
+\bf Problem & \bf Instance & \mcolc{2}{CEGO}              & \mcolc{2}{UMM}                                                                  &           &           & \bf CEGO                  & \bf UMM    \\\midrule
+LOP         & N-p40-01     & 10496.6                      & (116.8)                                                                         & 12369.1   & (389.9)   & {[}-2157.1,& -1587.9{]}    & 19.9 & 0.3 \\
+LOP         & N-p40-02     & 11714.7                      & (84.7)                                                                          & 13160.5   & (454.8)   & {[}-1773.5,& -1118.1{]}    & 20.1 & 0.3 \\
+LOP         & N-t59b11xx   & 102408.0                     & (3800.2)                                                                        & 114075.5  & (5389.2)  & {[}-16084.2,& -7250.8{]}   & 21.1 & 0.4 \\
+LOP         & N-t59d11xx   & 12692.2                      & (1097.9)                                                                        & 16662.9   & (2155.3)  & {[}-5618.5,& -2322.9{]}    & 21.2 & 0.4 \\
+LOP         & N-p50-02     & 19090.5                      & (285.4)                                                                         & 21746.6   & (645.4)   & {[}-3140.6,& -2171.6{]}    & 22.9 & 0.5 \\
+LOP         & N-p50-01     & 18197.7                      & (225.6)                                                                         & 21161.8   & (657.4)   & {[}-3447.4,& -2480.8{]}    & 23.2 & 0.5 \\
+LOP         & N-sgb75.01   & 1893533.2                    & (28594.6)                                                                       & 1992059.9 & (40235.7) & {[}-131577.1,& -65476.3{]} & 33.2 & 1.0 \\
+LOP         & N-sgb75.02   & 1895801.8                    & (82847.9)                                                                       & 1967132.3 & (49228.1) & {[}-136421.2,& -6239.8{]}  & 33.3 & 1.1 \\
+PFSP        & rec05        & 1276.2                       & (15.1)                                                                          & 1276.3    & (8.4)     & {[}-11.8,& 11.6{]}         & 18.3 & 0.1 \\
+PFSP        & rec13        & 1992.9                       & (27.6)                                                                          & 2057.8    & (22.8)    & {[}-88.8,& -41.0{]}        & 18.5 & 0.1 \\
+PFSP        & rec19        & 2264.5                       & (45.0)                                                                          & 2304.8    & (37.7)    & {[}-79.4,& -1.2{]}         & 18.2 & 0.2 \\
+PFSP        & rec31        & 3415.5                       & (62.1)                                                                          & 3419.3    & (37.0)    & {[}-52.6,& 45.0{]}         & 23.1 & 0.5 \\
 \bottomrule

### 2021-02-04 paper/main.tex
table

@@
 to the best (minimum) fitness found at the end of the run by each algorithm. We
-show in Table~\ref{tab:results}
+show in Table~\ref{tab:results} the mean fitness of the best solution found
+over the 10 runs of each algorithm.
 
@@
   \resizebox{\textwidth}{!}{%
-\begin{tabular}{r@{\hskip 0ex}*{5}{r}rl@{\hskip -2ex}*{2}{r}}
+\begin{tabular}{r*{5}{r}r@{\hskip 0.5ex}l*{2}{r}}
  \toprule
             &              & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the                                                                                       \\mean difference}}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){9-10}
-\bf Problem & \bf Instance & \mcolc{2}{CEGO}              & \mcolc{2}{UMM}                                                                  &           &           & \bf CEGO                  & \bf UMM    \\\midrule
+\bf Problem & \bf Instance & \mcolc{2}{\CEGOinv}              & \mcolc{2}{UMM$_{400}$}                                                                  &           &           & \bf \CEGOinv                  & \bf UMM$_{400}$    \\\midrule
 LOP         & N-p40-01     & 10496.6                      & (116.8)                                                                         & 12369.1   & (389.9)   & {[}-2157.1,& -1587.9{]}    & 19.9 & 0.3 \\
-LOP         & N-p40-02     & 11714.7                      & (84.7)                                                                          & 13160.5   & (454.8)   & {[}-1773.5,& -1118.1{]}    & 20.1 & 0.3 \\
-LOP         & N-t59b11xx   & 102408.0                     & (3800.2)                                                                        & 114075.5  & (5389.2)  & {[}-16084.2,& -7250.8{]}   & 21.1 & 0.4 \\
-LOP         & N-t59d11xx   & 12692.2                      & (1097.9)                                                                        & 16662.9   & (2155.3)  & {[}-5618.5,& -2322.9{]}    & 21.2 & 0.4 \\
-LOP         & N-p50-02     & 19090.5                      & (285.4)                                                                         & 21746.6   & (645.4)   & {[}-3140.6,& -2171.6{]}    & 22.9 & 0.5 \\
-LOP         & N-p50-01     & 18197.7                      & (225.6)                                                                         & 21161.8   & (657.4)   & {[}-3447.4,& -2480.8{]}    & 23.2 & 0.5 \\
-LOP         & N-sgb75.01   & 1893533.2                    & (28594.6)                                                                       & 1992059.9 & (40235.7) & {[}-131577.1,& -65476.3{]} & 33.2 & 1.0 \\
-LOP         & N-sgb75.02   & 1895801.8                    & (82847.9)                                                                       & 1967132.3 & (49228.1) & {[}-136421.2,& -6239.8{]}  & 33.3 & 1.1 \\
+         & N-p40-02     & 11714.7                      & (84.7)                                                                          & 13160.5   & (454.8)   & {[}-1773.5,& -1118.1{]}    & 20.1 & 0.3 \\
+         & N-t59b11xx   & 102408.0                     & (3800.2)                                                                        & 114075.5  & (5389.2)  & {[}-16084.2,& -7250.8{]}   & 21.1 & 0.4 \\
+         & N-t59d11xx   & 12692.2                      & (1097.9)                                                                        & 16662.9   & (2155.3)  & {[}-5618.5,& -2322.9{]}    & 21.2 & 0.4 \\
+         & N-p50-02     & 19090.5                      & (285.4)                                                                         & 21746.6   & (645.4)   & {[}-3140.6,& -2171.6{]}    & 22.9 & 0.5 \\
+         & N-p50-01     & 18197.7                      & (225.6)                                                                         & 21161.8   & (657.4)   & {[}-3447.4,& -2480.8{]}    & 23.2 & 0.5 \\
+         & N-sgb75.01   & 1893533.2                    & (28594.6)                                                                       & 1992059.9 & (40235.7) & {[}-131577.1,& -65476.3{]} & 33.2 & 1.0 \\
+         & N-sgb75.02   & 1895801.8                    & (82847.9)                                                                       & 1967132.3 & (49228.1) & {[}-136421.2,& -6239.8{]}  & 33.3 & 1.1 \\
+         \midrule
 PFSP        & rec05        & 1276.2                       & (15.1)                                                                          & 1276.3    & (8.4)     & {[}-11.8,& 11.6{]}         & 18.3 & 0.1 \\
-PFSP        & rec13        & 1992.9                       & (27.6)                                                                          & 2057.8    & (22.8)    & {[}-88.8,& -41.0{]}        & 18.5 & 0.1 \\
-PFSP        & rec19        & 2264.5                       & (45.0)                                                                          & 2304.8    & (37.7)    & {[}-79.4,& -1.2{]}         & 18.2 & 0.2 \\
-PFSP        & rec31        & 3415.5                       & (62.1)                                                                          & 3419.3    & (37.0)    & {[}-52.6,& 45.0{]}         & 23.1 & 0.5 \\
+       & rec13        & 1992.9                       & (27.6)                                                                          & 2057.8    & (22.8)    & {[}-88.8,& -41.0{]}        & 18.5 & 0.1 \\
+       & rec19        & 2264.5                       & (45.0)                                                                          & 2304.8    & (37.7)    & {[}-79.4,& -1.2{]}         & 18.2 & 0.2 \\
+       & rec31        & 3415.5                       & (62.1)                                                                          & 3419.3    & (37.0)    & {[}-52.6,& 45.0{]}         & 23.1 & 0.5 \\
 \bottomrule

### 2021-02-04 paper/main.tex
table
paper s5 conflics solved

@@
 %
-\MANUEL{Reviewer 3:  Justify the form of the formula used to learn the "learning rate" $\rho$ (equation 4). Why this particular approach was chosen?}
 % 
@@
 where $|S'| = r_2|S| \land \forall \tuple{\sigma',f(\sigma')}\in S', \tuple{\sigma,f(\sigma)}\in S\setminus S' : f(\sigma')\leq f(\sigma)$, that is, $S'$ contains the $r_2|S|$ elements of $S$ with the best  fitness values.
+
 %The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification.
@@
 \myparagraph{Comparison of UMM and CEGO}
-
+%
 When comparing the UMM and CEGO configurations shown in Fig.~\ref{fig:results},
@@
 
-\myparagraph{Comparison of the best solutions found} We now turn our analysis
-to the best (minimum) fitness found at the end of the run by each algorithm. We
-show in Table~\ref{tab:results} the mean fitness of the best solution found
-over the 10 runs of each algorithm.
+\myparagraph{Comparison of the best solutions found}
+%
+We now turn our analysis to the best (minimum) fitness found at the end of the
+run by the two best algorithms, that is, \CEGOinv and UMM$_{400}$. We show in
+Table~\ref{tab:results} the mean fitness (and standard deviation) of the best
+solution found over the 10 runs of each algorithm, together with a 95\%
+confidence interval (CI) around their mean fitness difference. Except for PFSP
+instances \texttt{rec05} and \texttt{rec31}, the differences are statistically
+significant in favour of \CEGOinv (the intervals do not contain the value
+zero). Nevertheless, the CIs indicate that the differences, although
+consistent, are relatively small.
+
+\myparagraph{Runtime comparison}
+%
+Table~\ref{tab:results} also shows the mean
+runtime of \CEGOinv and UMM$_{400}$ in hours. The difference in computational
+effort between \CEGOinv and UMM is at least two orders of magnitude.
 
@@
 \begin{table*}[tb]
-  \caption{Mean fitness (and standard deviation) of the best solution found.
-% at the end of each run, over 10 independent runs for each problem instance.
-The 95\% confidence interval corresponds to the two independent samples t-test on the mean difference between the fitness of UMM minus the one of CEGO.
+  \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance.
+The 95\% confidence interval corresponds to the two independent samples t-test on the mean difference between the fitness of \CEGOinv minus the one of UMM$_{400}$.
   Mean runtime is shown in hours. 
@@
   \label{tab:results}}
-  \resizebox{\textwidth}{!}{%
-\begin{tabular}{r*{5}{r}r@{\hskip 0.5ex}l*{2}{r}}
+ % \resizebox{\textwidth}{!}{%
+\begin{tabular}{r@{\hskip -1ex}r*{2}{r@{\hskip 1ex}l}r@{\hskip 0.5ex}lr@{\hskip 0.75ex}r}
  \toprule
@@
 \bottomrule
-\end{tabular}}
+\end{tabular}
+%}
 \end{table*}

### 2021-02-04 paper/main.tex
paper s5 conflics solved

--- 2021-02-04 paper/main.tex
@@
 permutation spaces. This algorithm is a probabilistic method based on the
-Mallows model and uBorda learning.\MANUEL{Perhaps we should call it UMM Algorithm? otherwise people would complain. Or Unbalanced Mallows Algorithm (UMA)?}
-
-\EKHINE{10.
-In Algorithm 1 'S' and 'F' are *sets* (thus, unordered).
-Therefore, when adding sigma and f(sigma) the correspondence between the two elements seems to be lost.
-This mapping would be better represented as a set of *pairs* < sigma, f(sigma) >.
-You always add these elements to both sets at the same time in step 8, anyway.
-MANUEL: FIXED
-}
+Mallows model and uBorda learning.%\MANUEL{Perhaps we should call it UMM Algorithm? otherwise people would complain. Or Unbalanced Mallows Algorithm (UMA)?}
 
@@
 A decreasing $\theta$ parameter along iterations (line~\ref{line:theta}) implies a decreasing variance of the Mallows model (Eq.~\ref{eq:MM}), as explained in  Section~\ref{sec:sampling} below. 
-Then, we sample a new candidate permutation from  $MM(\hat\sigma_0, \theta)$ (line~\ref{line:sample}).
+Then, we sample a new candidate permutation from  $MM(\hat\sigma_0, \theta)$ (line~\ref{line:sample}, also  in  Section~\ref{sec:sampling}).
 This candidate permutation (solution) is evaluated and added to the sample $S$.
 The process is repeated up to a given maximum budget of evaluations $\FEmax$.
-\EKHINE{(3) In section 3, second half of the second paragraph (from "Then, we sample a new candidate [...]"): the algorithm 1 is not properly described.
+%\EKHINE{(3) In section 3, second half of the second paragraph (from "Then, we sample a new candidate [...]"): the algorithm 1 is not properly described.
 Please provide a description in line with algorithm 1 and, above all, describe the rationale behind the choice of line 7 of algorithm 1.
-I like it, but it deserves to be better described.
+%I like it, but it deserves to be better described.
 It seems to me a sort of "novelty-driven" optimization.(4) A small detail, but requires a clarification.
@@
 Clearly the authors are working with minimization problems and want to increase the weight when fitness improves/decreases.
-This is ok, but this only happens with their formula when fitnesses are greater than 1.
+This is ok, but this only happens with their formula when fitnesses are greater than 1 NOOO.
 I know this is quite usual in the problems analyzed by the authors, but it may not be so usual in general.
@@
 
-As a consequence of this important distinction, when dealing with \manuel{data in the form of}{} permutations, in particular in 
+As a consequence of this important distinction, when dealing with  permutations, in particular in 
 combinatorial problems, we must convert rankings to orderings by inverting the
@@
 with a linear trend along iterations. The (close to) linear decrease is given
-by the linear decrease in the expected distance\MANUEL{distance to the sampling model or variance of the sampling model?} of the sampling model, as described in Section~\ref{sec:sampling}.
-Overall, the rate of improvement
+by the linear decrease in the expected distance of the model, as described in Section~\ref{sec:sampling}. Overall, the rate of improvement
 does not seem to converge, not even for the longer runs of 400 evaluations.
@@
 evaluations.}{}\MANUEL{We should discuss the time when discussing the tables.}
+\EKHINE{luego el detalle, que la gente no llega al final }
 
@@
 \myparagraph{Comparison of UMM and CEGO}
-
 When comparing the UMM and CEGO configurations shown in Fig.~\ref{fig:results},
@@
 Moreover, UMM seems to scale good for larger instances in quality performance. Note that the instances are order from top to bottom in the size of the permutations. 
-The computational complexity of UMM is \ekhine{}{ $O(n \log n)$ }at each iteration while both versions of CEGO need \manuel{tu sabes}{?}.
+The computational complexity of UMM is $O(n \log n)$ at each iteration while both versions of CEGO need \manuel{tu sabes}{?}.
 This difference is evident in the time performance of the algorithms.
@@
   \texttt{N-p50-01} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
+  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-02}
 \end{minipage}
@@
   \resizebox{\textwidth}{!}{%
-\begin{tabular}{r*{5}{r}r@{\hskip 0.5ex}l*{2}{r}}
+\begin{tabular}{r@{\hskip -1ex}r*{2}{r@{\hskip 1ex}l}r@{\hskip 0.5ex}lr@{\hskip 0.75ex}r}
  \toprule

### 2021-02-04 paper/main.tex
paper full

@@
 For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.\MANUEL{necesitamos las ecuaciones?
-Swap de una permutación es simplemente la distancia de kendall de la permutación inversa.} %The concepts of rankings and orderings are usually confused of ignored.
+Swap de una permutación es simplemente la distancia de kendall de la permutación inversa.
+.}
+\EKHINE{asi mas claro, no se van a perder con esto}
+ %The concepts of rankings and orderings are usually confused of ignored.
 An example of such confusion is the claim that Swap distance is the same metric as  the \ken distance, which is false except for particular permutations.
@@
 During our experiments, we realized that CEGO claimed to use the Swap distance. Instead, CEGO implements the \ken distance, yet CEGO did
-not invert the permutation before evaluating the PFSP.\MANUEL{no me convence esta forma de expresarlo porque la crítica obvia sería que por qué invertimos la función objectivo y no la función de distancia?} We verified that this is
+not invert the permutation before evaluating the PFSP.\MANUEL{no me convence esta forma de expresarlo porque la crítica obvia sería que por qué invertimos la función objectivo y no la función de distancia?} \EKHINE{es lo mismo}We verified that this is
 indeed how the implementation works and that those results match the published
@@
 
-In the following, we compare 5 algorithm configurations. Three configurations
-of UMM (UMM$_{100}$, UMM$_{200}$, and UMM$_{400}$) that differ in their budget
-of evaluations ($m$). As mentioned in Sec.~\ref{sec:sampling}, the behavior of
-UMM depends on the value of $m$. The other two configurations correspond 
-to \CEGOorig and \CEGOinv, as defined above.
+%In the following, we compare 5 algorithm configurations.
+% Three configurations
+%of UMM (UMM$_{100}$, UMM$_{200}$, and UMM$_{400}$) that differ in their budget
+%of evaluations ($m$). As mentioned in Sec.~\ref{sec:sampling}, the behavior of
+%UMM depends on the value of $m$. The other two configurations correspond 
+%to \CEGOorig and \CEGOinv, as defined above.
+
+Figure~\ref{fig:results} shows the results on the real-world instances from LOLIB and PFSP.
+There is one figure per instance.
+In each figure there is one horizontal line displaying the best known solutions for the problem.
+Besides, there are 5 different lines representing each of the different algorithm/versions considered, in particular:
+\begin{itemize}
+\item UMM\_100, UMM\_200 and UMM\_400 are three different versions for UMM, each considering a different total number of function evaluations, 100, 200 and 400.
+Obviously, the lines corresponding to UMM\_100, UMM\_200 wont reach the right corner of the figure.
+As mentioned in Sec.~\ref{sec:sampling}, the behavior of
+UMM depends on the value of $m$. 
+\item CEGO\_orig and CEGO\_inv are, respectively, the original version of CEGO and the modification of Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$.
+\end{itemize}
 
@@
 evaluated solution provides better insights on the search dynamics of the
-algorithms analyzed here. Please note that the algorithms do not return the
+algorithms analyzed here, despite it is not directly comparable with most literature results. Please note that the algorithms do not return the
 final solution shown in the plots, but rather the best one (minimum fitness) of
@@
 fitness of the best-so-far solution up to that point in the run. We provide
-later an analysis of the best solution found by each run of the algorithm.
+later an analysis of the best solution found by each run of the algorithm, which are comparable with the original paper of CEGO.
 
@@
 Moreover, a non-linear decrease in the expected distance should result in a faster convergence for larger evaluation budgets.
-\manuel{The
-average runtime for all the instances is around 0.3 minutes, 1.15 minutes and
-4.5 minutes respectively for the versions with 100, 200 and 400 fitness
-evaluations.}{}\MANUEL{We should discuss the time when discussing the tables.}
-\EKHINE{luego el detalle, que la gente no llega al final }
 
@@
 
-\myparagraph{Comparison of the best solutions found} We now turn our analysis
-to the best (minimum) fitness found at the end of the run by each algorithm. We
-show in Table~\ref{tab:results} the mean fitness of the best solution found
-over the 10 runs of each algorithm.
-
-For the reason for the abrupt change between both version for
-every instance, we refer the reader to Section~\ref{sec:exper_ro}. For the
-original version, there is a global decrease in the fitness function. However,
-a closer analysis of each iteration does not show a real convergence to any
-point. However, with the modified version in which rankings are modeled and
-orderings are evaluated we can appreciate the a decrease in terms of fitness
-per evaluation. Overall, the modified CEGO has a very abrupt decrease at the
-first 50 iterations. At one point, the algorithm seems to have converged and
-very little improvement is don from the 150-th iteration onwards.  The average
+\myparagraph{Comparison of the best solutions found}
+%
+We now turn our analysis to the best (minimum) fitness found at the end of the
+run by the two best algorithms, that is, \CEGOinv and UMM$_{400}$. We show in
+Table~\ref{tab:results} the mean fitness (and standard deviation) of the best
+solution found over the 10 runs of each algorithm, together with a 95\%
+confidence interval (CI) around their mean fitness difference. Except for PFSP
+instances \texttt{rec05} and \texttt{rec31}, the differences are statistically
+significant in favour of \CEGOinv (the intervals do not contain the value
+zero). Nevertheless, the CIs indicate that the differences, although
+consistent, are relatively small.
+
+\myparagraph{Runtime comparison}
+%
+The average
 running time is approximately 22 hours for the 400 function evaluations, which
@@
 
-\manuel{%
-This result summarization is not the classical.
-However, due to the limited number of function evaluations, we can have a better picture of the behaviour of the algorithm at each step.
-As a consequence, results displayed in the figures are not comparable with the original paper of CEGO: for this comparison we provide in Table~\ref{table:fitness} the summary of the best solutions found by each algorithm, which are similar to the original study.
-}{}
-
-\paragraph{Figure descriptiopn} Figure~\ref{fig:results} shows the results on the real-world instances from LOLIB and PFSP.
-There is one figure per instance.
-In each figure there is one horizontal line displaying the best known solutions for the problem.
-Besides, there are 5 different lines representing each of the different algorithm/versions considered, in particular:
-\begin{itemize}
-\item UMM\_100, UMM\_200 and UMM\_400 are three different versions for UMM, each considering a different total number of function evaluations, 100, 200 and 400.
-Obviously, the lines corresponding to UMM\_100, UMM\_200 wont reach the right corner of the figure.
-\item CEGO\_orig and CEGO\_inv are, respectively, the original version of CEGO and the modification of Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$.
-\end{itemize}
-
-\paragraph{Analysis} We start our analysis with the three version of UMM.
-We can see a general trend for all the instances: After the 10 random initial permutations, the fitness of the new solutions decreases with a linear trend along iterations.
-The (close to) linear decrease is given by the linear decrease in the expected distance of the model which is sampled at each step, described in Section~\ref{sec:sampling}.
-Overall, the improvement does not seem to converge not even for the longer runs of 400 evaluations.
+Table~\ref{tab:results} shows a brief comparison of time and the full ones are given in Table~\ref{tab:time}. The difference is time in impressive. 
 The average runtime for all the instances is around 0.3 minutes, 1.15 minutes and 4.5 minutes respectively for the  versions with 100, 200 and 400 fitness evaluations. 
-
-The red and the gold lines correspond to CEGO, the original and our modified versions.
-For the reason for the abrupt change between both version for every instance, we refer the reader to Section~\ref{sec:exper_ro}.
-For the original version, there is a global decrease in the fitness function.
-However, a closer analysis of each iteration does not show a real convergence to any point.
-However, with the modified version in which rankings are modeled and orderings are evaluated we can appreciate the a decrease in terms of fitness per evaluation.
-Overall, the modified CEGO has a very abrupt decrease at the first 50 iterations.
-At one point, the algorithm seems to have converged and very little improvement is don from the 150-th iteration onwards.
-The average running time is approximately 22 hours for the 400 function evaluations, which is 300 times slower than UMM for the same number of function evaluations.
-
-The performance of the UMM is clearly better than the performance of the original CEGO for the same problem.
-For CEGO\_inv on the other hand improves the results of the UMM with the current experimental setting.
-However, UMM seems to be improving the quality of the solutions until the last iteration while CEGO\_inv seems to be stucked at a local optimum.
-Further analysis lines will consider a larger budget of fitness evaluation.
-
-Moreover, UMM seems to scale good for larger instances in quality performance. Note that the instances are order from top to bottom in the size of the permutations. 
-The computational complexity of UMM is $O(n \log n)$ at each iteration while both versions of CEGO need \manuel{tu sabes}{?}.
+For CEGO in both versions the average time for 400 evaluations is more than 22 hour. This is 300 times slower than UMM for the same number of function evaluations. 
+Moreover, UMM scales good for larger instances in quality performance. 
+The computational complexity of UMM is $O(n \log n)$ at each iteration while both versions of CEGO need \ekhine{MANUEL, tu sabes}{?}.
 This difference is evident in the time performance of the algorithms.
@@
 \begin{minipage}{0.49\linewidth}
-  \texttt{N-p50-01} (LOP)\\[-0.5ex]
+  \texttt{N-p40-04} (LOP)\\[-0.5ex]
   \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-02}
@@
 
-\paragraph{Discussion on other problems: the case of QAP}
+%
+%\paragraph{Figure description} Figure~\ref{fig:results} shows the results on the real-world instances from LOLIB and PFSP.
+There is one figure per instance.
+In each figure there is one horizontal line displaying the best known solutions for the problem.
+Besides, there are 5 different lines representing each of the different algorithm/versions considered, in particular:
+%\begin{itemize}
+%\item UMM\_100, UMM\_200 and UMM\_400 are three different versions for UMM, each considering a different total number of function evaluations, 100, 200 and 400.
+Obviously, the lines corresponding to UMM\_100, UMM\_200 wont reach the right corner of the figure.
+%\item CEGO\_orig and CEGO\_inv are, respectively, the original version of CEGO and the modification of Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$.
+%\end{itemize}
+%
+%\paragraph{Analysis} We start our analysis with the three version of UMM.
+We can see a general trend for all the instances: After the 10 random initial permutations, the fitness of the new solutions decreases with a linear trend along iterations.
+The (close to) linear decrease is given by the linear decrease in the expected distance of the model which is sampled at each step, described in Section~\ref{sec:sampling}.
+Overall, the improvement does not seem to converge not even for the longer runs of 400 evaluations.
+%The average runtime for all the instances is around 0.3 minutes, 1.15 minutes and 4.5 minutes respectively for the  versions with 100, 200 and 400 fitness evaluations. 
+%
+%
+%The red and the gold lines correspond to CEGO, the original and our modified versions.
+For the reason for the abrupt change between both version for every instance, we refer the reader to Section~\ref{sec:exper_ro}.
+For the original version, there is a global decrease in the fitness function.
+However, a closer analysis of each iteration does not show a real convergence to any point.
+However, with the modified version in which rankings are modeled and orderings are evaluated we can appreciate the a decrease in terms of fitness per evaluation.
+Overall, the modified CEGO has a very abrupt decrease at the first 50 iterations.
+At one point, the algorithm seems to have converged and very little improvement is don from the 150-th iteration onwards.
+The average running time is approximately 22 hours for the 400 function evaluations, which is 300 times slower than UMM for the same number of function evaluations.
+%
+%The performance of the UMM is clearly better than the performance of the original CEGO for the same problem.
+For CEGO\_inv on the other hand improves the results of the UMM with the current experimental setting.
+However, UMM seems to be improving the quality of the solutions until the last iteration while CEGO\_inv seems to be stucked at a local optimum.
+Further analysis lines will consider a larger budget of fitness evaluation.
+%
+%Moreover, UMM seems to scale good for larger instances in quality performance. Note that the instances are order from top to bottom in the size of the permutations. 
+%The computational complexity of UMM is $O(n \log n)$ at each iteration while both versions of CEGO need \manuel{tu sabes}{?}.
+This difference is evident in the time performance of the algorithms.
+The details can be seen in Table~\ref{table:time}.
+%
+
+\newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
+\begin{table*}[tb]
+  \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance.
+The 95\% confidence interval corresponds to the two independent samples t-test on the mean difference between the fitness of \CEGOinv minus the one of UMM$_{400}$.
+  Mean runtime is shown in hours. 
+  %CEGO runs were stopped after 144 hours. This was the case for all LOP instances.
+  \label{tab:results}}
+ % \resizebox{\textwidth}{!}{%
+\begin{tabular}{r@{\hskip -1ex}r*{2}{r@{\hskip 1ex}l}r@{\hskip 0.5ex}lr@{\hskip 0.75ex}r}
+ \toprule
+            &              & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the                                                                                       \\mean difference}}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){9-10}
+\bf Problem & \bf Instance & \mcolc{2}{\CEGOinv}              & \mcolc{2}{UMM$_{400}$}                                                                  &           &           & \bf \CEGOinv                  & \bf UMM$_{400}$    \\\midrule
+LOP         & N-p40-01     & 10496.6                      & (116.8)                                                                         & 12369.1   & (389.9)   & {[}-2157.1,& -1587.9{]}    & 19.9 & 0.3 \\
+         & N-p40-02     & 11714.7                      & (84.7)                                                                          & 13160.5   & (454.8)   & {[}-1773.5,& -1118.1{]}    & 20.1 & 0.3 \\
+         & N-t59b11xx   & 102408.0                     & (3800.2)                                                                        & 114075.5  & (5389.2)  & {[}-16084.2,& -7250.8{]}   & 21.1 & 0.4 \\
+         & N-t59d11xx   & 12692.2                      & (1097.9)                                                                        & 16662.9   & (2155.3)  & {[}-5618.5,& -2322.9{]}    & 21.2 & 0.4 \\
+         & N-sgb75.01   & 1893533.2                    & (28594.6)                                                                       & 1992059.9 & (40235.7) & {[}-131577.1,& -65476.3{]} & 33.2 & 1.0 \\
+         & N-sgb75.02   & 1895801.8                    & (82847.9)                                                                       & 1967132.3 & (49228.1) & {[}-136421.2,& -6239.8{]}  & 33.3 & 1.1 \\
+         \midrule
+PFSP        & rec05        & 1276.2                       & (15.1)                                                                          & 1276.3    & (8.4)     & {[}-11.8,& 11.6{]}         & 18.3 & 0.1 \\
+       & rec13        & 1992.9                       & (27.6)                                                                          & 2057.8    & (22.8)    & {[}-88.8,& -41.0{]}        & 18.5 & 0.1 \\
+       & rec19        & 2264.5                       & (45.0)                                                                          & 2304.8    & (37.7)    & {[}-79.4,& -1.2{]}         & 18.2 & 0.2 \\
+       & rec31        & 3415.5                       & (62.1)                                                                          & 3419.3    & (37.0)    & {[}-52.6,& 45.0{]}         & 23.1 & 0.5 \\
+\bottomrule
+\end{tabular}
+%}
+\end{table*}
 
@@
 
-The overall conclusion from Table~\ref{tab:results} is that there is not a
-significant difference in performance between UMM and CEGO. For the larger LOP
-instances, CEGO reaches the time limit of 144 hours before consuming the budget
-of 400 evaluations and thus, obtains quite poor solutions compared to UMM,
-which is at least an order of magnitude faster.
-
-\newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
-\begin{table*}[tb]
-  \caption{Mean fitness (and standard deviation) of the best solution found.
-% at the end of each run, over 10 independent runs for each problem instance.
-The 95\% confidence interval corresponds to the two independent samples t-test on the mean difference between the fitness of UMM minus the one of CEGO.
-  Mean runtime is shown in hours. 
-  %CEGO runs were stopped after 144 hours. This was the case for all LOP instances.
-  \label{tab:results}}
-  \resizebox{\textwidth}{!}{%
-\begin{tabular}{r@{\hskip -1ex}r*{2}{r@{\hskip 1ex}l}r@{\hskip 0.5ex}lr@{\hskip 0.75ex}r}
- \toprule
-            &              & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI of the                                                                                       \\mean difference}}} & \mcolc{2}{Mean runtime} \\\cmidrule(lr){3-6}\cmidrule(lr){9-10}
-\bf Problem & \bf Instance & \mcolc{2}{\CEGOinv}              & \mcolc{2}{UMM$_{400}$}                                                                  &           &           & \bf \CEGOinv                  & \bf UMM$_{400}$    \\\midrule
-LOP         & N-p40-01     & 10496.6                      & (116.8)                                                                         & 12369.1   & (389.9)   & {[}-2157.1,& -1587.9{]}    & 19.9 & 0.3 \\
-         & N-p40-02     & 11714.7                      & (84.7)                                                                          & 13160.5   & (454.8)   & {[}-1773.5,& -1118.1{]}    & 20.1 & 0.3 \\
-         & N-t59b11xx   & 102408.0                     & (3800.2)                                                                        & 114075.5  & (5389.2)  & {[}-16084.2,& -7250.8{]}   & 21.1 & 0.4 \\
-         & N-t59d11xx   & 12692.2                      & (1097.9)                                                                        & 16662.9   & (2155.3)  & {[}-5618.5,& -2322.9{]}    & 21.2 & 0.4 \\
-         & N-p50-02     & 19090.5                      & (285.4)                                                                         & 21746.6   & (645.4)   & {[}-3140.6,& -2171.6{]}    & 22.9 & 0.5 \\
-         & N-p50-01     & 18197.7                      & (225.6)                                                                         & 21161.8   & (657.4)   & {[}-3447.4,& -2480.8{]}    & 23.2 & 0.5 \\
-         & N-sgb75.01   & 1893533.2                    & (28594.6)                                                                       & 1992059.9 & (40235.7) & {[}-131577.1,& -65476.3{]} & 33.2 & 1.0 \\
-         & N-sgb75.02   & 1895801.8                    & (82847.9)                                                                       & 1967132.3 & (49228.1) & {[}-136421.2,& -6239.8{]}  & 33.3 & 1.1 \\
-         \midrule
-PFSP        & rec05        & 1276.2                       & (15.1)                                                                          & 1276.3    & (8.4)     & {[}-11.8,& 11.6{]}         & 18.3 & 0.1 \\
-       & rec13        & 1992.9                       & (27.6)                                                                          & 2057.8    & (22.8)    & {[}-88.8,& -41.0{]}        & 18.5 & 0.1 \\
-       & rec19        & 2264.5                       & (45.0)                                                                          & 2304.8    & (37.7)    & {[}-79.4,& -1.2{]}         & 18.2 & 0.2 \\
-       & rec31        & 3415.5                       & (62.1)                                                                          & 3419.3    & (37.0)    & {[}-52.6,& 45.0{]}         & 23.1 & 0.5 \\
-\bottomrule
-\end{tabular}}
-\end{table*}
-
 \begin{table*}[tb]%time
@@
 \midrule
-    LOP &  N-t59b11xx &   1265.98 &    1240.36 &    22.24 &     5.78 &     1.37 \\
-    LOP &  N-t59d11xx &   1269.93 &    1252.89 &    21.84 &     5.62 &     1.37 \\
-    LOP &    N-p40-01 &   1195.90 &    1170.81 &    18.13 &     4.61 &     1.14 \\
-    LOP &    N-p40-02 &   1207.24 &    1183.09 &    18.11 &     4.76 &     1.15 \\
-    LOP &    N-p50-01 &   1391.55 &        NaN &    27.36 &      NaN &      NaN \\
-    LOP &    N-p50-02 &   1372.71 &        NaN &    28.36 &      NaN &      NaN \\
-    LOP &  N-sgb75.01 &   1994.22 &    2004.70 &    62.75 &    15.77 &     3.97 \\
-    LOP &  N-sgb75.02 &   1995.09 &    1975.41 &    64.58 &    15.90 &     4.03 \\
-   PFSP &       rec05 &   1099.63 &     917.81 &     4.54 &     1.14 &     0.30 \\
-   PFSP &       rec13 &   1109.51 &    1043.08 &     4.55 &     1.07 &     0.31 \\
-   PFSP &       rec19 &   1090.06 &    1087.38 &    10.51 &     2.59 &     0.63 \\
-   PFSP &       rec31 &   1387.09 &    1367.26 &    27.84 &     6.90 &     1.79 \\
+    LOP &    N-p40-01 &   1195.90 &    1170.81 &     3.12 &  0.850000 &     0.21 \\
+    LOP &    N-p40-02 &   1207.24 &    1183.09 &     3.11 &  0.830000 &     0.20 \\
+    LOP &  N-t59b11xx &   1265.98 &    1240.36 &     3.78 &  0.950000 &     0.23 \\
+    LOP &  N-t59d11xx &   1269.93 &    1252.89 &     3.93 &  0.966667 &     0.25 \\
+    LOP &  N-sgb75.01 &   1994.22 &    2004.70 &    10.88 &  2.730000 &     0.72 \\
+    LOP &  N-sgb75.02 &   1995.09 &    1975.41 &    10.88 &  2.760000 &     0.69 \\
+   PFSP &       rec05 &   1099.63 &     917.81 &     0.80 &  0.200000 &     0.10 \\
+   PFSP &       rec13 &   1109.51 &    1043.08 &     0.85 &  0.200000 &     0.10 \\
+   PFSP &       rec19 &   1090.06 &    1087.38 &     1.76 &  0.470000 &     0.10 \\
+   PFSP &       rec31 &   1387.09 &    1367.26 &     4.77 &  1.280000 &     0.32 \\
 \bottomrule
@@
 \midrule
-    LOP &  N-t59b11xx &          102408.0 &           126204.0 &         114075.5 &         115941.5 &         128195.5 \\
-    LOP &  N-t59d11xx &           12692.2 &            30199.8 &          16662.9 &          22117.4 &          28050.4 \\
-    LOP &    N-p40-01 &           10496.6 &            15027.7 &          12369.1 &          13043.7 &          14459.3 \\
-    LOP &    N-p40-02 &           11714.7 &            15150.2 &          13160.5 &          13923.2 &          15027.7 \\
-    LOP &    N-p50-01 &           18197.7 &                NaN &          21161.8 &              NaN &              NaN \\
-    LOP &    N-p50-02 &           19090.5 &                NaN &          21746.6 &              NaN &              NaN \\
-    LOP &  N-sgb75.01 &         1893533.2 &          2095899.6 &        1992059.9 &        2011271.7 &        2074111.6 \\
-    LOP &  N-sgb75.02 &         1895801.8 &          2069286.9 &        1967132.3 &        2011004.2 &        2045556.5 \\
-   PFSP &       rec05 &            1276.2 &             1318.7 &           1276.3 &           1283.1 &           1300.4 \\
-   PFSP &       rec13 &            1992.9 &             2096.0 &           2057.8 &           2113.6 &           2134.9 \\
-   PFSP &       rec19 &            2264.5 &             2328.4 &           2304.8 &           2346.5 &           2392.4 \\
-   PFSP &       rec31 &            3415.5 &             3512.3 &           3419.3 &           3438.9 &           3511.9 \\
+    LOP &    N-p40-01 &    10496.6 &    15027.7 &    12453.7 &  1.336170e+04 &    14649.9 \\
+    LOP &    N-p40-02 &    11714.7 &    15150.2 &    13191.9 &  1.420900e+04 &    15206.2 \\
+    LOP &  N-t59b11xx &   102408.0 &   126204.0 &   114726.5 &  1.202020e+05 &   132039.5 \\
+    LOP &  N-t59d11xx &    12692.2 &    30199.8 &    16775.4 &  2.352089e+04 &    29884.1 \\
+    LOP &  N-sgb75.01 &  1893533.2 &  2095899.6 &  1983521.2 &  2.004006e+06 &  2046895.1 \\
+    LOP &  N-sgb75.02 &  1895801.8 &  2069286.9 &  1973083.9 &  2.001244e+06 &  2039356.1 \\
+   PFSP &       rec05 &     1276.2 &     1318.7 &     1282.1 &  1.287000e+03 &     1320.2 \\
+   PFSP &       rec13 &     1992.9 &     2096.0 &     2077.7 &  2.108500e+03 &     2131.6 \\
+   PFSP &       rec19 &     2264.5 &     2328.4 &     2318.5 &  2.352100e+03 &     2386.5 \\
+   PFSP &       rec31 &     3415.5 &     3512.3 &     3420.5 &  3.490900e+03 &     3474.3 \\
 \bottomrule
-\end{tabular}
-\end{table*}
+\end{tabular}\end{table*}
 
@@
 \section{Conclusions}\label{sec:conclusions}
-
 In this paper, we have introduced UMM, a population-based probabilistic
@@
 While certain functions are
-this discussion is usually ommited in the literature.
+this discussion is usually omitted in the literature.
 We discuss the different concepts and show, experimentally, the consequences of misunderstanding both.
@@
 are still significant advances yet to be made. Thus, we hope that this paper
-will motivate further research. % on such problems.
-%\MANUEL{I'm thinking about finishing here and commenting out the following paragraph}
+will motivate further research.
 
 In this paper, we focus on the UMM under the \ken distance.
-Therefore, we consider appropiate problems for it, such as the Linear Ordering Problem and the Permutation Flowshop Scheduling Problem, where the pairwise ordering among the items is relevant (as opposed to the actual exat position of each item in the permutation)
+Therefore, we consider appropriate problems for it, such as the Linear Ordering Problem and the Permutation Flow-shop Scheduling Problem, where the pairwise ordering among the items is relevant (as opposed to the actual exact position of each item in the permutation)
 
@@
 possible to extend UMM to other distance metrics, which will allow us to
-dynamically select among various distance metrics for an unknown black-box
-% \ekhine{}{estoy totalmente en contra de esto :) }
-permutation landscape~\citep{ZaeStoBar2014:ppsn}. Also, as shown by our
+dynamically select among various distance metrics for an unknown black-box permutation landscape.
+Also, as shown by our
 experiments, the behavior of the parameters ($r_1$ and $r_2$) of UMM is not
@@
 generally good static values or an online adaptation approach. Finally, we have
-used here three very different combinatorial problems (LOP, PFSP and QAP) as
+used here  different combinatorial problems (LOP, PFSP) as
 black-box benchmarks. However, an even more diverse range of problems would be

### 2021-02-04 paper/main.tex
figures

@@
 \newcommand{\supplement}{\href{http://doi.org/10.5281/zenodo.4500974}{doi:~10.5281/zenodo.4500974}}
-\newcommand{\ken}{Kendall's-$\tau$}
+
+\newcommand{\ken}{Kendall's-$\tau$\xspace}
+
 %% Revision documents
@@
 %Finally, we note that learning Budget-constrained Combinatorial is a hot topic for a variety of structures besides permutation~\cite{manchanda2020}.
-\MANUEL{- the authors assigned to black-box optimization problems characteristics that they do not own. For instance they say that "local search cannot be used".
- they mixed expensive to evaluate functions with black-box functions and of course they are different. In fact much of the metaheuristics do not mind
- if the problem is withe-box or black-box as they do not use static information from the function in the optimization process. In this case I understand
- that they have a black-box problem where the objective function is costly to evaluate. This should be clear and I would recommend to delete other comments
- related with white-box and black-box. MANUEL: FIXED
-}
 
@@
 population-based probabilistic algorithm specifically designed for optimizing
-black-box problems over a permutation landscape under a limited budget. There\EKHINE{En realidad está disenado para esto  pero podríamos probarlo para otros casos. Trabajo futuro}
+black-box problems over a permutation landscape under a limited budget. There
+%\EKHINE{En realidad está disenado para esto  pero podríamos probarlo para otros casos. Trabajo futuro}
 exists a wide body of research on probabilistic models for permutations, such
@@
 
-\manuel{9. Please use some newer references instead of [11] and, especially, [14]}{}
+%\MANUEL{9. Please use some newer references instead of [11] and, especially, [14]}
 
@@
 The set of all permutations of $n$ values is denoted as $S_n$ and has cardinality $n!$.
-We will use the one-line notation\MANUEL{what is one-line?} and denote permutations with Greek letters, $\sigma  = (\sigma(1), \ldots , \sigma(n))$.
-We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi = (\sigma(\pi(1)), \ldots, \sigma(\pi(n)))$\MANUEL{Is this correct?} and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.
-
-\manuel{P}{Throughout this paper and for consistency, p}ermutation $\sigma$ represents a \emph{ranking}:  $\sigma(i)=j$ denotes that item $i$ has rank $j$.
+We will use the one-line notation
+% \MANUEL{what is one-line?}
+and denote permutations with Greek letters, $\sigma  = (\sigma(1), \ldots , \sigma(n))$. We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi = (\sigma(\pi(1)), \ldots, \sigma(\pi(n)))$
+% \MANUEL{Is this correct?}
+and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.
+
+Throughout this paper and for consistency, permutation $\sigma$ represents a \emph{ranking}:  $\sigma(i)=j$ denotes that item $i$ has rank $j$.
 The \emph{ordering} associated with $\sigma$ is given by its inverse $\sigma^{-1}$, therefore,  $\sigma(i)=j \Leftrightarrow  \sigma^{-1}(j)=i$.
-This distinction is in order since statistical methodologies \manuel{are devoted for}{typically assume} rankings \manuel{and}{whereas the} optimization literature \manuel{for}{often assumes} orderings.
+This distinction is importante since statistical methodologies typically assume rankings whereas the optimization literature often assumes orderings.
 Mixing both concepts is a very common confusion with grave consequences, as we will illustrate in the experimental section.
 
-Kendall's-$\tau$ distance is the metric for rankings that counts the pairs of items ranked in different order in two permutations:
+\ken distance is the metric for rankings that counts the pairs of items ranked in different order in two permutations:
 %
@@
 
-An equivalent definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made to convert $\sigma_1^{-1}$ into $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
+An equivalent definition for the \ken distance counts the number of adjacent swaps that have to be made to convert $\sigma_1^{-1}$ into $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
 %\EKHINE{Mirandolo bien, lo que pasa es que dicen que es lo mismo la swap distance que la kendall!}
@@
 \subsection{Unbalanced Borda}\label{sec:uborda}
-We introduce now the core for our probabilistic algorithm, the uBorda algorithm.
+We introduce now the core of our probabilistic algorithm, the uBorda algorithm.
 Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking.
@@
 
-\subsection{Black-box Combinatorial Optimization under a limited budget}
+\subsection{Black-box combinatorial optimization under a limited budget}
 
@@
 controlled by parameter $0 < \rho < 1$. 
-\MANUEL{Don't we have to say something about $f(\sigma) > 1$?}
-\EKHINE{es decreciente igual}
+% \MANUEL{Don't we have to say something about $f(\sigma) > 1$?}
+% \EKHINE{es decreciente igual}
 In UMM, the value of $\rho$ depends on the set of true fitness evaluations $F$
@@
 \begin{minipage}{0.49\linewidth}
-  \texttt{reC19} (PFSP)\\[-0.5ex]
+  \texttt{reC05} (PFSP)\\[-0.5ex]
   \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec05_txt}
@@
 \begin{minipage}{0.49\linewidth}
-  \texttt{reC31} (PFSP)\\[-0.5ex]
+  \texttt{reC13} (PFSP)\\[-0.5ex]
   \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec13_txt}

### 2021-02-04 paper/main.tex
update GA

@@
 CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use a GA to optimize
-\ekhine{}{hay que actualizar essto??}
 the surrogate models with population size of 20, crossover rate of 0.5,
@@
 cycle crossover for permutations. To generate each new solution, CEGO runs the
-GA using the model as a surrogate of the true fitness function.  The authors of
-CEGO used a budget of $10^4$ surrogate-model evaluations for each run of the
-GA and we use the same confuguration parameters.% However, with such a setting, a single run of CEGO with a budget of
+GA with a budget of $10^4$ evaluations of the model as a surrogate of the true fitness function.  
+%and we use the same configuration parameters.% However, with such a setting, a single run of CEGO with a budget of
 %$\FEmax=400$ true fitness evaluations requires \emph{more than a week} of

### 2021-02-04 paper/main.tex
paper full
update conclusions

@@
 A purely explotation approach will set $\Edist = 0$  generating $\sigma_0$ in the sampling stage.
-However, the setting of a decreasing variance   controls the exploration-explotation tradeoff as a function of the number of function evaluations ($m$) that the algorithm performs, thus \manuel{it follows that}{} the behaviour is different for different choices of $m$.
+However, the setting of a decreasing variance   controls the exploration-explotation tradeoff as a function of the number of function evaluations ($m$) that the algorithm performs, thus the behaviour is different for different choices of $m$.
 Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.
@@
 
-\subsubsection{Computing environment}
+\myparagraph{Computing environment}
 % Due to the computational effort required, we used several computing systems to
@@
 
-\paragraph{Convergence analysis of CEGO}
+\myparagraph{Convergence analysis of CEGO}
 The red and the gold lines in the plots shown in Fig.~\ref{fig:results}
@@
   \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-02}
+\end{minipage}
+\\
+\begin{minipage}{0.49\linewidth}
+  \texttt{N-t59b11xx} (LOP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+  \texttt{N-sgb75-01} (LOP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
 \end{minipage}
@@
 \section{Conclusions}\label{sec:conclusions}
+
 In this paper, we have introduced UMM, a population-based probabilistic
 algorithm based on an unbalanced Mallows model. The algorithm is designed for
-black-box combinatorial problems on permutation landscapes and when the budget
-of fitness evaluations is severely limited (here, up to $100$, $200$ and $400$ evaluations).
-Our experiments presented here show that UMM is able to match the results of
-the state-of-the-art Bayesian optimizer for expensive black-box
-optimization if the computation time of the optimizers is not a concern.
+expensive black-box combinatorial problems on permutation landscapes when the
+budget of fitness evaluations is severely limited (here, up to $100$, $200$ and
+$400$ evaluations).  Our experiments presented here show that UMM is not far
+away from, and sometimes  is able to match, the results of the state-of-the-art
+Bayesian optimizer for expensive black-box optimization.
+%if the computation time of the optimizers is not a concern.
 
@@
 be no more than a few minutes added to each fitness evaluation. By comparison,
-CEGO may require up to one hour per fitness evaluation, with the time
+CEGO may require around one day to perform 400 fitness evaluations, with the time
 increasing non-linearly with permutation size and larger budgets.  Hence, UMM

### 2021-02-04 paper/main.tex
conclusions

@@
 algorithm based on an unbalanced Mallows model. The algorithm is designed for
-expensive black-box combinatorial problems on permutation landscapes when the
-budget of fitness evaluations is severely limited (here, up to $100$, $200$ and
-$400$ evaluations).  Our experiments presented here show that UMM is not far
-away from, and sometimes  is able to match, the results of the state-of-the-art
-Bayesian optimizer for expensive black-box optimization.
-%if the computation time of the optimizers is not a concern.
-
-In addition, UMM is specially well-suited for budget-limited scenarios that are
-still time-sensitive, e.g., when the overhead incurred by the optimizer should
-be no more than a few minutes added to each fitness evaluation. By comparison,
-CEGO may require around one day to perform 400 fitness evaluations, with the time
-increasing non-linearly with permutation size and larger budgets.  Hence, UMM
-is a computationally feasible alternative for relatively large problem sizes.
-
-In this paper, we discuss the importance of the representation of permutations in applied mathematics problems.
-In the context considered in this paper, permutations can bee seen as rankings or orderings of items.
-In other problems they could be matchings or cycles.
-While certain functions are
-this discussion is usually omitted in the literature.
-We discuss the different concepts and show, experimentally, the consequences of misunderstanding both.
-
+expensive black-box combinatorial problems on permutation landscapes when the budget
+of fitness evaluations is severely limited (here, up to $100$, $200$ and $400$ evaluations).
 Although we were inspired by previous work on ACO for expensive black-box
 combinatorial problems~\citep{PerLopStu2015si}, UMM is, to the best of our
-knowledge, the first EDA specifically designed for such problems that is able
-to match (and sometimes even improve) the state-of-the-art
-optimizer. Moreover, it is overwhelmingly faster than the alternatives for
-large problem sizes.  The results presented here show that, despite the
-intrinsic difficulty of the expensive black-box combinatorial scenario, there
-are still significant advances yet to be made. Thus, we hope that this paper
-will motivate further research.
-
-In this paper, we focus on the UMM under the \ken distance.
-Therefore, we consider appropriate problems for it, such as the Linear Ordering Problem and the Permutation Flow-shop Scheduling Problem, where the pairwise ordering among the items is relevant (as opposed to the actual exact position of each item in the permutation)
-
-However, we are certain it is
-possible to extend UMM to other distance metrics, which will allow us to
+knowledge, the first EDA specifically designed for such problems.
+
+We also discuss the importance of the representation of permutations in applied mathematics problems.
+Permutations can bee seen as rankings or orderings of items and this difference is often neglected.
+Based on this differentiation, we propose an improvement for CEGO over the original version.
+The results of UMM clearly improve over the original CEGO and they are competitive with the improved version of CEGO proposed here.
+
+In addition, UMM is specially well-suited for budget-limited scenarios that are still time-sensitive, e.g., when the overhead incurred by the optimizer should
+be no more than a few second per fitness evaluation (no more than one hour for $400$ evaluations), due to its quasi-linear time complexity.
+By comparison, CEGO may require several minutes per fitness evaluation (around 22 hours for $400$ evaluations), with the time increasing non-linearly with permutation size and larger budgets.
+Hence, UMM is a computationally feasible alternative for relatively large problem sizes.
+
+The current version of UMM uses the \ken distance and, therefore, it is appropriate for problems where the pairwise ordering among the items is relevant (as opposed to the actual exact position of each item in the permutation), such as the Linear Ordering Problem and the Permutation Flow-shop Scheduling.
+However, we are certain that it is
+possible to extend UMM to other distance metrics such as Hamming, which will allow us to
  dynamically select among various distance metrics for an unknown black-box permutation landscape.
-Also, as shown by our
-experiments, the behavior of the parameters ($r_1$ and $r_2$) of UMM is not
-always obvious, thus a more detailed analysis would be needed to provide either
-generally good static values or an online adaptation approach. Finally, we have
-used here  different combinatorial problems (LOP, PFSP) as
-black-box benchmarks. However, an even more diverse range of problems would be
+%Also, as shown by our
+%experiments, the behavior of the parameters ($r_1$ and $r_2$) of UMM is not always obvious, thus
+In addition, a more detailed analysis would be needed of the parameters ($r_1$ and $r_2$) of UMM to provide either
+generally good static values or an online adaptation approach.
+Finally, an even more diverse range of problems would be
 needed to understand the behavior of UMM on real-world black-box combinatorial
 landscapes.
+
+The results presented here show that, despite the intrinsic difficulty of the expensive black-box combinatorial scenario, there are still significant advances yet to be made.
+Thus, we hope that this paper will motivate further research.
 

### 2021-02-04 paper/main.tex
fixed table

@@
 A natural question is what happens if we did invert the ranking.
-We have carried out all the experimentation with this modification of CEGO and we have denote this version CEGO\_inv.
-The original version of CEGO is denoted CEGO\_orig.
+We have carried out all the experimentation with this modification of CEGO and we have denote this version \CEGOinv.
+The original version of CEGO is denoted \CEGOorig.
 %
@@
 \begin{itemize}
-\item UMM\_100, UMM\_200 and UMM\_400 are three different versions for UMM, each considering a different total number of function evaluations, 100, 200 and 400.
-Obviously, the lines corresponding to UMM\_100, UMM\_200 wont reach the right corner of the figure.
+\item UMM$_{100}$, UMM$_{200}$ and UMM$_{400}$ are three different versions for UMM, each considering a different total number of function evaluations, 100, 200 and 400.
+Obviously, the lines corresponding to UMM$_{100}$, UMM$_{200}$ wont reach the right corner of the figure.
 As mentioned in Sec.~\ref{sec:sampling}, the behavior of
 UMM depends on the value of $m$. 
-\item CEGO\_orig and CEGO\_inv are, respectively, the original version of CEGO and the modification of Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$.
+\item \CEGOorig and \CEGOinv are, respectively, the original version of CEGO and the modification of Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$.
 \end{itemize}
@@
 %\begin{itemize}
-%\item UMM\_100, UMM\_200 and UMM\_400 are three different versions for UMM, each considering a different total number of function evaluations, 100, 200 and 400.
-Obviously, the lines corresponding to UMM\_100, UMM\_200 wont reach the right corner of the figure.
-%\item CEGO\_orig and CEGO\_inv are, respectively, the original version of CEGO and the modification of Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$.
+%\item UMM$_{100}$, UMM$_{200}$ and UMM$_{400}$ are three different versions for UMM, each considering a different total number of function evaluations, 100, 200 and 400.
+Obviously, the lines corresponding to UMM$_{100}$, UMM$_{200}$ wont reach the right corner of the figure.
+%\item \CEGOorig and \CEGOinv are, respectively, the original version of CEGO and the modification of Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$.
 %\end{itemize}
@@
 %The performance of the UMM is clearly better than the performance of the original CEGO for the same problem.
-For CEGO\_inv on the other hand improves the results of the UMM with the current experimental setting.
-However, UMM seems to be improving the quality of the solutions until the last iteration while CEGO\_inv seems to be stucked at a local optimum.
+For \CEGOinv on the other hand improves the results of the UMM with the current experimental setting.
+However, UMM seems to be improving the quality of the solutions until the last iteration while \CEGOinv seems to be stucked at a local optimum.
 Further analysis lines will consider a larger budget of fitness evaluation.
@@
 \begin{table*}[tb]%time
+  \caption{Mean fitness (and standard deviation), over 10 independent runs for each problem instance, of the best solution found at the end of each run of the original CEGO (\CEGOorig) and its improved version (\CEGOinv) and three configurations of UMM that different in the budget of evaluations $m \in \{100, 200, 400\}$.
+The runs of CEGO used a budget of $m=400$.\label{tab:results_fitness}}
 \begin{tabular}{llrrrrr}
 \toprule
-Problem &    Instance &  CEGO\_inv &  CEGO\_orig &  UMM\_400 &   UMM\_200 &  UMM\_100 \\
+\bf Problem & \bf   Instance &  \CEGOinv &  \CEGOorig &  UMM$_{400}$ &   UMM$_{200}$ &  UMM$_{100}$ \\
 \midrule
@@
 \toprule
-Problem &    Instance &   CEGO\_inv &  CEGO\_orig &    UMM\_400 &       UMM\_200 &    UMM\_100 \\
+Problem &    Instance &   \CEGOinv &  \CEGOorig &    UMM$_{400}$ &       UMM$_{200}$ &    UMM$_{100}$ \\
 \midrule

### 2021-02-04 paper/main.tex
fixed table

@@
 \begin{table*}[tb]%time
-  \caption{Mean fitness (and standard deviation), over 10 independent runs for each problem instance, of the best solution found at the end of each run of the original CEGO (\CEGOorig) and its improved version (\CEGOinv) and three configurations of UMM that different in the budget of evaluations $m \in \{100, 200, 400\}$.
-The runs of CEGO used a budget of $m=400$.\label{tab:results_fitness}}
+  \caption{Mean runtime  in minutes, over 10 independent runs for each problem instance, of the original CEGO (\CEGOorig) and its improved version (\CEGOinv) and three configurations of UMM that differ in the budget of evaluations $m \in \{100, 200, 400\}$.
+The runs of CEGO used a budget of $m=400$.\label{tab:results_time}}
 \begin{tabular}{llrrrrr}
@@
 \begin{table*}[tb]%fitness
+    \caption{Mean fitness, over 10 independent runs for each problem instance, of the best solution found at the end of each run of the original CEGO (\CEGOorig) and its improved version (\CEGOinv) and three configurations of UMM that differ in the budget of evaluations $m \in \{100, 200, 400\}$.
+The runs of CEGO used a budget of $m=400$.\label{tab:results_fitness}}
 \begin{tabular}{llrrrrr}

### 2021-02-04 paper/main.tex
paper full

--- 2021-02-04 paper/main.tex
@@
 Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually builds upon orderings.
-For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.\MANUEL{necesitamos las ecuaciones?
-Swap de una permutación es simplemente la distancia de kendall de la permutación inversa.
-.}
-\EKHINE{asi mas claro, no se van a perder con esto}
+For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.
 %The concepts of rankings and orderings are usually confused of ignored.
@@
 During our experiments, we realized that CEGO claimed to use the Swap distance. Instead, CEGO implements the \ken distance, yet CEGO did
-not invert the permutation before evaluating the PFSP.\MANUEL{no me convence esta forma de expresarlo porque la crítica obvia sería que por qué invertimos la función objectivo y no la función de distancia?} \EKHINE{es lo mismo}We verified that this is
+not invert the permutation before evaluating the PFSP.
+%\MANUEL{no me convence esta forma de expresarlo porque la crítica obvia sería que por qué invertimos la función objectivo y no la función de distancia?} \EKHINE{es lo mismo}
+We verified that this is
 indeed how the implementation works and that those results match the published
-ones~\citep{ZaeStoBar2014:ppsn}. Of course, the obvious question arises about the effect of inverting the permutation before calling the objective
+ones~\citep{ZaeStoBar2014:ppsn}. 
+Note that using the Swap distance as claimed by CEGO is equivalent to using \ken distance and inverting the permutation before calling the objective
+function. Thus, the obvious question arises about the effect of inverting the permutation before calling the objective
 function, so that the rankings are transformed to ordering, as done by our UMM
-algorithm. Therefore, in the following, we show results of both variants, that is, the original CEGO (\CEGOorig) and our modified version that inverts the permutation before evaluation (\CEGOinv).
+algorithm. In the following, we show results of both variants, that is, the original CEGO (\CEGOorig) and our modified version that inverts the permutation before evaluation (\CEGOinv).
 
@@
 solution found over the 10 runs of each algorithm, together with a 95\%
-confidence interval (CI) around their mean fitness difference. Except for PFSP
+confidence interval (CI) around their mean fitness difference. 
+The 
+Except for PFSP
 instances \texttt{rec05} and \texttt{rec31}, the differences are statistically
@@
 Moreover, UMM scales good for larger instances in quality performance. 
-The computational complexity of UMM is $O(n \log n)$ at each iteration while both versions of CEGO need \ekhine{MANUEL, tu sabes}{?}.
-This difference is evident in the time performance of the algorithms.
-The details can be seen in Table~\ref{table:time}.
+The computational complexity of UMM is $O(n \log n)$ at each iteration which causes the abrupt differencein the time performance of the algorithms.
+The details can be seen in Table~\ref{tab:results_time}.
 
@@
 \begin{minipage}{0.49\linewidth}
-  \texttt{N-p40-04} (LOP)\\[-0.5ex]
+  \texttt{N-p40-02} (LOP)\\[-0.5ex]
   \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-02}
@@
 \begin{minipage}{0.49\linewidth}
-  \texttt{N-sgb75-01} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
+  \texttt{N-t59d11xx} (LOP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59d11xx}
 \end{minipage}
@@
 \begin{minipage}{0.49\linewidth}
-  \texttt{N-t59b11xx} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
+  \texttt{SGB\_N-sgb75\_01} (LOP)\\[-0.5ex]
+  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_01}
 \end{minipage}
 \begin{minipage}{0.49\linewidth}
-  \texttt{N-sgb75-01} (LOP)\\[-0.5ex]
+  \texttt{SGB\_N-sgb75\_01} (LOP)\\[-0.5ex]
   \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}

### 2021-02-04 paper/main.tex
paper full

@@
 confidence interval (CI) around their mean fitness difference. 
-The 
 Except for PFSP
@@
 consistent, are relatively small.
+The detailed results of UMM$_{100}$  and UMM$_{200}$  and the original \CEGOorig are given in Table~\ref{tab:results_fit}.
+In this table, it can be seen that the quality of \CEGOorig  solutions is similar to the UMM$_{100}$ version of UMM.
 

### 2021-02-04 paper/main.tex

--- 2021-02-04 paper/main.tex
@@
 solution found over the 10 runs of each algorithm, together with a 95\%
-confidence interval (CI) around their mean fitness difference. Except for PFSP
+confidence interval (CI) around their mean fitness difference. 
+Except for PFSP
 instances \texttt{rec05} and \texttt{rec31}, the differences are statistically

### 2021-02-04 paper/main.tex
fixed abstract

@@
   time or resources, only a limited number of evaluations is possible,
-  typically several orders of magnitude smaller than in non-expensive problems. In
-  this scenario, classical optimization methods such as mixed-integer
-  programming and local search are not useful.  In the continuous case,
-  Bayesian optimization, in particular using Gaussian processes, has proven
-  very effective under these conditions. Much less research is available in the
-  combinatorial case. In this paper, we propose and analyze UMM, an
+  typically several orders of magnitude smaller than in non-expensive
+  problems. In this scenario, classical optimization methods such as
+  mixed-integer programming and local search are not useful.  In the continuous
+  case, Bayesian optimization, in particular using Gaussian processes, has
+  proven very effective under these conditions. Much less research is available
+  in the combinatorial case. In this paper, we propose and analyze UMM, an
   estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
   model and unbalanced rank aggregation (uBorda). Experimental results on
-  black-box versions of LOP and PFSP show that UMM is able to match, and
-  sometimes surpass, the solutions obtained by CEGO, a Bayesian optimization
-  algorithm for combinatorial
-  optimization. %, specially when the budget of solutions evaluated approaches 400.
-  Moreover, the computational complexity of UMM increases linearly with both
-  the number of function evaluations and the permutation size.
-%  an order of magnitude smaller than CEGO, which makes it specially useful when
-  %both computation time and number of evaluations are of concern.
-  \sloppy
-  %\textcolor{red}{16 pages max, deadline: 1 November 2020}
+  black-box versions of LOP and PFSP show that UMM outperforms the solutions
+  obtained by CEGO, a Bayesian optimization algorithm for combinatorial
+  optimization. Nevertheless, a small change to CEGO proposed here,
+  significantly improves its performance, thus producing solutions that are
+  slightly better than those of UMM.
+  % , specially when the budget of solutions evaluated approaches 400.
+  As another benefit, the computational complexity of UMM increases linearly
+  with both the number of function evaluations and the permutation size, which
+  results in computation times an order of magnitude shorter than CEGO, which
+  makes it specially useful when both computation time and number of
+  evaluations are limited.  \sloppy
 \end{abstract}

### 2021-02-04 paper/main.tex
paper full

--- 2021-02-04 paper/main.tex
@@
 consistent, are relatively small.
-The detailed results of UMM$_{100}$  and UMM$_{200}$  and the original \CEGOorig are given in Table~\ref{tab:results_fit}.
+The detailed results of UMM$_{100}$  and UMM$_{200}$  and the original \CEGOorig are given in Table~\ref{tab:results_fitness}.
 In this table, it can be seen that the quality of \CEGOorig  solutions is similar to the UMM$_{100}$ version of UMM.
@@
 
-Table~\ref{tab:results} shows a brief comparison of time and the full ones are given in Table~\ref{tab:time}. The difference is time in impressive. 
+Table~\ref{tab:results} shows a brief comparison of time and the full ones are given in Table~\ref{tab:results_time}. The difference is time in impressive. 
 The average runtime for all the instances is around 0.3 minutes, 1.15 minutes and 4.5 minutes respectively for the  versions with 100, 200 and 400 fitness evaluations. 
@@
 Moreover, UMM scales good for larger instances in quality performance. 
-The computational complexity of UMM is $O(n \log n)$ at each iteration which causes the abrupt differencein the time performance of the algorithms.
+The computational complexity of UMM is $O(n \log n)$ at each iteration which causes the abrupt difference in the time performance of the algorithms.
 The details can be seen in Table~\ref{tab:results_time}.

### 2021-02-04 paper/main.tex
fixed abstract

@@
   estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
-  model and unbalanced rank aggregation (uBorda). Experimental results on
-  black-box versions of LOP and PFSP show that UMM outperforms the solutions
-  obtained by CEGO, a Bayesian optimization algorithm for combinatorial
-  optimization. Nevertheless, a small change to CEGO proposed here,
+  model and unbalanced rank aggregation (uBorda).
+
+  Experimental results on black-box versions of LOP and PFSP show that UMM
+  outperforms the solutions obtained by CEGO, a Bayesian optimization algorithm
+  for combinatorial optimization. Nevertheless, a slight modification to CEGO,
+  based on the different interpretations for rankings and orderings,
   significantly improves its performance, thus producing solutions that are
-  slightly better than those of UMM.
-  % , specially when the budget of solutions evaluated approaches 400.
-  As another benefit, the computational complexity of UMM increases linearly
-  with both the number of function evaluations and the permutation size, which
-  results in computation times an order of magnitude shorter than CEGO, which
-  makes it specially useful when both computation time and number of
-  evaluations are limited.  \sloppy
+  slightly better than those of UMM and dramatically better than the original
+  version.  As another benefit, the computational complexity of UMM increases
+  linearly with both the number of function evaluations and the permutation
+  size, which results in computation times an order of magnitude shorter than
+  CEGO, which makes it specially useful when both computation time and number
+  of evaluations are limited.  \sloppy
 \end{abstract}

### 2021-02-04 paper/main.tex
fixed emptyset

@@
     % \\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
-    \State $S \assign \{\}$
+    \State $S \assign \emptyset$
     \For{$i \assign 0$ \textbf{to} $\minit$}
@@
 %
-UMM uses a budget of evaluations ($\FEmax$) of  $\minit=10$ solutions sampled uniformly.
+UMM initially samples $\minit=10$ solutions uniformly.
 The other two parameters of UMM are $r_1$ and $r_2$ that determine the learning rate $\rho$ in uBorda, and consequently, determine the
@@
 
-We consider 3 different values for the maximum budget of $\FEmax=100, 200,400$ evaluations of the
+We consider three different values $\FEmax \in\{100, 200,400\}$ for the maximum budget of  evaluations of the
 actual objective function. In a white-box context, state-of-the-art algorithms

### 2021-02-04 paper/main.tex
minor edit

@@
 
+\myparagraph{CEGO}
+%
 Combinatorial Efficient Global Optimization (CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
@@
 EGO, with a distance measure appropriate to combinatorial
-landscapes~\citep{ZaeStoBar2014:ppsn}, such as Kendall's-$\tau$ distance for
-permutations (Eq.~\ref{eq:kendall_decomp}). In CEGO, the surrogate model is
+landscapes~\citep{ZaeStoBar2014:ppsn}.
+% , such as Kendall's-$\tau$ distance for permutations
+% (Eq.~\ref{eq:kendall_decomp}).
+In CEGO, the surrogate model is
 explored by a GA with crossover and mutation operators appropriate for the

### 2021-02-05 paper/main.tex
final updates

@@
 \usepackage{afterpage,tabularx}
-\hypersetup{draft}
+%\hypersetup{draft}
 %\usepackage[colorlinks=true, allcolors=blue]{hyperref}
@@
 
-We consider three different values $\FEmax \in\{100, 200,400\}$ for the maximum budget of  evaluations of the
-actual objective function. In a white-box context, state-of-the-art algorithms
-for the LOP and PFSP typically evaluate thousands of solutions,
-thus, the budget considered here for the black-box context is extremely
-limited.
-
 \myparagraph{CEGO}
@@
 % (Eq.~\ref{eq:kendall_decomp}).
+
 In CEGO, the surrogate model is
 explored by a GA with crossover and mutation operators appropriate for the
-particular combinatorial problem. The original paper notes that coupling the GA
-with local search does not improve the results significantly since the model is
-anyway an inexact estimation of the original fitness
-function~\citep[p.~875]{ZaeStoFriFisNauBar2014}. A thorough comparison of CEGO
+particular combinatorial problem. A thorough comparison of CEGO
 with other optimizers for expensive black-box combinatorial optimization ranked
@@
 
-We use the original implementation of GECO.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
+We use here the original implementation of
+GECO.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
 Although it is never stated in the original paper, the implementation of CEGO
@@
 Following the authors of
-CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use a GA to optimize
-the surrogate models with population size of 20, crossover rate of 0.5,
+CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, the surrogate model is optimized by a GA with  population size of 20, crossover rate of 0.5,
 mutation rate of $1/n$, tournament selection of size 2 with probability of 0.9,
@@
 cycle crossover for permutations. To generate each new solution, CEGO runs the
-GA with a budget of $10^4$ evaluations of the model as a surrogate of the true fitness function.  
+GA with a budget of $10^4$ evaluations of the model as a surrogate of the true
+fitness function.  The original paper~\citep[p.~875]{ZaeStoFriFisNauBar2014} notes that coupling the GA with local
+search does not improve the results significantly since the model is anyway an
+inexact estimation of the original fitness
+function.
+
 %and we use the same configuration parameters.% However, with such a setting, a single run of CEGO with a budget of
@@
 %to understand CEGO parameters.
-
-\myparagraph{Other settings}
-%
-For simplicity, we use Kendall's-$\tau$
-distance~(Eq.~\ref{eq:kendall_decomp}) in all experiments.
-%Nevertheless,
-%\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
-%when using Hamming distance and the same may apply to UMM.
-We plan to extend UMM to other distance measures in the future.
 
@@
 % later, we also consider QAP instances \texttt{kra30a} and \texttt{kra30b}.
+
+\myparagraph{Other settings}
+%
+We consider three different values $\FEmax \in\{100, 200,$ $400\}$ for the maximum
+budget of evaluations of the actual objective function in UMM. In the case of
+CEGO, we only run experiments with the largest budget $m=400$.  In a white-box
+context, state-of-the-art algorithms for the LOP and PFSP typically evaluate
+thousands of solutions, thus, the budget considered here for the black-box
+context is extremely limited.
+%
+For simplicity, we use Kendall's-$\tau$
+distance~(Eq.~\ref{eq:kendall_decomp}) in all experiments.
+%Nevertheless,
+%\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
+%when using Hamming distance and the same may apply to UMM.
+We plan to extend UMM to other distance measures in the future.
 
@@
 From this preliminary experiment, it is not obvious how the best settings are
-related to problem features, such as permutation length ($n$). On the other
+related to problem features, such as permutation length $n$. On the other
 hand, given the sharp transitions in the plots, it is clear that further
@@
 
-\subsection{Rankings and orderings: experiments}\label{sec:exper_ro}
+\subsection{Experiments with rankings and orderings}\label{sec:exper_ro}
 
@@
 
-%During our experiments, we realized that the \emph{swap} distance used by CEGO
-%assumes rankings, i.e., it actually measures the \ken distance, yet CEGO did
 % not invert the permutation before evaluating the PFSP.
-During our experiments, we realized that CEGO claimed to use the Swap distance. Instead, CEGO implements the \ken distance, yet CEGO did
+% During our experiments, we realized that CEGO claimed to use the Swap distance. Instead, CEGO implements the \ken distance, yet CEGO did
+During our experiments, we realized that distance metric called Swap in CEGO~\citep{ZaeStoBar2014:ppsn} actually assumes rankings, that is, it actually implements the \ken distance, yet CEGO does
 not invert the permutation before evaluating the PFSP.
@@
 ones~\citep{ZaeStoBar2014:ppsn}. 
-Note that using the Swap distance as claimed by CEGO is equivalent to using \ken distance and inverting the permutation before calling the objective
-function. Thus, the obvious question arises about the effect of inverting the permutation before calling the objective
-function, so that the rankings are transformed to ordering, as done by our UMM
+%Note that using the Swap distance as claimed by CEGO is equivalent to using \ken distance and inverting the permutation before calling the objective
+%function.
+Thus, the obvious question arises about the effect of inverting the permutation before calling the objective
+function, so that the rankings are transformed to orderings, as done by our UMM
 algorithm. In the following, we show results of both variants, that is, the original CEGO (\CEGOorig) and our modified version that inverts the permutation before evaluation (\CEGOinv).
@@
 
-\subsection{Experimental Analysis} % on LOP and PFSP instances}
-
-%In the following, we compare 5 algorithm configurations.
-% Three configurations
-%of UMM (UMM$_{100}$, UMM$_{200}$, and UMM$_{400}$) that differ in their budget
-%of evaluations ($m$). As mentioned in Sec.~\ref{sec:sampling}, the behavior of
-%UMM depends on the value of $m$. The other two configurations correspond 
-%to \CEGOorig and \CEGOinv, as defined above.
-
-Figure~\ref{fig:results} shows the results on the real-world instances from LOLIB and PFSP.
-There is one figure per instance.
-In each figure there is one horizontal line displaying the best known solutions for the problem.
-Besides, there are 5 different lines representing each of the different algorithm/versions considered, in particular:
-\begin{itemize}
-\item UMM$_{100}$, UMM$_{200}$ and UMM$_{400}$ are three different versions for UMM, each considering a different total number of function evaluations, 100, 200 and 400.
-Obviously, the lines corresponding to UMM$_{100}$, UMM$_{200}$ wont reach the right corner of the figure.
-As mentioned in Sec.~\ref{sec:sampling}, the behavior of
-UMM depends on the value of $m$. 
-\item \CEGOorig and \CEGOinv are, respectively, the original version of CEGO and the modification of Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$.
-\end{itemize}
-
-Since there is a small number of evaluations and the algorithms evaluate just
-one solution at each step, we chose to record and plot the fitness of the
-solution evaluated at each step of each run, instead of the usual plots of the
-best-so-far solution over number of evaluations. We believe that plotting each
-evaluated solution provides better insights on the search dynamics of the
-algorithms analyzed here, despite it is not directly comparable with most literature results. Please note that the algorithms do not return the
-final solution shown in the plots, but rather the best one (minimum fitness) of
-each run. Nevertheless, the minimum values achieved in the plot up to a
-particular number of function evaluations gives an estimation of the mean
-fitness of the best-so-far solution up to that point in the run. We provide
-later an analysis of the best solution found by each run of the algorithm, which are comparable with the original paper of CEGO.
-
-Figure~\ref{fig:results} shows such plots for a few LOP instances from LOLIB
-and the PFSP.  For each problem instance and each algorithm, we plot the mean
-fitness of the solution evaluated at each step over 10 runs with different
-random seeds. The horizontal line indicates the best-known fitness for that instance.
-% 
-The shaded area shows one standard deviation around the mean. Ideally, each new
-evaluation will monotonically improve in fitness, since each solution evaluated
-helps to refine the model and leads to a better solution being evaluated in the
-next step.  However, it is possible that new solutions are worse than their
-predecessors. %Nevertheless, in this example, there is a clear
-% overall improvement as more evaluations are performed.
-
-\myparagraph{Convergence analysis of UMM} We start our analysis with the three
-configurations of UMM. We can see a common trend in all instances: After
-the 10 random initial permutations, the fitness of new solutions decreases
-with a linear trend along iterations. The (close to) linear decrease is given
-by the linear decrease in the expected distance of the model, as described in Section~\ref{sec:sampling}. Overall, the rate of improvement
-does not seem to converge, not even for the longer runs of 400 evaluations.
-Therefore, we conjecture that longer runs are likely to produce further improved solutions.
-Moreover, a non-linear decrease in the expected distance should result in a faster convergence for larger evaluation budgets.
-
-\myparagraph{Convergence analysis of CEGO}
-The red and the gold lines in the plots shown in Fig.~\ref{fig:results}
-correspond to the two variants of CEGO the original \CEGOorig and our modified
-version \CEGOinv. The different between these two variants is overwhelmingly in
-favour of \CEGOinv. With the exception of the \texttt{reC19} PFSP instance, the
-search behavior of \CEGOorig resembles a random search, after a brief
-improvement phase at the start of the run. By contrast, \CEGOinv shows a very
-fast improvement in fitness within the first 50 evaluations that tends to
-converge around the 150$^\text{th}$ evaluation, showing little improvement
-after that. Since the only difference between these two variants is the
-transformation from rankings to orderings before evaluating the objective
-function (Sec.~\ref{sec:exper_ro}), we must conclude that this difference has
-tremendous effect in performance. Therefore, we will focus our analysis on
-\CEGOinv.
-
-\myparagraph{Comparison of UMM and CEGO}
-When comparing the UMM and CEGO configurations shown in Fig.~\ref{fig:results},
-we can observe that all UMM configurations perform much better than the
-original CEGO (\CEGOorig), while they converge slower than our modified
-\CEGOinv. However, the rate of improvement of the UMM variants does not seem to
-slow down near the maximum budget given and, in several instances such as
-\texttt{rec31} and \texttt{N-sgb75-01}, matches and surpasses the solutions
-found by \CEGOinv. Interestingly, this is the case, even for the shorter budget
-configurations of UMM, i.e., UMM$_{100}$ and UMM$_{200}$. We also observe a
-pattern that UMM performs better, or \CEGOinv performs worse, on larger
-instances, as can be seen by comparing \texttt{rec19} with \texttt{rec31}.
-
-\myparagraph{Comparison of the best solutions found}
-%
-We now turn our analysis to the best (minimum) fitness found at the end of the
-run by the two best algorithms, that is, \CEGOinv and UMM$_{400}$. We show in
-Table~\ref{tab:results} the mean fitness (and standard deviation) of the best
-solution found over the 10 runs of each algorithm, together with a 95\%
-confidence interval (CI) around their mean fitness difference. 
-Except for PFSP
-instances \texttt{rec05} and \texttt{rec31}, the differences are statistically
-significant in favour of \CEGOinv (the intervals do not contain the value
-zero). Nevertheless, the CIs indicate that the differences, although
-consistent, are relatively small.
-The detailed results of UMM$_{100}$  and UMM$_{200}$  and the original \CEGOorig are given in Table~\ref{tab:results_fitness}.
-In this table, it can be seen that the quality of \CEGOorig  solutions is similar to the UMM$_{100}$ version of UMM.
-
-\myparagraph{Runtime comparison}
-%
-The average
-running time is approximately 22 hours for the 400 function evaluations, which
-is 300 times slower than UMM for the same number of function evaluations.
-
-Table~\ref{tab:results} shows a brief comparison of time and the full ones are given in Table~\ref{tab:results_time}. The difference is time in impressive. 
-The average runtime for all the instances is around 0.3 minutes, 1.15 minutes and 4.5 minutes respectively for the  versions with 100, 200 and 400 fitness evaluations. 
-For CEGO in both versions the average time for 400 evaluations is more than 22 hour. This is 300 times slower than UMM for the same number of function evaluations. 
-Moreover, UMM scales good for larger instances in quality performance. 
-The computational complexity of UMM is $O(n \log n)$ at each iteration which causes the abrupt difference in the time performance of the algorithms.
-The details can be seen in Table~\ref{tab:results_time}.
-
-\begin{figure*}[tb]
+\afterpage{
+\begin{figure*}
 \small
@@
 \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:results}}
+\vspace{10em}
 \end{figure*}
+}
+
+\subsection{Experimental Analysis} % on LOP and PFSP instances}
+
+%In the following, we compare 5 algorithm configurations.
+% Three configurations
+%of UMM (UMM$_{100}$, UMM$_{200}$, and UMM$_{400}$) that differ in their budget
+%of evaluations ($m$). As mentioned in Sec.~\ref{sec:sampling}, the behavior of
+%UMM depends on the value of $m$. The other two configurations correspond 
+%to \CEGOorig and \CEGOinv, as defined above.
+
+Since the number of evaluations is  small and the algorithms evaluate just
+one solution at each step, we chose to record and plot the fitness of the
+solution evaluated at each step of each run, instead of the usual plots of the
+best-so-far solution over number of evaluations. We believe that plotting each
+evaluated solution provides better insights on the search dynamics of the
+algorithms analyzed here. %, despite it is not directly comparable with most literature results.
+Please note that the algorithms do not return the
+final solution shown in the plots, but rather the best one (minimum fitness) of
+each run. Nevertheless, the minimum values achieved in the plot up to a
+particular number of function evaluations gives an estimation of the mean
+fitness of the best-so-far solution up to that point in the run. We provide
+later an analysis of the best solution found by each run of the algorithms.
+%, which are comparable with the original paper of CEGO.
+
+Figure~\ref{fig:results} shows the results on the real-world instances from
+LOLIB and PFSP.  There is one plot per instance.  The horizontal line indicates
+the best-known fitness for that instance.  Besides, there are 5 different lines
+representing each of the different algorithm variants tested, in particular:
+\begin{itemize}
+\item UMM$_{100}$, UMM$_{200}$ and UMM$_{400}$ are three different versions of UMM, each considering a different total budget of function evaluations.
+As can be expected, the lines corresponding to UMM$_{100}$ and UMM$_{200}$ do not reach the right corner of the figure.
+As mentioned in Sec.~\ref{sec:sampling}, the behavior of
+UMM depends on the value of $m$. 
+\item \CEGOorig and \CEGOinv are, respectively, the original version of CEGO and the modification discussed in Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$.
+\end{itemize}
+For each algorithm, we plot the mean
+fitness of the solution evaluated at each step over 10 runs with different
+random seeds.
+%Figure~\ref{fig:results} shows such plots for a few LOP instances from LOLIB
+%and the PFSP.
+% 
+The shaded area shows one standard deviation around the mean. Ideally, each new
+evaluation will monotonically improve in fitness, since each solution evaluated
+helps to refine the model and leads to a better solution being evaluated in the
+next step.  However, it is possible that new solutions are worse than their
+predecessors. %Nevertheless, in this example, there is a clear
+% overall improvement as more evaluations are performed.
+
+\myparagraph{Convergence analysis of UMM} We start our analysis with the three
+variants of UMM. We can see a common trend in all instances: After
+the 10 random initial permutations, the fitness of new solutions decreases
+with a linear trend along iterations. The (close to) linear decrease is given
+by the linear decrease in the expected distance of the model, as described in Section~\ref{sec:sampling}. Overall, the rate of improvement
+does not seem to converge, not even for the longer runs of 400 evaluations.
+Therefore, we conjecture that longer runs are likely to produce further improved solutions.
+Moreover, a non-linear decrease in the expected distance should result in a faster convergence for larger evaluation budgets.
+
+\myparagraph{Convergence analysis of CEGO}
+The red and the gold lines in the plots shown in Fig.~\ref{fig:results}
+correspond to the two variants of CEGO the original \CEGOorig and our modified
+version \CEGOinv. The difference between these two variants is overwhelmingly in
+favour of \CEGOinv. With the exception of the \texttt{reC19} PFSP instance, the
+search behavior of \CEGOorig resembles a random search, after a brief
+improvement phase at the start of the run. By contrast, \CEGOinv shows a very
+fast improvement in fitness within the first 50 evaluations and tends to
+converge around the 150$^\text{th}$ evaluation, showing little improvement
+after that. Since the only difference between these two variants is the
+transformation from rankings to orderings before evaluating the objective
+function (Sec.~\ref{sec:exper_ro}), we must conclude that this difference has
+tremendous effect in performance.
+%Therefore, we will focus our analysis on
+%\CEGOinv.
+
+\myparagraph{Comparison of UMM and CEGO}
+When comparing the UMM and CEGO variants shown in Fig.~\ref{fig:results},
+we can observe that all UMM variants perform much better than the
+original CEGO (\CEGOorig), while they converge slower than our modified
+\CEGOinv. However, the rate of improvement of the UMM variants does not seem to
+slow down near the maximum budget given and, in several instances such as
+\texttt{rec31} and \texttt{N-sgb75-01}, matches and surpasses the solutions
+found by \CEGOinv. Interestingly, this is the case, even for the shorter budget
+variants of UMM, i.e., UMM$_{100}$ and UMM$_{200}$. We also observe a
+pattern that UMM performs better, or \CEGOinv performs worse, on larger
+instances, as can be seen by comparing \texttt{rec19} with \texttt{rec31}.
+
+\myparagraph{Comparison of the best solutions found}
+%
+We now turn our analysis to the best (minimum) fitness found at the end of the
+run by the two best algorithms, that is, \CEGOinv and UMM$_{400}$. We show, in
+Table~\ref{tab:results}, the mean fitness (and standard deviation), over the 10
+runs of each algorithm, of the best solution found at the end of the run,
+together with a 95\% confidence interval (CI) around their mean fitness
+difference.  Except for PFSP instances \texttt{rec05} and \texttt{rec31}, the
+differences are statistically significant in favour of \CEGOinv (the intervals
+do not contain the value zero). Nevertheless, the CIs indicate that the
+differences, although consistent, are relatively small.  Detailed results
+of UMM$_{100}$ and UMM$_{200}$ and the original \CEGOorig are given in
+Table~\ref{tab:results_fitness} for completeness. In this table, it can be seen that the quality
+of \CEGOorig solutions is similar to the UMM$_{100}$ version of UMM.
+
+\myparagraph{Runtime comparison}
+%
+Table~\ref{tab:results} also shows the runtime (in hours) of \CEGOinv and UMM when both
+are run for 400 evaluations. For the same number of function evaluations, CEGO requires almost a day, whereas UMM
+never requires more than one hour, often much less than that.
+%
+Details for other algorithm variants are given in Table~\ref{tab:results_time}.
+The average runtime for all the instances is around 0.3 minutes, 1.15 minutes
+and 4.5 minutes respectively for the UMM variants with 100, 200 and 400 fitness
+evaluations.  For CEGO, in both versions, the average runtime for 400
+evaluations is more than 22 hours. This is 300 times slower than UMM for the
+same number of function evaluations.  Moreover, the runtime of UMM scales well
+for larger instances since its computational complexity  is $O(n \log n)$ at
+each iteration.
 
@@
 \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
-\begin{table*}[tb]
+\afterpage{%
+\begin{table*}
   \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance.
@@
 
-% The situation is very similar for the QAP, where both algorithms show the same
-% behaviour as in the PFSP for all instances tested, except for
-% \texttt{kra} instances. As shown in Fig~\ref{fig:kra32}, for \texttt{kra32} both
-% algorithms are able to improve over the number of evaluations, although not
-% with ease. As in the synthetic LOP instances, we again observe that CEGO does
-% improve faster than UMM initially, yet CEGO appears to get stuck around
-% evaluation 250, and UMM overtakes it around evaluation 300. We believe the
-% reason is the same as before, that is, the CEGO model is not able to adequately
-% reflect the ruggedness of the actual landscape and, hence, the best solutions
-% found by the GA searching the model are not improved solutions for the actual
-% problem. Interestingly, there are well-known differences between
-% \texttt{kra30a}, \texttt{kra30b}, \texttt{kra32}, on one hand, and \texttt{nug12}, \texttt{nug30} and
-% \texttt{tho32}, on the other~\citep{AnsBriGou2002qap}. However, we cannot at
-% this moment point out the root cause of this different behavior.
-
-% \begin{figure}[tp]
-%   \centering%
-%   \includegraphics[width=\textwidth]{../img/fitness_real_qap_kra32_dat}
-%   \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the QAP instance \texttt{kra32}.\label{fig:kra32}}
-%   \vspace*{-1em}
-% \end{figure}
-
-%Table~\ref{tab:results} shows results over all instances evaluated (excluding
-%the synthetic LOP instances). In particular, we show the mean fitness (and
-%standard deviation) of the best solution found at the end of each run.
-We also report the mean runtime in hours and, since CEGO sometimes reached the limit of 144 hours, also the number of fitness evaluations reached by CEGO.
-
-%In the LOP, differences are mostly in favour of UMM, due to the time limit of
-%144 hours, which resulted in CEGO typically achieving less than 200
-%evaluations. In contrast, UMM was always able to reach 400 evaluations in less
-%than 12.5 hours. If UMM was allowed to reach the 144 hours limit, the differences
-%would be even stronger.
-%%
-%In the PFSP, the differences are minor and/or not statistically
-%significant, with the qualification that both algorithms appear to be
-%performing a random search.
-
-\begin{table*}[tb]%time
-  \caption{Mean runtime  in minutes, over 10 independent runs for each problem instance, of the original CEGO (\CEGOorig) and its improved version (\CEGOinv) and three configurations of UMM that differ in the budget of evaluations $m \in \{100, 200, 400\}$.
+\begin{table*}
+    \caption{Mean fitness, over 10 independent runs for each problem instance, of the best solution found at the end of each run of the original CEGO (\CEGOorig) and its improved version (\CEGOinv) and three variants of UMM that differ in the budget of evaluations $m \in \{100, 200, 400\}$.
+The runs of CEGO used a budget of $m=400$.\label{tab:results_fitness}}
+\begin{tabular}{llrrrrr}
+\toprule
+\bf Problem & \bf   Instance &  \CEGOinv &  \CEGOorig &  UMM$_{400}$ &   UMM$_{200}$ &  UMM$_{100}$ \\
+\midrule
+    LOP &    N-p40-01 &    10496.6 &    15027.7 &    12453.7 &  1.336170e+04 &    14649.9 \\
+    &    N-p40-02 &    11714.7 &    15150.2 &    13191.9 &  1.420900e+04 &    15206.2 \\
+    &  N-t59b11xx &   102408.0 &   126204.0 &   114726.5 &  1.202020e+05 &   132039.5 \\
+    &  N-t59d11xx &    12692.2 &    30199.8 &    16775.4 &  2.352089e+04 &    29884.1 \\
+    &  N-sgb75.01 &  1893533.2 &  2095899.6 &  1983521.2 &  2.004006e+06 &  2046895.1 \\
+    &  N-sgb75.02 &  1895801.8 &  2069286.9 &  1973083.9 &  2.001244e+06 &  2039356.1 \\
+    \midrule
+    PFSP &       rec05 &     1276.2 &     1318.7 &     1282.1 &  1.287000e+03 &     1320.2 \\
+    &       rec13 &     1992.9 &     2096.0 &     2077.7 &  2.108500e+03 &     2131.6 \\
+    &       rec19 &     2264.5 &     2328.4 &     2318.5 &  2.352100e+03 &     2386.5 \\
+    &       rec31 &     3415.5 &     3512.3 &     3420.5 &  3.490900e+03 &     3474.3 \\
+\bottomrule
+\end{tabular}
+\end{table*}
+
+\begin{table*}
+  \caption{Mean runtime  in minutes, over 10 independent runs for each problem instance, of the original CEGO (\CEGOorig) and its improved version (\CEGOinv) and three variants of UMM that differ in the budget of evaluations $m \in \{100, 200, 400\}$.
 The runs of CEGO used a budget of $m=400$.\label{tab:results_time}}
@@
 \midrule
-    LOP &    N-p40-01 &   1195.90 &    1170.81 &     3.12 &  0.850000 &     0.21 \\
-    LOP &    N-p40-02 &   1207.24 &    1183.09 &     3.11 &  0.830000 &     0.20 \\
-    LOP &  N-t59b11xx &   1265.98 &    1240.36 &     3.78 &  0.950000 &     0.23 \\
-    LOP &  N-t59d11xx &   1269.93 &    1252.89 &     3.93 &  0.966667 &     0.25 \\
-    LOP &  N-sgb75.01 &   1994.22 &    2004.70 &    10.88 &  2.730000 &     0.72 \\
-    LOP &  N-sgb75.02 &   1995.09 &    1975.41 &    10.88 &  2.760000 &     0.69 \\
-   PFSP &       rec05 &   1099.63 &     917.81 &     0.80 &  0.200000 &     0.10 \\
-   PFSP &       rec13 &   1109.51 &    1043.08 &     0.85 &  0.200000 &     0.10 \\
-   PFSP &       rec19 &   1090.06 &    1087.38 &     1.76 &  0.470000 &     0.10 \\
-   PFSP &       rec31 &   1387.09 &    1367.26 &     4.77 &  1.280000 &     0.32 \\
+    LOP     & N-p40-01       & 1195.90  & 1170.81   & 3.12        & 0.85    & 0.21        \\
+            & N-p40-02       & 1207.24  & 1183.09   & 3.11        & 0.83    & 0.20        \\
+            & N-t59b11xx     & 1265.98  & 1240.36   & 3.78        & 0.95    & 0.23        \\
+            & N-t59d11xx     & 1269.93  & 1252.89   & 3.93        & 0.97    & 0.25        \\
+            & N-sgb75.01     & 1994.22  & 2004.70   & 10.88       & 2.73    & 0.72        \\
+            & N-sgb75.02     & 1995.09  & 1975.41   & 10.88       & 2.76    & 0.69        \\
+\midrule
+    PFSP    & rec05          & 1099.63  & 917.81    & 0.80        & 0.20    & 0.10        \\
+            & rec13          & 1109.51  & 1043.08   & 0.85        & 0.20    & 0.10        \\
+            & rec19          & 1090.06  & 1087.38   & 1.76        & 0.47    & 0.10        \\
+            & rec31          & 1387.09  & 1367.26   & 4.77        & 1.28    & 0.32        \\
 \bottomrule
@@
 
-\begin{table*}[tb]%fitness
-    \caption{Mean fitness, over 10 independent runs for each problem instance, of the best solution found at the end of each run of the original CEGO (\CEGOorig) and its improved version (\CEGOinv) and three configurations of UMM that differ in the budget of evaluations $m \in \{100, 200, 400\}$.
-The runs of CEGO used a budget of $m=400$.\label{tab:results_fitness}}
-\begin{tabular}{llrrrrr}
-\toprule
-Problem &    Instance &   \CEGOinv &  \CEGOorig &    UMM$_{400}$ &       UMM$_{200}$ &    UMM$_{100}$ \\
-\midrule
-    LOP &    N-p40-01 &    10496.6 &    15027.7 &    12453.7 &  1.336170e+04 &    14649.9 \\
-    LOP &    N-p40-02 &    11714.7 &    15150.2 &    13191.9 &  1.420900e+04 &    15206.2 \\
-    LOP &  N-t59b11xx &   102408.0 &   126204.0 &   114726.5 &  1.202020e+05 &   132039.5 \\
-    LOP &  N-t59d11xx &    12692.2 &    30199.8 &    16775.4 &  2.352089e+04 &    29884.1 \\
-    LOP &  N-sgb75.01 &  1893533.2 &  2095899.6 &  1983521.2 &  2.004006e+06 &  2046895.1 \\
-    LOP &  N-sgb75.02 &  1895801.8 &  2069286.9 &  1973083.9 &  2.001244e+06 &  2039356.1 \\
-   PFSP &       rec05 &     1276.2 &     1318.7 &     1282.1 &  1.287000e+03 &     1320.2 \\
-   PFSP &       rec13 &     1992.9 &     2096.0 &     2077.7 &  2.108500e+03 &     2131.6 \\
-   PFSP &       rec19 &     2264.5 &     2328.4 &     2318.5 &  2.352100e+03 &     2386.5 \\
-   PFSP &       rec31 &     3415.5 &     3512.3 &     3420.5 &  3.490900e+03 &     3474.3 \\
-\bottomrule
-\end{tabular}\end{table*}
+}
 

### 2021-02-05 paper/main.tex
shorten the abstract

@@
   real-world experiment. Since each fitness evaluation is expensive in terms of
-  time or resources, only a limited number of evaluations is possible,
-  typically several orders of magnitude smaller than in non-expensive
-  problems. In this scenario, classical optimization methods such as
-  mixed-integer programming and local search are not useful.  In the continuous
-  case, Bayesian optimization, in particular using Gaussian processes, has
-  proven very effective under these conditions. Much less research is available
-  in the combinatorial case. In this paper, we propose and analyze UMM, an
-  estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
-  model and unbalanced rank aggregation (uBorda).
-
+  time or resources, the number of possible evaluations is typically several
+  orders of magnitude smaller than in non-expensive problems. Classical
+  optimization methods are not useful in this scenario.
+  In this paper, we propose and analyze UMM, an estimation-of-distribution
+  (EDA) algorithm based on a Mallows probabilistic model and unbalanced rank
+  aggregation (uBorda).
   Experimental results on black-box versions of LOP and PFSP show that UMM
@@
   slightly better than those of UMM and dramatically better than the original
-  version.  As another benefit, the computational complexity of UMM increases
-  linearly with both the number of function evaluations and the permutation
-  size, which results in computation times an order of magnitude shorter than
-  CEGO, which makes it specially useful when both computation time and number
-  of evaluations are limited.  \sloppy
+  version.  Another benefit of UMM is that its computational complexity
+  increases linearly with both the number of function evaluations and the
+  permutation size, which results in computation times an order of magnitude
+  shorter than CEGO, making it specially useful when both computation time and
+  number of evaluations are limited.
+  \sloppy
 \end{abstract}
@@
 % (Eq.~\ref{eq:kendall_decomp}).
-
+%
 In CEGO, the surrogate model is

### 2021-03-26 paper/main.tex
notebook
implement reviewer comments

@@
 \usepackage{relsize}
-\usepackage{amsmath,amssymb}
+\usepackage{amsmath}
 \usepackage{xspace}
@@
 (CEGO)~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}. However, the
-ruggedness of combinatorial landscapes, which makes local search particularly
-effective for combinatorial problems, lessens the effectiveness of global
+ruggedness of combinatorial landscapes lessens the effectiveness of global
 surrogate models~\citep{EriPeaGar2019scalable}. Moreover, surrogate models are
@@
 %
-\begin{equation}
+\begin{equation}\label{eq:kendall_decomp}
 \begin{split}
-d(\sigma _{1},\sigma_{2})=\left\lvert\Bigl\{(i,j):i<j \land \phantom{}\Bigl(\right.&\bigl(\sigma_{1}(i) < \sigma_{1}(j)\land \sigma_{2}(i)>\sigma_{2}(j)\bigr)\lor \phantom{} \\
- &\bigl(\sigma_{1}(i)>\sigma_{1}(j)\land \sigma_{2}(i)<\sigma_{2}(j)\bigr)\Bigr)\Bigr\}\Bigr\rvert\\
- \end{split}
- \label{eq:kendall_decomp}
+\hspace*{-0.5em}d(\sigma _{1},\sigma_{2})=\left\lvert\Bigl\{(i,j):i<j \land \phantom{}\Bigl(\right.&\bigl(\sigma_{1}(i) < \sigma_{1}(j)\land \sigma_{2}(i)>\sigma_{2}(j)\bigr)\lor \phantom{} \\
+ \bigl(&\sigma_{1}(i)>\sigma_{1}(j)\land \sigma_{2}(i)<\sigma_{2}(j)\bigr)\Bigr)\Bigr\}\Bigr\rvert\\
+ \end{split}\hspace*{-0.6em}
  \end{equation}
 
-An equivalent definition for the \ken distance counts the number of adjacent swaps that have to be made to convert $\sigma_1^{-1}$ into $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}.
+An equivalent definition for the \ken distance counts the number of adjacent swaps that have to be made to convert $\sigma_1^{-1}$ into $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn} or adjacent swap distance.
 %\EKHINE{Mirandolo bien, lo que pasa es que dicen que es lo mismo la swap distance que la kendall!}
@@
 The Borda ranking of a sample of permutations $S \subset S_n$ is computed as follows. 
-First, for each item $i \in [n]$, compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$.
-Then, assign label 1 to the item with the smallest Borda score, label 2 to the item with the second smallest score and so on, i.e., order the positions by Borda score increasingly.
+First, for each item $i \in [n]$, compute its Borda score $B(i) =  \sum_{\sigma\in S}  \sigma(i)$.
+Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the positions by Borda score increasingly.
 
@@
 \begin{equation}\label{eq:uborda_score}
-B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i) \enspace,
+B(i) =  \sum_{\sigma\in S}  w(\sigma) \sigma(i) \enspace,
 \end{equation}
@@
 
-As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are generated uniformly at random  (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated.
+As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are generated uniformly at random  (although it would be possible to employ other sampling methods such as max-min-distance sequential design) and evaluated.
 Then, using the current sample $S$, the uBorda algorithm is applied to learn a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}).
@@
 % \EKHINE{es decreciente igual}
-In UMM, the value of $\rho$ depends on the set of true fitness evaluations $F$
+In UMM, the value of $\rho$ depends on the set of true fitness evaluations saved in $S$
 and it is set at each iteration such that the largest $100r_1$\% of the mass of
@@
 %  \begin{split}
-    r_1 \cdot \sum_{\tuple{\sigma,f(\sigma)}\in S}\rho^{f(\sigma)} =  \sum_{\tuple{\sigma', f(\sigma')}\in S'}\rho^{f(\sigma')}  %\\
+    r_1 \cdot \sum_{\tuple{\sigma,f(\sigma)}\in S}\rho^{\text{scale}(f(\sigma))} =  \sum_{\tuple{\sigma', f(\sigma')}\in S'}\rho^{\text{scale}(f(\sigma'))}  %\\
   %    & \text{where}\;|S'| = r_2|S| \land \forall \sigma'\in S', \sigma\in S\setminus S' : f(\sigma')\leq f(\sigma).
@@
 where $|S'| = r_2|S| \land \forall \tuple{\sigma',f(\sigma')}\in S', \tuple{\sigma,f(\sigma)}\in S\setminus S' : f(\sigma')\leq f(\sigma)$, that is, $S'$ contains the $r_2|S|$ elements of $S$ with the best fitness values.
+The function $\text{scale}\colon \mathbb{R} \to [0,1]$ is defined as $\text{scale}(f) = (f - f_\text{min})/(f_\text{max} - f_\text{min})$, where $f_\text{min}$ and $f_\text{max}$ are, respectively, the minimum and maximum fitness values in $S$.
+A possible alternative to scaling would be ranking the elements in $S$ with respect to their fitness values, which would make UMM invariant to non-linear monotonic transformations of the fitness function.
+However, in most applications, the fitness values have a real-world meaning (e.g., monetary value) and the relative scale of fitness differences is relevant.
 
@@
 Parameter $\theta$ is set automatically by the algorithm to control the diversification and intensification trade off.
-At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$, i.e., half of the expected distance for the uniform distribution.
+At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{2} / 2$, i.e., half of the expected distance for the uniform distribution.
 Then the expected distance linearly decreases until the last iteration $m$ where $\theta$ is set such that $\Edist = 1$.
@@
 Following the authors of
-CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, the surrogate model is optimized by a GA with  population size of 20, crossover rate of 0.5,
-mutation rate of $1/n$, tournament selection of size 2 with probability of 0.9,
+CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, the surrogate model is
+optimized by a GA with population size of 20, crossover rate of 0.5, mutation
+rate of $1/n$, tournament selection of size 2 with probability of 0.9,
 interchange mutation (i.e., exchanging two randomly selected elements) and
-cycle crossover for permutations. To generate each new solution, CEGO runs the
-GA with a budget of $10^4$ evaluations of the model as a surrogate of the true
-fitness function.  The original paper~\citep[p.~875]{ZaeStoFriFisNauBar2014} notes that coupling the GA with local
-search does not improve the results significantly since the model is anyway an
-inexact estimation of the original fitness
-function.
+cycle crossover for permutations. CEGO runs the GA with a budget of $10^4$
+evaluations of the surrogate model to generate a new solution, which is then
+evaluated on the true fitness function.  The original
+paper~\citep[p.~875]{ZaeStoFriFisNauBar2014} notes that coupling the GA with
+local search does not improve the results significantly since the model is
+anyway an inexact estimation of the original fitness function.
 
@@
 Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually builds upon orderings.
-For example, the Swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.
+For example, the (adjacent) swap distance $d_s(\sigma,\pi)$ is the \ken distance counterpart in optimization, $d_s(\sigma,\pi) = d(\sigma^{-1}, \pi^{-1})$.
 %The concepts of rankings and orderings are usually confused of ignored.
@@
 %
-Table~\ref{tab:results} also shows the runtime (in hours) of \CEGOinv and UMM when both
+Table~\ref{tab:results} also shows the runtime (in minutes) of \CEGOinv and UMM when both
 are run for 400 evaluations. For the same number of function evaluations, CEGO requires almost a day, whereas UMM

### 2021-03-26 umm.py

@@ -79,6 +79,7 @@ def UMM(instance, seed, budget,
 
     for m in range(budget - m_ini):
         ws = np.asarray(fitnesses).copy()
+        # FIXME: We could use rankings for invariance
         # FIXME: For maximization, this need to be changed.
         ws = ws - ws.min()
         # FIXME: Handle if ws.max() == 0.

### 2021-03-26 umm.py
umm.py: Implement maxmindist sampling.

@@ -1,4 +1,4 @@
-from imp import reload
+# from imp import reload
 import numpy as np
 import mallows_kendall as mk
 from scipy.spatial import distance
@@ -56,7 +56,34 @@ def get_expected_distance(iterat, n, budget):
   a = f_ini - jump * iterat
   return max(a, f_end)
 
+def remove_duplicates(s):
+  d = {a.tostring(): a for a in s}
+  return list(d.values())
 
+def design_random(m, n):
+  """
+  m: number of permutations to generate
+  n: permutation size"""
+  return remove_duplicates([ np.random.permutation(n) for _ in range(m)])
+
+def min_distance(x, s, dist):
+  return np.apply_along_axis(dist, -1, np.asarray(s), B=x).min()
+  
+def design_maxmindist(m, n, dist = mk.kendallTau, budget = 100):
+  sample = [ np.random.permutation(n) ]
+  while len(sample) < m:
+    best = np.random.permutation(n)
+    best_d = min_distance(best, sample, dist)
+    for i in range(budget):
+      xnew = np.random.permutation(n)
+      xnew_d = min_distance(xnew, sample, dist)
+      if xnew_d > best_d:
+        best, best_d = xnew, xnew_d
+    sample.append(best)
+  return remove_duplicates(sample)
+
+
+  
 def UMM(instance, seed, budget,
         m_ini, budgetMM,
         ratio_samples_learn,
@@ -71,7 +98,7 @@ def UMM(instance, seed, budget,
       f_eval = lambda p: instance.fitness(np.argsort(p))
 
     n = instance.n
-    sample = [ np.random.permutation(np.arange(n)) for _ in range(m_ini)]
+    sample = design_random(m_ini, n)
     fitnesses = [f_eval(perm) for perm in sample]
     # ['rho','phi_estim','phi_sample','Distance']
     res = [ [np.nan, np.nan, np.nan,

### 2021-03-28 3-launch_exps.sh
implement --init option in UMM

@@ -89,10 +89,10 @@ pfsp/rec31.txt \
 "
 
 INSTANCES="\
-pfsp/rec05.txt \
-pfsp/rec13.txt \
-pfsp/rec19.txt \
-pfsp/rec31.txt \
+pfsp_cmax/rec05.txt \
+pfsp_cmax/rec13.txt \
+pfsp_cmax/rec19.txt \
+pfsp_cmax/rec31.txt \
 "
 
 
@@ -118,6 +118,8 @@ r_2=0.9
 beta=10
 beta=1
 umm_m_ini=10
+init="random"
+init="maxmindist"
 
 counter=0
 for m in $budget; do
@@ -126,10 +128,8 @@ for m in $budget; do
 	    counter=$((counter+1))
 	    RESULTS="$OUTDIR/results/m${m}-er${er}/$instance"
 	    mkdir -p "$RESULTS"
-	    $LAUNCHER umm "${RESULTS}/umm-b${beta}" $instance --m_ini $umm_m_ini --budgetMM $beta --rsl $r_1 --wml $r_2 --budget $m --eval_ranks $er
+	    $LAUNCHER umm "${RESULTS}/umm-{init}-b${beta}" $instance --m_ini $umm_m_ini --budgetMM $beta --rsl $r_1 --wml $r_2 --budget $m --init $init --eval_ranks $er
 	    #$LAUNCHER cego "${RESULTS}/cego" $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $m --eval_ranks $er
 	done
     done
 done
-
-

### 2021-03-28 runner.py

@@ -61,5 +61,5 @@ def run_once(algo_name, instance_name, seed, out_filename = None,
     df['Solver'] = algo_name
     if out_filename is not None:
         df.to_csv(out_filename + '.csv.xz', index=False, compression = "xz")
-        df.to_pickle(out_filename + '.pkl.xz', compression = "xz")
+        # df.to_pickle(out_filename + '.pkl.xz', compression = "xz")
     return df

### 2021-03-28 target-runner-umm.py

@@ -36,6 +36,7 @@ parser.add_argument('--budgetMM', type=int, default=0, help='budgetMM')
 parser.add_argument('--rsl', type=float, default=0, help='rsl')
 parser.add_argument('--wml', type=float, default=0, help='wml')
 parser.add_argument('--eval_ranks', type=int, default=0, help='eval_ranks')
+parser.add_argument('--init', choices = ['random', 'maxmindist'], help='init')
 
 args = parser.parse_args()
 
@@ -46,8 +47,8 @@ stdout = sys.stdout
 outfilename = f'c{args.configuration_id}-{args.instance_id}-{args.seed}.stdout'
 with open(outfilename, 'w') as sys.stdout:
     df = runner.run_once("UMM", args.instance_name, args.seed, budget = budget, m_ini = args.m_ini, 
-                         budgetMM = args.budgetMM, ratio_samples_learn = args.rsl, weight_mass_learn = args.wml, eval_ranks = args.eval_ranks, out_filename = args.output)
-            
+                         budgetMM = args.budgetMM, ratio_samples_learn = args.rsl, weight_mass_learn = args.wml, eval_ranks = args.eval_ranks, init = args.init, out_filename = args.output)
+
 sys.stdout = stdout
 print(df["Fitness"].min())
 # remove tmp file.

### 2021-03-28 umm.py

@@ -69,7 +69,7 @@ def design_random(m, n):
 def min_distance(x, s, dist):
   return np.apply_along_axis(dist, -1, np.asarray(s), B=x).min()
   
-def design_maxmindist(m, n, dist = mk.kendallTau, budget = 100):
+def design_maxmindist(m, n, dist = mk.kendallTau, budget = 1000):
   sample = [ np.random.permutation(n) ]
   while len(sample) < m:
     best = np.random.permutation(n)
@@ -87,7 +87,7 @@ def design_maxmindist(m, n, dist = mk.kendallTau, budget = 100):
 def UMM(instance, seed, budget,
         m_ini, budgetMM,
         ratio_samples_learn,
-        weight_mass_learn, eval_ranks):
+        weight_mass_learn, eval_ranks, init):
 
     np.random.seed(seed)
 
@@ -98,7 +98,13 @@ def UMM(instance, seed, budget,
       f_eval = lambda p: instance.fitness(np.argsort(p))
 
     n = instance.n
-    sample = design_random(m_ini, n)
+    if init == "random":
+      sample = design_random(m_ini, n)
+    elif init == "maxmindist":
+      sample = design_maxmindist(m_ini, n)
+    else:
+      raise f"Invalid init: {init}"
+
     fitnesses = [f_eval(perm) for perm in sample]
     # ['rho','phi_estim','phi_sample','Distance']
     res = [ [np.nan, np.nan, np.nan,
@@ -154,4 +160,5 @@ def UMM(instance, seed, budget,
     df['ratio_samples_learn'] = ratio_samples_learn
     df['weight_mass_learn'] = weight_mass_learn
     df['eval_ranks'] = eval_ranks
+    df['init'] = init
     return df

### 2021-03-28 paper/main.tex
add authors
sync with iridia master, update acks

@@
 % Double-blind
-\author{Double-blind}
+%\author{Double-blind}
 % \authornote{Dr.~Trovato insisted his name be first.}
@@
 
-% \author{Ekhiñe Irurozki\inst{1} \and Manuel López-Ibáñez\inst{2}\orcidID{0000-0001-9974-1295}}
-% \institute{
-%    Basque Center for Applied Mathematics\\
-%    \email{eirurozki@bcamath.org}
-%    \and
-%    University of Málaga, Málaga, Spain\\
-%    \email{manuel.lopez-ibanez@uma.es}
-%  }
+\author{Ekhiñe Irurozki}
+\affiliation{
+  \institution{Basque Center for Applied Mathematics}
+}
+  \email{eirurozki@bcamath.org}
+
+\author{Manuel López-Ibáñez}
+\orcid{0000-0001-9974-1295}
+\affiliation{
+  \institution{University of Málaga}
+  \city{Málaga}
+  \state{Spain}
+}
+\email{manuel.lopez-ibanez@uma.es}
+
 %    \renewcommand{\shortauthors}{B. Trovato et al.}
@@
 %  \vspace*{-1em}
-%\begin{acks}
-%% MANUEL: Blind, so no acknowledgments yet.
+\begin{acks}
 %\paragraph*{Acknowledgements.} Blind
-% M.\@ L\'opez-Ib\'a\~nez is a ``Beatriz Galindo'' Senior Distinguished Researcher (BEAGAL 18/00053) funded by the Ministry of Science and Innovation of the Spanish Government.
-%
-%\manuel{}{: We should also thank the COST ACTION: Ask Carlos?} \ekhine{}{Sure!, esto no habra q ponerlo aun por ser blind, no?}
-%Thanks to: Hao Wang (Leiden University) for pointing us to the arguments of
-%\citep{EriPeaGar2019scalable}; Thomas Stützle (Université libre de Bruxelles) for suggesting works
+We would like to thank Hao Wang (Leiden University) for pointing us to the arguments of
+\citep{EriPeaGar2019scalable}.
+%
+\manuel{}{: We should also thank the COST ACTION}
+%
+M.\@ L\'opez-Ib\'a\~nez is a ``Beatriz Galindo'' Senior Distinguished Researcher (BEAGAL 18/00053) funded by the Ministry of Science and Innovation of the Spanish Government. 
+%Thomas Stützle (Université libre de Bruxelles) for suggesting works
 %about differences between QAPLIB instances; and Ahmed Boujaada for the experimental support. 
@@
   %   Scientists' Support Program}.
-%\end{acks}
+\end{acks}
 

### 2021-03-28 paper/main.tex
update paper

@@
 learned from the current sample $S$ by using the uBorda algorithm with a weight
-$w(\sigma)=\rho^{f(\sigma)}$ (Section~\ref{sec:uborda}). The learning rate is
+$w(\sigma)=\rho^{\text{scale}(f(\sigma))}$ (Section~\ref{sec:uborda}).
+The function $\text{scale}\colon \mathbb{R} \to [0,1]$ is defined as $\text{scale}(f) = (f - f_\text{min})/(f_\text{max} - f_\text{min})$, where $f_\text{min}$ and $f_\text{max}$ are, respectively, the minimum and maximum fitness values in $S$.
+A possible alternative to scaling would be ranking the elements in $S$ with respect to their fitness values, which would make UMM invariant to non-linear monotonic transformations of the fitness function.
+However, in most applications, the fitness values have a real-world meaning (e.g., monetary value) and the relative scale of fitness differences is relevant.
+
+The learning rate is
 controlled by parameter $0 < \rho < 1$. 
@@
 where $|S'| = r_2|S| \land \forall \tuple{\sigma',f(\sigma')}\in S', \tuple{\sigma,f(\sigma)}\in S\setminus S' : f(\sigma')\leq f(\sigma)$, that is, $S'$ contains the $r_2|S|$ elements of $S$ with the best fitness values.
-The function $\text{scale}\colon \mathbb{R} \to [0,1]$ is defined as $\text{scale}(f) = (f - f_\text{min})/(f_\text{max} - f_\text{min})$, where $f_\text{min}$ and $f_\text{max}$ are, respectively, the minimum and maximum fitness values in $S$.
-A possible alternative to scaling would be ranking the elements in $S$ with respect to their fitness values, which would make UMM invariant to non-linear monotonic transformations of the fitness function.
-However, in most applications, the fitness values have a real-world meaning (e.g., monetary value) and the relative scale of fitness differences is relevant.
-
 %The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification.

### 2021-04-05 mallows_kendall.py
hamming v0

@@ -145,12 +145,14 @@ def likelihood_mm(perms, s0, theta):
     # print(probs,m,n)
     return probs.sum()
 
-def samplingMM(m,n,theta=None, phi=None, k=None):
+def samplingMM(m,n,theta=None, phi=None, k=None,sigma0=None):
     # k return partial orderings
     theta, phi = check_theta_phi(theta, phi)
     if k == n:
         k = None
-    return samplingGMM(m, [theta] * (n-1), topk = k)
+    sample =  samplingGMM(m, [theta] * (n-1), topk = k)
+    sample = [perm[sigma0] for perm in sample]
+    return sample
 
 def samplingGMM(m,theta, topk=None):
     #  returns RANKINGS!!!!!!!*****
@@ -290,4 +292,3 @@ def u_phi(sample, s0, ws):
 
 def uborda(sample, ws):
     return borda(sample * ws[:, None])
-

### 2021-04-05 pfsp.py

@@ -14,7 +14,8 @@ class PFSP(Problem):
             print(f"header: {header}")
             if "tai" in filename:
                 # number of jobs, number of machines, initial seed, upper bound and lower bound :
-                n,m = f.readline().strip().split()
+                # print(f.readline().strip().split())
+                n,m = f.readline().strip().split()[:2]
                 n,m = int(n),int(m)
                 f.readline()
                 P = np.loadtxt(f, max_rows=n)
@@ -27,10 +28,10 @@ class PFSP(Problem):
                 P = np.loadtxt(f, max_rows=n)
                 # Processing times: each even column is useless.
                 P = P[:,1::2]
-        
+
         assert P.shape[0] == n
         assert P.shape[1] == m
- 
+
         best_sol = None
         if opt_filename is not None:
             with open(opt_filename) as f:
@@ -68,7 +69,7 @@ class PFSP(Problem):
     def sum_completion(self, x):
         """Also known as sum of completion times or sum of flowtimes"""
         return self.completion_times(x)[:, self.m - 1].sum()
-        
+
     def fitness_nosave(self, x):
         # In case it is not numpy array.
         x = np.asarray(x, dtype=int)

### 2021-04-05 umm.py

@@ -1,13 +1,16 @@
 # from imp import reload
 import numpy as np
 import mallows_kendall as mk
+import mallows_hamming as mh
+from imp import reload
+reload(mh)
 from scipy.spatial import distance
 import pandas as pd
 
-def binary_search_rho(w, ratio_samples_learn, weight_mass_learn,
-                      # 0 <= w_i <= 1, w is sorted increasingly,
-                      rho_ini=1, rho_end=0, tol=0.001):
+
+def binary_search_rho(w, ratio_samples_learn, weight_mass_learn, rho_ini=1, rho_end=0, tol=0.001):
   w = np.asarray(w)
+   # 0 <= w_i <= 1, w is sorted increasingly,
   assert np.all(w >= 0.0)
   assert np.all(w <= 1.0)
 
@@ -30,7 +33,6 @@ def binary_search_rho(w, ratio_samples_learn, weight_mass_learn,
       # If the differenc eot the target weight_mass is very small, just return.
       if abs(a / b - weight_mass_learn) < tol:
           return rho_med
-
       if a / b > weight_mass_learn:
           mid, last = rho_ini, rho_med
       else:
@@ -68,7 +70,7 @@ def design_random(m, n):
 
 def min_distance(x, s, dist):
   return np.apply_along_axis(dist, -1, np.asarray(s), B=x).min()
-  
+
 def design_maxmindist(m, n, dist = mk.kendallTau, budget = 1000):
   sample = [ np.random.permutation(n) ]
   while len(sample) < m:
@@ -83,7 +85,7 @@ def design_maxmindist(m, n, dist = mk.kendallTau, budget = 1000):
   return remove_duplicates(sample)
 
 
-  
+
 def UMM(instance, seed, budget,
         m_ini, budgetMM,
         ratio_samples_learn,
@@ -112,44 +114,29 @@ def UMM(instance, seed, budget,
 
     for m in range(budget - m_ini):
         ws = np.asarray(fitnesses).copy()
-        # FIXME: We could use rankings for invariance
-        # FIXME: For maximization, this need to be changed.
         ws = ws - ws.min()
-        # FIXME: Handle if ws.max() == 0.
         ws = ws / ws.max()
         co = ws.copy()
         co.sort()
-
         rho = binary_search_rho(co, ratio_samples_learn, weight_mass_learn)
-        # print(fitnesses)
-        # print(ws)
         ws = rho ** ws #MINIMIZE
-        # print(ws)
-        # ws = rho ** (1-ws) #MAXIMIZE
[…]

### 2021-04-10 mallows_kendall.py
* mallows_kendall.py: Improve error.

@@ -105,10 +105,11 @@ def fit_MM_phi(n, dist_avg): #returns sigma, phi
     try:
         theta = optimize.newton(mle_theta_mm_f, 0.01, fprime=mle_theta_mm_fdev, args=(n, dist_avg), tol=1.48e-08, maxiter=500, fprime2=None)
     except:
+        # FIXME: This returns before giving an error???
         if dist_avg == 0.0:
             # FIXME: Why 5?
             return s0, np.exp(-5)#=phi
-        print("error. fit_mm. dist_avg=",dist_avg, dist_avg == 0.0)
+        print(f"error: fit_MM_phi. dist_avg={dist_avg}", dist_avg == 0.0)
         print(rankings)
         print(s0)
         raise
@@ -276,7 +277,7 @@ def random_perm_at_dist(n, dist, sk):
 
 
 def u_phi(sample, s0, ws):
-    m, n = np.asarray(sample).shape
+    #m, n = np.asarray(sample).shape
     #if s0 is None: s0 = np.argsort(np.argsort(rankings.sum(axis=0))) #borda
     dist_avg = np.asarray([kendallTau(perm, s0) for perm in sample]*ws).sum()/ws.sum() #np.mean(np.array([kendallTau(s0, perm) for perm in rankings]))
     try:
@@ -284,8 +285,7 @@ def u_phi(sample, s0, ws):
         theta = optimize.newton(mle_theta_mm_f, 0.01, fprime=mle_theta_mm_fdev, args=(n, dist_avg), tol=1.48e-08, maxiter=500, fprime2=None)
     except:
         #if dist_avg == 0.0: return s0, np.exp(-5)#=phi
-        print("error. fit_mm. dist_avg=",dist_avg, dist_avg == 0.0)
-        print(s0)
+        print(f"error: u_phi: s0={s0}, dist_avg={dist_avg}", dist_avg == 0.0)
         raise
     if theta < 0:
         theta = 0.001

### 2021-04-10 3-launch_exps.sh
various tests

@@ -72,7 +72,7 @@ gen_lop_synthetic() {
 nruns=10
 
 LAUNCHER=qsub_job
-#LAUNCHER=launch_local
+LAUNCHER=launch_local
 
 ## For QAP, PFSP instances
 INSTANCES="\
@@ -128,7 +128,7 @@ for m in $budget; do
 	    counter=$((counter+1))
 	    RESULTS="$OUTDIR/results/m${m}-er${er}/$instance"
 	    mkdir -p "$RESULTS"
-	    $LAUNCHER umm "${RESULTS}/umm-{init}-b${beta}" $instance --m_ini $umm_m_ini --budgetMM $beta --rsl $r_1 --wml $r_2 --budget $m --init $init --eval_ranks $er
+	    $LAUNCHER umm "${RESULTS}/umm-sm001-${init}-b${beta}" $instance --m_ini $umm_m_ini --budgetMM $beta --rsl $r_1 --wml $r_2 --budget $m --init $init --eval_ranks $er
 	    #$LAUNCHER cego "${RESULTS}/cego" $instance --m_ini $cego_m_ini --budgetGA $budgetGA --budget $m --eval_ranks $er
 	done
     done

### 2021-04-14 paper/main.tex
hamming v0
Update template, fix affiliation errors
fit in 8 pages + ref

@@
 
-\author{Ekhiñe Irurozki}
+\author{Ekhine Irurozki}
+\orcid{0000-0003-3218-5735}
 \affiliation{
-  \institution{Basque Center for Applied Mathematics}
+  \institution{Telecom Paris}
+  \city{Paris}
+  \country{France}
 }
-  \email{eirurozki@bcamath.org}
+\additionalaffiliation{Basque Center for Applied Mathematics}
+
+\email{irurozki@telecom-paris.fr}
 
@@
   \city{Málaga}
-  \state{Spain}
+  \country{Spain}
 }
+\additionalaffiliation{University of Manchester, UK}
 \email{manuel.lopez-ibanez@uma.es}
@@
 
-This paper is structured as follows.
-Section~\ref{sec:backgroud} summarizes fundamental concepts about permutation spaces, uBorda, black-box combinatorial optimization under a limited budget and CEGO, a Bayesian optimizer for such problems.
+This paper is structured as follows. Section~\ref{sec:backgroud} summarizes fundamental concepts.
+% about permutation spaces, uBorda, black-box combinatorial optimization under a limited budget and CEGO, a Bayesian optimizer for such problems.
 Section~\ref{sec:umm} presents the UMM algorithm, which is the main contribution of this paper.
@@
 
-Permutations are defined as bijections of the set $[n]$ of integers onto itself.
-The set of all permutations of $n$ values is denoted as $S_n$ and has cardinality $n!$.
-We will use the one-line notation
+Permutations are %defined as
+bijections of the set $[n]$ of integers onto itself. The set of all permutations of $n$ values is denoted by $S_n$ and has cardinality $n!$. We will use the one-line notation
 % \MANUEL{what is one-line?}
-and denote permutations with Greek letters, $\sigma  = (\sigma(1), \ldots , \sigma(n))$. We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi = (\sigma(\pi(1)), \ldots, \sigma(\pi(n)))$
+and denote permutations with Greek letters, $\sigma  = (\sigma(1), \ldots , \sigma(n))$. The composition of $\sigma$ and $\pi$ is $\sigma\pi = (\sigma(\pi(1)), \ldots, \sigma(\pi(n)))$
 % \MANUEL{Is this correct?}
-and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.
+and the inverse of $\sigma$ is $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=(1, 2, 3, \dotsc,n)$ always holds.
 
@@
 %
-where $|S'| = r_2|S| \land \forall \tuple{\sigma',f(\sigma')}\in S', \tuple{\sigma,f(\sigma)}\in S\setminus S' : f(\sigma')\leq f(\sigma)$, that is, $S'$ contains the $r_2|S|$ elements of $S$ with the best fitness values.
+where $|S'| = r_2|S| \land \forall \tuple{\sigma',f(\sigma')}\in S', \tuple{\sigma,f(\sigma)}\in S\setminus S' : f(\sigma')\leq f(\sigma)$, i.e., $S'$ contains the $r_2|S|$ elements of $S$ with the best fitness.
 %The values of $r_1, r_2 \in (0, 1)$ are user-defined and control the balance between exploration and intensification.
@@
 
-\begin{figure}[tb]
+\begin{figure}[!tb]
   \centering%

### 2021-04-14 paper/main.tex
update refs
update acks

@@
 %
-\manuel{}{: We should also thank the COST ACTION}
+This article is based upon work from COST Action CA15140 `Improving Applicability of Nature-Inspired Optimisation by Joining Theory and Practice (ImAppNIO)' supported by COST (European Cooperation in Science and Technology).
 %
@@
 \bibliographystyle{ACM-Reference-Format-abbrv}
-\bibliography{optbib/abbrev,optbib/authors,optbib/abbrevshort,optbib/journals,optbib/biblio,optbib/crossref}%,./mendeley}
+\bibliography{optbib/abbrev,optbib/authors,optbib/journals,optbib/biblio,optbib/crossref}%,./mendeley}
 

### 2021-04-16 paper/main.tex
add bits

@@
 % Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-
-% ISBN
-\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-
-% Conference
-\acmConference[GECCO '21]{the Genetic and Evolutionary Computation Conference 2021}{July 10--14, 2021}{Lille, France}
+\copyrightyear{2021}
 \acmYear{2021}
-\copyrightyear{2021}
-
-%\acmArticle{4}
+\setcopyright{acmlicensed}
+\acmConference[GECCO '21]{2021 Genetic and Evolutionary Computation Conference}{July 10--14, 2021}{Lille, France}
+\acmBooktitle{2021 Genetic and Evolutionary Computation Conference (GECCO
+'21), July 10--14, 2021, Lille, France}
 \acmPrice{15.00}
+\acmDOI{10.1145/3449639.3459366}
+\acmISBN{978-1-4503-8350-9/21/07}
 

### 2021-04-16 paper/main.tex
add bits
paper
update

@@
 }
-\additionalaffiliation{University of Manchester, UK}
+%\additionalaffiliation{University of Manchester, UK}
 \email{manuel.lopez-ibanez@uma.es}
@@
     \texttt{N-t59d11xx} (LOP)\\[-0.2ex]
-    \includegraphics[width=\textwidth]{../img/heatmap_lop_IO_N-t59d11xx}
+    \includegraphics[width=\textwidth]{./img/heatmap_lop_IO_N-t59d11xx}
   \end{minipage}
@@
     \texttt{N-sgb75.02} (LOP)\\[-0.2ex]
- \includegraphics[width=\textwidth]{../img/heatmap_lop_SGB_N-sgb75_02}
+ \includegraphics[width=\textwidth]{./img/heatmap_lop_SGB_N-sgb75_02}
 \end{minipage}\\\vskip 1em
@@
   \texttt{reC13} (PFSP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec13_txt}
+  \includegraphics[width=\textwidth]{./img/heatmap_pfsp_rec13_txt}
 \end{minipage}
@@
   \texttt{reC19} (PFSP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec19_txt}
+  \includegraphics[width=\textwidth]{./img/heatmap_pfsp_rec19_txt}
 \end{minipage}
@@
 %   \texttt{kra32} (QAP)\\[-0.5ex]
-% \includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
+% \includegraphics[width=\textwidth]{./img/heatmap_kra32_dat}
 % \end{minipage}
@@
 %  \centering%
-%  \includegraphics[width=\textwidth]{../img/synthetic_LOP_combined}
+%  \includegraphics[width=\textwidth]{./img/synthetic_LOP_combined}
 %  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on LOP synthetic instances.\label{fig:lop_synth}}
@@
 %  \centering%
-%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandB_N-p50-01}\\
-%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d100_01}\\
-%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d150_01}\\
-%%  \includegraphics[width=\textwidth]{../img/fitness_real_lop_RandA1_N-t1d200_01}\\
+%  \includegraphics[width=\textwidth]{./img/fitness_real_lop_RandB_N-p50-01}\\
+%  \includegraphics[width=\textwidth]{./img/fitness_real_lop_RandA1_N-t1d100_01}\\
+%  \includegraphics[width=\textwidth]{./img/fitness_real_lop_RandA1_N-t1d150_01}\\
+%%  \includegraphics[width=\textwidth]{./img/fitness_real_lop_RandA1_N-t1d200_01}\\
 %    \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on three instances from LOLIB.\label{fig:lolib}}
@@
   \texttt{N-p40-01} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-01}
+  \includegraphics[width=\textwidth]{./img/fitness_lop_RandB_N-p40-01}
 \end{minipage}
@@
   \texttt{N-p40-02} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-02}
+  \includegraphics[width=\textwidth]{./img/fitness_lop_RandB_N-p40-02}
 \end{minipage}
@@
   \texttt{N-t59b11xx} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
+  \includegraphics[width=\textwidth]{./img/fitness_lop_IO_N-t59b11xx}
 \end{minipage}
@@
   \texttt{N-t59d11xx} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59d11xx}
+  \includegraphics[width=\textwidth]{./img/fitness_lop_IO_N-t59d11xx}
 \end{minipage}
@@
   \texttt{SGB\_N-sgb75\_01} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_01}
+  \includegraphics[width=\textwidth]{./img/fitness_lop_SGB_N-sgb75_01}
 \end{minipage}
@@
   \texttt{SGB\_N-sgb75\_01} (LOP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
+  \includegraphics[width=\textwidth]{./img/fitness_lop_SGB_N-sgb75_02}
 \end{minipage}
@@
   \texttt{reC05} (PFSP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec05_txt}
+  \includegraphics[width=\textwidth]{./img/fitness_pfsp_rec05_txt}
 \end{minipage}
@@
   \texttt{reC13} (PFSP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec13_txt}
+  \includegraphics[width=\textwidth]{./img/fitness_pfsp_rec13_txt}
 \end{minipage}
@@
   \texttt{reC19} (PFSP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec19_txt}
+  \includegraphics[width=\textwidth]{./img/fitness_pfsp_rec19_txt}
 \end{minipage}
@@
   \texttt{reC31} (PFSP)\\[-0.5ex]
-  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec31_txt}
+  \includegraphics[width=\textwidth]{./img/fitness_pfsp_rec31_txt}
 \end{minipage}
@@
 %\paragraph*{Acknowledgements.} Blind
+  \smaller%
   We would like to thank Hao Wang (Leiden University) for pointing us to the arguments of
@@
 M.\@ L\'opez-Ib\'a\~nez is a ``Beatriz Galindo'' Senior Distinguished Researcher (BEAGAL 18/00053) funded by the Ministry of Science and Innovation of the Spanish Government. 
+This work is also partially funded by the Elkartek program (Basque Government) and the Industrial Chair ``Machine Learning for Big Data'' from Télécom Paris, France. 
 %Thomas Stützle (Université libre de Bruxelles) for suggesting works

### 2021-04-16 paper/main.tex
paper
fix double label
update
fix margin

@@
 I wanted to explain why we use synthetic datasets.}
-We consider real LOP datasets from the LOLIB
+We consider LOP datasets from the LOLIB
 repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} In
@@
 \end{minipage}
-\label{fig:results}
 \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:results}}
@@
 \begin{table*}
+%\smaller
   \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance.
@@
 \begin{table*}
+%\smaller
     \caption{Mean fitness, over 10 independent runs for each problem instance, of the best solution found at the end of each run of the original CEGO (\CEGOorig) and its improved version (\CEGOinv) and three variants of UMM that differ in the budget of evaluations $m \in \{100, 200, 400\}$.
@@
 \begin{table*}
+%\smaller
   \caption{Mean runtime  in minutes, over 10 independent runs for each problem instance, of the original CEGO (\CEGOorig) and its improved version (\CEGOinv) and three variants of UMM that differ in the budget of evaluations $m \in \{100, 200, 400\}$.
